### 2021-02-15 paper/main.tex
outline

\documentclass[runningheads]{llncs}

\usepackage{amssymb,amsmath,mathtools,empheq,fancybox}

\usepackage{paralist}
\usepackage{url}
\usepackage{color}

\usepackage{textcomp,listings}

\newif\ifcomments
\commentstrue

\newcommand{\pr}[1]{{\color{red}``PR: #1''}}
\newcommand{\as}[1]{{\color{blue}``AS: #1''}}

\newif\ifoutline
\outlinetrue

\newcommand{\contents}[1]{\ifoutline{\color{blue}
    \begin{itemize}
    #1
    \end{itemize}
  }\fi}

\allowdisplaybreaks[1]

\title{Generalised Parikh Images of Regular Languages}
\author{some cool authors}
\institute{Uppsala University, Sweden}

\begin{document}
\maketitle

\begin{abstract}
  bla
\end{abstract}

\section{Introduction}

\contents{
\item key challenges: (1) complicated QE problem that was part of
  previous definitions; (2) languages with large alphabets described
  by symbolic automata; (3) regular languages described by products
  of automata
}

\subsection{Related Work}

\section{Overview}

\subsection{Generalised Parikh Images}

\contents{
\item Preliminaries, standard definition of the Parikh image of a
  regular/context-free language
\item Generalised version, considering arbitrary homomorphisms
\item Examples
}

\subsection{Lazy Computation of Parikh Images for Regular Languages}

\contents{
\item Lazy expansion of the Parikh conditions for a symbolic automaton
\item Finding elements vs.\ computing the complete Parikh image
\item Lazy product computation
}

\section{A Calculus of Parikh Images}

\contents{
\item Preliminaries, the underlying calculus, what are rules
\item Predicates used to represent Parikh images
\item Our calculus rules
\item Statement of properties, correctness, complexity
}

\section{Parikh Images from Products of Automata}

\contents{
\item additional rules needed for products
\item backjumping and conflict-driven learning
}

\section{Implementation and Experiments}

\contents{
\item length constraints
\item Parikh automata, automata with registers
\item Model-checking examples
}

\section{Conclusions}

\clearpage
%\bibliographystyle{splncs03}
%\bibliography{refs}

\end{document}

### 2021-07-16 paper/main.tex
Quickly rattle off a draft before vacation

@@
 \usepackage{textcomp,listings}
+\usepackage{biblatex}
+\usepackage{mymacros}
+\usepackage{mathpartir}
+\addbibresource{../bibliography.bib}
 
@@
 \begin{abstract}
-  bla
+  The image of the Parikh map is important in automata theory, offering a compact characterisation of an automaton.
+We contribute a novel understanding of how Parikh maps can be combined with arbitrary commutative monoid morphisms to efficiently represent a wide range of logics on automata and automata-like structures.
+Cases studied as examples include epistemic logic and string-length constraints in a string constraint solver.
+Furthermore, we show how this formulation can be efficiently implemented as a calculus in a theorem prover for succinct formulations of \Fudge{several constraints on strings}.
+In particular, our calculus is versatile enough to efficiently compute the Parikh image of a product of two automata, allowing our solver, \Ostrich{}, to solve \Fudge{X new instances} where it was previously constrained by the memory required to materialise the product.
+Finally, we show that this implementation in addition to being \Fudge{Z better} also offers \Fudge{X performance} improvements on \Fudge{Y real-world instances}, and in particular \Fudge{yields no regressions} in performance \Fudge{and, in fact, cures cancer, brings about world peace, and ends global hunger.}
 \end{abstract}
@@
 
-\subsection{Generalised Parikh Images}
+\subsection{Generalised Parikh Images}\label{sec:generalised}
+
+Formally, the \textit{Parikh map} over a context-free language $\Sigma = \left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:
+
+$$
+\begin{aligned}
+& \psi: \Sigma^* \rightarrow \mathbb{N}^k \\
+& \psi(s) = \left[\#a_1(s), \#a_2(s), \ldots, \#a_k(s)\right]
+\end{aligned}
+$$
+
+That is, $\psi(s)$ is a vector of the number of occurrences of each character in the language for a given string $s$.
+For example, for  $\Sigma = \left \{ a, b\right\}$, we would have $\psi(abb) = \left[1, 2\right]$.
+
+We define the image of this map, the \textit{Parikh image}, of some subset of the language $A \subseteq \Sigma^*$ as:
+
+$$
+\psi(A) = \left\{ \psi(x) | x \in A \right\}
+$$
+
+Thus we would have $\psi(\left\{ab, abb\right\}) = \left\{\left[1, 1\right], \left[1, 2\right]\right\}$.
+
+\subsubsection{The Parikh image of a regular language expressed in Presburger arithmetic.}
+Following~\cite{generate-parikh-image}, we define the Parikh Image of a regular language recognised by a DFA $\mathcal{A} =  \langle Q, \Sigma,\delta, I, F \rangle$ as:
+
+$$
+\begin{aligned}
+\psi(\mathcal{A}) := 
+& \bigwedge_{\alpha \in \Sigma}
+c_\alpha = \sum_{\delta \in \delta} t_\delta  
+\text{ where $\alpha \in \delta$}\\
+&\bigwedge_{q \in Q} \left (\text{$1$ if $q \in I$} +
+\sum_{\delta = q' \xrightarrow{} q} t_\delta 
+- \sum_{\delta = q\xrightarrow{}q'} t_\delta \right)
+\begin{cases}
+\geq 0 & \text{if $q \in I$} \\
+= 0 & \text{ otherwise}
+\end{cases}\\
+& \bigwedge_{\delta = q \xrightarrow{} q'} t_\delta > 0 
+\implies z_{'q} > 0 \\
+& \bigwedge_{q, q' \in Q} z_{q'} > 0 
+\implies 
+\begin{cases}
+\bigvee\limits_{\delta = q \xrightarrow{} q'} z_{q'} = z_{q} + 1 \land t_\delta > 0 \land z_{q} > 0 & \text{if $q \not\in  I$} \\
+z_{q'} = 1 \land t_\delta > 0& \text{if $q \in I$}
+\end{cases}
+\end{aligned}
+$$
+
+where all variables $z_i, t_i$ are existentially quantified and the variables $c_\alpha$ make up the actual image.
+
+\subsubsection{Generalising the Parikh image}
+
+Another way of viewing the Parikh map is as a monoid homomorphism $p:\: \left(\Sigma^*, \cdot, \epsilon \right) \to (\mathbb{Z}^\Sigma, +, \vec{0})$, where $\cdot$ is the string concatenation operation, the objects of the right-hand-side monoid are character counts, and $+$ is standard vector addition.
+Note that while the left monoid does not commute, the right one does.
+
+This viewpoint enables us to generalise the Parikh map and its image further to arbitrary monoid morphisms $h:\: \Sigma^* \to M$ where $M$ is a commutative monoid.
+It then follows from the universal mapping property that any such morphism $h$ can also be expressed in terms of the Parikh map, as $h' \circ p$.
+A useful example of such a morphism might be computing the length of a string, which could easily be recast in terms of the Parikh map by summing the individual character counts of the vector.
 
 \contents{
-\item Preliminaries, standard definition of the Parikh image of a
-  regular/context-free language
-\item Generalised version, considering arbitrary homomorphisms
 \item Examples
+\item Fix the formalised version
+\item Verify that the category theory is correct
 }
@@
 
+We assume an NFA $\Automaton = \Tuple{\States, \Alphabet, \Transitions, \Initial,
+\Accepting}$ with $\NrTransitions$ transitions $\Transitions =
+\Set{\EllipsisSequence{\Transitions}{\NrTransitions}}$ where we describe each
+such transition $\Transitions_i$ from node $q$ to node $q'$ with label $\Label \in \Alphabet$ as
+$\Transitions_i = \FromLabelTo{q}{\Label}{q'}$.
+
+Treating $\Automaton$ as a graph with vertices $Q$ and edges $\delta$, we use
+the term \textit{separating cut} of a set of \textit{transitions} $T =
+\Set{\Transitions_i, \ldots, \Transitions_n}, \SeparatingCut(T)$ to refer to any set
+of transitions whose removal causes $T$ to be unreachable from any state in
+$\Initial$, with the meaning that a transition is reachable iff its starting
+node is. Note that if $T$ contains a transition $e =
+\FromLabelTo{v}{\Label}{v'}$ such that $v \in \Initial$, $\SeparatingCut(T) =
+\emptyset$.
+
+We will follow the notation of~\cite{generate-parikh-image} and simultaneously talk
+about $\Automaton$ as a graph and an automaton. Moreover, we will continuously
+refer to the subgraph produced by keeping only the transitions/edges whose
+corresponding variables in $\TransitionVec$ are positive ($> 0$). An edge will
+be called \textit{selected} if it is in this subgraph. An edge that is known to be zero
+will conversely be called \textit{deselected}. An edge whose corrsponding term has no
+known status is called unknown. Formally, we define these as follows:
+
+\begin{definition}
+  $\Selected(x)(\SomeClause)$ means that $x > a$ for some $a \geq 0$ is a clause
+  in $\SomeClause$.
+  \end{definition}
+
+  \begin{definition}
+  $\Deselected(x)(\SomeClause)$ means that $x = 0$ or $x < a$ for some $a \leq 1$
+  is a clause in $\SomeClause$.
+  \end{definition}
+
+  \begin{definition}
+  $\Unknown(x)(\SomeClause) = \lnot\Selected(x)(\SomeClause)\land\lnot\Deselected(x)(\SomeClause)$.
+  \end{definition}
+
+  Additionally, we define a helper predicate:
+
+  \begin{definition}
+    $\Connected(\Automaton)(\SomeClause)$ holds when every $\Selected$ or $\Unknown$ transition in $\Automaton$ is reachable from its initial state.
+  \end{definition}
+
+  With these preliminaries out of the way, we can define our main predicate:
+
+\begin{definition}
+$\PredicateInstance$ is true exactly when:
+\begin{itemize}
+\item $h:\: \Alphabet^* \rightarrow \SomeMonoid$, a morphism to a commutative product monoid $\SomeMonoid = \prod_{i} \SomeMonoid_i$.
+\item $\MonoidElement$ is an element of $\SomeMonoid$.
+\item $\TransitionVec$ is a vector of terms such that each term correspond to a
+  transition in $\Automaton$.
+\end{itemize}
+\end{definition}
+
+To simplify equations, we let $h(t)$ for some transition $t \in \Transitions$ to
+mean the application of $h$ to $t$'s label. Similarly, we use the notation
+$\TransitionVec_t$ to refer to transition $t \in \Transitions$'s corresponding
+term, allowing $\TransitionVec$ to be directly indexed by the transitions
+w.l.o.g from integer indices.
+
+\begin{mathpar}
+  \inferrule*[left=Propagate, right=\textnormal{$C = \SeparatingCut(T)$, $\exists t \in T \::\: \lnot \Deselected(t)$}]
+    {\AndComp
+      {\Transitions_t \in T}
+      {\TransitionVec_{\Transitions_t} = 0} 
+    \land \PredicateInstance \land \\
+    \AndComp{\Transitions_c \in C}{\TransitionVec_{\Transitions_c} = 0} \land \SomeClause}
+    {\PredicateInstance \land \AndComp{\Transitions_c \in C}{\TransitionVec_{\Transitions_c} = 0} \land \SomeClause}
+
+  \inferrule*[left=Expand, right=\textnormal{Precisely once}]
+    {\FlowEq(\Automaton) \land \Connected(\Automaton )\: \land \\ 
+    y = \sum_{t \in \Transitions}{\TransitionVec_t \MonoidProduct h(t)} \land \AndComp{i \in 1,\ldots,t}{\TransitionVec_i \geq 0} \land \PredicateInstance \land \SomeClause}
+    {\PredicateInstance \land \SomeClause}
+
+  \inferrule*[left=Split, right=\textnormal{if $\Unknown(\TransitionVec_i)(\SomeClause)$}]
+  {\TransitionVec_i = 0 \land \PredicateInstance \land \SomeClause \\ | 
+  \\ \TransitionVec_i > 0 \land \PredicateInstance \land \SomeClause}
+    {\PredicateInstance{} \land \SomeClause}
+
+\inferrule*[left=Subsume-Main, right=\textnormal{if no instances of $\Connected(\Automaton)$ in $\SomeClause$}]
+  {\SomeClause}
+  {\PredicateInstance{} \land \SomeClause}
+
+  \inferrule*[left=Subsume-Connected, right=\textnormal{if \KnownConnected}]
+  {\SomeClause}
+  {\Connected({\Automaton}) \land \SomeClause}
+\end{mathpar}
+where $\FlowEq(\Automaton)$ are the flow-balancing part of the Parikh image from Section~\ref{sec:generalised} for an automaton $\Automaton = \Tuple{\States, \Alphabet, \Transitions, \Initial,\Accepting}$:
+
+$$
+\begin{aligned}
+& \In(q) = \text{$1$ if $q \in I$} + \sum_{i \in 1,\ldots,t \SuchThat \Transitions_i = \FromLabelTo{*}{}{q}} \TransitionVec_i\\
+& \Out(q) = \sum_{i \in 1,\ldots,t \SuchThat \Transitions_i = \FromLabelTo{q}{}{*}} \TransitionVec_i\\
+& \FlowEq = \AndComp{q \in Q}{\In(q) - \Out(q)}
+\text{ $\geq 0$ if $q \in F$, $= 0$ otherwise}
+\end{aligned}
+$$
+
+and $\KnownConnected$ corresponds to the following, implying guaranteed
+connectedness (or, conversely, the non-existence of cuts disagreeing with $\TransitionVec$):
+$$
+\forall{C, T}\: C = \SeparatingCut(T) \implies \forall{i} \: \Transitions_i \in T \land \TransitionVec_i > 0 \implies \forall{j} \: \Transitions_j \in C \implies \TransitionVec_j = 0
+$$
+
+Additionally, we assume the existence of a rule \PresburgerClose{},
+corresponding to a sound and complete solver for Presburger formulae.
+
+The $\Propagate{}$ rule allows us to propagate connectedness across
+$\Automaton$. It states that we are only allowed to "use" transitions attached
+to a reachable state, and is necessary to ensure connectedness in the presence
+of cycles. \textsc{Expand} expands the predicate into its most basic rules; one
+set of linear equations connecting $\TransitionVec$ and $\MonoidElement$,
+and the linear flow equations of the standard Parikh image formulation.
+
+Finally, $\Split{}$ allows us to branch on the proof tree by first trying to
+exclude a contested edge from a potential solution and then concluding that it
+must be included.
+
+A decision procedure for our predicate in a tableau-based automated theorem
+prover would start by expanding the predicate using the $\Expand{}$ rule. For
+many instances of the predicate, this would be enough to induce subsumption; as
+as long as the DFA contains no loops that could be disconnected from a minimum
+spanning tree (MST) of the automaton.
+
+\subsection{Generalising to products of automata}
+
+To generalise the calculus to calculations on products of automata, we change the main predicate to arbitrarily many automata:
+\begin{definition}
+  $\ProductPredicateInstance$ is true exactly when:
+  \begin{itemize}
+  \item $h:\: \Alphabet^* \rightarrow \SomeMonoid$, a morphism to a commutative product monoid $\SomeMonoid = \prod_{i} \SomeMonoid_i$.
+  \item $\MonoidElement$ is a vector of elements of $\SomeMonoid$.
+  \item $\Automaton = \prod_{i} \Automaton_i$
+  \item each $\TransitionVec_i$ is a vector of terms such that each term correspond to a
+    transition in $\Automaton_i$.
+  \item $y = \sum_{t \in \Transitions} h(t)$
+  \end{itemize}
+  \end{definition}
+
+  % FIXME y is a vector, and the plus of the sum is on the commutative monoid
+  % FIXME backport this to the previous predicate's definition
+
+  First we extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton:
+  \begin{mathpar}
+    \inferrule*[left=Expand]
+      {\AndComp{i}{\FlowEq(\Automaton_i) \land \Connected(\Automaton_i)}\ldots}
+      {\ProductPredicateInstance \land \SomeClause}
+  \end{mathpar}
+
+\contents{
+\item the rule MATERIALISE allows us to materialise a new partial product
+\item SUBSUME now terminates iff we have fully materialised
+}
+
 \section{Implementation and Experiments}
@@
 %\bibliographystyle{splncs03}
-%\bibliography{refs}
+\printbibliography
 

### 2021-09-23 paper/main.tex
Some writing...

@@
 
-\contents{
-\item key challenges: (1) complicated QE problem that was part of
-  previous definitions; (2) languages with large alphabets described
-  by symbolic automata; (3) regular languages described by products
-  of automata
-}
+Parikh maps and their image appears naturally as part of many operations in model checking and the solving of string constraints in automata-based string solvers such as \Ostrich{}.
+While it is possible to compute the Parikh image of an automaton using the method described in \cite{generate-parikh-image}, this method produces large existentially quantified clauses which are costly to eliminate and in practice make many real-world problems intractable.
+Furthermore, many modern applications require automata with symbolic labels to handle large alphabets \Fudge{citation needed}.
+They also compute Parikh images on products of automata, which would require the computation of the product before the Parikh image can be computed, running the risk of an exponential blow-up.
+
+Addressing these concerns, our approach allows us to extend the computation of Parikh images to handle symbolic transition labels naturally, while also allowing us to interleave the computation of arbitrarily deep products of automata with the Parikh image of their product.
+This allows us to let both calculations inform each other, eliminating unnecessary work.
+
+\Fudge{It's also versatile as heck!}
+
+\subsection{Motivating Example}
 
@@
 \section{Overview}
-
-\subsection{Generalised Parikh Images}\label{sec:generalised}
 
@@
 
-\subsubsection{Generalising the Parikh image}
+\subsection{An Example}
+
+\subsection{Generalised Parikh Images}\label{sec:generalised}
 
@@
 It then follows from the universal mapping property that any such morphism $h$ can also be expressed in terms of the Parikh map, as $h' \circ p$.
-A useful example of such a morphism might be computing the length of a string, which could easily be recast in terms of the Parikh map by summing the individual character counts of the vector.
 
-\contents{
-\item Examples
-\item Fix the formalised version
-\item Verify that the category theory is correct
-}
+A useful example of such a morphism might be computing the length of a string, which could be recast in terms of the Parikh map by summing the individual character counts of the vector: $h':\: (\mathbb{Z}^\Sigma, +, \vec{0}) \to (\mathbb{Z}, +, 0) = \vec{x} \to \sum_{i \in \Sigma} x_i$, where the respective operators $+$ is element-wise vector addition and integer addition respectively.
+
+\subsubsection{An Example}
 
@@
 
-\section{A Calculus of Parikh Images}
-
 \contents{
@@
 
-\section{Parikh Images from Products of Automata}
-
   \contents{
-\item additional rules needed for products
-\item backjumping and conflict-driven learning
+  \item remove the monoid map!!!
 }
@@
 the term \textit{separating cut} of a set of \textit{transitions} $T =
-\Set{\Transitions_i, \ldots, \Transitions_n}, \SeparatingCut(T)$ to refer to any set
-of transitions whose removal causes $T$ to be unreachable from any state in
-$\Initial$, with the meaning that a transition is reachable iff its starting
-node is. Note that if $T$ contains a transition $e =
-\FromLabelTo{v}{\Label}{v'}$ such that $v \in \Initial$, $\SeparatingCut(T) =
-\emptyset$.
+\Set{\Transitions_i, \ldots, \Transitions_n}, \SeparatingCut(T)$ to refer to any
+set of transitions whose removal causes $T$ to be unreachable from any state in
+$\Initial$, with the meaning that $e = \FromLabelTo{v}{\Label}{v'}$ is reachable
+if $v$ is. Note that if $T$ contains a transition  such that $v \in \Initial$,
+$\SeparatingCut(T) = \emptyset$, since we cannot disconnect an initial state by
+removing any transition.
 
@@
 be called \textit{selected} if it is in this subgraph. An edge that is known to be zero
-will conversely be called \textit{deselected}. An edge whose corrsponding term has no
+will conversely be called \textit{deselected}. An edge whose corresponding term has no
 known status is called unknown. Formally, we define these as follows:
 
+TODO what happens to $\TransitionVec$ now that we have $\TransitionMask$?
+
+TODO how do we work around the presburger logic here?
+
 \begin{definition}
-  $\Selected(x)(\SomeClause)$ means that $x > a$ for some $a \geq 0$ is a clause
-  in $\SomeClause$.
+  $\Selected(\Automaton, t)(\SomeClause)$ means that $\SomeClause$ contains $\TransitionMask(\Automaton, t, x)$ and that $\SomeClause \vDash_{\textsc{Presburger}} x > 0$.
   \end{definition}
@@
   \begin{definition}
-  $\Deselected(x)(\SomeClause)$ means that $x = 0$ or $x < a$ for some $a \leq 1$
-  is a clause in $\SomeClause$.
+    $\Deselected(\Automaton, t)(\SomeClause)$ means that $\SomeClause$ contains $\TransitionMask(\Automaton, t, x)$ and that $\SomeClause \vDash_{\textsc{Presburger}} x \leq 0$.
     \end{definition}
@@
   \begin{definition}
-  $\Unknown(x)(\SomeClause) = \lnot\Selected(x)(\SomeClause)\land\lnot\Deselected(x)(\SomeClause)$.
+  $\Unknown(\Automaton, t)(\SomeClause) = \lnot\Selected(\Automaton, t)(\SomeClause)\land\lnot\Deselected(\Automaton, t)(\SomeClause)$, while $\TransitionMask(\Automaton, t, x)$ is still a term in $\SomeClause$.
   \end{definition}
 
-  Additionally, we define a helper predicate:
+  For all of these cases, we say that a transition $t$ is $\Selected,
+  \Deselected$ or $\Unknown$ respectively, usually leaving out the clause
+  $\SomeClause$ and automaton $\Automaton$. Additionally, we define two helper
+  predicates, one of which we have already used:
 
   \begin{definition}
-    $\Connected(\Automaton)(\SomeClause)$ holds when every $\Selected$ or $\Unknown$ transition in $\Automaton$ is reachable from its initial state.
+    $\Connected(\Automaton)$ holds when every $\Selected$ or $\Unknown$ transition in $\Automaton$ is reachable from its initial state.
+  \end{definition}
+
+  \begin{definition}
+    $\TransitionTerm(\Automaton, t, x)$ holds when $\Automaton$'s transition $t$ is taken $x$ times (an integer value).
   \end{definition}
@@
 w.l.o.g from integer indices.
+We will also consistently assume that we have only one instance of our main predicate.
+In the actual implementation, described in \Fudge{the north pole}, we will use an additional, existentially quantified, integer variable to identify the instance of a predicate.
 
@@
 prover would start by expanding the predicate using the $\Expand{}$ rule. For
-many instances of the predicate, this would be enough to induce subsumption; as
+many instances of the predicate, this would be enough to induce subsumption;
 as long as the DFA contains no loops that could be disconnected from a minimum
@@
 
-\subsection{Generalising to products of automata}
+\subsection{An Example}
 
-To generalise the calculus to calculations on products of automata, we change the main predicate to arbitrarily many automata:
+\section{Generalising the calculus modulo arbitrary commutative monoid maps}
+
+\subsection{An Example}
+
+\section{Parikh Images from Products of Automata}
+
+\contents{
+    \item additional rules needed for products
+    \item backjumping and conflict-driven learning
+    }
+
+To generalise the calculus to calculations on products of automata, we change the main predicate take arbitrarily many automata:
 \begin{definition}
@@
   \item $\MonoidElement$ is a vector of elements of $\SomeMonoid$.
-  \item $\Automaton = \prod_{i} \Automaton_i$
+  \item $\AutomatonVector$ is a vector of automata.
+  \item $\Automaton = \prod_{i} \AutomatonVector_i$
   \item each $\TransitionVec_i$ is a vector of terms such that each term correspond to a
-    transition in $\Automaton_i$.
+    transition in $\AutomatonVector_i$, and $\TransitionVectors$ is a vector of those vectors.
+We use the terminology $\TransitionTerm(t)$ to refer to the term of the transition vector for a given transition $t$.
   \item $y = \sum_{t \in \Transitions} h(t)$
+  \item $\SourceTransitions(\Automaton, t)$ is the set of transitions from the
+  term automata used to produce the transition $t$ in the product automaton
+  $\Automaton$.
   \end{itemize}
@@
     \inferrule*[left=Expand]
-      {\AndComp{i}{\FlowEq(\Automaton_i) \land \Connected(\Automaton_i)}\ldots}
+      {\AndComp{\Automaton_i \in \AutomatonVector}{\FlowEq(\Automaton_i) \land \Connected(\Automaton_i)}\ldots}
       {\ProductPredicateInstance \land \SomeClause}
@@
 
-\contents{
-\item the rule MATERIALISE allows us to materialise a new partial product
-\item SUBSUME now terminates iff we have fully materialised
-}
+  %  x(t) = sum(e : termProductEdges(t, default=0))
+Then we introduce the rule $\Materialise$, used to compute a partial product between two terms $\Automaton_i, \Automaton_j$:
+\begin{mathpar}
+  \inferrule*[left=Materialise, 
+  right=\textnormal{where $\Automaton_{k} = \Automaton_i \times \Automaton_j, \AutomatonVector' = (\AutomatonVector \cup \{  \Automaton_k \}) \setminus {\Automaton_i, \Automaton_j}$}]
+    {\AndComp{t \in \delta(\Automaton_{k})}{\TransitionTerm(t) = \sum_{t' \in \SourceTransitions(t, \Automaton_k)} t'}\\
+    \FlowEq(\Automaton_{k}) \land \Connected(\Automaton_{k}) \land
+    \\ \MonoidMap(\AutomatonVector', \TransitionVectors', \MonoidElement) \land \SomeClause} {\MonoidMap(\AutomatonVector, \TransitionVectors, \MonoidElement) \land \SomeClause}
+\end{mathpar}
+
+Note that we implicitly existentially quantify the fresh transition terms $\TransitionVec_{k+1}$ for the product.
+
+$\Materialise$ comes with its companion rule $\RemoveTermConnected$ that simply
+states we are allowed to remove any instances of $\Connected$ from terms that
+have been used in a product, and whose connectedness we therefore no longer care about:
+\begin{mathpar}
+  \inferrule*[left=$\RemoveTermConnected$, right=\textnormal{If $\Automaton_a$ not in $\Automaton_1\ldots\Automaton_k$}]
+    {\ProductPredicateInstance \land \SomeClause}
+    {\Connected(\Automaton_a) \land \ProductPredicateInstance \land \SomeClause}
+\end{mathpar}
+
+Finally, we add a side condition to $\Subsume$ to only allow subsumption for
+instances of $\MonoidMap$ with just one automaton, ensuring that we do not bail
+out before having computed the entire product.
+
+\subsection{An Example}
+
+\section{Solving the Travelling Salesperson Problem}
+
+If we take an instance of the (decision version of the) TSP as a graph with vertices $N$, edges $E$ and a distance function $\delta(e)$, we can solve it using our method in the following way:
+\begin{enumerate}
+  \item Construct an automaton $\Automaton$ such that all its states are accepting, all the vertices and edges from the TSP graph are present, and the initial state has a transition to each state with weight 0.
+This encodes the movements in the graph, capturing the constraint that we may take any path through the graph, and that we may end anywhere.
+  \item Use a monoid morphism to an $n +1$-element vector, where the first element is the length of the current path, and the $n = |N|$ elements following it is how many times we visit each node.
+  \item The monoid morphism $m(\langle a, w, b\rangle) = \left[w, 0, \ldots, 1, \ldots, 0 \right] $ would simply map to the transition weight $w$ followed by a 1 for node $b$ and zeroes for all other nodes.
+The corresponding (commutative) monoid operation would then be simple elementwise addition of the bookkeeping vector.
+  \item Constrain the terms corresponding to the visited status for each state to be at least 1.
+  \item Iteratively query the solver for a looser and looser upper bound on the length. Once we have a solution, we know it to be minimal.
+\end{enumerate}
 

### 2021-10-06 abstract/main.tex
Add abstract

--- 2021-02-15 paper/main.tex
@@
 
-\title{Generalised Parikh Images of Regular Languages}
+\title{A Solver for Generalised Parikh Images of Conjunctions of Regular Languages}
 \author{some cool authors}
@@
 
-\begin{abstract}
-  bla
-\end{abstract}
-
 \section{Introduction}
 
-\contents{
-\item key challenges: (1) complicated QE problem that was part of
-  previous definitions; (2) languages with large alphabets described
-  by symbolic automata; (3) regular languages described by products
-  of automata
-}
+Parikh maps and their image appears naturally as part of many operations in
+model checking and the solving of string constraints in automata-based string
+solvers such as \Ostrich{} \cite{ostrich}. While it is possible to compute the Parikh image of
+an automaton using the method described in \cite{generate-parikh-image}, this
+method produces large existentially quantified clauses which are costly to
+eliminate and in practice make many real-world problems intractable.
+Furthermore, many modern applications require automata with symbolic labels to
+handle large alphabets, which has no immediate correspondence in the methods described. Automata-based string solvers like \Ostrich{} also compute
+Parikh images on products of automata, derived from conjunctions of string
+constraints. Under previous methods this would require the computation of the
+product before the Parikh image can be computed, running the risk of an
+exponential blow-up.
 
-\subsection{Related Work}
+Addressing these concerns, our approach allows us to extend the computation of
+Parikh images to handle symbolic transition labels naturally, while also
+allowing us to interleave the computation of arbitrarily deep products of
+automata with the Parikh image of their product. This allows us to let both
+calculations inform each other, allowing the computation of partial products to
+inform the computation of the product and vice versa. Moreover, the scheme
+allows us to learn interesting facts (implied clauses) about the problem.
 
-\section{Overview}
+Based on these insights, we present a tool to solve linear constraints on
+symbolic automata with counters, amounting to constraints on the Parikh image of
+their product. We are able to generate symbolic as well as concrete Parikh images.
 
-\subsection{Generalised Parikh Images}
+\section{Background}
 
-\contents{
-\item Preliminaries, standard definition of the Parikh image of a
-  regular/context-free language
-\item Generalised version, considering arbitrary homomorphisms
-\item Examples
-}
+Formally, the \textit{Parikh map} over a context-free language $\Sigma = \left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:
 
-\subsection{Lazy Computation of Parikh Images for Regular Languages}
+$$
+\begin{aligned}
+& \psi: \Sigma^* \rightarrow \mathbb{N}^k \\
+& \psi(s) = \left[\#a_1(s), \#a_2(s), \ldots, \#a_k(s)\right]
+\end{aligned}
+$$
 
-\contents{
-\item Lazy expansion of the Parikh conditions for a symbolic automaton
-\item Finding elements vs.\ computing the complete Parikh image
-\item Lazy product computation
-}
+That is, $\psi(s)$ is a vector of the number of occurrences of each character in the language for a given string $s$.
+For example, for  $\Sigma = \left \{ a, b\right\}$, we would have $\psi(abb) = \left[1, 2\right]$.
 
-\section{A Calculus of Parikh Images}
+We define the image of this map, the \textit{Parikh image}, of some subset of the language $A \subseteq \Sigma^*$ as:
 
-\contents{
-\item Preliminaries, the underlying calculus, what are rules
-\item Predicates used to represent Parikh images
-\item Our calculus rules
-\item Statement of properties, correctness, complexity
-}
+$$
+\psi(A) = \left\{ \psi(x) | x \in A \right\}
+$$
 
-\section{Parikh Images from Products of Automata}
+Thus we would have $\psi(\left\{ab, abb\right\}) = \left\{\left[1, 1\right], \left[1, 2\right]\right\}$.
 
-\contents{
-\item additional rules needed for products
-\item backjumping and conflict-driven learning
-}
+\subsection{Generalised Parikh Images}\label{sec:generalised}
 
-\section{Implementation and Experiments}
+Another way of viewing the Parikh map is as a monoid homomorphism $p:\: \left(\Sigma^*, \cdot, \epsilon \right) \to (\mathbb{Z}^\Sigma, +, \vec{0})$, where $\cdot$ is the string concatenation operation, the objects of the right-hand-side monoid are character counts, and $+$ is standard vector addition.
+Note that while the left monoid does not commute, the right one does.
 
-\contents{
-\item length constraints
-\item Parikh automata, automata with registers
-\item Model-checking examples
-}
+This viewpoint enables us to generalise the Parikh map and its image further to arbitrary monoid morphisms $h:\: \Sigma^* \to M$ where $M$ is a commutative monoid.
+It then follows from the universal mapping property that any such morphism $h$ can also be expressed in terms of the Parikh map, as $h' \circ p$.
 
-\section{Conclusions}
+A useful example of such a morphism might be computing the length of a string,
+which could be recast in terms of the Parikh map by summing the individual
+character counts of the vector: $h':\: (\mathbb{Z}^\Sigma, +, \vec{0}) \to
+(\mathbb{Z}, +, 0) = \vec{x} \to \sum_{i \in \Sigma} x_i$, where the respective
+operators $+$ is element-wise vector addition and integer addition respectively.
+This allows us to apply selective logics, symbolic logics, and arithmetic on
+counter variabels to the Parikh image calculus.
 
-\clearpage
+\section{Lazy Computation of Parikh Images for Regular Languages}
+
 %\bibliographystyle{splncs03}
-%\bibliography{refs}
+\printbibliography
 

### 2021-10-07 abstract/main.tex
Update abstract

@@
 This allows us to apply selective logics, symbolic logics, and arithmetic on
-counter variabels to the Parikh image calculus.
+counter variables to the Parikh image calculus.
 
 \section{Lazy Computation of Parikh Images for Regular Languages}
+
+Starting with the automaton in~\ref{fig:start}, we seek any value in its Parikh image, i.e. a character count.
+
+\begin{figure}[h]
+  \centering
+  \label{fig:start}
+  \includegraphics[width=0.5\textwidth]{trace-0}
+  \end{figure}
+
+  In this case, our calculation will introduce a free (and fresh) integer constant per character ($a$, $b$, $c$, and $d$).
+It will then start by introducing equations preserving flow through the automaton corresponding to how many times that transition is used in a path from the initial state to an accepting state.
+Each transition will be given an existentially quantified variable in these equations, and each of our target variables set equal to the sum of their corresponding transitions.
+This corresponds to the first portion of the formula described in~\cite{generate-parikh-image}, but crucially misses the constraints to ensure connectivity of a path in the presence of cycles.
+Instead, we enforce these constraints lazily using our own calculus rules to.
+
+  Note that this formulation allows us to propagate information between the terms of the product without (and before) computing the full product.
+For many unsatisfiable instances, this corresponds to the method recently introduced in~\cite{approximate-parikh}, where the Parikh-image of a product of automata is over-approximated to the conjunction of their Parikh images.
+
+Before computation using our rules begin, our automated theorem prover performs reasoning on the linear flow equations to simplify them into:
+$$
+\exists_{x_{0}} : d + a = 1 \land
+c = a
+$$
+
+Note the disappearance of all but one of the transition variables and their replacement with the corresponding linear equations:
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\textwidth]{trace-0-aut-0}
+\end{figure}
+
+We now have a choice of two paths through the automaton and execute a case split: \textsc{Split}: $a \leq 0$ $\mid$ $a > 0$.
+The split is performed as close to the initial state as possible and favours deselecting an edge by constraining its corresponding term to be 0.
+
+Since this decision disconnects a loop, the $b$~transition, from the initial state, we evaluate the rule \textsc{Propagate-Connected} and add the constraint $b = 0$.
+Following this, we can also subsume the propagating machinery for the connectedness constraint, as there are no nonzero loop transitions left.
+This leaves us with an automaton with just one nonzero transition, as in figure~\ref{fig:final}.
+
+\begin{figure}[h]
+  \centering
+  \label{fig:final}
+  \includegraphics[width=\textwidth]{trace-3-aut-0}
+  \end{figure}
+
+Since this automaton is trivially connected and allows for only one solution, we
+present the solution $a = b = c = 0, d=1$.
+
+\section{Possible Applications}
+Bring your own!
 

### 2021-10-10 abstract/main.tex
Some last-minute revisions

--- 2021-10-06 abstract/main.tex
@@
 \usepackage{mathpartir}
+\usepackage{csquotes}
+\usepackage[font={small,it}]{caption}
 \addbibresource{../bibliography.bib}
@@
 \allowdisplaybreaks[1]
-
-\title{A Solver for Generalised Parikh Images of Conjunctions of Regular Languages}
-\author{some cool authors}
+\title{A Solver for Generalised Parikh Images of Regular Languages}
+\author{Philipp Rümmer \& Amanda Stjerna}
 \institute{Uppsala University, Sweden}
@@
 
-Parikh maps and their image appears naturally as part of many operations in
-model checking and the solving of string constraints in automata-based string
-solvers such as \Ostrich{} \cite{ostrich}. While it is possible to compute the Parikh image of
+Parikh images appears naturally as part of many operations in model checking or
+solving of string constraints in automata-based string solvers such as
+\Ostrich{} \cite{ostrich}. While it is possible to compute the Parikh image of
 an automaton using the method described in \cite{generate-parikh-image}, this
-method produces large existentially quantified clauses which are costly to
-eliminate and in practice make many real-world problems intractable.
-Furthermore, many modern applications require automata with symbolic labels to
-handle large alphabets, which has no immediate correspondence in the methods described. Automata-based string solvers like \Ostrich{} also compute
-Parikh images on products of automata, derived from conjunctions of string
-constraints. Under previous methods this would require the computation of the
-product before the Parikh image can be computed, running the risk of an
-exponential blow-up.
+method produces clauses with many existentially quantified variables which are
+costly to eliminate, in practice making many real-world applications
+intractable. Furthermore, many applications require automata with symbolic
+labels to handle large alphabets like Unicode which would increase the number of
+variables even more. Finally, automata-based string solvers compute Parikh
+images on products of automata, derived from conjunctions of string constraints.
+Using \cite{generate-parikh-image} this would require the up-front computation
+of the product before its Parikh image, running the risk of an exponential
+blow-up.
 
 Addressing these concerns, our approach allows us to extend the computation of
-Parikh images to handle symbolic transition labels naturally, while also
-allowing us to interleave the computation of arbitrarily deep products of
-automata with the Parikh image of their product. This allows us to let both
-calculations inform each other, allowing the computation of partial products to
-inform the computation of the product and vice versa. Moreover, the scheme
-allows us to learn interesting facts (implied clauses) about the problem.
+Parikh images to handle symbolic transition labels ergonomically, while also
+allowing us to interleave the computations of the Parikh image of a product and
+the product itself. This allows both calculations to inform each other, thereby
+eliminating unneccesary work. Moreover, the scheme allows us to learn
+interesting facts (in the form of implied clauses) about the problem.
 
-Based on these insights, we present a tool to solve linear constraints on
-symbolic automata with counters, amounting to constraints on the Parikh image of
-their product. We are able to generate symbolic as well as concrete Parikh images.
+Based on these insights, we present a work-in-progress tool to solve linear
+constraints on symbolic automata with counters, amounting to constraints on the
+Parikh image of their product. We are able to generate symbolic (\enquote{there are
+twice as many \texttt{a}:s as \texttt{b}:s}) as well as concrete Parikh images.
 
@@
 character counts of the vector: $h':\: (\mathbb{Z}^\Sigma, +, \vec{0}) \to
-(\mathbb{Z}, +, 0) = \vec{x} \to \sum_{i \in \Sigma} x_i$, where the respective
-operators $+$ is element-wise vector addition and integer addition respectively.
-This allows us to apply selective logics, symbolic logics, and arithmetic on
-counter variabels to the Parikh image calculus.
+(\mathbb{Z}, +, 0) = \vec{x} \to \sum_{i \in \Sigma} x_i$. This allows us to
+apply e.g. arithmetic on register variables to the Parikh image calculus.
 
 \section{Lazy Computation of Parikh Images for Regular Languages}
+
+Starting with the automaton shown to the left in figure~\ref{fig:automaton}, we seek
+any value in its Parikh image, i.e. a character count.
+
+\begin{figure}[h]
+  \caption{Left to right: the starting automaton, the automaton with its corresponding linear equations, and the solution.
+Note the replacement of the existentially quantified transition variables with character count constants, and the introduction of a linear equation across the edge for \texttt{d} representing a choice between either an~\texttt{a} or a~\texttt{d}.
+Transitions with a zero value are not used in the solution.}\label{fig:automaton}
+  \begin{minipage}[t]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{trace-0}
+    \end{minipage}
+    \begin{minipage}[t]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{trace-0-aut-0}
+    \end{minipage}
+    \begin{minipage}[t]{0.3\textwidth}
+      \includegraphics[width=\textwidth]{trace-3-aut-0}
+      \end{minipage}
+  \end{figure}
+
+  In this case, our calculation will introduce a free integer constant per
+  character ($a$, $b$, $c$, and $d$). It will then introduce equations
+  preserving flow through the automaton corresponding to how many times that
+  transition is used in a run of the automaton. Each transition will be given an existentially
+  quantified variable in these equations, and each of our target variables is
+  constrained to the sum of the transitions where it occurs as a label. This
+  corresponds to the first portion of the formula described
+  in~\cite{generate-parikh-image}, but crucially misses the constraints to
+  ensure connectedness in the presence of cycles. Instead, we enforce this
+  constraint lazily using our own deduction rules.
+
+  Note that this formulation allows us to propagate information between terms of
+  a product before materialising it. For many unsatisfiable instances, this
+  corresponds to the method recently introduced in~\cite{approximate-parikh},
+  where the Parikh image of a product of automata is over-approximated as the
+  conjunction of their Parikh images.
+
+Before computation using our rules begin, our prover performs reasoning on the flow equations to simplify them into $d + a = 1 \land c = a$.
+
+We now have a choice of two paths for our run and execute a case split: \textsc{Split}: $a \leq 0$ $\mid$ $a > 0$.
+The split is performed as close to the initial state as possible and favours deselecting an edge by constraining its corresponding term to be 0.
+
+Since this decision disconnects a loop, the $b$~transition, from the initial state, we evaluate the rule \textsc{Propagate-Connected} to add the constraint $b = 0$.
+Afterwards we can subsume the propagating machinery for the connectedness constraint, as no more loop transitions can be positive.
+This leaves us with an automaton with just one nonzero transition, as seen in figure~\ref{fig:automaton}.
+
+As there is just one automaton left, there are no products to compute and this the
+calculation is complete. We present the solution $a = b = c = 0, d=1$.
 

### 2021-10-11 abstract/main.tex
Update abstract

@@
-\documentclass[runningheads]{llncs}
+\documentclass{easychair}
 
@@
 \title{A Solver for Generalised Parikh Images of Regular Languages}
-\author{Philipp Rümmer \& Amanda Stjerna}
+\author{Amanda Stjerna \and Philipp Rümmer}
 \institute{Uppsala University, Sweden}
@@
 
-Parikh images appears naturally as part of many operations in model checking or
-solving of string constraints in automata-based string solvers such as
-\Ostrich{} \cite{ostrich}. While it is possible to compute the Parikh image of
-an automaton using the method described in \cite{generate-parikh-image}, this
-method produces clauses with many existentially quantified variables which are
-costly to eliminate, in practice making many real-world applications
-intractable. Furthermore, many applications require automata with symbolic
-labels to handle large alphabets like Unicode which would increase the number of
-variables even more. Finally, automata-based string solvers compute Parikh
-images on products of automata, derived from conjunctions of string constraints.
-Using \cite{generate-parikh-image} this would require the up-front computation
-of the product before its Parikh image, running the risk of an exponential
-blow-up.
+Parikh images have a wide application within automata, including for the solving
+of string constraints in automata-based solvers such as \Ostrich{}
+\cite{ostrich}. While it is possible to compute the Parikh image of an automaton
+by adapting the method for CFGs of~\cite{generate-parikh-image}, this method
+produces clauses with many existentially quantified variables which are costly
+to eliminate. This makes many real-world applications intractable. Furthermore,
+applications like Unicode require automata with symbolic labels to handle their
+large alphabets, increasing the number of variables even more.
+Finally, automata-based string solvers compute Parikh images on products of
+automata, derived from conjunctions of string constraints. Using
+\cite{generate-parikh-image} this would require the up-front computation of the
+product before its Parikh image, running the risk of an exponential blow-up.
 
-Addressing these concerns, our approach allows us to extend the computation of
-Parikh images to handle symbolic transition labels ergonomically, while also
-allowing us to interleave the computations of the Parikh image of a product and
-the product itself. This allows both calculations to inform each other, thereby
-eliminating unneccesary work. Moreover, the scheme allows us to learn
-interesting facts (in the form of implied clauses) about the problem.
+Addressing these concerns, we introduce a lazy approach to the computation of
+Parikh images of products of automata that also generalises operations on
+labels. This allows us to extend the computation of Parikh images to handle
+symbolic transition labels ergonomically, while also allowing us to interleave
+the computations of the Parikh image of a product and the product itself. This
+allows both calculations to inform each other, thereby eliminating unnecessary
+work. Moreover, the scheme allows us to learn interesting facts (in the form of
+implied clauses) about the problem.
 
 Based on these insights, we present a work-in-progress tool to solve linear
-constraints on symbolic automata with counters, amounting to constraints on the
-Parikh image of their product. We are able to generate symbolic (\enquote{there are
-twice as many \texttt{a}:s as \texttt{b}:s}) as well as concrete Parikh images.
+constraints on symbolic automata with counters, aka Parikh automata, amounting
+to constraints on the Parikh image of their product. We are able to generate
+both a Presburger formula representing the image and find satisfying assignments
+within it.
 
@@
 
-Thus we would have $\psi(\left\{ab, abb\right\}) = \left\{\left[1, 1\right], \left[1, 2\right]\right\}$.
+Thus, we would have $\psi(\left\{ab, abb\right\}) = \left\{\left[1, 1\right], \left[1, 2\right]\right\}$.
+An interesting property of this map is that it can always be expressed as a symbolic Presburger formula.
 
@@
 
-Another way of viewing the Parikh map is as a monoid homomorphism $p:\: \left(\Sigma^*, \cdot, \epsilon \right) \to (\mathbb{Z}^\Sigma, +, \vec{0})$, where $\cdot$ is the string concatenation operation, the objects of the right-hand-side monoid are character counts, and $+$ is standard vector addition.
+Another way of viewing the Parikh map is as a monoid homomorphism $p:\: \left(\Sigma^*, \cdot, \epsilon \right) \to (\mathbb{Z}^\Sigma, +, \vec{0})$, where $\cdot$ is the string concatenation operation, the objects of the right-hand-side monoid are character counts, and $+$ is element-wise vector addition.
 Note that while the left monoid does not commute, the right one does.
 
-This viewpoint enables us to generalise the Parikh map and its image further to arbitrary monoid morphisms $h:\: \Sigma^* \to M$ where $M$ is a commutative monoid.
+This viewpoint enables us to generalise the Parikh map further to arbitrary monoid morphisms $h:\: \Sigma^* \to M$ where $M$ is a commutative monoid.
 It then follows from the universal mapping property that any such morphism $h$ can also be expressed in terms of the Parikh map, as $h' \circ p$.
 
-A useful example of such a morphism might be computing the length of a string,
-which could be recast in terms of the Parikh map by summing the individual
-character counts of the vector: $h':\: (\mathbb{Z}^\Sigma, +, \vec{0}) \to
-(\mathbb{Z}, +, 0) = \vec{x} \to \sum_{i \in \Sigma} x_i$. This allows us to
-apply e.g. arithmetic on register variables to the Parikh image calculus.
+An example is computing the length of a string, which is expressed in terms of
+the Parikh map by summing the individual character counts of the vector: $h':\:
+(\mathbb{Z}^\Sigma, +, \vec{0}) \to (\mathbb{Z}, +, 0) = \vec{x} \to \sum_{i \in
+\Sigma} x_i$. More generally, this approach extends to multiple counters to
+produce Parikh automata with one such counter per symbol in the language, as
+used in implementations of many string
+constraints~\cite{path-feasibility-strings}.
 
 \section{Lazy Computation of Parikh Images for Regular Languages}
+The calculus is based on four principles: linear equations to preserve flow
+through the automaton, case-splitting when there are multiple possible paths a
+run can take, propagation of connectedness constraints, and materialisation of
+products. Notably, ensuring connectedness of a run in the presence of cycles is
+the computationally most difficult constraint to enforce, as the flow equations
+of any loop will self-balance.
 
-Starting with the automaton shown to the left in figure~\ref{fig:automaton}, we seek
-any value in its Parikh image, i.e. a character count.
+Starting with the automaton shown to the left in figure~\ref{fig:automaton}, we
+seek a satisfying assignment as an illustration of our calculus.
 
@@
 
-  In this case, our calculation will introduce a free integer constant per
+  In this case, our calculus will introduce a free integer constant per
   character ($a$, $b$, $c$, and $d$). It will then introduce equations
   preserving flow through the automaton corresponding to how many times that
-  transition is used in a run of the automaton. Each transition will be given an existentially
-  quantified variable in these equations, and each of our target variables is
-  constrained to the sum of the transitions where it occurs as a label. This
-  corresponds to the first portion of the formula described
+  transition is used in a run. Each transition will be given an
+  existentially quantified variable in these equations, and each of our target
+  variables is constrained to the sum of the transitions where it occurs as a
+  label. This corresponds to the first portion of the formula described
   in~\cite{generate-parikh-image}, but crucially misses the constraints to
   ensure connectedness in the presence of cycles. Instead, we enforce this
-  constraint lazily using our own deduction rules.
+  constraint lazily.
 
   Note that this formulation allows us to propagate information between terms of
-  a product before materialising it. For many unsatisfiable instances, this
-  corresponds to the method recently introduced in~\cite{approximate-parikh},
-  where the Parikh image of a product of automata is over-approximated as the
-  conjunction of their Parikh images.
+  a product before materialising it to exclude infeasible parts of a product from the computation.
 
-Before computation using our rules begin, our prover performs reasoning on the flow equations to simplify them into $d + a = 1 \land c = a$.
+Before computation begins, our prover performs reasoning on the flow equations
+to simplify them into $d + a = 1 \land c = a$. This leaves us with a choice of
+two paths for our run so we execute a case split: \textsc{Split}: $a \leq 0$
+$\mid$ $a > 0$. The split is performed as close to the initial state as possible
+and favours deselecting an edge by constraining its corresponding term to be 0.
+The linear inequalities of the flow equations will ensure that the choice
+selects precisely one of the upper and lower paths.
 
-We now have a choice of two paths for our run and execute a case split: \textsc{Split}: $a \leq 0$ $\mid$ $a > 0$.
-The split is performed as close to the initial state as possible and favours deselecting an edge by constraining its corresponding term to be 0.
+Since choosing the \texttt{d}~transition over \texttt{a} disconnects a loop, the \texttt{b}~transition, from the initial state, we evaluate the rule \textsc{Propagate-Connected} to add the constraint $b = 0$ and exclude the unreachable \texttt{b} from our run.
+Afterwards we can subsume the propagator for the connectedness constraint as all loop transitions terms are negative, and our run is therefore guaranteed to be connected.
+This leaves us with an automaton with just one nonzero transition, as seen in the rightmost image of figure~\ref{fig:automaton}.
 
-Since this decision disconnects a loop, the $b$~transition, from the initial state, we evaluate the rule \textsc{Propagate-Connected} to add the constraint $b = 0$.
-Afterwards we can subsume the propagating machinery for the connectedness constraint, as no more loop transitions can be positive.
-This leaves us with an automaton with just one nonzero transition, as seen in figure~\ref{fig:automaton}.
-
-As there is just one automaton left, there are no products to compute and this the
-calculation is complete. We present the solution $a = b = c = 0, d=1$.
+As there is just one automaton left, there are no products to materialise and so
+the calculation is complete. We present the assignment $a = b = c = 0, d=1$. To
+find the Parikh image rather than an assignment, we would iteratively query
+our calculus for partial solutions and perform quantifier elimination on the
+free constants.
 

### 2021-10-11 abstract/main.tex
improved intro a bit

@@
 
-Parikh images have a wide application within automata, including for the solving
+Parikh images are a central notion in automata theory and context-free grammars,
+with wide-ranging applications, including the solving
 of string constraints in automata-based solvers such as \Ostrich{}
-\cite{ostrich}. While it is possible to compute the Parikh image of an automaton
-by adapting the method for CFGs of~\cite{generate-parikh-image}, this method
+\cite{ostrich}. In this work, we focus on the efficient computation
+of Parikh images, or points within the Parikh image, of finite-state
+automata. While it is possible to compute the Parikh image of an automaton
+by adapting the method for context-free grammars
+ of~\cite{generate-parikh-image}, this method
 produces clauses with many existentially quantified variables which are costly

### 2021-10-11 abstract/main.tex
conclusions

@@
 
-Parikh images are a central notion in automata theory and context-free grammars,
+Parikh images are a central notion in automata and language theory,
 with wide-ranging applications, including the solving
-of string constraints in automata-based solvers such as \Ostrich{}
-\cite{ostrich}. In this work, we focus on the efficient computation
+of string constraints in automata-based solvers such as \Ostrich{}~\cite{ostrich}. In this work, we focus on the efficient computation
 of Parikh images, or points within the Parikh image, of finite-state
@@
 produce Parikh automata with one such counter per symbol in the language, as
-used in implementations of many string
-constraints~\cite{path-feasibility-strings}.
+used in some string
+constraint solvers~\cite{path-feasibility-strings}.
 
 \section{Lazy Computation of Parikh Images for Regular Languages}
+In the scope of this abstract, we introduce our calculus for computing
+Parikh images through an example.
 The calculus is based on four principles: linear equations to preserve flow
@@
 
+\section{Conclusions}
+
+We are currently in the process of finishing the first version of our
+Parikh image tool. As an evaluation, we will use the tool to compute
+length abstractions of regular expression constraints in string
+solving, and to handle the automata augmented with counters introduced
+in \cite{path-feasibility-strings}. We will compare with the approach
+proposed in \cite{generate-parikh-image}, and with a translation to a
+model checking problem, which can then be solved using (more general)
+symbolic and abstraction-based model checkers.
+
 %\bibliographystyle{splncs03}

### 2022-03-11 paper/main.tex
Some rules for one automaton

--- 2021-02-15 paper/main.tex
@@
-\documentclass[runningheads]{llncs}
+\documentclass{easychair}
 
+\usepackage{amsthm}
 \usepackage{amssymb,amsmath,mathtools,empheq,fancybox}
@@
 \usepackage{textcomp,listings}
+\usepackage[backend=biber]{biblatex}
+\usepackage{mymacros}
+
+\usepackage{mathpartir}
+\addbibresource{../bibliography.bib}
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
 
@@
 
-\title{Generalised Parikh Images of Regular Languages}
+\title{An Efficient Calculus for Generalised Parikh Images of Regular Languages}
 \author{some cool authors}
@@
 \begin{abstract}
-  bla
+  The image of the Parikh map is important in automata theory, offering a
+  compact characterisation of an automaton. We contribute a novel understanding
+  of how Parikh maps can be combined with arbitrary commutative monoid morphisms
+  to efficiently represent a wide range of logics on automata and automata-like
+  structures. Furthermore, we show how this formulation can be efficiently
+  implemented as a calculus in a theorem prover supporting Presburger logic. In
+  particular, our calculus allows us to solve the common problem in string
+  solvers of computing the Parikh image of products of arbitrarily many
+  automata. We demonstrate this by implementing a tool called \Catra{}, which we
+  use to solve a set of constraints produced by \Fudge{the PyEx benchmarks} when
+  solved by the \Ostrich{} string constraint solver. We show that this
+  implementation in addition to having a lower memory footprint than the
+  standard eager approach executed on the same underlying solver also
+  outperforms the nuxmv model checker on the same problem\Fudge{, as well as the
+  over-approximation recently described in~\cite{approximate-parikh}.}
 \end{abstract}
@@
 
-\contents{
-\item key challenges: (1) complicated QE problem that was part of
-  previous definitions; (2) languages with large alphabets described
-  by symbolic automata; (3) regular languages described by products
-  of automata
-}
+Parikh maps and their image appears naturally as part of many operations in model checking and the solving of string constraints in automata-based string solvers such as \Ostrich{}, notably in representing constraints on string lengths.
+While it is possible to compute the Parikh image of an automaton using the method described in \cite{generate-parikh-image}, this method produces large existentially quantified clauses which are costly to eliminate and in practice make many real-world problems intractable.
+Furthermore, applications to string analysis in theorem provers require symbolic labels to handle Unicode alphabets \Fudge{citation needed}.
+Conjunctions of string constraints finally, and most crucially, lead to the computation of Parikh images of automata products.
+Applying the approach in \cite{generate-parikh-image} would require the computation of the product before the Parikh image can be computed.
+In several instances we have observed, the materialisation of the product of the automata exhausts the memory of any reasonable machine due to the exponential blow-up of the product.
 
-\subsection{Related Work}
+Addressing these concerns, our specialised calculus for Parikh images allows us to handle symbolic transition labels naturally, while also allowing us to interleave the computation of arbitrarily deep products of automata with the Parikh image of their product.
+This allows us to let both calculations inform each other, eliminating unnecessary work, and pruning the size of the partial products used in the computation.
+Moreover, the method can be used iteratively to, as it were, chew off bite-sized chunks of the product, thereby avoiding the problem of memory-outs.
+
+\subsection{Motivating Example}
 
@@
 
-\subsection{Generalised Parikh Images}
+We define an non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$
+as $\Automaton = \AutomatonTuple$ where
+$\Transitions = \States \times \Alphabet \times \States$, $\States$ is its
+states, $\InitialState$ is w.l.o.g. assumed to be the single initial state, and
+$\Accepting$ is its set of accepting states. Note that we will later generalise
+the labels of transitions further.
 
-\contents{
-\item Preliminaries, standard definition of the Parikh image of a
-  regular/context-free language
-\item Generalised version, considering arbitrary homomorphisms
-\item Examples
-}
+\subsection{The Parikh Map of a Regular language}
+Formally, the \textit{Parikh map} over a context-free language $\Strings=
+\left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:
+
+$$
+\begin{aligned}
+& \ParikhMap: \MapFromTo{\Strings}{\natural^k} \\
+& \ParikhMap(s) = \left[\#a_1(s), \#a_2(s), \ldots, \#a_k(s)\right]
+\end{aligned}
+$$
+
+That is, $\ParikhMap(s)$ is a vector of the number of occurrences of each
+character in the language for a given string $s$. For example, for  $\Strings =
+\Set{a, b}$, we would have $\ParikhMap(abb) = \VectorLiteral{1, 2}$.
+
+We define the image of this map, the \textit{Parikh image}, of some subset of
+the language $\Language \subseteq \Strings$ as:
+
+$$
+\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
+$$
+
+Thus we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1, 1\right],
+\left[1, 2\right]\right\}$.
+
+\subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
+
+Following~\cite{generate-parikh-image}, we define the Parikh Image of a regular
+language recognised by a DFA $\Automaton = \AutomatonTuple$ as:
+
+$$
+\begin{aligned}
+\psi(\mathcal{A}) := 
+& \bigwedge_{\alpha \in \Sigma}
+c_\alpha = \sum_{\delta \in \delta} t_\delta  
+\text{ where $\alpha \in \delta$}\\
+&\bigwedge_{q \in Q} \left (\text{$1$ if $q \in I$} +
+\sum_{\delta = q' \xrightarrow{} q} t_\delta 
+- \sum_{\delta = q\xrightarrow{}q'} t_\delta \right)
+\begin{cases}
+\geq 0 & \text{if $q \in I$} \\
+= 0 & \text{ otherwise}
+\end{cases}\\
+& \bigwedge_{\delta = q \xrightarrow{} q'} t_\delta > 0 
+\implies z_{'q} > 0 \\
+& \bigwedge_{q, q' \in Q} z_{q'} > 0 
+\implies 
+\begin{cases}
+\bigvee\limits_{\delta = q \xrightarrow{} q'} z_{q'} = z_{q} + 1 \land t_\delta > 0 \land z_{q} > 0 & \text{if $q \not\in  I$} \\
+z_{q'} = 1 \land t_\delta > 0& \text{if $q \in I$}
+\end{cases}
+\end{aligned}
+$$
+
+where all variables $z_i, t_i$ are existentially quantified and the variables
+$c_\alpha$ make up the actual image. Feeding this definition into a Presburger
+solver and performing quantifier elimination on $z_i, t_i$ would produce the
+Parikh image in its Presburger form. In this paper we use this example
+(sometimes appropriately modified) as the baseline approach.
+
+\subsubsection{An Example}
+
+\Fudge{AN EXAMPLE!}
+
+\subsection{Generalised Parikh Images}\label{sec:generalised}
+
+To generalise the logic over Parikh images, we intrdouce a morphism $\Map:\: \Sigma
+\to \Monoid$ where $\Monoid$ is a commutative monoid where the following holds:
+\begin{itemize}
+  \item $\Map(\epsilon) = 0$, where $0$ is the neutral element of $\Monoid$
+  \item $\Map(a \Concat b) = \Map(a) + \Map(b)$
+\end{itemize}
+
+\subsubsection{Example~1: String length}
+
+A useful example of such a morphism can express string length. We let $\Monoid =
+\left(\mathbb{Z};+;0\right)$ and let $\Map(x) = 1$. Then the length of a string $s
+= s_1 \RepeatSum{\Concat}  s_n$ is given by $\Map(s) = \sum \Map(s_i) = 1
+\RepeatSum{+} 1$. Similarly, vectors and vector addition can be used to obtain
+the regular Parikh map.
+
+\Fudge{We also expect more efficient computation}
+
+\subsubsection{Example~2: Parikh Automata}
+
+\Fudge{We apply it to Parikh Automata and Magic Happens}
 
@@
 
-\contents{
-\item Lazy expansion of the Parikh conditions for a symbolic automaton
-\item Finding elements vs.\ computing the complete Parikh image
-\item Lazy product computation
-}
+We assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
+transitions $\Transitions =
+\Set{\EllipsisSequence{\Transitions}{\NrTransitions}}$ where we describe each
+such transition $\Transitions_i$ from state $\State$ to state $\State'$ with
+label $\Label \in \Alphabet$ as $\Transitions_i =
+\FromLabelTo{\State}{\Label}{\State'}$. When one or more of the states and
+labels of a transition are uninteresting we will omit it.
 
-\section{A Calculus of Parikh Images}
+For convenience, we introduce the following supporting concepts and notations:
 
-\contents{
-\item Preliminaries, the underlying calculus, what are rules
-\item Predicates used to represent Parikh images
-\item Our calculus rules
-\item Statement of properties, correctness, complexity
-}
+\begin{definition}
+  A \textit{path} $\Path = \PathEnumeration$ of an
+  automaton $\Automaton$ represents a valid run of the automaton as a path
+  starting in its initial state and taking zero or more transitions to arrive at
+  a final accepting state $\State_n \in \Accepting$, while passing a number of labels
+  $\Label_1, \ldots, \Label_n$. If $\InitialState \in \Accepting$, $\Tuple{\InitialState}$
+  is a valid path.
+
+  More formally, we require that:
+  \begin{enumerate}
+    \item $\forall_{k} \Transitions(\State_k, \Label_{k+1}, \State_{k+1})$
+    \item $\State_n \in \Accepting$
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}
+  The \textit{word} of a path $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat} \Label_n$ is
+  the word read out on its labels.
+\end{definition}
+
+\begin{definition}
+ The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
+ number of times a transition $\Transition =
+ \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
+ $p$.
+\end{definition}
+
+We then introduce the two predicates into our calculus with the following
+definitions:
+
+\begin{definition}
+  $\SinglePredicateInstance$ where:
+  \begin{itemize}
+    \item $\MonoidElement \in \Monoid$, for some commutative monoid $\Monoid$,
+    as described in Section~\ref{sec:generalised}.
+    \item $\Filter:\: \Transitions \to \Naturals$
+  \end{itemize}
+  holds when $\exists \Path = \PathEnumeration \in \Automaton$ such that:
+  \begin{itemize}
+    \item $\forall_{\Transition_i \in \Path} \Filter(\Transition_i) =
+    \TransitionCount(\Transition_i, \Path)$
+    \item $\WordOf(\Path) \in \Language(\Automaton)$
+    \item $\MonoidElement = \Map(\WordOf(\Path))$
+  \end{itemize}
+\end{definition}
+
+\begin{definition}
+  $\Connected(\Automaton, \Filter)$ holds when
+  $\exists \Path = \PathEnumeration \in \Automaton$ such that:
+  \begin{itemize}
+    \item $\forall_{\Transition_i \in \Transitions} \Filter(\Transition_i) = \TransitionCount(\Transition_i, \Path)$
+    \item $\WordOf(p) \in \Language(\Automaton)$
+  \end{itemize}
+
+Intuitively, it represents the condition that $\Automaton$ is connected with
+respect to the selection function $\Filter$. It is redundant to $\Image$ by
+design.
+\end{definition}
+
+Finally, we present the rules of our calculus for one automaton. Note that we
+operate on sets of symbols (terms, clauses). Additionally, we also use the
+convention of splitting the terms into linear inequalities ($\SomeInequalities$)
+and any other clauses ($\SomeClause$).
+
+\begin{mathpar}
+  \inferrule*[left=Expand]
+  {\Connected(\Automaton, \Filter) \land \FlowEq(\Automaton, \Filter) \land \MonoidElement = \sum_{\Transition \in \Transitions} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
+  {\SinglePredicateInstance, \SomeInequalities, \SomeClause}
+
+  \inferrule*[left=Split, right=\textnormal{if $\Filter(\Transition) = 0 \not\in \SomeInequalities, \Filter(\Transition) > 0 \not\in \SomeInequalities$}]{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause}
+
+  \inferrule*[left=Propagate, right=\textnormal{if $\Filter(\Transition) = 0 \not\in\SomeInequalities, \forall \Path \in \Automaton \HoldsThat \exists c \in \Path \HoldsThat \Filter(c) = 0 \in \SomeInequalities $}]{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause}
+
+  \inferrule*[left=Subsume, right=\textnormal{if $\forall \Transition = \FromLabelTo{s_0}{}{s_1} \in \Transitions, \Filter(\Transition) = 0 \not \in \SomeInequalities \HoldsThat \exists \Path \in \Automaton, s_0 \in \Path \HoldsThat \forall c \in \TransitionsOf(\Path) \HoldsThat \Filter(c) > 0 \in \SomeInequalities$}]{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause}
+\end{mathpar}
+
+We use the pseudo-function $\FlowEq(\Automaton, \Filter)$ that generates a set
+of existentially quantified linear inequalities with the following definition:
+$$
+\begin{aligned}
+  & \FlowEq(\Automaton, \Filter) = \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)}
+\text{ $\geq 0$ if $\State \in \Accepting$, $= 0$ otherwise}\\
+  & \In(\State, \Filter) = \left(\text{$1$ if $\State = \InitialState$, otherwise $0$}\right) + \sum_{\Transition = \FromLabelTo{\State_0}{}{\State} \in \Transitions} \Filter(\Transition)\\
+  & \Out(\State, \Filter) = \sum_{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions} \Filter(\Transition)
+\end{aligned}
+$$
+Where we transparently assign fresh variables to each state $\State \in \States$
+and existentially quantify them in the whole clause.
+
+Additionally, we assume the existence of a rule \PresburgerClose{},
+corresponding to a sound and complete solver for Presburger formulae, modulo the
+monoid~$\Monoid$.
+
+The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
+$\Automaton$. It states that we are only allowed to use transitions attached to
+a reachable state, and is necessary to ensure connectedness in the presence of
+cycles in~$\Automaton$.
+
+\textsc{Expand} expands the predicate into its most basic rules; one set of
+linear equations synchronising the transitions mentioned by $\Filter$ to the
+corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
+the standard Parikh image formulation, as described by~\FlowEq. Since
+$\Connected$ and $\Image$ are partially redundant and the difference is covered
+by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
+this sense, we split the semantics of the $\Image$~predicate into its counting
+aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
+$\Connected$).
+
+Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
+contested transition from a potential solution before concluding that it must be
+included. Intuitively, this is what guarantees our ability to make forward
+progress by eliminating paths through~$\Automaton$.
+
+A decision procedure for our predicate in a tableau-based automated theorem
+prover would start by expanding the predicate using the $\Expand{}$~rule.
+If~$\Automaton$ contains no loops that could be disconnected from a minimum
+spanning tree (MST) with its root in~$\InitialState$, the initial state, this
+would be sufficient to reach subsumption immediately.
+
+\subsection{An Example}
 
@@
     }
+
+% To generalise the calculus to calculations on products of automata, we change the main predicate take arbitrarily many automata:
+% \begin{definition}
+%   $\ProductPredicateInstance$ is true exactly when:
+%   \begin{itemize}
+%   \item $h:\: \Alphabet^* \rightarrow \SomeMonoid$, a morphism to a commutative product monoid $\SomeMonoid = \prod_{i} \SomeMonoid_i$.
+%   \item $\MonoidElement$ is a vector of elements of $\SomeMonoid$.
+%   \item $\AutomatonVector$ is a vector of automata.
+%   \item $\Automaton = \prod_{i} \AutomatonVector_i$
+%   \item each $\TransitionVec_i$ is a vector of terms such that each term correspond to a
+%     transition in $\AutomatonVector_i$, and $\TransitionVectors$ is a vector of those vectors.
+We use the terminology $\TransitionTerm(t)$ to refer to the term of the transition vector for a given transition $t$.
+%   \item $y = \sum_{t \in \Transitions} h(t)$
+%   \item $\SourceTransitions(\Automaton, t)$ is the set of transitions from the
+%   term automata used to produce the transition $t$ in the product automaton
+%   $\Automaton$.
+%   \end{itemize}
+%   \end{definition}
+
+%   First we extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton:
+%   \begin{mathpar}
+%     \inferrule*[left=Expand]
+%       {\AndComp{\Automaton_i \in \AutomatonVector}{\FlowEq(\Automaton_i) \land \Connected(\Automaton_i)}\ldots}
+%       {\ProductPredicateInstance \land \SomeClause}
+%   \end{mathpar}
+
+%   %  x(t) = sum(e : termProductEdges(t, default=0))
+% Then we introduce the rule $\Materialise$, used to compute a partial product between two terms $\Automaton_i, \Automaton_j$:
+% \begin{mathpar}
+%   \inferrule*[left=Materialise, 
+%   right=\textnormal{where $\Automaton_{k} = \Automaton_i \times \Automaton_j, \AutomatonVector' = (\AutomatonVector \cup \{  \Automaton_k \}) \setminus {\Automaton_i, \Automaton_j}$}]
+%     {\AndComp{t \in \delta(\Automaton_{k})}{\TransitionTerm(t) = \sum_{t' \in \SourceTransitions(t, \Automaton_k)} t'}\\
+%     \FlowEq(\Automaton_{k}) \land \Connected(\Automaton_{k}) \land
+%     \\ \MonoidMap(\AutomatonVector', \TransitionVectors', \MonoidElement) \land \SomeClause} {\MonoidMap(\AutomatonVector, \TransitionVectors, \MonoidElement) \land \SomeClause}
+% \end{mathpar}
+
+% Note that we implicitly existentially quantify the fresh transition terms $\TransitionVec_{k+1}$ for the product.
+
+% $\Materialise$ comes with its companion rule $\RemoveTermConnected$ that simply
+% states we are allowed to remove any instances of $\Connected$ from terms that
+% have been used in a product, and whose connectedness we therefore no longer care about:
+% \begin{mathpar}
+%   \inferrule*[left=$\RemoveTermConnected$, right=\textnormal{If $\Automaton_a$ not in $\Automaton_1\ldots\Automaton_k$}]
+%     {\ProductPredicateInstance \land \SomeClause}
+%     {\Connected(\Automaton_a) \land \ProductPredicateInstance \land \SomeClause}
+% \end{mathpar}
+
+% Finally, we add a side condition to $\Subsume$ to only allow subsumption for
+% instances of $\MonoidMap$ with just one automaton, ensuring that we do not bail
+% out before having computed the entire product.
+
+\subsection{An Example}
+
+\subsection{The calculus is correct}
+
+To be as clear as possible about what it means for our calculus to be correct,
+we use more exact definitions than the traditional soundness and completeness.
+
+\subsubsection{The calculus terminates}
+\subsubsection{The calculus allows no values outside the Parikh image}
+\subsubsection{The calculus allows all values in the Parikh image}
+
+\section{Extending and generalising the calculus}
 

### 2022-04-04 paper/main.tex
Today's edits

@@
-\documentclass{easychair}
+%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
+\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=true}
+%% For double-blind review submission, w/ CCS and ACM Reference
+%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
+%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For single-blind review submission, w/ CCS and ACM Reference
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
+%% For final camera-ready submission, w/ required CCS and ACM Reference
+%\documentclass[acmsmall]{acmart}\settopmatter{}
+
+%% Journal information
+%% Supplied to authors by publisher for camera-ready submission;
+%% use defaults for review submission.
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
+\acmArticle{1}
+\acmYear{2023}
+\acmMonth{1}
+\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+
+%% Copyright information
+%% Supplied to authors (based on authors' rights management selection;
+%% see authors.acm.org) by publisher for camera-ready submission;
+%% use 'none' for review submission.
+\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\copyrightyear{2018}           %% If different from \acmYear
+
+%% Bibliography style
+\bibliographystyle{ACM-Reference-Format}
+%% Citation style
+%% Note: author/year citations are required for papers published as an
+%% issue of PACMPL.
+\citestyle{acmauthoryear}   %% For author/year citations
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Note: Authors migrating a paper from PACMPL format to traditional
+%% SIGPLAN proceedings format must update the '\documentclass' and
+%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Some recommended packages.
+\usepackage{booktabs}   %% For formal tables:
+                        %% http://ctan.org/pkg/booktabs
+\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
+                        %% http://ctan.org/pkg/subcaption
 
 \usepackage{amsthm}
-\usepackage{amssymb,amsmath,mathtools,empheq,fancybox}
-
+\usepackage{amsmath,mathtools,empheq,fancybox}
 \usepackage{paralist}
@@
 \usepackage{color}
-
 \usepackage{textcomp,listings}
-\usepackage[backend=biber]{biblatex}
+\usepackage{array}
 \usepackage{mymacros}
@@
 \usepackage{mathpartir}
-\addbibresource{../bibliography.bib}
-
 \theoremstyle{definition}
@@
 \commentstrue
-
-\newcommand{\pr}[1]{{\color{red}``PR: #1''}}
-\newcommand{\as}[1]{{\color{blue}``AS: #1''}}
-
 \newif\ifoutline
@@
 
+\begin{document}
+
+%% Title information
 \title{An Efficient Calculus for Generalised Parikh Images of Regular Languages}
-\author{some cool authors}
-\institute{Uppsala University, Sweden}
+                                        %% when present, will be used in
+                                        %% header instead of Full Title.
+\titlenote{with title note}             %% \titlenote is optional;
+                                        %% can be repeated if necessary;
+                                        %% contents suppressed with 'anonymous'
 
-\begin{document}
-\maketitle
+%% Author information
+%% Contents and number of authors suppressed with 'anonymous'.
+%% Each author should be introduced by \author, followed by
+%% \authornote (optional), \orcid (optional), \affiliation, and
+%% \email.
+%% An author may have multiple affiliations and/or emails; repeat the
+%% appropriate command.
+%% Many elements are not rendered, but should be provided for metadata
+%% extraction tools.
 
+%% Author with single affiliation.
+\author{First1 Last1}
+\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  \position{Position1}
+  \department{Department1}              %% \department is recommended
+  \institution{Institution1}            %% \institution is required
+  \streetaddress{Street1 Address1}
+  \city{City1}
+  \state{State1}
+  \postcode{Post-Code1}
+  \country{Country1}                    %% \country is recommended
+}
+\email{first1.last1@inst1.edu}          %% \email is recommended
+
+%% Author with two affiliations and emails.
+\author{First2 Last2}
+\authornote{with author2 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  \position{Position2a}
+  \department{Department2a}             %% \department is recommended
+  \institution{Institution2a}           %% \institution is required
+  \streetaddress{Street2a Address2a}
+  \city{City2a}
+  \state{State2a}
+  \postcode{Post-Code2a}
+  \country{Country2a}                   %% \country is recommended
+}
+\email{first2.last2@inst2a.com}         %% \email is recommended
+\affiliation{
+  \position{Position2b}
+  \department{Department2b}             %% \department is recommended
+  \institution{Institution2b}           %% \institution is required
+  \streetaddress{Street3b Address2b}
+  \city{City2b}
+  \state{State2b}
+  \postcode{Post-Code2b}
+  \country{Country2b}                   %% \country is recommended
+}
+\email{first2.last2@inst2b.org}         %% \email is recommended
+
+%% Abstract
+%% Note: \begin{abstract}...\end{abstract} environment must come
+%% before \maketitle command
 \begin{abstract}
@@
   to efficiently represent a wide range of logics on automata and automata-like
-  structures. Furthermore, we show how this formulation can be efficiently
-  implemented as a calculus in a theorem prover supporting Presburger logic. In
-  particular, our calculus allows us to solve the common problem in string
-  solvers of computing the Parikh image of products of arbitrarily many
-  automata. We demonstrate this by implementing a tool called \Catra{}, which we
-  use to solve a set of constraints produced by \Fudge{the PyEx benchmarks} when
-  solved by the \Ostrich{} string constraint solver. We show that this
-  implementation in addition to having a lower memory footprint than the
-  standard eager approach executed on the same underlying solver also
+  structures. This allows us to \Fudge{ensure world peace forever and express
+  loop invariants that solve the halting problem for any practical program}.
+  Furthermore, we show how this formulation can be efficiently implemented as a
+  calculus in a theorem prover supporting Presburger logic. In particular, our
+  calculus allows us to solve the common problem in string solvers of computing
+  the Parikh image of products of arbitrarily many automata. We demonstrate this
+  by implementing a tool called \Catra{}, which we use to solve a set of
+  constraints produced by \Fudge{the PyEx benchmarks} when solved by the
+  \Ostrich{} string constraint solver. We show that this implementation in
+  addition to \Fudge{having a lower memory footprint} than the standard eager
+  approach executed on the same underlying solver also \Fudge{strictly}
   outperforms the nuxmv model checker on the same problem\Fudge{, as well as the
-  over-approximation recently described in~\cite{approximate-parikh}.}
+  over-approximation recently described by~\citeauthor{approximate-parikh} and
+  the original approach described by \citeauthor{generate-parikh-image}.}
+
 \end{abstract}
+
+%% 2012 ACM Computing Classification System (CSS) concepts
+%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011008</concept_id>
+<concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003456.10003457.10003521.10003525</concept_id>
+<concept_desc>Social and professional topics~History of programming languages</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Software and its engineering~General programming languages}
+\ccsdesc[300]{Social and professional topics~History of programming languages}
+%% End of generated code
+
+%% Keywords
+%% comma separated list
+\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
+
+%% \maketitle
+%% Note: \maketitle command must come after title commands, author
+%% commands, abstract environment, Computing Classification System
+%% environment and commands, and keywords command.
+\maketitle
 
@@
 
-Parikh maps and their image appears naturally as part of many operations in model checking and the solving of string constraints in automata-based string solvers such as \Ostrich{}, notably in representing constraints on string lengths.
-While it is possible to compute the Parikh image of an automaton using the method described in \cite{generate-parikh-image}, this method produces large existentially quantified clauses which are costly to eliminate and in practice make many real-world problems intractable.
-Furthermore, applications to string analysis in theorem provers require symbolic labels to handle Unicode alphabets \Fudge{citation needed}.
-Conjunctions of string constraints finally, and most crucially, lead to the computation of Parikh images of automata products.
-Applying the approach in \cite{generate-parikh-image} would require the computation of the product before the Parikh image can be computed.
-In several instances we have observed, the materialisation of the product of the automata exhausts the memory of any reasonable machine due to the exponential blow-up of the product.
+Parikh maps and their image appears naturally as part of many operations in
+model checking and the solving of string constraints in automata-based string
+solvers such as \Ostrich{}, notably in representing constraints on string
+lengths. While it is possible to compute the Parikh image of any context-free
+language using the method described in~\cite{generate-parikh-image}, this method
+produces existentially quantified clauses and take at most linear time to
+produce~\cite{muscholl-linear}. However, the resulting existentially quantified
+clauses are costly to eliminate, in practice making many real-world problems
+intractable. Furthermore, applications to string analysis in theorem provers
+require symbolic labels to handle Unicode alphabets \Fudge{citation needed}.
+Conjunctions of string constraints finally, and most crucially, lead to the
+computation of Parikh images of products of regular languages. Applying the
+approach in \cite{generate-parikh-image} would require the computation of the
+product before the Parikh image can be computed. In several instances we have
+observed, the materialisation of the product of the automata exhausts the memory
+of any reasonable machine due to the exponential blow-up of the product.
 
-Addressing these concerns, our specialised calculus for Parikh images allows us to handle symbolic transition labels naturally, while also allowing us to interleave the computation of arbitrarily deep products of automata with the Parikh image of their product.
-This allows us to let both calculations inform each other, eliminating unnecessary work, and pruning the size of the partial products used in the computation.
-Moreover, the method can be used iteratively to, as it were, chew off bite-sized chunks of the product, thereby avoiding the problem of memory-outs.
+Addressing these concerns, our specialised calculus for Parikh images of
+products of regular languages allows us to handle symbolic transition labels
+naturally, while also allowing us to interleave the computation of arbitrarily
+deep products of automata with the Parikh image of their product. This allows us
+to let both calculations inform each other, eliminating unnecessary work, and
+pruning the size of the partial products used in the computation. Moreover, the
+method can be used iteratively to, as it were, chew off bite-sized chunks of the
+product, thereby avoiding the problem of memory-outs.
 
@@
 
-We define an non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$
-as $\Automaton = \AutomatonTuple$ where
-$\Transitions = \States \times \Alphabet \times \States$, $\States$ is its
-states, $\InitialState$ is w.l.o.g. assumed to be the single initial state, and
-$\Accepting$ is its set of accepting states. Note that we will later generalise
-the labels of transitions further.
+We define a non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$
+as $\Automaton = \AutomatonTuple$ where $\Transitions = \States \times \Alphabet
+\times \States$, $\States$ is its states, $\InitialState$ is w.l.o.g. assumed to
+be the single initial state, and $\AcceptingStates$ is its set of accepting
+states. Note that we will later generalise the labels of transitions further.
 
-\subsection{The Parikh Map of a Regular language}
+\subsection{The Parikh Map}
 Formally, the \textit{Parikh map} over a context-free language $\Strings=
@@
 & \ParikhMap: \MapFromTo{\Strings}{\natural^k} \\
-& \ParikhMap(s) = \left[\#a_1(s), \#a_2(s), \ldots, \#a_k(s)\right]
+& \ParikhMap(s) = \VectorLiteral{\#a_1(s), \#a_2(s), \ldots, \#a_k(s)}
 \end{aligned}
@@
 
-$$
+\[
 \ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
-$$
+\]
 
-Thus we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1, 1\right],
-\left[1, 2\right]\right\}$.
+Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
+1\right], \left[1, 2\right]\right\}$.
+
+Parikh's theorem states that ignoring order, any context-free grammar has a
+letter-equivalent regular language (c.f.~\cite{construction} for a construction of
+such automata from context-free grammars). However, there are languages that are
+not context-free that also have semilinear images under~$\ParikhMap$ (e.g.
+$\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) = \ParikhMap((abc)^*) = \#a =
+\#b = \#c \land \#a \geq 0$).
+
+\contents {
+  \item parikh images of regular images are semilinear, therefore presburger
+  \item presburger is semilinear
+}
 
@@
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
@@
 solver and performing quantifier elimination on $z_i, t_i$ would produce the
-Parikh image in its Presburger form. In this paper we use this example
-(sometimes appropriately modified) as the baseline approach.
+Parikh image in its Presburger form.
+
+In this paper we use this example (sometimes appropriately modified) as the
+baseline approach. Notably, the approach presented in this paper can at its core
+be described as lazily enforcing the connectedness constraint of this method
+(the final two clauses) while also interleaving the (similarly lazy) computation
+of products of automata.
 
@@
 
-To generalise the logic over Parikh images, we intrdouce a morphism $\Map:\: \Sigma
+To generalise the logic over Parikh images, we introdouce a morphism $\Map:\: \Sigma
 \to \Monoid$ where $\Monoid$ is a commutative monoid where the following holds:
@@
 \end{itemize}
+
+\contents{
+  \item The Anca Muscholl paper?
+  \item Describe how a "simpler" function can accelerate solving
+}
 
@@
 
-\Fudge{We also expect more efficient computation}
-
 \subsubsection{Example~2: Parikh Automata}
 
-\Fudge{We apply it to Parikh Automata and Magic Happens}
+Another application for this generalisation is to express Parikh
+automata~\cite{parikh-automata}, and this is indeed what the implementation
+described in Section~\ref{sec:implementation} does.
+
+\contents{
+ \item what is parikh automata
+ \item Give an actual example of a parikh automata
+}
 
@@
 \begin{definition}
-  A \textit{path} $\Path = \PathEnumeration$ of an
-  automaton $\Automaton$ represents a valid run of the automaton as a path
-  starting in its initial state and taking zero or more transitions to arrive at
-  a final accepting state $\State_n \in \Accepting$, while passing a number of labels
-  $\Label_1, \ldots, \Label_n$. If $\InitialState \in \Accepting$, $\Tuple{\InitialState}$
-  is a valid path.
+  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
+  transitions $\Transitions$ represents a path through $\Automaton$ using
+  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
+  \State_{k+1})$ holds for every $k$), passing zero or more labels $\Label_1,
+  \ldots, \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
+  \InitialState$. However, the ending state, $\State_n$, is not necessarily an
+  accepting state.
+  \end{definition}
 
-  More formally, we require that:
-  \begin{enumerate}
-    \item $\forall_{k} \Transitions(\State_k, \Label_{k+1}, \State_{k+1})$
-    \item $\State_n \in \Accepting$
-  \end{enumerate}
+\begin{definition}
+  Moreover, we also talk about the \textit{set of paths} of an automaton,
+  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
+  valid paths through $\Automaton$. Additionally, we use the
+  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
+  state~$\State$.
 \end{definition}
@@
   the word read out on its labels.
+\end{definition}
+
+\begin{definition}
+  The \textit{states} of a path $\StatesOf(\Path) = \Set{\State \SuchThat
+  \FromLabelTo{\State}{}{\State'} \in \Path \text{ or }
+  \FromLabelTo{\State'}{}{\State} \in \Path}$ are the states visited along a
+  path. Note that $\InitialState \in \StatesOf(\Path)$ for every path since all
+  paths start in the initial state.
+\end{definition}
+
+\begin{definition}
+  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$,
+  $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of transitions whose
+  removal from~$\Transitions$ would cause~$\State \in \States$ to be unreachable
+  from the initial state~$\InitialState$. $\SeparatingCut(\Automaton,
+  \InitialState)$ is crucially undefined since it cannot be made unreachable
+  from itself by definition.
 \end{definition}
@@
 \begin{definition}
-  $\SinglePredicateInstance$ where:
+  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
+  $\Automaton = \AutomatonTuple$, modulo some map $\Map$ to a commutative monoid
+  $\Monoid$ as described in Section~\ref{sec:generalised} and with a transition
+  selection function $\Filter:\: \Transitions \to \Naturals$ holds when
+  $\MonoidElement \in \Monoid$ is the Parikh image of $\Automaton$ modulo
+  $\Map$, or more formally when there is an accepting path $\Path =
+  \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such that:
   \begin{itemize}
-    \item $\MonoidElement \in \Monoid$, for some commutative monoid $\Monoid$,
-    as described in Section~\ref{sec:generalised}.
-    \item $\Filter:\: \Transitions \to \Naturals$
-  \end{itemize}
-  holds when $\exists \Path = \PathEnumeration \in \Automaton$ such that:
-  \begin{itemize}
-    \item $\forall_{\Transition_i \in \Path} \Filter(\Transition_i) =
-    \TransitionCount(\Transition_i, \Path)$
-    \item $\WordOf(\Path) \in \Language(\Automaton)$
+    \item $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
     \item $\MonoidElement = \Map(\WordOf(\Path))$
@@
 \begin{definition}
-  $\Connected(\Automaton, \Filter)$ holds when
-  $\exists \Path = \PathEnumeration \in \Automaton$ such that:
-  \begin{itemize}
-    \item $\forall_{\Transition_i \in \Transitions} \Filter(\Transition_i) = \TransitionCount(\Transition_i, \Path)$
-    \item $\WordOf(p) \in \Language(\Automaton)$
-  \end{itemize}
-
-Intuitively, it represents the condition that $\Automaton$ is connected with
-respect to the selection function $\Filter$. It is redundant to $\Image$ by
-design.
+  $\Connected(\Automaton, \Filter)$ for some automaton $\Automaton =
+  \AutomatonTuple$ holds when for every $\Transition =
+  \FromLabelTo{\State}{}{\State'} \in \Transitions$, $\Filter(\Transition) > 0
+  \implies \exists \Path \in \Paths(\Automaton)$ $\Filter(\Transition_i) > 0$
+  for $\Transition_i \in p$, and $\State \in \StatesOf(p)$, or in words that
+  there exists some $\Filter$-selected valid path that reaches $\Transition$'s
+  starting state, $\State$. Intuitively, it represents the condition that
+  $\Automaton$ is connected with respect to the selection function $\Filter$ for
+  every transition. It is redundant to $\Image$ by design.
 \end{definition}
 
-Finally, we present the rules of our calculus for one automaton. Note that we
+Finally, we present the rules of our calculus for one automaton in Table~\ref{tbl:rules:single}. Note that we
 operate on sets of symbols (terms, clauses). Additionally, we also use the
-convention of splitting the terms into linear inequalities ($\SomeInequalities$)
+convention of splitting the formulas into linear inequalities ($\SomeInequalities$)
 and any other clauses ($\SomeClause$).
 
-\begin{mathpar}
-  \inferrule*[left=Expand]
-  {\Connected(\Automaton, \Filter) \land \FlowEq(\Automaton, \Filter) \land \MonoidElement = \sum_{\Transition \in \Transitions} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
-  {\SinglePredicateInstance, \SomeInequalities, \SomeClause}
+We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
+transitions of an automaton~$\Automaton$. Additionally, for an
+automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
+like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
+\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
+$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
+instance, the basis for the matched linear inequalities is
+implicitly~$\SomeInequalities$.
 
-  \inferrule*[left=Split, right=\textnormal{if $\Filter(\Transition) = 0 \not\in \SomeInequalities, \Filter(\Transition) > 0 \not\in \SomeInequalities$}]{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause}
+These rules are meant to be executed in an automated theorem prover, and are
+read bottom-up. Since a large part of our rules operate by adding and literally
+matching linear (in)equalities in a proof goal, we use the shorthand of listing
+the matched inequalities as antecedents. An example of this can be seen in
+the~\Propagate{} rule.
 
-  \inferrule*[left=Propagate, right=\textnormal{if $\Filter(\Transition) = 0 \not\in\SomeInequalities, \forall \Path \in \Automaton \HoldsThat \exists c \in \Path \HoldsThat \Filter(c) = 0 \in \SomeInequalities $}]{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause}
+Please note that the filtering function~$\Filter$ is evaluated symbolically in
+these rules, and can in practice be read as a symbolic function from transitions
+to terms (e.g.~\texttt{t} or~\texttt{t+1}), not dissimilar to~$\FlowEq$ below.
+In our implementation, $\Filter$~is a vector of fresh unknown constants with the
+same size as~$\Transitions$.
 
-  \inferrule*[left=Subsume, right=\textnormal{if $\forall \Transition = \FromLabelTo{s_0}{}{s_1} \in \Transitions, \Filter(\Transition) = 0 \not \in \SomeInequalities \HoldsThat \exists \Path \in \Automaton, s_0 \in \Path \HoldsThat \forall c \in \TransitionsOf(\Path) \HoldsThat \Filter(c) > 0 \in \SomeInequalities$}]{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause}
-\end{mathpar}
+% Use a table!
 
-We use the pseudo-function $\FlowEq(\Automaton, \Filter)$ that generates a set
+We assume that every rule only fires when it would add a new clause. For
+example, this means that we cannot use~\Split{} to split on the same term twice.
+This suggests a proof strategy where you \Propagate{} when you can, \Split{}
+when you must, and \Subsume{} when neither is possible anymore.
+
+\begin{table}[h]
+\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
+  Name & Rule & Prerequisites\\
+  \midrule
+
+  % EXPAND
+  \Expand & 
+    \inferrule
+  {\Connected(\Automaton, \Filter) \land \FlowEq(\Automaton, \Filter) \land \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
+  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
+  None \\
+
+  % SPLIT
+  \Split & 
+  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
+  if $\Transition \in \Transitions_\Automaton$ \\
+
+  % PROPAGATE
+  \Propagate &
+  \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
+  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$\\
+
+  % SUBSUME
+  \Subsume &
+  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
+  \Split{} and \Propagate{} cannot be applied \\
+  \bottomrule
+  \end{tabular}
+  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
+\end{table}
+
+We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
 of existentially quantified linear inequalities with the following definition:
@@
   & \FlowEq(\Automaton, \Filter) = \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)}
-\text{ $\geq 0$ if $\State \in \Accepting$, $= 0$ otherwise}\\
+\text{ $\geq 0$ if $\State \in \AcceptingStates$, $= 0$ otherwise}\\
   & \In(\State, \Filter) = \left(\text{$1$ if $\State = \InitialState$, otherwise $0$}\right) + \sum_{\Transition = \FromLabelTo{\State_0}{}{\State} \in \Transitions} \Filter(\Transition)\\
@@
 \textsc{Expand} expands the predicate into its most basic rules; one set of
-linear equations synchronising the transitions mentioned by $\Filter$ to the
+linear equations synchronising the transitions mentioned by~$\Filter$ to the
 corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
@@
 A decision procedure for our predicate in a tableau-based automated theorem
-prover would start by expanding the predicate using the $\Expand{}$~rule.
-If~$\Automaton$ contains no loops that could be disconnected from a minimum
-spanning tree (MST) with its root in~$\InitialState$, the initial state, this
-would be sufficient to reach subsumption immediately.
+prover would start by expanding the predicate using the $\Expand{}$~rule. A
+modestly clever theorem prover would perform algebraic substitution on the
+underlying constants of~$\Filter$, boiling them down to choices of branches,
+which depend on one single variable, and loop transitions. This logic
+corresponds to the placement of counters for optimally edge-profiling the CFG of
+a program, making up a MST of the automaton~\cite{path-profiling}.
 
 \subsection{An Example}
+
+\subsection{Finding the Presburger representation of a Parikh Image}
+\contents{
+  \item lazy qe
+}
+
+\subsection{The calculus is correct}
+
+To be as clear as possible about what it means for our calculus to be correct,
+we use more exact definitions than the traditional soundness and completeness.
+
+\subsubsection{The calculus terminates}
+\subsubsection{The calculus allows no values outside the Parikh image}
+\subsubsection{The calculus excludes no values in the Parikh image}
+
+\section{Extending and generalising the calculus}
 
@@
     \item additional rules needed for products
-    \item backjumping and conflict-driven learning
     }
+
+\section{Backjumping and Learning No-Goods}
 
@@
 
-\subsection{The calculus is correct}
-
-To be as clear as possible about what it means for our calculus to be correct,
-we use more exact definitions than the traditional soundness and completeness.
-
-\subsubsection{The calculus terminates}
-\subsubsection{The calculus allows no values outside the Parikh image}
-\subsubsection{The calculus allows all values in the Parikh image}
-
-\section{Extending and generalising the calculus}
-
 \section{Implementation and Experiments}
@@
 \item Parikh automata, automata with registers
-\item Model-checking examples
+\item Model-checking examples e.g. \cite{epistemic-logic}
 }
@@
 
-\clearpage
-%\bibliographystyle{splncs03}
-\printbibliography
+%% Acknowledgments
+\begin{acks}                            %% acks environment is optional
+                                        %% contents suppressed with 'anonymous'
+  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
+  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
+  %% acknowledge financial support and will be used by metadata
+  %% extraction tools.
+  This material is based upon work supported by the
+  \grantsponsor{GS100000001}{National Science
+    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+  conclusions or recommendations expressed in this material are those
+  of the author and do not necessarily reflect the views of the
+  National Science Foundation.
+\end{acks}
+
+%% Bibliography
+\bibliography{bibliography}
+
+%% Appendix
+\appendix
+\section{Appendix}
+
+Text of appendix \ldots
 

### 2022-04-05 paper/main.tex
Work work

@@
 \usepackage{mymacros}
+\usepackage{microtype}
 
@@
 
-% Use a table!
-
 We assume that every rule only fires when it would add a new clause. For
@@
 
-\contents{
-    \item additional rules needed for products
-    }
-
-\section{Backjumping and Learning No-Goods}
-
 % To generalise the calculus to calculations on products of automata, we change the main predicate take arbitrarily many automata:
@@
 
+\contents{
+    \item additional rules needed for products
+    }
+
 \subsection{An Example}
 
-\section{Implementation and Experiments}
+\section{Backjumping and Learning No-Goods}
+
+\subsection{An Example}
+
+\section{Implementation and Experiments}\label{sec:implementation}
+
+We implement a calculus for Parikh automata as described in
+Section~\ref{sec:parikh-automata}. The artefact submitted along with this paper
+is a program that reads one or more products of one or more DFA, the
+register incrementations performed along the edges of their transitions, and the
+labels of the transitions as ranges of Unicode characters, along with a set of
+constraints on the final values of their registers expressed as Presburger
+arithmetic. We call this program \Catra{}.\footnote{If you really must read it
+as an acronym, please read it as CAtegory Theory on Register Automata, or
+alternatively if you find this whole paper to be too much of a theoretical
+exercise as Can Anyone Think of A Real Application.}
+
+\Catra{} is implemented in Scala, with the calculus described in this paper
+implemented as a theory plug-in for the \Princess{} automated theorem
+prover~\cite{princess}. For comparison, we also provide an implementation of the
+baseline method from~\cite{generate-parikh-image}, a direct translation that
+uses the~\Nuxmv{} symbolic model checker~\cite{nuxmv} to solve our constraints,
+and \Fudge{the approximation described in~\cite{approximate-parikh}}. An example
+of an input file \Fudge{illustrating every feature of the file format} can be
+seen in Listing~\ref{lst:input-example}.
+
+\contents{
+  \item Listing: a nice input example
+  \item Do we need to say any more about the nuxmv backend?
+  \item Optimisations and tricks
+}
+
+In standard mode, supported by all back-ends, \Catra{} tries to satisfy the
+constraints expressed by the input file, reporting satisfiable or unsatisfiable
+much like traditional~SAT- or SMT solvers.
+Additionally, \Fudge{the baseline and} our own back-end also supports generating the Presburger formula describing the constraints of the input file \Fudge{using the method described in Section~\ref{}}
+
+\subsection{\Catra{} is Versatile: Encoding Epistemic Logic Problems}
+
+\contents{
+\item Model-checking examples e.g. \cite{epistemic-logic}
+}
+
+\subsection{\Catra{} is Faster Than The Competition: Benchmarks and Their Results}
 
@@
 \item Parikh automata, automata with registers
-\item Model-checking examples e.g. \cite{epistemic-logic}
+\item Where do the instances come from!? Why?
+\item Result table
+\item Definition of winning
 }
@@
 \section{Conclusions}
+
+\contents{
+  \item Future extensions: more logics
+  \item Loop invariants?
+  \item Integration into string solvers?
+}
 

### 2022-04-05 paper/progress.sh

@@ -1,0 +1,3 @@
+#!/bin/sh
+
+echo "$(date): $(wc -w main.tex)" >> progress.txt

### 2022-04-05 paper/main.tex
Minor fudge changes

@@
   The image of the Parikh map is important in automata theory, offering a
-  compact characterisation of an automaton. We contribute a novel understanding
-  of how Parikh maps can be combined with arbitrary commutative monoid morphisms
-  to efficiently represent a wide range of logics on automata and automata-like
-  structures. This allows us to \Fudge{ensure world peace forever and express
-  loop invariants that solve the halting problem for any practical program}.
-  Furthermore, we show how this formulation can be efficiently implemented as a
-  calculus in a theorem prover supporting Presburger logic. In particular, our
-  calculus allows us to solve the common problem in string solvers of computing
-  the Parikh image of products of arbitrarily many automata. We demonstrate this
-  by implementing a tool called \Catra{}, which we use to solve a set of
-  constraints produced by \Fudge{the PyEx benchmarks} when solved by the
-  \Ostrich{} string constraint solver. We show that this implementation in
-  addition to \Fudge{having a lower memory footprint} than the standard eager
-  approach executed on the same underlying solver also \Fudge{strictly}
-  outperforms the nuxmv model checker on the same problem\Fudge{, as well as the
-  over-approximation recently described by~\citeauthor{approximate-parikh} and
-  the original approach described by \citeauthor{generate-parikh-image}.}
+  compact characterisation of an automaton. We contribute \Fudge{a novel}
+  understanding of how Parikh maps can be combined with arbitrary commutative
+  monoid morphisms to efficiently represent a wide range of logics on automata
+  and automata-like structures. This allows us to \Fudge{ensure world peace
+  forever and express loop invariants that solve the halting problem for any
+  practical program}. Furthermore, we show how this formulation can be
+  efficiently implemented as a calculus in a theorem prover supporting
+  Presburger logic. In particular, our calculus allows us to solve the common
+  problem in string solvers of computing the Parikh image of products of
+  arbitrarily many automata. We demonstrate this by implementing a tool called
+  \Catra{}, which we use to solve a set of constraints produced by \Fudge{the
+  PyEx benchmarks} when solved by the \Ostrich{} string constraint solver. We
+  show that this implementation in addition to \Fudge{having a lower memory
+  footprint} than the standard eager approach executed on the same underlying
+  solver also \Fudge{strictly} outperforms the \Nuxmv{} model checker on the
+  same problem\Fudge{, as well as both the over-approximation recently described
+  by~\citeauthor{approximate-parikh} and the original approach described by
+  \citeauthor{generate-parikh-image}.}
 
@@
 
-\Catra{} is implemented in Scala, with the calculus described in this paper
+\Catra{} is written in Scala, with the calculus described in this paper
 implemented as a theory plug-in for the \Princess{} automated theorem
-prover~\cite{princess}. For comparison, we also provide an implementation of the
-baseline method from~\cite{generate-parikh-image}, a direct translation that
-uses the~\Nuxmv{} symbolic model checker~\cite{nuxmv} to solve our constraints,
-and \Fudge{the approximation described in~\cite{approximate-parikh}}. An example
-of an input file \Fudge{illustrating every feature of the file format} can be
-seen in Listing~\ref{lst:input-example}.
+prover~\cite{princess}, which also performs the Presburger reasoning. For
+comparison, we also provide an implementation of the baseline method
+from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
+symbolic model checker~\cite{nuxmv} to solve our constraints, and \Fudge{the
+approximation described in~\cite{approximate-parikh} on top of the standard
+Verma back-end}. An example of an input file \Fudge{illustrating every feature
+of the file format} can be seen in Listing~\ref{lst:input-example}.
 
@@
 
-In standard mode, supported by all back-ends, \Catra{} tries to satisfy the
-constraints expressed by the input file, reporting satisfiable or unsatisfiable
-much like traditional~SAT- or SMT solvers.
-Additionally, \Fudge{the baseline and} our own back-end also supports generating the Presburger formula describing the constraints of the input file \Fudge{using the method described in Section~\ref{}}
+In satisfaction mode, supported by all back-ends, \Catra{} tries to satisfy the
+constraints expressed by the input file, reporting satisfiable with register
+assignments or unsatisfiable much like traditional~SAT- or SMT solvers would.
+Additionally, \Fudge{the baseline and} our own back-end also supports generating
+the Presburger formula describing the constraints of the input file \Fudge{using
+the method described in Section~\ref{sec:presburger-representation}}
 
@@
   \item Integration into string solvers?
+  \item We have shown fastest
+  \item We have shown versatile
+  \item We have shown concretely useful
 }

### 2022-05-03 paper/main.tex
More structure

@@
 \usepackage{microtype}
+\usepackage{listings}
+\usepackage{csquotes}
 
@@
 %% comma separated list
-\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
+\keywords{parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
-Parikh maps and their image appears naturally as part of many operations in
-model checking and the solving of string constraints in automata-based string
-solvers such as \Ostrich{}, notably in representing constraints on string
-lengths. While it is possible to compute the Parikh image of any context-free
-language using the method described in~\cite{generate-parikh-image}, this method
-produces existentially quantified clauses and take at most linear time to
+The Parikh map is a characterisation of a context-free language as a vector of
+its character counts, famously a semilinear set~\cite{parikh-theorem}. It
+appears naturally as part of many operations in model checking and the solving
+of string constraints in automata-based string solvers such as \Ostrich{},
+notably in representing constraints on string lengths.
+
+It is possible to compute the Parikh image of any context-free language using
+the method described in~\cite{generate-parikh-image}. Later improvements have
+produced a construction taking at most linear time to
 produce~\cite{muscholl-linear}. However, the resulting existentially quantified
 clauses are costly to eliminate, in practice making many real-world problems
-intractable. Furthermore, applications to string analysis in theorem provers
-require symbolic labels to handle Unicode alphabets \Fudge{citation needed}.
-Conjunctions of string constraints finally, and most crucially, lead to the
-computation of Parikh images of products of regular languages. Applying the
-approach in \cite{generate-parikh-image} would require the computation of the
-product before the Parikh image can be computed. In several instances we have
-observed, the materialisation of the product of the automata exhausts the memory
-of any reasonable machine due to the exponential blow-up of the product.
+intractable.
+
+In addition to costly elimination of the existentially quantified variables,
+another problem arises for the common case of solving length constraints in
+string solvers. Conjunctions of string constraints lead to the computation of
+Parikh images of products of regular languages. Applying the approach
+in~\cite{generate-parikh-image} would require the computation of the product
+before its Parikh image. In several instances we have observed while solving
+real-world string constraints, the computation the product of automata exhausts
+the memory of any machine due to the exponential blow-up in size of the product,
+quickly becoming intractable as the number of terms in the product increases.
 
 Addressing these concerns, our specialised calculus for Parikh images of
-products of regular languages allows us to handle symbolic transition labels
-naturally, while also allowing us to interleave the computation of arbitrarily
-deep products of automata with the Parikh image of their product. This allows us
-to let both calculations inform each other, eliminating unnecessary work, and
-pruning the size of the partial products used in the computation. Moreover, the
-method can be used iteratively to, as it were, chew off bite-sized chunks of the
-product, thereby avoiding the problem of memory-outs.
+products of regular languages allows us to interleave the computation of
+arbitrarily deep products of automata with the product's Parikh image. This
+allows us to let both calculations inform each other, eliminating unnecessary
+work, and pruning the size of the partial products considered in the
+computation. Moreover, the method can be used iteratively to tackle smaller
+chunks of the product incrementally, thereby avoiding running out of memory.
 
@@
 letter-equivalent regular language (c.f.~\cite{construction} for a construction of
-such automata from context-free grammars). However, there are languages that are
+such automata from context-free grammars and~\cite{bounds} for bounds on its size). However, there are languages that are
 not context-free that also have semilinear images under~$\ParikhMap$ (e.g.
@@
   \item Describe how a "simpler" function can accelerate solving
+  \item Describe the relationship of this generalisation to Parikh matrices!
 }
@@
 
-\section{Extending and generalising the calculus}
-
 \section{Parikh Images from Products of Automata}
 
-% To generalise the calculus to calculations on products of automata, we change the main predicate take arbitrarily many automata:
-% \begin{definition}
-%   $\ProductPredicateInstance$ is true exactly when:
-%   \begin{itemize}
-%   \item $h:\: \Alphabet^* \rightarrow \SomeMonoid$, a morphism to a commutative product monoid $\SomeMonoid = \prod_{i} \SomeMonoid_i$.
-%   \item $\MonoidElement$ is a vector of elements of $\SomeMonoid$.
-%   \item $\AutomatonVector$ is a vector of automata.
-%   \item $\Automaton = \prod_{i} \AutomatonVector_i$
-%   \item each $\TransitionVec_i$ is a vector of terms such that each term correspond to a
-%     transition in $\AutomatonVector_i$, and $\TransitionVectors$ is a vector of those vectors.
-We use the terminology $\TransitionTerm(t)$ to refer to the term of the transition vector for a given transition $t$.
-%   \item $y = \sum_{t \in \Transitions} h(t)$
-%   \item $\SourceTransitions(\Automaton, t)$ is the set of transitions from the
-%   term automata used to produce the transition $t$ in the product automaton
-%   $\Automaton$.
-%   \end{itemize}
-%   \end{definition}
-
-%   First we extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton:
-%   \begin{mathpar}
-%     \inferrule*[left=Expand]
-%       {\AndComp{\Automaton_i \in \AutomatonVector}{\FlowEq(\Automaton_i) \land \Connected(\Automaton_i)}\ldots}
-%       {\ProductPredicateInstance \land \SomeClause}
-%   \end{mathpar}
-
-%   %  x(t) = sum(e : termProductEdges(t, default=0))
-% Then we introduce the rule $\Materialise$, used to compute a partial product between two terms $\Automaton_i, \Automaton_j$:
-% \begin{mathpar}
-%   \inferrule*[left=Materialise, 
-%   right=\textnormal{where $\Automaton_{k} = \Automaton_i \times \Automaton_j, \AutomatonVector' = (\AutomatonVector \cup \{  \Automaton_k \}) \setminus {\Automaton_i, \Automaton_j}$}]
-%     {\AndComp{t \in \delta(\Automaton_{k})}{\TransitionTerm(t) = \sum_{t' \in \SourceTransitions(t, \Automaton_k)} t'}\\
-%     \FlowEq(\Automaton_{k}) \land \Connected(\Automaton_{k}) \land
-%     \\ \MonoidMap(\AutomatonVector', \TransitionVectors', \MonoidElement) \land \SomeClause} {\MonoidMap(\AutomatonVector, \TransitionVectors, \MonoidElement) \land \SomeClause}
-% \end{mathpar}
-
-% Note that we implicitly existentially quantify the fresh transition terms $\TransitionVec_{k+1}$ for the product.
-
-% $\Materialise$ comes with its companion rule $\RemoveTermConnected$ that simply
-% states we are allowed to remove any instances of $\Connected$ from terms that
-% have been used in a product, and whose connectedness we therefore no longer care about:
-% \begin{mathpar}
-%   \inferrule*[left=$\RemoveTermConnected$, right=\textnormal{If $\Automaton_a$ not in $\Automaton_1\ldots\Automaton_k$}]
-%     {\ProductPredicateInstance \land \SomeClause}
-%     {\Connected(\Automaton_a) \land \ProductPredicateInstance \land \SomeClause}
-% \end{mathpar}
-
-% Finally, we add a side condition to $\Subsume$ to only allow subsumption for
-% instances of $\MonoidMap$ with just one automaton, ensuring that we do not bail
-% out before having computed the entire product.
+To generalise the calculus to calculations on products of automata, we change the main predicate to take arbitrarily many automata:
+\begin{definition}
+  $\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}$
+  is true exactly when the single-automaton version of the predicate would hold
+  for the automaton.
 
   \contents{
-    \item additional rules needed for products
+    \item Is the enumeration correct?
+    \item Do we need a new name for the predicate?
+    \item Should the automata still be subscripted?
   }
+\end{definition}
+
+\begin{definition}
+When computing a product of two automata, $\Automaton_1 \times \Automaton_2$, we
+need to be able to reason about the sources of a transition. For that, we
+introduce the relation $\ComesFrom(\Transition, \Transition',
+\Automaton_1\times\Automaton_2)$. The predicate holds when $\Transition \in
+\Transitions_{\Automaton_2 \times \Automaton_2}$ is created from $\Transition', \Transition''
+\in \Transitions_{\Automaton_1} \cup \Transitions_{\Automaton_2}$.\Fudge{Do I need to be more specific?}
+\end{definition}
+
+  First we extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton:
+  \begin{mathpar}
+    \inferrule*[left=ExpandM]
+      {\Set{\Automaton_i}{\FlowEq(\Automaton_i), \Connected(\Automaton_i)}, \ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
+      {\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
+  \end{mathpar}
+
+\ExpandM{} must be applied before any other rule, like \Expand{}, but unlike \Expand{}, \ExpandM{} does not remove the $\Image$~predicate since it is needed to keep track of the product.
+
+Then we introduce the rule $\Materialise$, used to compute a partial product between two terms:
+\begin{mathpar}
+  \inferrule*[left=Materialise,right=\textnormal{where $\Automaton' = \Automaton_1\times\Automaton_2$}]
+    {\Set{\Filter(\Transition) = \sum \Filter(\Transition') \HoldsThat \Transition' \in \Transitions_{\Automaton'}, \ComesFrom(\Transition', \Transition, \Automaton') \SuchThat \Transition \in \Transitions_{\Automaton_1} \cup \Transitions_{\Automaton_2}},\\\ImagePredicate{\Automaton'\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
+    {\ImagePredicate{\Automaton_1\times\Automaton_2\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
+\end{mathpar}
+
+Note that we implicitly map $\Filter$ to fresh terms for each transition in the
+product just like before.
+
+Finally, for instances of precisely one automaton, neither rule applies and we
+perform the calculus as before.
 
 \subsection{An Example}
+
+\subsection{The extended calculus is also correct}
+\subsubsection{The calculus terminates}
+\subsubsection{The calculus allows no values outside the Parikh image}
+\subsubsection{The calculus excludes no values in the Parikh image}
 
@@
 
+\begin{lstlisting}[caption={An example input file for \Catra{}, illustrating every major syntax element.
+From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), range labels, and their single-character shorthand syntax, register increments, and constraints on the final values of their counters.}, label=lst:input-example]
+  /* Declare counters. Initially only integers, but other data-types
+ could be considered as well. All counters start at zero. */
+counter int x, y, z, length;
+
+// List automata. Automata transitions are labelled with intervals,
+// and can specify counter increments/decrements. A counter can only
+// be used in at most one automaton.
+
+synchronised {
+    automaton A {
+    init s0;
+
+    s0 -> s1 [0, 10] { x += 1};
+
+    accepting s1, s2;
+    };
+
+    automaton B {
+    init t1;
+
+    t1 -> t2 [11, 22] { y += 1};
+
+    accepting t1, t2;
+    };
+};
+
+automaton C {
+    init s0;
+
+    s0 -> s1 [15] { z += 1 };
+    s1 -> s1 [15] { z += 1};
+
+    accepting s1;
+};
+
+// Specify a constraint on the final values of counters; check whether
+// there is a word accepted by all automata for which the final counter
+// values satisfy this constraints.
+constraint z = 2x + 2y && !(x = 0);
+\end{lstlisting}
+
 \contents{
-  \item Listing: a nice input example
   \item Do we need to say any more about the nuxmv backend?
@@
 
-\subsection{\Catra{} is Versatile: Encoding Epistemic Logic Problems}
+\subsection{\Catra{} is Versatile: Encoding Epistemic Logic Problems As Parikh Automata}
 
@@
 \item Definition of winning
+\item We compare to an implementation using the same automated theorem prover, the same product implementation
 }

### 2022-05-13 main.tex
Clean up repository

--- 2022-05-03 paper/main.tex
@@
 \usepackage{csquotes}
+\usepackage{proof}
 
@@
 \newtheorem{definition}{Definition}[section]
+
+\lstset{
+    columns=fullflexible,
+    showspaces=false,
+    showtabs=false,
+    breaklines=true,
+    showstringspaces=false,
+    breakatwhitespace=true,
+    escapeinside={(*@}{@*)},
+    commentstyle=\color{greencomments},
+    keywordstyle=\color{bluekeywords},
+    stringstyle=\color{redstrings},
+    numberstyle=\color{graynumbers},
+    basicstyle=\ttfamily\small,
+    framesep=12pt,
+    xleftmargin=12pt,
+    tabsize=4,
+    captionpos=b
+}
 
@@
 
-\section{Overview}
+\section{Preliminaries}
 
+\subsection{Finite-state Automata and their Products}
 We define a non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$
@@
 be the single initial state, and $\AcceptingStates$ is its set of accepting
-states. Note that we will later generalise the labels of transitions further.
+states.  We describe a transition $\Transition$ from state $\State$ to state
+$\State'$ with label $\Label \in \Alphabet$ as $\Transition =
+\FromLabelTo{\State}{\Label}{\State'}$. When one or more of the states and
+labels of a transition are uninteresting we will omit it.
 
-\subsection{The Parikh Map}
+We will let variables $\Transition, \Transition', \Transition_1, \ldots, \Transition_n$ etc describe transitions, $\State, \ldots, \State_n$ states, and $\Automaton, \ldots, \Automaton_n$ automata, and use subscript indexing ($\Transitions_\Automaton$) to refer to the transitions, states, etc of a given automaton.
+
+By a \emph{product} of two automata $\Automaton_1, \Automaton_2$, written
+$\Automaton_1 \times \Automaton_2$, we mean an automaton constructed to run
+$\Automaton_1$ and $\Automaton_2$ in parallel on an input and only accept the
+input if both automata would do so.
+
+We refer to the resulting product states as tuples, $\Tuple{\State, \State'}$,
+which represent the state of the product automaton where $\Automaton_1$ would be
+in $\State$ and $\Automaton_2$ would be in $\State'$. Note that since we use
+ordered tuples the product is technically (but w.l.o.g) not commutative; the
+left-hand-side must come from the left-hand term. The sole purpose of this
+matching is to allow us to speak with precision about the origins of components
+in a product.
+
+\subsection{The Parikh Map and its Image}
 Formally, the \textit{Parikh map} over a context-free language $\Strings=
@@
 
-\contents {
-  \item parikh images of regular images are semilinear, therefore presburger
-  \item presburger is semilinear
-}
-
 \subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
 
-Following~\cite{generate-parikh-image}, we define the Parikh Image of a regular
-language recognised by a DFA $\Automaton = \AutomatonTuple$ as:
+Since Parikh images are semilinear, they can be fully expressed by Presburger
+arithmetic. This means that any Parikh image can be written as a set of linear
+equation. Following~\cite{generate-parikh-image}, we define the Parikh Image of
+a regular language recognised by a DFA $\Automaton = \AutomatonTuple$ as:
 
-\[
+\begin{equation}
 \begin{aligned}
@@
 \end{aligned}
-\]
+\label{eq:generate-parikh}
+\end{equation}
 
@@
 
+\section{A Calculus for Generalised Parikh Images}
+
 \subsection{Generalised Parikh Images}\label{sec:generalised}
@@
 To generalise the logic over Parikh images, we introdouce a morphism $\Map:\: \Sigma
-\to \Monoid$ where $\Monoid$ is a commutative monoid where the following holds:
+\to \Monoid$ where $\Monoid = \left(X;+_{\Monoid};0_{\Monoid}\right)$ is a commutative monoid where the following holds:
 \begin{itemize}
-  \item $\Map(\epsilon) = 0$, where $0$ is the neutral element of $\Monoid$
-  \item $\Map(a \Concat b) = \Map(a) + \Map(b)$
+  \item $\Map(\epsilon) = 0_{\Monoid}$
+  \item $\Map(a \Concat b) = \Map(a) +_{\Monoid} \Map(b)$, where $\Concat$ is
+  the string concatenation operation
 \end{itemize}
 
-\contents{
-  \item The Anca Muscholl paper?
-  \item Describe how a "simpler" function can accelerate solving
-  \item Describe the relationship of this generalisation to Parikh matrices!
-}
+Many other generalisations of the Parikh map have been studied, most of them
+less plain than this one. Prominent examples include generalising the Parikh map
+to segments of a fixed length \cite{KARHUMAKI1980155} and the more general
+Parikh matrix, which gives more information about a word than the standard
+Parikh image \cite{parikh-matrix}. Another notable generalisation is the
+p-vector, introduced in~\cite{infinite-words}, which denotes the position of
+each letter in the word rather than the number of their occurrences and allows
+for generalisations into infinite alphabets. All of these, in some sense, add
+something which is not already there in the Parikh map. By contrast, the main
+utility of the formulation introduced here is that it allows us to \emph{remove}
+something, thereby obtaining an easier problem.
 
@@
 
-A useful example of such a morphism can express string length. We let $\Monoid =
-\left(\mathbb{Z};+;0\right)$ and let $\Map(x) = 1$. Then the length of a string $s
-= s_1 \RepeatSum{\Concat}  s_n$ is given by $\Map(s) = \sum \Map(s_i) = 1
-\RepeatSum{+} 1$. Similarly, vectors and vector addition can be used to obtain
-the regular Parikh map.
+A useful example of such a simplifying morphism can express string length, the
+problem that originally motivated our study of the Parikh map. We let $\Monoid =
+\left(\mathbb{Z};+;0\right)$ and let $\Map(x) = 1$. Then the length of a string
+$s = s_1 \RepeatSum{\Concat}  s_n$ is given by $\Map(s) = \sum \Map(s_i) = 1
+\RepeatSum{+} 1$. Note that if we substitute vectors and vector addition for
+this monoid, we obtain the regular Parikh map.
+
+Intuitively, this produces a simpler problem than computing the Parikh image, as
+we now ignore not just the order of characters but also their differences. This
+means that the branching automaton of Figure~\ref{fig:len-branch} becomes
+equivalent to the non-branching automaton of Figure~\ref{fig:len-branch-free}.
+Note that the elimination of the choice of paths also effectively eliminates the
+possibility of disconnecting one of the self-loops. This, in turn, means that we
+can eliminate all of the connectedness implications of the quantified formula,
+which would in turn simplify it (and the quantifier elimination of it)
+considerably.
+
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{choice}
+  \caption{An example automaton with a choice of two paths: either we start with an a or a b.}
+  \label{fig:len-branch}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{no-choice}
+  \caption{An equivalent automaton to the one in Figure~\ref{fig:len-branch}
+  when treating any letter the same as would be the case with a length
+  constraint.}\label{fig:len-branch-free}
+\end{figure}
+
+\contents{
+  \item Formally analyse the complexity of how much easier the length problem is
+}
 
@@
 transitions $\Transitions =
-\Set{\EllipsisSequence{\Transitions}{\NrTransitions}}$ where we describe each
-such transition $\Transitions_i$ from state $\State$ to state $\State'$ with
-label $\Label \in \Alphabet$ as $\Transitions_i =
-\FromLabelTo{\State}{\Label}{\State'}$. When one or more of the states and
-labels of a transition are uninteresting we will omit it.
-
-For convenience, we introduce the following supporting concepts and notations:
+\Set{\EllipsisSequence{\Transitions}{\NrTransitions}}$. For convenience, we then introduce the following supporting notations:
 
@@
 
-  \contents{
-    \item Is the enumeration correct?
-    \item Do we need a new name for the predicate?
-    \item Should the automata still be subscripted?
-  }
-\end{definition}
-
-\begin{definition}
-When computing a product of two automata, $\Automaton_1 \times \Automaton_2$, we
-need to be able to reason about the sources of a transition. For that, we
-introduce the relation $\ComesFrom(\Transition, \Transition',
-\Automaton_1\times\Automaton_2)$. The predicate holds when $\Transition \in
-\Transitions_{\Automaton_2 \times \Automaton_2}$ is created from $\Transition', \Transition''
-\in \Transitions_{\Automaton_1} \cup \Transitions_{\Automaton_2}$.\Fudge{Do I need to be more specific?}
 \end{definition}
@@
     \inferrule*[left=ExpandM]
-      {\Set{\Automaton_i}{\FlowEq(\Automaton_i), \Connected(\Automaton_i)}, \ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
+      {\Set{\FlowEq(\Automaton_i), \Connected(\Automaton_i) \SuchThat \Automaton_i \in \Automaton_1,\ldots,\Automaton_k}, \ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
       {\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
@@
 
+Before we can begin to define our final rule, we need to talk about product states.
+
 Then we introduce the rule $\Materialise$, used to compute a partial product between two terms:
@@
   \inferrule*[left=Materialise,right=\textnormal{where $\Automaton' = \Automaton_1\times\Automaton_2$}]
-    {\Set{\Filter(\Transition) = \sum \Filter(\Transition') \HoldsThat \Transition' \in \Transitions_{\Automaton'}, \ComesFrom(\Transition', \Transition, \Automaton') \SuchThat \Transition \in \Transitions_{\Automaton_1} \cup \Transitions_{\Automaton_2}},\\\ImagePredicate{\Automaton'\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
+    {    \BindingSum(\Automaton', \Filter),\ImagePredicate{\Automaton'\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
     {\ImagePredicate{\Automaton_1\times\Automaton_2\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
@@
 
-Note that we implicitly map $\Filter$ to fresh terms for each transition in the
-product just like before.
+  With the following helper symbolic function:
+
+  $$
+  \BindingSum(\Automaton_1 \times \Automaton_2, \Filter) = \bigcup
+  \begin{aligned}
+  & \Set{
+    \Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_1}
+  }, \\ 
+  & \Set{
+    \Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_2}
+  }
+  \end{aligned}
+$$
+
+Note that this definition implies that $\Filter(\Transition) =
+\Filter(\Transition')$ whenever two transitions $\Transition \in
+\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
+produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
+\times \Automaton_2}$. This corresponds to our intuition that the terms of the
+product must agree on the value they accept. As before, we implicitly map
+$\Filter$ to fresh terms for each transition in the product.
 
@@
 
-\contents{
-  \item Do we need to say any more about the nuxmv backend?
-  \item Optimisations and tricks
-}
+Please note that \Catra{} uses symbolic labels for automata. A symbolic label is
+defined as a range of matching single Unicode code points. This allows succinct
+representation of many regular expression patterns such as \lstinline{(a-z).*}
+which would have otherwise required $27$~nearly identical transitions. In this
+instance, the transition would be defined as \lstinline{a -> b [97, 122]}, as the
+interval is closed from both edges.
 
@@
 assignments or unsatisfiable much like traditional~SAT- or SMT solvers would.
-Additionally, \Fudge{the baseline and} our own back-end also supports generating
+Additionally, the baseline and our own back-end also support generating
 the Presburger formula describing the constraints of the input file \Fudge{using
-the method described in Section~\ref{sec:presburger-representation}}
+the method described in Section~\ref{sec:presburger-representation}}.
+
+Note that \Catra{} can solve equations of Parikh images. An equation like
+$\ParikhMap(\Automaton_1 \times \Automaton_2) = \ParikhMap(\Automaton_3 \times
+\Automaton_4)$ would be expressed as one \lstinline{synchronised} block per side of
+the equation containing their respective terms followed by a
+\lstinline{constraint} requiring all their counters to agree (e.g
+\lstinline|Aa = Ba && Ab = Bb ...|).
+Note that the product construction will implicitly require the counters of each term to agree with each other (and the product), so equalities between all counters of all automata are not necessary.
+
+\subsection{Implementing the Baseline}
+
+We implement the baseline approach using the same Presburger solver
+(\Princess{}), input file parser, and automaton implementation as \Catra{}. We
+do this in order to make the comparison between methods as fair as possible, and
+give us the ability to compare the efficiency of the calculus rules themselves.
+Using the formula of Equation~\ref{eq:generate-parikh}, we produce quantified
+Presburger formulae for each successive term and add them to Princess. In this
+fashion we compute the product incrementally term by term, checking
+satisfiability in each step. We use \Fudge{a priority queue} to select which
+term to use for the next step of the product, and order it by \Fudge{the number
+of transitions} as a heuristic for the size of the term. We use this heuristic
+to avoid computing large (and therefore slow) products until we have to, banking
+our hopes on finding a source of unsatisfiability earlier.
+
+It should be noted that our automata (including the successive products) are
+always by construction forward- and backward- reachable-minimal. We avoid
+computing fully minimal automata since it is unclear when an automaton with
+registers can be safely minimised. This guarantees that any automaton we produce
+only contains states that are both reachable from the initial state and has a
+path to an accepting state.
+
+\subsection{Tuning and optimising our Implementation}
+\contents{
+  \item Optimisations and tricks
+}
 
@@
 \contents{
+  \item Do we need to say any more about the nuxmv backend?
+  \item How did we run the Benchmarks
+
 \item length constraints
@@
 \item We compare to an implementation using the same automated theorem prover, the same product implementation
+\item nuXmv has weird differences in runtime
 }
@@
   \item We have shown concretely useful
+}
+
+\section{Discussions}
+
+\contents{
+\item cf Parikh Images of Grammars: Complexity and Applications -- what is the best complexity we can do? https://ieeexplore.ieee.org/abstract/document/5571050
+
+}
+
+\section{Threats to Validity}
+\contents{
+  \item There might be better implementations
+  \item We might have over-biased
+  \item Our benchmarks might be biased
+  \item The underlying automata implementation might be bad and therefore skew
+  the results.
+
 }

### 2022-06-06 main.tex
Veckans text tm

@@
 \usepackage{proof}
+\usepackage{algorithm2e}
 
@@
 
-Addressing these concerns, our specialised calculus for Parikh images of
-products of regular languages allows us to interleave the computation of
-arbitrarily deep products of automata with the product's Parikh image. This
-allows us to let both calculations inform each other, eliminating unnecessary
-work, and pruning the size of the partial products considered in the
-computation. Moreover, the method can be used iteratively to tackle smaller
-chunks of the product incrementally, thereby avoiding running out of memory.
+Addressing these concerns, our calculus for Parikh images of products of regular
+languages allows us to interleave the computation of arbitrarily deep products
+of automata with the product's Parikh image. This allows us to let both
+calculations inform each other, eliminating unnecessary work, and pruning the
+size of the partial products considered in the computation. Moreover, the method
+can be used iteratively to tackle smaller chunks of the product incrementally,
+thereby avoiding running out of memory.
 
-\subsection{Motivating Example}
+\subsection{A demonstration of our approach}
 
@@
 \begin{aligned}
-\psi(\mathcal{A}) := 
-& \bigwedge_{\alpha \in \Sigma}
-c_\alpha = \sum_{\delta \in \delta} t_\delta  
-\text{ where $\alpha \in \delta$}\\
-&\bigwedge_{q \in Q} \left (\text{$1$ if $q \in I$} +
-\sum_{\delta = q' \xrightarrow{} q} t_\delta 
-- \sum_{\delta = q\xrightarrow{}q'} t_\delta \right)
-\begin{cases}
-\geq 0 & \text{if $q \in I$} \\
-= 0 & \text{ otherwise}
-\end{cases}\\
-& \bigwedge_{\delta = q \xrightarrow{} q'} t_\delta > 0 
-\implies z_{'q} > 0 \\
-& \bigwedge_{q, q' \in Q} z_{q'} > 0 
-\implies 
-\begin{cases}
-\bigvee\limits_{\delta = q \xrightarrow{} q'} z_{q'} = z_{q} + 1 \land t_\delta > 0 \land z_{q} > 0 & \text{if $q \not\in  I$} \\
-z_{q'} = 1 \land t_\delta > 0& \text{if $q \in I$}
+\ParikhMap(\Automaton) := 
+& \AndComp{\Letter \in \Alphabet}{\LetterVar_{\Letter} = \sum_{\Transition = \FromLabelTo{}{\Letter}{} \in \Transitions} \TransitionVar_{\Transition}
+}\\
+& \AndComp{\State \in \AcceptingStates}{\FinalStateVar_{\State} > 0 \implies \StateVar_{\State} > 0}\\
+& \AndComp{\State \in \States}{
+  \left(\FinalStateVar_{\State} \text{ if } \State \in \AcceptingStates \text{ otherwise } 0 \right) +
+  \sum_{\Transition = \FromLabelTo{\State}{}{}} - \sum_{\Transition = \FromLabelTo{}{}{\State}}
+= \begin{cases}
+    1 \text{  if $\State = \InitialState$} \\
+    0 \text{ otherwise}
   \end{cases}
+}\\
+& \AndComp{\Transition = \FromLabelTo{}{}{\State} \in \Transitions}{
+  \TransitionVar_{\Transition} > 0 \implies \StateVar_{\State}
+} \\
+& \AndComp{\State \in \States}{
+  \StateVar_{\State} > 0 \implies
+  \left(\FinalStateVar_{\State} = 1 \text{ if } \State \in \AcceptingStates\right) \OrComp{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions}{
+    \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
+    \TransitionVar_{\Transition} \geq 1 \land
+  \StateVar_{\State'} \geq 1
+    }
+}
 \end{aligned}
@@
 
-where all variables $z_i, t_i$ are existentially quantified and the variables
-$c_\alpha$ make up the actual image. Feeding this definition into a Presburger
-solver and performing quantifier elimination on $z_i, t_i$ would produce the
+where all variables $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ are existentially quantified and the variables
+$\LetterVar_\Letter$ make up the actual image. Feeding this definition into a Presburger
+solver and performing quantifier elimination on $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ with $\LetterVar_\Letter$ as free variables would produce the
 Parikh image in its Presburger form.
@@
 \subsubsection{An Example}
+We apply Equation~\ref{eq:generate-parikh} to the automaton of Figure~\ref{fig:len-branch}.
+We then get the following quantified Presburger formula.
+As before, we omit the implicit existential quantifiers for brevity, along with the constranints requiring all variables to be nonnegative.
+We also omit the $\land$-clauses between each line.
+In other words, in this example, only the variables $\LetterVar_a, \LetterVar_b, \LetterVar_c, \LetterVar_d, \LetterVar_e$, and $\LetterVar_f$ are free, and corresponds to the counts of the respective subscript letter.
+Note that since they here uniquely identify a transition due to the one-to-one mapping between character labels and transitions in the example, the first clause can be trivially eliminated by substituting $\LetterVar_a$ for $\TransitionVar_{\FromLabelTo{0}{a}{1}}$, etc.
+This is not generally the case.
 
-\Fudge{AN EXAMPLE!}
+\begin{equation}
+  \begin{aligned}
+  \ParikhMap(\Automaton):=\:
+  &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} \land \LetterVar_{d} = \TransitionVar_{\FromLabelTo{2}{d}{3}} \land \LetterVar_{e} = \TransitionVar_{\FromLabelTo{2}{e}{2}} \land \LetterVar_{f} = \TransitionVar_{\FromLabelTo{1}{f}{3}}\\
+  &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
+  & \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{0}{a}{1}} - 0 = 1 \land  \TransitionVar_{\FromLabelTo{1}{f}{3}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
+  & \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3} + 0 - \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{1}{f}{3}} = 0\\
+  &\TransitionVar_{\FromLabelTo{0}{a}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} > 0 \implies \StateVar_{2} > 0\\
+  &\TransitionVar_{\FromLabelTo{1}{c}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{2}{e}{2}} > 0 \implies \StateVar_{2} > 0\\
+  &\TransitionVar_{\FromLabelTo{1}{f}{3}} > 0 \implies \StateVar_{3} > 0 \land \TransitionVar_{\FromLabelTo{2}{d}{3}} > 0 \implies \StateVar_{3} > 0\\
+  &\StateVar_{0} > 0 \implies \left(\StateVar_{0} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{0}{a}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{0} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} \geq 1 \land \StateVar_{2} \geq 1\right)\\
+  &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{1}{c}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{1} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{1}{f}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
+  &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{2}{e}{2}} \geq 1 \land \StateVar_{2} \geq 1\right) \lor \left(\StateVar_{2} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{2}{d}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
+  &\StateVar_{3} > 0 \implies \FinalStateVar_{3} = 1
+  \end{aligned}
+  \end{equation}
 
-\section{A Calculus for Generalised Parikh Images}
+\section{Overview of our generalised Parikh images}
 
@@
 something, thereby obtaining an easier problem.
+
+\subsection{Applications}
 
@@
 
-\subsection{Lazy Computation of Parikh Images for Regular Languages}
+\section{A Calculus for Generalised Parikh Images}\label{sec:calculus}
 
@@
 
-\subsection{Finding the Presburger representation of a Parikh Image}
-\contents{
-  \item lazy qe
-}
-
 \subsection{The calculus is correct}
@@
 \subsubsection{The calculus excludes no values in the Parikh image}
+
+\subsection{Finding the Presburger representation of a Parikh Image}
+\contents{
+  \item lazy qe
+}
 
@@
 
-\section{Backjumping and Learning No-Goods}
-
-\subsection{An Example}
+\section{Extensions}
+\subsection{Backjumping and Learning No-Goods}
+\subsubsection{An Example}
+\subsection{Symbolic Automata}\label{sec:ext:symbolic}
 
@@
 to avoid computing large (and therefore slow) products until we have to, banking
-our hopes on finding a source of unsatisfiability earlier.
+our hopes on finding a source of unsatisfiability earlier. The pseudocode for
+our implementation can be seen in Algorithm~\ref{alg:baseline}.
+
+\begin{algorithm}
+  \caption{An algorithm with caption}\label{alg:baseline}
+  \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
+  \KwResult{\textsc{Sat} or \textsc{Unsat}}
+  \SetKwFunction{NewTheoremProver}{newTheoremProver}
+  \SetKwFunction{NewPriorityQueue}{newPriorityQueue}
+  \SetKwFunction{Dequeue}{dequeue}
+  \SetKwFunction{Enqueue}{enqueue}
+  \SetKwFunction{Assert}{assert}
+
+  $p \gets \NewTheoremProver{}$
+
+  \Assert{$p$, $\SomeClause$}
+
+  \ForEach{$\Automaton_i$}{
+    \Assert{$p, \ParikhMap(\Automaton_i)$}
+
+    \If{$p$ is \textsc{Unsat}}{break}
+
+  }
+
+  $q \gets \NewPriorityQueue{}$
+
+  \While{$p$ not \textsc{Unsat} and $|q| > 1$}{
+    $\Automaton, \Automaton' \gets \Dequeue{q}$
+
+    \Assert{$p, \ParikhMap(\Automaton \times \Automaton')$}
+
+    \Enqueue{$q, \Automaton \times \Automaton'$}
+  }
+
+  \KwRet{$p$'s SAT status}
+
+  \end{algorithm}
 
@@
 
-\subsection{Tuning and optimising our Implementation}
+\subsection{Heuristics and search strategies}
 \contents{
@@
 
+\subsection{Threats to Validity}
+\contents{
+  \item There might be better implementations
+  \item We might have over-biased
+  \item Our benchmarks might be biased
+  \item The underlying automata implementation might be bad and therefore skew
+  the results.
+
+}
+
 \section{Conclusions}
@@
 \item cf Parikh Images of Grammars: Complexity and Applications -- what is the best complexity we can do? https://ieeexplore.ieee.org/abstract/document/5571050
-
-}
-
-\section{Threats to Validity}
-\contents{
-  \item There might be better implementations
-  \item We might have over-biased
-  \item Our benchmarks might be biased
-  \item The underlying automata implementation might be bad and therefore skew
-  the results.
 
@@
 \appendix
-\section{Appendix}
+% \section{Appendix}
 
-Text of appendix \ldots
+% Text of appendix \ldots
 

### 2022-06-06 bin/verma-to-tex.py

@@ -1,0 +1,205 @@
+#!/usr/bin/env python3
+
+from re import A
+
+
+initial_state = 0
+accepting_states = {3}
+transitions = [
+    (0, "a", 1),
+    (0, "b", 2),
+    (1, "c", 1),
+    (2, "e", 2),
+    (1, "f", 3),
+    (2, "d", 3),
+]
+
+states = {
+    initial_state,
+    *accepting_states,
+    *{s for s, _, _ in transitions},
+    *{s for _, _, s in transitions},
+}
+
+
+def chunks(l, n):
+    n = max(1, n)
+    return (l[i : i + n] for i in range(0, len(l), n))
+
+
+def subscript(x, s):
+    return x + r"_{" + s + "}"
+
+
+def fmt_transition(t):
+    _from, label, to = t
+    return r"\FromLabelTo{" + str(_from) + "}{" + label + "}{" + str(to) + "}"
+
+
+alphabet = sorted({l for _, l, _ in transitions})
+var = (
+    {ch: subscript(r"\LetterVar", ch) for ch in alphabet}
+    | {
+        t: subscript(
+            r"\TransitionVar",
+            fmt_transition(t),
+        )
+        for t_id, t in enumerate(transitions)
+    }
+    | {s: subscript(r"\StateVar", str(s)) for s in states}
+)
+
+final_var = {s: subscript(r"\FinalStateVar", str(s)) for s in accepting_states}
+
+automaton_symb = r"\Automaton"
+defined_as = r":=\:"
+align_here = "&"
+
+
+def parikh_map_of(something):
+    return r"\ParikhMap(" + something + ")"
+
+
+def tex_env(name, contents):
+    out = r"\begin{" + name + "}\n"
+    out += contents
+    out += "\n"
+    out += r"\end{" + name + r"}"
+    return out
+
+
+def equation(contents):
+    return tex_env("equation", contents)
+
+
+def aligned(things):
+    return tex_env("aligned", ("\\\\\n").join(things))
+
+
+def cases(outcomes):
+    return tex_env("cases", ("\\\\\n").join(outcomes))
+
+
+def and_join(clauses):
+    return r" \land ".join(clauses)
+
+
+def or_join(clauses):
+    return r" \lor ".join(clauses)
+
+
+def sum_of(things):
+    return 0 if not things else " + ".join(things)
+
+
+def sub_of(things):
+    return 0 if not things else " - ".join(things)
+
+
[…]

### 2022-06-11 main.tex
Dagens text

@@
   The image of the Parikh map is important in automata theory, offering a
-  compact characterisation of an automaton. We contribute \Fudge{a novel}
-  understanding of how Parikh maps can be combined with arbitrary commutative
-  monoid morphisms to efficiently represent a wide range of logics on automata
-  and automata-like structures. This allows us to \Fudge{ensure world peace
-  forever and express loop invariants that solve the halting problem for any
-  practical program}. Furthermore, we show how this formulation can be
-  efficiently implemented as a calculus in a theorem prover supporting
-  Presburger logic. In particular, our calculus allows us to solve the common
-  problem in string solvers of computing the Parikh image of products of
-  arbitrarily many automata. We demonstrate this by implementing a tool called
-  \Catra{}, which we use to solve a set of constraints produced by \Fudge{the
-  PyEx benchmarks} when solved by the \Ostrich{} string constraint solver. We
-  show that this implementation in addition to \Fudge{having a lower memory
-  footprint} than the standard eager approach executed on the same underlying
-  solver also \Fudge{strictly} outperforms the \Nuxmv{} model checker on the
-  same problem\Fudge{, as well as both the over-approximation recently described
-  by~\citeauthor{approximate-parikh} and the original approach described by
-  \citeauthor{generate-parikh-image}.}
-
+  compact characterisation of an automaton. We contribute a novel understanding
+  of how Parikh maps can be combined with arbitrary commutative monoid morphisms
+  to efficiently represent real-world constraints found in String solvers.
+  Furthermore, we show how this formulation can be efficiently implemented as a
+  calculus that we call \Calculus{} in a theorem prover supporting Presburger
+  logic. In particular, our calculus allows us to solve the common problem in
+  string solvers of computing the Parikh image of products of arbitrarily many
+  automata. We demonstrate this by implementing a tool called \Catra{}, which we
+  evaluate on constraints produced by \Fudge{the PyEx benchmarks} when solved by
+  the \Ostrich{} string constraint solver. We show that \Calculus{} strictly
+  outperforms the standard eager approach described
+  in~\citeauthor{generate-parikh-image} as well as the improved
+  over-approximating method recently described
+  by~\citeauthor{approximate-parikh}. \Calculus{} typically also outperforms
+  the~\Nuxmv{} model checker, in particularly on unsatisfiable instances.
+  Finally, \Fudge{\Calculus{} and therefore \Catra{} provides an efficient
+  method for deriving the Presburger formula of a given Parikh image that
+  outperforms any previously published method.}
 \end{abstract}
@@
 
-In addition to costly elimination of the existentially quantified variables,
+In addition to the costly elimination of the existentially quantified variables,
 another problem arises for the common case of solving length constraints in
@@
 
-Addressing these concerns, our calculus for Parikh images of products of regular
-languages allows us to interleave the computation of arbitrarily deep products
-of automata with the product's Parikh image. This allows us to let both
-calculations inform each other, eliminating unnecessary work, and pruning the
-size of the partial products considered in the computation. Moreover, the method
-can be used iteratively to tackle smaller chunks of the product incrementally,
-thereby avoiding running out of memory.
+Addressing these concerns, we have developed a calculus for Parikh images of
+products of regular languages that we call \Calculus{}. It allows us to
+interleave the computation of arbitrarily deep products of automata with the
+product's Parikh image. This enables us to let both calculations inform each
+other, eliminating unnecessary work, and pruning the size of the partial
+products considered in the computation for a smaller memory footprint. Moreover,
+the method can be used iteratively to tackle smaller chunks of the product
+incrementally, thereby avoiding running out of memory. We implement this
+calculus, along with the approximate method of~\cite{approximate-parikh}, its
+fall-back variant adapted from~\cite{generate-parikh-image}, and an adapter for
+the \Nuxmv{} model checker, in a tool that we call \Catra{}. \Catra{} allows
+solving constraints over groups of synchronised (products of) Parikh automata,
+that is symbolic automata with transitions that increment and decrement integer
+registers and that allows checking the constraints on these registers after
+executing automata.
+
+Additionally, \Catra{} is able to efficiently generate a quantifier-free
+Presburger formula form of such an instance of constraints and automata,
+something beyond the reach of general model checkers like~\Nuxmv{}.
+
+In general we contribute:
+\begin{itemize}
+\item A Calculus
+\item QE-implementation
+\end{itemize}
 
@@
 
-\subsection{\Catra{} is Faster Than The Competition: Benchmarks and Their Results}
+\subsection{\Catra{} is Faster Than The Competition: Benchmarks and Their Results}\label{sec:experiments}
 
-\contents{
-  \item Do we need to say any more about the nuxmv backend?
-  \item How did we run the Benchmarks
+We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances generated by the \Ostrich{} string constraint solver when solving the \Fudge{pyex benchmarks}.
+\Fudge{The instances are all enormous and have some on average 900 products of 400000 automata with 97 counters each and 9999 constraints}.
+After generating an initial~\InitialNrBenchmarks{}, we remove~\NrBroken{} misgenerated instances that did not parse, as well as~\NrTrivial{} instances solved in under five seconds by the baseline backend.
 
-\item length constraints
-\item Parikh automata, automata with registers
-\item Where do the instances come from!? Why?
-\item Result table
-\item Definition of winning
-\item We compare to an implementation using the same automated theorem prover, the same product implementation
-\item nuXmv has weird differences in runtime
-}
+The benchmarks are executed in parallel, since they are mostly single-threaded and initial results showed negligible interference on performance, on \BenchmarkRig{}.
+We compiled the code using Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a maximum heap of~\MaxHeapSize{}.
+We used \Nuxmv{} version~\NuxmvVersion{} invoked as a subprocess for each instance.
+Experiments were executed on a fresh JVM per backend, but the JVM was kept active for all benchmarks, meaning that JIT compilation would make subsequent executions faster.
+We believe this represents a realistic use case where \Calculus{} is integrated into a string solver or other system and be called repeatedly.
+To mitigate systemic impacts of this decision on our results, experiments were executed in random order for all backends.
+We run all  with a timeout of~\RuntimeTimeout{}.
+
+As a way of gauging the differences in overhead between the native backends and \Nuxmv{}, we executed a small experiment where we ran the same trivial instance 30 times and observed the runtimes for all solvers.
+All backends saw an improvement from JIT compilation within the first three runs, but subsequent runs only improved the native backends.
+The final measured overhead for the fully warm JVMs was less than~\OverheadSeconds{}.
+From this we draw the conclusion that major differences in runtime between the native backends and \Nuxmv{} is more likely to stem from innate differences rather than unfair overhead resulting from our particular configuration.
+
+All runtimes are measured in wall-clock time as observed by the JVM when executing the instance, and exclude time spent parsing, which in all observed cases was minimal in comparison to the runtime, usually far below 0.1 second.
+
+It should be noted that despite running in the standard, deterministic configuration, we observed some degree of nondeterminism in \Nuxmv{}.
+Instances that were solved or unsolved would sometimes be solved or timed out in subsequent runs.
+This observation was made even without randomised execution order, suggesting that the nondeterminism likely stemmed either from our encoding of the instances into \Nuxmv{}'s format, the execution environment.
+Similarly, \Calculus{} showed similar signs of nondeterminism, but only for satisfiable instances.
+To address this, we increasded the runtime from our initial experiments to give us a wider marigin.
+By contrast, the baseline implementation showed no signs of nondeterminism.
+It should be noted that for both backends exhibiting nondeterminism, the nondeterminism did not change the general trend of the experiments, even for experiments with few instances.
+Therefore, we believe that our large set of benchmarks will protect us from bias since any backend is unlikely to be neither consistently lucky nor consistently unlucky.
+
+\subsubsection{Execution Time and Ability to Solve Instances}\label{sec:runtime}
+
+In Figures~\ref{fig:solve-nr} and~\ref{fig:solve-division} we show how many of the~\NrBenchmarks{} the respective back-ends could solve and with which status.
+A full summary of their outcomes is also available in Table~\ref{tab:solve-status}.
+We see here \Fudge{that some solver is better than the others}.
+
+\begin{figure}
+  \caption{The number of solved instances per backend.}
+  \label{fig:solve-nr}
+\end{figure}
+
+\begin{figure}
+  \caption{The division of statuses per backend.}
+  \label{fig:solve-division}
+\end{figure}
+
+\subsubsection{Scalability}\label{sec:scaling}
+
+A cactus plot showing the number of instances solved within a given timeout for each backend can be seen in Figure~\ref{fig:cactus}. \Fudge{We see here that some solver is fast and some other is not}.
+
+\begin{figure}
+  \caption{The number of instances solved within a given number of seconds per backend.}
+  \label{fig:cactus}
+\end{figure}
+
+\subsubsection{Finding a Presburger Formula}
+
+For baseline and \Calculus{}, \Catra{} offers the ability to find the equivalent Presburger formula representing a given instance.
+For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} \Fudge{we use the specially tailored approach described in some section}.
+For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in Sections~\ref{sec:scaling} and~\ref{sec:runtime}.
+
+To make sure baseline puts up as much competition as possible, we disable
+checking intermittent satisfiability and configure \Catra{} to run in the
+maximally eager mode where the product is first computed before any
+satisifiability check is performed. We run the experiments with a timeout of~\ImageTimeout{}. The results of the experiment is summarised
+in Figure~\ref{fig:cactus:image} and Table~\ref{tab:image-results}. \Fudge{We see here that something happens}.
+
+\begin{figure}
+  \caption{The number of instances solved within a given number of seconds per backend.}
+  \label{fig:cactus:image}
+\end{figure}
 

### 2022-06-13 main.tex
Dagens text

@@
   for the automaton.
-
 \end{definition}
 
-  First we extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton:
-  \begin{mathpar}
-    \inferrule*[left=ExpandM]
+  For the calculus, we first extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton.
+
+  \begin{table}[h]
+    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
+      Name & Rule & Prerequisites\\
+      \midrule
+
+      % EXPAND
+      \ExpandM & 
+      \inferrule
       {\Set{\FlowEq(\Automaton_i), \Connected(\Automaton_i) \SuchThat \Automaton_i \in \Automaton_1,\ldots,\Automaton_k}, \ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
-      {\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
-  \end{mathpar}
+      {\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
+      None \\
+      \Materialise &
+      \inferrule
+      {    \BindingSum(\Automaton', \Filter),\ImagePredicate{\Automaton'\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
+      {\ImagePredicate{\Automaton_1\times\Automaton_2\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} &
+      $\Automaton' = \Automaton_1\times\Automaton_2$ \\
+      \bottomrule
+      \end{tabular}
+      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
+    \end{table}
 
@@
 Then we introduce the rule $\Materialise$, used to compute a partial product between two terms:
-\begin{mathpar}
-  \inferrule*[left=Materialise,right=\textnormal{where $\Automaton' = \Automaton_1\times\Automaton_2$}]
-    {    \BindingSum(\Automaton', \Filter),\ImagePredicate{\Automaton'\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
-    {\ImagePredicate{\Automaton_1\times\Automaton_2\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
-\end{mathpar}
 

### 2022-06-14 main.tex
Dagens text

@@
 Presburger formula form of such an instance of constraints and automata,
-something beyond the reach of general model checkers like~\Nuxmv{}.
+something beyond the reach of general model checkers like~\Nuxmv.
 
@@
 
-where all variables $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ are existentially quantified and the variables
-$\LetterVar_\Letter$ make up the actual image. Feeding this definition into a Presburger
-solver and performing quantifier elimination on $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ with $\LetterVar_\Letter$ as free variables would produce the
-Parikh image in its Presburger form.
-
-In this paper we use this example (sometimes appropriately modified) as the
-baseline approach. Notably, the approach presented in this paper can at its core
-be described as lazily enforcing the connectedness constraint of this method
-(the final two clauses) while also interleaving the (similarly lazy) computation
-of products of automata.
+where all variables $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ are
+existentially quantified and the free variables $\LetterVar_\Letter$ make up the
+actual image. Performing standard quantifier elimination on $\TransitionVar_i,
+\StateVar_i, \FinalStateVar_i$ with $\LetterVar_\Letter$ would produce the
+Parikh image in its Presburger form. Note that compared to the method described
+in~\cite{generate-parikh-image}, ours is backwards. Intuitively the approaches
+are equivalent since the Parikh image, by definition, disregards order. The
+approach can be described as either following paths through the automaton
+forwards from an (in our case, \emph{the}) initial state, or conversely as in
+the case here, choose an accepting state and trace a path backwards to the
+initial state.
+
+\contents{
+  \item mention the bug here!
+}
+
+In this paper we refer to this model as the baseline approach, though we also
+apply some optimisations that allow us to detect unsatisfiability early,
+described in Section~\ref{sec:implementing-baseline}. Notably, the calculus
+introduced in this paper can at its core be described as lazily enforcing the
+connectedness constraint of this method (the final two clauses) while also
+interleaving the (similarly lazy) computation of products of automata.
 
@@
 We apply Equation~\ref{eq:generate-parikh} to the automaton of Figure~\ref{fig:len-branch}.
-We then get the following quantified Presburger formula.
+We then get the quantified Presburger formula seen in Equation~\ref{eq:parikh-formula}.
 As before, we omit the implicit existential quantifiers for brevity, along with the constranints requiring all variables to be nonnegative.
@@
   \begin{aligned}
-  \ParikhMap(\Automaton):=\:
   &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} \land \LetterVar_{d} = \TransitionVar_{\FromLabelTo{2}{d}{3}} \land \LetterVar_{e} = \TransitionVar_{\FromLabelTo{2}{e}{2}} \land \LetterVar_{f} = \TransitionVar_{\FromLabelTo{1}{f}{3}}\\
   &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
-  & \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{0}{a}{1}} - 0 = 1 \land  \TransitionVar_{\FromLabelTo{1}{f}{3}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
+  & \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{0}{a}{1}} = 1 \land  \TransitionVar_{\FromLabelTo{1}{f}{3}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
   & \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3} + 0 - \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{1}{f}{3}} = 0\\
@@
   \end{aligned}
+  \label{eq:parikh-formula}
   \end{equation}
+
+  We see here that the formula encodes the choice of path at the start of the
+  automaton ($\TransitionVar_{\FromLabelTo{0}{b}{2}} +
+  \TransitionVar_{\FromLabelTo{0}{a}{1}} = 1$), and the dependency between that
+  choice and the following outgoing transition, e.g. in
+  $\TransitionVar_{\FromLabelTo{2}{d}{3}} -
+  \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0$. Here we only have one choice of
+  incoming flow; from $\FinalStateVar_3$, corresponding to the only accepting
+  state. Simplifying the equation using standard algebra gives us this smaller
+  set of clauses:
+
+  \begin{equation}
+    \begin{aligned}
+    &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
+    & \LetterVar_{d} + \LetterVar_{a} = 1\\
+    & \FinalStateVar_{3}  > \LetterVar_{d} \implies \StateVar_{1} > 0 \land \LetterVar_{d} > 0 \implies \StateVar_{2} > 0\\
+    &\LetterVar_{c} > 0 \implies \StateVar_{1} > 0 \land \LetterVar_{e} > 0 \implies \StateVar_{2} > 0\\
+    &\FinalStateVar_{3}  > \LetterVar_{d} \implies \StateVar_{3} > 0 \land \LetterVar_{d} > 0 \implies \StateVar_{3} > 0\\
+    &\StateVar_{0} > 0 \implies \left(\StateVar_{0} = \StateVar_{1} + 1 \land (\FinalStateVar_{3}  - \LetterVar_{d}) \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{0} = \StateVar_{2} + 1 \land \LetterVar_{d} \geq 1 \land \StateVar_{2} \geq 1\right)\\
+    &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{3} + 1 \land (\FinalStateVar_{3}  - \LetterVar_{d}) \geq 1 \land \StateVar_{3} \geq 1\right)\\
+    &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{3} + 1 \land \LetterVar_{d} \geq 1 \land \StateVar_{3} \geq 1\right)\\
+    &\StateVar_{3} > 0 \implies \FinalStateVar_{3} = 1
+    \end{aligned}
+    \label{eq:parikh-formula:reduced}
+    \end{equation}
+
+Note that some of the or-clauses disappear since they contain false
+inequalities of the type $x = x + 1$, arising from cycles. These clauses encode
+the distance in a spanning tree from an accepting state, and the
+unsatisfiability is by design, essentially ensuring that two states in a cycle
+cannot vouch for each other's connectedness.
+
+Assume we are really interested in occurrences with the letter e, i.e. we solve
+for $\LetterVar_e > 0$. This implies that $\StateVar_{2} > 0$ (essentially, that
+the state our transition of interest started in is reachable), in turn implying
+that $\StateVar_{2} = \StateVar_{3} + 1 \land \LetterVar_{d} \geq 1 \land
+\StateVar_{3} \geq 1$. Armed with this knowledge, we now know that $\StateVar_3
+> 0$, which means that $\FinalStateVar_3 = 1$. Since we also know that
+$\LetterVar_d \geq 1$, we know that $\LetterVar_a = 0$. This corresponds to
+choosing the bottom path through the automaton. Applying this to the equation
+and performing standard boolean logic propagation gives the following:
+
+\begin{equation}
+  \begin{aligned}
+  &\LetterVar_{d} = 1 \land \LetterVar_{a} = 0 \\
+  & \StateVar_{3} > -1 \\
+  &\LetterVar_{c} = 0\\
+  &\StateVar_{0} > 0 \implies \StateVar_{0} = \StateVar_{3} + 2\\
+  \end{aligned}
+  \label{eq:parikh-formula:choose-a}
+  \end{equation}
+
+  Since the final implication holds (we know $\StateVar_3 \geq 0$, and therefore
+  $\StateVar_3 + 2 > 0$), we can disregard it. Thereby we have eliminated all
+  quantified variables and are left with a formula containing only the free
+  ones. Note that all constraints on $\LetterVar_e$ have disappeared after we
+  asserted that it was strictly positive. This is no coincidence: since it is a
+  loop, it can take any positive value.
+
+  This example, in addition to showing how the encoding works, has hopefully
+  also illustrated one of the key insights behind the design of \Calculus: the
+  fact that even few additional constraints on the Parikh image can dramatically
+  reduce the difficulty of computing it, as in this extreme case where only one
+  additional choice in practice fixes the values of all other symbols in the image.
 
@@
 Section~\ref{sec:parikh-automata}. The artefact submitted along with this paper
-is a program that reads one or more products of one or more DFA, the
-register incrementations performed along the edges of their transitions, and the
-labels of the transitions as ranges of Unicode characters, along with a set of
+is a program that reads one or more products of one or more DFA, the register
+incrementations performed along the edges of their transitions, and the labels
+of the transitions as ranges of Unicode characters, along with a set of
 constraints on the final values of their registers expressed as Presburger
 arithmetic. We call this program \Catra{}.\footnote{If you really must read it
-as an acronym, please read it as CAtegory Theory on Register Automata, or
-alternatively if you find this whole paper to be too much of a theoretical
-exercise as Can Anyone Think of A Real Application.}
+as an acronym, please read it as CAtegory Theory on Register Automata, or if you
+object to the somewhat nonstandard use of register automata and category theory,
+as Check Assignments of The Registers Afterwards, or alternatively if you find it
+all to be too much of a theoretical exercise as Can Anyone Think of A Real
+Application.}
 
@@
 approximation described in~\cite{approximate-parikh} on top of the standard
-Verma back-end}. An example of an input file \Fudge{illustrating every feature
+baseline back-end}. An example of an input file \Fudge{illustrating every feature
 of the file format} can be seen in Listing~\ref{lst:input-example}.
@@
 
-\subsection{Implementing the Baseline}
+\subsection{Implementing the Baseline}\label{sec:implementing-baseline}
 
@@
 
-As a way of gauging the differences in overhead between the native backends and \Nuxmv{}, we executed a small experiment where we ran the same trivial instance 30 times and observed the runtimes for all solvers.
+As a way of gauging the differences in overhead between the native backends and \Nuxmv, we executed a small experiment where we ran the same trivial instance 30 times and observed the runtimes for all solvers.
 All backends saw an improvement from JIT compilation within the first three runs, but subsequent runs only improved the native backends.
-The final measured overhead for the fully warm JVMs was less than~\OverheadSeconds{}.
+The final measured overhead for the fully warm JVMs was less than~\OverheadSeconds.
 From this we draw the conclusion that major differences in runtime between the native backends and \Nuxmv{} is more likely to stem from innate differences rather than unfair overhead resulting from our particular configuration.
@@
 
-It should be noted that despite running in the standard, deterministic configuration, we observed some degree of nondeterminism in \Nuxmv{}.
+It should be noted that despite running in the standard, deterministic configuration, we observed some degree of nondeterminism in \Nuxmv.
 Instances that were solved or unsolved would sometimes be solved or timed out in subsequent runs.
-This observation was made even without randomised execution order, suggesting that the nondeterminism likely stemmed either from our encoding of the instances into \Nuxmv{}'s format, the execution environment.
+This observation was made even without randomised execution order, suggesting that the nondeterminism likely stemmed either from our encoding of the instances into \Nuxmv's format, or the execution environment.
 Similarly, \Calculus{} showed similar signs of nondeterminism, but only for satisfiable instances.
-To address this, we increasded the runtime from our initial experiments to give us a wider marigin.
+To address this, we increased the runtime from our initial experiments to give us a wider marigin.
 By contrast, the baseline implementation showed no signs of nondeterminism.

### 2022-06-17 main.tex
Walkthrough

@@
 Additionally, \Catra{} is able to efficiently generate a quantifier-free
-Presburger formula form of such an instance of constraints and automata,
+Presburger formula representing such an instance of constraints and automata,
 something beyond the reach of general model checkers like~\Nuxmv.
 
-In general we contribute:
+In summary, we contribute:
 \begin{itemize}
-\item A Calculus
-\item QE-implementation
+\item The \Calculus{} calculus to efficiently compute the Parikh image of products of Parikh automata.
+\item A specialised method for quantifier elimination that blends well with \Calculus{} to efficiently compute quantifier-free Presburger representations of a given Parikh image.
+\item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future studies.
+\item The \Catra{} tool for solving such instances, containing an implementation of \Calculus{}, the over-approximation described in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model checker~\cite{nuxmv}.
+\item Suggestions for how to efficiently implement \Calculus{} in a modern automated theorem prover, including strategies for case splitting, clause learning, and constraint propagation for connectedness.
+\item Finally, a published description of the well-known bug in~\cite{generate-parikh-image} previously only described in folklore.
 \end{itemize}
@@
 \subsection{A demonstration of our approach}
+
+Consider the two automata $\SomethingCSomething{}$
+(Figure~\ref{fig:something-c-something}) and $\AcaOrBc{}$
+(Figure~\ref{fig:aca-or-bc}). Assume we want to find a value in the Parikh image
+of their product where $\TransitionCount{a} > \TransitionCount{b}$, that is find
+a count of characters in a string accepted by both automata where there are more
+a:s than b:s.
+
+The na\"ive approach would be to first compute the product of the automata and
+then compute the Parikh image of its product. However, we can take advantage of
+the constraint to prune the automata before computing the product and help
+reduce the blow-up. To give an intuition for the core idea of \Calculus{}, we
+will annotate the transitions of $\SomethingCSomething$ and $\AcaOrBc$ with
+variables illustrating how many times they are taken, and perform informal
+reasoning on them corresponding to the reasoning performed by our calculus.
+
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.75\linewidth}
+  \centering 
+    \includegraphics[width=\textwidth]{aca_or_bc}
+    \caption{An automaton recognising the regular expression
+    $\mathtt{ac^*a|b(bb)^*c}$.}\label{fig:aca-or-bc}
+
+    \includegraphics[width=\textwidth]{something_c_something}
+    \caption{An automaton recognising the regular expression
+    $\mathtt{.^*c.^*}$.}\label{fig:something-c-something}
+  \end{minipage}
+  \end{figure}
+
+Note that we have already performed some reasoning in the starting figures -- we
+use the same symbol for the number of times each of the upper transitions of
+$\AcaOrBc$ are taken, since we know that they must be taken the same number of
+times, and we introduced the constant $1$ for the middle transition in
+$\SomethingCSomething$, since it must always be taken precisely once to reach
+the accepting state. We set the initial flow into the starting state to 1 and
+use linear equations to enforce balanced flow throughout the automaton to
+determine the relations between the counting variables on each transition (e.g.
+$1 + {l_b}' = l_a + l_b$, or $l_a + l_c' = 1$). Using standard linear equation
+reasoning loses additional variables and gives us the annotations in
+Figure~\ref{fig:example-simplify-1}.
+
+\begin{figure}[h]
+  \begin{minipage}[b]{0.75\linewidth}
+  \centering 
+    \includegraphics[width=\textwidth]{aca_or_bc_simplified}
+    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
+    simplification of their equations.}\label{fig:example-simplify-1}
+  \end{minipage}
+  \end{figure}
+
+  We now need to relate the number of times a transition was taken to the
+  character counts we are really interested in. Note that this also puts the two
+  automata into contact with each other, before any product has been computed.
+  Some consideration leads us to the definition that the number of occurrence of
+  a each character corresponds to the sum of usages of transitions containing
+  it, e.g. relations of the kind $\TransitionCount{a} = 2 - 2l_c$, and
+  $\TransitionCount{b} = 2 l_b' + l_c'$.\footnote{For the attentive reader: in this
+  example, we leave the catch-all transitions of $\SomethingCSomething$
+  unconstrained since they will be taken care of by the product and would
+  complicate the reasoning since we cannot yet know which character they will
+  correspond to in the final product. In other words, this is a safe
+  under-approximation that only limits our ability to reason before
+  computing the product.}
+
+This relation can be combined with the constraint that $\TransitionCount{a} >
+\TransitionCount{b}$ to obtain $2 - 2l_c > 2 l_b' + l_c'$. Combining this
+equation with the requirement from $\SomethingCSomething$ that
+$\TransitionCount{c} \geq 1$ and $\TransitionCount{c} = l_c + l_c'$ from
+$\AcaOrBc$ gives $0 \geq l_b'$, and since we cannot use a transition a negative
+number of times $l_b' = 0$, which gives the automaton in
+Figure~\ref{fig:example-simplify-2}.
+
+\begin{figure}[h]
+  \begin{minipage}[b]{0.75\linewidth}
+  \centering 
+    \includegraphics[width=\textwidth]{aca_or_bc_simplified_2}
+    \caption{$\AcaOrBc{}$ after removing the now unused b-transition.}\label{fig:example-simplify-2}
+  \end{minipage}
+  \end{figure}
+
+  Plugging in the same inequality again on the same principle, we get $2(1 -
+  l_c') > l_c' \equiv 1 > l_c'$ since we are working with integers, and
+  therefore that $l_c' = 0$. This gives us the even smaller automaton of
+  Figure~\ref{fig:example-simplify-3}.
+
+  \begin{figure}[h]
+    \begin{minipage}[b]{0.75\linewidth}
+    \centering 
+      \includegraphics[width=\textwidth]{aca_or_bc_simplified_3}
+      \caption{$\AcaOrBc{}$ after further constraint propagation.}\label{fig:example-simplify-3}
+    \end{minipage}
+    \end{figure}
+
+This will produce a straightforward product since both automata now have the
+same shape, where we essentially only need to expand the self-loops of
+$\SomethingCSomething$ and reduce its range labels to only capture a:s. Note
+that we got this result by only performing linear inequality reasoning on the
+automata of the product versus the number of times each transition would be
+used.
 

### 2022-06-20 main.tex
Finish motivating example + fix CI

@@
 This relation can be combined with the constraint that $\TransitionCount{a} >
-\TransitionCount{b}$ to obtain $2 - 2l_c > 2 l_b' + l_c'$. Combining this
+\TransitionCount{b}$, that is that there are more instances in the accepted
+string of letters a than b, to obtain $2 - 2l_c > 2 l_b' + l_c'$. Combining this
 equation with the requirement from $\SomethingCSomething$ that
@@
   \begin{figure}[h]
-    \begin{minipage}[b]{0.75\linewidth}
     \centering 
-      \includegraphics[width=\textwidth]{aca_or_bc_simplified_3}
+      \includegraphics[width=0.75\linewidth]{aca_or_bc_simplified_3}
       \caption{$\AcaOrBc{}$ after further constraint propagation.}\label{fig:example-simplify-3}
-    \end{minipage}
     \end{figure}
@@
 same shape, where we essentially only need to expand the self-loops of
-$\SomethingCSomething$ and reduce its range labels to only capture a:s. Note
-that we got this result by only performing linear inequality reasoning on the
-automata of the product versus the number of times each transition would be
-used.
+$\SomethingCSomething$ and reduce its range labels to only capture a:s, as seen
+in Figure~\ref{fig:final-example}. Note that we got this result by only
+performing linear inequality reasoning on the automata of the product versus the
+number of times each transition would be used.
+
+\begin{figure}[h]
+\centering
+    \includegraphics[width=0.75\linewidth]{aca_or_bc_times_something_c_something}
+    \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ under the constraint
+    $\TransitionCount{a} > \TransitionCount{b}$. Note the fresh variables
+    introduced for the transitions (for most transitions the same one, since
+    they are all linearly dependent).}\label{fig:final-example}
+  \end{figure}
+
+  \begin{figure}[h]
+\centering
+    \includegraphics[width=0.75\linewidth]{original_product}
+    \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ as it would have
+    appeared if we had computed it from the initial
+    automata.}\label{fig:original-product}
+  \end{figure}
+
+To obtain the final Parikh image, we need to tie the inequalities for the
+transitions of the automata of the product to the new transition counters we
+introduced in the product automaton. The only interesting transition is the one
+carrying the c, so we will choose that one for illustration. In that case, the
+bridging equation is $r_2 = l_c \land r_2 = r_{\Sigma}$ (and for all others
+$=1$), which gives the Parikh image $\exists_{r_2} \TransitionCount{a} = 1 \land
+\TransitionCount{b} = 0 \land \TransitionCount{c} = r_2 + 1$, and after trivial
+existence-elimination using reasoning on lower bounds (e.g. $r_2 \geq 0$) we
+arrive at $\TransitionCount{c} \geq 1$. For comparison, the $\AcaOrBc{} \times
+\SomethingCSomething{}$ we would have obtained if we had eagerly computed the
+product can be seen in Figure~\ref{fig:original-product}. This would have
+introduced a significantly more complicated existentially quantified formula.
 
@@
 
-We will let variables $\Transition, \Transition', \Transition_1, \ldots, \Transition_n$ etc describe transitions, $\State, \ldots, \State_n$ states, and $\Automaton, \ldots, \Automaton_n$ automata, and use subscript indexing ($\Transitions_\Automaton$) to refer to the transitions, states, etc of a given automaton.
+We will let variables $\Transition, \Transition', \Transition_1, \ldots,
+\Transition_n$ etc describe transitions, $\State, \ldots, \State_n$ states, and
+$\Automaton, \ldots, \Automaton_n$ automata, and use subscript indexing
+($\Transitions_\Automaton$) to refer to the transitions, states, etc of a given
+automaton.
 

### 2022-06-20 main.tex
Preliminary graphs

@@
 \usepackage{algorithm2e}      
-
+\usepackage{multirow}
 \usepackage{mathpartir}
@@
 
+\contents{
+  \item introducera homomorfismen
+  \item start with length
+  \item räkna bara a och b, strunta i c
+  \item ändra till >= för att illustrera splitting
+  }
+
 Consider the two automata $\SomethingCSomething{}$
@@
 After generating an initial~\InitialNrBenchmarks{}, we remove~\NrBroken{} misgenerated instances that did not parse, as well as~\NrTrivial{} instances solved in under five seconds by the baseline backend.
+The benchmarks are run on commit \texttt{\commit}.
 
@@
 
-In Figures~\ref{fig:solve-nr} and~\ref{fig:solve-division} we show how many of the~\NrBenchmarks{} the respective back-ends could solve and with which status.
+In Figures~\ref{fig:solve-division} we show how many of the~\NrBenchmarks{} the respective back-ends could solve and with which status.
 A full summary of their outcomes is also available in Table~\ref{tab:solve-status}.
@@
 
+\begin{table}[ht]
+  \centering
+  \input{graphs/solved_pivot_table.tex}
+  \caption{The result of running the respective back-ends by instance
+  satisifiability (satisifiable or unsatisifiable) with a timeout of
+  \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
+  from the table. }\label{tab:solve-status}
+\end{table}
+
 \begin{figure}
-  \caption{The number of solved instances per backend.}
-  \label{fig:solve-nr}
-\end{figure}
-
-\begin{figure}
+  \includegraphics[width=0.75\linewidth]{graphs/\commit-by-solver.pdf}
   \caption{The division of statuses per backend.}
@@
 
+\begin{figure}
+  \includegraphics[width=0.75\linewidth]{graphs/\commit-time-boxplot.pdf}
+  \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
+  \label{fig:runtime-boxplot}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=0.75\linewidth]{graphs/\commit-duels-scatter.pdf}
+  \caption{The pairwise runtime of each instance for instances solved by any backend. Timeouts are reported as a runtime of \RuntimeTimeout.}
+  \label{fig:duels}
+\end{figure}
+
 \subsubsection{Scalability}\label{sec:scaling}
@@
 \begin{figure}
-  \caption{The number of instances solved within a given number of seconds per backend.}
+  \includegraphics[width=0.75\linewidth]{graphs/\commit-cactus.pdf}
+  \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
   \label{fig:cactus}
@@
 \begin{figure}
-  \caption{The number of instances solved within a given number of seconds per backend.}
+  \caption{The number of instances \Catra{} was able to find the Presburger form of the image for within a given number of seconds per backend.}
   \label{fig:cactus:image}

### 2022-06-21 main.tex
Add duels
Mainly work on introduction

@@
 
-\contents{
-  \item introducera homomorfismen
-  \item start with length
-  \item räkna bara a och b, strunta i c
-  \item ändra till >= för att illustrera splitting
-  }
-
-Consider the two automata $\SomethingCSomething{}$
-(Figure~\ref{fig:something-c-something}) and $\AcaOrBc{}$
-(Figure~\ref{fig:aca-or-bc}). Assume we want to find a value in the Parikh image
-of their product where $\TransitionCount{a} > \TransitionCount{b}$, that is find
-a count of characters in a string accepted by both automata where there are more
-a:s than b:s.
-
-The na\"ive approach would be to first compute the product of the automata and
-then compute the Parikh image of its product. However, we can take advantage of
-the constraint to prune the automata before computing the product and help
-reduce the blow-up. To give an intuition for the core idea of \Calculus{}, we
-will annotate the transitions of $\SomethingCSomething$ and $\AcaOrBc$ with
-variables illustrating how many times they are taken, and perform informal
-reasoning on them corresponding to the reasoning performed by our calculus.
-
-\begin{figure}[ht]
+To illustrate the key features of our approach, we start with two regular
+languages, a commutative homomorphism $f$ on them that we care about, and a side
+constraint for which we want to find values in $f$'s image over the intersection
+of the wto languages. We will use the same two regular languages, recognised by
+two automata; $\SomethingCSomething{}$ (Figure~\ref{fig:something-c-something})
+and $\AcaOrBc{}$ (Figure~\ref{fig:aca-or-bc}) for both examples but change the
+constraint and homomorphism $f$.
+
+\subsubsection{Counting string lengths}
+
+To keep things simple, in this first example we
+let $f$ be the function that gives the length of a given string, e.g.
+$f\left(\text{curious and curiouser still}\right) = 27$, and $f\left(\text{homo}\right) +
+f\left(\text{morphism}\right) = f\left(\text{homomorphism}\right) = 12$.
+
+\begin{figure}[h]
   \begin{minipage}[b]{0.75\linewidth}
@@
 
+The na\"ive approach would be to first compute the product of the automata and
+then compute the length of its product. However, we can take advantage of the
+both constraints and definitions of $f$ to prune the automata before computing
+the product and help reduce the blow-up. To give an intuition for the core idea
+of \Calculus{}, we will annotate the transitions of $\SomethingCSomething$ and
+$\AcaOrBc$ with variables illustrating how many times they are taken, and
+perform informal reasoning on them corresponding to the reasoning performed by
+our calculus.
+
+\begin{figure}[h]
+  \centering 
+    \includegraphics[width=0.75\linewidth]{aca_or_bc_simplified}
+    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
+    simplification of their equations.}\label{fig:example-simplify-1}
+  \end{figure}
+
 Note that we have already performed some reasoning in the starting figures -- we
@@
 $\SomethingCSomething$, since it must always be taken precisely once to reach
-the accepting state. We set the initial flow into the starting state to 1 and
-use linear equations to enforce balanced flow throughout the automaton to
-determine the relations between the counting variables on each transition (e.g.
-$1 + {l_b}' = l_a + l_b$, or $l_a + l_c' = 1$). Using standard linear equation
-reasoning loses additional variables and gives us the annotations in
-Figure~\ref{fig:example-simplify-1}.
+the accepting state. We use standard graph-theoretical flow reasoning to set the
+initial flow into the starting state to 1 and use linear equations to enforce
+balanced flow throughout the automaton to determine the relations between the
+counting variables on each transition (e.g. $1 + {l_b}' = l_a + l_b$, or $l_a +
+l_c' = 1$). Strictly speaking, this is where the constant number $1$ in
+Figure~\ref{fig:something-c-something} comes: from $1 + r_\Sigma = x + r_\Sigma$
+solved for $x$. Using analogous linear equation reasoning loses additional
+variables and gives us the annotations in Figure~\ref{fig:example-simplify-1}.
+
+We sum the individual transition variables up to get the length we are actually
+interested in, e.g. the image of $f$ on $\SomethingCSomething =
+\exists_{r_\Sigma, r_\Sigma'} \: r_\Sigma + 1 + r_\Sigma'$. For
+$\SomethingCSomething{}$, this is all we need to produce the image since all
+transitions are always accessible from any path from the initial state to the
+accepting state.
+
+The story gets more complicated with $\AcaOrBc{}$, where we have a choice of two
+paths, and where that choice affects the possibility of taking either of the two
+loops. In that case (letting the existential quantifiers from now on be
+implicit), we would have the image $2l_a + l_c + l_c' + l_b + l_b'$. If we apply
+standard linear reasoning on the equations (e.g. $1 + {l_b}' = l_a + l_b$, or
+$l_a + l_c' = 1$), we can reduce the number of bound ($\exists$-quantified)
+variables somewhat, obtaining the relations in
+Figure~\ref{fig:example-simplify-1}, and the corresponding over-approximation on
+the $f$-image $2 + l_c + 2l_b'$. It is an over-approximation because it only
+considers the number of times each transition is taken and the existence of a
+path from an initial to an accepting state, but omits the connectivity of that
+path in the presence of loops. For example, no valid path would have both $l_c >
+0$ and $l_b' > 0$, but no such restriction exists in the formula above.
 
 \begin{figure}[h]
-  \begin{minipage}[b]{0.75\linewidth}
   \centering 
-    \includegraphics[width=\textwidth]{aca_or_bc_simplified}
-    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
-    simplification of their equations.}\label{fig:example-simplify-1}
-  \end{minipage}
-  \end{figure}
+    \includegraphics[width=0.75\linewidth]{something_c_something_length_simplified}
+    \caption{$\SomethingCSomething{}$ after using linear reasoning over lengths to replace a variable.}\label{fig:example-length-reduced}
+  \end{figure}
+
+Still, this approximate solution allows us to perform further reasoning. Since
+we know that the length of a string accepted by the intersection of the two
+automata's languages must be the same, we also know that their images (string
+lengths) must be the same under $f$, e.g. that $r_\Sigma + 1 + r_\Sigma' = 2 +
+l_c + 2l_b' \iff r_\Sigma + r_\Sigma' = 1 + l_c + 2l_b'$. This in turn allows us
+to replace one additional variable, e.g. $r_\Sigma = 1 + l_c + 2l_b' -
+r_\Sigma'$ (see Figure~\ref{fig:example-length-reduced}). Note that the two
+automata are now in direct contact. Any reasoning performed on one of them could
+be propagated to the other.
+
+Since we cannot take a transition a negative number of times, we have an
+implicit constraint on any transition variable that it must be at least $0$.
+This means that we have $1 + l_c + 2l_b' - r_\Sigma'$, implying an upper bound
+on $r_\Sigma'$: $r_\Sigma' \leq 1 + l_c + 2l_b'$.
+
+We talked initially about having a constraint on the image (or conversely, on
+the strings of the language), so let us add one. We are only interested in
+strings of odd length, e.g. $2 \nmid  f(\AcaOrBc \times \SomethingCSomething)$.
+This means that: $2 \nmid 2 + l_c + 2l_b' \iff 2 \nmid l_c$. To encode this condition, we simply replace $l_c$ with $2k + 1$ with an existentially quantified integer $k \geq 0$.
+
+This is as far as we can get with this reasoning. We must now choose: either we
+compute the product and continue our reasoning, or we perform a case split. Let
+us try a case split in order to put off computing a potentially large product.
+We pick an early transition in the largest automaton with loops, in our case
+$\FromLabelTo{start}{a}{sawA}$ of $\AcaOrBc$. This creates two cases: $1-l_c' =
+0$ and $1-l_c' > 0$.
+We start with the first case, and after propagating the now known values and removing any transition that becomes zero, we get the automaton of Figure~\ref{fig:aca-or-bc-length-split-1}.
+
+\begin{figure}[h]
+  \centering 
+    \includegraphics[width=0.75\linewidth]{aca_or_bc_length_split_1}
+    \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{start}{a}{sawA}$ is not used.}\label{fig:aca-or-bc-length-split-1}
+  \end{figure}
+
+  We immediately notice that the self-loop of state $sawA$ is now disconnected
+  and must be removed. This condition can be detected efficiently using standard
+  forwards/backwards reachability from the initial and accepting states
+  respectively.
+In other words, $2k+1 = 0$ for this case and is implied by the splitting constraint.
+However, no choice of integer $k$  will satisfy this equation.
+Therefore, we close this branch and proceed with the other one -- where $1-l_c' > 0$.
+Note that we do so without ever computing a potentially expensive product of automata.
+
+  \begin{figure}[h]
+    \centering 
+      \includegraphics[width=0.75\linewidth]{aca_or_bc_length_split_2}
+      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{start}{a}{sawA}$ is used.}\label{fig:aca-or-bc-length-split-2}
+    \end{figure}
+
+Plugging in the assumption $1-l_c' > 0$ (and its consequence that $l_c' = 0$),
+we also get a reduced variant of $\AcaOrBc$, as seen in
+Figure~\ref{fig:aca-or-bc-length-split-1}. Note that $sawB$ is no longer
+backwards-reachable from $final$ and therefore dead, along with both transitions
+into it. We add the conclusion $l_b' = 0$, which severs both transitions into
+the state and removes it from our automaton. Following this we are ready to
+compute the product, which can be seen in Figure~\ref{fig:length-split-product}.
+
+\begin{figure}[h]
+  \centering 
+    \includegraphics[width=0.75\linewidth]{length_split_product}
+    \caption{The product of $\AcaOrBc$ and $\SomethingCSomething$ after
+    case splitting and removal of dead states.}\label{fig:length-split-product}
+  \end{figure}
+
+Before continuing, we need to tie the inequalities for the transitions of the
+automata of the product to the new transition counting variables we introduced
+in the product automaton to ensure that they are consistent.
+In our case, we obtain the following equations from the principle that a transition variable in either automaton must be equal to the sum of variables in the resulting edges of the product, which gives the following two interesting inequality from $\AcaOrBc{}$'s transition $\FromLabelTo{sawA}{c}{sawA}$: $2k + 1 = r_1 + r_2 + 1$
+
+Note how this conserves the requirement that the number of loop iterations be
+odd even when the same original loop appears three times in the product, twice
+as a self-loop and once as a regular state transition. Using the same principle
+as before for computing the length by adding up the transition counters, we
+obtain the quantified formula $1 + r_1 + 1 = r_2 + 1 = r_1 + r_2 + 3 = 2k + r_1
++ 3$ for the length. Eliminating the final quantifier gives us the image of
+$f(\AcaOrBc \times \SomethingCSomething)$ modulo odd lengths: $3 + 2k$, for some
+integer $k$.
+
+\subsubsection{Producing the Parikh image}
+
+Now assume that rather than the length (e.g. the count of the characters) we
+want to find the whole Parikh image of the intersection of $\AcaOrBc{}$ and
+$\SomethingCSomething{}$. In other words, now $f$ is an object (usually a
+vector) counting how often each character occurs in a string. Since extracting
+values from a vector is tiresome, we will use the shorthand notation
+$\TransitionCount{a}$ to refer to the number of letters a that appear.
+E.g.
+$\TransitionCount{a}(\text{curious and curiouser still}) = 1$.
+Note that we still have the property that $\TransitionCount{o}(\text{homomorphism}) = \TransitionCount{o}(\text{homo}) + \TransitionCount{o}(\text{morphism}) = 2 + 1 = 3$.
+If you treat the addition as element-wise vector addition, this approach works for the vector version as well.
+
+For our constraint, we will use $\TransitionCount{a} > \TransitionCount{b}$,
+that is find a count of characters in a string accepted by both automata where
+there are more a:s than b:s. Since the underlying calculus is the same, we start immediately from Figure~\ref{fig:example-simplify-1}.
 
   We now need to relate the number of times a transition was taken to the
-  character counts we are really interested in. Note that this also puts the two
-  automata into contact with each other, before any product has been computed.
-  Some consideration leads us to the definition that the number of occurrence of
-  a each character corresponds to the sum of usages of transitions containing
-  it, e.g. relations of the kind $\TransitionCount{a} = 2 - 2l_c$, and
-  $\TransitionCount{b} = 2 l_b' + l_c'$.\footnote{For the attentive reader: in this
-  example, we leave the catch-all transitions of $\SomethingCSomething$
+  character counts we are really interested in, as opposed to just the length in
+  the previous example. Note that this, like the length calculus before, puts
+  the two automata into contact with each other, before any product has been
+  computed. Some consideration leads us to the definition that the number of
+  occurrence of each character corresponds to the sum of usages of transitions
+  containing it, e.g. relations of the kind $\TransitionCount{a} = 2 - 2l_c$,
+  and $\TransitionCount{b} = 2 l_b' + l_c'$.\footnote{For the attentive reader:
+  in this example, we leave the catch-all transitions of $\SomethingCSomething$
   unconstrained since they will be taken care of by the product and would
@@
   correspond to in the final product. In other words, this is a safe
-  under-approximation that only limits our ability to reason before
-  computing the product.}
+  under-approximation that only limits our ability to reason before computing
+  the product.}
 
@@
 \begin{figure}[h]
-  \begin{minipage}[b]{0.75\linewidth}
   \centering 
-    \includegraphics[width=\textwidth]{aca_or_bc_simplified_2}
+    \includegraphics[width=0.75\linewidth]{aca_or_bc_simplified_2}
     \caption{$\AcaOrBc{}$ after removing the now unused b-transition.}\label{fig:example-simplify-2}
-  \end{minipage}
   \end{figure}
@@
   Plugging in the same inequality again on the same principle, we get $2(1 -
-  l_c') > l_c' \equiv 1 > l_c'$ since we are working with integers, and
+  l_c') > l_c' \iff 1 > l_c'$ since we are working with integers, and
   therefore that $l_c' = 0$. This gives us the even smaller automaton of
@@
 \subsubsection{An Example}
-We apply Equation~\ref{eq:generate-parikh} to the automaton of Figure~\ref{fig:len-branch}.
+We apply Equation~\ref{eq:generate-parikh} to the automaton $\AcaOrBc{}$ of Figure~\ref{fig:aca-or-bc}.
 We then get the quantified Presburger formula seen in Equation~\ref{eq:parikh-formula}.
@@
 In other words, in this example, only the variables $\LetterVar_a, \LetterVar_b, \LetterVar_c, \LetterVar_d, \LetterVar_e$, and $\LetterVar_f$ are free, and corresponds to the counts of the respective subscript letter.
-Note that since they here uniquely identify a transition due to the one-to-one mapping between character labels and transitions in the example, the first clause can be trivially eliminated by substituting $\LetterVar_a$ for $\TransitionVar_{\FromLabelTo{0}{a}{1}}$, etc.
-This is not generally the case.
 
@@
   \begin{aligned}
-  &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} \land \LetterVar_{d} = \TransitionVar_{\FromLabelTo{2}{d}{3}} \land \LetterVar_{e} = \TransitionVar_{\FromLabelTo{2}{e}{2}} \land \LetterVar_{f} = \TransitionVar_{\FromLabelTo{1}{f}{3}}\\
+  \ParikhMap(\Automaton):=\:
+  &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} + \TransitionVar_{\FromLabelTo{1}{a}{3}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{2}{b}{1}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} + \TransitionVar_{\FromLabelTo{2}{c}{3}}\\
   &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
-  & \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{0}{a}{1}} = 1 \land  \TransitionVar_{\FromLabelTo{1}{f}{3}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
-  & \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3} + 0 - \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{1}{f}{3}} = 0\\
+  & \TransitionVar_{\FromLabelTo{0}{a}{1}} + \TransitionVar_{\FromLabelTo{0}{b}{2}} - 0 = 1 \land  \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{2}{b}{1}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
+  & \TransitionVar_{\FromLabelTo{2}{b}{1}} + \TransitionVar_{\FromLabelTo{2}{c}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3} + 0 - \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{2}{c}{3}} = 0\\
   &\TransitionVar_{\FromLabelTo{0}{a}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} > 0 \implies \StateVar_{2} > 0\\
-  &\TransitionVar_{\FromLabelTo{1}{c}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{2}{e}{2}} > 0 \implies \StateVar_{2} > 0\\
-  &\TransitionVar_{\FromLabelTo{1}{f}{3}} > 0 \implies \StateVar_{3} > 0 \land \TransitionVar_{\FromLabelTo{2}{d}{3}} > 0 \implies \StateVar_{3} > 0\\
+  &\TransitionVar_{\FromLabelTo{1}{c}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{1}{a}{3}} > 0 \implies \StateVar_{3} > 0\\
+  &\TransitionVar_{\FromLabelTo{2}{b}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} > 0 \implies \StateVar_{3} > 0\\
   &\StateVar_{0} > 0 \implies \left(\StateVar_{0} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{0}{a}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{0} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} \geq 1 \land \StateVar_{2} \geq 1\right)\\
-  &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{1}{c}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{1} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{1}{f}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
-  &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{2}{e}{2}} \geq 1 \land \StateVar_{2} \geq 1\right) \lor \left(\StateVar_{2} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{2}{d}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
+  &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{1}{c}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{1} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{1}{a}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
+  &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{2}{b}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{2} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
   &\StateVar_{3} > 0 \implies \FinalStateVar_{3} = 1
@@
   set of clauses:
+
+  \contents{\item FIXME: this needs rewriting for the new automaton}
 
@@
   & \Set{ 
-    \Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_1}
-  }, \\ 
+    \left<\Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter(\Transition')\right>
+  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_1} } , \\ 
   & \Set{
-    \Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_2}
+    \left<\Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter(\Transition')\right> \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_2}
   }
@@
 \begin{figure}
-  \includegraphics[width=0.75\linewidth]{graphs/\commit-duels-scatter.pdf}
+  \begin{minipage}[b]{0.75\linewidth}
+    \centering 
+      \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter}
+      \caption{Runtime duel: \Calculus{} versus baseline.}
+      \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter}
+      \caption{Runtime duel: \Calculus{} versus \Nuxmv.}
+    \end{minipage}  
   \caption{The pairwise runtime of each instance for instances solved by any backend. Timeouts are reported as a runtime of \RuntimeTimeout.}

### 2022-06-21 bin/verma-to-tex.py

@@ -9,9 +9,9 @@ transitions = [
     (0, "a", 1),
     (0, "b", 2),
     (1, "c", 1),
-    (2, "e", 2),
-    (1, "f", 3),
-    (2, "d", 3),
+    (1, "a", 3),
+    (2, "b", 1),
+    (2, "c", 3),
 ]
 
 states = {
@@ -203,3 +203,5 @@ def main():
             ]
         )
     )
+
+print(main())

### 2022-06-22 main.tex
Ignore generated PDFs

@@
 
-\subsubsection{Producing the Parikh image}
+\subsubsection{Producing the Parikh image}\label{sec:introduction:parikh}
 
@@
 want to find the whole Parikh image of the intersection of $\AcaOrBc{}$ and
-$\SomethingCSomething{}$. In other words, now $f$ is an object (usually a
-vector) counting how often each character occurs in a string. Since extracting
-values from a vector is tiresome, we will use the shorthand notation
-$\TransitionCount{a}$ to refer to the number of letters a that appear.
-E.g.
-$\TransitionCount{a}(\text{curious and curiouser still}) = 1$.
-Note that we still have the property that $\TransitionCount{o}(\text{homomorphism}) = \TransitionCount{o}(\text{homo}) + \TransitionCount{o}(\text{morphism}) = 2 + 1 = 3$.
-If you treat the addition as element-wise vector addition, this approach works for the vector version as well.
+$\SomethingCSomething{}$. In other words, now $f$ is a mapping to an object
+(usually seen as a vector) counting how often each character occurs in a string. Since
+extracting values from a vector is tiresome, we will use the shorthand notation
+$\TransitionCount{a}$ to refer to the number of letters a that appear. E.g.
+$\TransitionCount{a}(\text{curious and curiouser still}) = 1$. Note that we
+still have the property that $\TransitionCount{o}(\text{homomorphism}) =
+\TransitionCount{o}(\text{homo}) + \TransitionCount{o}(\text{morphism}) = 2 + 1
+= 3$. If you treat the addition as element-wise vector addition, this approach
+works for the vector version as well.
 
@@
 Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
-1\right], \left[1, 2\right]\right\}$.
+1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
+notation $\#l$ to talk about an individual letter $l$. For example, for the Parikh vector above, we would have $\CountOf{a} = 1$. You have already seen 
+this in Section~\ref{sec:introduction:parikh}.
 
@@
 not context-free that also have semilinear images under~$\ParikhMap$ (e.g.
-$\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) = \ParikhMap((abc)^*) = \#a =
-\#b = \#c \land \#a \geq 0$).
+$\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) = \ParikhMap((abc)^*) = \CountOf{a} =
+\CountOf{b} = \CountOf{c} \land \CountOf{a} \geq 0$).
 
@@
 
-\section{Parikh Images from Products of Automata}
+\section{Parikh Images from Products of Automata}\label{sec:multiple}
 
@@
 \section{Extensions}
-\subsection{Backjumping and Learning No-Goods}
+\subsection{Backjumping and Learning No-Goods}\label{sec:ext:backjumping}
 \subsubsection{An Example}
@@
 
+\contents{
+  \item the unused predicate
+  \item single-arity
+  \item a princess theory
+}
+
 \subsection{Implementing the Baseline}\label{sec:implementing-baseline}
@@
 We implement the baseline approach using the same Presburger solver
-(\Princess{}), input file parser, and automaton implementation as \Catra{}. We
+(\Princess{}), input file parser, and automaton implementation as \Catra. We
 do this in order to make the comparison between methods as fair as possible, and
@@
 only contains states that are both reachable from the initial state and has a
-path to an accepting state.
+path to an accepting state. We never perform any other minimisation on the
+automata for either backend. More complex minimisation was left out since
+performing minimisation on automata with counters is non-trivial, and indeed
+already minimising symbolic automata is itself risks exponential blowup
+\cite{minimising-symbolic}.
 
 \subsection{Heuristics and search strategies}
-\contents{
-  \item Optimisations and tricks
-}
+
+There are a number of choices left unspecified in \Calculus{} as described in
+Sections~\ref{sec:calculus} and~\ref{sec:multiple}. For example, the order of
+materialisation of intermediate products and the order of splitting. In
+addition, the splitting specified in the general calculus is correct but
+inefficient. In this section we describe additional implementation details used
+to enhance our solution.
+
+\subsubsection{Splitting, Materialisation, and Propagation}
+
+We order our rule applications so that we first propagate connectedness if
+possible, then perform materialisation if tractable (see below!), then finally
+resort to splitting if we must.
+
+In addition to the splitting on individual transition variables described in Table~\ref{tbl:rules:single}, where we randomly select a variable whose bounds are not known, we also prefer splitting to sever/include a strongly connected component.
+We randomly select an automaton where we can compute a cut to at least one strongly connecting component that can be severed from the initial state, e.g.
+where the strongly connected component does not contain the initial state and where the sum of the transition variables we have associated with the transitions of a minimal cut between the initial state and an arbitrary state in the component is not known to be greater than 0, but also not equal to 0.
+If there are multiple such strongly connected components we choose one randomly.
+We then proceed to split on the cut as if it were a regular transition, e.g.
+its sum being zero or nonzero.
+In this way we use our splitting strategy to drive the connectedness constraint towards propagation by cutting off loops from the automata.
+
+The implementation of the connectedness constraint is opportunistic and straightforward.
+We compute a set of dead states by performing forward and backwards reachability computations on an automaton, where we disregard any transition whose associated variable is known to be zero.
+After that we add clauses ensuring any transition variable associated with a transition starting in a dead state is zero.
+
+Product materialisation, then, is the final piece of the puzzle.
+In the current implementation, we put off computing intermediate products until at least all but 2 transition variables of one of them have a known status, e.g.
+either is known to be zero or known to be positive.
+The number was chosen experimentally, and we observe a consistent trend that to have a low number and therefore a high threshold for materialisation is better throughout the implementation of other features.
+The other automaton for the product is selected randomly.
+
+\subsubsection{Clause Learning}
+
+Since \Catra{} does not learn very sophisticated clauses yet (such as the ones
+described in Section~\ref{sec:ext:backjumping}), it is turned off by default.
+However, even in its underdeveloped state, clause learning has shown to
+significantly improve the performance on unsatisfiable instances, at the cost of
+a loss in performance for satisfiable instances. Since \Catra{} already
+struggles more with satisfiable instances than unsatisfiable ones, we opted to
+turn the option off by default.
+
+\subsubsection{Random Restarts}
+
+Finally, we perform restarts scaled by the Luby series~\cite{luby}. Experimental
+results have shown this to have a large improvement in performance, which is
+unsurprising given how many random choices we make during solving and how
+tail-heavy our problem is. It seems to be very easy to get stuck with a poor
+choice of product to materialise or transitions to split on.
 

### 2022-06-22 bin/verma-to-tex.py

@@ -3,15 +3,15 @@
 from re import A
 
 
-initial_state = 0
-accepting_states = {3}
+initial_state = "start"
+accepting_states = {"final"}
 transitions = [
-    (0, "a", 1),
-    (0, "b", 2),
-    (1, "c", 1),
-    (1, "a", 3),
-    (2, "b", 1),
-    (2, "c", 3),
+    ("start", "a", "sawA"),
+    ("start", "b", "sawB"),
+    ("sawA", "c", "sawA"),
+    ("sawA", "a", "final"),
+    ("sawB", "b", "start"),
+    ("sawB", "c", "final"),
 ]
 
 states = {

### 2022-06-23 main.tex
Dagens text

@@
 \usepackage{proof}
+\usepackage{cleveref}
 \usepackage{algorithm2e}      
@@
 before its Parikh image. In several instances we have observed while solving
-real-world string constraints, the computation the product of automata exhausts
-the memory of any machine due to the exponential blow-up in size of the product,
-quickly becoming intractable as the number of terms in the product increases.
+real-world string constraints, the computation of the product of automata
+exhausts the memory of any machine due to the exponential blow-up in size of the
+product, quickly becoming intractable as the number of automata in the product
+increases.
 
@@
 solving constraints over groups of synchronised (products of) Parikh automata,
-that is symbolic automata with transitions that increment and decrement integer
-registers and that allows checking the constraints on these registers after
-executing automata.
+that is (finite-alphabet) symbolic automata with transitions that increment and
+decrement integer registers and that allows checking linear constraints on these
+registers after running the automata. Parikh automata without constraints on the
+final values on their registers are also sometimes called cost-enriched
+automata, weighted automata or counter automata, depending on your definitions
+and side-constraints.
+
+\contents{
+  \item Automata intersection emptiness is PSPACE-complete
+  \item Producing the formula is linear, but
+  \item deciding a product of automata on a word is also PSPACE-hard
+  \item try proving that finding a word accepted by a product AND satisfying additional constraints is IN PSPACE, e.g. that we are PSPACE-complete
+}
 
@@
 
-\subsection{A demonstration of our approach}
+\subsection{A demonstration of our approach}\label{sec:introduction:motivation}
 
@@
 
-\subsubsection{Example~2: Parikh Automata}
-
-Another application for this generalisation is to express Parikh
-automata~\cite{parikh-automata}, and this is indeed what the implementation
-described in Section~\ref{sec:implementation} does.
-
-\contents{
- \item what is parikh automata
- \item Give an actual example of a parikh automata
-}
+\subsubsection{Example~2: Parikh Automata}\label{sec:parikh-automata}
+
+Another application for this generalisation is to express Parikh automata, and
+this is indeed what the implementation described in
+Section~\ref{sec:implementation} does. Following~\cite{parikh-automata} as
+filtered through~\cite{expressiveness} (which offers a better formalisation for
+our use), we define a Parikh automaton as an automaton with integer counters
+that are incremented at the transitions, and where we have linear
+constraints on the final values of the counters after running the automaton, or
+formally as in \cref{def:parikh-automata,def:projections}.
+
+In the following definitions, we use (as in~\cite{expressiveness}) the following notation:
+\begin{itemize}
+  \item $\Alphabet$ is an alphabet
+  \item $D$ is a finite subset of $\Naturals^d$, for some $d \in \Naturals^+$ (e.g. vectors of length $d$)
+\end{itemize}
+
+\begin{definition}\label{def:projections}
+  We define two projections on the composite alphabet $\left(\Alphabet \times D\right)^*$:
+  \begin{itemize}
+    \item $\pi_{\Alphabet} := (a, \Vector{v}) \mapsto a$ or the projection on $\Alphabet$ is the monoid morphism from $\left(\Alphabet \times D\right)^*$ to $\Alphabet^*$, and the opposite projection
+    \item $\pi_{v} := (a, \Vector{v}) \mapsto \Vector{v}$, called the
+    \emph{extended Parikh image}.
+  \end{itemize}
+\end{definition}
+
+\begin{definition}\label{def:parikh-automata} A Parikh automaton of dimension $d
+  \geq 0$ is a pair $\Tuple{\Automaton, C}$, where $C \subseteq \Naturals^d$ is a semi-linear set (or, equivalently, a Presburger formula), where $\Automaton$ is a finite automaton with the alphabet $\Alphabet \times D$, where $D \subseteq \Naturals^d$.
+
+  We say that $\Tuple{\Automaton, \varphi}$ \emph{recognises} the language, $\Language(\Automaton, C)$, that is it recognises the language crated by is the $\pi_\Alphabet$-projection on $\Alphabet$ whose $\pi_{v} \subseteq C$.
+\end{definition}
+
+\begin{figure}[ht]
+  \centering
+      \includegraphics[width=0.75\linewidth]{parikh_automaton}
+      \caption{A The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
+      $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
+      formula containing the constraints on the final register values cannot be
+      visualised.}\label{fig:parikh-automaton}
+    \end{figure}
+
+An example of a Parikh automata version of $\AcaOrBc{}$ can be seen in \cref{fig:parikh-automaton}.
+A corresponding Presburger version of the constraint seen in \cref{sec:introduction:motivation} is $\psi(\Vec{v}) = v_1 > v_2$.
+
+Generally, we can encode the problem as follows:
+\begin{enumerate}
+  \item The morphism ($\Map$) we would use is $\pi_{v}$
+  \item The monoid operation is element-wise vector addition, which is trivially commutative.
+  \item \Fudge{We would just add the constraints on the final values!?}
+\end{enumerate}
 
@@
 
-\section{Extensions}
+\section{Extensions}\label{sec:extensions}
+
+This section describes a number of possible extensions to the calculus. Note
+that some are partially or fully implemented in \Catra{} already. In particular,
+we have support for finite symbolic transitions over Unicode alphabets to keep
+practical automata under a reasonable size.
+
 \subsection{Backjumping and Learning No-Goods}\label{sec:ext:backjumping}
+\contents{
+  \item What do we have
+  \item A case where no-goods would be good
+  \item What to add to the calculus?
+}
+
 \subsubsection{An Example}
 \subsection{Symbolic Automata}\label{sec:ext:symbolic}
+
+\contents{
+  \item Should we also write about how to adapt to infinite alphabets?
+}
+
+\subsection{Transducers}\label{sec:ext:transducers}
+
+\contents{\item What actually goes here!?}
 
@@
 
-\begin{lstlisting}[caption={An example input file for \Catra{}, illustrating every major syntax element.
-From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), range labels, and their single-character shorthand syntax, register increments, and constraints on the final values of their counters.}, label=lst:input-example]
-  /* Declare counters. Initially only integers, but other data-types
- could be considered as well. All counters start at zero. */
-counter int x, y, z, length;
-
-// List automata. Automata transitions are labelled with intervals,
-// and can specify counter increments/decrements. A counter can only
-// be used in at most one automaton.
+\begin{lstlisting}[caption={An example input file for \Catra{} for the problem introduced in \cref{sec:introduction:motivation}, illustrating every major syntax element.
+From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]
+  // So far we only support integer counters. Note that we have individual
+  // counters for each automaton to avoid surprises for product construction.
+  counter int l_a, l_b, l_c, r_c;
 
   synchronised {
-    automaton A {
-    init s0;
-
-    s0 -> s1 [0, 10] { x += 1};
-
-    accepting s1, s2;
+  automaton aca_or_bc {
+    init start;
+
+    // We use ASCII values here, this is for lowercase a
+    start -> sawA [97] { l_a += 1 };
+    start -> sawB [98] { l_b += 1 };
+
+    sawB -> start  [98] { l_b += 1 };
+    sawB -> final [99] { l_c += 1 };
+
+    sawA -> sawA [99] { l_c += 1 };
+    sawA -> final [97] { l_a += 1 };
+
+    accepting final;
   };
 
-    automaton B {
-    init t1;
-
-    t1 -> t2 [11, 22] { y += 1};
-
-    accepting t1, t2;
+  automaton something_c_something {
+      init start;
+
+      // Special short-hand value for the whole Unicode alphabet.
+      start -> start [any] ;
+      start -> final [99] { r_c += 1 };
+
+      final -> final [any];
+
+      accepting final;
   };
@@
 
-automaton C {
-    init s0;
-
-    s0 -> s1 [15] { z += 1 };
-    s1 -> s1 [15] { z += 1};
-
-    accepting s1;
-};
-
-// Specify a constraint on the final values of counters; check whether
-// there is a word accepted by all automata for which the final counter
-// values satisfy this constraints.
-constraint z = 2x + 2y && !(x = 0);
+  // Constrain the number of a:s to be larger than the number of c:s. Since the
+  // automata are synchronised on every transition and counters are guaranteed
+  // to be consistent these constraints are sufficient without also constraining
+  // r_c.
+  constraint l_a > l_c;
 \end{lstlisting}
@@
 After generating an initial~\InitialNrBenchmarks{}, we remove~\NrBroken{} misgenerated instances that did not parse, as well as~\NrTrivial{} instances solved in under five seconds by the baseline backend.
-The benchmarks are run on commit \texttt{\commit}.
+The benchmarks are run on commit~\texttt{\commit}.
 
@@
 \subsection{Threats to Validity}
-\contents{
-  \item There might be better implementations
-  \item We might have over-biased
-  \item Our benchmarks might be biased
-  \item The underlying automata implementation might be bad and therefore skew
-  the results.
-
-}
+
+The most obvious threats to validity would be poor benchmarking or poor
+implementation, e.g. if the method described in \cref{sec:calculus,sec:multiple}
+deviate from what is actually benchmarked in section \cref{sec:experiments}, or
+if the methods used for benchmarking would be unsound. The first problem cannot
+easily be addressed, except by stating that we have been careful in our
+implementation and descriptions alike. The second is more practical. To increase
+the probability that our results are indeed representative both in the sense of
+representing runs on expected, real world input and in the statistical sense for
+a given run of \Catra, despite the fact that we use randomness in our
+implementation, we execute many experiments. During the development of \Catra{},
+we have also executed experimeints on different hardware and software
+configurations on smaller samples of instances to ensure that the general trend
+holds. The logs from these executions are distributed along with notes as part
+of the artefact in the interest of transparency. Moreover, we have validated
+\Fudge{all reported satisfiable instances and their witnesses} with \Nuxmv to
+ensure that \Calculus{} is indeed sound and does not cheat.
+
+Another threat to validity would be if our implementations of the competition
+(baseline and \Nuxmv) would deviate significantly from the expected
+implementations. For \Nuxmv{} we use the default configuration which we believe
+should be performant (or it should not be the default). Additionally, tweaking our invocation of \Nuxmv is explicitly made easy for artefact reviewers.
+
+The most important and most probable threat to validity is our choice of
+implementation platform and our choice of automata library, that is our choice
+not to use one. Since our automata are more advanced than anything supported by
+e.g. the BRICS library, we opted to use our own implementation. There are some
+signs that in particular our product computation is inefficient, notably the
+good performance of low-threshold automata materialisation that prioritises
+computing smaller products. This means that our implementation of \Calculus{}
+makes much less heavy use of our product construction library than the baseline
+implementation does, instead putting its weight into \Princess{}' theorem
+proving abilities. This situation would unfairly advantage \Calculus{} if our
+automata library was indeed the bottleneck. We believe this is unlikely since
+similar performance issues as the ones in baseline in our string solver was the
+instigating reason for this research. Additionally, while we may have a
+constant-factor performance impact from bad implementation, it is less likely
+for our potentially poor implementation to be asymptotically worse than e.g.
+BRICS. However, the performance improvement of \Calculus{} over baseline seen in
+\cref{sec:experiments} is generally not constant.
 

### 2022-06-29 main.tex
Dagens text

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=true}
+\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=true}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 
-\usepackage{amsthm}
-\usepackage{amsmath,mathtools,empheq,fancybox}
+\usepackage{amsmath,empheq,fancybox}
 \usepackage{paralist}
@@
 \usepackage{mathpartir}
+\usepackage{amsthm}
+
+\usepackage{mathtools} % Bonus
+
 \theoremstyle{definition}
@@
 utility of the formulation introduced here is that it allows us to \emph{remove}
-something, thereby obtaining an easier problem.
+something, thereby potentialy obtaining an easier problem.
 
@@
 
-\begin{definition}
+\begin{definition}\label{def:single-image}
   The Parikh predicate, $\SinglePredicateInstance$, for some automaton
@@
 
-\subsection{The calculus is correct}
+\subsection{\Calculus{} for one automaton is correct}\label{sec:single:correct}
 
@@
 we use more exact definitions than the traditional soundness and completeness.
-
-\subsubsection{The calculus terminates}
-\subsubsection{The calculus allows no values outside the Parikh image}
-\subsubsection{The calculus excludes no values in the Parikh image}
-
-\subsection{Finding the Presburger representation of a Parikh Image}
+Since correctness is guaranteed by the formulation of the predicate, our
+correctness proof means proving inductively that no rule would invalidate that
+definition.
+
+\subsubsection{\Calculus{} terminates}
+\begin{lemma}\label{lma:single-terminates}
+  A solver implementing the calculus described in \cref{sec:calculus} starting
+  with some decidable (modulo the solver) set of clauses $\SomeClause{}$ and an
+  instance of our predicate $\SinglePredicateInstance$ will eventually
+  terminate.
+\end{lemma}
+
+\begin{proof}
+The first remaining source of divergence is from leftover predicates. The
+$\Image$ predicate can always be trivially removed by applying \Expand{}. This
+means a leftover $\Image$ predicate can never cause divergence. The only
+remaining predicate then is $\Connected$, which can always be removed by
+applying $\Subsume$ when no other of the rules is applicable. Hence all
+predicates introduced in the calculus can always be removed.
+
+It remains to prove that each rule is only applied finitely many times. This is
+almost obvious since both \Propagate{} and \Split{} can only be applied to
+transition variables whose $\Filter$ value has no known bound. However, \Split{}
+can always be used to split the proof tree and add clauses in precisely the
+right shape, thereby guaranteeing the ability to make forward progress.
+Therefore, the number of transition variables whose bounds are unknown is
+decreasing monotonically for any given proof goal, and hence termination is
+guaranteed.
+\end{proof}
+
+\subsubsection{\Calculus{} computes the homomorphic image}
+\begin{lemma}\label{lma:single-correct}
+  Starting with $\SinglePredicateInstance$, we will close the goal with $\top$
+  if $\exists s \in \Language(\Automaton) \HoldsThat \Map(s) = \MonoidElement$
+  and $\bot$ otherwise.
+\end{lemma}
+\begin{proof}
+Since we are always guaranteed to terminate following
+\cref{lma:single-terminates}, we know that we will close the goal eventually.
+Additionally, the property of \cref{lma:single-terminates} is trivially obtained
+from the semantics of the $\Image$ predicate as defined in
+\cref{def:single-image}. This means that we only need to prove the preservation
+of those properties to prove \cref{lma:single-correct}.
+
+\begin{lemma}\label{lma:expand-preserves}
+  \Expand{} preserves the semantics of $\Image$.
+\end{lemma}
+\begin{proof}
+
+  After application of \Expand{}, $\FlowEq$ guarantees that any
+  $\Filter(\Transition)$  for every $\Transition \in \Automaton$ is consistent
+  with the flow through the automaton. After its application, the proof goal
+  either contains an unsatisfiable set of Presburger formulae if the flow is
+  infeasible, otherwise a satisfiable set. An infesaible flow can happen in
+  either of these conditions:
+  \begin{itemize}
+    \item if $\Filter$ demands that some transition $\Transition$ is visited more times than would be allowed by $\Automaton$, or
+    \item if the path described by $\Filter$ is not consistent with a path from
+    $\InitialState$ to some state $\State$ in $\AcceptingStates$.
+  \end{itemize}
+
+  Note that \Expand{} is taken directly from the existing formulation of
+  \cref{eq:generate-parikh}.
+\end{proof}
+
+\begin{lemma}\label{lma:split-preserves}
+  \Split{} preserves the semantics of $\Image$.
+\end{lemma}
+\begin{proof}
+  \Split{} allows us to case-split at most once on each transition $\Transition
+  \in \Transitions$. This means that our proof tree at its leaves will be
+  guaranteed to have all transitions of known bounds in the correct format. This
+  means that once \Split{} is at fixpoint, the status (present or absent) of any
+  transition is known. Therefore it preserves the semantics of the $\Image$
+  predicate.
+\end{proof}
+
+\begin{lemma}\label{lma:propagate-preserves}
+  \Propagate{} preserves the semantics of $\Image$.
+\end{lemma}
+\begin{proof}
+  \Propagate{} will lazily enforce connectedness to any state. If it can be
+  applied, there exists a cut $C$ in $\Automaton$ such that some state $\State$
+  for any incoming transition $\Transition \in \Transitions_\Automaton$ has
+  $\Filter(t) = 0$  as part of the equations of the proof goal, and such that
+  $\State$ has an outgoing transition $\Transition'$ whose lower bound is not
+  $0$. Adding this conclusion will make equations unsatisfiable precisely if
+  there exists a transition $\Transition \in \Transitions_\Automaton$ such that
+  $\Filter(\Transition)$ is inconsistent with the homomorphic image, since it
+  implies that the selection of transitions is disconnected.
+\end{proof}
+
+\begin{lemma}\label{lma:subsume-preserves}
+  \Subsume{} preserves the semantics of $\Image$.
+\end{lemma}
+\begin{proof}
+  Since this rule can \emph{only} be applied if no other rule can, it means that
+  the upper and lower bounds of any transition $\Transition \in \Transitions$ as
+  defined by $\Filter(\Transition)$ is known, meaning clauses of the shape
+  $\Filter(\Transition) > 0$ or $\Filter(\Transition) = 0$ are present in the
+  proof goal for every transition and that the $\Image$ predicate has been
+  consumed by \Expand{} to be replaced by $\Connected$. This in turns means that
+  the goal can be closed only if the assignment of $\Filter$ is consistent with
+  an accepting path through $\Automaton$.
+
+  It also means that \Propagate{} must have been applied whenever possible, meaning that for any transition $\Transition$ starting in a state $\State$ that is disconnected the corresponding $\Filter$ values, we must have applied $\Propagate{}$ to derive $\Filter(\Transition) = 0$.
+Note that for an infeasible assignment (such as an unsatisfiable instance) this would contradict other decisions, but it would preserve the semantics of $\Image$.
+\end{proof}
+
+Since the final step we execute is \Subsume, which we know from
+\cref{lma:subsume-preserves} preserves the semantics of $\Image$ and each step
+along the way
+(\cref{lma:expand-preserves,lma:split-preserves,lma:propagate-preserves}), it
+holds that \Calculus{} for one automaton also preserves the semantics.
+\end{proof}
+
+\subsection{Finding the Presburger representation of a homomorphic image}
 \contents{
@@
 To generalise the calculus to calculations on products of automata, we change the main predicate to take arbitrarily many automata:
-\begin{definition}
+\begin{definition}\label{def:multiple}
   $\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}$
@@
 \subsection{The extended calculus is also correct}
+
+Since the expanded calculus is in practice the addition of two rules with the
+purpose of reducing an~$\Image$ predicate containing a product to the form
+solved by the previous single-automaton rules of \cref{tbl:rules:single}, we
+extend the reasoning from \cref{sec:single:correct} to the single-automaton
+rules of \cref{tbl:rules:multi}.
+
 \subsubsection{The calculus terminates}
-\subsubsection{The calculus allows no values outside the Parikh image}
-\subsubsection{The calculus excludes no values in the Parikh image}
+\begin{lemma}
+  The addition of the \ExpandM{} and \Materialise{} rules in \cref{tbl:rules:multi}
+  maintains termination as shown in \cref{lma:single-terminates}.
+\end{lemma}
+\begin{proof}
+  Similarly to how we previously showed that the number of possible executions
+  of the \Split{} rule is bounded above by the number of transitions of an
+  automaton, we can bound the number of applications of \Materialise{} by the
+  (monotonically) decreasing number of automata in the product until we have
+  approached the starting state for the single-automaton calculus. Trivially, we
+  can eagerly apply \Materialise{} repeatedly to do so. Similarly, \ExpandM{} is
+  just the generalisation of \Expand{} to multiple automata. In other words there are no possibilities of the calculus diverging.
+\end{proof}
+
+\subsubsection{The calculus implemements the homomorphic image of the product}
+\begin{lemma}
+  Using \Calculus{} to compute $\Image_{\Automaton_1\times\ldots\times\Automaton_k, \Map}(\Filter, \MonoidElement)$ is equivalent to using the single-automaton rules to compute $\Image_{\Automaton', \Map}(\Filter, \MonoidElement)$ where $\Automaton' = \Automaton_1\times\ldots\times\Automaton_k$.
+\end{lemma}
+\begin{proof}
+
+  We perform this proof equivalent to set inclusion: by proving that neither predicate is stronger than the other, in the sense of rejecting an element that the other accepts and vice versa.
+
+\begin{lemma}\label{lma:multi:rinclude}
+  There exists no element $\MonoidElement$ such that $\Image_{\Automaton_1\times\ldots\times\Automaton_k, \Map}(\Filter, \MonoidElement) = \top$ but $\Image_{\Automaton', \Map}(\Filter, \MonoidElement) = \bot$.
+\end{lemma}
+\begin{proof}
+  Assume that such an $\MonoidElement$ exists with the goal of deriving a
+  contradiction. Under this assumption, would we be able to use the two new
+  rules to close the proof goal with $\top$?
+
+  The \ExpandM{} rule would allow us to close no goal since it only adds
+  additional clauses. Our only hope is in using \Materialise{} to get rid of the
+  product step by step without introducing an unsatisfiable clause. However, all
+  \Materialise{} does is computing the product step by step, eventually arriving
+  at $\Automaton'$ modulo permutations. Therefore, it can be no more satisfiable
+  than the single-automata predicate.
+
+  \contents{\item Do I need to worry about the fact that I have no extra
+  clauses/$\SomeClause$ here? If present, I would need to apply the definition
+  of \BindingSum{} since that is what transfers constraints on individual
+  automata up and down the materialisation.}
+\end{proof}
+
+\begin{lemma}\label{lma:multi:linclude}
+  There exists no element $\MonoidElement$ such that $\Image_{\Automaton_1\times\ldots\times\Automaton_k, \Map}(\Filter, \MonoidElement) = \top$ but $\Image_{\Automaton', \Map}(\Filter, \MonoidElement) = \bot$.
+\end{lemma}
+\begin{proof}
+  Assume that such an element exists. Then $\MonoidElement =
+  \Map(\WordOf(\Path))$ for some $\Path$ ending up in an accepting state in
+  $\Automaton'$, and consistent with $\Filter$ along the transitions of $\Path$.
+  Then a corresponding path must exist in each automata of $\Automaton'$,
+  although possibly with different labels.
+
+  For \ExpandM{} to introduce unsatisfiability, either of its automata must be
+  disconnected (or its corresponding $\Connected{}$ predicate can be removed
+  with \Propagate{}), or its flow must be infeasible. However, this contradicts
+  the existence of some corresponding path $\Path$ in each automaton. Hence,
+  \ExpandM{} cannot be the source of the unsatisfiability.
+
+  The same applies for the second half of \Materialise{}, if the corresponding
+  path exists in each individual automaton and in $\Automaton'$, it must also
+  exist in every intermittent product on the way to $\Automaton'$, and if $\Map$
+  is consistent for intermittent transition labels as it must be, then neither the intermittent $\Image$ predicates nor the final one can be the source of the unsatisfiability.
+
+  This leaves one final potential source of unsatisfiability: the first half of
+  \Materialise{}: \BindingSum{}. For that to be unsatisfiable, it requires
+  intermittent automata $\Automaton_l, \Automaton_r$ such that adding
+  $\BindingSum(\Automaton_l \times \Automaton_r)$ is unsatisfiable. The
+  generated equations state that the transitions of $\Automaton_l$ and
+  $\Automaton_r$ must be used precisely as often as their corresponding results
+  in $\Automaton_l \times \Automaton_r$. For this set of equations to be
+  unsatisfiable there must exist no path through $\Automaton_l \times
+  \Automaton_r$ that preserves this. However, this is true from the definition
+  of the product of automata: since we know the product to be nonempty, it must
+  in some sense contain at least one path whose equivalent exists in both
+  $\Automaton_l$ and $\Automaton_2$ and cannot use a corresponding transition
+  more often, or it would recognise a different langauge. Hence no such case
+  exists.
+\end{proof}
+
+Since it follows from \cref{lma:multi:linclude,lma:multi:rinclude} that both
+variants of the predicate are precisely equally strong, it must mean that they
+are equivalent, and we have proven the correctness as defined in
+\cref{def:multiple}.
+\end{proof}
 
@@
 baseline back-end}. An example of an input file \Fudge{illustrating every feature
-of the file format} can be seen in Listing~\ref{lst:input-example}.
-
-\begin{lstlisting}[caption={An example input file for \Catra{} for the problem introduced in \cref{sec:introduction:motivation}, illustrating every major syntax element.
-From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]
-  // So far we only support integer counters. Note that we have individual
-  // counters for each automaton to avoid surprises for product construction.
-  counter int l_a, l_b, l_c, r_c;
-
-  synchronised {
-  automaton aca_or_bc {
-    init start;
-
-    // We use ASCII values here, this is for lowercase a
-    start -> sawA [97] { l_a += 1 };
-    start -> sawB [98] { l_b += 1 };
-
-    sawB -> start  [98] { l_b += 1 };
-    sawB -> final [99] { l_c += 1 };
-
-    sawA -> sawA [99] { l_c += 1 };
-    sawA -> final [97] { l_a += 1 };
-
-    accepting final;
-  };
-
-  automaton something_c_something {
-      init start;
-
-      // Special short-hand value for the whole Unicode alphabet.
-      start -> start [any] ;
-      start -> final [99] { r_c += 1 };
-
-      final -> final [any];
-
-      accepting final;
-  };
-  };
-
-  // Constrain the number of a:s to be larger than the number of c:s. Since the
-  // automata are synchronised on every transition and counters are guaranteed
-  // to be consistent these constraints are sufficient without also constraining
-  // r_c.
-  constraint l_a > l_c;
-\end{lstlisting}
+of the file format} can be seen in \cref{lst:input-example}.
 
@@
 \appendix
-% \section{Appendix}
+\section{Appendix}
+
+\begin{lstlisting}[caption={An example input file for \Catra{} for the problem introduced in \cref{sec:introduction:motivation}, illustrating every major syntax element.
+From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]
+  // So far we only support integer counters. Note that we have individual
+  // counters for each automaton to avoid surprises for product construction.
+  counter int l_a, l_b, l_c, r_c;
+
+  synchronised {
+  automaton aca_or_bc {
+    init start;
+
+    // We use ASCII values here, this is for lowercase a
+    start -> sawA [97] { l_a += 1 };
+    start -> sawB [98] { l_b += 1 };
+
+    sawB -> start  [98] { l_b += 1 };
+    sawB -> final [99] { l_c += 1 };
+
+    sawA -> sawA [99] { l_c += 1 };
+    sawA -> final [97] { l_a += 1 };
+
+    accepting final;
+  };
+
+  automaton something_c_something {
+      init start;
+
+      // Special short-hand value for the whole Unicode alphabet.
+      start -> start [any] ;
+      start -> final [99] { r_c += 1 };
+
+      final -> final [any];
+
+      accepting final;
+  };
+  };
+
+  // Constrain the number of a:s to be larger than the number of c:s. Since the
+  // automata are synchronised on every transition and counters are guaranteed
+  // to be consistent these constraints are sufficient without also constraining
+  // r_c.
+  constraint l_a > l_c;
+\end{lstlisting}
 

### 2022-07-01 main.tex
Dagens text

@@
 \usepackage{amsthm}
+\usepackage{ebproof}
+\usepackage{stackengine}
 
@@
 produce~\cite{muscholl-linear}. However, the resulting existentially quantified
-clauses are costly to eliminate, in practice making many real-world problems
-intractable.
-
-In addition to the costly elimination of the existentially quantified variables,
-another problem arises for the common case of solving length constraints in
-string solvers. Conjunctions of string constraints lead to the computation of
-Parikh images of products of regular languages. Applying the approach
-in~\cite{generate-parikh-image} would require the computation of the product
-before its Parikh image. In several instances we have observed while solving
-real-world string constraints, the computation of the product of automata
-exhausts the memory of any machine due to the exponential blow-up in size of the
-product, quickly becoming intractable as the number of automata in the product
-increases.
+clauses are costly to eliminate as the number of variables increases, in
+practice making many real-world problems intractable.
+
+In addition to the elimination of existentially quantified variables, another
+problem arises for the common case of solving length constraints in string
+solvers. Conjunctions of string constraints lead to the computation of Parikh
+images of products (intersections) of regular languages represented as finite
+automata. Applying the approach in~\cite{generate-parikh-image} would require
+the eager computation of the product before its Parikh image. In several
+instances we have observed while solving real-world string constraints, the
+computation of the product of automata exhausts the memory of any machine
+available due to the exponential blow-up in size of the product, quickly
+becoming intractable as the number of automata in the product increases.
 
@@
 interleave the computation of arbitrarily deep products of automata with the
-product's Parikh image. This enables us to let both calculations inform each
-other, eliminating unnecessary work, and pruning the size of the partial
-products considered in the computation for a smaller memory footprint. Moreover,
-the method can be used iteratively to tackle smaller chunks of the product
-incrementally, thereby avoiding running out of memory. We implement this
-calculus, along with the approximate method of~\cite{approximate-parikh}, its
-fall-back variant adapted from~\cite{generate-parikh-image}, and an adapter for
-the \Nuxmv{} model checker, in a tool that we call \Catra{}. \Catra{} allows
-solving constraints over groups of synchronised (products of) Parikh automata,
-that is (finite-alphabet) symbolic automata with transitions that increment and
+product's Parikh image, and is generalised to an arbitrary Parikh-like
+homomorphism over automata labels. This enables us to let both calculations
+inform each other, eliminating unnecessary work, and pruning the size of the
+partial products considered in the computation for a smaller memory footprint.
+Moreover, the method can be used iteratively to tackle smaller chunks of the
+product incrementally, thereby avoiding running out of memory at all. We
+implement this calculus, along with the approximate method
+of~\cite{approximate-parikh}, its fall-back variant adapted
+from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
+checker, in a tool that we call \Catra{}. \Catra{} allows solving constraints
+over groups of synchronised (products of) Parikh automata, that is
+(finite-alphabet) symbolic automata with transitions that increment and
 decrement integer registers and that allows checking linear constraints on these
@@
 automata, weighted automata or counter automata, depending on your definitions
-and side-constraints.
-
-\contents{
-  \item Automata intersection emptiness is PSPACE-complete
-  \item Producing the formula is linear, but
-  \item deciding a product of automata on a word is also PSPACE-hard
-  \item try proving that finding a word accepted by a product AND satisfying additional constraints is IN PSPACE, e.g. that we are PSPACE-complete
-}
+and side-constraints. In the satisfaction mode, we are more formally solving the
+intersection problem of Parikh automata (under homomorphism), known to be
+PSPACE-complete~\cite{graph-queries}.
 
@@
 \begin{itemize}
-\item The \Calculus{} calculus to efficiently compute the Parikh image of products of Parikh automata.
+\item The \Calculus{} calculus to efficiently compute (a homomorphism on) the
+Parikh image of products of Parikh automata.
 \item A specialised method for quantifier elimination that blends well with \Calculus{} to efficiently compute quantifier-free Presburger representations of a given Parikh image.
-\item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future studies.
+\item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future study.
 \item The \Catra{} tool for solving such instances, containing an implementation of \Calculus{}, the over-approximation described in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model checker~\cite{nuxmv}.
 \item Suggestions for how to efficiently implement \Calculus{} in a modern automated theorem prover, including strategies for case splitting, clause learning, and constraint propagation for connectedness.
-\item Finally, a published description of the well-known bug in~\cite{generate-parikh-image} previously only described in folklore.
 \end{itemize}
@@
 constraint for which we want to find values in $f$'s image over the intersection
-of the wto languages. We will use the same two regular languages, recognised by
+of the two languages. We will use the same two regular languages, recognised by
 two automata; $\SomethingCSomething{}$ (Figure~\ref{fig:something-c-something})
@@
 $\SomethingCSomething{}$. In other words, now $f$ is a mapping to an object
-(usually seen as a vector) counting how often each character occurs in a string. Since
-extracting values from a vector is tiresome, we will use the shorthand notation
-$\TransitionCount{a}$ to refer to the number of letters a that appear. E.g.
-$\TransitionCount{a}(\text{curious and curiouser still}) = 1$. Note that we
-still have the property that $\TransitionCount{o}(\text{homomorphism}) =
-\TransitionCount{o}(\text{homo}) + \TransitionCount{o}(\text{morphism}) = 2 + 1
-= 3$. If you treat the addition as element-wise vector addition, this approach
-works for the vector version as well.
+(usually represented as a column vector) counting how often each character
+occurs in a string. Since extracting values from a vector is tiresome, we will
+use the shorthand notation $\TransitionCount{a}$ to refer to the number of
+letters a that appear. E.g. $\TransitionCount{a}(\text{curious and curiouser
+still}) = 1$. Note that we still have the property that
+$\TransitionCount{o}(\text{homomorphism}) = \TransitionCount{o}(\text{homo}) +
+\TransitionCount{o}(\text{morphism}) = 2 + 1 = 3$. If you treat the addition as
+element-wise vector addition, this approach works for the vector version as
+well.
 
@@
 & \AndComp{\Transition = \FromLabelTo{}{}{\State} \in \Transitions}{
-  \TransitionVar_{\Transition} > 0 \implies \StateVar_{\State}
+  \TransitionVar_{\Transition} > 0 \implies \StateVar_{\State} > 0
 } \\
@@
   \StateVar_{\State} > 0 \implies
-  \left(\FinalStateVar_{\State} = 1 \text{ if } \State \in \AcceptingStates\right) \OrComp{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions}{
+  \left(\FinalStateVar_{\State} = 1 \text{ if } \State \in \AcceptingStates \land \right) \OrComp{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions}{
     \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
@@
 
-\contents{
-  \item mention the bug here!
-}
-
 In this paper we refer to this model as the baseline approach, though we also
@@
   \ParikhMap(\Automaton):=\:
-  &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} + \TransitionVar_{\FromLabelTo{1}{a}{3}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{2}{b}{1}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} + \TransitionVar_{\FromLabelTo{2}{c}{3}}\\
+  &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} + \TransitionVar_{\FromLabelTo{1}{a}{3}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{2}{b}{0}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} + \TransitionVar_{\FromLabelTo{2}{c}{3}}\\
   &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
-  & \TransitionVar_{\FromLabelTo{0}{a}{1}} + \TransitionVar_{\FromLabelTo{0}{b}{2}} - 0 = 1 \land  \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{2}{b}{1}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
-  & \TransitionVar_{\FromLabelTo{2}{b}{1}} + \TransitionVar_{\FromLabelTo{2}{c}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3} + 0 - \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{2}{c}{3}} = 0\\
+  &\TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{0}{a}{1}} - \TransitionVar_{\FromLabelTo{2}{b}{0}} = 1 \land  \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
+  & \TransitionVar_{\FromLabelTo{2}{b}{0}} + \TransitionVar_{\FromLabelTo{2}{c}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3}  - \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{2}{c}{3}} = 0\\
   &\TransitionVar_{\FromLabelTo{0}{a}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} > 0 \implies \StateVar_{2} > 0\\
   &\TransitionVar_{\FromLabelTo{1}{c}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{1}{a}{3}} > 0 \implies \StateVar_{3} > 0\\
-  &\TransitionVar_{\FromLabelTo{2}{b}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} > 0 \implies \StateVar_{3} > 0\\
+  &\TransitionVar_{\FromLabelTo{2}{b}{0}} > 0 \implies \StateVar_{0} > 0 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} > 0 \implies \StateVar_{3} > 0\\
   &\StateVar_{0} > 0 \implies \left(\StateVar_{0} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{0}{a}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{0} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} \geq 1 \land \StateVar_{2} \geq 1\right)\\
   &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{1}{c}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{1} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{1}{a}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
-  &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{2}{b}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{2} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
+  &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{0} + 1 \land \TransitionVar_{\FromLabelTo{2}{b}{0}} \geq 1 \land \StateVar_{0} \geq 1\right) \lor \left(\StateVar_{2} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
   &\StateVar_{3} > 0 \implies \FinalStateVar_{3} = 1
@@
   automaton ($\TransitionVar_{\FromLabelTo{0}{b}{2}} +
-  \TransitionVar_{\FromLabelTo{0}{a}{1}} = 1$), and the dependency between that
+  \TransitionVar_{\FromLabelTo{0}{a}{1}} -
+  \TransitionVar_{\FromLabelTo{2}{b}{0}} = 1$), and the dependency between that
   choice and the following outgoing transition, e.g. in
-  $\TransitionVar_{\FromLabelTo{2}{d}{3}} -
-  \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0$. Here we only have one choice of
+  $\TransitionVar_{\FromLabelTo{1}{a}{3}} -
+  \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0$. Here we only have one choice of
   incoming flow; from $\FinalStateVar_3$, corresponding to the only accepting
-  state. Simplifying the equation using standard algebra gives us this smaller
-  set of clauses:
-
-  \contents{\item FIXME: this needs rewriting for the new automaton}
+  state.
+
+Note that one of the $\lor$-clauses can be immediately removed since it
+contains the contradiction $\StateVar_{1} = \StateVar_{1} + 1$, arising from
+cycles. These clauses encode the distance in a spanning tree from an accepting
+state, and the unsatisfiability is by design, essentially ensuring that two
+states in a cycle cannot vouch for each other's connectedness.
+
+Assume we are really interested in occurrences with the letter a, i.e. we solve
+for $\LetterVar_a > 0$. This implies that $\StateVar_{1} > 0 \land \StateVar_{3}
+> 0$ (essentially, that the state our transition of interest started in is
+reachable), in turn implying that $\FinalStateVar_3 = 1$. Plugging in these
+facts, we get:
 
@@
   \begin{aligned}
-    &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
-    & \LetterVar_{d} + \LetterVar_{a} = 1\\
-    & \FinalStateVar_{3}  > \LetterVar_{d} \implies \StateVar_{1} > 0 \land \LetterVar_{d} > 0 \implies \StateVar_{2} > 0\\
-    &\LetterVar_{c} > 0 \implies \StateVar_{1} > 0 \land \LetterVar_{e} > 0 \implies \StateVar_{2} > 0\\
-    &\FinalStateVar_{3}  > \LetterVar_{d} \implies \StateVar_{3} > 0 \land \LetterVar_{d} > 0 \implies \StateVar_{3} > 0\\
-    &\StateVar_{0} > 0 \implies \left(\StateVar_{0} = \StateVar_{1} + 1 \land (\FinalStateVar_{3}  - \LetterVar_{d}) \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{0} = \StateVar_{2} + 1 \land \LetterVar_{d} \geq 1 \land \StateVar_{2} \geq 1\right)\\
-    &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{3} + 1 \land (\FinalStateVar_{3}  - \LetterVar_{d}) \geq 1 \land \StateVar_{3} \geq 1\right)\\
-    &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{3} + 1 \land \LetterVar_{d} \geq 1 \land \StateVar_{3} \geq 1\right)\\
-    &\StateVar_{3} > 0 \implies \FinalStateVar_{3} = 1
-    \end{aligned}
-    \label{eq:parikh-formula:reduced}
-    \end{equation}
-
-Note that some of the or-clauses disappear since they contain false
-inequalities of the type $x = x + 1$, arising from cycles. These clauses encode
-the distance in a spanning tree from an accepting state, and the
-unsatisfiability is by design, essentially ensuring that two states in a cycle
-cannot vouch for each other's connectedness.
-
-Assume we are really interested in occurrences with the letter e, i.e. we solve
-for $\LetterVar_e > 0$. This implies that $\StateVar_{2} > 0$ (essentially, that
-the state our transition of interest started in is reachable), in turn implying
-that $\StateVar_{2} = \StateVar_{3} + 1 \land \LetterVar_{d} \geq 1 \land
-\StateVar_{3} \geq 1$. Armed with this knowledge, we now know that $\StateVar_3
-> 0$, which means that $\FinalStateVar_3 = 1$. Since we also know that
-$\LetterVar_d \geq 1$, we know that $\LetterVar_a = 0$. This corresponds to
-choosing the bottom path through the automaton. Applying this to the equation
-and performing standard boolean logic propagation gives the following:
-
-\begin{equation}
-  \begin{aligned}
-  &\LetterVar_{d} = 1 \land \LetterVar_{a} = 0 \\
-  & \StateVar_{3} > -1 \\
-  &\LetterVar_{c} = 0\\
+  &\LetterVar_{a} = 2 \land \LetterVar_{b} = 0 \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}}\\
+  & \StateVar_{3} > 0\\
   &\StateVar_{0} > 0 \implies \StateVar_{0} = \StateVar_{3} + 2\\
   \end{aligned}
-  \label{eq:parikh-formula:choose-a}
+  \label{eq:parikh-formula-calc}
   \end{equation}
 
-  Since the final implication holds (we know $\StateVar_3 \geq 0$, and therefore
-  $\StateVar_3 + 2 > 0$), we can disregard it. Thereby we have eliminated all
-  quantified variables and are left with a formula containing only the free
-  ones. Note that all constraints on $\LetterVar_e$ have disappeared after we
-  asserted that it was strictly positive. This is no coincidence: since it is a
-  loop, it can take any positive value.
+We can then proceed to eliminate the remaining existentially quantified
+variables to obtain the formula $\LetterVar_{a} = 2 \land \LetterVar_{b} = 0
+\land \LetterVar_{c} \geq 0$ without too much effort.
 
@@
   fact that even few additional constraints on the Parikh image can dramatically
-  reduce the difficulty of computing it, as in this extreme case where only one
-  additional choice in practice fixes the values of all other symbols in the image.
+  reduce the difficulty of computing it.
 
@@
 
-To generalise the logic over Parikh images, we introdouce a morphism $\Map:\: \Sigma
+To generalise the logic over Parikh images, we introduce a morphism $\Map:\: \Sigma
 \to \Monoid$ where $\Monoid = \left(X;+_{\Monoid};0_{\Monoid}\right)$ is a commutative monoid where the following holds:
@@
 \begin{aligned}
-  & \FlowEq(\Automaton, \Filter) = \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)}
-\text{ $\geq 0$ if $\State \in \AcceptingStates$, $= 0$ otherwise}\\
-  & \In(\State, \Filter) = \left(\text{$1$ if $\State = \InitialState$, otherwise $0$}\right) + \sum_{\Transition = \FromLabelTo{\State_0}{}{\State} \in \Transitions} \Filter(\Transition)\\
+  & \FlowEq(\Automaton, \Filter) = \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \\
+  & \Sink(\State) = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
+  & \In(\State, \Filter) = \StartFlow(\State) + \sum_{\Transition = \FromLabelTo{\State_0}{}{\State} \in \Transitions} \Filter(\Transition)\\
+  & \StartFlow(\State)  = 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
   & \Out(\State, \Filter) = \sum_{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions} \Filter(\Transition)
@@
 $$
-Where we transparently assign fresh variables to each state $\State \in \States$
-and existentially quantify them in the whole clause.
+Where we transparently assign fresh variables to each state $\State \in
+\States$ as well as every accepting state $\State' \in \AcceptingStates$ and
+existentially quantify them in the whole clause.
 
@@
 \subsection{An Example}
+
+Starting with~$\AcaOrBc{}$, where $\Map$ is the length function, in effect
+$\Transition \mapsto 1$ for transitions, and the constraints that the length is
+odd using the same trick as in \cref{sec:introduction:motivation}, we have the
+definitions in \cref{fig:example:single:equivalences} (omitting existential
+quantifiers and $x \geq 0$ for every variable to avoid clutter).
+
+\begin{figure}[ht]
+  We define $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$ as follows:
+
+  \begin{minipage}[b]{0.45\linewidth}
+    \begin{equation*}
+      \begin{aligned}
+        & \Filter(\FromLabelTo{start}{a}{sawA}) & = \TransitionVar_1 \\
+        & \Filter(\FromLabelTo{start}{b}{sawB}) & = \TransitionVar_2 \\
+        & \Filter(\FromLabelTo{sawA}{c}{sawA})  & = \TransitionVar_3  \\
+        & \Filter(\FromLabelTo{sawB}{b}{start}) & = \TransitionVar_4 \\
+        & \Filter(\FromLabelTo{sawA}{a}{final}) & = \TransitionVar_5 \\
+        & \Filter(\FromLabelTo{sawB}{c}{final}) & = \TransitionVar_6 \\
+      \end{aligned}
+    \end{equation*}    
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[b]{0.45\linewidth}
+    \begin{equation*}
+      \begin{aligned}
+        % start
+        &  \TransitionVar_4 = \TransitionVar_5 + \TransitionVar_2 - 1 \\
+        % sawA
+        & \TransitionVar_1 = \TransitionVar_5 \\
+        % sawB
+        & \TransitionVar_2 = \TransitionVar_4 + \TransitionVar_6 \\
+        % final
+        & \TransitionVar_5 + \TransitionVar_6 = \FinalStateVar_1 \\
+        & \FinalStateVar_1 = 1 \\
+      \end{aligned}  
+    \end{equation*}    
+  \end{minipage}
+  \caption{Equivalences defining $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$
+  respectively.}\label{fig:example:single:equivalences}
+  \end{figure}
+
+Propagating the equivalences from \FlowEq{} gives us (under the implicit
+assumption that every RHS is $\geq 0$):
+\begin{equation*}
+  \begin{aligned}
+    & \Filter(\FromLabelTo{start}{a}{sawA}) & = 1 - \TransitionVar_6 \\
+    & \Filter(\FromLabelTo{start}{b}{sawB}) & = \TransitionVar_4 + \TransitionVar_6 \\
+    & \Filter(\FromLabelTo{sawA}{c}{sawA})  & = \TransitionVar_3  \\
+    & \Filter(\FromLabelTo{sawB}{b}{start}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
+    & \Filter(\FromLabelTo{sawA}{a}{final}) & = 1 - \TransitionVar_6 \\
+    & \Filter(\FromLabelTo{sawB}{c}{final}) & = \TransitionVar_6 \\
+  \end{aligned}
+\end{equation*}
+
+\begin{figure}
+  \centering
+\begin{prooftree}
+  \infer0[\PresburgerClose]{
+    \begin{matrix}
+      1 - \TransitionVar_6 = 0 \land \\
+      \TransitionVar_6 = 1 \land \\
+      \TransitionVar_4 + 1 > 0 \land \\
+      \TransitionVar_3 = 0 \land \\
+      2 + \TransitionVar_4 > 0 \land \\
+      k = 1 + \TransitionVar_4 
+    \end{matrix}
+  }
+  \infer1[\Subsume{}]{
+    \begin{matrix}
+      \Connected(\AcaOrBc{}, \Filter) \land \\
+      1 - \TransitionVar_6 = 0 \land \\
+      \TransitionVar_6 = 1 \land \\
+      \TransitionVar_4 + 1 > 0 \land \\
+      \TransitionVar_3 = 0 \land \\
+      2 + \TransitionVar_4 > 0 \land \\
+      k = 1 + \TransitionVar_4 
+    \end{matrix}
+  }
+  \infer1[\Propagate]{
+    \begin{matrix}
+      \Connected(\AcaOrBc{}, \Filter) \land \\
+      1 - \TransitionVar_6 = 0 \land \\
+      \TransitionVar_6 = 1 \land \\
+      \TransitionVar_4 + 1 > 0 \land \\
+      \TransitionVar_3 \geq 0 \land \\
+      2 + \TransitionVar_4 \geq 0 \land \\
+      2k = 2 + 2\TransitionVar_4 + \TransitionVar_3
+    \end{matrix}
+  }
+  \infer1[Eq. reason]{
+    \begin{matrix}
+      \Connected(\AcaOrBc{}, \Filter) \land \\
+      1 - \TransitionVar_6 = 0 \land \\
+      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
+    \end{matrix}
+  }
+  % BRANCH:  = 0
+  \infer0[\PresburgerClose{}]{
+    \begin{matrix}
+      \TransitionVar_3 > 0 \\
+      \TransitionVar_6 = 0 \land \\
+      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
+      1 - \TransitionVar_6 > 0 \land \\
+      2k = 1 + \TransitionVar_3
+    \end{matrix}
+  }
+  \infer1[\Subsume{}]{
+    \begin{matrix}
+      \Connected(\AcaOrBc{}, \Filter) \land \\
+      \TransitionVar_3 > = 0 \\
+      \TransitionVar_6 = 0 \land \\
+      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
+      1 - \TransitionVar_6 > 0 \land \\
+      2k = 1 + \TransitionVar_3
+    \end{matrix}
+  }
+  \infer1[Eq. reason]{
+    \begin{matrix}
+      \Connected(\AcaOrBc{}, \Filter) \land \\
+      \TransitionVar_6 = 0 \land \\
+      1 - \TransitionVar_6 > 0 \land \\
+      2\TransitionVar_6 + \TransitionVar_4 = 0\\
+      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
+    \end{matrix}
+  }
+  \infer1[\Propagate{}]{
+    \begin{matrix}
+      \Connected(\AcaOrBc{}, \Filter) \land \\
+      \TransitionVar_6 = 0 \land \\
+      1 - \TransitionVar_6 > 0 \land \\
+      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
+    \end{matrix}
+  }
+  \infer1[Eq. reason]{
+    \begin{matrix}
+      \Connected(\AcaOrBc{}, \Filter) \land \\
+      1 - \TransitionVar_6 > 0 \land \\
+      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
+    \end{matrix} % BRANCH: > 0
+  } % SPLIT
+  \infer2[\Split{} $1 - \TransitionVar_6$]{ \Connected(\AcaOrBc{}, \Filter) \land 2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6 }
+  \infer1[\Expand{}]{\Image{}_{\AcaOrBc{}, \Transition \mapsto 1}(\Filter, 2k + 1)}
+\end{prooftree}
+\caption{A derivation for \Calculus{} computing odd lengths in $\AcaOrBc{}$.}\label{fig:derivation:single}
+\end{figure}
+
+In \cref{fig:derivation:single}, we see how we start by expanding the predicate
+using the (simplified) flow equations. We can also see how interleaving
+reasoning on the corresponding linear equations helps the production. In both
+branches the reasoning is similar: we conclude that when choosing either path at
+the starting state we must avoid the other one, perform propagation based on
+that fact, use some equational reasoning to derive the expected forms of the
+upper and lower bounds to propagate the now disconnected transitions, and
+finally subsume, removing the $\Connected{}$ predicate when we are unable to use
+either of the rules, leaving only the satisfiable remnants of the flow
+equations.
 
@@
 
-\subsection{Finding the Presburger representation of a homomorphic image}
-\contents{
-  \item lazy qe
-}
+\subsection{Finding the Presburger representation of a homomorphic image}\label{sec:finding-the-image}
+
+To find the Presburger form of the homomorphic image efficiently, we adapt the
+quantifier elimination approach of~\cite{qe} to our problem domain. The core
+method is the same: we incrementally use \Calculus{} to find models,
+\Generalise{} the models we find into a quantifier-free Presburger formula with
+the $\MonoidElement$ as the only free variable (algorithm~\ref{alg:generalise}), add the
+negated formula as a constraint, and continue enumerating models until none
+remain. The disjunction of the generalised models we enumerated is now our
+image.
+
+\begin{algorithm}
+  \caption{$\Generalise{}(\Automaton, \Filter, a)$ will generalise a final product $\Automaton$ and a model $a$ under the homomorphism $\Map$.}\label{alg:generalise}
+  \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
+  \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
+  \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
+
+  $\Automaton' \gets \Tuple{\Automaton_\States, \Automaton_{\InitialState}, \Automaton_\AcceptingStates, \Set{\Transition \in \Automaton_\Transitions \SuchThat a(\Filter(\Transition)) \neq 0}}$
+
+  \Fudge{// This notation might be too much but IT'S OBVIOUS DAMMIT}
+
+  \KwRet{\EliminateQuantifiers{$\Map(\ParikhMap(\Automaton'))$}}
+  \end{algorithm}
+
+  \begin{algorithm}
+    \DontPrintSemicolon
+    \caption{$\FindImage{}(\Automaton_1 \times \ldots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
+    \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
+    \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
+    \SetKwFunction{NewTheoremProver}{newTheoremProver}
+    \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
+    \SetKwFunction{FreshVariable}{freshVariable}
+    \SetKwFunction{Assert}{assert}
+    \SetKwFunction{GetModel}{getModel}
+    \SetKwData{ImageVar}{image}
+
+$p \gets \NewTheoremProver{}$\;
+$\Filter(\Transition) := \FreshVariable{p}$ for every $\Transition \in \Transitions_{\Automaton}$\;
+$\MonoidElement \gets$ \FreshVariable{$p$}\;
+\Assert{$p, \exists \MonoidElement, \Filter(\Transition) \text{ for every } \Transition \in \Transitions_\Automaton \HoldsThat \ImagePredicate{\Automaton}{\Map}{\Filter}{\MonoidElement} \land \AndComp{\Transition \in \Transitions}{\Filter(\Transition) \geq 0}$}\;
+$\ImageVar \gets \bot$\;
+\While{$p$ has more models}{
+  $\Tuple{\Automaton, a} \gets \GetModel(p)$\;
+  $G \gets \Generalise{}(\Automaton, \Filter, a)$\;
+  $\ImageVar \gets G \lor \Image$\;
+  \Assert{$p, \lnot G$}\;
+  }
+    \KwRet{\ImageVar}
+    \end{algorithm}
+
+    The attentive reader will notice that the \GetModel{} function is
+    nonstandard in that it returns both the model and the automaton used to
+    generate it. Since we implement algorithm~\ref{alg:find-image} inside a theorem
+    prover under our control, we are able to glean such internal data structures.
+This should be considered fair game, since we are in practice implementing a specialised form of quantifier elimination, which is an internal affair.
+
+    The performance of this approach is evaluated in
+    \cref{sec:evaluation:finding-image}, where we benchmark our implementation
+    of it.
 
@@
       \inferrule
-      {\Set{\FlowEq(\Automaton_i), \Connected(\Automaton_i) \SuchThat \Automaton_i \in \Automaton_1,\ldots,\Automaton_k}, \ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
+      {
+        {\begin{matrix}
+          \Set{ 
+            \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
+          \SuchThat \Automaton_i \in \Automaton_1,\ldots,\Automaton_k}, \\
+          \Set{\FlowEq(\Automaton_i), \Connected(\Automaton_i) \SuchThat \Automaton_i \in \Automaton_1,\ldots,\Automaton_k}, \\
+          \ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \\
+          \SomeInequalities, \SomeClause
+        \end{matrix}}
+        }
       {\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
@@
 \subsection{An Example}
+
+We return again to our example in \cref{sec:introduction:parikh}, where we
+compute the whole Parikh image of the product of $\AcaOrBc{}$ and
+$\SomethingCSomething$ under the constraint that there are more instances of
+letters a than b. We omit the definition of the map $\Map$ that maps each
+transition to the corresponding increment vector (e.g.
+\Map($\FromLabelTo{start}{a}{sawA}) = \begin{bmatrix}1\\0\\0\end{bmatrix}$),
+since it should be apparent by now.
+
+In the interest of space, we use a similar technique as in
+\cref{fig:example:single:equivalences} to combine the application of $\FlowEq$
+with the definition of $\Filter$.
+
+  We define $\Filter$ after substitutions from both initial applications of
+  $\FlowEq$ as follows:
+
+% Propagated equations
+    \begin{equation*}
+      \begin{aligned}
+        & \Filter(\FromLabelTo{start}{a}{sawA}) & = 1 - \TransitionVar_6 \\
+        & \Filter(\FromLabelTo{start}{b}{sawB}) & = \TransitionVar_4 + \TransitionVar_6 \\
+        & \Filter(\FromLabelTo{sawA}{c}{sawA})  & = \TransitionVar_3  \\
+        & \Filter(\FromLabelTo{sawB}{b}{start}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
+        & \Filter(\FromLabelTo{sawA}{a}{final}) & = 1 - \TransitionVar_6 \\
+        & \Filter(\FromLabelTo{sawB}{c}{final}) & = \TransitionVar_6 \\
+        % Other automaton
+        & \Filter(\FromLabelTo{start}{\Sigma}{start}) & = \TransitionVar_7 \\
+        & \Filter(\FromLabelTo{start}{c}{final}) & = 1 \\
+        & \Filter(\FromLabelTo{final}{\Sigma}{final}) & = \TransitionVar_9 \\
+      \end{aligned}
+    \end{equation*}
+
+\begin{prooftree}
+  \infer0[]{
+    \begin{matrix}
+      \begin{bmatrix}
+        a \\
+        b \\
+        c
+        \end{bmatrix} = \begin{bmatrix}
+          2 - 2\TransitionVar_6 \\
+          \TransitionVar_4 + \TransitionVar_6 + 2\TransitionVar_6 + \TransitionVar_4 \\
+          \TransitionVar_3 + \TransitionVar_6
+        \end{bmatrix}
+        \\
+        \begin{bmatrix} % What to do about sigma!?
+          a \\
+          b \\
+          c
+          \end{bmatrix} = \\
+      \Connected(\AcaOrBc{}, \Filter) \land \\
+      \Connected(\SomethingCSomething{}, \Filter) \land \\
+      \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
+      \begin{bmatrix}
+        a \\
+        b \\
+        c
+        \end{bmatrix}) \land \\
+        a > b
+    \end{matrix}
+  }
+  \infer1[\ExpandM]{\Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, \begin{bmatrix}
+    a \\
+    b \\
+    c
+    \end{bmatrix}) \land a > b}
+\end{prooftree}
 
@@
 that some are partially or fully implemented in \Catra{} already. In particular,
-we have support for finite symbolic transitions over Unicode alphabets to keep
-practical automata under a reasonable size.
+we have some level of support for symbolic transitions over Unicode alphabets to
+keep practical automata under a reasonable size, though we do not allow a full Boolean algebra over symbols as described e.g. in~\cite{symbolic-automata}.
 
 \subsection{Backjumping and Learning No-Goods}\label{sec:ext:backjumping}
-\contents{
-  \item What do we have
-  \item A case where no-goods would be good
-  \item What to add to the calculus?
-}
-
-\subsubsection{An Example}
+
+\Calculus{} can be accelerated for some instances by adding rules for
+backjumping. In particular, central connectedness constraints for an automaton
+can be learnt. In that case, when discovering that a state $\State$ of an
+automaton $\Automaton$ under a certain transition variable $\Filter$ has become
+unreachable, we can learn the clause $\Connected(\Automaton, \Filter) \land
+\sum_{\Transition \in \SeparatingCut(\Automaton, \State)} \Filter(\Transition) =
+0 \implies \sum_{\Transition' = \FromLabelTo{\State}{}{} \in
+\Transitions_{\Automaton}} = 0$. A similar rule can be used when a state becomes
+backwards-unreachable (but with a corresponding backwards cut). At the moment,
+only forward-cut learning is implemented in \Catra, and has provided a slight
+improvement in performance on some instances.
+
+The other source of clauses to learn is the \Materialise{} rule. Whenever an
+attempt to materialise a product of two ($\Filter$-filtered) automata
+$\Automaton_1 \times \Automaton_2$ produces an empty product, we can determine
+the cause of the failure with respect to the automata and their respective
+transition variables. \Fudge{How? I have no idea}
+
+\contents{\item dela upp på flera predikat
+\item lär no-good kombinationer}
+
 \subsection{Symbolic Automata}\label{sec:ext:symbolic}
 
-\contents{
-  \item Should we also write about how to adapt to infinite alphabets?
-}
+Extending \Calculus{} to support fully symbolic \Fudge{automata is straigthforward}. The
+only difficulty consists in handling the mapping of the homomorphism $\Map$ over
+symbolic labels, assuming it maps to finitely many monoid elements.
+
+\Fudge{We should probably expand here? I don't know}
 
@@
 
-\contents{\item What actually goes here!?}
+\contents{\item I have no idea what to write here
+\item multiple tracks}
 
@@
 We implement a calculus for Parikh automata as described in
-Section~\ref{sec:parikh-automata}. The artefact submitted along with this paper
+\cref{sec:parikh-automata}. The artefact submitted along with this paper
 is a program that reads one or more products of one or more DFA, the register
@@
 constraints on the final values of their registers expressed as Presburger
-arithmetic. We call this program \Catra{}.\footnote{If you really must read it
+arithmetic. We call this program \Catra.\footnote{If you really must read it
 as an acronym, please read it as CAtegory Theory on Register Automata, or if you
@@
 assignments or unsatisfiable much like traditional~SAT- or SMT solvers would.
-Additionally, the baseline and our own back-end also support generating
-the Presburger formula describing the constraints of the input file \Fudge{using
-the method described in Section~\ref{sec:presburger-representation}}.
-
-Note that \Catra{} can solve equations of Parikh images. An equation like
+Additionally, the baseline and our own back-end also support efficient
+generation of the Presburger formula describing the constraints of the input
+file using the method described in \cref{sec:finding-the-image}.
+
+\Catra{} can solve equations of Parikh images. An equation like
 $\ParikhMap(\Automaton_1 \times \Automaton_2) = \ParikhMap(\Automaton_3 \times
@@
 
-\contents{
-  \item the unused predicate
-  \item single-arity
-  \item a princess theory
-}
+We implement \Calculus{} as a theory plug-in for the \Princess{} theorem prover.
+Since \Princess{} does not support multiple-arity predicates like the ones we
+use in \Calculus{}, we have implemented variable-length arguments using
+additional helper predicates. These are $\Unused{}(\Automaton)$, which marks an
+automaton as unused in any product, and $\TransitionMask{}(\Automaton,
+\Transition, \Filter(\Transition))$ which associates a transition $\Transition$
+and automaton $\Automaton$ with its corresponding transition variable.
+Additionally, we associate each of our predicates with an instance variable in
+order to differentiate instances of the predicates.
 
@@
 fashion we compute the product incrementally term by term, checking
-satisfiability in each step. We use \Fudge{a priority queue} to select which
-term to use for the next step of the product, and order it by \Fudge{the number
-of transitions} as a heuristic for the size of the term. We use this heuristic
+satisfiability in each step. We use a priority queue to select which
+term to use for the next step of the product, and order it by the number
+of transitions as a heuristic for the size of the automaton. We use this heuristic
 to avoid computing large (and therefore slow) products until we have to, banking
-our hopes on finding a source of unsatisfiability earlier. The pseudocode for
+our hopes on finding a source of unsatisfiability early. The pseudocode for
 our implementation can be seen in Algorithm~\ref{alg:baseline}.
@@
 
-Since \Catra{} does not learn very sophisticated clauses yet (such as the ones
-described in Section~\ref{sec:ext:backjumping}), it is turned off by default.
-However, even in its underdeveloped state, clause learning has shown to
-significantly improve the performance on unsatisfiable instances, at the cost of
-a loss in performance for satisfiable instances. Since \Catra{} already
-struggles more with satisfiable instances than unsatisfiable ones, we opted to
-turn the option off by default.
+\Catra{} enables clause learning by default when using our backend, as it has
+been experimentally shown to increase the performance in aggregate (though not
+strictly). We do not currently implement all the proposed features of
+\cref{sec:ext:backjumping}, but we do implement forward-reachability cut
+learning (which had a very modest improvement in performance). No sophisticated
+clause learning for products has been implemented.
 
@@
 
-\subsubsection{Finding a Presburger Formula}
+\subsubsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
 
@@
 of the artefact in the interest of transparency. Moreover, we have validated
-\Fudge{all reported satisfiable instances and their witnesses} with \Nuxmv to
+\Fudge{all reported satisfiable instances and their witnesses} with \Nuxmv{} to
 ensure that \Calculus{} is indeed sound and does not cheat.
@@
 implementations. For \Nuxmv{} we use the default configuration which we believe
-should be performant (or it should not be the default). Additionally, tweaking our invocation of \Nuxmv is explicitly made easy for artefact reviewers.
+should be performant (or it should not be the default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made easy for artefact reviewers.
 

### 2022-07-01 bin/verma-to-tex.py

@@ -3,23 +3,23 @@
 from re import A
 
 
-initial_state = "start"
-accepting_states = {"final"}
+initial_state = 0
+accepting_states = {3}
 transitions = [
-    ("start", "a", "sawA"),
-    ("start", "b", "sawB"),
-    ("sawA", "c", "sawA"),
-    ("sawA", "a", "final"),
-    ("sawB", "b", "start"),
-    ("sawB", "c", "final"),
+    (0, "a", 1),
+    (0, "b", 2),
+    (1, "c", 1),
+    (1, "a", 3),
+    (2, "b", 0),
+    (2, "c", 3),
 ]
 
-states = {
+states = sorted({
     initial_state,
     *accepting_states,
     *{s for s, _, _ in transitions},
     *{s for _, _, s in transitions},
-}
+})
 
 
 def chunks(l, n):

### 2022-07-05 main.tex
Update data, extensions

@@
   removal from~$\Transitions$ would cause~$\State \in \States$ to be unreachable
-  from the initial state~$\InitialState$. $\SeparatingCut(\Automaton,
-  \InitialState)$ is crucially undefined since it cannot be made unreachable
-  from itself by definition.
+  from the initial state~$\InitialState$, or cut it off from reaching any
+  accepting state. $\SeparatingCut(\Automaton, \InitialState)$ is crucially
+  undefined if $\Automaton$ accepts any string since it cannot be made unreachable
+  from itself by definition and must have at least onepath to an accepting state
+  if it accepts any string.
 \end{definition}
@@
 
-\subsection{An Example}
+\subsection{An Example}\label{sec:multiple:example}
 
@@
   We define $\Filter$ after substitutions from both initial applications of
-  $\FlowEq$ as follows:
+  $\FlowEq$ as follows, note the element-wise expansion for the $\Sigma$ labels:
 
@@
         % Other automaton
-        & \Filter(\FromLabelTo{start}{\Sigma}{start}) & = \TransitionVar_7 \\
+        & \Filter(\FromLabelTo{start}{\Sigma}{start}) & = \TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c} \\
         & \Filter(\FromLabelTo{start}{c}{final}) & = 1 \\
-        & \Filter(\FromLabelTo{final}{\Sigma}{final}) & = \TransitionVar_9 \\
+        & \Filter(\FromLabelTo{final}{\Sigma}{final}) & = \TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c} \\
       \end{aligned}
@@
 \begin{prooftree}
-  \infer0[]{
+  \infer0[\PresburgerClose{}]{
+    \begin{matrix}
+      1 = \TransitionVar_{10} \\
+      \TransitionVar_3 = \TransitionVar_{11} \\
+      1 = \TransitionVar_{12} \\
+      1 = \TransitionVar_{11} \\
+      2 > 0
+    \end{matrix}  
+  }
+  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
+    \begin{matrix}
+      1 = \TransitionVar_{10} \\
+      \TransitionVar_3 = \TransitionVar_{11} \\
+      1 = \TransitionVar_{12} \\
+      1 = \TransitionVar_{11} \\
+      \Image{}_{\Automaton', \Map}(\Filter, 
+      \begin{bmatrix}
+        2 \\
+        0 \\
+        1
+        \end{bmatrix}) \land \\
+      \Connected(\SomethingCSomething{}, \Filter) \land \\
+        2 > 0
+    \end{matrix}  
+  }
+  \infer1[Eq. reasoning]{
+    \begin{matrix}
+      1 = \TransitionVar_{10} \\
+      \TransitionVar_3 = \TransitionVar_{11} \\
+      1 = \TransitionVar_{12} \\
+      1 = \TransitionVar_{11} \\
+      \Image{}_{\Automaton', \Map}(\Filter, 
+      \begin{bmatrix}
+        a \\
+        0 \\
+        c
+        \end{bmatrix}) \land \\
+      \TransitionVar_6 = 0 \land \\
+      \TransitionVar_4 = 0 \land \\
+      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
+      \begin{bmatrix}
+        a \\
+        b \\
+        c
+        \end{bmatrix} = \begin{bmatrix}
+          2  \\
+          0 \\
+          1
+        \end{bmatrix} \land
+        \\
+        \begin{bmatrix}
+          a \\
+          b \\
+          c
+          \end{bmatrix} = \begin{bmatrix}
+            \TransitionVar_{7a} + \TransitionVar_{9a} \\
+            0 \\
+            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
+          \end{bmatrix} \land \\
+      \Connected(\SomethingCSomething{}, \Filter) \land \\
+        a > 0
+    \end{matrix}  
+  }
+\end{prooftree}
+
+\begin{prooftree}
+  \infer0[\Materialise]{
+    \begin{matrix}
+      \TransitionVar_6 = 0 \\
+      \TransitionVar_4 = 0 \\
+      2\TransitionVar_6 + \TransitionVar_4 = 0 \\
+      \begin{bmatrix}
+        a \\
+        b \\
+        c
+        \end{bmatrix} = \begin{bmatrix}
+          2  \\
+          0 \\
+          \TransitionVar_3 
+        \end{bmatrix}
+        \\
+        \begin{bmatrix}
+          a \\
+          b \\
+          c
+          \end{bmatrix} = \begin{bmatrix}
+            \TransitionVar_{7a} + \TransitionVar_{9a} \\
+            0 \\
+            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
+          \end{bmatrix} \\
+      \Connected(\SomethingCSomething{}, \Filter) \land \\
+      \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
+      \begin{bmatrix}
+        a \\
+        0 \\
+        c
+        \end{bmatrix}) \land \\
+        a > 0
+    \end{matrix}  
+  }
+  \infer1[\Subsume{}]{
+  \begin{matrix}
+    \TransitionVar_6 + \TransitionVar_4 = 0 \\
+    1 - \TransitionVar_6 > 0 \\
+    \TransitionVar_6 = 0 \\
+    \TransitionVar_4 = 0 \\
+    \TransitionVar_3 > 0 \\
+    2\TransitionVar_6 + \TransitionVar_4 = 0 \\
+    \begin{bmatrix}
+      a \\
+      b \\
+      c
+      \end{bmatrix} = \begin{bmatrix}
+        2  \\
+        0 \\
+        \TransitionVar_3 
+      \end{bmatrix}
+      \\
+      \begin{bmatrix} % What to do about sigma!?
+        a \\
+        b \\
+        c
+        \end{bmatrix} = \begin{bmatrix}
+          \TransitionVar_{7a} + \TransitionVar_{9a} \\
+          0 \\
+          \TransitionVar_{7c} + \TransitionVar_{9c} + 1
+        \end{bmatrix} \\
+    \Connected(\AcaOrBc{}, \Filter) \land \\
+    \Connected(\SomethingCSomething{}, \Filter) \land \\
+    \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
+    \begin{bmatrix}
+      a \\
+      0 \\
+      c
+      \end{bmatrix}) \land \\
+      a > 0
+  \end{matrix}
+  }
+  \infer1[Eq. reasoning]{
     \begin{matrix}
@@
           2 - 2\TransitionVar_6 \\
-          \TransitionVar_4 + \TransitionVar_6 + 2\TransitionVar_6 + \TransitionVar_4 \\
+          2\TransitionVar_4 + 3\TransitionVar_6 \\
           \TransitionVar_3 + \TransitionVar_6
@@
           c
-          \end{bmatrix} = \\
+          \end{bmatrix} = \begin{bmatrix}
+            \TransitionVar_{7a} + \TransitionVar_{9a} \\
+            \TransitionVar_{7b} + \TransitionVar_{9b} \\
+            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
+          \end{bmatrix} \\
       \Connected(\AcaOrBc{}, \Filter) \land \\
@@
 the cause of the failure with respect to the automata and their respective
-transition variables. \Fudge{How? I have no idea}
-
-\contents{\item dela upp på flera predikat
-\item lär no-good kombinationer}
+transition variables, and learn no-good combinations that can never be part of a
+model. In order to be able to do this, we need a few semantic predicates to
+record the state of the calculation, as well as a system of disambiguating
+automata, since it is possible to arrive at the same automaton by multiple
+combinations of decisions and products. In \Catra{}, we use a number of
+additional predicates to record the status of the materialisation of products,
+to separate instances of our main predicate, and to register the mapping between
+automata and their respective transition $\Filter$ terms, described in
+\cref{sec:implementation}. However, at this point only rudimentary no-good
+learning is implemented.
 
@@
 
-Extending \Calculus{} to support fully symbolic \Fudge{automata is straigthforward}. The
-only difficulty consists in handling the mapping of the homomorphism $\Map$ over
-symbolic labels, assuming it maps to finitely many monoid elements.
-
-\Fudge{We should probably expand here? I don't know}
+Extending \Calculus{} to support fully symbolic automata is possible within the
+framework, depending on your choice of $\Map$. The difficulty consists in
+handling the mapping of the homomorphism $\Map$ over symbolic labels, assuming
+it maps to finitely many monoid elements. This is not always straightforward, as
+seen in the example of \cref{sec:multiple:example}. This complexity is inherent
+in computing the full Parikh image, and stems from the fact that we need to
+differentiate between the possible interpretations of the $\Sigma$ transitions
+without knowing ahead of times which ones will actually be materialied in the
+product. If, on the other hand, if our $\Map$ had been length-counting, which
+does not differentiate between values, it would have required no adaptation at
+all. With a somewhat liberal interpretation of what we are allowed to map to
+(e.g. fresh terms), it is possible to use $\Map$ to represent choice operations
+like the ones in a range label. In \Catra{}, we frontload this problem by
+requiring the user to encode their input as a Parikh automaton. \Fudge{How is
+the encoding of Ostrich Plus automata done?}
 
 \subsection{Transducers}\label{sec:ext:transducers}
+
+Another interesting application is applying \Calculus{} to automata with
+multiple tracks, e.g. transducers. One application of such a calculus would be
+to represent replace operations and other functions on regular languages, and to
+be able to answer questions like "does this operation change the length of the
+string". The difficulty in implementing it for transducers is first to perform
+the mapping on the labels, which for both the length and Parikh cases is
+straightforward; just do the same thing in two dimensions. The more complicated
+operation is defining the product of transducers. In some cases it would
+probably be desirable to perform synchronisation (e.g. requiring overlapping
+transitions) on only some transitions, for example the first track.
+
+Implementing such a calculus is straightforward in \Calculus, since the
+definition of products was left out of the definition. All that would be needed
+is an appropriate update of the definitions of products. Simiarly, \Catra was
+written with modularity in mind, and it should be straightforward to extend both
+the input grammar and automata implementations to accomodate multi-track
+automata with arbitrary synchronisation.
 
@@
 
-The benchmarks are executed in parallel, since they are mostly single-threaded and initial results showed negligible interference on performance, on \BenchmarkRig{}.
-We compiled the code using Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a maximum heap of~\MaxHeapSize{}.
-We used \Nuxmv{} version~\NuxmvVersion{} invoked as a subprocess for each instance.
-Experiments were executed on a fresh JVM per backend, but the JVM was kept active for all benchmarks, meaning that JIT compilation would make subsequent executions faster.
-We believe this represents a realistic use case where \Calculus{} is integrated into a string solver or other system and be called repeatedly.
-To mitigate systemic impacts of this decision on our results, experiments were executed in random order for all backends.
-We run all  with a timeout of~\RuntimeTimeout{}.
-
-As a way of gauging the differences in overhead between the native backends and \Nuxmv, we executed a small experiment where we ran the same trivial instance 30 times and observed the runtimes for all solvers.
-All backends saw an improvement from JIT compilation within the first three runs, but subsequent runs only improved the native backends.
-The final measured overhead for the fully warm JVMs was less than~\OverheadSeconds.
-From this we draw the conclusion that major differences in runtime between the native backends and \Nuxmv{} is more likely to stem from innate differences rather than unfair overhead resulting from our particular configuration.
+The benchmarks are executed in parallel using GNU Parallel~\cite{parallel},
+since they are mostly single-threaded and initial results showed negligible
+interference on performance, on \BenchmarkRig{}. We compiled the code using
+Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a
+maximum heap of~\MaxHeapSize{}. We used \Nuxmv{} version~\NuxmvVersion{} invoked
+as a subprocess for each instance. Experiments were executed on a fresh JVM per
+backend, but the JVM was kept active for each batch of \BatchSize{} instances,
+meaning that JIT compilation would make subsequent executions faster, while
+still periodically restarting the JVM. We believe this represents a realistic
+use case where \Calculus{} is integrated into a string solver or other system
+and is called repeatedly. To mitigate systemic impacts of this decision on our
+results, experiments were executed in random order for all backends. We run all
+benchmarks with a timeout of~\RuntimeTimeout{}.
+
+As a way of gauging the differences in overhead between the native backends and
+\Nuxmv, we executed a small experiment where we ran the same trivial instance 30
+times and observed the runtimes for all solvers. All backends saw an improvement
+from JIT compilation within the first three runs, but subsequent runs only
+improved the native backends. The final measured overhead for the fully warm
+JVMs was less than~\OverheadSeconds. From this we draw the conclusion that major
+differences in runtime between the native backends and \Nuxmv{} is more likely
+to stem from innate differences rather than unfair overhead resulting from our
+particular configuration. Additionally, one of the advantages of \Calculus{} to
+\Nuxmv{} is that the former can me integrated into an automated theorem prover
+while the latter has to be called as a separate process. In this sense, our
+benchmarking setup can be said to be realistic.
 
@@
 
-It should be noted that despite running in the standard, deterministic configuration, we observed some degree of nondeterminism in \Nuxmv.
-Instances that were solved or unsolved would sometimes be solved or timed out in subsequent runs.
-This observation was made even without randomised execution order, suggesting that the nondeterminism likely stemmed either from our encoding of the instances into \Nuxmv's format, or the execution environment.
-Similarly, \Calculus{} showed similar signs of nondeterminism, but only for satisfiable instances.
-To address this, we increased the runtime from our initial experiments to give us a wider marigin.
-By contrast, the baseline implementation showed no signs of nondeterminism.
-It should be noted that for both backends exhibiting nondeterminism, the nondeterminism did not change the general trend of the experiments, even for experiments with few instances.
-Therefore, we believe that our large set of benchmarks will protect us from bias since any backend is unlikely to be neither consistently lucky nor consistently unlucky.
+It should be noted that despite running in the standard, deterministic
+configuration, we observed some degree of nondeterminism in \Nuxmv. Instances
+that were solved or unsolved would sometimes be solved or timed out in
+subsequent runs. This observation was made even without randomised execution
+order, suggesting that the nondeterminism likely stemmed either from our
+encoding of the instances into \Nuxmv's format, or the execution environment.
+Similarly, \Calculus{} showed similar signs of nondeterminism, but only for
+satisfiable instances. To address this, we increased the runtime from our
+initial experiments to give us a wider marigin. By contrast, the baseline
+implementation showed no signs of nondeterminism. It should be noted that for
+both backends exhibiting nondeterminism, the nondeterminism did not change the
+general trend of the experiments, even for experiments with few instances.
+Therefore, we believe that our large set of benchmarks will protect us from bias
+since any backend is unlikely to be neither consistently lucky nor consistently
+unlucky.
 
@@
 
-In Figures~\ref{fig:solve-division} we show how many of the~\NrBenchmarks{} the respective back-ends could solve and with which status.
-A full summary of their outcomes is also available in Table~\ref{tab:solve-status}.
-We see here \Fudge{that some solver is better than the others}.
+In Figures~\ref{fig:solve-division} we show how many of the~\NrBenchmarks{} the
+respective back-ends could solve and with which status. A full summary of their
+outcomes is also available in Table~\ref{tab:solve-status}. We see here that
+\Calculus{} generally outperforms \Nuxmv{} on unsatisfiable instances, while
+being slightly worse at satisfiable instances. Both \Nuxmv{} and \Calculus{}
+massively and strictly outperform the baseline approach on every kind of
+instance, but most of all on satisfiable instances. In fact, the baseline
+approach was significantly harder to benchmark due to its much higher demand for
+RAM.
 

### 2022-07-05 main.tex
Intro

@@
 \usepackage{ebproof}
-\usepackage{stackengine}
 
@@
 \begin{abstract}
-  The image of the Parikh map is important in automata theory, offering a
-  compact characterisation of an automaton. We contribute a novel understanding
-  of how Parikh maps can be combined with arbitrary commutative monoid morphisms
-  to efficiently represent real-world constraints found in String solvers.
-  Furthermore, we show how this formulation can be efficiently implemented as a
-  calculus that we call \Calculus{} in a theorem prover supporting Presburger
-  logic. In particular, our calculus allows us to solve the common problem in
-  string solvers of computing the Parikh image of products of arbitrarily many
-  automata. We demonstrate this by implementing a tool called \Catra{}, which we
-  evaluate on constraints produced by \Fudge{the PyEx benchmarks} when solved by
-  the \Ostrich{} string constraint solver. We show that \Calculus{} strictly
-  outperforms the standard eager approach described
-  in~\citeauthor{generate-parikh-image} as well as the improved
+  A common problem in string constraint solvers is computing the Parikh image, a
+   set of linear equations that describe all possible combinations of character
+   counts in strings of a given language. Automata-based string solvers
+   frequently need to compute the Parikh image of large products (intersections)
+   of automata, which in many operations is both prohibitively slow and
+   memory-intensive. We contribute a novel understanding of how Parikh maps can
+   be combined with arbitrary commutative monoid morphisms to represent
+   real-world constraints on regular languages, most notably the length
+   constraint. Furthermore, we show how this formulation can be efficiently
+   implemented as a calculus, \Calculus{}, in an automated theorem prover
+   supporting Presburger logic. We also provide an efficient method for deriving
+   the Presburger formula of a given Parikh image, \Fudge{which is good for
+   something} based on techniques for quantifier elimination. We implement
+   \Calculus{} in a tool called \Catra{}, and evaluate it on constraints
+   produced by the \Ostrich{} string constraint solver when solving
+   \Fudge{standard string constraint benchmarks involving string lengths}. We
+   show that our solution strictly outperforms the standard eager approach
+   described in~\citeauthor{generate-parikh-image} as well as the
    over-approximating method recently described
-  by~\citeauthor{approximate-parikh}. \Calculus{} typically also outperforms
-  the~\Nuxmv{} model checker, in particularly on unsatisfiable instances.
-  Finally, \Fudge{\Calculus{} and therefore \Catra{} provides an efficient
-  method for deriving the Presburger formula of a given Parikh image that
-  outperforms any previously published method.}
+   by~\citeauthor{approximate-parikh}, and in many cases also the~\Nuxmv{} model
+   checker.
 \end{abstract}
@@
 \item A specialised method for quantifier elimination that blends well with \Calculus{} to efficiently compute quantifier-free Presburger representations of a given Parikh image.
+\Fudge{Why is this good?}
 \item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future study.
@@
 
-\subsection{A demonstration of our approach}\label{sec:introduction:motivation}
+\section{A demonstration of our approach}\label{sec:motivation}
 
@@
 constraint for which we want to find values in $f$'s image over the intersection
-of the two languages. We will use the same two regular languages, recognised by
-two automata; $\SomethingCSomething{}$ (Figure~\ref{fig:something-c-something})
-and $\AcaOrBc{}$ (Figure~\ref{fig:aca-or-bc}) for both examples but change the
-constraint and homomorphism $f$.
-
-\subsubsection{Counting string lengths}
+of the two languages. We will use the same two regular languages,
+$\mathtt{ac^*a|b(bb)^*c}$ and $\mathtt{.^*c.^*}$ for both examples but change
+the constraint $\varphi$ and homomorphism $f$. In essence, we solve the
+following set of constraints for all possible values of $s$ and
+$\MonoidElement$:
+
+\begin{equation*}
+  \AndComp{}{\begin{aligned}
+  & \mathtt{ac^*a|b(bb)^*c} \text{ accepts } s\\
+  & \mathtt{.^*c.^*} \text{ accepts } s \\
+  & \MonoidElement = f(s) \\
+  & \varphi(\MonoidElement)
+  \end{aligned}}
+ \end{equation*}
+
+We will represent the regular languages as the automata $\SomethingCSomething{}$
+ (\cref{fig:something-c-something}) and $\AcaOrBc{}$ (\cref{fig:aca-or-bc})
+ respectively.
+
+\subsection{Counting string lengths}
 
@@
 
-The na\"ive approach would be to first compute the product of the automata and
-then compute the length of its product. However, we can take advantage of the
-both constraints and definitions of $f$ to prune the automata before computing
-the product and help reduce the blow-up. To give an intuition for the core idea
-of \Calculus{}, we will annotate the transitions of $\SomethingCSomething$ and
-$\AcaOrBc$ with variables illustrating how many times they are taken, and
-perform informal reasoning on them corresponding to the reasoning performed by
-our calculus.
+The na\"ive approach would be to first compute the product of the automata, then
+use the standard approach to find its image under $f$, and then verify the
+constraints. However, we can take advantage of the both constraints and
+definitions of $f$ to prune the automata before computing the product and help
+reduce the blow-up. To give an intuition for the core idea of \Calculus{}, we
+will annotate the transitions of $\SomethingCSomething$ and $\AcaOrBc$ with
+variables illustrating how many times they are taken, and perform informal
+reasoning on them corresponding to the reasoning performed by our calculus.
 
@@
 l_c' = 1$). Strictly speaking, this is where the constant number $1$ in
-Figure~\ref{fig:something-c-something} comes: from $1 + r_\Sigma = x + r_\Sigma$
+\cref{fig:something-c-something} comes: from $1 + r_\Sigma = x + r_\Sigma$
 solved for $x$. Using analogous linear equation reasoning loses additional
-variables and gives us the annotations in Figure~\ref{fig:example-simplify-1}.
+variables and gives us the annotations in \cref{fig:example-simplify-1}.
 
@@
 
-\subsubsection{Producing the Parikh image}\label{sec:introduction:parikh}
+\subsection{Producing the Parikh image}\label{sec:introduction:parikh}
 
@@
 written with modularity in mind, and it should be straightforward to extend both
-the input grammar and automata implementations to accomodate multi-track
+the input grammar and automata implementations to accommodate multi-track
 automata with arbitrary synchronisation.
-
-\contents{\item I have no idea what to write here
-\item multiple tracks}
 

### 2022-07-05 main.tex
More text

@@
                                         %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-
 %% Author information
@@
 its character counts, famously a semilinear set~\cite{parikh-theorem}. It
-appears naturally as part of many operations in model checking and the solving
-of string constraints in automata-based string solvers such as \Ostrich{},
-notably in representing constraints on string lengths.
+appears naturally as part of \Fudge{many operations} in model checking and
+solving string constraints in automata-based solvers such as \Ostrich{}
+\cite{ostrich}, notably in representing constraints on string lengths.
+Specialised string solvers like e.g. Trau \cite{trau} also perform extensive
+reasoning using Parikh images, but do so on restricted automata that do not
+suffer from the blow-up we seek to address, though it might be possible to
+adapts aspects of our approach to theirs.
 
@@
 In addition to the elimination of existentially quantified variables, another
-problem arises for the common case of solving length constraints in string
-solvers. Conjunctions of string constraints lead to the computation of Parikh
-images of products (intersections) of regular languages represented as finite
-automata. Applying the approach in~\cite{generate-parikh-image} would require
-the eager computation of the product before its Parikh image. In several
-instances we have observed while solving real-world string constraints, the
-computation of the product of automata exhausts the memory of any machine
-available due to the exponential blow-up in size of the product, quickly
-becoming intractable as the number of automata in the product increases.
+problem arises from length constraints in string solvers. Conjunctions of string
+constraints lead to the computation of length images of products (intersections)
+of regular languages represented as finite automata. Applying the approach
+in~\cite{generate-parikh-image} would require the eager computation of the
+product before its length image. In several instances we have observed while
+solving real-world string constraints, the computation of the product of
+automata exhausts the memory of any machine available due to the exponential
+blow-up in size of the product, quickly becoming intractable as the number of
+automata in the product increases. The current best published mitigation for
+this problem is an over-approximation that works by approximating the Parikh
+image of a product of automata to be the conjunction of the image of the
+individual automata of the product \cite{approximate-parikh}. This approach only
+works for unsatisfiable instances, and comes with a harsh penalty on satisfiable
+instances.
 
@@
 product's Parikh image, and is generalised to an arbitrary Parikh-like
-homomorphism over automata labels. This enables us to let both calculations
-inform each other, eliminating unnecessary work, and pruning the size of the
-partial products considered in the computation for a smaller memory footprint.
-Moreover, the method can be used iteratively to tackle smaller chunks of the
-product incrementally, thereby avoiding running out of memory at all. We
-implement this calculus, along with the approximate method
-of~\cite{approximate-parikh}, its fall-back variant adapted
+homomorphism over automata labels, including string lengths. This enables us to
+let both calculations inform each other, eliminating unnecessary work, and
+pruning the size of the partial products considered in the computation for a
+smaller memory footprint. Moreover, the method can be used iteratively to tackle
+smaller chunks of the product incrementally, thereby avoiding running out of
+memory at all.
+
+The problem of computing constraints on Parikh images over products of regular
+languages under a given commutative homomorphism amounts to solving products of
+Parikh automata. Parikh automata are regular automata extended with integer
+counters with given increments and decrements for each transition, where we
+allow checking a set of linear constraints on the final values of the counters
+(but not their intermittent values) \cite{parikh-automata}. Parikh automata
+without constraints on the final values on their registers are also sometimes
+called cost-enriched automata, weighted automata or counter automata, depending
+on exact definitions and side-constraints. The decision problem tackled in this
+paper, determining the emptiness of an intersection of Parikh automata, was
+recently shown to be PSPACE-complete~\cite{graph-queries}.
+
+In addition to the decision problem of determining whether an intersection of
+Parikh automata is empty and the satisfaction problem of finding a value from
+the intersection if it is not, \Catra{} is able to efficiently generate a
+quantifier-free Presburger formula representing an intersection of Parikh
+automata, something beyond the reach of general model checkers like~\Nuxmv.
+\Fudge{This is useful because the moon is made of cheese.}
+
+We implement \Calculus{} as a theory for the \Princess{} automated theorem
+prover in a tool that we call \Catra{}. \Catra also supports using the
+approximate method of \cite{approximate-parikh}, its fall-back variant adapted
 from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
-checker, in a tool that we call \Catra{}. \Catra{} allows solving constraints
-over groups of synchronised (products of) Parikh automata, that is
-(finite-alphabet) symbolic automata with transitions that increment and
-decrement integer registers and that allows checking linear constraints on these
-registers after running the automata. Parikh automata without constraints on the
-final values on their registers are also sometimes called cost-enriched
-automata, weighted automata or counter automata, depending on your definitions
-and side-constraints. In the satisfaction mode, we are more formally solving the
-intersection problem of Parikh automata (under homomorphism), known to be
-PSPACE-complete~\cite{graph-queries}.
-
-Additionally, \Catra{} is able to efficiently generate a quantifier-free
-Presburger formula representing such an instance of constraints and automata,
-something beyond the reach of general model checkers like~\Nuxmv.
+checker~\cite{nuxmv}. Using \Catra{}, we compare \Calculus{} to the other two
+back-ends on \NrBenchmarks{} Parikh automata satisfaction problems generated
+\Fudge{by Ostrich+}.
 
@@
 
-To illustrate the key features of our approach, we start with two regular
+To illustrate the key features of \Calculus, we start with two regular
 languages, a commutative homomorphism $f$ on them that we care about, and a side
 constraint for which we want to find values in $f$'s image over the intersection
-of the two languages. We will use the same two regular languages,
-$\mathtt{ac^*a|b(bb)^*c}$ and $\mathtt{.^*c.^*}$ for both examples but change
-the constraint $\varphi$ and homomorphism $f$. In essence, we solve the
-following set of constraints for all possible values of $s$ and
-$\MonoidElement$:
+of the two languages using a heavily elided version of \Calculus. We will use
+the same two regular languages, $\mathtt{ac^*a|b(bb)^*c}$ and $\mathtt{.^*c.^*}$
+for both examples but change the constraint $\varphi$ and homomorphism $f$. In
+essence, we solve the following set of constraints for all possible values of
+$s$ and $\MonoidElement$:
 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \begin{minipage}[b]{0.75\linewidth}
@@
 The na\"ive approach would be to first compute the product of the automata, then
-use the standard approach to find its image under $f$, and then verify the
-constraints. However, we can take advantage of the both constraints and
-definitions of $f$ to prune the automata before computing the product and help
-reduce the blow-up. To give an intuition for the core idea of \Calculus{}, we
-will annotate the transitions of $\SomethingCSomething$ and $\AcaOrBc$ with
-variables illustrating how many times they are taken, and perform informal
-reasoning on them corresponding to the reasoning performed by our calculus.
-
-\begin{figure}[h]
+use the approach from \cite{generate-parikh-image} to find its image under $f$,
+and then verify the constraints. However, we can take advantage of both the
+constraints and definitions of $f$ to prune the automata before computing the
+product and help reduce the blow-up.
+
+To give an intuition for the core idea of \Calculus{}, we will annotate the
+transitions of $\SomethingCSomething$ and $\AcaOrBc$ with variables illustrating
+how many times they are taken, and perform informal reasoning on them
+corresponding to the reasoning performed by our calculus.
+
+\begin{figure}[ht]
   \centering 
@@
 
-Note that we have already performed some reasoning in the starting figures -- we
+Note that we have already performed some reasoning in the starting figures; we
 use the same symbol for the number of times each of the upper transitions of
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering 
@@
 
-  \begin{figure}[h]
+  \begin{figure}[ht]
     \centering 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 
-  \begin{figure}[h]
+  \begin{figure}[ht]
 \centering
@@
 \end{figure}
-
-\contents{
-  \item Formally analyse the complexity of how much easier the length problem is
-}
 
@@
 
-\section{Implementation and Experiments}\label{sec:implementation}
+\section{Implementation}\label{sec:implementation}
 
@@
 from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
-symbolic model checker~\cite{nuxmv} to solve our constraints, and \Fudge{the
+symbolic model checker~\cite{nuxmv} to solve our constraints, and the
 approximation described in~\cite{approximate-parikh} on top of the standard
-baseline back-end}. An example of an input file \Fudge{illustrating every feature
-of the file format} can be seen in \cref{lst:input-example}.
+baseline back-end. An example of an input file corresponding to our running
+example introduced in \cref{sec:motivation} can be seen in
+\cref{lst:input-example}.
 
@@
 \begin{algorithm}
-  \caption{An algorithm with caption}\label{alg:baseline}
+  \caption{How we implement the baseline approach}\label{alg:baseline}
   \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
@@
 
-\subsection{\Catra{} is Versatile: Encoding Epistemic Logic Problems As Parikh Automata}
-
-\contents{
-\item Model-checking examples e.g. \cite{epistemic-logic}
-}
-
-\subsection{\Catra{} is Faster Than The Competition: Benchmarks and Their Results}\label{sec:experiments}
-
-We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances generated by the \Ostrich{} string constraint solver when solving the \Fudge{pyex benchmarks}.
-\Fudge{The instances are all enormous and have some on average 900 products of 400000 automata with 97 counters each and 9999 constraints}.
-After generating an initial~\InitialNrBenchmarks{}, we remove~\NrBroken{} misgenerated instances that did not parse, as well as~\NrTrivial{} instances solved in under five seconds by the baseline backend.
-The benchmarks are run on commit~\texttt{\commit}.
+\section{Evaluation}\label{sec:experiments}
+
+We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances generated
+by the \Ostrich{} string constraint solver when solving the \Fudge{pyex
+benchmarks}. \Fudge{The instances are all enormous and have some on average 900
+products of 400000 automata with 97 counters each and 9999 constraints}. After
+generating an initial~\InitialNrBenchmarks{}, we remove~\NrBroken{} misgenerated
+instances that did not parse, as well as~\NrTrivial{} instances solved in under
+five seconds by the baseline backend. The benchmarks are run on
+commit~\texttt{\commit}.
 
@@
 
-\subsubsection{Execution Time and Ability to Solve Instances}\label{sec:runtime}
+\subsection{Execution Time and Ability to Solve Instances}\label{sec:runtime}
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
   \includegraphics[width=0.75\linewidth]{graphs/\commit-by-solver.pdf}
@@
 
-\begin{figure}
+\begin{figure}[ht]
   \includegraphics[width=0.75\linewidth]{graphs/\commit-time-boxplot.pdf}
@@
 
-\begin{figure}
+\begin{figure}[ht]
   \begin{minipage}[b]{0.75\linewidth}
@@
 
-\begin{figure}
+\begin{figure}[ht]
   \includegraphics[width=0.75\linewidth]{graphs/\commit-cactus.pdf}
@@
 
-\subsubsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
+\subsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
 
 For baseline and \Calculus{}, \Catra{} offers the ability to find the equivalent Presburger formula representing a given instance.
-For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} \Fudge{we use the specially tailored approach described in some section}.
-For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in Sections~\ref{sec:scaling} and~\ref{sec:runtime}.
+For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} we use the specially tailored approach described in \cref{sec:finding-the-image}.
+For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in \cref{sec:scaling,sec:runtime}.
 
@@
 maximally eager mode where the product is first computed before any
-satisifiability check is performed. We run the experiments with a timeout of~\ImageTimeout{}. The results of the experiment is summarised
-in Figure~\ref{fig:cactus:image} and Table~\ref{tab:image-results}. \Fudge{We see here that something happens}.
-
-\begin{figure}
+satisifiability check is performed. We run the experiments with a timeout
+of~\ImageTimeout{}. The results of the experiment is summarised in
+\cref{fig:cactus:image} and \cref{tab:image-results}. \Fudge{We see here that
+something happens}.
+
+\begin{figure}[ht]
   \caption{The number of instances \Catra{} was able to find the Presburger form of the image for within a given number of seconds per backend.}
@@
 
-\section{Conclusions}
+\section{Conclusion}
 
@@
 
-\section{Discussions}
-
-\contents{
-\item cf Parikh Images of Grammars: Complexity and Applications -- what is the best complexity we can do? https://ieeexplore.ieee.org/abstract/document/5571050
-
-}
-
 %% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
+% \begin{acks}                            %% acks environment is optional
+%                                         %% contents suppressed with 'anonymous'
+%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
+%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
+%   %% acknowledge financial support and will be used by metadata
+%   %% extraction tools.
+%   This material is based upon work supported by the
+%   \grantsponsor{GS100000001}{National Science
+%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+%   conclusions or recommendations expressed in this material are those
+%   of the author and do not necessarily reflect the views of the
+%   National Science Foundation.
+% \end{acks}
 

### 2022-07-05 main.tex
conclusion

@@
 
-\contents{
-  \item Future extensions: more logics
-  \item Loop invariants?
-  \item Integration into string solvers?
-  \item We have shown fastest
-  \item We have shown versatile
-  \item We have shown concretely useful
-}
+In this paper we have introduced an automated theorem prover-friendly calculus
+to compute commutating operations on intersections of regular languages that we
+call \Calculus{}. We have evaluated it on \NrBenchmarks{} Parikh automata
+intersection problems generated \Fudge{by the Ostrich+} string solver
+\cite{ostrich} solving \Fudge{standard string constraint benchmarks} using our
+Parikh automata solver \Catra{}.
+
+Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
+usage and solve-time compared to the baseline approach laid out in
+\cite{generate-parikh-image} when implemented on the same underlying automated
+theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
+\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
+and generally outperforming it for timeouts under 30 seconds with its advantage
+increasing drastically for even shorter timeouts. 30 seconds would generally be
+considered a long timeout for our intended use as supporting infrastructure to a
+string constraint solver.
+
+Future investigations involve two tracks. The first one is integration into
+existing string solvers (wich \Ostrich{} being a particularly promising
+candidate due to its shared use of \Princess{}), and further adaptation to that
+use case. Closer inspection of the instances where we currently time out should
+be useful to further improve our heuristics.
+
+The second track for future improvements is the extension into other problem
+domains, including other logics, model checking problems, as well as to more
+powerful automata such as transducers. In principle, we are also already able to
+express stronger constraints than Parikh automata, due to our use of a full
+automated theorem prover which allows adding arbitrary constraints in addition
+to the expected Presburger formulae.
 

### 2022-07-05 main.tex
Work on introduction

@@
    memory-intensive. We contribute a novel understanding of how Parikh maps can
-   be combined with arbitrary commutative monoid morphisms to represent
-   real-world constraints on regular languages, most notably the length
+   be tackled as a constraint-solving problem to solve real-world constraints
+   stemming from functions on regular languages, most notably the length
    constraint. Furthermore, we show how this formulation can be efficiently
@@
    \Calculus{} in a tool called \Catra{}, and evaluate it on constraints
-   produced by the \Ostrich{} string constraint solver when solving
+   produced by the \OstrichPlus{} string constraint solver when solving
    \Fudge{standard string constraint benchmarks involving string lengths}. We
@@
    over-approximating method recently described
-   by~\citeauthor{approximate-parikh}, and in many cases also the~\Nuxmv{} model
-   checker.
+   by~\citeauthor{approximate-parikh}, and for realistic timeouts for constraint
+   solving also the~\Nuxmv{} model checker.
 \end{abstract}
@@
 solving string constraints in automata-based solvers such as \Ostrich{}
-\cite{ostrich}, notably in representing constraints on string lengths.
-Specialised string solvers like e.g. Trau \cite{trau} also perform extensive
-reasoning using Parikh images, but do so on restricted automata that do not
-suffer from the blow-up we seek to address, though it might be possible to
-adapts aspects of our approach to theirs.
+\cite{ostrich}, notably in representing constraints on string lengths. The
+enhanced \OstrichPlus{} solver~\cite{ostrich-plus} makes even more extensive use
+of Parikh images.
 
@@
 In addition to the elimination of existentially quantified variables, another
-problem arises from length constraints in string solvers. Conjunctions of string
+problem arises in string solvers in particular. Conjunctions of string
 constraints lead to the computation of length images of products (intersections)
 of regular languages represented as finite automata. Applying the approach
-in~\cite{generate-parikh-image} would require the eager computation of the
-product before its length image. In several instances we have observed while
-solving real-world string constraints, the computation of the product of
+in~\cite{generate-parikh-image} would in this case require the eager computation
+of the product before its length image. In several instances we have observed
+while solving real-world string constraints, the computation of the product of
 automata exhausts the memory of any machine available due to the exponential
@@
 individual automata of the product \cite{approximate-parikh}. This approach only
-works for unsatisfiable instances, and comes with a harsh penalty on satisfiable
-instances.
+works for unsatisfiable instances, and comes with a harsh penalty for
+satisfiable instances.
 
@@
 interleave the computation of arbitrarily deep products of automata with the
-product's Parikh image, and is generalised to an arbitrary Parikh-like
-homomorphism over automata labels, including string lengths. This enables us to
-let both calculations inform each other, eliminating unnecessary work, and
-pruning the size of the partial products considered in the computation for a
-smaller memory footprint. Moreover, the method can be used iteratively to tackle
-smaller chunks of the product incrementally, thereby avoiding running out of
-memory at all.
+product's Parikh image, and is generalised to an arbitrary homomorphism over
+automata labels, including string lengths. This enables us to let both
+calculations inform each other, eliminating unnecessary work, and pruning the
+size of the partial products considered in the computation for a smaller memory
+footprint. Moreover, the method can be used iteratively to tackle smaller chunks
+of the product incrementally, thereby avoiding running out of memory at all.
 
@@
 back-ends on \NrBenchmarks{} Parikh automata satisfaction problems generated
-\Fudge{by Ostrich+}.
+\Fudge{by \OstrichPlus{}}.
 
@@
 
+\subsection{Related Work}
+
+Parikh image computations as well as Parikh automata \cite{parikh-automata}
+feature extensively in string solvers, including as mentioned above \Ostrich{}
+and \OstrichPlus{} \cite{ostrich,ostrich-plus}, but also formes the basis of
+Trau~\cite{trau-pldi}, and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images
+frequently appear when introducing cardinality constraints like length or string
+indexing. The state-of-the-art approach to handling Parikh image computation is
+to over-approximate the Parikh image of a product of $k$~automata
+$\ParikhMap(\Automaton_1 \times \ldots \times \Automaton_k)$ with the
+conjunction of the automata's parikh maps, $\AndComp{i=1}^{k}
+\ParikhMap(\Automaton_i)$. This approach works only for unsatisfiable instances,
+and will require falling back to computing the product of the automata before
+using the standard approach for finding its image originally presented in~\cite{generate-parikh-image}.
+
+Outside of string solvers, Parikh automata have been proposed as the basis of
+queries~\cite{graph-queries}, and for solving cardinalities in model checking
+problems involving epistemic logic~\cite{epistemic-logic}.
+
 \section{A demonstration of our approach}\label{sec:motivation}
 
-To illustrate the key features of \Calculus, we start with two regular
-languages, a commutative homomorphism $f$ on them that we care about, and a side
-constraint for which we want to find values in $f$'s image over the intersection
-of the two languages using a heavily elided version of \Calculus. We will use
-the same two regular languages, $\mathtt{ac^*a|b(bb)^*c}$ and $\mathtt{.^*c.^*}$
-for both examples but change the constraint $\varphi$ and homomorphism $f$. In
-essence, we solve the following set of constraints for all possible values of
-$s$ and $\MonoidElement$:
-
+To illustrate the key features of \Calculus, we will informally present our
+reasoning. We begin by considering the following set of string and other
+constraints, which we want to solve for all possible values of $\MonoidElement$,
+a value in some commutative monoid \Fudge{MONOID} and the string $s$, where
+$\varphi(\MonoidElement)$ is some set of constraints on $\MonoidElement$:
 \begin{equation*}
@@
 
-We will represent the regular languages as the automata $\SomethingCSomething{}$
- (\cref{fig:something-c-something}) and $\AcaOrBc{}$ (\cref{fig:aca-or-bc})
- respectively.
+We will represent the strings using the nondeterministic finite automata
+ $\SomethingCSomething{}$ (\cref{fig:something-c-something}) and $\AcaOrBc{}$
+ (\cref{fig:aca-or-bc}) respectively.
+
+ Throughout these examples we will present our three key ideas:
+ \emph{automata-aware splitting}, \emph{lazy enforcement of automata
+ connectivity}, and \emph{incremental materialisation of products}.
 
@@
 
-To keep things simple, in this first example we
-let $f$ be the function that gives the length of a given string, e.g.
-$f\left(\text{curious and curiouser still}\right) = 27$, and $f\left(\text{homo}\right) +
-f\left(\text{morphism}\right) = f\left(\text{homomorphism}\right) = 12$.
-
-\begin{figure}[ht]
-  \begin{minipage}[b]{0.75\linewidth}
+To keep things simple, in this first example we let $f$ be the function that
+gives the length of a given string, e.g. $f\left(\text{curious and curiouser
+still}\right) = 27$, and $f\left(\text{homo}\right) +
+f\left(\text{morphism}\right) = f\left(\text{homomorphism}\right) = 12$, and we
+let $\varphi(\MonoidElement) = \exists k \HoldsThat m = 2k + 1 \land k \geq 0$,
+i.e. require that the length of the string is odd.
+
+\begin{figure}[t]
+  \begin{minipage}[b]{0.50\linewidth}
   \centering 
@@
   \end{minipage}
+  \Description[Two deterministic finite automata, one above the other]{NFAs 
+    recognising the languages ac^*a|b(bb)^*c and ac^*a|b(bb)^*c respectively.
+    The first one has 4 states and the second one 2.}
   \end{figure}
@@
 
-To give an intuition for the core idea of \Calculus{}, we will annotate the
-transitions of $\SomethingCSomething$ and $\AcaOrBc$ with variables illustrating
-how many times they are taken, and perform informal reasoning on them
-corresponding to the reasoning performed by our calculus.
-
-\begin{figure}[ht]
+Similar to \cite{generate-parikh-image} we associate each transition
+$\Transition$ with an integer variable $\TransitionVar_\Transition \geq 0$
+representing how many times $\Transition$ is used that we leave implicitly
+existentially quantified. These variables are shown as annotations under the
+transition labels in \cref{fig:aca-or-bc,fig:something-c-something}. We then
+apply graph-like flow reasoning by adding linear equations (e.g. $1 + {l_b}' =
+l_a + l_b$, or $l_a + l_c' = 1$) that equate the incoming transitions into a
+state to the ones going out. For the initial and accepting states we add an
+incoming and outgoing flow of 1, respectively. We will use these linear
+equations to perform reasoning to reduce the size of the product before we
+compute it. Note that we have already performed some linear reasoning to
+eliminate some of the variables in the graphs above, for example to introduce
+the constant number $1$ in \cref{fig:something-c-something}, which comes from $1
++ r_\Sigma = x + r_\Sigma$ solved for $x$.
+
+\begin{figure}[t]
   \centering 
-    \includegraphics[width=0.75\linewidth]{aca_or_bc_simplified}
+    \includegraphics[width=0.50\linewidth]{aca_or_bc_simplified}
     \caption{$\AcaOrBc{}$ with its symbolic transition counters after
     simplification of their equations.}\label{fig:example-simplify-1}
+    \Description[The automaton $\AcaOrBc{}$ labelled with variables counting the number of times each transition is taken, after having simplified the equations.]{The starting state shows an initial flow of 1, with two outgoing transitions to $A$ ($1-l_c'$ on letter a, and $l_b' + l_c'$ on b respectively).
+$A$ has a flow of $l_c$, and its outgoing transition into the final state, $F$ is the same as the incoming one from $S$, the initial state.
+The $B$ transition has an outgoing transition labeled $c/ l_c'$ to the final state $F$, and one backwards transition $b / l_b'$ to the starting state.}
   \end{figure}
 
-Note that we have already performed some reasoning in the starting figures; we
-use the same symbol for the number of times each of the upper transitions of
-$\AcaOrBc$ are taken, since we know that they must be taken the same number of
-times, and we introduced the constant $1$ for the middle transition in
-$\SomethingCSomething$, since it must always be taken precisely once to reach
-the accepting state. We use standard graph-theoretical flow reasoning to set the
-initial flow into the starting state to 1 and use linear equations to enforce
-balanced flow throughout the automaton to determine the relations between the
-counting variables on each transition (e.g. $1 + {l_b}' = l_a + l_b$, or $l_a +
-l_c' = 1$). Strictly speaking, this is where the constant number $1$ in
-\cref{fig:something-c-something} comes: from $1 + r_\Sigma = x + r_\Sigma$
-solved for $x$. Using analogous linear equation reasoning loses additional
-variables and gives us the annotations in \cref{fig:example-simplify-1}.
-
-We sum the individual transition variables up to get the length we are actually
-interested in, e.g. the image of $f$ on $\SomethingCSomething =
-\exists_{r_\Sigma, r_\Sigma'} \: r_\Sigma + 1 + r_\Sigma'$. For
-$\SomethingCSomething{}$, this is all we need to produce the image since all
-transitions are always accessible from any path from the initial state to the
-accepting state.
+We sum the individual transition variables to get the length $\MonoidElement$ we
+are actually interested in. For example the image of $f$ on
+$\SomethingCSomething$ is given by $\exists_{r_\Sigma, r_\Sigma'} \: 2k + 1 =
+r_\Sigma + 1 + r_\Sigma' \land k \geq 0$. For $\SomethingCSomething{}$, this is
+all we need to produce the image since all transitions are always accessible
+from any path from the initial state to the accepting state. The difficult
+connectivity constraint is not necessary to enforce here, since any path through
+$\SomethingCSomething{}$ is connected, and therefore the flow equations
+described above are sufficient to capture its image.
 
 The story gets more complicated with $\AcaOrBc{}$, where we have a choice of two
-paths, and where that choice affects the possibility of taking either of the two
-loops. In that case (letting the existential quantifiers from now on be
-implicit), we would have the image $2l_a + l_c + l_c' + l_b + l_b'$. If we apply
-standard linear reasoning on the equations (e.g. $1 + {l_b}' = l_a + l_b$, or
-$l_a + l_c' = 1$), we can reduce the number of bound ($\exists$-quantified)
-variables somewhat, obtaining the relations in
-Figure~\ref{fig:example-simplify-1}, and the corresponding over-approximation on
+paths, up or down, and where that choice affects the possibility of taking
+either of the two loops. In that case, we would have the image $2l_a + l_c + l_c' + l_b + l_b'$.
+If we apply standard linear reasoning on the equations (e.g. $1 + {l_b}' = l_a +
+l_b$, or $l_a + l_c' = 1$), we can reduce the number of bound
+($\exists$-quantified) variables somewhat, obtaining the relations in
+\cref{fig:example-simplify-1}, and the corresponding over-approximation on
 the $f$-image $2 + l_c + 2l_b'$. It is an over-approximation because it only
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \centering 
-    \includegraphics[width=0.75\linewidth]{something_c_something_length_simplified}
+    \includegraphics[width=0.4\linewidth]{something_c_something_length_simplified}
     \caption{$\SomethingCSomething{}$ after using linear reasoning over lengths to replace a variable.}\label{fig:example-length-reduced}
+    \Description[The automaton $\SomethingCSomething{}$ after reasoning over lengths.]{The same automaton as before for $\SomethingCSomething{}$, but with the transition from start to final annotated with a constant number 1, and the leftmost self-loop transition replaced with the equation $1 + l_c + 2l_b' -
+    r_\Sigma'$.}
   \end{figure}
@@
 to replace one additional variable, e.g. $r_\Sigma = 1 + l_c + 2l_b' -
-r_\Sigma'$ (see Figure~\ref{fig:example-length-reduced}). Note that the two
-automata are now in direct contact. Any reasoning performed on one of them could
-be propagated to the other.
+r_\Sigma'$ (see \cref{fig:example-length-reduced}). Note that the two automata
+are now in contact, allowing us to propagate reasoning on one to the other.
 
@@
 
-We talked initially about having a constraint on the image (or conversely, on
-the strings of the language), so let us add one. We are only interested in
-strings of odd length, e.g. $2 \nmid  f(\AcaOrBc \times \SomethingCSomething)$.
-This means that: $2 \nmid 2 + l_c + 2l_b' \iff 2 \nmid l_c$. To encode this condition, we simply replace $l_c$ with $2k + 1$ with an existentially quantified integer $k \geq 0$.
-
 This is as far as we can get with this reasoning. We must now choose: either we
-compute the product and continue our reasoning, or we perform a case split. Let
-us try a case split in order to put off computing a potentially large product.
-We pick an early transition in the largest automaton with loops, in our case
-$\FromLabelTo{start}{a}{sawA}$ of $\AcaOrBc$. This creates two cases: $1-l_c' =
-0$ and $1-l_c' > 0$.
-We start with the first case, and after propagating the now known values and removing any transition that becomes zero, we get the automaton of Figure~\ref{fig:aca-or-bc-length-split-1}.
-
-\begin{figure}[ht]
+compute (materialise) the product and continue our reasoning, or we perform a
+case split. Let us try a case split in order to put off computing a potentially
+large product. 1 an early transition in the largest automaton with loops,
+in our case $\FromLabelTo{S}{a}{A}$ of $\AcaOrBc$. This creates two
+cases: $1-l_c' = 0$ and $1-l_c' > 0$. We start with the first case, and after
+propagating the now known values and removing any transition that becomes zero,
+we get the automaton of \cref{fig:aca-or-bc-length-split-1}.
+
+\begin{figure}[t]
   \centering 
     \includegraphics[width=0.75\linewidth]{aca_or_bc_length_split_1}
-    \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{start}{a}{sawA}$ is not used.}\label{fig:aca-or-bc-length-split-1}
+    \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ is not
+    used.}\label{fig:aca-or-bc-length-split-1}
+    \Description[The automaton \AcaOrBc{} where we have removed the transition
+    from $S$ to $A$.]{ The state $A$ hangs free without incoming transitions,
+    and only the lower part of the automaton remains.}
   \end{figure}
 
-  We immediately notice that the self-loop of state $sawA$ is now disconnected
-  and must be removed. This condition can be detected efficiently using standard
+  We immediately notice that the self-loop of state $A$ is now disconnected and
+  must be removed. This condition can be detected efficiently using standard
   forwards/backwards reachability from the initial and accepting states
-  respectively.
-In other words, $2k+1 = 0$ for this case and is implied by the splitting constraint.
-However, no choice of integer $k$  will satisfy this equation.
-Therefore, we close this branch and proceed with the other one -- where $1-l_c' > 0$.
-Note that we do so without ever computing a potentially expensive product of automata.
-
-  \begin{figure}[ht]
+  respectively. In other words, $2k+1 = 0$ for this case and is implied by the
+  splitting constraint. However, no choice of integer $k$  will satisfy this
+  equation. Therefore, we close this branch and proceed with the other one --
+  where $1-l_c' > 0$. Note that we do so without ever computing a potentially
+  expensive product of automata.
+
+  \begin{figure}[t]
     \centering 
-      \includegraphics[width=0.75\linewidth]{aca_or_bc_length_split_2}
-      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{start}{a}{sawA}$ is used.}\label{fig:aca-or-bc-length-split-2}
+      \includegraphics[width=0.5\linewidth]{aca_or_bc_length_split_2}
+      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ \emph{is} used.}\label{fig:aca-or-bc-length-split-2}
+      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ when we keep the transition from $S$ to $A$.
+This removes the transition between $B$ and $F$, but otherwise maintains the automaton.}
     \end{figure}
@@
 we also get a reduced variant of $\AcaOrBc$, as seen in
-Figure~\ref{fig:aca-or-bc-length-split-1}. Note that $sawB$ is no longer
-backwards-reachable from $final$ and therefore dead, along with both transitions
-into it. We add the conclusion $l_b' = 0$, which severs both transitions into
-the state and removes it from our automaton. Following this we are ready to
-compute the product, which can be seen in Figure~\ref{fig:length-split-product}.
-
-\begin{figure}[ht]
+\cref{fig:aca-or-bc-length-split-2}. Note that there is now no way to select any
+transition variable so as to disconnect a loop from a path between $S$ and $F$,
+and we can therefore stop worrying about enforcing connectivity.
+
+\begin{figure}[t]
   \centering 
-    \includegraphics[width=0.75\linewidth]{length_split_product}
+    \includegraphics[width=0.5\linewidth]{length_split_product}
     \caption{The product of $\AcaOrBc$ and $\SomethingCSomething$ after
@@
   occurrence of each character corresponds to the sum of usages of transitions
-  containing it, e.g. relations of the kind $\TransitionCount{a} = 2 - 2l_c$,
+  containing it, e.g. relations of the kind $\TransitionCount{a} = 2 - 2l_c'$,
   and $\TransitionCount{b} = 2 l_b' + l_c'$.\footnote{For the attentive reader:
@@
 \TransitionCount{b}$, that is that there are more instances in the accepted
-string of letters a than b, to obtain $2 - 2l_c > 2 l_b' + l_c'$. Combining this
+string of letters a than b, to obtain $2 - 2l_c' > 2 l_b' + l_c'$. Combining this
 equation with the requirement from $\SomethingCSomething$ that
@@
 $=1$), which gives the Parikh image $\exists_{r_2} \TransitionCount{a} = 1 \land
-\TransitionCount{b} = 0 \land \TransitionCount{c} = r_2 + 1$, and after trivial
+\TransitionCount{b} = 0 \land \TransitionCount{c} = r_2 + 1$, and after
 existence-elimination using reasoning on lower bounds (e.g. $r_2 \geq 0$) we
@@
 \section{Preliminaries}
+
+\contents{\item Monoid
+\item Alfabet
+\item ord oever alfabet
+\item konkatenering
+}
 
@@
 $\State'$ with label $\Label \in \Alphabet$ as $\Transition =
-\FromLabelTo{\State}{\Label}{\State'}$. When one or more of the states and
-labels of a transition are uninteresting we will omit it.
+\FromLabelTo{\State}{\Label}{\State'}$. \Fudge{When one or more of the states and
+labels of a transition are uninteresting we will omit them}.
 
@@
 \subsection{The Parikh Map and its Image}
-Formally, the \textit{Parikh map} over a context-free language $\Strings=
+Formally, the \textit{Parikh map} over an alphabet $\Alphabet=
 \left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:
-
 $$
@@
 That is, $\ParikhMap(s)$ is a vector of the number of occurrences of each
-character in the language for a given string $s$. For example, for  $\Strings =
+character in the language for a given string $s$. For example, for  $\Alphabet =
 \Set{a, b}$, we would have $\ParikhMap(abb) = \VectorLiteral{1, 2}$.
@@
 the language $\Language \subseteq \Strings$ as:
-
 \[
@@
 1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
-notation $\#l$ to talk about an individual letter $l$. For example, for the Parikh vector above, we would have $\CountOf{a} = 1$. You have already seen 
-this in Section~\ref{sec:introduction:parikh}.
-
-Parikh's theorem states that ignoring order, any context-free grammar has a
+notation $\#l(w)$ to talk about an individual letter $l$ in a word $w$. For
+example, for the Parikh vector above, we would have $\CountOf{a} = 1$. You have
+already seen this in Section~\ref{sec:introduction:parikh}.
+
+Parikh's theorem states that any context-free language has a
 letter-equivalent regular language (c.f.~\cite{construction} for a construction of
@@
 
+\contents{\item this means we can represent this as Presburger
+\item only interested in regular languages}
+
+\subsection{Generalised Parikh maps}\label{sec:generalised}
+
+To generalise the logic of the Parikh map, we introduce a morphism $\Map:\: \Sigma^*
+\to \Monoid$ where $\Monoid = \left(X;+_{\Monoid};0_{\Monoid}\right)$ is a commutative monoid where the following holds:
+\begin{itemize}
+  \item $\Map(\epsilon) = 0_{\Monoid}$
+  \item $\Map(a \Concat b) = \Map(a) +_{\Monoid} \Map(b)$, where $\Concat$ is
+  the string concatenation operation
+\end{itemize}
+
+Many other generalisations of the Parikh map have been studied. Prominent
+examples include generalising the Parikh map to segments of a fixed length
+\cite{KARHUMAKI1980155} and the more general Parikh matrix, which gives more
+information about a word than the standard Parikh image \cite{parikh-matrix}.
+Another notable generalisation is the p-vector, introduced
+in~\cite{infinite-words}, which denotes the position of each letter in the word
+rather than the number of their occurrences and allows for generalisations into
+infinite alphabets. All of these in some sense extend the Parikh map. By
+contrast, the main utility of the formulation introduced here is that it allows
+us to \emph{remove} something, thereby potentialy obtaining an easier problem.
+
 \subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
@@
 a regular language recognised by a DFA $\Automaton = \AutomatonTuple$ as:
+
+\contents{\item replace implies with singlearrow
+\item splitta sista implikationerna i tvaa!
+\item intuition for what the variables mean}
 
@@
   \left(\FinalStateVar_{\State} \text{ if } \State \in \AcceptingStates \text{ otherwise } 0 \right) +
-  \sum_{\Transition = \FromLabelTo{\State}{}{}} - \sum_{\Transition = \FromLabelTo{}{}{\State}}
+  \sum_{\Transition = \FromLabelTo{\State}{}{}} \Filter(\Transition) - \sum_{\Transition = \FromLabelTo{}{}{\State}} \Filter(\Transition)
 = \begin{cases}
@@
 \subsubsection{An Example}
-We apply Equation~\ref{eq:generate-parikh} to the automaton $\AcaOrBc{}$ of Figure~\ref{fig:aca-or-bc}.
-We then get the quantified Presburger formula seen in Equation~\ref{eq:parikh-formula}.
+We apply \eqref{eq:generate-parikh} to the automaton $\AcaOrBc{}$ of \cref{fig:aca-or-bc}.
+We then get the quantified Presburger formula seen in \eqref{eq:parikh-formula}.
 As before, we omit the implicit existential quantifiers for brevity, along with the constranints requiring all variables to be nonnegative.
@@
 
-\section{Overview of our generalised Parikh images}
-
-\subsection{Generalised Parikh Images}\label{sec:generalised}
-
-To generalise the logic over Parikh images, we introduce a morphism $\Map:\: \Sigma
-\to \Monoid$ where $\Monoid = \left(X;+_{\Monoid};0_{\Monoid}\right)$ is a commutative monoid where the following holds:
-\begin{itemize}
-  \item $\Map(\epsilon) = 0_{\Monoid}$
-  \item $\Map(a \Concat b) = \Map(a) +_{\Monoid} \Map(b)$, where $\Concat$ is
-  the string concatenation operation
-\end{itemize}
-
-Many other generalisations of the Parikh map have been studied, most of them
-less plain than this one. Prominent examples include generalising the Parikh map
-to segments of a fixed length \cite{KARHUMAKI1980155} and the more general
-Parikh matrix, which gives more information about a word than the standard
-Parikh image \cite{parikh-matrix}. Another notable generalisation is the
-p-vector, introduced in~\cite{infinite-words}, which denotes the position of
-each letter in the word rather than the number of their occurrences and allows
-for generalisations into infinite alphabets. All of these, in some sense, add
-something which is not already there in the Parikh map. By contrast, the main
-utility of the formulation introduced here is that it allows us to \emph{remove}
-something, thereby potentialy obtaining an easier problem.
-
-\subsection{Applications}
+\section{Applications of generalised Parikh images}\label{sec:applications}
 
 \subsubsection{Example~1: String length}
-
+\Fudge{We have already illustrated refer intro}
 A useful example of such a simplifying morphism can express string length, the
@@
 we now ignore not just the order of characters but also their differences. This
-means that the branching automaton of Figure~\ref{fig:len-branch} becomes
-equivalent to the non-branching automaton of Figure~\ref{fig:len-branch-free}.
+means that the branching automaton of \cref{fig:len-branch} becomes
+equivalent to the non-branching automaton of \cref{fig:len-branch-free}.
 Note that the elimination of the choice of paths also effectively eliminates the
@@
 
-\begin{figure}
-  \includegraphics[width=0.5\linewidth]{choice}
-  \caption{An example automaton with a choice of two paths: either we start with an a or a b.}
-  \label{fig:len-branch}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=0.5\linewidth]{no-choice}
-  \caption{An equivalent automaton to the one in Figure~\ref{fig:len-branch}
-  when treating any letter the same as would be the case with a length
-  constraint.}\label{fig:len-branch-free}
-\end{figure}
-
 \subsubsection{Example~2: Parikh Automata}\label{sec:parikh-automata}
+\contents{\item Ostrich+!}
 
@@
 
-An example of a Parikh automata version of $\AcaOrBc{}$ can be seen in \cref{fig:parikh-automaton}.
-A corresponding Presburger version of the constraint seen in \cref{sec:introduction:motivation} is $\psi(\Vec{v}) = v_1 > v_2$.
+An example of a Parikh automata version of $\AcaOrBc{}$ can be seen in
+\cref{fig:parikh-automaton}. A corresponding Presburger version of the
+constraint seen in \cref{sec:motivation} is $\psi(\Vec{v}) = v_1 > v_2$.
 
@@
 \begin{definition}
+  \Fudge{fixme: use pattern matching on path definition}
   The \textit{states} of a path $\StatesOf(\Path) = \Set{\State \SuchThat
@@
 \begin{definition}
+  \Fudge{if every valid path containing q overlaps with C}
   A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$,
@@
   $\Automaton = \AutomatonTuple$, modulo some map $\Map$ to a commutative monoid
-  $\Monoid$ as described in Section~\ref{sec:generalised} and with a transition
+  $\Monoid$ as described in \cref{sec:generalised} and with a transition
   selection function $\Filter:\: \Transitions \to \Naturals$ holds when
@@
 
-Finally, we present the rules of our calculus for one automaton in Table~\ref{tbl:rules:single}. Note that we
-operate on sets of symbols (terms, clauses). Additionally, we also use the
-convention of splitting the formulas into linear inequalities ($\SomeInequalities$)
-and any other clauses ($\SomeClause$).
+Finally, we present the rules of our calculus for one automaton in
+\cref{tbl:rules:single}. Note that we operate on sets of symbols (terms,
+clauses). Additionally, we also use the convention of splitting the formulas
+into linear inequalities ($\SomeInequalities$) and any other clauses
+($\SomeClause$).
 
@@
   {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
-  None \\
+  None \\[5ex]
 
@@
   \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  if $\Transition \in \Transitions_\Automaton$ \\
+  if $\Transition \in \Transitions_\Automaton$ \\[5ex]
 
@@
   \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
-  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$\\
+  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[5ex]
 
@@
 $\Transition \mapsto 1$ for transitions, and the constraints that the length is
-odd using the same trick as in \cref{sec:introduction:motivation}, we have the
+odd using the same trick as in \cref{sec:motivation}, we have the
 definitions in \cref{fig:example:single:equivalences} (omitting existential

### 2022-07-06 main.tex
intro

@@
 
-The Parikh map is a characterisation of a context-free language as a vector of
-its character counts, famously a semilinear set~\cite{parikh-theorem}. It
-appears naturally as part of \Fudge{many operations} in model checking and
-solving string constraints in automata-based solvers such as \Ostrich{}
-\cite{ostrich}, notably in representing constraints on string lengths. The
-enhanced \OstrichPlus{} solver~\cite{ostrich-plus} makes even more extensive use
-of Parikh images.
-
-It is possible to compute the Parikh image of any context-free language using
-the method described in~\cite{generate-parikh-image}. Later improvements have
-produced a construction taking at most linear time to
-produce~\cite{muscholl-linear}. However, the resulting existentially quantified
-clauses are costly to eliminate as the number of variables increases, in
-practice making many real-world problems intractable.
-
-In addition to the elimination of existentially quantified variables, another
-problem arises in string solvers in particular. Conjunctions of string
-constraints lead to the computation of length images of products (intersections)
-of regular languages represented as finite automata. Applying the approach
-in~\cite{generate-parikh-image} would in this case require the eager computation
-of the product before its length image. In several instances we have observed
-while solving real-world string constraints, the computation of the product of
-automata exhausts the memory of any machine available due to the exponential
-blow-up in size of the product, quickly becoming intractable as the number of
-automata in the product increases. The current best published mitigation for
-this problem is an over-approximation that works by approximating the Parikh
-image of a product of automata to be the conjunction of the image of the
-individual automata of the product \cite{approximate-parikh}. This approach only
-works for unsatisfiable instances, and comes with a harsh penalty for
-satisfiable instances.
+The Parikh image is a characterisation of formal languages in terms of
+their character counts. Given a language over an
+alphabet~$\{a_1, \ldots, a_k\}$, the Parikh image is a set of
+$k$-dimensional vectors that contains some
+vector~$\VectorLiteral{m_1, m_2, \ldots, m_k}$ if and only if the
+formal language contains a word in which each $a_i$ occurs $m_i$
+times. It is a classical result that the Parikh image of every
+context-free language (and, thus, also of every regular language) is a
+semilinear set~\cite{parikh-theorem}, i.e., Presburger-definable.
+
+Parikh images play a central role in many automata-based algorithms,
+for instance and notably in today's string solvers, which often have
+to process constraints that combine regular language membership with
+word length. To decide whether a simple formula like
+$x \in \mathcal{L}_1 \wedge y \in \mathcal{L}_2 \wedge |x| > |y|$,
+with string variables~$x, y$ and regular
+languages~$\mathcal{L}_1, \mathcal{L}_2$, is satisfiable, it is
+necessary to reason about the sets of word lengths induced by
+$\mathcal{L}_1, \mathcal{L}_2$, which is a special case of the Parikh
+image.  This required combined reasoning about strings and string
+length has long been identified as a major bottleneck in string
+solvers
+\cite{DBLP:conf/cav/AbdullaACHRRS15,length-aware-solver,approximate-parikh,more}.
+Other string solvers make use of Parikh automata~\cite{X}, and thus
+Parikh images in the general case, to handle operations including
+\verb!str.substr!  and \verb!str.at!, which comes at an even higher
+price in terms of computational complexity~\cite{ostrich-plus}.
+
+\iffalse
+It appears naturally as part of \Fudge{many operations} in model
+checking and solving string constraints in automata-based solvers such
+as \Ostrich{} \cite{ostrich}, notably in representing constraints on
+string lengths. The enhanced \OstrichPlus{} solver~\cite{ostrich-plus}
+makes even more extensive use of Parikh images.
+\fi
+
+It is possible to compute an existential Presburger formula describing
+the Parikh image of any context-free language in linear
+time~\cite{generate-parikh-image}; for the special case of regular
+languages, this result was also stated in
+\cite{muscholl-linear}. While theoretically elegant, this construction
+has several disadvantages, often making it unpractical for integration
+into algorithms. Firstly, the constructed Presburger formula contains
+a linear number of existential quantifiers in the size of the
+considered grammar, as well as complex Boolean structure, which is
+needed to express the connectedness of sets of productions considered
+in the construction. Eliminating those quantifiers to obtain a
+quantifier-free representation of the Parikh image has exponential
+complexity~\cite{X}, and is in practice often impossible in reasonable
+time. Just solving the Parikh image membership problem is NP-complete,
+as it corresponds to computing a satisfying assignment of the
+existential Presburger formula, and taxing for solvers as
+well~\cite{ostrich-plus}.
+
+\iffalse
+Later improvements have produced a construction taking at
+most linear time to produce~\cite{muscholl-linear}. However, the
+resulting existentially quantified clauses are costly to eliminate as
+the number of variables increases, in practice making many real-world
+problems intractable.
+\fi
+
+Secondly, in applications involving regular languages, it is typically
+necessary to consider the Parikh image not only of a single automaton,
+but of the intersection of multiple automata. This problem arises in
+string solvers in particular, as conjunctions of string constraints
+lead to the computation of length images of products (intersections)
+of regular languages represented as finite automata. Applying the
+approach in~\cite{generate-parikh-image} would in this case require
+the eager computation of the product before its length image, and
+result in an existential Presburger formula of exponential size (in
+the number of automata). In several instances we have observed while
+solving real-world string constraints, the computation of the product
+of automata exhausts the memory of any machine available due to the
+exponential blow-up in size of the product, quickly becoming
+intractable as the number of automata in the product increases. The
+current best published mitigation for this problem is an
+over-approximation that works by approximating the Parikh image of a
+product of automata to be the conjunction of the image of the
+individual automata of the product \cite{approximate-parikh}. This
+approach only works for unsatisfiable instances, and comes with a
+harsh penalty for satisfiable instances.
 

### 2022-07-06 main.tex
intro

@@
 path in the presence of loops. For example, no valid path would have both $l_c >
-0$ and $l_b' > 0$, but no such restriction exists in the formula above.
+0$ and $1 - l_c' = 0$, but no such restriction exists in the formula above.
 
@@
 
-Still, this approximate solution allows us to perform further reasoning. Since
-we know that the length of a string accepted by the intersection of the two
-automata's languages must be the same, we also know that their images (string
-lengths) must be the same under $f$, e.g. that $r_\Sigma + 1 + r_\Sigma' = 2 +
-l_c + 2l_b' \iff r_\Sigma + r_\Sigma' = 1 + l_c + 2l_b'$. This in turn allows us
-to replace one additional variable, e.g. $r_\Sigma = 1 + l_c + 2l_b' -
-r_\Sigma'$ (see \cref{fig:example-length-reduced}). Note that the two automata
-are now in contact, allowing us to propagate reasoning on one to the other.
+Still, this approximate view allows us to perform further reasoning. Since we
+know that the length of a string accepted by the intersection of the two
+automata's languages must be the same, we also know that their images must be
+the same under $f$, e.g. that $r_\Sigma + 1 + r_\Sigma' = 2 + l_c + 2l_b' \iff
+r_\Sigma + r_\Sigma' = 1 + l_c + 2l_b'$. This in turn allows us to replace one
+additional variable, e.g. $r_\Sigma = 1 + l_c + 2l_b' - r_\Sigma'$ (see
+\cref{fig:example-length-reduced}). Note that the two automata are now in
+contact, allowing us to propagate reasoning on one to the other.
 
@@
 case split. Let us try a case split in order to put off computing a potentially
-large product. 1 an early transition in the largest automaton with loops,
-in our case $\FromLabelTo{S}{a}{A}$ of $\AcaOrBc$. This creates two
-cases: $1-l_c' = 0$ and $1-l_c' > 0$. We start with the first case, and after
-propagating the now known values and removing any transition that becomes zero,
-we get the automaton of \cref{fig:aca-or-bc-length-split-1}.
+large product.
+
+We apply the principle of splitting and select an early transition in the
+largest automaton with loops, in our case $\FromLabelTo{S}{a}{A}$ of $\AcaOrBc$.
+This creates two cases: $1-l_c' = 0$ and $1-l_c' > 0$. We start with the first
+case, and after propagating the now known values and removing any transition
+that becomes zero, we get the automaton of \cref{fig:aca-or-bc-length-split-1}.
 
@@
 
-  We immediately notice that the self-loop of state $A$ is now disconnected and
-  must be removed. This condition can be detected efficiently using standard
-  forwards/backwards reachability from the initial and accepting states
-  respectively. In other words, $2k+1 = 0$ for this case and is implied by the
-  splitting constraint. However, no choice of integer $k$  will satisfy this
-  equation. Therefore, we close this branch and proceed with the other one --
-  where $1-l_c' > 0$. Note that we do so without ever computing a potentially
-  expensive product of automata.
+  We immediately notice that the self-loop of state $A$ is now disconnected from
+  the initial state and must be removed. This condition can be detected
+  efficiently using standard forwards/backwards reachability from the initial
+  and accepting states respectively. In other words, $2k+1 = 0$ for this case
+  and is implied by the splitting constraint. However, no choice of integer $k$
+  will satisfy this equation. Therefore, we close this branch and proceed with
+  the other one -- where $1-l_c' > 0$. Note that we do so without ever computing
+  a potentially expensive product of automata.
 
@@
 \cref{fig:aca-or-bc-length-split-2}. Note that there is now no way to select any
-transition variable so as to disconnect a loop from a path between $S$ and $F$,
-and we can therefore stop worrying about enforcing connectivity.
+transition variable to disconnect a loop from a path between $S$ and $F$,
+and we can therefore stop worrying about enforcing connectivity at all.
 
@@
     case splitting and removal of dead states.}\label{fig:length-split-product}
+    \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ with the transition from $B$ to $F$ removed. There now remains only one path to an accepting state.}
   \end{figure}
@@
 automata of the product to the new transition counting variables we introduced
-in the product automaton to ensure that they are consistent.
-In our case, we obtain the following equations from the principle that a transition variable in either automaton must be equal to the sum of variables in the resulting edges of the product, which gives the following two interesting inequality from $\AcaOrBc{}$'s transition $\FromLabelTo{sawA}{c}{sawA}$: $2k + 1 = r_1 + r_2 + 1$
+in the product automaton to ensure that they are consistent. In our case, we
+obtain the following equations from the principle that a transition variable in
+either automaton must be equal to the sum of variables in the resulting edges of
+the product, which gives the following two interesting inequalities from
+$\AcaOrBc{}$'s transition $\FromLabelTo{A}{c}{A}$: $2k + 1 = r_1 + r_2 + 1$
 
@@
 as before for computing the length by adding up the transition counters, we
-obtain the quantified formula $1 + r_1 + 1 = r_2 + 1 = r_1 + r_2 + 3 = 2k + r_1
-+ 3$ for the length. Eliminating the final quantifier gives us the image of
-$f(\AcaOrBc \times \SomethingCSomething)$ modulo odd lengths: $3 + 2k$, for some
-integer $k$.
+obtain the quantified formula $1 + 2k + 1 + 1 + r_b + r_b= 3 + 2k + 2r_b$ for
+the length. Eliminating the final quantifier gives us the image of $f(\AcaOrBc
+\times \SomethingCSomething)$ modulo odd lengths: $3 + 2k$, for some integer
+$k$.
 
@@
       \begin{aligned}
-        & \Filter(\FromLabelTo{start}{a}{sawA}) & = \TransitionVar_1 \\
-        & \Filter(\FromLabelTo{start}{b}{sawB}) & = \TransitionVar_2 \\
-        & \Filter(\FromLabelTo{sawA}{c}{sawA})  & = \TransitionVar_3  \\
-        & \Filter(\FromLabelTo{sawB}{b}{start}) & = \TransitionVar_4 \\
-        & \Filter(\FromLabelTo{sawA}{a}{final}) & = \TransitionVar_5 \\
-        & \Filter(\FromLabelTo{sawB}{c}{final}) & = \TransitionVar_6 \\
+        & \Filter(\FromLabelTo{S}{a}{A}) & = \TransitionVar_1 \\
+        & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_2 \\
+        & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
+        & \Filter(\FromLabelTo{B}{b}{S}) & = \TransitionVar_4 \\
+        & \Filter(\FromLabelTo{A}{a}{F}) & = \TransitionVar_5 \\
+        & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
       \end{aligned}
@@
       \begin{aligned}
-        % start
+        % S
         &  \TransitionVar_4 = \TransitionVar_5 + \TransitionVar_2 - 1 \\
-        % sawA
+        % A
         & \TransitionVar_1 = \TransitionVar_5 \\
-        % sawB
+        % B
         & \TransitionVar_2 = \TransitionVar_4 + \TransitionVar_6 \\
-        % final
+        % F
         & \TransitionVar_5 + \TransitionVar_6 = \FinalStateVar_1 \\
@@
   \begin{aligned}
-    & \Filter(\FromLabelTo{start}{a}{sawA}) & = 1 - \TransitionVar_6 \\
-    & \Filter(\FromLabelTo{start}{b}{sawB}) & = \TransitionVar_4 + \TransitionVar_6 \\
-    & \Filter(\FromLabelTo{sawA}{c}{sawA})  & = \TransitionVar_3  \\
-    & \Filter(\FromLabelTo{sawB}{b}{start}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-    & \Filter(\FromLabelTo{sawA}{a}{final}) & = 1 - \TransitionVar_6 \\
-    & \Filter(\FromLabelTo{sawB}{c}{final}) & = \TransitionVar_6 \\
+    & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
+    & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
+    & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
+    & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
+    & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
+    & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
   \end{aligned}
@@
 transition to the corresponding increment vector (e.g.
-\Map($\FromLabelTo{start}{a}{sawA}) = \begin{bmatrix}1\\0\\0\end{bmatrix}$),
+\Map($\FromLabelTo{S}{a}{A}) = \begin{bmatrix}1\\0\\0\end{bmatrix}$),
 since it should be apparent by now.
@@
       \begin{aligned}
-        & \Filter(\FromLabelTo{start}{a}{sawA}) & = 1 - \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{start}{b}{sawB}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{sawA}{c}{sawA})  & = \TransitionVar_3  \\
-        & \Filter(\FromLabelTo{sawB}{b}{start}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        & \Filter(\FromLabelTo{sawA}{a}{final}) & = 1 - \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{sawB}{c}{final}) & = \TransitionVar_6 \\
+        & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
+        & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
+        & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
+        & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
+        & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
+        & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
         % Other automaton
-        & \Filter(\FromLabelTo{start}{\Sigma}{start}) & = \TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c} \\
-        & \Filter(\FromLabelTo{start}{c}{final}) & = 1 \\
-        & \Filter(\FromLabelTo{final}{\Sigma}{final}) & = \TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c} \\
+        & \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c} \\
+        & \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
+        & \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c} \\
       \end{aligned}
@@
   automaton aca_or_bc {
-    init start;
+    init S;
 
     // We use ASCII values here, this is for lowercase a
-    start -> sawA [97] { l_a += 1 };
-    start -> sawB [98] { l_b += 1 };
-
-    sawB -> start  [98] { l_b += 1 };
-    sawB -> final [99] { l_c += 1 };
-
-    sawA -> sawA [99] { l_c += 1 };
-    sawA -> final [97] { l_a += 1 };
+    S -> A [97] { l_a += 1 };
+    S -> B [98] { l_b += 1 };
+
+    B -> S  [98] { l_b += 1 };
+    B -> final [99] { l_c += 1 };
+
+    A -> A [99] { l_c += 1 };
+    A -> final [97] { l_a += 1 };
 
@@
   automaton something_c_something {
-      init start;
+      init S;
 
       // Special short-hand value for the whole Unicode alphabet.
-      start -> start [any] ;
-      start -> final [99] { r_c += 1 };
-
-      final -> final [any];
-
-      accepting final;
+      S -> S [any] ;
+      S -> F [99] { r_c += 1 };
+
+      F -> F [any];
+
+      accepting F;
   };

### 2022-07-06 main.tex
add marker
motivating example

@@
 want to find the whole Parikh image of the intersection of $\AcaOrBc{}$ and
-$\SomethingCSomething{}$. In other words, now $f$ is a mapping to an object
-(usually represented as a column vector) counting how often each character
-occurs in a string. Since extracting values from a vector is tiresome, we will
-use the shorthand notation $\TransitionCount{a}$ to refer to the number of
-letters a that appear. E.g. $\TransitionCount{a}(\text{curious and curiouser
-still}) = 1$. Note that we still have the property that
-$\TransitionCount{o}(\text{homomorphism}) = \TransitionCount{o}(\text{homo}) +
-\TransitionCount{o}(\text{morphism}) = 2 + 1 = 3$. If you treat the addition as
-element-wise vector addition, this approach works for the vector version as
-well.
+$\SomethingCSomething{}$. In other words, now $f$ is a mapping to a 3-vector
+counting how often each character occurs in a string. Since extracting values
+from a vector is tiresome, we will use the shorthand notation
+$\TransitionCount{a}$ to refer to the number of letters a that appear. E.g.
+$\TransitionCount{a}(\text{curious and curiouser still}) = 1$. Note that we
+still have the property that $\TransitionCount{o}(\text{homomorphism}) =
+\TransitionCount{o}(\text{homo}) + \TransitionCount{o}(\text{morphism}) = 2 + 1
+= 3$. If you treat the addition as element-wise vector addition, this approach
+works for the vector version as well.
 
 For our constraint, we will use $\TransitionCount{a} > \TransitionCount{b}$,
-that is find a count of characters in a string accepted by both automata where
+that is finding a count of characters in a string accepted by both automata where
 there are more a:s than b:s. Since the underlying calculus is the same, we start immediately from Figure~\ref{fig:example-simplify-1}.
@@
   character counts we are really interested in, as opposed to just the length in
-  the previous example. Note that this, like the length calculus before, puts
-  the two automata into contact with each other, before any product has been
-  computed. Some consideration leads us to the definition that the number of
-  occurrence of each character corresponds to the sum of usages of transitions
-  containing it, e.g. relations of the kind $\TransitionCount{a} = 2 - 2l_c'$,
-  and $\TransitionCount{b} = 2 l_b' + l_c'$.\footnote{For the attentive reader:
-  in this example, we leave the catch-all transitions of $\SomethingCSomething$
-  unconstrained since they will be taken care of by the product and would
-  complicate the reasoning since we cannot yet know which character they will
-  correspond to in the final product. In other words, this is a safe
-  under-approximation that only limits our ability to reason before computing
-  the product.}
-
-This relation can be combined with the constraint that $\TransitionCount{a} >
-\TransitionCount{b}$, that is that there are more instances in the accepted
-string of letters a than b, to obtain $2 - 2l_c' > 2 l_b' + l_c'$. Combining this
-equation with the requirement from $\SomethingCSomething$ that
-$\TransitionCount{c} \geq 1$ and $\TransitionCount{c} = l_c + l_c'$ from
-$\AcaOrBc$ gives $0 \geq l_b'$, and since we cannot use a transition a negative
-number of times $l_b' = 0$, which gives the automaton in
-Figure~\ref{fig:example-simplify-2}.
-
-\begin{figure}[ht]
+  the previous example. For the $\Sigma$ labels, since we do not know which
+  letter to use, we introduce a sum $r_{a1} + r_{b1} + r_{c1}$ per $\Sigma$
+  transition.
+
+  As in the previous example, we sum the occurrences for each transition to
+  obtain the counts, here with the element-wise vector addition unpacked:
+  \[
+\MonoidElement =   \begin{bmatrix}
+  r_{a1} + r_{a2} \\
+  r_{b1} + r_{b2} \\
+  r_{c1} + r_{c2} + 1 \\
+  \end{bmatrix} =
+  \begin{bmatrix}
+    2 -2l_c' \\
+    2l_b' + l_c' \\
+    l_c + l_c' \\
+    \end{bmatrix} =
+    \begin{bmatrix}
+      \TransitionCount{a} \\
+      \TransitionCount{b} \\
+      \TransitionCount{c}
+      \end{bmatrix}
+  \]
+
+These equalities can be combined with the constraint that $\TransitionCount{a} >
+\TransitionCount{b}$ to obtain $2 - l_c' > l_b'$. Since we have $1-l_c' \geq 0$
+from one of the transition labels, we know $l_c' \geq 0$, and therefore that $1
+> l_b' \implies l_b' = 0$. Plugging in the same inequality again on the same
+principle, we get $2 - 2l_c' > l_c' \iff l_c' = 0$. This gives us the much
+smaller automaton of \cref{fig:example-simplify-3}.
+
+% \begin{figure}[t]
+%   \centering 
+%     \includegraphics[width=0.70\linewidth]{aca_or_bc_simplified_2}
+%     \caption{$\AcaOrBc{}$ after removing the now unused b-transition.}\label{fig:example-simplify-2}
+%     \Description[$\AcaOrBc$ without the returning transition from $B$ to $S$]{$\AcaOrBc$ without the returning transition from $B$ to $S$, which now makes the transition from $S$ to $B$ also have the same associated variable $l_c'$.}
+%   \end{figure}
+
+  \begin{figure}[t]
     \centering 
-    \includegraphics[width=0.75\linewidth]{aca_or_bc_simplified_2}
-    \caption{$\AcaOrBc{}$ after removing the now unused b-transition.}\label{fig:example-simplify-2}
-  \end{figure}
-
-  Plugging in the same inequality again on the same principle, we get $2(1 -
-  l_c') > l_c' \iff 1 > l_c'$ since we are working with integers, and
-  therefore that $l_c' = 0$. This gives us the even smaller automaton of
-  Figure~\ref{fig:example-simplify-3}.
-
-  \begin{figure}[ht]
-    \centering 
-      \includegraphics[width=0.75\linewidth]{aca_or_bc_simplified_3}
+      \includegraphics[width=0.7\linewidth]{aca_or_bc_simplified_3}
       \caption{$\AcaOrBc{}$ after further constraint propagation.}\label{fig:example-simplify-3}
+      \Description[$\AcaOrBc$ without the lower section.]{$\AcaOrBc$ with only its upper section left and the $a$ transitions now set to static 1.}
     \end{figure}
@@
 same shape, where we essentially only need to expand the self-loops of
-$\SomethingCSomething$ and reduce its range labels to only capture a:s, as seen
-in Figure~\ref{fig:final-example}. Note that we got this result by only
-performing linear inequality reasoning on the automata of the product versus the
-number of times each transition would be used.
-
-\begin{figure}[ht]
+$\SomethingCSomething$ and reduce its range labels to only capture a's, as seen
+in \cref{fig:final-example}. Note that we got this result by only performing
+linear inequality reasoning on the automata of the product versus the number of
+times each transition would be used.
+
+\begin{figure}[t]
 \centering
@@
     introduced for the transitions (for most transitions the same one, since
-    they are all linearly dependent).}\label{fig:final-example}
+    they are all linearly dependent).}\label{fig:final-example} \Description[The
+    final product, still a boring stick automaton with only one loop for an
+    arbitrary number of c's.]{A product of the intermittent automata, which
+    accepts \lstinline{ac.*ca}}
   \end{figure}
 
-  \begin{figure}[ht]
-\centering
-    \includegraphics[width=0.75\linewidth]{original_product}
-    \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ as it would have
-    appeared if we had computed it from the initial
-    automata.}\label{fig:original-product}
-  \end{figure}
-
-To obtain the final Parikh image, we need to tie the inequalities for the
+%   \begin{figure}[t]
+% \centering
+%     \includegraphics[width=0.75\linewidth]{original_product}
+%     \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ as it would have
+%     appeared if we had computed it from the initial
+%     automata.}\label{fig:original-product}
+%   \end{figure}
+
+To obtain the final Parikh image, we again need to tie the inequalities for the
 transitions of the automata of the product to the new transition counters we
@@
 existence-elimination using reasoning on lower bounds (e.g. $r_2 \geq 0$) we
-arrive at $\TransitionCount{c} \geq 1$. For comparison, the $\AcaOrBc{} \times
-\SomethingCSomething{}$ we would have obtained if we had eagerly computed the
-product can be seen in Figure~\ref{fig:original-product}. This would have
-introduced a significantly more complicated existentially quantified formula.
+arrive at $\TransitionCount{c} \geq 1$.
 

### 2022-07-06 main.tex
Work out the encoding

@@
 
+Many other generalisations of the Parikh than the ones we use here have been
+studied. Prominent examples include generalising the Parikh map to segments of a
+fixed length \cite{KARHUMAKI1980155} and the more general Parikh matrix, which
+gives more information about a word than the standard Parikh image
+\cite{parikh-matrix}. Another notable generalisation is the p-vector, introduced
+in~\cite{infinite-words}, which denotes the position of each letter in the word
+rather than the number of their occurrences and allows for generalisations into
+infinite alphabets. All of these in some sense extend the Parikh map. By
+contrast, the main utility of the formulation introduced here is that it allows
+us to \emph{remove} something, thereby potentialy obtaining an easier problem.
+
 \section{A demonstration of our approach}\label{sec:motivation}
@@
 constraints, which we want to solve for all possible values of $\MonoidElement$,
-a value in some commutative monoid \Fudge{MONOID} and the string $s$, where
+a value in some commutative monoid and the string $s$, where
 $\varphi(\MonoidElement)$ is some set of constraints on $\MonoidElement$:
@@
 
-\contents{\item Monoid
-\item Alfabet
-\item ord oever alfabet
-\item konkatenering
-}
-
-\subsection{Finite-state Automata and their Products}
-We define a non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$
-as $\Automaton = \AutomatonTuple$ where $\Transitions = \States \times \Alphabet
-\times \States$, $\States$ is its states, $\InitialState$ is w.l.o.g. assumed to
-be the single initial state, and $\AcceptingStates$ is its set of accepting
-states.  We describe a transition $\Transition$ from state $\State$ to state
-$\State'$ with label $\Label \in \Alphabet$ as $\Transition =
-\FromLabelTo{\State}{\Label}{\State'}$. \Fudge{When one or more of the states and
-labels of a transition are uninteresting we will omit them}.
+\subsection{Monoids}
+
+We write commutative monoids as $\Monoid =
+\left(X;+_{\Monoid};0_{\Monoid}\right)$, where we mean $X$ to be the underlying
+set of elements, $+_{\Monoid}$ to be the (commutative) monoid map, with
+$0_{\Monoid}$ as its neutral element.
+
+\subsection{Languages, Finite-state Automata and their Products}
+
+We define an alphabet as a set of symbols $\Alphabet$ with words $\Strings$, and
+the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
+
+A non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$ is
+$\AutomatonTuple$ where $\Transitions = \States \times \Alphabet \times
+\States$, $\States$ is its states, $\InitialState$ is w.l.o.g. assumed to be the
+single initial state, and $\AcceptingStates$ is its set of accepting states.  We
+write a transition $\Transition = \Tuple{\State, \Label, \State'} \in
+\Transitions$ as $\Transition = \FromLabelTo{\State}{\Label}{\State'}$.
+Similarly, we use the notation $\FromLabelTo{\State}{}{}$ to refer to the set of
+transitions starting in $\State$, and $\FromLabelTo{}{}{\State}$ to refer to the
+set of transitions coming into $\State$, whenever the automaton is clear from
+the context.
 
@@
 example, for the Parikh vector above, we would have $\CountOf{a} = 1$. You have
-already seen this in Section~\ref{sec:introduction:parikh}.
-
-Parikh's theorem states that any context-free language has a
-letter-equivalent regular language (c.f.~\cite{construction} for a construction of
-such automata from context-free grammars and~\cite{bounds} for bounds on its size). However, there are languages that are
-not context-free that also have semilinear images under~$\ParikhMap$ (e.g.
-$\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) = \ParikhMap((abc)^*) = \CountOf{a} =
-\CountOf{b} = \CountOf{c} \land \CountOf{a} \geq 0$).
-
-\contents{\item this means we can represent this as Presburger
-\item only interested in regular languages}
+already seen this in \cref{sec:introduction:parikh}.
+
+Parikh's theorem states that any context-free language has a letter-equivalent
+regular language (c.f.~\cite{construction} for a construction of such automata
+from context-free grammars and~\cite{bounds} for bounds on its size). However,
+there are languages that are not context-free that also have semilinear images
+under~$\ParikhMap$ (e.g. $\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) =
+\ParikhMap((abc)^*) = \CountOf{a} = \CountOf{b} = \CountOf{c} \land \CountOf{a}
+\geq 0$). This means they can be represented as a quantifier-free Presburger
+formula.
+
+Note that while Parikh's theorem applies to context-free languages, in this
+paper we are only interested in regular languages.
 
@@
   \item $\Map(\epsilon) = 0_{\Monoid}$
-  \item $\Map(a \Concat b) = \Map(a) +_{\Monoid} \Map(b)$, where $\Concat$ is
-  the string concatenation operation
+  \item $\Map(a \Concat b) = \Map(a) +_{\Monoid} \Map(b)$
 \end{itemize}
 
-Many other generalisations of the Parikh map have been studied. Prominent
-examples include generalising the Parikh map to segments of a fixed length
-\cite{KARHUMAKI1980155} and the more general Parikh matrix, which gives more
-information about a word than the standard Parikh image \cite{parikh-matrix}.
-Another notable generalisation is the p-vector, introduced
-in~\cite{infinite-words}, which denotes the position of each letter in the word
-rather than the number of their occurrences and allows for generalisations into
-infinite alphabets. All of these in some sense extend the Parikh map. By
-contrast, the main utility of the formulation introduced here is that it allows
-us to \emph{remove} something, thereby potentialy obtaining an easier problem.
-
 \subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
 
-Since Parikh images are semilinear, they can be fully expressed by Presburger
-arithmetic. This means that any Parikh image can be written as a set of linear
-equation. Following~\cite{generate-parikh-image}, we define the Parikh Image of
-a regular language recognised by a DFA $\Automaton = \AutomatonTuple$ as:
-
-\contents{\item replace implies with singlearrow
-\item splitta sista implikationerna i tvaa!
-\item intuition for what the variables mean}
-
+Since Parikh images are semilinear, any Parikh image can be written as a set of linear
+equations. The following construction, here adapted to work on an NFA $\Automaton = \AutomatonTuple$, was presented in\cite{generate-parikh-image}:
 \begin{equation}
@@
 }\\
-& \AndComp{\State \in \AcceptingStates}{\FinalStateVar_{\State} > 0 \implies \StateVar_{\State} > 0}\\
+& \AndComp{\State \in \AcceptingStates}{\FinalStateVar_{\State} > 0 \longrightarrow  \StateVar_{\State} > 0}\\
 & \AndComp{\State \in \States}{
@@
 }\\
-& \AndComp{\Transition = \FromLabelTo{}{}{\State} \in \Transitions}{
-  \TransitionVar_{\Transition} > 0 \implies \StateVar_{\State} > 0
+& \AndComp{\Transition = \FromLabelTo{\State'}{}{\State} \in \Transitions}{
+  \TransitionVar_{\Transition} > 0 \longrightarrow \StateVar_{\State} > 0
 } \\
-& \AndComp{\State \in \States}{
-  \StateVar_{\State} > 0 \implies
+& \AndComp{\State \in F}{
+  \StateVar_{\State} > 0 \longrightarrow \FinalStateVar_{\State} = 1
+} \\
+&\AndComp{\State \in \States}{
+  \StateVar_{\State} > 0 \longrightarrow
   \left(\FinalStateVar_{\State} = 1 \text{ if } \State \in \AcceptingStates \land \right) \OrComp{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions}{
@@
 
-where all variables $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ are
+All variables $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ are
 existentially quantified and the free variables $\LetterVar_\Letter$ make up the
-actual image. Performing standard quantifier elimination on $\TransitionVar_i,
-\StateVar_i, \FinalStateVar_i$ with $\LetterVar_\Letter$ would produce the
-Parikh image in its Presburger form. Note that compared to the method described
-in~\cite{generate-parikh-image}, ours is backwards. Intuitively the approaches
-are equivalent since the Parikh image, by definition, disregards order. The
-approach can be described as either following paths through the automaton
-forwards from an (in our case, \emph{the}) initial state, or conversely as in
-the case here, choose an accepting state and trace a path backwards to the
-initial state.
+image. $\StateVar_\State$ represents the distance of state $\State$ from
+$\InitialState$ in a spanning tree, $\TransitionVar_\Transition$ how many times
+a transition $\Transition$ is used, and $\FinalStateVar_\State$ whether a given
+accepting state $\State \in \AcceptingStates$ is the actually used final state.
 
 In this paper we refer to this model as the baseline approach, though we also
-apply some optimisations that allow us to detect unsatisfiability early,
-described in Section~\ref{sec:implementing-baseline}. Notably, the calculus
-introduced in this paper can at its core be described as lazily enforcing the
-connectedness constraint of this method (the final two clauses) while also
-interleaving the (similarly lazy) computation of products of automata.
+apply optimisations as described in \cref{sec:implementing-baseline}. The
+calculus introduced in this paper, by contrast, lazily enforces the
+connectedness constraint of this encoding (the final three clauses) while also
+interleaving the computation of products of automata and propagating information
+between the steps to reduce the amount of work that needs to be done.
 
@@
 
-Note that one of the $\lor$-clauses can be immediately removed since it
-contains the contradiction $\StateVar_{1} = \StateVar_{1} + 1$, arising from
-cycles. These clauses encode the distance in a spanning tree from an accepting
-state, and the unsatisfiability is by design, essentially ensuring that two
-states in a cycle cannot vouch for each other's connectedness.
+Note that one of the $\lor$-clauses can be immediately removed since it contains
+the contradiction $\StateVar_{1} = \StateVar_{1} + 1$, arising from cycles.The
+unsatisfiability is by design, and ensures that two states in a cycle cannot
+vouch for each other's connectedness.
 
@@
 facts, we get:
-
 \begin{equation}

### 2022-07-06 main.tex
Continue fixing red

@@
 
-\subsubsection{Example~1: String length}
+\subsection{Example~1: String length}
 \Fudge{We have already illustrated refer intro}
@@
 
-Intuitively, this produces a simpler problem than computing the Parikh image, as
-we now ignore not just the order of characters but also their differences. This
-means that the branching automaton of \cref{fig:len-branch} becomes
-equivalent to the non-branching automaton of \cref{fig:len-branch-free}.
-Note that the elimination of the choice of paths also effectively eliminates the
-possibility of disconnecting one of the self-loops. This, in turn, means that we
-can eliminate all of the connectedness implications of the quantified formula,
-which would in turn simplify it (and the quantifier elimination of it)
-considerably.
-
-\subsubsection{Example~2: Parikh Automata}\label{sec:parikh-automata}
+Intuitively, this often produces a simpler problem than computing the Parikh
+image, as we now ignore not just the order of characters but also their
+differences.
+
+\subsection{Example~2: Parikh Automata}\label{sec:parikh-automata}
 \contents{\item Ostrich+!}
@@
 \begin{definition}\label{def:parikh-automata} A Parikh automaton of dimension $d
-  \geq 0$ is a pair $\Tuple{\Automaton, C}$, where $C \subseteq \Naturals^d$ is a semi-linear set (or, equivalently, a Presburger formula), where $\Automaton$ is a finite automaton with the alphabet $\Alphabet \times D$, where $D \subseteq \Naturals^d$.
-
-  We say that $\Tuple{\Automaton, \varphi}$ \emph{recognises} the language, $\Language(\Automaton, C)$, that is it recognises the language crated by is the $\pi_\Alphabet$-projection on $\Alphabet$ whose $\pi_{v} \subseteq C$.
+  \geq 0$ is a pair $\Tuple{\Automaton, C}$, where $C \subseteq \Naturals^d$ is
+  a semi-linear set (or, equivalently, a Presburger formula), and $\Automaton$
+  is a finite automaton with the alphabet $\Alphabet \times D$, where $D
+  \subseteq \Naturals^d$. We say that $\Tuple{\Automaton, \varphi}$
+  \emph{recognises} the language, $\Language(\Automaton, C)$, that is it
+  recognises the language created by is the $\pi_\Alphabet$-projection on
+  $\Alphabet$ whose $\pi_{v} \subseteq C$.
 \end{definition}
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \centering
-      \includegraphics[width=0.75\linewidth]{parikh_automaton}
-      \caption{A The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
+      \includegraphics[width=0.5\linewidth]{parikh_automaton}
+      \caption{The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
       $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
@@
 
-We assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
-transitions $\Transitions =
-\Set{\EllipsisSequence{\Transitions}{\NrTransitions}}$. For convenience, we then introduce the following supporting notations:
+We start by defining our calculus, \Calculus{}, for one automaton, and only
+extend it to products of automata in \cref{sec:multiple}. Assume an NFA
+$\Automaton = \AutomatonTuple$ with $\NrTransitions$ transitions $\Transitions =
+\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. For convenience, we then
+introduce the following supporting notations:
 
@@
   transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
-  \State_{k+1})$ holds for every $k$), passing zero or more labels $\Label_1,
-  \ldots, \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
-  \InitialState$. However, the ending state, $\State_n$, is not necessarily an
-  accepting state.
+  \State_{k+1})$ holds), passing zero or more labels $\Label_1, \ldots,
+  \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
+  \InitialState$. The end state, $\State_n$, is not necessarily accepting. If it
+  is, we say the path is also \textit{accepting}.
   \end{definition}
@@
 \begin{definition}
-  Moreover, we also talk about the \textit{set of paths} of an automaton,
+  Moreover, we talk about the \textit{set of paths} of an automaton,
   $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
@@
 \begin{definition}
-  The \textit{word} of a path $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat} \Label_n$ is
-  the word read out on its labels.
+  For a path $\Path = \Tuple{\State_0 \Label_1 \State_1 \RepeatSum{,}
+  \State_n}$, its \textit{word} $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat}
+  \Label_n$ is the word read out on its labels.
 \end{definition}
@@
 \begin{definition}
-  \Fudge{fixme: use pattern matching on path definition}
-  The \textit{states} of a path $\StatesOf(\Path) = \Set{\State \SuchThat
-  \FromLabelTo{\State}{}{\State'} \in \Path \text{ or }
-  \FromLabelTo{\State'}{}{\State} \in \Path}$ are the states visited along a
-  path. Note that $\InitialState \in \StatesOf(\Path)$ for every path since all
-  paths start in the initial state.
+  The \textit{states} of a path $\Path = \Tuple{\State_0 \Label_1 \State_1
+  \RepeatSum{,} \State_n}$, $\StatesOf(\Path) = \Set{\State_0\RepeatSum{,}
+  \State_n}$ are the states visited along $\Path$. Note that $\InitialState \in
+  \StatesOf(\Path)$ for every path since all paths start in the initial state.
 \end{definition}
@@
 \begin{definition}
-  \Fudge{if every valid path containing q overlaps with C}
-  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$,
-  $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of transitions whose
-  removal from~$\Transitions$ would cause~$\State \in \States$ to be unreachable
-  from the initial state~$\InitialState$, or cut it off from reaching any
-  accepting state. $\SeparatingCut(\Automaton, \InitialState)$ is crucially
-  undefined if $\Automaton$ accepts any string since it cannot be made unreachable
-  from itself by definition and must have at least onepath to an accepting state
-  if it accepts any string.
+  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$ to state
+  $\State$, written $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of
+  transitions such that every accepting path $p$ where $\State \in
+  \StatesOf(\Path)$ contains a transition $\Transition \in C$. A cut does not
+  exist for every state; notably it never exists for $\InitialState$.
 \end{definition}
@@
 
-Finally, we present the rules of our calculus for one automaton in
+We present the rules of \Calculus{} for one automaton in
 \cref{tbl:rules:single}. Note that we operate on sets of symbols (terms,
@@
 ($\SomeClause$).
+
+Note that the filtering function~$\Filter$ is evaluated symbolically in these
+rules, and can in practice be read as a symbolic function from transitions to
+terms (e.g.~\texttt{t} or~\texttt{t+1}). In our implemendation \Catra{},
+described in \cref{sec:implementation}, $\Filter$~is a vector of fresh variables
+with the same size as~$\Transitions$.
 
@@
 instance, the basis for the matched linear inequalities is
-implicitly~$\SomeInequalities$.
-
-These rules are meant to be executed in an automated theorem prover, and are
-read bottom-up. Since a large part of our rules operate by adding and literally
-matching linear (in)equalities in a proof goal, we use the shorthand of listing
-the matched inequalities as antecedents. An example of this can be seen in
-the~\Propagate{} rule.
-
-Please note that the filtering function~$\Filter$ is evaluated symbolically in
-these rules, and can in practice be read as a symbolic function from transitions
-to terms (e.g.~\texttt{t} or~\texttt{t+1}), not dissimilar to~$\FlowEq$ below.
-In our implementation, $\Filter$~is a vector of fresh unknown constants with the
-same size as~$\Transitions$.
-
-We assume that every rule only fires when it would add a new clause. For
+implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
+\left(X;+_{\Monoid};0_{\Monoid}\right)$ and the map into it $\Map : \Strings
+\rightarrow X$, we also allow mapping over transitions:
+$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$.
+
+The rules of \cref{tbl:rules:single} are meant to be used in tableaux-style
+proofs, and are read bottom-up, where the syntax at the bottom is replaced with
+the one at the top until \Fudge{it is possible to close the proof goal}. Since
+our rules operate by adding and matching linear (in)equalities in a proof goal,
+we use the shorthand of listing the matched inequalities as antecedents. An
+example of this can be seen in the~\Propagate{} rule. \Fudge{We pull solutions
+out of a hat? I have no idea.}
+
+We require that every rule only fires when it would add a new clause. For
 example, this means that we cannot use~\Split{} to split on the same term twice.
@@
 
-\begin{table}[h]
+\begin{table}
 \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
@@
   {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
-  None \\[5ex]
+  None \\[3ex]
 
@@
   \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  if $\Transition \in \Transitions_\Automaton$ \\[5ex]
+  if $\Transition \in \Transitions_\Automaton$ \\[3ex]
 
@@
   \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
-  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[5ex]
+  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[3ex]
 
@@
 \end{table}
-
 We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
-of existentially quantified linear inequalities with the following definition:
-$$
+of existentially quantified linear inequalities with the following definition,
+where we assign fresh, existentially quantified variables to
+$\FinalStateVar_\State, \Filter(\Transition)$ for every $\State \in
+\AcceptingStates, \Transition \in \Transitions$:
+\[
 \begin{aligned}
-  & \FlowEq(\Automaton, \Filter) = \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \\
+  & \FlowEq(\Automaton, \Filter) = \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
+  \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}\\
   & \Sink(\State) = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
-  & \In(\State, \Filter) = \StartFlow(\State) + \sum_{\Transition = \FromLabelTo{\State_0}{}{\State} \in \Transitions} \Filter(\Transition)\\
+  & \In(\State, \Filter) = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
   & \StartFlow(\State)  = 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
-  & \Out(\State, \Filter) = \sum_{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions} \Filter(\Transition)
+  & \Out(\State, \Filter) = \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
 \end{aligned}
-$$
-Where we transparently assign fresh variables to each state $\State \in
-\States$ as well as every accepting state $\State' \in \AcceptingStates$ and
-existentially quantify them in the whole clause.
-
-Additionally, we assume the existence of a rule \PresburgerClose{},
-corresponding to a sound and complete solver for Presburger formulae, modulo the
-monoid~$\Monoid$.
+\]
+
+In addition to \cref{tbl:rules:single}, we assume the existence of a rule
+\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
+formulae, and for the elements of~$\Monoid$.
 
@@
 prover would start by expanding the predicate using the $\Expand{}$~rule. A
-modestly clever theorem prover would perform algebraic substitution on the
-underlying constants of~$\Filter$, boiling them down to choices of branches,
-which depend on one single variable, and loop transitions. This logic
-corresponds to the placement of counters for optimally edge-profiling the CFG of
-a program, making up a MST of the automaton~\cite{path-profiling}.
+theorem prover would perform algebraic substitution on the underlying constants
+of~$\Filter$, boiling them down to choices of branches, which depend on one
+single variable, and loop transitions. This logic corresponds to the placement
+of counters for optimally edge-profiling the CFG of a program, making up a
+minimum-spanning tree of the automaton~\cite{path-profiling}.
+
+In order to make the examples below tractable, we will assume the existence of a
+rule \EquationReasoning{} that allows us to perform standard algebraic reasoning
+on linear inequalities. This rule is not necessary for correctness or
+completeness, but shortens the examples considerably.
 
@@
 \begin{figure}[ht]
-  We define $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$ as follows:
-
-  \begin{minipage}[b]{0.45\linewidth}
+  We define $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$ as in the following two
+  equations, and then apply \EquationReasoning{} (under the implicit assumption
+  that every RHS is $\geq 0$) to obtain the equivalent third definition :
+  \begin{minipage}[b]{0.3\linewidth}
     \begin{equation*}
@@
   \hspace{0.5cm}
-  \begin{minipage}[b]{0.45\linewidth}
+  \begin{minipage}[b]{0.3\linewidth}
     \begin{equation*}
@@
   \end{minipage}
-  \caption{Equivalences defining $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$
-  respectively.}\label{fig:example:single:equivalences}
-  \end{figure}
-
-Propagating the equivalences from \FlowEq{} gives us (under the implicit
-assumption that every RHS is $\geq 0$):
+  \begin{minipage}[b]{0.3\linewidth}
     \begin{equation*}
@@
     \end{equation*}    
+  \end{minipage}
+  \caption{Equivalences defining $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$
+  respectively.}\label{fig:example:single:equivalences}
+  \end{figure}
 
@@
   }
-  \infer1[Eq. reason]{
+  \infer1[\EquationReasoning{}]{
     \begin{matrix}
@@
   \infer2[\Split{} $1 - \TransitionVar_6$]{ \Connected(\AcaOrBc{}, \Filter) \land 2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6 }
+  \infer1[\EquationReasoning{}]{
+    \Connected(\AcaOrBc{}, \Filter) \land
+    2k + 1 =
+    \sum\limits_{\Transition \in \Transitions} (\Transition \mapsto 1)(\Label) \cdot \Filter(\Transition)
+  }
   \infer1[\Expand{}]{\Image{}_{\AcaOrBc{}, \Transition \mapsto 1}(\Filter, 2k + 1)}
@@
 
-To be as clear as possible about what it means for our calculus to be correct,
-we use more exact definitions than the traditional soundness and completeness.
-Since correctness is guaranteed by the formulation of the predicate, our
-correctness proof means proving inductively that no rule would invalidate that
-definition.
+Since correctness is defined by the formulation of the $\Image{}$~predicate, our
+correctness proof amounts to proving inductively that no rule would invalidate
+that definition.
 
 \subsubsection{\Calculus{} terminates}
-\begin{lemma}\label{lma:single-terminates}
-  A solver implementing the calculus described in \cref{sec:calculus} starting
-  with some decidable (modulo the solver) set of clauses $\SomeClause{}$ and an
-  instance of our predicate $\SinglePredicateInstance$ will eventually
-  terminate.
+\begin{lemma}\label{lma:single-terminates} A solver implementing the calculus
+  described in \cref{sec:calculus} starting with some decidable set of clauses
+  $\SomeClause{}$ and an instance of our predicate $\SinglePredicateInstance$
+  will eventually terminate.
 \end{lemma}
@@
 It remains to prove that each rule is only applied finitely many times. This is
-almost obvious since both \Propagate{} and \Split{} can only be applied to
-transition variables whose $\Filter$ value has no known bound. However, \Split{}
-can always be used to split the proof tree and add clauses in precisely the
-right shape, thereby guaranteeing the ability to make forward progress.
-Therefore, the number of transition variables whose bounds are unknown is
-decreasing monotonically for any given proof goal, and hence termination is
-guaranteed.
+ obvious since both \Propagate{} and \Split{} can only be applied to transition
+ variables whose $\Filter$ value has no known lower bound. However, \Split{} can
+ always be used to split the proof tree and add clauses in precisely the right
+ shape, thereby guaranteeing the ability to make forward progress. Therefore,
+ the number of transition variables whose bounds are unknown is decreasing
+ monotonically for any given proof goal, and hence termination is guaranteed.
 \end{proof}
@@
 \subsubsection{\Calculus{} computes the homomorphic image}
+\Fudge{I am unsure of the definitions here}
 \begin{lemma}\label{lma:single-correct}

### 2022-07-06 main.tex
reformulated some of the text about generalised Pa images

--- 2022-07-06 main.tex
@@
 Note that while Parikh's theorem applies to context-free languages, in this
-paper we are only interested in regular languages.
-
-\subsection{Generalised Parikh maps}\label{sec:generalised}
-
-To generalise the logic of the Parikh map, we introduce a morphism $\Map:\: \Sigma^*
-\to \Monoid$ where $\Monoid = \left(X;+_{\Monoid};0_{\Monoid}\right)$ is a commutative monoid where the following holds:
-\begin{itemize}
-  \item $\Map(\epsilon) = 0_{\Monoid}$
-  \item $\Map(a \Concat b) = \Map(a) +_{\Monoid} \Map(b)$
-\end{itemize}
+paper we focus only on regular languages.
 
@@
 Since Parikh images are semilinear, any Parikh image can be written as a set of linear
-equations. The following construction, here adapted to work on an NFA $\Automaton = \AutomatonTuple$, was presented in\cite{generate-parikh-image}:
+equations. The following construction, here adapted to work on an NFA $\Automaton = \AutomatonTuple$, was presented in \cite{generate-parikh-image}:
 \begin{equation}
@@
 
-\section{Applications of generalised Parikh images}\label{sec:applications}
-
-\subsubsection{Example~1: String length}
-\Fudge{We have already illustrated refer intro}
-A useful example of such a simplifying morphism can express string length, the
-problem that originally motivated our study of the Parikh map. We let $\Monoid =
-\left(\mathbb{Z};+;0\right)$ and let $\Map(x) = 1$. Then the length of a string
-$s = s_1 \RepeatSum{\Concat}  s_n$ is given by $\Map(s) = \sum \Map(s_i) = 1
-\RepeatSum{+} 1$. Note that if we substitute vectors and vector addition for
-this monoid, we obtain the regular Parikh map.
+\section{Generalised Parikh Maps}\label{sec:generalised}
+
+It is easy to see that the Parikh map~$\ParikhMap$ represents a
+homomorphism from the (free) non-commutative monoid~$\Strings$ to the
+(free) commutative monoid~$\Naturals^k$. As we are often not
+interested in the full Parikh image, but only in some projections of
+it, for the purposes of this paper we use a generalised version of the
+Parikh map. We consider \emph{arbitrary} homomorphisms
+$\Map:\: \Sigma^* \to \Monoid$, where
+$\Monoid = \left(X;+_{\Monoid};0_{\Monoid}\right)$ is a commutative
+monoid; this means that $\Map(\epsilon) = 0_{\Monoid}$ and
+$\Map(u \Concat w) = \Map(u) +_{\Monoid} \Map(w)$. We give several
+examples of such generalised Parikh image later in this section.
+
+Observe that every homomorphism~$\Map:\: \Sigma^* \to \Monoid$ can be
+represented as the composition~$h' \circ \ParikhMap$, for some
+homomorphism~$h' : \Naturals^k \to M$. One of the insights underlying our
+approach is that it is more efficient to directly compute a
+generalised Parikh image~$\Map(\Language)$, than to first compute the
+standard image~$\ParikhMap(\Language)$ followed by projection to
+some property of interest.
+
+%\subsection{Applications of generalised Parikh images}\label{sec:applications}
+
+\subsection{Example~1: Reasoning about string length}
+
+A useful example of such a simplifying morphism can express string
+length, the problem that originally motivated our study of the Parikh
+map. We let $\Monoid = \left(\mathbb{Z};+;0\right)$, and define the
+homomorphism~$\Map$ by $\Map(a) = 1$ for all
+characters~$a \in \Alphabet$. Then the length of a string
+$s = s_1 \RepeatSum{\Concat} s_n$ is given by
+$\Map(s) = \sum \Map(s_i) = 1 \RepeatSum{+} 1 = n$.
 
@@
 
-\subsubsection{Example~2: Parikh Automata}\label{sec:parikh-automata}
+\subsection{Example~2: Parikh Automata}\label{sec:parikh-automata}
 \contents{\item Ostrich+!}

### 2022-07-06 main.tex

--- 2022-07-06 main.tex
@@
 
-\section{Applications of generalised Parikh images}\label{sec:applications}
-
-\subsection{Example~1: String length}
-\Fudge{We have already illustrated refer intro}
-A useful example of such a simplifying morphism can express string length, the
-problem that originally motivated our study of the Parikh map. We let $\Monoid =
-\left(\mathbb{Z};+;0\right)$ and let $\Map(x) = 1$. Then the length of a string
-$s = s_1 \RepeatSum{\Concat}  s_n$ is given by $\Map(s) = \sum \Map(s_i) = 1
-\RepeatSum{+} 1$. Note that if we substitute vectors and vector addition for
-this monoid, we obtain the regular Parikh map.
-
-Intuitively, this often produces a simpler problem than computing the Parikh
-image, as we now ignore not just the order of characters but also their
-differences.
+\section{Generalised Parikh Maps}\label{sec:generalised}
+
+It is easy to see that the Parikh map~$\ParikhMap$ represents a
+homomorphism from the (free) non-commutative monoid~$\Strings$ to the
+(free) commutative monoid~$\Naturals^k$. As we are often not
+interested in the full Parikh image, but only in some projections of
+it, for the purposes of this paper we use a generalised version of the
+Parikh map. We consider \emph{arbitrary} homomorphisms
+$\Map:\: \Sigma^* \to \Monoid$, where
+$\Monoid = \left(X;+_{\Monoid};0_{\Monoid}\right)$ is a commutative
+monoid; this means that $\Map(\epsilon) = 0_{\Monoid}$ and
+$\Map(u \Concat w) = \Map(u) +_{\Monoid} \Map(w)$. We give several
+examples of such generalised Parikh image later in this section.
+
+Observe that every homomorphism~$\Map:\: \Sigma^* \to \Monoid$ can be
+represented as the composition~$h' \circ \ParikhMap$, for some
+homomorphism~$h' : \Naturals^k \to M$. One of the insights underlying our
+approach is that it is more efficient to directly compute a
+generalised Parikh image~$\Map(\Language)$, than to first compute the
+standard image~$\ParikhMap(\Language)$ followed by projection to
+some property of interest.
+
+%\subsection{Applications of generalised Parikh images}\label{sec:applications}
+
+\subsection{Example~1: Reasoning about string length}
+
+A useful example of such a simplifying morphism can express string
+length, the problem that originally motivated our study of the Parikh
+map. We let $\Monoid = \left(\mathbb{Z};+;0\right)$, and define the
+homomorphism~$\Map$ by $\Map(a) = 1$ for all
+characters~$a \in \Alphabet$. Then the length of a string
+$s = s_1 \RepeatSum{\Concat} s_n$ is given by
+$\Map(s) = \sum \Map(s_i) = 1 \RepeatSum{+} 1 = n$.
+
+Intuitively, this produces a simpler problem than computing the Parikh image, as
+we now ignore not just the order of characters but also their differences.
+\iffalse
+This
+means that the branching automaton of \cref{fig:len-branch} becomes
+equivalent to the non-branching automaton of \cref{fig:len-branch-free}.
+Note that the elimination of the choice of paths also effectively eliminates the
+possibility of disconnecting one of the self-loops. This, in turn, means that we
+can eliminate all of the connectedness implications of the quantified formula,
+which would in turn simplify it (and the quantifier elimination of it)
+considerably.
+\fi
 

### 2022-07-06 main.tex
string length

@@
 word length. To decide whether a simple formula like
-$x \in \mathcal{L}_1 \wedge y \in \mathcal{L}_2 \wedge |x| > |y|$,
+$x \in \Language_1 \wedge y \in \Language_2 \wedge |x| > |y|$,
 with string variables~$x, y$ and regular
-languages~$\mathcal{L}_1, \mathcal{L}_2$, is satisfiable, it is
+languages~$\Language_1, \Language_2$, is satisfiable, it is
 necessary to reason about the sets of word lengths induced by
-$\mathcal{L}_1, \mathcal{L}_2$, which is a special case of the Parikh
+$\Language_1, \Language_2$, which is a special case of the Parikh
 image.  This required combined reasoning about strings and string
@@
 
-\subsection{Example~1: Reasoning about string length}
+\subsection{Example~1: Reasoning about String Length}
 
@@
 length, the problem that originally motivated our study of the Parikh
-map. We let $\Monoid = \left(\mathbb{Z};+;0\right)$, and define the
-homomorphism~$\Map$ by $\Map(a) = 1$ for all
-characters~$a \in \Alphabet$. Then the length of a string
+map. This mapping is relevant when solving constraints that combine
+language membership as well as word length, for instance the formula
+\begin{equation}\label{eq:stringLength}
+x \in \Language_1 \wedge y \in \Language_2 \wedge |x| > |y|
+\end{equation}
+mentioned in the introduction. To solve this formula, let
+$\Monoid = \Naturals$, and define the homomorphism~$L$ by $L(a) = 1$
+for all characters~$a \in \Alphabet$. The length of a string
 $s = s_1 \RepeatSum{\Concat} s_n$ is given by
-$\Map(s) = \sum \Map(s_i) = 1 \RepeatSum{+} 1 = n$.
+$L(s) = \sum L(s_i) = 1 \RepeatSum{+} 1 = n$, and to solve
+\eqref{eq:stringLength} we can instead solve the equi-satisfiable
+formula~$\alpha \in L(\Language_1) \wedge \beta \in L(\Language_2)
+\wedge \alpha > \beta$. This paper proposes efficient native
+procedures to reason about membership constraints
+like~$\alpha \in L(\Language_1)$, avoiding the computation of
+the complete image~$L(\Language_1)$.
 
@@
 we now ignore not just the order of characters but also their differences.
-\iffalse
 This
@@
 considerably.
-\fi
-
-\subsection{Example~2: Parikh Automata}\label{sec:parikh-automata}
+
+\subsection{Example~2: Generalised String Constraints with Integer
+  Datatype}\label{sec:parikh-automata}
 \contents{\item Ostrich+!}

### 2022-07-06 main.tex
Examples

@@
 
+\Fudge{I don't understand.}
+  % \[
+  % \begin{aligned}
+  %   & \inferrule
+  %   {\Connected(\Automaton, \Filter) \land \FlowEq(\Automaton, \Filter) \land \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
+  %   {\SinglePredicateInstance, \SomeInequalities, \SomeClause} &\beta \models R_0 \iff \exists i : \beta \models R_i
+  % \end{aligned}
+  % \]
+
 \begin{lemma}\label{lma:expand-preserves}
@@
   Note that \Expand{} is taken directly from the existing formulation of
-  \cref{eq:generate-parikh}.
+  \eqref{eq:generate-parikh}.
 \end{proof}
@@
 $\SomethingCSomething$ under the constraint that there are more instances of
-letters a than b. We omit the definition of the map $\Map$ that maps each
-transition to the corresponding increment vector (e.g.
-\Map($\FromLabelTo{S}{a}{A}) = \begin{bmatrix}1\\0\\0\end{bmatrix}$),
-since it should be apparent by now.
-
-In the interest of space, we use a similar technique as in
-\cref{fig:example:single:equivalences} to combine the application of $\FlowEq$
-with the definition of $\Filter$.
-
-  We define $\Filter$ after substitutions from both initial applications of
-  $\FlowEq$ as follows, note the element-wise expansion for the $\Sigma$ labels:
-
-% Propagated equations
+letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
+element-wise addition, and $\Map$ that maps each transition to the corresponding
+increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.
+
+In the interest of space, we refer back to
+\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
+$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
+substitutions as follows. Note the expansion of the $\Sigma$ labels:
     \begin{equation*}
       \begin{aligned}
-        & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
+        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
+        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
+        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
+        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
+        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
+        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
         % Other automaton
-        & \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c} \\
+        & \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
         & \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        & \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c} \\
+        & \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
       \end{aligned}
@@
     \begin{matrix}
-      1 = \TransitionVar_{10} \\
-      \TransitionVar_3 = \TransitionVar_{11} \\
-      1 = \TransitionVar_{12} \\
-      1 = \TransitionVar_{11} \\
+      1 = \TransitionVar_{10} \land 
+      \TransitionVar_3 = \TransitionVar_{11} \land \\
+      1 = \TransitionVar_{12} \land 
+      1 = \TransitionVar_{11} \land 
       2 > 0
@@
     \begin{matrix}
-      1 = \TransitionVar_{10} \\
-      \TransitionVar_3 = \TransitionVar_{11} \\
-      1 = \TransitionVar_{12} \\
+      1 = \TransitionVar_{10} \land 
+      \TransitionVar_3 = \TransitionVar_{11} \land  \\
+      1 = \TransitionVar_{12} \land 
+      1 = \TransitionVar_{11} \land \\
+      \Image{}_{\Automaton', \Map}(\Filter, 
+      \VectorLiteral{2, 0, 1}) \land 
+      \Connected(\SomethingCSomething{}, \Filter) \land
+        2 > 0
+    \end{matrix}  
+  }
+  \infer1[\EquationReasoning{}]{
+    \begin{matrix}
+      1 = \TransitionVar_{10} \land
+      \TransitionVar_3 = \TransitionVar_{11} 
+      1 = \TransitionVar_{12} \land
       1 = \TransitionVar_{11} \\
       \Image{}_{\Automaton', \Map}(\Filter, 
-      \begin{bmatrix}
-        2 \\
-        0 \\
-        1
-        \end{bmatrix}) \land \\
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-        2 > 0
+      \VectorLiteral{
+        \TransitionVar_{7a} + \TransitionVar_{9a},
+        0,
+        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \\
+        \TransitionVar_6 + \TransitionVar_4 = 0 \land 
+        1 - \TransitionVar_6 > 0 \land
+        \TransitionVar_6 = 0 \land 
+        \TransitionVar_4 = 0 \land
+        \TransitionVar_3 > 0 \land
+        2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
+          \VectorLiteral{
+        \TransitionVar_{7a} + \TransitionVar_{9a},
+        0,
+        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
+        = \VectorLiteral{
+          2,
+          0,
+          \TransitionVar_3} \land \\
+      \Connected(\SomethingCSomething{}, \Filter) \land
+      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
     \end{matrix}  
   }
-  \infer1[Eq. reasoning]{
+  \infer1[\Materialise]{
     \begin{matrix}
-      1 = \TransitionVar_{10} \\
-      \TransitionVar_3 = \TransitionVar_{11} \\
-      1 = \TransitionVar_{12} \\
-      1 = \TransitionVar_{11} \\
-      \Image{}_{\Automaton', \Map}(\Filter, 
-      \begin{bmatrix}
-        a \\
-        0 \\
-        c
-        \end{bmatrix}) \land \\
-      \TransitionVar_6 = 0 \land \\
-      \TransitionVar_4 = 0 \land \\
+      \TransitionVar_6 + \TransitionVar_4 = 0 \land 
+      1 - \TransitionVar_6 > 0 \land
+      \TransitionVar_6 = 0 \land 
+      \TransitionVar_4 = 0 \land
+      \TransitionVar_3 > 0 \land
       2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      \begin{bmatrix}
-        a \\
-        b \\
-        c
-        \end{bmatrix} = \begin{bmatrix}
-          2  \\
-          0 \\
-          1
-        \end{bmatrix} \land
-        \\
-        \begin{bmatrix}
-          a \\
-          b \\
-          c
-          \end{bmatrix} = \begin{bmatrix}
-            \TransitionVar_{7a} + \TransitionVar_{9a} \\
-            0 \\
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          \end{bmatrix} \land \\
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-        a > 0
-    \end{matrix}  
-  }
-\end{prooftree}
-
-\begin{prooftree}
-  \infer0[\Materialise]{
-    \begin{matrix}
-      \TransitionVar_6 = 0 \\
-      \TransitionVar_4 = 0 \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \\
-      \begin{bmatrix}
-        a \\
-        b \\
-        c
-        \end{bmatrix} = \begin{bmatrix}
-          2  \\
-          0 \\
-          \TransitionVar_3 
-        \end{bmatrix}
-        \\
-        \begin{bmatrix}
-          a \\
-          b \\
-          c
-          \end{bmatrix} = \begin{bmatrix}
-            \TransitionVar_{7a} + \TransitionVar_{9a} \\
-            0 \\
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          \end{bmatrix} \\
+      \VectorLiteral{
+        \TransitionVar_{7a} + \TransitionVar_{9a},
+        0,
+        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
+        = \VectorLiteral{
+          2,
+          0,
+          \TransitionVar_3} \land \\
       \Connected(\SomethingCSomething{}, \Filter) \land \\
       \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
-      \begin{bmatrix}
-        a \\
-        0 \\
-        c
-        \end{bmatrix}) \land \\
-        a > 0
+      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
     \end{matrix}
@@
   \begin{matrix}
-    \TransitionVar_6 + \TransitionVar_4 = 0 \\
-    1 - \TransitionVar_6 > 0 \\
-    \TransitionVar_6 = 0 \\
-    \TransitionVar_4 = 0 \\
-    \TransitionVar_3 > 0 \\
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \\
-    \begin{bmatrix}
-      a \\
-      b \\
-      c
-      \end{bmatrix} = \begin{bmatrix}
-        2  \\
-        0 \\
-        \TransitionVar_3 
-      \end{bmatrix}
-      \\
-      \begin{bmatrix} % What to do about sigma!?
-        a \\
-        b \\
-        c
-        \end{bmatrix} = \begin{bmatrix}
-          \TransitionVar_{7a} + \TransitionVar_{9a} \\
-          0 \\
-          \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-        \end{bmatrix} \\
-    \Connected(\AcaOrBc{}, \Filter) \land \\
+    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
+    1 - \TransitionVar_6 > 0 \land
+    \TransitionVar_6 = 0 \land 
+    \TransitionVar_4 = 0 \land
+    \TransitionVar_3 > 0 \land
+    2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
+    \VectorLiteral{
+      \TransitionVar_{7a} + \TransitionVar_{9a},
+      0,
+      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
+      = \VectorLiteral{
+        2,
+        0,
+        \TransitionVar_3} \land \\
+    \Connected(\AcaOrBc{}, \Filter) \land 
     \Connected(\SomethingCSomething{}, \Filter) \land \\
     \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
-    \begin{bmatrix}
-      a \\
-      0 \\
-      c
-      \end{bmatrix}) \land \\
-      a > 0
+    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
   \end{matrix}
   }
-  \infer1[Eq. reasoning]{
+  \infer1[\EquationReasoning]{
     \begin{matrix}
-      \begin{bmatrix}
-        a \\
-        b \\
-        c
-        \end{bmatrix} = \begin{bmatrix}
-          2 - 2\TransitionVar_6 \\
-          2\TransitionVar_4 + 3\TransitionVar_6 \\
+      \VectorLiteral{a, b, c} = \VectorLiteral{
+          2 - 2\TransitionVar_6,
+          2\TransitionVar_4 + 3\TransitionVar_6,
           \TransitionVar_3 + \TransitionVar_6
-        \end{bmatrix}
+        } \land
         \\
-        \begin{bmatrix} % What to do about sigma!?
-          a \\
-          b \\
-          c
-          \end{bmatrix} = \begin{bmatrix}
-            \TransitionVar_{7a} + \TransitionVar_{9a} \\
-            \TransitionVar_{7b} + \TransitionVar_{9b} \\
+        \VectorLiteral{a, b, c} = \VectorLiteral{
+            \TransitionVar_{7a} + \TransitionVar_{9a},
+            \TransitionVar_{7b} + \TransitionVar_{9b},
             \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          \end{bmatrix} \\
-      \Connected(\AcaOrBc{}, \Filter) \land \\
+          } \land \\
+      \Connected(\AcaOrBc{}, \Filter) \land 
       \Connected(\SomethingCSomething{}, \Filter) \land \\
       \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
-      \begin{bmatrix}
-        a \\
-        b \\
-        c
-        \end{bmatrix}) \land \\
-        a > b
+      \VectorLiteral{a, b, c}) \land a > b
     \end{matrix}
   }
-  \infer1[\ExpandM]{\Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, \begin{bmatrix}
-    a \\
-    b \\
-    c
-    \end{bmatrix}) \land a > b}
+  \infer1[\ExpandM]{\Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
 \end{prooftree}

### 2022-07-06 main.tex
started rewriting Ostrich+ section

@@
 \begin{figure}[t]
-  \begin{minipage}[b]{0.50\linewidth}
+  \begin{minipage}[b]{0.70\linewidth}
   \centering 
-    \includegraphics[width=\textwidth]{aca_or_bc}
+  \includegraphics[scale=0.9]{aca_or_bc}
   \caption{An automaton recognising the regular expression
-    $\mathtt{ac^*a|b(bb)^*c}$.}\label{fig:aca-or-bc}
-
-    \includegraphics[width=\textwidth]{something_c_something}
+    $\mathtt{ac^*a|b(bb)^*c}$. \Fudge{this is actually not the right
+      regular expression. the automaton also accepts words in which a
+      b is followed by a.}}\label{fig:aca-or-bc}
+
+    \includegraphics[scale=0.9]{something_c_something}
     \caption{An automaton recognising the regular expression
@@
   Datatype}\label{sec:parikh-automata}
-\contents{\item Ostrich+!}
-
-Another application for this generalisation is to express Parikh automata, and
-this is indeed what the implementation described in
-Section~\ref{sec:implementation} does. Following~\cite{parikh-automata} as
-filtered through~\cite{expressiveness} (which offers a better formalisation for
-our use), we define a Parikh automaton as an automaton with integer counters
-that are incremented at the transitions, and where we have linear
-constraints on the final values of the counters after running the automaton, or
-formally as in \cref{def:parikh-automata,def:projections}.
+
+Parikh images are also applicable to decide more general classes of
+string constraints~\cite{ostrich-plus}. Consider, for instance, a
+constraint involving the substring operation:
+\begin{equation}
+  \label{eq:substring}
+  x \in \Language_1 \wedge \; 0 \leq n \leq m < |x| \wedge x[n:m] \in \Language_2
+\end{equation}
+in which $x$ is a string variable, $n, m$ are integer variables, and
+$x[n:m]$ denotes the substring of $x$ starting at position~$n$ running
+until position~$m$. To illustrate the decision procedure in
+\cite{ostrich-plus} for string constraints of this kind, suppose that
+$\Language_2$ is defined by the regular
+expression~$\mathtt{(ab|c)*}$. An automaton recognising this language
+is shown in Fig.~\ref{fig:parikh-automata} (left).
+
+We can model \eqref{eq:substring} using the notion of a \emph{Parikh
+  automaton}~\cite{parikh-automata,expressiveness}. A Parikh automaton
+is an automaton in which transitions are labelled both with characters
+from the alphabet~$\Alphabet$, and with offset vectors defining the
+increments of a finite number of counters.
 
@@
 
+\iffalse
 \begin{figure}[t]
@@
       visualised.}\label{fig:parikh-automaton}
+    \end{figure}
+\fi
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.9]{counter_automaton_pre}
+  \hfill
+  \includegraphics[scale=0.9]{counter_automaton}
+  \caption{An automaton recognising the language $\mathtt{(ab|c)*}$
+    (left), and a Parikh automaton describing the pre-image of the
+    substring operation (right).}\label{fig:parikh-automata}
     \end{figure}

### 2022-07-07 main.tex
skip duels

@@
   \label{eq:substring}
-  x \in \Language_1 \wedge \; 0 \leq n \leq m < |x| \wedge x[n:m] \in \Language_2
+  x \in \Language_1 \wedge \; 0 \leq n \leq m \leq |x| \wedge x[n:m] \in \Language_2
 \end{equation}
@@
 $x[n:m]$ denotes the substring of $x$ starting at position~$n$ running
-until position~$m$. To illustrate the decision procedure in
+until position~$m$. This constraint belongs to an expressive fragment
+of string logic that cannot be decided by most state-of-the-art string
+solvers. To illustrate the decision procedure proposed in
 \cite{ostrich-plus} for string constraints of this kind, suppose that
@@
 from the alphabet~$\Alphabet$, and with offset vectors defining the
-increments of a finite number of counters.
-
-In the following definitions, we use (as in~\cite{expressiveness}) the following notation:
-\begin{itemize}
-  \item $\Alphabet$ is an alphabet
-  \item $D$ is a finite subset of $\Naturals^d$, for some $d \in \Naturals^+$ (e.g. vectors of length $d$)
-\end{itemize}
-
-\begin{definition}\label{def:projections}
-  We define two projections on the composite alphabet $\left(\Alphabet \times D\right)^*$:
-  \begin{itemize}
-    \item $\pi_{\Alphabet} := (a, \Vector{v}) \mapsto a$ or the projection on $\Alphabet$ is the monoid morphism from $\left(\Alphabet \times D\right)^*$ to $\Alphabet^*$, and the opposite projection
-    \item $\pi_{v} := (a, \Vector{v}) \mapsto \Vector{v}$, called the
-    \emph{extended Parikh image}.
-  \end{itemize}
+increments of a finite number of counters. This means that Parikh
+automata recognise words over an extended
+alphabet~$\Alphabet \times D$, where $D \subseteq \Naturals^d$ is a
+finite set of increment vectors (notation as
+in~\cite{expressiveness}). We use the symbols~$\pi_\Alphabet, \pi_D$
+to denote projections to the first and the second component of a
+composite letter~$(a, \Vector{v})$, respectively, and extend those
+projections to words:
+\begin{equation*}
+  \pi_\Alphabet((a_1, \Vector{v}_1) \RepeatSum{\Concat} (a_k, \Vector{v}_k))
+  ~=~ a_1 \RepeatSum{\Concat} a_k,
+  \qquad
+  \pi_D((a_1, \Vector{v}_1) \RepeatSum{\Concat} (a_k, \Vector{v}_k))
+  ~=~ \Vector{v}_1 \RepeatSum{+} \Vector{v}_k.
+\end{equation*}
+
+\begin{definition}\label{def:parikh-automata} A Parikh automaton of dimension $d
+  \geq 0$ is a pair $\Tuple{\Automaton, C}$, where
+  $C \subseteq \Naturals^d$ is a semi-linear set (or, equivalently, a
+  Presburger formula), and $\Automaton$ is a finite automaton with the
+  alphabet $\Alphabet \times D$, where $D \subseteq \Naturals^d$. We
+  say that $\Tuple{\Automaton, \varphi}$ recognises a
+  word~$w \in \Alphabet^*$ if and only if the automaton has a run
+  accepting an extended word~$w' \in (\Alphabet \times D)^*$ such that
+  $\pi_\Alphabet(w') = w$ and $\pi_D(w') \in C$.
 \end{definition}
 
-\begin{definition}\label{def:parikh-automata} A Parikh automaton of dimension $d
-  \geq 0$ is a pair $\Tuple{\Automaton, C}$, where $C \subseteq \Naturals^d$ is
-  a semi-linear set (or, equivalently, a Presburger formula), and $\Automaton$
-  is a finite automaton with the alphabet $\Alphabet \times D$, where $D
-  \subseteq \Naturals^d$. We say that $\Tuple{\Automaton, \varphi}$
-  \emph{recognises} the language, $\Language(\Automaton, C)$, that is it
-  recognises the language created by is the $\pi_\Alphabet$-projection on
-  $\Alphabet$ whose $\pi_{v} \subseteq C$.
-\end{definition}
+The decision procedure in \cite{ostrich-plus} for formulas like
+\eqref{eq:substring} will construct a pre-image of $\Language_2$ under
+the substring operation~$\cdot[n:m]$, and check whether this pre-image
+is consistent with the constraint~$x \in \Language_1$. Because the
+substring operation depends on the values of the integer
+variables~$n, m$, it is appropriate to describe the pre-image as
+Parikh automaton of dimension~$2$, resulting in the automaton in
+Fig.~\ref{fig:parikh-automata} (right). The first component of the
+increment vectors is used to count the number of letters eliminated in
+the beginning of the string (value~$n$), while the second component
+records the beginning of the eliminated suffix (value~$m$).
+
+\Fudge{continue}
 
@@
   \centering
-  \includegraphics[scale=0.9]{counter_automaton_pre}
+  \raisebox{5ex}{\includegraphics[scale=0.9]{counter_automaton_pre}}
   \hfill
@@
   \caption{An automaton recognising the language $\mathtt{(ab|c)*}$
-    (left), and a Parikh automaton describing the pre-image of the
-    substring operation (right).}\label{fig:parikh-automata}
+    (left), and a Parikh automaton describing the pre-image of this
+    language under the substring operation~$\cdot[n:m]$ (right). For
+    sake of presentation, the Parikh automaton contains
+    $\epsilon$-transitions, which could be eliminated in the standard
+    way.}\label{fig:parikh-automata}
     \end{figure}
 
+\iffalse
 An example of a Parikh automata version of $\AcaOrBc{}$ can be seen in
@@
 \end{enumerate}
+\fi
 
@@
 
-\begin{figure}[ht]
-  \begin{minipage}[b]{0.75\linewidth}
-    \centering 
-      \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter}
-      \caption{Runtime duel: \Calculus{} versus baseline.}
-      \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter}
-      \caption{Runtime duel: \Calculus{} versus \Nuxmv.}
-    \end{minipage}  
-  \caption{The pairwise runtime of each instance for instances solved by any backend. Timeouts are reported as a runtime of \RuntimeTimeout.}
-  \label{fig:duels}
-\end{figure}
+% \begin{figure}[ht]
+%   \begin{minipage}[b]{0.75\linewidth}
+%     \centering 
+%       \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter}
+%       \caption{Runtime duel: \Calculus{} versus baseline.}
+%       \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter}
+%       \caption{Runtime duel: \Calculus{} versus \Nuxmv.}
+%     \end{minipage}  
+%   \caption{The pairwise runtime of each instance for instances solved by any backend. Timeouts are reported as a runtime of \RuntimeTimeout.}
+%   \label{fig:duels}
+% \end{figure}
 

### 2022-07-07 main.tex
more about Parikh automata
New graphs, figure placement, etc

@@
   \AndComp{}{\begin{aligned}
-  & \mathtt{ac^*a|b(bb)^*c} \text{ accepts } s\\
-  & \mathtt{.^*c.^*} \text{ accepts } s \\
+  & \mathtt{/}\SndRegex{}\mathtt{/} \text{ accepts } s\\
+  & \mathtt{/}\FstRegex{}\mathtt{/} \text{ accepts } s \\
   & \MonoidElement = f(s) \\
@@
 
-\begin{figure}[t]
-  \begin{minipage}[b]{0.70\linewidth}
+\begin{figure}[p]
   \centering 
-  \includegraphics[scale=0.9]{aca_or_bc}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{aca_or_bc}
     \caption{An automaton recognising the regular expression
-    $\mathtt{ac^*a|b(bb)^*c}$. \Fudge{this is actually not the right
-      regular expression. the automaton also accepts words in which a
-      b is followed by a.}}\label{fig:aca-or-bc}
-
-    \includegraphics[scale=0.9]{something_c_something}
+      $\FstRegex{}$.}\label{fig:aca-or-bc}
+      \Description[An NFA recognising $\FstRegex{}$]{An NFA recognising $\FstRegex{}$. It has 4 states.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth} 
+    \includegraphics[width=\textwidth]{something_c_something}
     \caption{An automaton recognising the regular expression
-    $\mathtt{.^*c.^*}$.}\label{fig:something-c-something}
-  \end{minipage}
-  \Description[Two deterministic finite automata, one above the other]{NFAs 
-    recognising the languages ac^*a|b(bb)^*c and ac^*a|b(bb)^*c respectively.
-    The first one has 4 states and the second one 2.}
+    $\SndRegex{}$.}\label{fig:something-c-something}
+    \Description[An NFA recognising $\SndRegex{}$]{An NFA recognising $\FstRegex{}$. It has 2 states.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{aca_or_bc_simplified}
+    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
+    simplification of their equations.}\label{fig:example-simplify-1}
+    \Description[The automaton $\AcaOrBc{}$ labelled with variables counting the number of times each transition is taken, after having simplified the equations.]{The starting state shows an initial flow of 1, with two outgoing transitions to $A$ ($1-l_c'$ on letter a, and $l_b' + l_c'$ on b respectively).
+$A$ has a flow of $l_c$, and its outgoing transition into the final state, $F$ is the same as the incoming one from $S$, the initial state.
+The $B$ transition has an outgoing transition labeled $c/ l_c'$ to the final state $F$, and one backwards transition $b / l_b'$ to the starting state.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{something_c_something_length_simplified}
+    \caption{$\SomethingCSomething{}$ after using linear reasoning over lengths to replace a variable.}\label{fig:example-length-reduced}
+    \Description[The automaton $\SomethingCSomething{}$ after reasoning over lengths.]{The same automaton as before for $\SomethingCSomething{}$, but with the transition from start to final annotated with a constant number 1, and the leftmost self-loop transition replaced with the equation $1 + l_c + 2l_b' -
+    r_\Sigma'$.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth}
+      \includegraphics[width=\textwidth]{aca_or_bc_length_split_1}
+      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ is not
+      used.}\label{fig:aca-or-bc-length-split-1}
+      \Description[The automaton \AcaOrBc{} where we have removed the transition
+      from $S$ to $A$.]{ The state $A$ hangs free without incoming transitions,
+      and only the lower part of the automaton remains.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth}
+      \includegraphics[width=\textwidth]{aca_or_bc_length_split_2}
+      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ \emph{is} used.}\label{fig:aca-or-bc-length-split-2}
+      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ when we keep the transition from $S$ to $A$.
+This removes the transition between $B$ and $F$, but otherwise maintains the automaton.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth}
+      \includegraphics[width=\textwidth]{length_split_product}
+      \caption{The product of $\AcaOrBc$ and $\SomethingCSomething$ after
+      case splitting and removal of dead states.}\label{fig:length-split-product}
+      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ with the transition from $B$ to $F$ removed. There now remains only one path to an accepting state.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth}
+      \includegraphics[width=\textwidth]{aca_or_bc_simplified_3}
+      \caption{$\AcaOrBc{}$ after further constraint propagation.}\label{fig:example-simplify-3}
+      \Description[$\AcaOrBc$ without the lower section.]{$\AcaOrBc$ with only its upper section left and the $a$ transitions now set to static 1.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.6\textwidth}
+        \includegraphics[width=\textwidth]{aca_or_bc_times_something_c_something}
+        \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ under the constraint
+        $\TransitionCount{a} > \TransitionCount{b}$.}\label{fig:final-example} \Description[The
+        final product, still a boring stick automaton with only one loop for an
+        arbitrary number of c's.]{A product of the intermittent automata, which
+        accepts \lstinline{ac.*ca}}
+  \end{subfigure}    
   \end{figure}
@@
 
-\begin{figure}[t]
-  \centering 
-    \includegraphics[width=0.50\linewidth]{aca_or_bc_simplified}
-    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
-    simplification of their equations.}\label{fig:example-simplify-1}
-    \Description[The automaton $\AcaOrBc{}$ labelled with variables counting the number of times each transition is taken, after having simplified the equations.]{The starting state shows an initial flow of 1, with two outgoing transitions to $A$ ($1-l_c'$ on letter a, and $l_b' + l_c'$ on b respectively).
-$A$ has a flow of $l_c$, and its outgoing transition into the final state, $F$ is the same as the incoming one from $S$, the initial state.
-The $B$ transition has an outgoing transition labeled $c/ l_c'$ to the final state $F$, and one backwards transition $b / l_b'$ to the starting state.}
-  \end{figure}
-
 We sum the individual transition variables to get the length $\MonoidElement$ we
@@
 
-\begin{figure}[t]
-  \centering 
-    \includegraphics[width=0.4\linewidth]{something_c_something_length_simplified}
-    \caption{$\SomethingCSomething{}$ after using linear reasoning over lengths to replace a variable.}\label{fig:example-length-reduced}
-    \Description[The automaton $\SomethingCSomething{}$ after reasoning over lengths.]{The same automaton as before for $\SomethingCSomething{}$, but with the transition from start to final annotated with a constant number 1, and the leftmost self-loop transition replaced with the equation $1 + l_c + 2l_b' -
-    r_\Sigma'$.}
-  \end{figure}
-
 Still, this approximate view allows us to perform further reasoning. Since we
@@
 
-\begin{figure}[t]
-  \centering 
-    \includegraphics[width=0.75\linewidth]{aca_or_bc_length_split_1}
-    \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ is not
-    used.}\label{fig:aca-or-bc-length-split-1}
-    \Description[The automaton \AcaOrBc{} where we have removed the transition
-    from $S$ to $A$.]{ The state $A$ hangs free without incoming transitions,
-    and only the lower part of the automaton remains.}
-  \end{figure}
-
 We immediately notice that the self-loop of state $A$ is now disconnected from
@@
 
-  \begin{figure}[t]
-    \centering 
-      \includegraphics[width=0.5\linewidth]{aca_or_bc_length_split_2}
-      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ \emph{is} used.}\label{fig:aca-or-bc-length-split-2}
-      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ when we keep the transition from $S$ to $A$.
-This removes the transition between $B$ and $F$, but otherwise maintains the automaton.}
-    \end{figure}
-
 Plugging in the assumption $1-l_c' > 0$ (and its consequence that $l_c' = 0$),
@@
 and we can therefore stop worrying about enforcing connectivity at all.
-
-\begin{figure}[t]
-  \centering 
-    \includegraphics[width=0.5\linewidth]{length_split_product}
-    \caption{The product of $\AcaOrBc$ and $\SomethingCSomething$ after
-    case splitting and removal of dead states.}\label{fig:length-split-product}
-    \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ with the transition from $B$ to $F$ removed. There now remains only one path to an accepting state.}
-  \end{figure}
 
@@
 
-  \begin{figure}[t]
-    \centering 
-      \includegraphics[width=0.7\linewidth]{aca_or_bc_simplified_3}
-      \caption{$\AcaOrBc{}$ after further constraint propagation.}\label{fig:example-simplify-3}
-      \Description[$\AcaOrBc$ without the lower section.]{$\AcaOrBc$ with only its upper section left and the $a$ transitions now set to static 1.}
-    \end{figure}
-
 This will produce a straightforward product since both automata now have the
@@
 times each transition would be used.
-
-\begin{figure}[t]
-\centering
-    \includegraphics[width=0.75\linewidth]{aca_or_bc_times_something_c_something}
-    \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ under the constraint
-    $\TransitionCount{a} > \TransitionCount{b}$. Note the fresh variables
-    introduced for the transitions (for most transitions the same one, since
-    they are all linearly dependent).}\label{fig:final-example} \Description[The
-    final product, still a boring stick automaton with only one loop for an
-    arbitrary number of c's.]{A product of the intermittent automata, which
-    accepts \lstinline{ac.*ca}}
-  \end{figure}
 
@@
 
-\subsubsection{An Example}
-We apply \eqref{eq:generate-parikh} to the automaton $\AcaOrBc{}$ of \cref{fig:aca-or-bc}.
-We then get the quantified Presburger formula seen in \eqref{eq:parikh-formula}.
-As before, we omit the implicit existential quantifiers for brevity, along with the constranints requiring all variables to be nonnegative.
-We also omit the $\land$-clauses between each line.
-In other words, in this example, only the variables $\LetterVar_a, \LetterVar_b, \LetterVar_c, \LetterVar_d, \LetterVar_e$, and $\LetterVar_f$ are free, and corresponds to the counts of the respective subscript letter.
-
-\begin{equation}
-  \begin{aligned}
-  \ParikhMap(\Automaton):=\:
-  &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} + \TransitionVar_{\FromLabelTo{1}{a}{3}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{2}{b}{0}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} + \TransitionVar_{\FromLabelTo{2}{c}{3}}\\
-  &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
-  &\TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{0}{a}{1}} - \TransitionVar_{\FromLabelTo{2}{b}{0}} = 1 \land  \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
-  & \TransitionVar_{\FromLabelTo{2}{b}{0}} + \TransitionVar_{\FromLabelTo{2}{c}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3}  - \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{2}{c}{3}} = 0\\
-  &\TransitionVar_{\FromLabelTo{0}{a}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} > 0 \implies \StateVar_{2} > 0\\
-  &\TransitionVar_{\FromLabelTo{1}{c}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{1}{a}{3}} > 0 \implies \StateVar_{3} > 0\\
-  &\TransitionVar_{\FromLabelTo{2}{b}{0}} > 0 \implies \StateVar_{0} > 0 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} > 0 \implies \StateVar_{3} > 0\\
-  &\StateVar_{0} > 0 \implies \left(\StateVar_{0} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{0}{a}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{0} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} \geq 1 \land \StateVar_{2} \geq 1\right)\\
-  &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{1}{c}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{1} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{1}{a}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
-  &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{0} + 1 \land \TransitionVar_{\FromLabelTo{2}{b}{0}} \geq 1 \land \StateVar_{0} \geq 1\right) \lor \left(\StateVar_{2} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
-  &\StateVar_{3} > 0 \implies \FinalStateVar_{3} = 1
-  \end{aligned}
-  \label{eq:parikh-formula}
-  \end{equation}
-
-  We see here that the formula encodes the choice of path at the start of the
-  automaton ($\TransitionVar_{\FromLabelTo{0}{b}{2}} +
-  \TransitionVar_{\FromLabelTo{0}{a}{1}} -
-  \TransitionVar_{\FromLabelTo{2}{b}{0}} = 1$), and the dependency between that
-  choice and the following outgoing transition, e.g. in
-  $\TransitionVar_{\FromLabelTo{1}{a}{3}} -
-  \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0$. Here we only have one choice of
-  incoming flow; from $\FinalStateVar_3$, corresponding to the only accepting
-  state.
-
-Note that one of the $\lor$-clauses can be immediately removed since it contains
-the contradiction $\StateVar_{1} = \StateVar_{1} + 1$, arising from cycles.The
-unsatisfiability is by design, and ensures that two states in a cycle cannot
-vouch for each other's connectedness.
-
-Assume we are really interested in occurrences with the letter a, i.e. we solve
-for $\LetterVar_a > 0$. This implies that $\StateVar_{1} > 0 \land \StateVar_{3}
-> 0$ (essentially, that the state our transition of interest started in is
-reachable), in turn implying that $\FinalStateVar_3 = 1$. Plugging in these
-facts, we get:
-\begin{equation}
-  \begin{aligned}
-  &\LetterVar_{a} = 2 \land \LetterVar_{b} = 0 \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}}\\
-  & \StateVar_{3} > 0\\
-  &\StateVar_{0} > 0 \implies \StateVar_{0} = \StateVar_{3} + 2\\
-  \end{aligned}
-  \label{eq:parikh-formula-calc}
-  \end{equation}
-
-We can then proceed to eliminate the remaining existentially quantified
-variables to obtain the formula $\LetterVar_{a} = 2 \land \LetterVar_{b} = 0
-\land \LetterVar_{c} \geq 0$ without too much effort.
-
-  This example, in addition to showing how the encoding works, has hopefully
-  also illustrated one of the key insights behind the design of \Calculus: the
-  fact that even few additional constraints on the Parikh image can dramatically
-  reduce the difficulty of computing it.
-
 \section{Generalised Parikh Maps}\label{sec:generalised}
@@
 the complete image~$L(\Language_1)$.
-
-Intuitively, this produces a simpler problem than computing the Parikh image, as
-we now ignore not just the order of characters but also their differences.
-This
-means that the branching automaton of \cref{fig:len-branch} becomes
-equivalent to the non-branching automaton of \cref{fig:len-branch-free}.
-Note that the elimination of the choice of paths also effectively eliminates the
-possibility of disconnecting one of the self-loops. This, in turn, means that we
-can eliminate all of the connectedness implications of the quantified formula,
-which would in turn simplify it (and the quantifier elimination of it)
-considerably.
 
@@
     \end{figure}
-
-\iffalse
-An example of a Parikh automata version of $\AcaOrBc{}$ can be seen in
-\cref{fig:parikh-automaton}. A corresponding Presburger version of the
-constraint seen in \cref{sec:motivation} is $\psi(\Vec{v}) = v_1 > v_2$.
-
-Generally, we can encode the problem as follows:
-\begin{enumerate}
-  \item The morphism ($\Map$) we would use is $\pi_{v}$
-  \item The monoid operation is element-wise vector addition, which is trivially commutative.
-  \item \Fudge{We would just add the constraints on the final values!?}
-\end{enumerate}
-\fi
 
@@
 
-\subsection{Finding the Presburger representation of a homomorphic image}\label{sec:finding-the-image}
-
-To find the Presburger form of the homomorphic image efficiently, we adapt the
-quantifier elimination approach of~\cite{qe} to our problem domain. The core
-method is the same: we incrementally use \Calculus{} to find models,
-\Generalise{} the models we find into a quantifier-free Presburger formula with
-the $\MonoidElement$ as the only free variable (algorithm~\ref{alg:generalise}), add the
-negated formula as a constraint, and continue enumerating models until none
-remain. The disjunction of the generalised models we enumerated is now our
-image.
-
-\begin{algorithm}
-  \caption{$\Generalise{}(\Automaton, \Filter, a)$ will generalise a final product $\Automaton$ and a model $a$ under the homomorphism $\Map$.}\label{alg:generalise}
-  \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-  \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-  \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-
-  $\Automaton' \gets \Tuple{\Automaton_\States, \Automaton_{\InitialState}, \Automaton_\AcceptingStates, \Set{\Transition \in \Automaton_\Transitions \SuchThat a(\Filter(\Transition)) \neq 0}}$
-
-  \Fudge{// This notation might be too much but IT'S OBVIOUS DAMMIT}
-
-  \KwRet{\EliminateQuantifiers{$\Map(\ParikhMap(\Automaton'))$}}
-  \end{algorithm}
-
-  \begin{algorithm}
-    \DontPrintSemicolon
-    \caption{$\FindImage{}(\Automaton_1 \times \ldots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
-    \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-    \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-    \SetKwFunction{NewTheoremProver}{newTheoremProver}
-    \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-    \SetKwFunction{FreshVariable}{freshVariable}
-    \SetKwFunction{Assert}{assert}
-    \SetKwFunction{GetModel}{getModel}
-    \SetKwData{ImageVar}{image}
-
-$p \gets \NewTheoremProver{}$\;
-$\Filter(\Transition) := \FreshVariable{p}$ for every $\Transition \in \Transitions_{\Automaton}$\;
-$\MonoidElement \gets$ \FreshVariable{$p$}\;
-\Assert{$p, \exists \MonoidElement, \Filter(\Transition) \text{ for every } \Transition \in \Transitions_\Automaton \HoldsThat \ImagePredicate{\Automaton}{\Map}{\Filter}{\MonoidElement} \land \AndComp{\Transition \in \Transitions}{\Filter(\Transition) \geq 0}$}\;
-$\ImageVar \gets \bot$\;
-\While{$p$ has more models}{
-  $\Tuple{\Automaton, a} \gets \GetModel(p)$\;
-  $G \gets \Generalise{}(\Automaton, \Filter, a)$\;
-  $\ImageVar \gets G \lor \Image$\;
-  \Assert{$p, \lnot G$}\;
-  }
-    \KwRet{\ImageVar}
-    \end{algorithm}
-
-    The attentive reader will notice that the \GetModel{} function is
-    nonstandard in that it returns both the model and the automaton used to
-    generate it. Since we implement algorithm~\ref{alg:find-image} inside a theorem
-    prover under our control, we are able to glean such internal data structures.
-This should be considered fair game, since we are in practice implementing a specialised form of quantifier elimination, which is an internal affair.
-
-    The performance of this approach is evaluated in
-    \cref{sec:evaluation:finding-image}, where we benchmark our implementation
-    of it.
-
 \section{Parikh Images from Products of Automata}\label{sec:multiple}
@@
 
+\subsection{Finding the Presburger representation of a homomorphic image}\label{sec:finding-the-image}
+
+To find the Presburger form of the homomorphic image efficiently, we adapt the
+quantifier elimination approach of~\cite{qe} to our problem domain. The core
+method is the same: we incrementally use \Calculus{} to find models,
+\Generalise{} the models we find into a quantifier-free Presburger formula with
+the $\MonoidElement$ as the only free variable (algorithm~\ref{alg:generalise}), add the
+negated formula as a constraint, and continue enumerating models until none
+remain. The disjunction of the generalised models we enumerated is now our
+image.
+
+\begin{algorithm}
+  \caption{$\Generalise{}(\Automaton, \Filter, a)$ will generalise a final product $\Automaton$ and a model $a$ under the homomorphism $\Map$.}\label{alg:generalise}
+  \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
+  \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
+  \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
+
+  $\Automaton' \gets \Tuple{\Automaton_\States, \Automaton_{\InitialState}, \Automaton_\AcceptingStates, \Set{\Transition \in \Automaton_\Transitions \SuchThat a(\Filter(\Transition)) \neq 0}}$
+
+  \KwRet{\EliminateQuantifiers{$\Map(\ParikhMap(\Automaton'))$}}
+  \end{algorithm}
+
+  \begin{algorithm}
+    \DontPrintSemicolon
+    \caption{$\FindImage{}(\Automaton_1 \times \ldots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
+    \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
+    \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
+    \SetKwFunction{NewTheoremProver}{newTheoremProver}
+    \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
+    \SetKwFunction{FreshVariable}{freshVariable}
+    \SetKwFunction{Assert}{assert}
+    \SetKwFunction{GetModel}{getModel}
+    \SetKwData{ImageVar}{image}
+
+$p \gets \NewTheoremProver{}$\;
+$\Filter(\Transition) := \FreshVariable{p}$ for every $\Transition \in \Transitions_{\Automaton}$\;
+$\MonoidElement \gets$ \FreshVariable{$p$}\;
+\Assert{$p, \exists \MonoidElement, \Filter(\Transition) \text{ for every } \Transition \in \Transitions_\Automaton \HoldsThat \ImagePredicate{\Automaton}{\Map}{\Filter}{\MonoidElement} \land \AndComp{\Transition \in \Transitions}{\Filter(\Transition) \geq 0}$}\;
+$\ImageVar \gets \bot$\;
+\While{$p$ has more models}{
+  $\Tuple{\Automaton, a} \gets \GetModel(p)$\;
+  $G \gets \Generalise{}(\Automaton, \Filter, a)$\;
+  $\ImageVar \gets G \lor \Image$\;
+  \Assert{$p, \lnot G$}\;
+  }
+    \KwRet{\ImageVar}
+    \end{algorithm}
+
+    The \GetModel{} function is nonstandard in that it returns both the model
+    and its associated automaton. Since algorithm~\ref{alg:find-image} is
+    essentially a form of quantifier elimination and therefore an internal affair
+    to the theorem prover, this should be considered fair game.
+
+    A preliminary implementation of this approach is available as part
+    of~\Catra, but has been neither optimised nor fully tested.
+
 \section{Implementation}\label{sec:implementation}
@@
 assignments or unsatisfiable much like traditional~SAT- or SMT solvers would.
-Additionally, the baseline and our own back-end also support efficient
-generation of the Presburger formula describing the constraints of the input
-file using the method described in \cref{sec:finding-the-image}.
+Additionally, the baseline and our own back-end also support generation of the
+Presburger formula describing the constraints of the input file.
 
@@
 
-\begin{table}[ht]
+\begin{table}
   \centering
@@
 
-\begin{figure}[ht]
-  \includegraphics[width=0.75\linewidth]{graphs/\commit-by-solver.pdf}
+\begin{figure}[t]
+  \begin{minipage}[b]{0.49\linewidth}
+    \centering
+      \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
       \caption{The division of statuses per backend.}
       \label{fig:solve-division}
-\end{figure}
-
-\begin{figure}[ht]
-  \includegraphics[width=0.75\linewidth]{graphs/\commit-time-boxplot.pdf}
+      \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
+Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly more in total than lazy.}
+  \end{minipage}
+  \begin{minipage}[b]{0.49\linewidth}
+    \centering
+      \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
       \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
       \label{fig:runtime-boxplot}
+      \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline.
+The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy.
+On the other hand, lazy has  a lot of outliers.}
+  \end{minipage}
+  \vspace{0.5cm}
+  \begin{minipage}{\linewidth}
+    \centering
+    \includegraphics[width=0.75\textwidth]{graphs/\commit-cactus.pdf}
+    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
+    \label{fig:cactus}
+    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
+Lazy solves much more instances than nuxmv on shorter runtimes, but is overtaken by just before the 60-second timeout used in these experiments.}
+  \end{minipage}
 \end{figure}
@@
 
-\begin{figure}[ht]
-  \includegraphics[width=0.75\linewidth]{graphs/\commit-cactus.pdf}
-  \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
-  \label{fig:cactus}
-\end{figure}
-
+\iffalse
 \subsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
@@
 \end{figure}
+\fi
 

### 2022-07-07 main.tex
Fix dependency parse issue
Image descriptions, reference fixes

@@
   \centering
-  \raisebox{5ex}{\includegraphics[scale=0.9]{counter_automaton_pre}}
+  \raisebox{5ex}{
+    \includegraphics[scale=0.9]{counter_automaton_pre}
+    }
   \hfill
@@
     way.}\label{fig:parikh-automata}
+    \Description[An 2-state automaton (left) and a slightly larger automaton with counters (right).]{
+      The rightmost automaton is the starts with an initial state with a self-loop on any character that increments both registers by one, followed by an epsilon transition to a state with a self-loop on c that increments the lower register by one.
+From that state, the automaton has an epsilon-transition into the accepting state, as well as a transition on a into a state with only a transition back again, both incrementing the lower register by one.
+The accepting state has a self-loop on any character that does nothing to the registers.
+      }
     \end{figure}
@@
 
-\begin{lstlisting}[caption={An example input file for \Catra{} for the problem introduced in \cref{sec:introduction:motivation}, illustrating every major syntax element.
+\begin{lstlisting}[caption={An example input file for \Catra{} for the problem introduced in \cref{sec:motivation}, illustrating every major syntax element.
 From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]

### 2022-07-07 main.tex
Tweak intro, clean up bibliography

@@
    frequently need to compute the Parikh image of large products (intersections)
-   of automata, which in many operations is both prohibitively slow and
-   memory-intensive. We contribute a novel understanding of how Parikh maps can
-   be tackled as a constraint-solving problem to solve real-world constraints
-   stemming from functions on regular languages, most notably the length
-   constraint. Furthermore, we show how this formulation can be efficiently
-   implemented as a calculus, \Calculus{}, in an automated theorem prover
-   supporting Presburger logic. We also provide an efficient method for deriving
-   the Presburger formula of a given Parikh image, \Fudge{which is good for
-   something} based on techniques for quantifier elimination. We implement
-   \Calculus{} in a tool called \Catra{}, and evaluate it on constraints
-   produced by the \OstrichPlus{} string constraint solver when solving
-   \Fudge{standard string constraint benchmarks involving string lengths}. We
-   show that our solution strictly outperforms the standard eager approach
-   described in~\citeauthor{generate-parikh-image} as well as the
-   over-approximating method recently described
-   by~\citeauthor{approximate-parikh}, and for realistic timeouts for constraint
-   solving also the~\Nuxmv{} model checker.
+   of nondeterministic automata, which in many operations is both prohibitively
+   slow and memory-intensive. We contribute a novel understanding of how Parikh
+   maps can be tackled as a constraint solving problem to solve real-world
+   constraints stemming from functions on regular languages, most notably the
+   length constraint. Furthermore, we show how this formulation can be
+   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
+   prover supporting Presburger logic.  We implement \Calculus{} in a tool
+   called \Catra{}, and evaluate it on constraints produced by the
+   \OstrichPlus{} string constraint solver when solving \Fudge{standard string
+   constraint benchmarks involving string lengths}. We show that our solution
+   strictly outperforms the standard eager approach described
+   in~\citeauthor{generate-parikh-image} as well as the over-approximating
+   method recently described by~\citeauthor{approximate-parikh} by a wuide
+   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
+   model checker.
 \end{abstract}
@@
 footprint. Moreover, the method can be used iteratively to tackle smaller chunks
-of the product incrementally, thereby avoiding running out of memory at all.
+of the product incrementally, thereby decreasing the memory footprint.
 
@@
 
-In addition to the decision problem of determining whether an intersection of
-Parikh automata is empty and the satisfaction problem of finding a value from
-the intersection if it is not, \Catra{} is able to efficiently generate a
-quantifier-free Presburger formula representing an intersection of Parikh
-automata, something beyond the reach of general model checkers like~\Nuxmv.
-\Fudge{This is useful because the moon is made of cheese.}
-
 We implement \Calculus{} as a theory for the \Princess{} automated theorem
-prover in a tool that we call \Catra{}. \Catra also supports using the
+prover in the tool \Catra{}. \Catra also supports using the
 approximate method of \cite{approximate-parikh}, its fall-back variant adapted
@@
 Parikh image of products of Parikh automata.
-\item A specialised method for quantifier elimination that blends well with \Calculus{} to efficiently compute quantifier-free Presburger representations of a given Parikh image.
-\Fudge{Why is this good?}
 \item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future study.
@@
 \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-  Name & Rule & Prerequisites\\
+  Name & \text{Rule} & Prerequisites\\
   \midrule
@@
   {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
-  None \\[3ex]
+  None \\[4ex]
 
@@
   \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  if $\Transition \in \Transitions_\Automaton$ \\[3ex]
+  if $\Transition \in \Transitions_\Automaton$ \\[4ex]
 
@@
   \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
-  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[3ex]
+  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[4ex]
 
@@
 
-  \begin{table}[h]
+  \begin{table}[t]
     \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-      Name & Rule & Prerequisites\\
+      Name & \text{Rule} & Prerequisites\\
       \midrule
@@
       {\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
-      None \\
+      None \\[5ex]
       \Materialise &
@@
 \begin{figure}[t]
-  \begin{minipage}[b]{0.49\linewidth}
   \centering
+  \begin{subfigure}[b]{0.49\textwidth}
     \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
@@
 Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly more in total than lazy.}
-  \end{minipage}
-  \begin{minipage}[b]{0.49\linewidth}
-    \centering
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth}
     \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
@@
 On the other hand, lazy has  a lot of outliers.}
-  \end{minipage}
-  \vspace{0.5cm}
-  \begin{minipage}{\linewidth}
-    \centering
-    \includegraphics[width=0.75\textwidth]{graphs/\commit-cactus.pdf}
+  \end{subfigure}
+  \vfill
+  \begin{subfigure}[b]{0.75\textwidth}
+    \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
     \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
@@
 Lazy solves much more instances than nuxmv on shorter runtimes, but is overtaken by just before the 60-second timeout used in these experiments.}
-  \end{minipage}
+  \end{subfigure}
 \end{figure}

### 2022-07-07 main.tex
more ostrich+

@@
 
-The decision procedure in \cite{ostrich-plus} for formulas like
-\eqref{eq:substring} will construct a pre-image of $\Language_2$ under
-the substring operation~$\cdot[n:m]$, and check whether this pre-image
-is consistent with the constraint~$x \in \Language_1$. Because the
-substring operation depends on the values of the integer
-variables~$n, m$, it is appropriate to describe the pre-image as
-Parikh automaton of dimension~$2$, resulting in the automaton in
-Fig.~\ref{fig:parikh-automata} (right). The first component of the
-increment vectors is used to count the number of letters eliminated in
-the beginning of the string (value~$n$), while the second component
-records the beginning of the eliminated suffix (value~$m$).
-
-\Fudge{continue}
-
-\iffalse
-\begin{figure}[t]
-  \centering
-      \includegraphics[width=0.5\linewidth]{parikh_automaton}
-      \caption{The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
-      $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
-      formula containing the constraints on the final register values cannot be
-      visualised.}\label{fig:parikh-automaton}
-    \end{figure}
-\fi
-
 \begin{figure}[t]
@@
     \end{figure}
+
+    Applied to \eqref{eq:substring}, the decision procedure in
+\cite{ostrich-plus} will construct a pre-image of $\Language_2$ under
+the substring operation~$\cdot[n:m]$, and check whether this pre-image
+is consistent with the constraint~$x \in \Language_1$. Because the
+substring operation depends on the values of the integer
+variables~$n, m$, a Parikh automaton of dimension~$2$ is a suitable
+formalism to describe the pre-image, resulting
+in the automaton in Fig.~\ref{fig:parikh-automata} (right). The first
+component of the increment vectors is used to count the number of
+letters eliminated in the beginning of the string (value~$n$), while
+the second component records the beginning of the eliminated suffix
+(value~$m$).
+
+Denoting the language described by Fig.~\ref{fig:parikh-automata}
+(right) as $\Language_{pre}$, we can then replace \eqref{eq:substring}
+with an equi-satisfiable formula that no longer contains any explicit
+substring operation:
+%
+\begin{equation}
+  \label{eq:parikh-constraint}
+  p \in \Language_{pre} \wedge~
+  \pi_\Alphabet(p) \in \Language_1
+  \wedge~~ \pi_D(p) =
+  \begin{bmatrix}
+    n \\ m
+  \end{bmatrix}~~
+  \wedge 0 \leq n \leq m \leq |p|
+\end{equation}
+%
+To check the satisfiability of \eqref{eq:parikh-constraint}, we need a
+decision procedure that can process intersections of regular languages
+(in this case, of $\Language_{pre}$ and $\Language_1$, synchronising
+on $\Alphabet$), while imposing the side
+condition~$0 \leq n \leq m \leq |p|$ on the increment sum. In
+\cite{ostrich-plus}, this decision procedure turned out to be main
+bottleneck of the string solver, which was one of the motivations to
+develop the lazy algorithm proposed in this paper.
+
+\iffalse
+\begin{figure}[t]
+  \centering
+      \includegraphics[width=0.5\linewidth]{parikh_automaton}
+      \caption{The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
+      $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
+      formula containing the constraints on the final register values cannot be
+      visualised.}\label{fig:parikh-automaton}
+    \end{figure}
+\fi
 

### 2022-07-07 main.tex
termination proof

@@
 We present the rules of \Calculus{} for one automaton in
-\cref{tbl:rules:single}. Note that we operate on sets of symbols (terms,
-clauses). Additionally, we also use the convention of splitting the formulas
-into linear inequalities ($\SomeInequalities$) and any other clauses
-($\SomeClause$).
+\cref{tbl:rules:single}. Note that we operate on sets of
+formulas. Additionally, we also use the convention of splitting the
+formulas into linear equations and inequalities ($\SomeInequalities$)
+and other formulas ($\SomeClause$). \Fudge{need to say that we assume
+  that the predicates (Im, etc.)  only occur positively.}
 
@@
 
-We require that every rule only fires when it would add a new clause. For
-example, this means that we cannot use~\Split{} to split on the same term twice.
+We require that rules can only be applied when they add new formulas
+on every created branch (the notion of \emph{regularity} of a proof is
+required~\cite{Fitting96a}). For example, this means that \Split{} can
+only be applied to proof goals that contain neither
+$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
+never be applied to split on the same term twice on the same branch.
 This suggests a proof strategy where you \Propagate{} when you can, \Split{}
@@
 \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-  Name & \text{Rule} & Prerequisites\\
+  Name & \text{Rule} & Side conditions\\
   \midrule
@@
 
-\subsection{\Calculus{} for one automaton is correct}\label{sec:single:correct}
-
-Since correctness is defined by the formulation of the $\Image{}$~predicate, our
-correctness proof amounts to proving inductively that no rule would invalidate
-that definition.
+\subsection{Correctness of \Calculus{}}\label{sec:single:correct}
+
+Our correctness proof of \Calculus{} consists of two main parts: we
+first show that the construction of a proof always terminates, and
+then that each of the proof rules in Table~\ref{tbl:rules:single} is
+an equivalence transformation, i.e., does not change the set of
+satisfying assignments of a formula. In combination, those two results
+immediately imply that \Calculus{} gives rise to a decision procedure.
 
 \subsubsection{\Calculus{} terminates}
-\begin{lemma}\label{lma:single-terminates} A solver implementing the calculus
-  described in \cref{sec:calculus} starting with some decidable set of clauses
-  $\SomeClause{}$ and an instance of our predicate $\SinglePredicateInstance$
-  will eventually terminate.
+\begin{lemma}\label{lma:single-terminates}
+  Suppose $\SomeClause{}$ is a set of formulas in which the predicates
+  $\Image$ and $\Connected$ only occur positively. There is no
+  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
+  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
+  $P_i$ by applying one of the rules in Table~\ref{tbl:rules:single}.
 \end{lemma}
@@
 \begin{proof}
-The first remaining source of divergence is from leftover predicates. The
-$\Image$ predicate can always be trivially removed by applying \Expand{}. This
-means a leftover $\Image$ predicate can never cause divergence. The only
-remaining predicate then is $\Connected$, which can always be removed by
-applying $\Subsume$ when no other of the rules is applicable. Hence all
-predicates introduced in the calculus can always be removed.
-
-It remains to prove that each rule is only applied finitely many times. This is
- obvious since both \Propagate{} and \Split{} can only be applied to transition
- variables whose $\Filter$ value has no known lower bound. However, \Split{} can
- always be used to split the proof tree and add clauses in precisely the right
- shape, thereby guaranteeing the ability to make forward progress. Therefore,
- the number of transition variables whose bounds are unknown is decreasing
- monotonically for any given proof goal, and hence termination is guaranteed.
+  The rule~\Expand{} can only be applied finitely often, since each
+  application removes one $\Image$ predicate, and none of the rules
+  introduce new instances of the predicate. The rule~\Subsume{} can
+  only be applied finitely often, since it strictly decreases the
+  combined number of $\Image$ and $\Connected$ predicates in sets of
+  formulas, and none of the rules increases that number.
+
+  To show termination of \Split{} and \Propagate{}, observe that the
+  $\Filter$ in a predicate~$\Connected(\Automaton, \Filter)$ is never
+  updated on a proof branch, which means that the set of terms
+  $\Filter(t)$ for $t \in \Automaton$ on every branch is finite. Each
+  application of \Split{} and \Propagate{} adds a new
+  formula~$\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$
+  to a proof goal, which can only happen finitely often.
 \end{proof}
@@
     \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-      Name & \text{Rule} & Prerequisites\\
+      Name & \text{Rule} & Side conditions\\
       \midrule

### 2022-07-07 main.tex
Monoid definitions

@@
 
-We write commutative monoids as $\Monoid =
-\left(X;+_{\Monoid};0_{\Monoid}\right)$, where we mean $X$ to be the underlying
-set of elements, $+_{\Monoid}$ to be the (commutative) monoid map, with
-$0_{\Monoid}$ as its neutral element.
+A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
+consisting of the carrier set of elements, $X$, an associative binary operation
+$X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,
+c \in X$, $(a \MonoidOp b) \MonoidOp c = a \MonoidOp (b \MonoidOp c)$. Finally,
+$\Monoid$ must have an identity element $0_{\Monoid} \in X$ such that
+$0_{\Monoid} \MonoidOp a = a \MonoidOp 0_{\Monoid} =   a$ for every $a \in X$.
+We sometimes use integer multiplication to represent repeated application of
+$\MonoidOp$, e.g. $3a = a \MonoidOp a \MonoidOp a$, for $a \in X$. $\Monoid$ is
+called \textit{commutative} if $\MonoidOp$ also commutes, that is if $a
+\MonoidOp b = b \MonoidOp a$ for all $a, b \in X$.
+
+Finally, we sometimes refer to a structure-preserving map between two monoids
+$\Monoid_1, \Monoid_2$ as a \textit{homomorphism} between $\Monoid_1$ and
+$\Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
+\MonoidOp_1 b) = \Map(a) \MonoidOp_2 \Map(b)$, where $S_1, S_2$ are the carrier
+sets of $\Monoid_1$ and $\Monoid_2$ respectively, and $\MonoidOp_1, \MonoidOp_2$
+their binary operations.
 
@@
 the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
+Note that $\Strings = \Tuple{\Alphabet;\Concat{};\epsilon}$, is a
+non-commutative monoid, referred to as the free monoid. The string length
+function, $\Length{s}$ is an example of a homomorphism between $\Strings$
+and~$\mathbb{Z}$.
 
@@
 $\Map:\: \Sigma^* \to \Monoid$, where
-$\Monoid = \left(X;+_{\Monoid};0_{\Monoid}\right)$ is a commutative
+$\Monoid = \left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative
 monoid; this means that $\Map(\epsilon) = 0_{\Monoid}$ and
-$\Map(u \Concat w) = \Map(u) +_{\Monoid} \Map(w)$. We give several
+$\Map(u \Concat w) = \Map(u) \MonoidOp \Map(w)$. We give several
 examples of such generalised Parikh image later in this section.
@@
 implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
-\left(X;+_{\Monoid};0_{\Monoid}\right)$ and the map into it $\Map : \Strings
+\left(X;\MonoidOp;0_{\Monoid}\right)$ and the map into it $\Map : \Strings
 \rightarrow X$, we also allow mapping over transitions:

### 2022-07-07 main.tex
Minor tweaks

@@
 first show that the construction of a proof always terminates, and
-then that each of the proof rules in Table~\ref{tbl:rules:single} is
+then that each of the proof rules in \cref{tbl:rules:single} is
 an equivalence transformation, i.e., does not change the set of
@@
   has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in Table~\ref{tbl:rules:single}.
+  $P_i$ by applying one of the rules in \cref{tbl:rules:single}.
 \end{lemma}
@@
 implementation and descriptions alike. The second is more practical. To increase
-the probability that our results are indeed representative both in the sense of
+the probability that our results are representative both in the sense of
 representing runs on expected, real world input and in the statistical sense for
-a given run of \Catra, despite the fact that we use randomness in our
-implementation, we execute many experiments. During the development of \Catra{},
-we have also executed experimeints on different hardware and software
-configurations on smaller samples of instances to ensure that the general trend
-holds. The logs from these executions are distributed along with notes as part
-of the artefact in the interest of transparency. Moreover, we have validated
-\Fudge{all reported satisfiable instances and their witnesses} with \Nuxmv{} to
-ensure that \Calculus{} is indeed sound and does not cheat.
-
-Another threat to validity would be if our implementations of the competition
-(baseline and \Nuxmv) would deviate significantly from the expected
-implementations. For \Nuxmv{} we use the default configuration which we believe
-should be performant (or it should not be the default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made easy for artefact reviewers.
-
-The most important and most probable threat to validity is our choice of
-implementation platform and our choice of automata library, that is our choice
-not to use one. Since our automata are more advanced than anything supported by
-e.g. the BRICS library, we opted to use our own implementation. There are some
-signs that in particular our product computation is inefficient, notably the
-good performance of low-threshold automata materialisation that prioritises
-computing smaller products. This means that our implementation of \Calculus{}
-makes much less heavy use of our product construction library than the baseline
-implementation does, instead putting its weight into \Princess{}' theorem
-proving abilities. This situation would unfairly advantage \Calculus{} if our
-automata library was indeed the bottleneck. We believe this is unlikely since
-similar performance issues as the ones in baseline in our string solver was the
-instigating reason for this research. Additionally, while we may have a
-constant-factor performance impact from bad implementation, it is less likely
-for our potentially poor implementation to be asymptotically worse than e.g.
-BRICS. However, the performance improvement of \Calculus{} over baseline seen in
-\cref{sec:experiments} is generally not constant.
+a given run of \Catra despite the fact that we use randomness in our
+implementation we execute many experiments. Moreover, we have validated all
+reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
+is indeed sound and does not cheat.
+
+Another threat to validity would be if the competition (baseline and \Nuxmv)
+would be disadvantaged in our comparison. For \Nuxmv{} we use the default
+configuration which we believe should be performant (or it should not be the
+default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
+easy for artefact reviewers. For baseline, our best argument is that the
+solution we use is close to the one found in the \Ostrich{} string solver and
+that it therefore should be realistic.
+
+The most probable threat to validity is our choice of implementation platform
+and our choice of automata library, that is our choice not to use one. Since our
+automata are more advanced than anything supported by e.g. the BRICS library, we
+opted to use our own implementation. There are some signs that in particular our
+product computation is inefficient, notably the good performance of
+low-threshold automata materialisation that prioritises computing smaller
+products. This means that our implementation of \Calculus{} makes much less
+heavy use of our product construction library than the baseline implementation
+does, instead putting its weight into \Princess{}' theorem proving abilities.
+This situation would unfairly advantage \Calculus{} if our automata library was
+indeed the bottleneck. We believe this is unlikely since similar performance
+issues as the ones in baseline in our string solver was the instigating reason
+for this research. Additionally, while we may have a constant-factor performance
+impact from bad implementation, it is less likely for our potentially poor
+implementation to be asymptotically worse than e.g. BRICS. However, the
+performance improvement of \Calculus{} over baseline seen in
+\cref{sec:experiments} is generally not constant. Additionally, non-systematic
+profiling suggests that both \Princess{}-based backends spend most of their time
+in \Princess{}, suggesting that the automaton implementation is not the
+bottleneck.
 

### 2022-07-07 main.tex
expand on examples

@@
 %% Title information
-\title{An Efficient Calculus for Generalised Parikh Images of Regular Languages}
+\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
                                         %% when present, will be used in
@@
    in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wuide
+   method recently described by~\citeauthor{approximate-parikh} by a wide
    marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
@@
 \begin{prooftree}
-  \infer0[\PresburgerClose]{
+  \hypo{\bigstar}
+  \infer1[]{
     \begin{matrix}
@@
   % BRANCH:  = 0
-  \infer0[\PresburgerClose{}]{
+  \hypo{\bigstar}
+  \infer1[]{
     \begin{matrix}
@@
 the starting state we must avoid the other one, perform propagation based on
-that fact, use some equational reasoning to derive the expected forms of the
+that fact, use some algebraic reasoning to derive the expected forms of the
 upper and lower bounds to propagate the now disconnected transitions, and
 finally subsume, removing the $\Connected{}$ predicate when we are unable to use
-either of the rules, leaving only the satisfiable remnants of the flow
-equations.
+either of the rules, leaving only the satisfiable remnants of the flow equations
+at the respective leaves marked with~$\bigstar$. To obtain a solution from the
+proof, one would need to perform standard model generation on the remaining
+constraints to obtain values for the sought variable $\MonoidElement$. The full
+Parikh image is the disjunction of the constraints left at the two
+$\bigstar$-marked leaves, after quantifier elimination.
 
@@
 
+In the derivation tree of \cref{fig:derivation:multi} we start as before, but
+with the product version of the $\Image$ predicate. The only possible rule here
+is $\ExpandM$, which we use to add the corresponding constraints on each
+automata of the product as we would have had in the single-automaton version. We
+perform algebraic reasoning on those equations, along with the constraint on
+$\MonoidElement$ to determine bounds on transition variables that will enable us
+to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
+finished doing so, we use \Materialise{} to compute the product of the now
+filtered automata. Further algebraic reasoning allows us to remove the final
+instance of $\Connected$. We then perform \Expand{} to get rid of the final
+$\Image$ instance, and are immediately able to $\Subsume$ the resulting
+$\Connected{}$ predicate. This, again, leaves us with a set of linear
+inequalities that we can use to obtain a model of our final $\MonoidElement$
+values (now a vector). Performing quantifier elimination on the remaining clauses would also produce the Presburger formula of the Parikh image under the constraint we started with.
+
+\begin{figure}
+  \centering
 \begin{prooftree}
-  \infer0[\PresburgerClose{}]{
+  \hypo{\bigstar}
+  \infer1[]{
     \begin{matrix}
@@
 \end{prooftree}
+\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
+\end{figure}
 
@@
 representing runs on expected, real world input and in the statistical sense for
-a given run of \Catra despite the fact that we use randomness in our
+a given run of \Catra{} despite the use of randomness in our
 implementation we execute many experiments. Moreover, we have validated all

### 2022-07-07 main.tex
more proof

--- 2022-07-07 main.tex
@@
 
-A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
-consisting of the carrier set of elements, $X$, an associative binary operation
-$X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,
-c \in X$, $(a \MonoidOp b) \MonoidOp c = a \MonoidOp (b \MonoidOp c)$. Finally,
-$\Monoid$ must have an identity element $0_{\Monoid} \in X$ such that
-$0_{\Monoid} \MonoidOp a = a \MonoidOp 0_{\Monoid} =   a$ for every $a \in X$.
-We sometimes use integer multiplication to represent repeated application of
-$\MonoidOp$, e.g. $3a = a \MonoidOp a \MonoidOp a$, for $a \in X$. $\Monoid$ is
-called \textit{commutative} if $\MonoidOp$ also commutes, that is if $a
-\MonoidOp b = b \MonoidOp a$ for all $a, b \in X$.
-
-Finally, we sometimes refer to a structure-preserving map between two monoids
-$\Monoid_1, \Monoid_2$ as a \textit{homomorphism} between $\Monoid_1$ and
-$\Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
-\MonoidOp_1 b) = \Map(a) \MonoidOp_2 \Map(b)$, where $S_1, S_2$ are the carrier
-sets of $\Monoid_1$ and $\Monoid_2$ respectively, and $\MonoidOp_1, \MonoidOp_2$
-their binary operations.
+We write commutative monoids as $\Monoid =
+\left(X;+_{\Monoid};0_{\Monoid}\right)$, where we mean $X$ to be the underlying
+set of elements, $+_{\Monoid}$ to be the (commutative) monoid map, with
+$0_{\Monoid}$ as its neutral element.
 
@@
 the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
-Note that $\Strings = \Tuple{\Alphabet;\Concat{};\epsilon}$, is a
-non-commutative monoid, referred to as the free monoid. The string length
-function, $\Length{s}$ is an example of a homomorphism between $\Strings$
-and~$\mathbb{Z}$.
 
@@
 \subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
+\label{sec:verma}
 
@@
 $\Map:\: \Sigma^* \to \Monoid$, where
-$\Monoid = \left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative
+$\Monoid = \left(X;+_{\Monoid};0_{\Monoid}\right)$ is a commutative
 monoid; this means that $\Map(\epsilon) = 0_{\Monoid}$ and
-$\Map(u \Concat w) = \Map(u) \MonoidOp \Map(w)$. We give several
+$\Map(u \Concat w) = \Map(u) +_{\Monoid} \Map(w)$. We give several
 examples of such generalised Parikh image later in this section.
@@
   The Parikh predicate, $\SinglePredicateInstance$, for some automaton
-  $\Automaton = \AutomatonTuple$, modulo some map $\Map$ to a commutative monoid
-  $\Monoid$ as described in \cref{sec:generalised} and with a transition
-  selection function $\Filter:\: \Transitions \to \Naturals$ holds when
-  $\MonoidElement \in \Monoid$ is the Parikh image of $\Automaton$ modulo
-  $\Map$, or more formally when there is an accepting path $\Path =
-  \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such that:
-  \begin{itemize}
-    \item $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
-    \item $\MonoidElement = \Map(\WordOf(\Path))$
-  \end{itemize}
+  $\Automaton = \AutomatonTuple$, modulo some map $\Map$ to a
+  commutative monoid $\Monoid$ as described in \cref{sec:generalised}
+  and with a transition selection function
+  $\Filter:\: \Transitions \to \Naturals$ holds when
+  $\MonoidElement \in \Monoid$ is the Parikh image of $\Automaton$
+  modulo $\Map$, or more formally when there is an accepting path
+  $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
+  that $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
+  for all $t \in \Transitions$, and $\MonoidElement = \Map(\WordOf(\Path))$.
 \end{definition}
@@
 Note that the filtering function~$\Filter$ is evaluated symbolically in these
-rules, and can in practice be read as a symbolic function from transitions to
-terms (e.g.~\texttt{t} or~\texttt{t+1}). In our implemendation \Catra{},
+rules, and can in practice be read as a function from transitions to
+$\Naturals$-valued terms (e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{},
 described in \cref{sec:implementation}, $\Filter$~is a vector of fresh variables
@@
 implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
-\left(X;\MonoidOp;0_{\Monoid}\right)$ and the map into it $\Map : \Strings
+\left(X;+_{\Monoid};0_{\Monoid}\right)$ and the map into it $\Map : \Strings
 \rightarrow X$, we also allow mapping over transitions:
@@
 
-The rules of \cref{tbl:rules:single} are meant to be used in tableaux-style
-proofs, and are read bottom-up, where the syntax at the bottom is replaced with
-the one at the top until \Fudge{it is possible to close the proof goal}. Since
+Each of the rules of \cref{tbl:rules:single} should be read and
+applied bottom-up, and relates
+premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
+conclusion~$\SomeClause$. When constructing a proof, we start from some
+root~$\SomeClause$, and then apply proof rules to the goals of the proof
+until a goal can be closed, or no more rule is applicable. Since
 our rules operate by adding and matching linear (in)equalities in a proof goal,
@@
 
-\subsubsection{\Calculus{} computes the homomorphic image}
-\Fudge{I am unsure of the definitions here}
+\subsubsection{The rules in Table~\ref{tbl:rules:single} are solution-preserving}
+
 \begin{lemma}\label{lma:single-correct}
-  Starting with $\SinglePredicateInstance$, we will close the goal with $\top$
-  if $\exists s \in \Language(\Automaton) \HoldsThat \Map(s) = \MonoidElement$
-  and $\bot$ otherwise.
+  Consider an application of one of the rules in
+  Table~\ref{tbl:rules:single}, with
+  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
+  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
+  conclusion~$\SomeClause$ if and only if it satisfies one of the
+  premises~$\SomeClause_i$.
 \end{lemma}
+
 \begin{proof}
-Since we are always guaranteed to terminate following
-\cref{lma:single-terminates}, we know that we will close the goal eventually.
-Additionally, the property of \cref{lma:single-terminates} is trivially obtained
-from the semantics of the $\Image$ predicate as defined in
-\cref{def:single-image}. This means that we only need to prove the preservation
-of those properties to prove \cref{lma:single-correct}.
-
-\Fudge{I don't understand.}
-  % \[
-  % \begin{aligned}
-  %   & \inferrule
-  %   {\Connected(\Automaton, \Filter) \land \FlowEq(\Automaton, \Filter) \land \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
-  %   {\SinglePredicateInstance, \SomeInequalities, \SomeClause} &\beta \models R_0 \iff \exists i : \beta \models R_i
-  % \end{aligned}
-  % \]
-
-\begin{lemma}\label{lma:expand-preserves}
-  \Expand{} preserves the semantics of $\Image$.
-\end{lemma}
-\begin{proof}
-
-  After application of \Expand{}, $\FlowEq$ guarantees that any
-  $\Filter(\Transition)$  for every $\Transition \in \Automaton$ is consistent
-  with the flow through the automaton. After its application, the proof goal
-  either contains an unsatisfiable set of Presburger formulae if the flow is
-  infeasible, otherwise a satisfiable set. An infesaible flow can happen in
-  either of these conditions:
+  This property has to be shown by analysing the possible applications
+  of each proof rule.
+
+  \Expand{} unfolds the definition of the $\Image$ predicate. To show
+  that the rule is solution-preserving we prove the equivalence of the
+  upper and lower sets of formulas:
   \begin{itemize}
-    \item if $\Filter$ demands that some transition $\Transition$ is visited more times than would be allowed by $\Automaton$, or
-    \item if the path described by $\Filter$ is not consistent with a path from
-    $\InitialState$ to some state $\State$ in $\AcceptingStates$.
+  \item Assume that $\beta$ satisfies the conclusion, which means that
+    there is some accepting path
+    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ with
+    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
+    \Path)$ and $\val_\beta(\MonoidElement) =
+    \Map(\WordOf(\Path))$. Since immediately implies that $\beta$
+    satisfies $\Connected(\Automaton, \Filter)$, since a path is
+    connected, and $\FlowEq(\Automaton, \Filter)$ since an accepting
+    path satisfies the flow equations. The
+    equation~$\MonoidElement = \sum_{\Transition \in
+      \Transitions_\Automaton}\Filter(\Transition) \cdot
+    \Map(\Transition)$ holds because of
+    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
+    \Path)$.
+  \item Assume that $\beta$ satisfies the premise, which implies that
+    $\val_\beta(\Filter)$ describes a consistent, connected flow of
+    the automaton. By the same argument as in
+    \cite{generate-parikh-image} (Section~\ref{sec:verma}), this flow
+    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
+    that each transition~$\Transition$ occurs on $\Path$ exactly
+    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
+    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
+    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
   \end{itemize}
 
-  Note that \Expand{} is taken directly from the existing formulation of
-  \eqref{eq:generate-parikh}.
-\end{proof}
-
-\begin{lemma}\label{lma:split-preserves}
-  \Split{} preserves the semantics of $\Image$.
-\end{lemma}
-\begin{proof}
-  \Split{} allows us to case-split at most once on each transition $\Transition
-  \in \Transitions$. This means that our proof tree at its leaves will be
-  guaranteed to have all transitions of known bounds in the correct format. This
-  means that once \Split{} is at fixpoint, the status (present or absent) of any
-  transition is known. Therefore it preserves the semantics of the $\Image$
-  predicate.
-\end{proof}
-
-\begin{lemma}\label{lma:propagate-preserves}
-  \Propagate{} preserves the semantics of $\Image$.
-\end{lemma}
-\begin{proof}
-  \Propagate{} will lazily enforce connectedness to any state. If it can be
-  applied, there exists a cut $C$ in $\Automaton$ such that some state $\State$
-  for any incoming transition $\Transition \in \Transitions_\Automaton$ has
-  $\Filter(t) = 0$  as part of the equations of the proof goal, and such that
-  $\State$ has an outgoing transition $\Transition'$ whose lower bound is not
-  $0$. Adding this conclusion will make equations unsatisfiable precisely if
-  there exists a transition $\Transition \in \Transitions_\Automaton$ such that
-  $\Filter(\Transition)$ is inconsistent with the homomorphic image, since it
-  implies that the selection of transitions is disconnected.
-\end{proof}
-
-\begin{lemma}\label{lma:subsume-preserves}
-  \Subsume{} preserves the semantics of $\Image$.
-\end{lemma}
-\begin{proof}
-  Since this rule can \emph{only} be applied if no other rule can, it means that
-  the upper and lower bounds of any transition $\Transition \in \Transitions$ as
-  defined by $\Filter(\Transition)$ is known, meaning clauses of the shape
-  $\Filter(\Transition) > 0$ or $\Filter(\Transition) = 0$ are present in the
-  proof goal for every transition and that the $\Image$ predicate has been
-  consumed by \Expand{} to be replaced by $\Connected$. This in turns means that
-  the goal can be closed only if the assignment of $\Filter$ is consistent with
-  an accepting path through $\Automaton$.
-
-  It also means that \Propagate{} must have been applied whenever possible, meaning that for any transition $\Transition$ starting in a state $\State$ that is disconnected the corresponding $\Filter$ values, we must have applied $\Propagate{}$ to derive $\Filter(\Transition) = 0$.
-Note that for an infeasible assignment (such as an unsatisfiable instance) this would contradict other decisions, but it would preserve the semantics of $\Image$.
-\end{proof}
-
-Since the final step we execute is \Subsume, which we know from
-\cref{lma:subsume-preserves} preserves the semantics of $\Image$ and each step
-along the way
-(\cref{lma:expand-preserves,lma:split-preserves,lma:propagate-preserves}), it
-holds that \Calculus{} for one automaton also preserves the semantics.
+  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
+  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
+  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
+  be satisfied, implying the property.
+
+  For \Propagate{}, suppose that
+  $\SeparatingCut(C, \Automaton, \State)$, which means that every
+  accepting path visiting~$q$ contains at least one of the transitions
+  in $C$. For a $\beta$ satisfying $\Connected(\Automaton, \Filter)$
+  and $\Filter(\Transition') = 0$ for $\Transition' \in C$, this means
+  that also $\val_\beta(\Filter(\Transition)) = 0$ has to hold.
+
+  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
+  then a goal must contain~$\Filter(\Transition) = 0$ or
+  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
+  formulas in $\SomeInequalities$ are inconsistent, an application of
+  \Subsume{} is trivially solution-preserving; therefore assume that
+  $\SomeInequalities$ is consistent, which means that it contains
+  exactly one of $\Filter(\Transition) = 0$ or
+  $\Filter(\Transition) > 0$ for each $\Transition$.
 \end{proof}

### 2022-07-07 main.tex

@@
 implementation and descriptions alike. The second is more practical. To increase
-the probability that our results are indeed representative both in the sense of
+the probability that our results are representative both in the sense of
 representing runs on expected, real world input and in the statistical sense for
-a given run of \Catra, despite the fact that we use randomness in our
-implementation, we execute many experiments. During the development of \Catra{},
-we have also executed experimeints on different hardware and software
-configurations on smaller samples of instances to ensure that the general trend
-holds. The logs from these executions are distributed along with notes as part
-of the artefact in the interest of transparency. Moreover, we have validated
-\Fudge{all reported satisfiable instances and their witnesses} with \Nuxmv{} to
-ensure that \Calculus{} is indeed sound and does not cheat.
-
-Another threat to validity would be if our implementations of the competition
-(baseline and \Nuxmv) would deviate significantly from the expected
-implementations. For \Nuxmv{} we use the default configuration which we believe
-should be performant (or it should not be the default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made easy for artefact reviewers.
-
-The most important and most probable threat to validity is our choice of
-implementation platform and our choice of automata library, that is our choice
-not to use one. Since our automata are more advanced than anything supported by
-e.g. the BRICS library, we opted to use our own implementation. There are some
-signs that in particular our product computation is inefficient, notably the
-good performance of low-threshold automata materialisation that prioritises
-computing smaller products. This means that our implementation of \Calculus{}
-makes much less heavy use of our product construction library than the baseline
-implementation does, instead putting its weight into \Princess{}' theorem
-proving abilities. This situation would unfairly advantage \Calculus{} if our
-automata library was indeed the bottleneck. We believe this is unlikely since
-similar performance issues as the ones in baseline in our string solver was the
-instigating reason for this research. Additionally, while we may have a
-constant-factor performance impact from bad implementation, it is less likely
-for our potentially poor implementation to be asymptotically worse than e.g.
-BRICS. However, the performance improvement of \Calculus{} over baseline seen in
-\cref{sec:experiments} is generally not constant.
+a given run of \Catra{} despite the use of randomness in our
+implementation we execute many experiments. Moreover, we have validated all
+reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
+is indeed sound and does not cheat.
+
+Another threat to validity would be if the competition (baseline and \Nuxmv)
+would be disadvantaged in our comparison. For \Nuxmv{} we use the default
+configuration which we believe should be performant (or it should not be the
+default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
+easy for artefact reviewers. For baseline, our best argument is that the
+solution we use is close to the one found in the \Ostrich{} string solver and
+that it therefore should be realistic.
+
+The most probable threat to validity is our choice of implementation platform
+and our choice of automata library, that is our choice not to use one. Since our
+automata are more advanced than anything supported by e.g. the BRICS library, we
+opted to use our own implementation. There are some signs that in particular our
+product computation is inefficient, notably the good performance of
+low-threshold automata materialisation that prioritises computing smaller
+products. This means that our implementation of \Calculus{} makes much less
+heavy use of our product construction library than the baseline implementation
+does, instead putting its weight into \Princess{}' theorem proving abilities.
+This situation would unfairly advantage \Calculus{} if our automata library was
+indeed the bottleneck. We believe this is unlikely since similar performance
+issues as the ones in baseline in our string solver was the instigating reason
+for this research. Additionally, while we may have a constant-factor performance
+impact from bad implementation, it is less likely for our potentially poor
+implementation to be asymptotically worse than e.g. BRICS. However, the
+performance improvement of \Calculus{} over baseline seen in
+\cref{sec:experiments} is generally not constant. Additionally, non-systematic
+profiling suggests that both \Princess{}-based backends spend most of their time
+in \Princess{}, suggesting that the automaton implementation is not the
+bottleneck.
 

### 2022-07-07 main.tex
Evaluation

@@
 \usepackage{amsthm}
+\usepackage{numprint}
 \usepackage{ebproof}
@@
 
-We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances generated
-by the \Ostrich{} string constraint solver when solving the \Fudge{pyex
-benchmarks}. \Fudge{The instances are all enormous and have some on average 900
-products of 400000 automata with 97 counters each and 9999 constraints}. After
-generating an initial~\InitialNrBenchmarks{}, we remove~\NrBroken{} misgenerated
-instances that did not parse, as well as~\NrTrivial{} instances solved in under
-five seconds by the baseline backend. The benchmarks are run on
-commit~\texttt{\commit}.
+We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
+automata intersection problems generated by the \Ostrich{} string constraint
+solver when solving the \Fudge{pyex benchmarks}. After generating an
+initial~\InitialNrBenchmarks{}, we remove~\NrBroken{} misgenerated instances
+that did not parse, as well as~\NrTrivial{} instances solved in under five
+seconds by baseline. The benchmarks are run on commit~\texttt{\commit}.
 
@@
 maximum heap of~\MaxHeapSize{}. We used \Nuxmv{} version~\NuxmvVersion{} invoked
-as a subprocess for each instance. Experiments were executed on a fresh JVM per
-backend, but the JVM was kept active for each batch of \BatchSize{} instances,
-meaning that JIT compilation would make subsequent executions faster, while
-still periodically restarting the JVM. We believe this represents a realistic
-use case where \Calculus{} is integrated into a string solver or other system
-and is called repeatedly. To mitigate systemic impacts of this decision on our
-results, experiments were executed in random order for all backends. We run all
-benchmarks with a timeout of~\RuntimeTimeout{}.
-
-As a way of gauging the differences in overhead between the native backends and
-\Nuxmv, we executed a small experiment where we ran the same trivial instance 30
-times and observed the runtimes for all solvers. All backends saw an improvement
-from JIT compilation within the first three runs, but subsequent runs only
-improved the native backends. The final measured overhead for the fully warm
-JVMs was less than~\OverheadSeconds. From this we draw the conclusion that major
-differences in runtime between the native backends and \Nuxmv{} is more likely
-to stem from innate differences rather than unfair overhead resulting from our
-particular configuration. Additionally, one of the advantages of \Calculus{} to
-\Nuxmv{} is that the former can me integrated into an automated theorem prover
-while the latter has to be called as a separate process. In this sense, our
-benchmarking setup can be said to be realistic.
-
-All runtimes are measured in wall-clock time as observed by the JVM when executing the instance, and exclude time spent parsing, which in all observed cases was minimal in comparison to the runtime, usually far below 0.1 second.
-
-It should be noted that despite running in the standard, deterministic
-configuration, we observed some degree of nondeterminism in \Nuxmv. Instances
-that were solved or unsolved would sometimes be solved or timed out in
-subsequent runs. This observation was made even without randomised execution
-order, suggesting that the nondeterminism likely stemmed either from our
-encoding of the instances into \Nuxmv's format, or the execution environment.
-Similarly, \Calculus{} showed similar signs of nondeterminism, but only for
-satisfiable instances. To address this, we increased the runtime from our
-initial experiments to give us a wider marigin. By contrast, the baseline
-implementation showed no signs of nondeterminism. It should be noted that for
-both backends exhibiting nondeterminism, the nondeterminism did not change the
-general trend of the experiments, even for experiments with few instances.
-Therefore, we believe that our large set of benchmarks will protect us from bias
-since any backend is unlikely to be neither consistently lucky nor consistently
-unlucky.
+as a subprocess for each instance. Instances were executed in batches of
+\BatchSize{}, each given a fresh JVM. We believe this represents a realistic use
+case where \Calculus{} is used to support e.g a string solver. Experiments were
+executed in random order for all backends. Each instance got a time budget
+of~\RuntimeTimeout.
+
+All runtimes are measured in wall-clock time as observed by the JVM when
+executing the instance, and exclude time spent parsing (usually far below
+\numprint{0.1}s).
 
@@
 
-In Figures~\ref{fig:solve-division} we show how many of the~\NrBenchmarks{} the
-respective back-ends could solve and with which status. A full summary of their
-outcomes is also available in Table~\ref{tab:solve-status}. We see here that
+In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances
+the respective back-ends could solve and with which status. A full summary of
+their outcomes is also available in \cref{tab:solve-status}. We see that
 \Calculus{} generally outperforms \Nuxmv{} on unsatisfiable instances, while
-being slightly worse at satisfiable instances. Both \Nuxmv{} and \Calculus{}
-massively and strictly outperform the baseline approach on every kind of
-instance, but most of all on satisfiable instances. In fact, the baseline
-approach was significantly harder to benchmark due to its much higher demand for
-RAM.
+being worse at satisfiable instances. Both \Nuxmv{} and \Calculus{} strictly
+outperform baseline on every kind of instance, but most of all on satisfiable
+instances. \Cref{tab:solve-status} also shows a small number of bugs in
+\Calculus{}: 27~satisfiable instances are misreported as \Unsat{}, and 20
+additional instances fail with an error. Both problems are due to
+non-deterministic bugs in the clause learning. Note that \Calculus{} is sound:
+it never misreports an unsatisfiable instance as \Sat{}, and all
+\Fudge{\numprint{6218}} satisfying assignments have been verified by \Nuxmv{}.
 
@@
 
-% \begin{figure}[ht]
-%   \begin{minipage}[b]{0.75\linewidth}
-%     \centering 
-%       \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter}
-%       \caption{Runtime duel: \Calculus{} versus baseline.}
-%       \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter}
-%       \caption{Runtime duel: \Calculus{} versus \Nuxmv.}
-%     \end{minipage}  
-%   \caption{The pairwise runtime of each instance for instances solved by any backend. Timeouts are reported as a runtime of \RuntimeTimeout.}
-%   \label{fig:duels}
-% \end{figure}
-
 \subsubsection{Scalability}\label{sec:scaling}
 
-A cactus plot showing the number of instances solved within a given timeout for each backend can be seen in Figure~\ref{fig:cactus}. \Fudge{We see here that some solver is fast and some other is not}.
+A cactus plot showing the number of instances solved within a given timeout for
+each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
+\Calculus{} outperforms \Nuxmv{} for shorter timeouts, and that \Nuxmv{} is
+only able to catch up once we increase the timeout to far beyond what would be
+normal for a string solver.
 
@@
 deviate from what is actually benchmarked in section \cref{sec:experiments}, or
-if the methods used for benchmarking would be unsound. The first problem cannot
-easily be addressed, except by stating that we have been careful in our
-implementation and descriptions alike. The second is more practical. To increase
-the probability that our results are representative both in the sense of
-representing runs on expected, real world input and in the statistical sense for
-a given run of \Catra{} despite the use of randomness in our
-implementation we execute many experiments. Moreover, we have validated all
+if the methods used for benchmarking would be unsound.  To increase the
+probability that our results are representative both in the sense of
+representing  on expected inputs and in the statistical sense for a given run of
+\Catra{} despite the use of randomness in our implementation we execute many
+experiments. Moreover, to address the issue of correctness we have validated all
 reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
-is indeed sound and does not cheat.
+is indeed sound.
 
@@
 
-The most probable threat to validity is our choice of implementation platform
-and our choice of automata library, that is our choice not to use one. Since our
-automata are more advanced than anything supported by e.g. the BRICS library, we
-opted to use our own implementation. There are some signs that in particular our
-product computation is inefficient, notably the good performance of
-low-threshold automata materialisation that prioritises computing smaller
-products. This means that our implementation of \Calculus{} makes much less
-heavy use of our product construction library than the baseline implementation
-does, instead putting its weight into \Princess{}' theorem proving abilities.
-This situation would unfairly advantage \Calculus{} if our automata library was
-indeed the bottleneck. We believe this is unlikely since similar performance
-issues as the ones in baseline in our string solver was the instigating reason
-for this research. Additionally, while we may have a constant-factor performance
-impact from bad implementation, it is less likely for our potentially poor
-implementation to be asymptotically worse than e.g. BRICS. However, the
-performance improvement of \Calculus{} over baseline seen in
-\cref{sec:experiments} is generally not constant. Additionally, non-systematic
-profiling suggests that both \Princess{}-based backends spend most of their time
-in \Princess{}, suggesting that the automaton implementation is not the
-bottleneck.
+The greatest threat to validity is our choice of implementation platform and
+automata library. There are some signs that product computation is inefficient,
+notably the good performance of low-threshold automata materialisation that
+prioritises computing smaller products. This means that \Calculus{} makes less
+heavy use of product computation than baseline does, instead relying more on
+\Princess{}'. This situation would unfairly advantage \Calculus{} if our
+automata library was the bottleneck due to our automata library being
+unnaturally poor. We believe this is unlikely since similar performance issues
+have been reported for string solvers. Additionally, profiling suggests that
+both \Princess{}-based backends spend most of their time in \Princess{},
+suggesting that the automaton implementation is not the bottleneck.
 
@@
 
-In this paper we have introduced an automated theorem prover-friendly calculus
-to compute commutating operations on intersections of regular languages that we
-call \Calculus{}. We have evaluated it on \NrBenchmarks{} Parikh automata
-intersection problems generated \Fudge{by the Ostrich+} string solver
-\cite{ostrich} solving \Fudge{standard string constraint benchmarks} using our
-Parikh automata solver \Catra{}.
+In this paper we have introduced a calculus to compute commutating operations on
+intersections of regular languages that we call \Calculus{}. We have evaluated
+it on \NrBenchmarks{} Parikh automata intersection problems generated \Fudge{by
+the Ostrich+} string solver \cite{ostrich} solving \Fudge{standard string
+constraint benchmarks} using our Parikh automata solver \Catra{}.
 

### 2022-07-07 main.tex
Work on implementation, shortening

@@
 
-Please note that \Catra{} uses symbolic labels for automata. A symbolic label is
-defined as a range of matching single Unicode code points. This allows succinct
-representation of many regular expression patterns such as \lstinline{(a-z).*}
-which would have otherwise required $27$~nearly identical transitions. In this
-instance, the transition would be defined as \lstinline{a -> b [97, 122]}, as the
-interval is closed from both edges.
-
-In satisfaction mode, supported by all back-ends, \Catra{} tries to satisfy the
-constraints expressed by the input file, reporting satisfiable with register
-assignments or unsatisfiable much like traditional~SAT- or SMT solvers would.
-Additionally, the baseline and our own back-end also support generation of the
-Presburger formula describing the constraints of the input file.
-
-\Catra{} can solve equations of Parikh images. An equation like
-$\ParikhMap(\Automaton_1 \times \Automaton_2) = \ParikhMap(\Automaton_3 \times
-\Automaton_4)$ would be expressed as one \lstinline{synchronised} block per side of
-the equation containing their respective terms followed by a
-\lstinline{constraint} requiring all their counters to agree (e.g
-\lstinline|Aa = Ba && Ab = Bb ...|).
-Note that the product construction will implicitly require the counters of each term to agree with each other (and the product), so equalities between all counters of all automata are not necessary.
-
-We implement \Calculus{} as a theory plug-in for the \Princess{} theorem prover.
+\Catra{} uses symbolic labels for automata. A symbolic label is defined as a
+finite range of Unicode code points. This allows succinct representation of many
+regular expression patterns such as \lstinline{(a-z).*} which would have
+otherwise required $27$~transitions. The transition for the range would be
+written \lstinline{a -> b [97, 122]}.
+
+In satisfaction mode, supported by all backends, \Catra{} tries to satisfy the
+constraints expressed by the input file, reporting \Sat{} with register
+assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
+Additionally, baseline and \Calculus{} also support generation of the Presburger
+formula describing the constraints of the input file. Baseline uses standard
+quantifier elimination, and \Calculus{} uses the method described in
+\cref{sec:finding-the-image}.
+
 Since \Princess{} does not support multiple-arity predicates like the ones we
@@
 
-We implement the baseline approach using the same Presburger solver
-(\Princess{}), input file parser, and automaton implementation as \Catra. We
-do this in order to make the comparison between methods as fair as possible, and
-give us the ability to compare the efficiency of the calculus rules themselves.
-Using the formula of Equation~\ref{eq:generate-parikh}, we produce quantified
-Presburger formulae for each successive term and add them to Princess. In this
-fashion we compute the product incrementally term by term, checking
-satisfiability in each step. We use a priority queue to select which
-term to use for the next step of the product, and order it by the number
-of transitions as a heuristic for the size of the automaton. We use this heuristic
-to avoid computing large (and therefore slow) products until we have to, banking
-our hopes on finding a source of unsatisfiability early. The pseudocode for
-our implementation can be seen in Algorithm~\ref{alg:baseline}.
+We baseline using the same Presburger solver (\Princess{}), input file parser,
+and automaton implementation as \Catra. We do this in order to better analyse
+the impact of the calculus rules themselves. Using the formula of
+\eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
+successive term and add them to \Princess. We compute the product incrementally
+term by term, checking satisfiability at each step. We use a priority queue to
+select automata for each step, and order it by the number of transitions as a
+heuristic for the size of the automata. We use this heuristic to avoid computing
+large (and therefore slow) products until we have to, banking our hopes on
+computing an empty intermittent product early. The pseudocode for our
+implementation can be seen in \cref{alg:baseline}.
 
@@
 
-It should be noted that our automata (including the successive products) are
-always by construction forward- and backward- reachable-minimal. We avoid
-computing fully minimal automata since it is unclear when an automaton with
-registers can be safely minimised. This guarantees that any automaton we produce
-only contains states that are both reachable from the initial state and has a
-path to an accepting state. We never perform any other minimisation on the
-automata for either backend. More complex minimisation was left out since
-performing minimisation on automata with counters is non-trivial, and indeed
-already minimising symbolic automata is itself risks exponential blowup
-\cite{minimising-symbolic}.
+As an optimisation, our automata (including intermittent products) are created
+forward- and backward- reachable-minimal. Any automaton we produce only contains
+states that are both reachable from the initial state and has a path to an
+accepting state. We never perform any other minimisation on the automata for
+either backend. More complex minimisation was left out since performing
+minimisation on automata with counters is non-trivial, and minimising symbolic
+automata risks exponential blowup \cite{minimising-symbolic}.
 
@@
 There are a number of choices left unspecified in \Calculus{} as described in
-Sections~\ref{sec:calculus} and~\ref{sec:multiple}. For example, the order of
-materialisation of intermediate products and the order of splitting. In
-addition, the splitting specified in the general calculus is correct but
-inefficient. In this section we describe additional implementation details used
-to enhance our solution.
+\cref{sec:calculus,sec:multiple}. For example, the order of materialisation of
+intermediate products and the order of splitting. In this section we describe
+additional implementation details and techniques used to enhance \Catra.
 
@@
 We order our rule applications so that we first propagate connectedness if
-possible, then perform materialisation if tractable (see below!), then finally
-resort to splitting if we must.
-
-In addition to the splitting on individual transition variables described in Table~\ref{tbl:rules:single}, where we randomly select a variable whose bounds are not known, we also prefer splitting to sever/include a strongly connected component.
-We randomly select an automaton where we can compute a cut to at least one strongly connecting component that can be severed from the initial state, e.g.
-where the strongly connected component does not contain the initial state and where the sum of the transition variables we have associated with the transitions of a minimal cut between the initial state and an arbitrary state in the component is not known to be greater than 0, but also not equal to 0.
-If there are multiple such strongly connected components we choose one randomly.
-We then proceed to split on the cut as if it were a regular transition, e.g.
-its sum being zero or nonzero.
-In this way we use our splitting strategy to drive the connectedness constraint towards propagation by cutting off loops from the automata.
+possible, then perform materialisation if tractable as defined below, then
+finally resort to splitting if we must.
+
+In addition to applying \Split{} as described in \cref{tbl:rules:single} to
+randomly selected transitions, we prefer splitting to sever a strongly connected
+from the initial state. We randomly select an automaton where we can compute a
+cut between an SCC and the initial state, that is where the SCC does not contain
+the initial state and where the sum of the transition variables of the
+transitions in the cut is not known to be positive. If there are multiple such
+strongly connected components we choose one randomly. We then proceed to split
+on the sum of the transition variables of the cut as if it were a regular
+transition, e.g. its sum being zero or nonzero. In this way we drive \Calculus{}
+towards applying \Propagate{}.
 
@@
 
-Product materialisation, then, is the final piece of the puzzle.
-In the current implementation, we put off computing intermediate products until at least all but 2 transition variables of one of them have a known status, e.g.
-either is known to be zero or known to be positive.
-The number was chosen experimentally, and we observe a consistent trend that to have a low number and therefore a high threshold for materialisation is better throughout the implementation of other features.
-The other automaton for the product is selected randomly.
+Product materialisation is the final piece of the puzzle. In the current
+implementation we put off computing intermediate products until at least all but
+two transition variables of one of the automata is known to be either present or
+absent. The number was chosen experimentally, and we observe a consistent trend
+towards lower numbers being better. The other automaton for the product is
+selected randomly.
 
@@
 \cref{sec:ext:backjumping}, but we do implement forward-reachability cut
-learning (which had a very modest improvement in performance). No sophisticated
-clause learning for products has been implemented.
+learning. No sophisticated clause learning for products has been implemented.
 
@@
 unsurprising given how many random choices we make during solving and how
-tail-heavy our problem is. It seems to be very easy to get stuck with a poor
-choice of product to materialise or transitions to split on.
+tail-heavy our problem is.
 
@@
 \Fudge{\numprint{6218}} satisfying assignments have been verified by \Nuxmv{}.
+Baseline performs worse on satisfiable instances because it executes a heuristic
+meant to detect unsatisfiability early at a heavy penalty to satisfiable
+instances.
 

### 2022-07-07 main.tex
Cut some more stuff

@@
 
-We implement a calculus for Parikh automata as described in
-\cref{sec:parikh-automata}. The artefact submitted along with this paper
-is a program that reads one or more products of one or more DFA, the register
-incrementations performed along the edges of their transitions, and the labels
-of the transitions as ranges of Unicode characters, along with a set of
-constraints on the final values of their registers expressed as Presburger
-arithmetic. We call this program \Catra.\footnote{If you really must read it
-as an acronym, please read it as CAtegory Theory on Register Automata, or if you
-object to the somewhat nonstandard use of register automata and category theory,
-as Check Assignments of The Registers Afterwards, or alternatively if you find it
-all to be too much of a theoretical exercise as Can Anyone Think of A Real
-Application.}
+We implement \Calculus{} for Parikh automata as described in
+\cref{sec:parikh-automata}. The artefact submitted along with this paper is a
+program that reads an instance file with one or more products of one or more
+Parikh automaton with transition labels defined as ranges of Unicode characters,
+along with a set of constraints on the final values of their registers expressed
+as Presburger arithmetic in a C-like syntax. We call this program
+\Catra.\footnote{If you really must read it as an acronym, please read it as
+CAtegory Theory on Register Automata, or if you object to the somewhat
+nonstandard use of register automata and category theory, as Check Assignments
+of The Registers Afterwards. Or alternatively if you find it all to be too much
+of a theoretical exercise, as Can Anyone Think of A Real Application.}
 

### 2022-07-07 main.tex
PYEX

@@
 We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
-automata intersection problems generated by the \Ostrich{} string constraint
-solver when solving the \Fudge{pyex benchmarks}. After generating an
-initial~\InitialNrBenchmarks{}, we remove~\NrBroken{} misgenerated instances
-that did not parse, as well as~\NrTrivial{} instances solved in under five
-seconds by baseline. The benchmarks are run on commit~\texttt{\commit}.
+automata intersection problems generated by the \OstrichPlus{} string constraint
+solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
+initial~\InitialNrBenchmarks{}, we remove misgenerated instances that did not
+parse, as well as~\NrTrivial{} instances solved in under five seconds by
+baseline. The benchmarks are run on commit~\texttt{\commit}.
 
@@
 intersections of regular languages that we call \Calculus{}. We have evaluated
-it on \NrBenchmarks{} Parikh automata intersection problems generated \Fudge{by
-the Ostrich+} string solver \cite{ostrich} solving \Fudge{standard string
-constraint benchmarks} using our Parikh automata solver \Catra{}.
+it on \NrBenchmarks{} Parikh automata intersection problems generated by the
+\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
+suite \cite{pyex} using our Parikh automata solver \Catra.
 

### 2022-07-07 main.tex
Keywords

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=true}
+\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
   <concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
+  <concept_id>10003752.10003766.10003773.10003775</concept_id>
+  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
   <concept_significance>500</concept_significance>
@@
   <concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
+  <concept_id>10003752.10003766.10003776</concept_id>
+  <concept_desc>Theory of computation~Regular languages</concept_desc>
+  <concept_significance>300</concept_significance>
+  </concept>
+  <concept>
+  <concept_id>10003752.10003790.10003794</concept_id>
+  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
+  <concept_significance>500</concept_significance>
+  </concept>
+  <concept>
+  <concept_id>10003752.10010124.10010138.10010142</concept_id>
+  <concept_desc>Theory of computation~Program verification</concept_desc>
+  <concept_significance>300</concept_significance>
+  </concept>
+  <concept>
+  <concept_id>10003752.10003790.10011192</concept_id>
+  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
   <concept_significance>300</concept_significance>
@@
 
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
+\ccsdesc[500]{Theory of computation~Quantitative automata}
+\ccsdesc[300]{Theory of computation~Regular languages}
+\ccsdesc[500]{Theory of computation~Automated reasoning}
+\ccsdesc[300]{Theory of computation~Program verification}
+\ccsdesc[300]{Theory of computation~Verification by model checking}
 %% End of generated code
@@
 \OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
-suite \cite{pyex} using our Parikh automata solver \Catra.
+suite \cite{pyex} using our Parikh automata solver \Catra{}.
 

### 2022-07-07 main.tex
First stab at correctness proof

@@
    called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving \Fudge{standard string
-   constraint benchmarks involving string lengths}. We show that our solution
-   strictly outperforms the standard eager approach described
+   \OstrichPlus{} string constraint solver when solving Parikh automata
+   intersection problems produced when solving standard string constraint
+   benchmarks involving string length constraints. We show that our solution
+   strictly outperforms the standard approach described
    in~\citeauthor{generate-parikh-image} as well as the over-approximating
@@
 
+The key ideas of \Calculus{} is a combination of problem-aware case splitting,
+lazy enforcement of automata connectivity, and lazy computation of products.
+
 The problem of computing constraints on Parikh images over products of regular
@@
 We implement \Calculus{} as a theory for the \Princess{} automated theorem
-prover in the tool \Catra{}. \Catra also supports using the
-approximate method of \cite{approximate-parikh}, its fall-back variant adapted
+prover in the tool \Catra. \Catra also supports the approximate method of
+\cite{approximate-parikh}, its fall-back variant adapted
 from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
-checker~\cite{nuxmv}. Using \Catra{}, we compare \Calculus{} to the other two
-back-ends on \NrBenchmarks{} Parikh automata satisfaction problems generated
-\Fudge{by \OstrichPlus{}}.
+checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
+back-ends on \NrBenchmarks{} Parikh automata intersection problems generated by
+\OstrichPlus{} when solving the PyEx string constraint benchmark suite involving
+string length constraints \cite{pyex}.
 
@@
 To keep things simple, in this first example we let $f$ be the function that
-gives the length of a given string, e.g. $f\left(\text{curious and curiouser
+gives the length of a given string; $f\left(\text{curious and curiouser
 still}\right) = 27$, and $f\left(\text{homo}\right) +
@@
 use the approach from \cite{generate-parikh-image} to find its image under $f$,
-and then verify the constraints. However, we can take advantage of both the
-constraints and definitions of $f$ to prune the automata before computing the
-product and help reduce the blow-up.
+and then solve the constraints in an automated theorem prover. However, we can
+take advantage of both the constraints and definitions of $f$ to prune the
+automata before computing the product and help reduce the blow-up.
 
@@
 that is finding a count of characters in a string accepted by both automata where
-there are more a:s than b:s. Since the underlying calculus is the same, we start immediately from Figure~\ref{fig:example-simplify-1}.
+there are more a:s than b:s. Since the underlying calculus is the same, we start immediately from \cref{fig:example-simplify-1}.
 
@@
 
-\section{Generalised Parikh Maps}\label{sec:generalised}
+\section{Projections on Parikh Maps}\label{sec:generalised}
 
@@
 approach is that it is more efficient to directly compute a
-generalised Parikh image~$\Map(\Language)$, than to first compute the
+projection on the Parikh image~$\Map(\Language)$, than to first compute the
 standard image~$\ParikhMap(\Language)$ followed by projection to
@@
 
-A useful example of such a simplifying morphism can express string
+One such simplifying homomorphism can express string
 length, the problem that originally motivated our study of the Parikh
@@
 
-Parikh images are also applicable to decide more general classes of
+Parikh images are also applicable for deciding more general classes of
 string constraints~\cite{ostrich-plus}. Consider, for instance, a
@@
 expression~$\mathtt{(ab|c)*}$. An automaton recognising this language
-is shown in Fig.~\ref{fig:parikh-automata} (left).
+is shown in \cref{fig:parikh-automata} (left).
 
@@
 
-\section{A Calculus for Generalised Parikh Images}\label{sec:calculus}
+\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
 
@@
 formulas into linear equations and inequalities ($\SomeInequalities$)
-and other formulas ($\SomeClause$). \Fudge{need to say that we assume
-  that the predicates (Im, etc.)  only occur positively.}
-
-Note that the filtering function~$\Filter$ is evaluated symbolically in these
-rules, and can in practice be read as a function from transitions to
-$\Naturals$-valued terms (e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{},
-described in \cref{sec:implementation}, $\Filter$~is a vector of fresh variables
-with the same size as~$\Transitions$.
+and other formulas ($\SomeClause$). We assume that our predicates $\Image$ and $\Connected$ only occur positively.
+
+The filtering function~$\Filter$ is evaluated symbolically, and can in practice
+be read as a function from transitions to $\Naturals$-valued terms
+(e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{}, described in
+\cref{sec:implementation}, $\Filter$~is a vector of fresh variables with the
+same size as~$\Transitions$.
 
@@
 
-\subsubsection{The rules in Table~\ref{tbl:rules:single} are solution-preserving}
+\subsubsection{The rules in \cref{tbl:rules:single} are solution-preserving}
 
@@
   Consider an application of one of the rules in
-  Table~\ref{tbl:rules:single}, with
+  \cref{tbl:rules:single}, with
   premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
@@
     the automaton. By the same argument as in
-    \cite{generate-parikh-image} (Section~\ref{sec:verma}), this flow
+    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
     can be mapped to an accepting path~$\Path$ of $\Automaton$ such
@@
 
-\subsection{The extended calculus is also correct}
-
-Since the expanded calculus is in practice the addition of two rules with the
-purpose of reducing an~$\Image$ predicate containing a product to the form
-solved by the previous single-automaton rules of \cref{tbl:rules:single}, we
-extend the reasoning from \cref{sec:single:correct} to the single-automaton
-rules of \cref{tbl:rules:multi}.
-
-\subsubsection{The calculus terminates}
-\begin{lemma}
-  The addition of the \ExpandM{} and \Materialise{} rules in \cref{tbl:rules:multi}
-  maintains termination as shown in \cref{lma:single-terminates}.
+\subsection{Correctness of \Calculus{} for Products of Automata}
+
+Since \cref{tbl:rules:multi} only extends the existing rules of
+\cref{tbl:rules:single}, we only have to prove termination and solution
+preservation for the new rules.
+
+\subsubsection{\Calculus{} for products of automata terminates}
+\begin{lemma}\label{lma:multi-terminates}
+  Suppose $\SomeClause{}$ is a set of formulas in which the product
+  version of $\Image$ only occurs positively. There is no
+  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
+  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
+  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
 \end{lemma}
+
 \begin{proof}
-  Similarly to how we previously showed that the number of possible executions
-  of the \Split{} rule is bounded above by the number of transitions of an
-  automaton, we can bound the number of applications of \Materialise{} by the
-  (monotonically) decreasing number of automata in the product until we have
-  approached the starting state for the single-automaton calculus. Trivially, we
-  can eagerly apply \Materialise{} repeatedly to do so. Similarly, \ExpandM{} is
-  just the generalisation of \Expand{} to multiple automata. In other words there are no possibilities of the calculus diverging.
+  The rule~\ExpandM{} can only be applied finitely often since its side condition only allows applying it once per instance of an $\Image$ predicate containing a product, and the only rule that introduces a new $\Image$ predicate with a product, $\Materialise$ does so with non-overlapping automata from the $\Image$ predicate it removes.
+Therefore, the number of automata in $\Image$ predicates is strictly decreasing.
+
+  The rule \Materialise{} can similarly only be used finitely many times, as
+  each application reduces the number of automata in the product of $\Image$ by
+  one automaton, until only one remains and \cref{lma:single-terminates} for
+  single-automaton instances applies.
 \end{proof}
 
-\subsubsection{The calculus implemements the homomorphic image of the product}
-\begin{lemma}
-  Using \Calculus{} to compute $\Image_{\Automaton_1\times\ldots\times\Automaton_k, \Map}(\Filter, \MonoidElement)$ is equivalent to using the single-automaton rules to compute $\Image_{\Automaton', \Map}(\Filter, \MonoidElement)$ where $\Automaton' = \Automaton_1\times\ldots\times\Automaton_k$.
+\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
+
+\begin{lemma}\label{lma:multi-correct}
+  Consider an application of one of the rules in
+  \cref{tbl:rules:multi}, with
+  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
+  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
+  conclusion~$\SomeClause$ if and only if it satisfies one of the
+  premises~$\SomeClause_i$.
 \end{lemma}
+
 \begin{proof}
-
-  We perform this proof equivalent to set inclusion: by proving that neither predicate is stronger than the other, in the sense of rejecting an element that the other accepts and vice versa.
-
-\begin{lemma}\label{lma:multi:rinclude}
-  There exists no element $\MonoidElement$ such that $\Image_{\Automaton_1\times\ldots\times\Automaton_k, \Map}(\Filter, \MonoidElement) = \top$ but $\Image_{\Automaton', \Map}(\Filter, \MonoidElement) = \bot$.
-\end{lemma}
-\begin{proof}
-  Assume that such an $\MonoidElement$ exists with the goal of deriving a
-  contradiction. Under this assumption, would we be able to use the two new
-  rules to close the proof goal with $\top$?
-
-  The \ExpandM{} rule would allow us to close no goal since it only adds
-  additional clauses. Our only hope is in using \Materialise{} to get rid of the
-  product step by step without introducing an unsatisfiable clause. However, all
-  \Materialise{} does is computing the product step by step, eventually arriving
-  at $\Automaton'$ modulo permutations. Therefore, it can be no more satisfiable
-  than the single-automata predicate.
-
-  \contents{\item Do I need to worry about the fact that I have no extra
-  clauses/$\SomeClause$ here? If present, I would need to apply the definition
-  of \BindingSum{} since that is what transfers constraints on individual
-  automata up and down the materialisation.}
-\end{proof}
-
-\begin{lemma}\label{lma:multi:linclude}
-  There exists no element $\MonoidElement$ such that $\Image_{\Automaton_1\times\ldots\times\Automaton_k, \Map}(\Filter, \MonoidElement) = \top$ but $\Image_{\Automaton', \Map}(\Filter, \MonoidElement) = \bot$.
-\end{lemma}
-\begin{proof}
-  Assume that such an element exists. Then $\MonoidElement =
-  \Map(\WordOf(\Path))$ for some $\Path$ ending up in an accepting state in
-  $\Automaton'$, and consistent with $\Filter$ along the transitions of $\Path$.
-  Then a corresponding path must exist in each automata of $\Automaton'$,
-  although possibly with different labels.
-
-  For \ExpandM{} to introduce unsatisfiability, either of its automata must be
-  disconnected (or its corresponding $\Connected{}$ predicate can be removed
-  with \Propagate{}), or its flow must be infeasible. However, this contradicts
-  the existence of some corresponding path $\Path$ in each automaton. Hence,
-  \ExpandM{} cannot be the source of the unsatisfiability.
-
-  The same applies for the second half of \Materialise{}, if the corresponding
-  path exists in each individual automaton and in $\Automaton'$, it must also
-  exist in every intermittent product on the way to $\Automaton'$, and if $\Map$
-  is consistent for intermittent transition labels as it must be, then neither the intermittent $\Image$ predicates nor the final one can be the source of the unsatisfiability.
-
-  This leaves one final potential source of unsatisfiability: the first half of
-  \Materialise{}: \BindingSum{}. For that to be unsatisfiable, it requires
-  intermittent automata $\Automaton_l, \Automaton_r$ such that adding
-  $\BindingSum(\Automaton_l \times \Automaton_r)$ is unsatisfiable. The
-  generated equations state that the transitions of $\Automaton_l$ and
-  $\Automaton_r$ must be used precisely as often as their corresponding results
-  in $\Automaton_l \times \Automaton_r$. For this set of equations to be
-  unsatisfiable there must exist no path through $\Automaton_l \times
-  \Automaton_r$ that preserves this. However, this is true from the definition
-  of the product of automata: since we know the product to be nonempty, it must
-  in some sense contain at least one path whose equivalent exists in both
-  $\Automaton_l$ and $\Automaton_2$ and cannot use a corresponding transition
-  more often, or it would recognise a different langauge. Hence no such case
-  exists.
-\end{proof}
-
-Since it follows from \cref{lma:multi:linclude,lma:multi:rinclude} that both
-variants of the predicate are precisely equally strong, it must mean that they
-are equivalent, and we have proven the correctness as defined in
-\cref{def:multiple}.
+  This property has to be shown by analysing the possible applications
+  of each proof rule.
+
+  \ExpandM{} applies \Expand{} to every automata $\Automaton_i$ of the product. 
+  \begin{itemize}
+  \item Assume that $\beta$ satisfies the conclusion, which means that
+    there is some accepting path in the product $\Automaton_1 \times \ldots \times \Automaton_k$ such that
+    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton_1 \times \ldots \times \Automaton_k)}$ with
+    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
+    \Path)$ and $\val_\beta(\MonoidElement) =
+    \Map(\WordOf(\Path))$.
+That immediately implies that every automaton $\Automaton_i$ would satisfy its $\Image{}$ constraints individually, and hence also the constraints generated by \ExpandM{}.
+\Fudge{Missing Materialise!}
+  \item Assume that $\beta$ satisfies the premise, which implies that
+    $\val_\beta(\Filter)$ describes a consistent, connected flow of
+    the automaton. By the same argument as in
+    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
+    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
+    that each transition~$\Transition$ occurs on $\Path$ exactly
+    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
+    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
+    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
+  \end{itemize}
+
+  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
+  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
+  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
+  be satisfied, implying the property.
+
+  For \Propagate{}, suppose that
+  $\SeparatingCut(C, \Automaton, \State)$, which means that every
+  accepting path visiting~$q$ contains at least one of the transitions
+  in $C$. For a $\beta$ satisfying $\Connected(\Automaton, \Filter)$
+  and $\Filter(\Transition') = 0$ for $\Transition' \in C$, this means
+  that also $\val_\beta(\Filter(\Transition)) = 0$ has to hold.
+
+  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
+  then a goal must contain~$\Filter(\Transition) = 0$ or
+  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
+  formulas in $\SomeInequalities$ are inconsistent, an application of
+  \Subsume{} is trivially solution-preserving; therefore assume that
+  $\SomeInequalities$ is consistent, which means that it contains
+  exactly one of $\Filter(\Transition) = 0$ or
+  $\Filter(\Transition) > 0$ for each $\Transition$.
 \end{proof}

### 2022-07-07 main.tex
missing proof step

@@
   $\Image$ and $\Connected$ only occur positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
+  infinite sequence of proofs~$P_r0, P_1, P_2, \ldots$ in which $P_0$
   has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
@@
   exactly one of $\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for each $\Transition$.
+  $\Filter(\Transition) > 0$ for each $\Transition$. Since
+  \Propagate{} is not applicable, the transitions~$\Transition$ with
+  $\Filter(\Transition) > 0$ must form a connect sub-graph of the
+  automaton; this means that $\Connected(\Automaton, \Filter)$ is
+  redundant as it is implied by $\SomeInequalities$.
 \end{proof}

### 2022-07-08 main.tex
That must have been a mistake
references

--- 2022-07-07 main.tex
@@
 solvers
-\cite{DBLP:conf/cav/AbdullaACHRRS15,length-aware-solver,approximate-parikh,more}.
-Other string solvers make use of Parikh automata~\cite{X}, and thus
+\cite{DBLP:conf/cav/AbdullaACHRRS15,length-aware-solver,approximate-parikh,DBLP:journals/corr/BerzishZG17}.
+Other string solvers make use of Parikh automata~\cite{parikh-automata}, and thus
 Parikh images in the general case, to handle operations including
@@
 quantifier-free representation of the Parikh image has exponential
-complexity~\cite{X}, and is in practice often impossible in reasonable
+complexity~\cite{DBLP:conf/issac/Weispfenning97}, and is in practice often impossible in reasonable
 time. Just solving the Parikh image membership problem is NP-complete,

### 2022-07-08 main.tex
Scatter plots, new figures

@@
   \input{graphs/solved_pivot_table.tex}
-  \caption{The result of running the respective back-ends by instance
-  satisifiability (satisifiable or unsatisifiable) with a timeout of
+  \caption{The result of running the respective back-ends with a timeout of
   \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
@@
 \section{Appendix}
-
+\subsection{Per-instance comparison between backends}
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter.pdf}
+    \caption{Per-instance runtime between baseline and \Calculus{}}
+    \label{fig:duels:baseline}
+    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and baseline]{The plot shows most dots at the bottom of the plot, with some plots collected at the right-hand side.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter.pdf}
+    \caption{Per-instance runtime between baseline and \Calculus{}}
+    \label{fig:duels:nuxmv}
+    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and \Nuxmv{}]{The plot shows flames of dots between the bottom and top, with a line of dots near the right-hand side corresponding to timeouts.}
+  \end{subfigure}
+\end{figure}
+\subsection{Example instance for \Catra{}}
 \begin{lstlisting}[caption={An example input file for \Catra{} for the problem introduced in \cref{sec:motivation}, illustrating every major syntax element.

### 2022-07-08 main.tex
Experiments

@@
 solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
-initial~\InitialNrBenchmarks{}, we remove misgenerated instances that did not
-parse, as well as~\NrTrivial{} instances solved in under five seconds by
-baseline. The benchmarks are run on commit~\texttt{\commit}.
+initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
+five seconds by baseline.
+We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial.
+For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s.
+The benchmarks are run on commit~\texttt{\commit}.
 
@@
 
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances
+In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
 the respective back-ends could solve and with which status. A full summary of
-their outcomes is also available in \cref{tab:solve-status}. We see that
-\Calculus{} generally outperforms \Nuxmv{} on unsatisfiable instances, while
-being worse at satisfiable instances. Both \Nuxmv{} and \Calculus{} strictly
-outperform baseline on every kind of instance, but most of all on satisfiable
-instances. \Cref{tab:solve-status} also shows a small number of bugs in
-\Calculus{}: 27~satisfiable instances are misreported as \Unsat{}, and 20
-additional instances fail with an error. Both problems are due to
-non-deterministic bugs in the clause learning. Note that \Calculus{} is sound:
-it never misreports an unsatisfiable instance as \Sat{}, and all
-\Fudge{\numprint{6218}} satisfying assignments have been verified by \Nuxmv{}.
+their outcomes is also available in \cref{tab:solve-status}. Note that we lack
+ground truth for all instances. We see that \Calculus{} generally outperforms
+\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
+instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
+kind of instance, but most of all on satisfiable instances, and \Calculus{}
+outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
+\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
+caused by a non-deterministically triggered bug in the clause learning of
+\Calculus{}, but have not had time to further investigate it. We have however
+validated all \numprint{3908} satisfying assignments from \Calculus{} with
+\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
+our implementation.
+
 Baseline performs worse on satisfiable instances because it executes a heuristic
 meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances.
+instances. The heuristic is enabled since it is unrealistic to assume that we
+know in advance whether an instance is satisfiable or not, but we do know that
+unsatisfiable instances are more common.
 
@@
     \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
-Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly more in total than lazy.}
+Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
   \end{subfigure}
@@
     \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
-Lazy solves much more instances than nuxmv on shorter runtimes, but is overtaken by just before the 60-second timeout used in these experiments.}
+Lazy solves much more instances than nuxmv.}
   \end{subfigure}
@@
 each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
-\Calculus{} outperforms \Nuxmv{} for shorter timeouts, and that \Nuxmv{} is
-only able to catch up once we increase the timeout to far beyond what would be
-normal for a string solver.
+\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
+long-term scaling properties.
 

### 2022-07-08 main.tex
formalisms

@@
 %% comma separated list
-\keywords{parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
+\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
+We implement \Calculus{} as a theory for the \Princess{} automated theorem
+prover in the tool \Catra. \Catra also supports the approximate method of
+\cite{approximate-parikh}, its fall-back variant adapted
+from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
+checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
+back-ends on \NrBenchmarks{} Parikh automata intersection problems generated by
+\OstrichPlus{} when solving the PyEx string constraint benchmark suite involving
+string length constraints \cite{pyex}.
+
+In summary, we contribute:
+\begin{itemize}
+\item The \Calculus{} calculus to efficiently compute (a homomorphism on) the
+Parikh image of products of Parikh automata.
+\item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future study.
+\item The \Catra{} tool for solving such instances, containing an implementation of \Calculus{}, the over-approximation described in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model checker~\cite{nuxmv}.
+\item Suggestions for how to efficiently implement \Calculus{} in a modern automated theorem prover, including strategies for case splitting, clause learning, and constraint propagation for connectedness.
+\end{itemize}
+
+\subsection{Related Work}
+
 The problem of computing constraints on Parikh images over products of regular
@@
 
-We implement \Calculus{} as a theory for the \Princess{} automated theorem
-prover in the tool \Catra. \Catra also supports the approximate method of
-\cite{approximate-parikh}, its fall-back variant adapted
-from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
-checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
-back-ends on \NrBenchmarks{} Parikh automata intersection problems generated by
-\OstrichPlus{} when solving the PyEx string constraint benchmark suite involving
-string length constraints \cite{pyex}.
-
-In summary, we contribute:
-\begin{itemize}
-\item The \Calculus{} calculus to efficiently compute (a homomorphism on) the
-Parikh image of products of Parikh automata.
-\item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future study.
-\item The \Catra{} tool for solving such instances, containing an implementation of \Calculus{}, the over-approximation described in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model checker~\cite{nuxmv}.
-\item Suggestions for how to efficiently implement \Calculus{} in a modern automated theorem prover, including strategies for case splitting, clause learning, and constraint propagation for connectedness.
-\end{itemize}
-
-\subsection{Related Work}
-
-Parikh image computations as well as Parikh automata \cite{parikh-automata}
-feature extensively in string solvers, including as mentioned above \Ostrich{}
-and \OstrichPlus{} \cite{ostrich,ostrich-plus}, but also formes the basis of
-Trau~\cite{trau-pldi}, and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images
-frequently appear when introducing cardinality constraints like length or string
-indexing. The state-of-the-art approach to handling Parikh image computation is
-to over-approximate the Parikh image of a product of $k$~automata
+Parikh image computations as well as Parikh automata feature extensively in
+string solvers, including as mentioned above \Ostrich{} and \OstrichPlus{}
+\cite{ostrich,ostrich-plus}, but also formes the basis of Trau~\cite{trau-pldi},
+and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images frequently appear when
+introducing cardinality constraints like length or string indexing. The
+state-of-the-art approach to handling Parikh image computation is to
+over-approximate the Parikh image of a product of $k$~automata
 $\ParikhMap(\Automaton_1 \times \ldots \times \Automaton_k)$ with the
@@
 and will require falling back to computing the product of the automata before
-using the standard approach for finding its image originally presented in~\cite{generate-parikh-image}.
+using the standard approach for finding its image originally presented
+in~\cite{generate-parikh-image}.
 
@@
 A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
-consisting of the carrier set of elements, $X$, an associative binary operation
+consisting of the non-empty carrier set of elements, $X$, an associative binary operation
 $X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,

### 2022-07-08 main.tex
intro

@@
 \usepackage{proof}
-\usepackage{cleveref}
+\usepackage[capitalise]{cleveref}
 \usepackage{algorithm2e}      
@@
 We implement \Calculus{} as a theory for the \Princess{} automated theorem
-prover in the tool \Catra. \Catra also supports the approximate method of
+prover in the tool \Catra. \Catra{} also supports the approximate method of
 \cite{approximate-parikh}, its fall-back variant adapted
@@
 string solvers, including as mentioned above \Ostrich{} and \OstrichPlus{}
-\cite{ostrich,ostrich-plus}, but also formes the basis of Trau~\cite{trau-pldi},
+\cite{ostrich,ostrich-plus}, but also forms the basis of Trau~\cite{trau-pldi},
 and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images frequently appear when
@@
 over-approximate the Parikh image of a product of $k$~automata
-$\ParikhMap(\Automaton_1 \times \ldots \times \Automaton_k)$ with the
-conjunction of the automata's parikh maps, $\AndComp{i=1}^{k}
+$\ParikhMap(\Language(\Automaton_1) \cap \ldots \cap \Language(\Automaton_k))$ with the
+conjunction of the automata's parikh maps, $\bigwedge_{i=1}^{k}
 \ParikhMap(\Automaton_i)$. This approach works only for unsatisfiable instances,
@@
 
-Many other generalisations of the Parikh than the ones we use here have been
+Many other generalisations of the Parikh image than the projections we use here have been
 studied. Prominent examples include generalising the Parikh map to segments of a
@@
 $\varphi(\MonoidElement)$ is some set of constraints on $\MonoidElement$:
-\begin{equation*}
-  \AndComp{}{\begin{aligned}
-  & \mathtt{/}\SndRegex{}\mathtt{/} \text{ accepts } s\\
-  & \mathtt{/}\FstRegex{}\mathtt{/} \text{ accepts } s \\
-  & \MonoidElement = f(s) \\
-  & \varphi(\MonoidElement)
-  \end{aligned}}
- \end{equation*}
+\[
+   s \in \Language(\mathtt{/}\SndRegex{}\mathtt{/})  \land
+   s \in \Language(\mathtt{/}\FstRegex{}\mathtt{/})  \land
+   \MonoidElement = f(s)  \land
+   \varphi(\MonoidElement)
+\]
 
@@
   \end{subfigure}
+  \caption{The collection of automata we use as running examples in this paper, with annotations illustrating intermittent states of our calculus.}    
   \end{figure}
@@
 
-\subsection{Producing the Parikh image}\label{sec:introduction:parikh}
+\subsection{Satisfying Constraints on Occurrence of Letters}\label{sec:introduction:parikh}
 

### 2022-07-08 main.tex
Preliminaries

@@
 
-\subsection{Counting string lengths}
+\subsection{Counting String Lengths}
 
@@
 
-\subsection{Satisfying Constraints on Occurrence of Letters}\label{sec:introduction:parikh}
+\subsection{Counting Letters}\label{sec:introduction:parikh}
 
@@
 
-Finally, we sometimes refer to a structure-preserving map between two monoids
-$\Monoid_1, \Monoid_2$ as a \textit{homomorphism} between $\Monoid_1$ and
-$\Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
+Finally, a \textit{homomorphism} is a structure-preserving map between two monoids
+$\Monoid_1, \Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
 \MonoidOp_1 b) = \Map(a) \MonoidOp_2 \Map(b)$, where $S_1, S_2$ are the carrier
@@
 
-We define an alphabet as a set of symbols $\Alphabet$ with words $\Strings$, and
+We define an alphabet as a finite set of symbols $\Alphabet$ with words $\Strings$, and
 the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
@@
 
-A non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$ is
-$\AutomatonTuple$ where $\Transitions = \States \times \Alphabet \times
-\States$, $\States$ is its states, $\InitialState$ is w.l.o.g. assumed to be the
-single initial state, and $\AcceptingStates$ is its set of accepting states.  We
+An automaton~$\Automaton$ with alphabet~$\Alphabet$ is
+$\AutomatonTuple$ where $\Transitions \subseteq \States \times \Alphabet \times
+\States$, $\States$ is its states, $\InitialState$ its
+initial state, and $\AcceptingStates$ is its set of accepting states.  We
 write a transition $\Transition = \Tuple{\State, \Label, \State'} \in

### 2022-07-08 main.tex
3.1 klar

@@
 Since Parikh images are semilinear, any Parikh image can be written as a set of linear
-equations. The following construction, here adapted to work on an NFA $\Automaton = \AutomatonTuple$, was presented in \cite{generate-parikh-image}:
+equations. The following construction for an NFA $\Automaton = \AutomatonTuple$ is based on the version for context-free grammars presented in \cite{generate-parikh-image}:
 \begin{equation}
@@
   \StateVar_{\State} > 0 \longrightarrow
-  \left(\FinalStateVar_{\State} = 1 \text{ if } \State \in \AcceptingStates \land \right) \OrComp{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions}{
+  \OrComp{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions}{
     \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
@@
 
-All variables $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ are
+All variables $\TransitionVar_\Transition, \StateVar_\State, \FinalStateVar_\State$ are
 existentially quantified and the free variables $\LetterVar_\Letter$ make up the

### 2022-07-08 main.tex
Fitting till preliminaries

@@
   \left(\FinalStateVar_{\State} \text{ if } \State \in \AcceptingStates \text{ otherwise } 0 \right) +
-  \sum_{\Transition = \FromLabelTo{\State}{}{}} \Filter(\Transition) - \sum_{\Transition = \FromLabelTo{}{}{\State}} \Filter(\Transition)
+  \sum_{\Transition \in \FromLabelTo{\State}{}{}} \Filter(\Transition) - \sum_{\Transition \in \FromLabelTo{}{}{\State}} \Filter(\Transition)
 = \begin{cases}
@@
   \StateVar_{\State} > 0 \longrightarrow
-  \OrComp{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions}{
+  \OrComp{\Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions}{
     \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
@@
 
-\section{Projections on Parikh Maps}\label{sec:generalised}
+\subsection{Theorem Proving}
+
+We assume basic familiarity with theorem provers, in particular sequent calculus
+or similar. For reference, see e.g. \cite{Fitting96a}.
+
+\section{Projections on Parikh Images}\label{sec:generalised}
 
@@
 interested in the full Parikh image, but only in some projections of
-it, for the purposes of this paper we use a generalised version of the
-Parikh map. We consider \emph{arbitrary} homomorphisms
+it. We consider \emph{arbitrary} homomorphisms
 $\Map:\: \Sigma^* \to \Monoid$, where
 $\Monoid = \left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative
-monoid; this means that $\Map(\epsilon) = 0_{\Monoid}$ and
-$\Map(u \Concat w) = \Map(u) \MonoidOp \Map(w)$. We give several
-examples of such generalised Parikh image later in this section.
+monoid. We give several
+examples of such projections on Parikh images later in this section.
 

### 2022-07-08 main.tex
rewrote definitions of product images

--- 2022-07-08 main.tex
@@
 \usepackage{proof}
-\usepackage[capitalise]{cleveref}
+\usepackage{cleveref}
 \usepackage{algorithm2e}      
@@
 %% comma separated list
-\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
+\keywords{parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
-We implement \Calculus{} as a theory for the \Princess{} automated theorem
-prover in the tool \Catra. \Catra{} also supports the approximate method of
-\cite{approximate-parikh}, its fall-back variant adapted
-from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
-checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
-back-ends on \NrBenchmarks{} Parikh automata intersection problems generated by
-\OstrichPlus{} when solving the PyEx string constraint benchmark suite involving
-string length constraints \cite{pyex}.
-
-In summary, we contribute:
-\begin{itemize}
-\item The \Calculus{} calculus to efficiently compute (a homomorphism on) the
-Parikh image of products of Parikh automata.
-\item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future study.
-\item The \Catra{} tool for solving such instances, containing an implementation of \Calculus{}, the over-approximation described in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model checker~\cite{nuxmv}.
-\item Suggestions for how to efficiently implement \Calculus{} in a modern automated theorem prover, including strategies for case splitting, clause learning, and constraint propagation for connectedness.
-\end{itemize}
-
-\subsection{Related Work}
-
 The problem of computing constraints on Parikh images over products of regular
@@
 
-Parikh image computations as well as Parikh automata feature extensively in
-string solvers, including as mentioned above \Ostrich{} and \OstrichPlus{}
-\cite{ostrich,ostrich-plus}, but also forms the basis of Trau~\cite{trau-pldi},
-and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images frequently appear when
-introducing cardinality constraints like length or string indexing. The
-state-of-the-art approach to handling Parikh image computation is to
-over-approximate the Parikh image of a product of $k$~automata
-$\ParikhMap(\Language(\Automaton_1) \cap \ldots \cap \Language(\Automaton_k))$ with the
-conjunction of the automata's parikh maps, $\bigwedge_{i=1}^{k}
+We implement \Calculus{} as a theory for the \Princess{} automated theorem
+prover in the tool \Catra. \Catra also supports the approximate method of
+\cite{approximate-parikh}, its fall-back variant adapted
+from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
+checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
+back-ends on \NrBenchmarks{} Parikh automata intersection problems generated by
+\OstrichPlus{} when solving the PyEx string constraint benchmark suite involving
+string length constraints \cite{pyex}.
+
+In summary, we contribute:
+\begin{itemize}
+\item The \Calculus{} calculus to efficiently compute (a homomorphism on) the
+Parikh image of products of Parikh automata.
+\item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future study.
+\item The \Catra{} tool for solving such instances, containing an implementation of \Calculus{}, the over-approximation described in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model checker~\cite{nuxmv}.
+\item Suggestions for how to efficiently implement \Calculus{} in a modern automated theorem prover, including strategies for case splitting, clause learning, and constraint propagation for connectedness.
+\end{itemize}
+
+\subsection{Related Work}
+
+Parikh image computations as well as Parikh automata \cite{parikh-automata}
+feature extensively in string solvers, including as mentioned above \Ostrich{}
+and \OstrichPlus{} \cite{ostrich,ostrich-plus}, but also formes the basis of
+Trau~\cite{trau-pldi}, and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images
+frequently appear when introducing cardinality constraints like length or string
+indexing. The state-of-the-art approach to handling Parikh image computation is
+to over-approximate the Parikh image of a product of $k$~automata
+$\ParikhMap(\Automaton_1 \times \ldots \times \Automaton_k)$ with the
+conjunction of the automata's parikh maps, $\AndComp{i=1}^{k}
 \ParikhMap(\Automaton_i)$. This approach works only for unsatisfiable instances,
 and will require falling back to computing the product of the automata before
-using the standard approach for finding its image originally presented
-in~\cite{generate-parikh-image}.
+using the standard approach for finding its image originally presented in~\cite{generate-parikh-image}.
 
@@
 
-Many other generalisations of the Parikh image than the projections we use here have been
+Many other generalisations of the Parikh than the ones we use here have been
 studied. Prominent examples include generalising the Parikh map to segments of a
@@
 $\varphi(\MonoidElement)$ is some set of constraints on $\MonoidElement$:
-\[
-   s \in \Language(\mathtt{/}\SndRegex{}\mathtt{/})  \land
-   s \in \Language(\mathtt{/}\FstRegex{}\mathtt{/})  \land
-   \MonoidElement = f(s)  \land
-   \varphi(\MonoidElement)
-\]
+\begin{equation*}
+  \AndComp{}{\begin{aligned}
+  & \mathtt{/}\SndRegex{}\mathtt{/} \text{ accepts } s\\
+  & \mathtt{/}\FstRegex{}\mathtt{/} \text{ accepts } s \\
+  & \MonoidElement = f(s) \\
+  & \varphi(\MonoidElement)
+  \end{aligned}}
+ \end{equation*}
 
@@
 
-\subsection{Counting String Lengths}
+\subsection{Counting string lengths}
 
@@
   \end{subfigure}    
-  \caption{The collection of automata we use as running examples in this paper, with annotations illustrating intermittent states of our calculus.}    
   \end{figure}
@@
 
-\subsection{Counting Letters}\label{sec:introduction:parikh}
+\subsection{Producing the Parikh image}\label{sec:introduction:parikh}
 
@@
 A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
-consisting of the non-empty carrier set of elements, $X$, an associative binary operation
+consisting of the carrier set of elements, $X$, an associative binary operation
 $X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,
@@
 
-Finally, a \textit{homomorphism} is a structure-preserving map between two monoids
-$\Monoid_1, \Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
+Finally, we sometimes refer to a structure-preserving map between two monoids
+$\Monoid_1, \Monoid_2$ as a \textit{homomorphism} between $\Monoid_1$ and
+$\Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
 \MonoidOp_1 b) = \Map(a) \MonoidOp_2 \Map(b)$, where $S_1, S_2$ are the carrier
@@
 
-We define an alphabet as a finite set of symbols $\Alphabet$ with words $\Strings$, and
+We define an alphabet as a set of symbols $\Alphabet$ with words $\Strings$, and
 the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
@@
 
-An automaton~$\Automaton$ with alphabet~$\Alphabet$ is
-$\AutomatonTuple$ where $\Transitions \subseteq \States \times \Alphabet \times
-\States$, $\States$ is its states, $\InitialState$ its
-initial state, and $\AcceptingStates$ is its set of accepting states.  We
+A non-deterministic automaton~$\Automaton$ with alphabet~$\Alphabet$ is
+$\AutomatonTuple$ where $\Transitions = \States \times \Alphabet \times
+\States$, $\States$ is its states, $\InitialState$ is w.l.o.g. assumed to be the
+single initial state, and $\AcceptingStates$ is its set of accepting states.  We
 write a transition $\Transition = \Tuple{\State, \Label, \State'} \in
@@
   The Parikh predicate, $\SinglePredicateInstance$, for some automaton
-  $\Automaton = \AutomatonTuple$, modulo some map $\Map$ to a
+  $\Automaton = \AutomatonTuple$, modulo some map $\Map : \Alphabet^* \to \Monoid$ to a
   commutative monoid $\Monoid$ as described in \cref{sec:generalised}
@@
   $\Filter:\: \Transitions \to \Naturals$ holds when
-  $\MonoidElement \in \Monoid$ is the Parikh image of $\Automaton$
+  $\MonoidElement \in \Monoid$ is an element of the Parikh image of
+  $\Language(\Automaton)$
   modulo $\Map$, or more formally when there is an accepting path
@@
     \inferrule
-  {\Connected(\Automaton, \Filter) \land \FlowEq(\Automaton, \Filter) \land \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
+  {\Connected(\Automaton, \Filter) ,~ \FlowEq(\Automaton, \Filter) ,~ \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
   {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
@@
 
-To generalise the calculus to calculations on products of automata, we change the main predicate to take arbitrarily many automata:
+We now generalise our calculus to natively work with intersections of
+regular languages, or equivalently products of automata. For this
+extension, we change the main predicate~$\Image$ to be indexed by a
+vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
+simplicity, we assume that the sets of states of the $k$ automata (and
+therefore also the transition sets) are pairwise disjoint. 
 \begin{definition}\label{def:multiple}
-  $\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}$
-  is true exactly when the single-automaton version of the predicate would hold
-  for the automaton.
+  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
+  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
+  $\Monoid$,
+  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
+  is a transition selection function, and
+  $\MonoidElement \in \Monoid$. The predicate
+  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
+      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
+  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
+  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
+  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
+  \begin{itemize}
+  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
+    $\Filter$, that is,
+    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
+  \item the automata all accept the same
+    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
+  \item the accepted word is mapped to
+    $\MonoidElement = \Map(\WordOf(\Path_i)) =
+    \Map(\WordOf(\Path_1))$.
+  \end{itemize}
 \end{definition}
-
-  For the calculus, we first extend $\Expand$ to generate flow equations and instances of $\Connected$ for each automaton.
 
@@
       {
-        {\begin{matrix}
-          \Set{ 
+       {\begin{array}{c}
+           \left\{
+           \FlowEq(\Automaton_i, \Filter),~~
+           \Connected(\Automaton_i, \Filter),~~
            \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
-          \SuchThat \Automaton_i \in \Automaton_1,\ldots,\Automaton_k}, \\
-          \Set{\FlowEq(\Automaton_i), \Connected(\Automaton_i) \SuchThat \Automaton_i \in \Automaton_1,\ldots,\Automaton_k}, \\
-          \ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \\
+           \right\}_{i=1}^k,
+           \\[4ex]
+          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
           \SomeInequalities, \SomeClause
-        \end{matrix}}
-        }
-      {\ImagePredicate{\Automaton_1\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
+        \end{array}}
+        }
+      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
       None \\[5ex]
@@
       \inferrule
-      {    \BindingSum(\Automaton', \Filter),\ImagePredicate{\Automaton'\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause}
-      {\ImagePredicate{\Automaton_1\times\Automaton_2\times\ldots\times\Automaton_k}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} &
-      $\Automaton' = \Automaton_1\times\Automaton_2$ \\
+      {{
+         \begin{array}{c}
+           \FlowEq(\Automaton', \Filter'),~
+           \Connected(\Automaton', \Filter'),~
+           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
+           \\\ImagePredicate{
+                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
+                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
+                     \Automaton'},~ \SomeInequalities, \SomeClause
+           }{\Map}{\Filter'}{\MonoidElement}
+           \end{array}}
+           }
+      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
+      $\begin{array}{@{}l@{}}
+         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
+         1 \leq i < j \leq k, \\
+      \Automaton' = \Automaton_i \times \Automaton_j,\\
+      \end{array}$ \\
       \bottomrule
@@
 
-\ExpandM{} must be applied before any other rule, like \Expand{}, but unlike \Expand{}, \ExpandM{} does not remove the $\Image$~predicate since it is needed to keep track of the product.
-
-Before we can begin to define our final rule, we need to talk about product states.
-
-Then we introduce the rule $\Materialise$, used to compute a partial product between two terms:
-
-  With the following helper symbolic function:
-
+    For the calculus (\cref{tbl:rules:multi}), we first extend
+    $\Expand$ to generate flow equations and instances of $\Connected$
+    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
+    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
+    since it is needed to keep track of the currently considered
+    partial products.
+
+The rule~$\Materialise$ introduces the product of two individual
+automata~$\Automaton_i, \Automaton_j$; this step eliminates
+$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
+instead adds the product~$\Automaton_i \times \Automaton_j$, while
+also introducing the flow equations and the $\Connected$ predicate.
+
+The rule~$\Materialise$ also has to connect the newly introduced
+product~$\Automaton_i \times \Automaton_j$ to the previous automata
+$\Automaton_i, \Automaton_j$. This is done by extending the
+selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
+to fresh variables~$x_{\Transition}$:
+\begin{equation*}
+  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
+  ~~=~~
+  \begin{cases}
+    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
+    \\
+    \Filter(t) & \text{otherwise}
+  \end{cases}
+\end{equation*}
+
+The multiplicity of transitions in the product then has to be related
+to the multiplicities in the individual automata, modelled using the
+$\BindingSum$ predicate. The predicate expresses that the multiplicity
+of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
+$\Automaton_i$ has to coincide with the sum of the multiplicities of
+transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
+and similarly for $\Automaton_j$:
+%
   $$
-  \BindingSum(\Automaton_1 \times \Automaton_2, \Filter) = \bigcup
+  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
   \begin{aligned}
   & \Set{ 
-    \left<\Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter(\Transition')\right>
-  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_1} } , \\ 
+    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
+  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
   & \Set{
-    \left<\Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter(\Transition')\right> \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_2}
+    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
   }
@@
 
+\iffalse
 Note that this definition implies that $\Filter(\Transition) =
@@
 $\Filter$ to fresh terms for each transition in the product.
-
-Finally, for instances of precisely one automaton, neither rule applies and we
+\fi
+
+For instances of precisely one automaton, neither rule applies and we
 perform the calculus as before.
@@
 solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
-initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
-five seconds by baseline.
-We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial.
-For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s.
-The benchmarks are run on commit~\texttt{\commit}.
+initial~\InitialNrBenchmarks{}, we remove misgenerated instances that did not
+parse, as well as~\NrTrivial{} instances solved in under five seconds by
+baseline. The benchmarks are run on commit~\texttt{\commit}.
 
@@
 
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
+In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances
 the respective back-ends could solve and with which status. A full summary of
-their outcomes is also available in \cref{tab:solve-status}. Note that we lack
-ground truth for all instances. We see that \Calculus{} generally outperforms
-\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
-instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
-kind of instance, but most of all on satisfiable instances, and \Calculus{}
-outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
-\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
-caused by a non-deterministically triggered bug in the clause learning of
-\Calculus{}, but have not had time to further investigate it. We have however
-validated all \numprint{3908} satisfying assignments from \Calculus{} with
-\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
-our implementation.
-
+their outcomes is also available in \cref{tab:solve-status}. We see that
+\Calculus{} generally outperforms \Nuxmv{} on unsatisfiable instances, while
+being worse at satisfiable instances. Both \Nuxmv{} and \Calculus{} strictly
+outperform baseline on every kind of instance, but most of all on satisfiable
+instances. \Cref{tab:solve-status} also shows a small number of bugs in
+\Calculus{}: 27~satisfiable instances are misreported as \Unsat{}, and 20
+additional instances fail with an error. Both problems are due to
+non-deterministic bugs in the clause learning. Note that \Calculus{} is sound:
+it never misreports an unsatisfiable instance as \Sat{}, and all
+\Fudge{\numprint{6218}} satisfying assignments have been verified by \Nuxmv{}.
 Baseline performs worse on satisfiable instances because it executes a heuristic
 meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances. The heuristic is enabled since it is unrealistic to assume that we
-know in advance whether an instance is satisfiable or not, but we do know that
-unsatisfiable instances are more common.
+instances.
 
@@
   \input{graphs/solved_pivot_table.tex}
-  \caption{The result of running the respective back-ends with a timeout of
+  \caption{The result of running the respective back-ends by instance
+  satisifiability (satisifiable or unsatisifiable) with a timeout of
   \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
@@
     \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
-Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
+Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly more in total than lazy.}
   \end{subfigure}
@@
     \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
-Lazy solves much more instances than nuxmv.}
+Lazy solves much more instances than nuxmv on shorter runtimes, but is overtaken by just before the 60-second timeout used in these experiments.}
   \end{subfigure}
@@
 each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
-\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
-long-term scaling properties.
+\Calculus{} outperforms \Nuxmv{} for shorter timeouts, and that \Nuxmv{} is
+only able to catch up once we increase the timeout to far beyond what would be
+normal for a string solver.
 
@@
 \section{Appendix}
-\subsection{Per-instance comparison between backends}
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter.pdf}
-    \caption{Per-instance runtime between baseline and \Calculus{}}
-    \label{fig:duels:baseline}
-    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and baseline]{The plot shows most dots at the bottom of the plot, with some plots collected at the right-hand side.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter.pdf}
-    \caption{Per-instance runtime between baseline and \Calculus{}}
-    \label{fig:duels:nuxmv}
-    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and \Nuxmv{}]{The plot shows flames of dots between the bottom and top, with a line of dots near the right-hand side corresponding to timeouts.}
-  \end{subfigure}
-\end{figure}
-\subsection{Example instance for \Catra{}}
+
 \begin{lstlisting}[caption={An example input file for \Catra{} for the problem introduced in \cref{sec:motivation}, illustrating every major syntax element.

### 2022-07-08 main.tex
updated notation
Single example

@@
 \begin{prooftree}
-  \hypo{\bigstar}
-  \infer1[]{
+  \hypo[]{
     \begin{matrix}
@@
   }
-  % BRANCH:  = 0
-  \hypo{\bigstar}
-  \infer1[]{
+  % BRANCH:  = 0'
+  \hypo{
     \begin{matrix}
-      \TransitionVar_3 > 0 \\
+      \TransitionVar_3 > 0 \land \\
       \TransitionVar_6 = 0 \land \\
@@
       \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_3 > = 0 \\
+      \TransitionVar_3 > = 0 \land \\
       \TransitionVar_6 = 0 \land \\
@@
 using the (simplified) flow equations. We can also see how interleaving
-reasoning on the corresponding linear equations helps the production. In both
+reasoning on the corresponding linear equations helps the deduction. In both
 branches the reasoning is similar: we conclude that when choosing either path at
@@
 finally subsume, removing the $\Connected{}$ predicate when we are unable to use
-either of the rules, leaving only the satisfiable remnants of the flow equations
-at the respective leaves marked with~$\bigstar$. To obtain a solution from the
+either of the rules, leaving only a set of linear equations. To obtain a solution from the
 proof, one would need to perform standard model generation on the remaining
@@
 Parikh image is the disjunction of the constraints left at the two
-$\bigstar$-marked leaves, after quantifier elimination.
+leaves after quantifier elimination.
+
+Following the values of $\Filter$, we can read out the path corresponding to our solution in the product, and with that a word.
+In our case, for the right branch we have $\Filter(\FromLabelTo{S}{b}{B}) = 0,
+\Filter(\FromLabelTo{A}{c}{A})  > 0,
+\Filter(\FromLabelTo{B}{b}{S}) = 0,
+\Filter(\FromLabelTo{S}{a}{A}) = 1 ,
+\Filter(\FromLabelTo{A}{a}{F}) = 1 ,
+\Filter(\FromLabelTo{B}{c}{F}) = 0$.
+We can read out a word along it: aca, accca, etc, and the corresponding constraint: $\exists k : \TransitionVar_3 = 2k -1 \land \TransitionVar_3 > 0 \land k \geq 0$.
 
@@
       \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
+      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
       \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
@@
     \Connected(\SomethingCSomething{}, \Filter) \land \\
-    \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
+    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
     \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
@@
       \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
+      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
       \VectorLiteral{a, b, c}) \land a > b
@@
   }
-  \infer1[\ExpandM]{\Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
+  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
 \end{prooftree}

### 2022-07-08 main.tex
Fix fudge
cleanup

@@
 $\Connected{}$ predicate. This, again, leaves us with a set of linear
-inequalities that we can use to obtain a model of our final $\MonoidElement$
-values (now a vector). Performing quantifier elimination on the remaining clauses would also produce the Presburger formula of the Parikh image under the constraint we started with.
+inequalities at the leaf that we can use to obtain a model of our final
+$\MonoidElement$ values (now a vector).
 
@@
 \begin{prooftree}
-  \hypo{\bigstar}
-  \infer1[]{
+  \hypo{
     \begin{matrix}
@@
 like the ones in a range label. In \Catra{}, we frontload this problem by
-requiring the user to encode their input as a Parikh automaton. \Fudge{How is
-the encoding of Ostrich Plus automata done?}
+requiring the user to encode their input as a Parikh automaton. For an example of how automata can be encoded, see \cref{sec:parikh-automata}.
 

### 2022-07-08 main.tex
updated proofs

--- 2022-07-08 main.tex
@@
 Since \cref{tbl:rules:multi} only extends the existing rules of
-\cref{tbl:rules:single}, we only have to prove termination and solution
-preservation for the new rules.
-
-\subsubsection{\Calculus{} for products of automata terminates}
+\cref{tbl:rules:single}, we focus on the differences compared
+to the calculus for a single automaton.
+
+%\subsubsection{\Calculus{} for products of automata terminates}
 \begin{lemma}\label{lma:multi-terminates}
@@
 \begin{proof}
-  The rule~\ExpandM{} can only be applied finitely often since its side condition only allows applying it once per instance of an $\Image$ predicate containing a product, and the only rule that introduces a new $\Image$ predicate with a product, $\Materialise$ does so with non-overlapping automata from the $\Image$ predicate it removes.
-Therefore, the number of automata in $\Image$ predicates is strictly decreasing.
-
   The rule \Materialise{} can similarly only be used finitely many times, as
@@
   single-automaton instances applies.
+
+  This implies that also the rule~\ExpandM{} can only be applied
+  finitely often since its side condition only allows applying it once
+  per instance of an $\Image$ predicate containing a product, and only finitely
+  many instances of $\Image$ can be introduced on each branch.
 \end{proof}
 
-\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
-
+%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
+
+Since our calculus now includes a rule introducing new variables, the
+\Materialise{} rule, we have to slightly generalise the notion of
+solution-preservation:
+%
 \begin{lemma}\label{lma:multi-correct}
@@
   premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
-  conclusion~$\SomeClause$ if and only if it satisfies one of the
+  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
+  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
+  there is an extension~$\beta' of \beta$ satisfying one of the
   premises~$\SomeClause_i$.
@@
 \begin{proof}
-  This property has to be shown by analysing the possible applications
-  of each proof rule.
-
-  \ExpandM{} applies \Expand{} to every automata $\Automaton_i$ of the product. 
-  \begin{itemize}
-  \item Assume that $\beta$ satisfies the conclusion, which means that
-    there is some accepting path in the product $\Automaton_1 \times \ldots \times \Automaton_k$ such that
-    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton_1 \times \ldots \times \Automaton_k)}$ with
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$ and $\val_\beta(\MonoidElement) =
-    \Map(\WordOf(\Path))$.
-That immediately implies that every automaton $\Automaton_i$ would satisfy its $\Image{}$ constraints individually, and hence also the constraints generated by \ExpandM{}.
-\Fudge{Missing Materialise!}
-  \item Assume that $\beta$ satisfies the premise, which implies that
-    $\val_\beta(\Filter)$ describes a consistent, connected flow of
-    the automaton. By the same argument as in
-    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
-    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
-    that each transition~$\Transition$ occurs on $\Path$ exactly
-    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
-  \end{itemize}
-
-  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
-  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
-  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
-  be satisfied, implying the property.
-
-  For \Propagate{}, suppose that
-  $\SeparatingCut(C, \Automaton, \State)$, which means that every
-  accepting path visiting~$q$ contains at least one of the transitions
-  in $C$. For a $\beta$ satisfying $\Connected(\Automaton, \Filter)$
-  and $\Filter(\Transition') = 0$ for $\Transition' \in C$, this means
-  that also $\val_\beta(\Filter(\Transition)) = 0$ has to hold.
-
-  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
-  then a goal must contain~$\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
-  formulas in $\SomeInequalities$ are inconsistent, an application of
-  \Subsume{} is trivially solution-preserving; therefore assume that
-  $\SomeInequalities$ is consistent, which means that it contains
-  exactly one of $\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for each $\Transition$.
+  We have to consider the two new rules in \cref{tbl:rules:multi}. The
+  result is immediate for \ExpandM{}, since this rule does not remove the
+  $\Image$ predicate from a proof goal, and the newly introduced formulas
+  are all implied by the $\Image$ predicate.
+
+  For \Materialise{}, observe that the existence of an accepting path
+  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
+  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
+  same word. The path in the product will satisfy the flow equations
+  and connectedness, and it will be related to the individual paths as
+  stipulated by the \BindingSum{} predicate.
 \end{proof}

### 2022-07-08 main.tex
Example 2
Samma skala för alla automator

@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{aca_or_bc}
+    \centering
+    \includegraphics[scale=\autscale]{aca_or_bc}
     \caption{An automaton recognising the regular expression
@@
   \begin{subfigure}[b]{0.49\textwidth} 
-    \includegraphics[width=\textwidth]{something_c_something}
+    \centering
+    \includegraphics[scale=\autscale]{something_c_something}
     \caption{An automaton recognising the regular expression
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{aca_or_bc_simplified}
+    \centering
+    \includegraphics[scale=\autscale]{aca_or_bc_simplified}
     \caption{$\AcaOrBc{}$ with its symbolic transition counters after
@@
   \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{something_c_something_length_simplified}
+    \centering
+    \includegraphics[scale=\autscale]{something_c_something_length_simplified}
     \caption{$\SomethingCSomething{}$ after using linear reasoning over lengths to replace a variable.}\label{fig:example-length-reduced}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-      \includegraphics[width=\textwidth]{aca_or_bc_length_split_1}
+    \centering
+      \includegraphics[scale=\autscale]{aca_or_bc_length_split_1}
       \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ is not
@@
   \begin{subfigure}[b]{0.49\textwidth}
-      \includegraphics[width=\textwidth]{aca_or_bc_length_split_2}
+    \centering
+      \includegraphics[scale=\autscale]{aca_or_bc_length_split_2}
       \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ \emph{is} used.}\label{fig:aca-or-bc-length-split-2}
@@
   \begin{subfigure}[b]{0.49\textwidth}
-      \includegraphics[width=\textwidth]{length_split_product}
+    \centering
+      \includegraphics[scale=\autscale]{length_split_product}
       \caption{The product of $\AcaOrBc$ and $\SomethingCSomething$ after
@@
   \begin{subfigure}[b]{0.49\textwidth}
-      \includegraphics[width=\textwidth]{aca_or_bc_simplified_3}
+    \centering
+      \includegraphics[scale=\autscale]{aca_or_bc_simplified_3}
       \caption{$\AcaOrBc{}$ after further constraint propagation.}\label{fig:example-simplify-3}
@@
   \begin{subfigure}[b]{0.6\textwidth}
-        \includegraphics[width=\textwidth]{aca_or_bc_times_something_c_something}
+    \centering
+        \includegraphics[scale=\autscale]{aca_or_bc_times_something_c_something}
         \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ under the constraint
@@
 inequalities at the leaf that we can use to obtain a model of our final
-$\MonoidElement$ values (now a vector).
+$\MonoidElement$ values (now a vector). This time we can read out aca, or
+$\VectorLiteral{2, 0, 1}$.
 

### 2022-07-08 main.tex
how to find solutions

--- 2022-07-08 main.tex
@@
 necessary to reason about the sets of word lengths induced by
-$\Language_1, \Language_2$, which is a special case of the Parikh
+$\Language_1, \Language_2$, which are special cases of the Parikh
 image.  This required combined reasoning about strings and string
@@
 quantifier-free representation of the Parikh image has exponential
-complexity~\cite{DBLP:conf/issac/Weispfenning97}, and is in practice often impossible in reasonable
+complexity~\cite{DBLP:conf/issac/Weispfenning97}, and is often impossible in reasonable
 time. Just solving the Parikh image membership problem is NP-complete,
@@
 
-We assume basic familiarity with theorem provers, in particular sequent calculus
-or similar. For reference, see e.g. \cite{Fitting96a}.
+We assume basic familiarity with first-order logic, Presburger
+arithmetic, and the classical sequent calculus. For reference, see
+e.g. \cite{Fitting96a}.
 
@@
 We present the rules of \Calculus{} for one automaton in
-\cref{tbl:rules:single}. Note that we operate on sets of
-formulas. Additionally, we also use the convention of splitting the
-formulas into linear equations and inequalities ($\SomeInequalities$)
-and other formulas ($\SomeClause$). We assume that our predicates $\Image$ and $\Connected$ only occur positively.
-
+\cref{tbl:rules:single}. The rules operate on sets of formulas, and
+can be interpreted as rules of a one-sided sequent calculus, in
+which all formulas are located in the antecedent~\cite{Fitting96a}.
+Each of the rules should be read and
+applied bottom-up, and relates
+premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
+conclusion~$\SomeClause$. When constructing a proof, we start from some
+root~$\SomeClause$, and then apply proof rules to the goals of the proof
+until a goal can be closed, or no more rule is applicable.
+
+A proof in which all proof goals could be closed shows that the formulas
+in the root~$\SomeClause$ of the proof are inconsistent, and do not
+have any solutions. A proof goal that could not be closed, but to which no
+more rules are applicable, gives rise to  a solution of the formulas in the
+root~$\SomeClause$. Such an open proof goal will only contain formulas
+in Presburger arithmetic, so that a solution can be computed using
+standard algorithms~\cite{Fitting96a}.
+
+We use the convention of splitting the
+formulas into linear (in)equalities ($\SomeInequalities$)
+and other formulas ($\SomeClause$), and assume that predicates
+$\Image$ and $\Connected$ only occur positively.
+ Since
+our rules operate by adding and matching linear (in)equalities in a proof goal,
+we use the shorthand of listing the matched inequalities as antecedents
+e.g., in \Propagate{}).
 The filtering function~$\Filter$ is evaluated symbolically, and can in practice
@@
 $\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$.
-
-Each of the rules of \cref{tbl:rules:single} should be read and
-applied bottom-up, and relates
-premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
-conclusion~$\SomeClause$. When constructing a proof, we start from some
-root~$\SomeClause$, and then apply proof rules to the goals of the proof
-until a goal can be closed, or no more rule is applicable. Since
-our rules operate by adding and matching linear (in)equalities in a proof goal,
-we use the shorthand of listing the matched inequalities as antecedents. An
-example of this can be seen in the~\Propagate{} rule. \Fudge{We pull solutions
-out of a hat? I have no idea.}
 

### 2022-07-08 main.tex
removed an unnecessary footnote

@@
 as Presburger arithmetic in a C-like syntax. We call this program
-\Catra.\footnote{If you really must read it as an acronym, please read it as
-CAtegory Theory on Register Automata, or if you object to the somewhat
-nonstandard use of register automata and category theory, as Check Assignments
-of The Registers Afterwards. Or alternatively if you find it all to be too much
-of a theoretical exercise, as Can Anyone Think of A Real Application.}
+\Catra.
+% \footnote{If you really must read it as an acronym, please read it as
+%CAtegory Theory on Register Automata, or if you object to the somewhat
+%nonstandard use of register automata and category theory, as Check Assignments
+%of The Registers Afterwards. Or alternatively if you find it all to be too much
+%of a theoretical exercise, as Can Anyone Think of A Real Application.}
 

### 2022-11-28 main.tex
Its ALIGNED
rename MATHS, fix alignment in examples
Align 2nd example
Some uncommitted work, apparently

@@
 \begin{aligned}
-  & \FlowEq(\Automaton, \Filter) = \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
+  \FlowEq(\Automaton, \Filter) &= \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
   \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}\\
-  & \Sink(\State) = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
-  & \In(\State, \Filter) = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
-  & \StartFlow(\State)  = 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
-  & \Out(\State, \Filter) = \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
+  \Sink(\State) & = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
+  \In(\State, \Filter) & = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
+  \StartFlow(\State)  &= 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
+  \Out(\State, \Filter) &= \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
 \end{aligned}
@@
   equations, and then apply \EquationReasoning{} (under the implicit assumption
-  that every RHS is $\geq 0$) to obtain the equivalent third definition :
+  that every RHS is $\geq 0$) to obtain the third definition:
   \begin{minipage}[b]{0.3\linewidth}
@@
       \begin{aligned}
-        & \Filter(\FromLabelTo{S}{a}{A}) & = \TransitionVar_1 \\
-        & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_2 \\
-        & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        & \Filter(\FromLabelTo{B}{b}{S}) & = \TransitionVar_4 \\
-        & \Filter(\FromLabelTo{A}{a}{F}) & = \TransitionVar_5 \\
-        & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
+        \Filter(\FromLabelTo{S}{a}{A}) & = \TransitionVar_1 \\
+        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_2 \\
+        \Filter(\FromLabelTo{A}{c}{A}) & = \TransitionVar_3  \\
+        \Filter(\FromLabelTo{B}{b}{S}) & = \TransitionVar_4 \\
+        \Filter(\FromLabelTo{A}{a}{F}) & = \TransitionVar_5 \\
+        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
       \end{aligned}
@@
         % S
-        &  \TransitionVar_4 = \TransitionVar_5 + \TransitionVar_2 - 1 \\
+        \TransitionVar_4 &= \TransitionVar_5 + \TransitionVar_2 - 1 \\
         % A
-        & \TransitionVar_1 = \TransitionVar_5 \\
+        \TransitionVar_1 &= \TransitionVar_5 \\
         % B
-        & \TransitionVar_2 = \TransitionVar_4 + \TransitionVar_6 \\
+        \TransitionVar_2 &= \TransitionVar_4 + \TransitionVar_6 \\
         % F
-        & \TransitionVar_5 + \TransitionVar_6 = \FinalStateVar_1 \\
-        & \FinalStateVar_1 = 1 \\
+        \TransitionVar_5 + \TransitionVar_6 &= \FinalStateVar_1 \\
+        \FinalStateVar_1 &= 1 \\
       \end{aligned}  
@@
       \begin{aligned}
-        & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
+        \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
+        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
+        \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
+        \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
+        \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
+        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
       \end{aligned}
@@
   \hypo[]{
-    \begin{matrix}
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 = 0 \land \\
-      2 + \TransitionVar_4 > 0 \land \\
+{    \begin{array}{lll}
+      1 - \TransitionVar_6 = 0 & \land \TransitionVar_6 = 1 & \land  \TransitionVar_4 + 1 > 0 \\
+      \land \TransitionVar_3 = 0 &\land  2 + \TransitionVar_4 > 0 & \land k = 1 + \TransitionVar_4 
+    \end{array}}
+  }
+  \infer1[\Subsume{}]{
+{    \begin{array}{lll}
+      \Connected(\AcaOrBc{}, \Filter) & \land 1 - \TransitionVar_6 = 0 \\
+      \land \TransitionVar_6 = 1 & \land \TransitionVar_4 + 1 > 0 \land \\
+      \land \TransitionVar_3 = 0 & \land 2 + \TransitionVar_4 > 0 \land \\
       k = 1 + \TransitionVar_4 
-    \end{matrix}
-  }
-  \infer1[\Subsume{}]{
-    \begin{matrix}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 = 0 \land \\
-      2 + \TransitionVar_4 > 0 \land \\
-      k = 1 + \TransitionVar_4 
-    \end{matrix}
+    \end{array}}
   }
   \infer1[\Propagate]{
-    \begin{matrix}
+    \begin{aligned}
       \Connected(\AcaOrBc{}, \Filter) \land \\
@@
       2k = 2 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{matrix}
+    \end{aligned}
   }
   \infer1[\EquationReasoning{}]{
-    \begin{matrix}
+    \begin{aligned}
       \Connected(\AcaOrBc{}, \Filter) \land \\
@@
       2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{matrix}
+    \end{aligned}
   }
@@
   \hypo{
-    \begin{matrix}
-      \TransitionVar_3 > 0 \land \\
-      \TransitionVar_6 = 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
+    {\begin{array}{lll}
+      \TransitionVar_3 > 0 & \land 
+      \TransitionVar_6 = 0 \\ \land
+      2\TransitionVar_6 + \TransitionVar_4 = 0 & \land
+      1 - \TransitionVar_6 > 0\\  \land 
       2k = 1 + \TransitionVar_3
-    \end{matrix}
+    \end{array}}
   }
   \infer1[\Subsume{}]{
-    \begin{matrix}
+    \begin{aligned}
       \Connected(\AcaOrBc{}, \Filter) \land \\
@@
       2k = 1 + \TransitionVar_3
-    \end{matrix}
+    \end{aligned}
   }
   \infer1[\EquationReasoning{}]{
-    \begin{matrix}
+    \begin{aligned}
       \Connected(\AcaOrBc{}, \Filter) \land \\
@@
       2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{matrix}
+    \end{aligned}
   }
   \infer1[\Propagate{}]{
-    \begin{matrix}
+    \begin{aligned}
       \Connected(\AcaOrBc{}, \Filter) \land \\
@@
       2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{matrix}
+    \end{aligned}
   }
   \infer1[\EquationReasoning{}]{
-    \begin{matrix}
+    \begin{aligned}
       \Connected(\AcaOrBc{}, \Filter) \land \\
@@
       2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{matrix} % BRANCH: > 0
+    \end{aligned} % BRANCH: > 0
   } % SPLIT
@@
         % Other automaton
-        & \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
-        & \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        & \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
+        \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
+        \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
+        \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
       \end{aligned}
@@
   \hypo{
-    \begin{matrix}
-      1 = \TransitionVar_{10} \land 
-      \TransitionVar_3 = \TransitionVar_{11} \land \\
-      1 = \TransitionVar_{12} \land 
-      1 = \TransitionVar_{11} \land 
-      2 > 0
-    \end{matrix}  
+    {\begin{array}{lll}
+      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
+      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
+    \end{array}}  
   }
   \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
-    \begin{matrix}
-      1 = \TransitionVar_{10} \land 
-      \TransitionVar_3 = \TransitionVar_{11} \land  \\
-      1 = \TransitionVar_{12} \land 
-      1 = \TransitionVar_{11} \land \\
-      \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{2, 0, 1}) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land
-        2 > 0
-    \end{matrix}  
+    {\begin{array}{ll}
+      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
+      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
+      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
+    \end{array}}  
   }
   \infer1[\EquationReasoning{}]{
-    \begin{matrix}
-      1 = \TransitionVar_{10} \land
-      \TransitionVar_3 = \TransitionVar_{11} 
-      1 = \TransitionVar_{12} \land
-      1 = \TransitionVar_{11} \\
-      \Image{}_{\Automaton', \Map}(\Filter, 
+  {  \begin{array}{ll}
+      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
+      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
+      \land \Image{}_{\Automaton', \Map}(\Filter, 
       \VectorLiteral{
@@
         0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \\
-        \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-        1 - \TransitionVar_6 > 0 \land
-        \TransitionVar_6 = 0 \land 
-        \TransitionVar_4 = 0 \land
-        \TransitionVar_3 > 0 \land
-        2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
+        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
+        \TransitionVar_6 + \TransitionVar_4 = 0 \\
+        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
+        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
+        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
           \VectorLiteral{
@@
           0,
-          \TransitionVar_3} \land \\
-      \Connected(\SomethingCSomething{}, \Filter) \land
+          \TransitionVar_3} \\ \land 
+      \Connected(\SomethingCSomething{}, \Filter) &
       \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{matrix}  
+    \end{array} } 
   }
   \infer1[\Materialise]{
-    \begin{matrix}
-      \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-      1 - \TransitionVar_6 > 0 \land
-      \TransitionVar_6 = 0 \land 
-      \TransitionVar_4 = 0 \land
-      \TransitionVar_3 > 0 \land
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
+    {\begin{array}{ll}
+      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
+      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
+      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
+      \land 
       \VectorLiteral{
@@
           0,
-          \TransitionVar_3} \land \\
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
+          \TransitionVar_3} & \land
+      \Connected(\SomethingCSomething{}, \Filter) \\
+      \land 
       \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{matrix}
+      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
+    \end{array}}
   }
   \infer1[\Subsume{}]{
-  \begin{matrix}
+  {\begin{array}{ll}
     \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-    1 - \TransitionVar_6 > 0 \land
+    1 - \TransitionVar_6 > 0 &
+    \land
     \TransitionVar_6 = 0 \land 
-    \TransitionVar_4 = 0 \land
-    \TransitionVar_3 > 0 \land
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
+    \TransitionVar_4 = 0 \\
+    \land
+    \TransitionVar_3 > 0 & \land
+    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
+    \land
     \VectorLiteral{
@@
         0,
-        \TransitionVar_3} \land \\
-    \Connected(\AcaOrBc{}, \Filter) \land 
-    \Connected(\SomethingCSomething{}, \Filter) \land \\
+        \TransitionVar_3} & \land 
+    \Connected(\AcaOrBc{}, \Filter) \\
+    \land 
+    \Connected(\SomethingCSomething{}, \Filter) \\
+    \land 
     \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-  \end{matrix}
+    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
+  \end{array}}
   }
   \infer1[\EquationReasoning]{
-    \begin{matrix}
+    \begin{aligned}
       \VectorLiteral{a, b, c} = \VectorLiteral{
@@
       \VectorLiteral{a, b, c}) \land a > b
-    \end{matrix}
+    \end{aligned}
   }

### 2023-05-17 introduction.tex
Slightly refactored version of the paper

--- 2022-07-08 main.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2023}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\usepackage{amsmath,empheq,fancybox}
-\usepackage{paralist}
-\usepackage{url}
-\usepackage{color}
-\usepackage{textcomp,listings}
-\usepackage{array}
-\usepackage{mymacros}
-\usepackage{microtype}
-\usepackage{listings}
-\usepackage{csquotes}
-\usepackage{proof}
-\usepackage[capitalise]{cleveref}
-\usepackage{algorithm2e}      
-\usepackage{multirow}
-\usepackage{mathpartir}
-\usepackage{amsthm}
-\usepackage{numprint}
-\usepackage{ebproof}
-
-\usepackage{mathtools} % Bonus
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-
-\lstset{
-    columns=fullflexible,
-    showspaces=false,
-    showtabs=false,
-    breaklines=true,
-    showstringspaces=false,
-    breakatwhitespace=true,
-    escapeinside={(*@}{@*)},
-    commentstyle=\color{greencomments},
-    keywordstyle=\color{bluekeywords},
-    stringstyle=\color{redstrings},
-    numberstyle=\color{graynumbers},
-    basicstyle=\ttfamily\small,
-    framesep=12pt,
-    xleftmargin=12pt,
-    tabsize=4,
-    captionpos=b
-}
-
-\newif\ifcomments
-\commentstrue
-\newif\ifoutline
-\outlinetrue
-
-\newcommand{\contents}[1]{\ifoutline{\color{blue}
-    \begin{itemize}
-    #1
-    \end{itemize}
-  }\fi}
-
-\allowdisplaybreaks[1]
-
-\begin{document}
-
-%% Title information
-\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
-   counts in strings of a given language. Automata-based string solvers
-   frequently need to compute the Parikh image of large products (intersections)
-   of nondeterministic automata, which in many operations is both prohibitively
-   slow and memory-intensive. We contribute a novel understanding of how Parikh
-   maps can be tackled as a constraint solving problem to solve real-world
-   constraints stemming from functions on regular languages, most notably the
-   length constraint. Furthermore, we show how this formulation can be
-   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
-   prover supporting Presburger logic.  We implement \Calculus{} in a tool
-   called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving Parikh automata
-   intersection problems produced when solving standard string constraint
-   benchmarks involving string length constraints. We show that our solution
-   strictly outperforms the standard approach described
-   in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wide
-   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
-   model checker.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-  <ccs2012>
-  <concept>
-  <concept_id>10003752.10003766.10003773.10003775</concept_id>
-  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003766.10003776</concept_id>
-  <concept_desc>Theory of computation~Regular languages</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10003794</concept_id>
-  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10010124.10010138.10010142</concept_id>
-  <concept_desc>Theory of computation~Program verification</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10011192</concept_id>
-  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  </ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Quantitative automata}
-\ccsdesc[300]{Theory of computation~Regular languages}
-\ccsdesc[500]{Theory of computation~Automated reasoning}
-\ccsdesc[300]{Theory of computation~Program verification}
-\ccsdesc[300]{Theory of computation~Verification by model checking}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}
-
 The Parikh image is a characterisation of formal languages in terms of
@@
 us to \emph{remove} something, thereby potentialy obtaining an easier problem.
-
-\section{A demonstration of our approach}\label{sec:motivation}
-
-To illustrate the key features of \Calculus, we will informally present our
-reasoning. We begin by considering the following set of string and other
-constraints, which we want to solve for all possible values of $\MonoidElement$,
-a value in some commutative monoid and the string $s$, where
-$\varphi(\MonoidElement)$ is some set of constraints on $\MonoidElement$:
-\[
-   s \in \Language(\mathtt{/}\SndRegex{}\mathtt{/})  \land
-   s \in \Language(\mathtt{/}\FstRegex{}\mathtt{/})  \land
-   \MonoidElement = f(s)  \land
-   \varphi(\MonoidElement)
-\]
-
-We will represent the strings using the nondeterministic finite automata
- $\SomethingCSomething{}$ (\cref{fig:something-c-something}) and $\AcaOrBc{}$
- (\cref{fig:aca-or-bc}) respectively.
-
- Throughout these examples we will present our three key ideas:
- \emph{automata-aware splitting}, \emph{lazy enforcement of automata
- connectivity}, and \emph{incremental materialisation of products}.
-
-\subsection{Counting String Lengths}
-
-To keep things simple, in this first example we let $f$ be the function that
-gives the length of a given string; $f\left(\text{curious and curiouser
-still}\right) = 27$, and $f\left(\text{homo}\right) +
-f\left(\text{morphism}\right) = f\left(\text{homomorphism}\right) = 12$, and we
-let $\varphi(\MonoidElement) = \exists k \HoldsThat m = 2k + 1 \land k \geq 0$,
-i.e. require that the length of the string is odd.
-
-\begin{figure}[p]
-  \centering 
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{aca_or_bc}
-    \caption{An automaton recognising the regular expression
-      $\FstRegex{}$.}\label{fig:aca-or-bc}
-      \Description[An NFA recognising $\FstRegex{}$]{An NFA recognising $\FstRegex{}$. It has 4 states.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth} 
-    \centering
-    \includegraphics[scale=\autscale]{something_c_something}
-    \caption{An automaton recognising the regular expression
-    $\SndRegex{}$.}\label{fig:something-c-something}
-    \Description[An NFA recognising $\SndRegex{}$]{An NFA recognising $\FstRegex{}$. It has 2 states.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{aca_or_bc_simplified}
-    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
-    simplification of their equations.}\label{fig:example-simplify-1}
-    \Description[The automaton $\AcaOrBc{}$ labelled with variables counting the number of times each transition is taken, after having simplified the equations.]{The starting state shows an initial flow of 1, with two outgoing transitions to $A$ ($1-l_c'$ on letter a, and $l_b' + l_c'$ on b respectively).
-$A$ has a flow of $l_c$, and its outgoing transition into the final state, $F$ is the same as the incoming one from $S$, the initial state.
-The $B$ transition has an outgoing transition labeled $c/ l_c'$ to the final state $F$, and one backwards transition $b / l_b'$ to the starting state.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{something_c_something_length_simplified}
-    \caption{$\SomethingCSomething{}$ after using linear reasoning over lengths to replace a variable.}\label{fig:example-length-reduced}
-    \Description[The automaton $\SomethingCSomething{}$ after reasoning over lengths.]{The same automaton as before for $\SomethingCSomething{}$, but with the transition from start to final annotated with a constant number 1, and the leftmost self-loop transition replaced with the equation $1 + l_c + 2l_b' -
-    r_\Sigma'$.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-      \includegraphics[scale=\autscale]{aca_or_bc_length_split_1}
-      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ is not
-      used.}\label{fig:aca-or-bc-length-split-1}
-      \Description[The automaton \AcaOrBc{} where we have removed the transition
-      from $S$ to $A$.]{ The state $A$ hangs free without incoming transitions,
-      and only the lower part of the automaton remains.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-      \includegraphics[scale=\autscale]{aca_or_bc_length_split_2}
-      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ \emph{is} used.}\label{fig:aca-or-bc-length-split-2}
-      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ when we keep the transition from $S$ to $A$.
-This removes the transition between $B$ and $F$, but otherwise maintains the automaton.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-      \includegraphics[scale=\autscale]{length_split_product}
-      \caption{The product of $\AcaOrBc$ and $\SomethingCSomething$ after
-      case splitting and removal of dead states.}\label{fig:length-split-product}
-      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ with the transition from $B$ to $F$ removed. There now remains only one path to an accepting state.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-      \includegraphics[scale=\autscale]{aca_or_bc_simplified_3}
-      \caption{$\AcaOrBc{}$ after further constraint propagation.}\label{fig:example-simplify-3}
-      \Description[$\AcaOrBc$ without the lower section.]{$\AcaOrBc$ with only its upper section left and the $a$ transitions now set to static 1.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.6\textwidth}
-    \centering
-        \includegraphics[scale=\autscale]{aca_or_bc_times_something_c_something}
-        \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ under the constraint
-        $\TransitionCount{a} > \TransitionCount{b}$.}\label{fig:final-example} \Description[The
-        final product, still a boring stick automaton with only one loop for an
-        arbitrary number of c's.]{A product of the intermittent automata, which
-        accepts \lstinline{ac.*ca}}
-  \end{subfigure}
-  \caption{The collection of automata we use as running examples in this paper, with annotations illustrating intermittent states of our calculus.}    
-  \end{figure}
-
-The na\"ive approach would be to first compute the product of the automata, then
-use the approach from \cite{generate-parikh-image} to find its image under $f$,
-and then solve the constraints in an automated theorem prover. However, we can
-take advantage of both the constraints and definitions of $f$ to prune the
-automata before computing the product and help reduce the blow-up.
-
-Similar to \cite{generate-parikh-image} we associate each transition
-$\Transition$ with an integer variable $\TransitionVar_\Transition \geq 0$
-representing how many times $\Transition$ is used that we leave implicitly
-existentially quantified. These variables are shown as annotations under the
-transition labels in \cref{fig:aca-or-bc,fig:something-c-something}. We then
-apply graph-like flow reasoning by adding linear equations (e.g. $1 + {l_b}' =
-l_a + l_b$, or $l_a + l_c' = 1$) that equate the incoming transitions into a
-state to the ones going out. For the initial and accepting states we add an
-incoming and outgoing flow of 1, respectively. We will use these linear
-equations to perform reasoning to reduce the size of the product before we
-compute it. Note that we have already performed some linear reasoning to
-eliminate some of the variables in the graphs above, for example to introduce
-the constant number $1$ in \cref{fig:something-c-something}, which comes from $1
-+ r_\Sigma = x + r_\Sigma$ solved for $x$.
-
-We sum the individual transition variables to get the length $\MonoidElement$ we
-are actually interested in. For example the image of $f$ on
-$\SomethingCSomething$ is given by $\exists_{r_\Sigma, r_\Sigma'} \: 2k + 1 =
-r_\Sigma + 1 + r_\Sigma' \land k \geq 0$. For $\SomethingCSomething{}$, this is
-all we need to produce the image since all transitions are always accessible
-from any path from the initial state to the accepting state. The difficult
-connectivity constraint is not necessary to enforce here, since any path through
-$\SomethingCSomething{}$ is connected, and therefore the flow equations
-described above are sufficient to capture its image.
-
-The story gets more complicated with $\AcaOrBc{}$, where we have a choice of two
-paths, up or down, and where that choice affects the possibility of taking
-either of the two loops. In that case, we would have the image $2l_a + l_c + l_c' + l_b + l_b'$.
-If we apply standard linear reasoning on the equations (e.g. $1 + {l_b}' = l_a +
-l_b$, or $l_a + l_c' = 1$), we can reduce the number of bound
-($\exists$-quantified) variables somewhat, obtaining the relations in
-\cref{fig:example-simplify-1}, and the corresponding over-approximation on
-the $f$-image $2 + l_c + 2l_b'$. It is an over-approximation because it only
-considers the number of times each transition is taken and the existence of a
-path from an initial to an accepting state, but omits the connectivity of that
-path in the presence of loops. For example, no valid path would have both $l_c >
-0$ and $1 - l_c' = 0$, but no such restriction exists in the formula above.
-
-Still, this approximate view allows us to perform further reasoning. Since we
-know that the length of a string accepted by the intersection of the two
-automata's languages must be the same, we also know that their images must be
-the same under $f$, e.g. that $r_\Sigma + 1 + r_\Sigma' = 2 + l_c + 2l_b' \iff
-r_\Sigma + r_\Sigma' = 1 + l_c + 2l_b'$. This in turn allows us to replace one
-additional variable, e.g. $r_\Sigma = 1 + l_c + 2l_b' - r_\Sigma'$ (see
-\cref{fig:example-length-reduced}). Note that the two automata are now in
-contact, allowing us to propagate reasoning on one to the other.
-
-Since we cannot take a transition a negative number of times, we have an
-implicit constraint on any transition variable that it must be at least $0$.
-This means that we have $1 + l_c + 2l_b' - r_\Sigma'$, implying an upper bound
-on $r_\Sigma'$: $r_\Sigma' \leq 1 + l_c + 2l_b'$.
-
-This is as far as we can get with this reasoning. We must now choose: either we
-compute (materialise) the product and continue our reasoning, or we perform a
-case split. Let us try a case split in order to put off computing a potentially
-large product.
-
-We apply the principle of splitting and select an early transition in the
-largest automaton with loops, in our case $\FromLabelTo{S}{a}{A}$ of $\AcaOrBc$.
-This creates two cases: $1-l_c' = 0$ and $1-l_c' > 0$. We start with the first
-case, and after propagating the now known values and removing any transition
-that becomes zero, we get the automaton of \cref{fig:aca-or-bc-length-split-1}.
-
-We immediately notice that the self-loop of state $A$ is now disconnected from
-the initial state and must be removed. This condition can be detected
-efficiently using standard forwards/backwards reachability from the initial
-and accepting states respectively. In other words, $2k+1 = 0$ for this case
-and is implied by the splitting constraint. However, no choice of integer $k$
-will satisfy this equation. Therefore, we close this branch and proceed with
-the other one -- where $1-l_c' > 0$. Note that we do so without ever computing
-a potentially expensive product of automata.
-
-Plugging in the assumption $1-l_c' > 0$ (and its consequence that $l_c' = 0$),
-we also get a reduced variant of $\AcaOrBc$, as seen in
-\cref{fig:aca-or-bc-length-split-2}. Note that there is now no way to select any
-transition variable to disconnect a loop from a path between $S$ and $F$,
-and we can therefore stop worrying about enforcing connectivity at all.
-
-Before continuing, we need to tie the inequalities for the transitions of the
-automata of the product to the new transition counting variables we introduced
-in the product automaton to ensure that they are consistent. In our case, we
-obtain the following equations from the principle that a transition variable in
-either automaton must be equal to the sum of variables in the resulting edges of
-the product, which gives the following two interesting inequalities from
-$\AcaOrBc{}$'s transition $\FromLabelTo{A}{c}{A}$: $2k + 1 = r_1 + r_2 + 1$
-
-Note how this conserves the requirement that the number of loop iterations be
-odd even when the same original loop appears three times in the product, twice
-as a self-loop and once as a regular state transition. Using the same principle
-as before for computing the length by adding up the transition counters, we
-obtain the quantified formula $1 + 2k + 1 + 1 + r_b + r_b= 3 + 2k + 2r_b$ for
-the length. Eliminating the final quantifier gives us the image of $f(\AcaOrBc
-\times \SomethingCSomething)$ modulo odd lengths: $3 + 2k$, for some integer
-$k$.
-
-\subsection{Counting Letters}\label{sec:introduction:parikh}
-
-Now assume that rather than the length (e.g. the count of the characters) we
-want to find the whole Parikh image of the intersection of $\AcaOrBc{}$ and
-$\SomethingCSomething{}$. In other words, now $f$ is a mapping to a 3-vector
-counting how often each character occurs in a string. Since extracting values
-from a vector is tiresome, we will use the shorthand notation
-$\TransitionCount{a}$ to refer to the number of letters a that appear. E.g.
-$\TransitionCount{a}(\text{curious and curiouser still}) = 1$. Note that we
-still have the property that $\TransitionCount{o}(\text{homomorphism}) =
-\TransitionCount{o}(\text{homo}) + \TransitionCount{o}(\text{morphism}) = 2 + 1
-= 3$. If you treat the addition as element-wise vector addition, this approach
-works for the vector version as well.
-
-For our constraint, we will use $\TransitionCount{a} > \TransitionCount{b}$,
-that is finding a count of characters in a string accepted by both automata where
-there are more a:s than b:s. Since the underlying calculus is the same, we start immediately from \cref{fig:example-simplify-1}.
-
-  We now need to relate the number of times a transition was taken to the
-  character counts we are really interested in, as opposed to just the length in
-  the previous example. For the $\Sigma$ labels, since we do not know which
-  letter to use, we introduce a sum $r_{a1} + r_{b1} + r_{c1}$ per $\Sigma$
-  transition.
-
-  As in the previous example, we sum the occurrences for each transition to
-  obtain the counts, here with the element-wise vector addition unpacked:
-  \[
-\MonoidElement =   \begin{bmatrix}
-  r_{a1} + r_{a2} \\
-  r_{b1} + r_{b2} \\
-  r_{c1} + r_{c2} + 1 \\
-  \end{bmatrix} =
-  \begin{bmatrix}
-    2 -2l_c' \\
-    2l_b' + l_c' \\
-    l_c + l_c' \\
-    \end{bmatrix} =
-    \begin{bmatrix}
-      \TransitionCount{a} \\
-      \TransitionCount{b} \\
-      \TransitionCount{c}
-      \end{bmatrix}
-  \]
-
-These equalities can be combined with the constraint that $\TransitionCount{a} >
-\TransitionCount{b}$ to obtain $2 - l_c' > l_b'$. Since we have $1-l_c' \geq 0$
-from one of the transition labels, we know $l_c' \geq 0$, and therefore that $1
-> l_b' \implies l_b' = 0$. Plugging in the same inequality again on the same
-principle, we get $2 - 2l_c' > l_c' \iff l_c' = 0$. This gives us the much
-smaller automaton of \cref{fig:example-simplify-3}.
-
-% \begin{figure}[t]
-%   \centering 
-%     \includegraphics[width=0.70\linewidth]{aca_or_bc_simplified_2}
-%     \caption{$\AcaOrBc{}$ after removing the now unused b-transition.}\label{fig:example-simplify-2}
-%     \Description[$\AcaOrBc$ without the returning transition from $B$ to $S$]{$\AcaOrBc$ without the returning transition from $B$ to $S$, which now makes the transition from $S$ to $B$ also have the same associated variable $l_c'$.}
-%   \end{figure}
-
-This will produce a straightforward product since both automata now have the
-same shape, where we essentially only need to expand the self-loops of
-$\SomethingCSomething$ and reduce its range labels to only capture a's, as seen
-in \cref{fig:final-example}. Note that we got this result by only performing
-linear inequality reasoning on the automata of the product versus the number of
-times each transition would be used.
-
-%   \begin{figure}[t]
-% \centering
-%     \includegraphics[width=0.75\linewidth]{original_product}
-%     \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ as it would have
-%     appeared if we had computed it from the initial
-%     automata.}\label{fig:original-product}
-%   \end{figure}
-
-To obtain the final Parikh image, we again need to tie the inequalities for the
-transitions of the automata of the product to the new transition counters we
-introduced in the product automaton. The only interesting transition is the one
-carrying the c, so we will choose that one for illustration. In that case, the
-bridging equation is $r_2 = l_c \land r_2 = r_{\Sigma}$ (and for all others
-$=1$), which gives the Parikh image $\exists_{r_2} \TransitionCount{a} = 1 \land
-\TransitionCount{b} = 0 \land \TransitionCount{c} = r_2 + 1$, and after
-existence-elimination using reasoning on lower bounds (e.g. $r_2 \geq 0$) we
-arrive at $\TransitionCount{c} \geq 1$.
-
-\section{Preliminaries}
-
-\subsection{Monoids}
-
-A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
-consisting of the non-empty carrier set of elements, $X$, an associative binary operation
-$X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,
-c \in X$, $(a \MonoidOp b) \MonoidOp c = a \MonoidOp (b \MonoidOp c)$. Finally,
-$\Monoid$ must have an identity element $0_{\Monoid} \in X$ such that
-$0_{\Monoid} \MonoidOp a = a \MonoidOp 0_{\Monoid} =   a$ for every $a \in X$.
-We sometimes use integer multiplication to represent repeated application of
-$\MonoidOp$, e.g. $3a = a \MonoidOp a \MonoidOp a$, for $a \in X$. $\Monoid$ is
-called \textit{commutative} if $\MonoidOp$ also commutes, that is if $a
-\MonoidOp b = b \MonoidOp a$ for all $a, b \in X$.
-
-Finally, a \textit{homomorphism} is a structure-preserving map between two monoids
-$\Monoid_1, \Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
-\MonoidOp_1 b) = \Map(a) \MonoidOp_2 \Map(b)$, where $S_1, S_2$ are the carrier
-sets of $\Monoid_1$ and $\Monoid_2$ respectively, and $\MonoidOp_1, \MonoidOp_2$
-their binary operations.
-
-\subsection{Languages, Finite-state Automata and their Products}
-
-We define an alphabet as a finite set of symbols $\Alphabet$ with words $\Strings$, and
-the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
-Note that $\Strings = \Tuple{\Alphabet;\Concat{};\epsilon}$, is a
-non-commutative monoid, referred to as the free monoid. The string length
-function, $\Length{s}$ is an example of a homomorphism between $\Strings$
-and~$\mathbb{Z}$.
-
-An automaton~$\Automaton$ with alphabet~$\Alphabet$ is
-$\AutomatonTuple$ where $\Transitions \subseteq \States \times \Alphabet \times
-\States$, $\States$ is its states, $\InitialState$ its
-initial state, and $\AcceptingStates$ is its set of accepting states.  We
-write a transition $\Transition = \Tuple{\State, \Label, \State'} \in
-\Transitions$ as $\Transition = \FromLabelTo{\State}{\Label}{\State'}$.
-Similarly, we use the notation $\FromLabelTo{\State}{}{}$ to refer to the set of
-transitions starting in $\State$, and $\FromLabelTo{}{}{\State}$ to refer to the
-set of transitions coming into $\State$, whenever the automaton is clear from
-the context.
-
-We will let variables $\Transition, \Transition', \Transition_1, \ldots,
-\Transition_n$ etc describe transitions, $\State, \ldots, \State_n$ states, and
-$\Automaton, \ldots, \Automaton_n$ automata, and use subscript indexing
-($\Transitions_\Automaton$) to refer to the transitions, states, etc of a given
-automaton.
-
-By a \emph{product} of two automata $\Automaton_1, \Automaton_2$, written
-$\Automaton_1 \times \Automaton_2$, we mean an automaton constructed to run
-$\Automaton_1$ and $\Automaton_2$ in parallel on an input and only accept the
-input if both automata would do so.
-
-We refer to the resulting product states as tuples, $\Tuple{\State, \State'}$,
-which represent the state of the product automaton where $\Automaton_1$ would be
-in $\State$ and $\Automaton_2$ would be in $\State'$. Note that since we use
-ordered tuples the product is technically (but w.l.o.g) not commutative; the
-left-hand-side must come from the left-hand term. The sole purpose of this
-matching is to allow us to speak with precision about the origins of components
-in a product.
-
-\subsection{The Parikh Map and its Image}
-Formally, the \textit{Parikh map} over an alphabet $\Alphabet=
-\left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:
-$$
-\begin{aligned}
-& \ParikhMap: \MapFromTo{\Strings}{\natural^k} \\
-& \ParikhMap(s) = \VectorLiteral{\#a_1(s), \#a_2(s), \ldots, \#a_k(s)}
-\end{aligned}
-$$
-
-That is, $\ParikhMap(s)$ is a vector of the number of occurrences of each
-character in the language for a given string $s$. For example, for  $\Alphabet =
-\Set{a, b}$, we would have $\ParikhMap(abb) = \VectorLiteral{1, 2}$.
-
-We define the image of this map, the \textit{Parikh image}, of some subset of
-the language $\Language \subseteq \Strings$ as:
-\[
-\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
-\]
-
-Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
-1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
-notation $\#l(w)$ to talk about an individual letter $l$ in a word $w$. For
-example, for the Parikh vector above, we would have $\CountOf{a} = 1$. You have
-already seen this in \cref{sec:introduction:parikh}.
-
-Parikh's theorem states that any context-free language has a letter-equivalent
-regular language (c.f.~\cite{construction} for a construction of such automata
-from context-free grammars and~\cite{bounds} for bounds on its size). However,
-there are languages that are not context-free that also have semilinear images
-under~$\ParikhMap$ (e.g. $\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) =
-\ParikhMap((abc)^*) = \CountOf{a} = \CountOf{b} = \CountOf{c} \land \CountOf{a}
-\geq 0$). This means they can be represented as a quantifier-free Presburger
-formula.
-
-Note that while Parikh's theorem applies to context-free languages, in this
-paper we focus only on regular languages.
-
-\subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
-\label{sec:verma}
-
-Since Parikh images are semilinear, any Parikh image can be written as a set of linear
-equations. The following construction for an NFA $\Automaton = \AutomatonTuple$ is based on the version for context-free grammars presented in \cite{generate-parikh-image}:
-\begin{equation}
-\begin{aligned}
-\ParikhMap(\Automaton) := 
-& \AndComp{\Letter \in \Alphabet}{\LetterVar_{\Letter} = \sum_{\Transition = \FromLabelTo{}{\Letter}{} \in \Transitions} \TransitionVar_{\Transition}
-}\\
-& \AndComp{\State \in \AcceptingStates}{\FinalStateVar_{\State} > 0 \longrightarrow  \StateVar_{\State} > 0}\\
-& \AndComp{\State \in \States}{
-  \left(\FinalStateVar_{\State} \text{ if } \State \in \AcceptingStates \text{ otherwise } 0 \right) +
-  \sum_{\Transition \in \FromLabelTo{\State}{}{}} \Filter(\Transition) - \sum_{\Transition \in \FromLabelTo{}{}{\State}} \Filter(\Transition)
-= \begin{cases}
-    1 \text{  if $\State = \InitialState$} \\
-    0 \text{ otherwise}
-  \end{cases}
-}\\
-& \AndComp{\Transition = \FromLabelTo{\State'}{}{\State} \in \Transitions}{
-  \TransitionVar_{\Transition} > 0 \longrightarrow \StateVar_{\State} > 0
-} \\
-& \AndComp{\State \in F}{
-  \StateVar_{\State} > 0 \longrightarrow \FinalStateVar_{\State} = 1
-} \\
-&\AndComp{\State \in \States}{
-  \StateVar_{\State} > 0 \longrightarrow
-  \OrComp{\Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions}{
-    \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
-    \TransitionVar_{\Transition} \geq 1 \land
-  \StateVar_{\State'} \geq 1
-    }
-}
-\end{aligned}
-\label{eq:generate-parikh}
-\end{equation}
-
-All variables $\TransitionVar_\Transition, \StateVar_\State, \FinalStateVar_\State$ are
-existentially quantified and the free variables $\LetterVar_\Letter$ make up the
-image. $\StateVar_\State$ represents the distance of state $\State$ from
-$\InitialState$ in a spanning tree, $\TransitionVar_\Transition$ how many times
-a transition $\Transition$ is used, and $\FinalStateVar_\State$ whether a given
-accepting state $\State \in \AcceptingStates$ is the actually used final state.
-
-In this paper we refer to this model as the baseline approach, though we also
-apply optimisations as described in \cref{sec:implementing-baseline}. The
-calculus introduced in this paper, by contrast, lazily enforces the
-connectedness constraint of this encoding (the final three clauses) while also
-interleaving the computation of products of automata and propagating information
-between the steps to reduce the amount of work that needs to be done.
-
-\subsection{Theorem Proving}
-
-We assume basic familiarity with first-order logic, Presburger
-arithmetic, and the classical sequent calculus. For reference, see
-e.g. \cite{Fitting96a}.
-
-\section{Projections on Parikh Images}\label{sec:generalised}
-
-It is easy to see that the Parikh map~$\ParikhMap$ represents a
-homomorphism from the (free) non-commutative monoid~$\Strings$ to the
-(free) commutative monoid~$\Naturals^k$. As we are often not
-interested in the full Parikh image, but only in some projections of
-it. We consider \emph{arbitrary} homomorphisms
-$\Map:\: \Sigma^* \to \Monoid$, where
-$\Monoid = \left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative
-monoid. We give several
-examples of such projections on Parikh images later in this section.
-
-Observe that every homomorphism~$\Map:\: \Sigma^* \to \Monoid$ can be
-represented as the composition~$h' \circ \ParikhMap$, for some
-homomorphism~$h' : \Naturals^k \to M$. One of the insights underlying our
-approach is that it is more efficient to directly compute a
-projection on the Parikh image~$\Map(\Language)$, than to first compute the
-standard image~$\ParikhMap(\Language)$ followed by projection to
-some property of interest.
-
-%\subsection{Applications of generalised Parikh images}\label{sec:applications}
-
-\subsection{Example~1: Reasoning about String Length}
-
-One such simplifying homomorphism can express string
-length, the problem that originally motivated our study of the Parikh
-map. This mapping is relevant when solving constraints that combine
-language membership as well as word length, for instance the formula
-\begin{equation}\label{eq:stringLength}
-x \in \Language_1 \wedge y \in \Language_2 \wedge |x| > |y|
-\end{equation}
-mentioned in the introduction. To solve this formula, let
-$\Monoid = \Naturals$, and define the homomorphism~$L$ by $L(a) = 1$
-for all characters~$a \in \Alphabet$. The length of a string
-$s = s_1 \RepeatSum{\Concat} s_n$ is given by
-$L(s) = \sum L(s_i) = 1 \RepeatSum{+} 1 = n$, and to solve
-\eqref{eq:stringLength} we can instead solve the equi-satisfiable
-formula~$\alpha \in L(\Language_1) \wedge \beta \in L(\Language_2)
-\wedge \alpha > \beta$. This paper proposes efficient native
-procedures to reason about membership constraints
-like~$\alpha \in L(\Language_1)$, avoiding the computation of
-the complete image~$L(\Language_1)$.
-
-\subsection{Example~2: Generalised String Constraints with Integer
-  Datatype}\label{sec:parikh-automata}
-
-Parikh images are also applicable for deciding more general classes of
-string constraints~\cite{ostrich-plus}. Consider, for instance, a
-constraint involving the substring operation:
-\begin{equation}
-  \label{eq:substring}
-  x \in \Language_1 \wedge \; 0 \leq n \leq m \leq |x| \wedge x[n:m] \in \Language_2
-\end{equation}
-in which $x$ is a string variable, $n, m$ are integer variables, and
-$x[n:m]$ denotes the substring of $x$ starting at position~$n$ running
-until position~$m$. This constraint belongs to an expressive fragment
-of string logic that cannot be decided by most state-of-the-art string
-solvers. To illustrate the decision procedure proposed in
-\cite{ostrich-plus} for string constraints of this kind, suppose that
-$\Language_2$ is defined by the regular
-expression~$\mathtt{(ab|c)*}$. An automaton recognising this language
-is shown in \cref{fig:parikh-automata} (left).
-
-We can model \eqref{eq:substring} using the notion of a \emph{Parikh
-  automaton}~\cite{parikh-automata,expressiveness}. A Parikh automaton
-is an automaton in which transitions are labelled both with characters
-from the alphabet~$\Alphabet$, and with offset vectors defining the
-increments of a finite number of counters. This means that Parikh
-automata recognise words over an extended
-alphabet~$\Alphabet \times D$, where $D \subseteq \Naturals^d$ is a
-finite set of increment vectors (notation as
-in~\cite{expressiveness}). We use the symbols~$\pi_\Alphabet, \pi_D$
-to denote projections to the first and the second component of a
-composite letter~$(a, \Vector{v})$, respectively, and extend those
-projections to words:
-\begin{equation*}
-  \pi_\Alphabet((a_1, \Vector{v}_1) \RepeatSum{\Concat} (a_k, \Vector{v}_k))
-  ~=~ a_1 \RepeatSum{\Concat} a_k,
-  \qquad
-  \pi_D((a_1, \Vector{v}_1) \RepeatSum{\Concat} (a_k, \Vector{v}_k))
-  ~=~ \Vector{v}_1 \RepeatSum{+} \Vector{v}_k.
-\end{equation*}
-
-\begin{definition}\label{def:parikh-automata} A Parikh automaton of dimension $d
-  \geq 0$ is a pair $\Tuple{\Automaton, C}$, where
-  $C \subseteq \Naturals^d$ is a semi-linear set (or, equivalently, a
-  Presburger formula), and $\Automaton$ is a finite automaton with the
-  alphabet $\Alphabet \times D$, where $D \subseteq \Naturals^d$. We
-  say that $\Tuple{\Automaton, \varphi}$ recognises a
-  word~$w \in \Alphabet^*$ if and only if the automaton has a run
-  accepting an extended word~$w' \in (\Alphabet \times D)^*$ such that
-  $\pi_\Alphabet(w') = w$ and $\pi_D(w') \in C$.
-\end{definition}
-
-\begin{figure}[t]
-  \centering
-  \raisebox{5ex}{
-    \includegraphics[scale=0.9]{counter_automaton_pre}
-    }
-  \hfill
-  \includegraphics[scale=0.9]{counter_automaton}
-  \caption{An automaton recognising the language $\mathtt{(ab|c)*}$
-    (left), and a Parikh automaton describing the pre-image of this
-    language under the substring operation~$\cdot[n:m]$ (right). For
-    sake of presentation, the Parikh automaton contains
-    $\epsilon$-transitions, which could be eliminated in the standard
-    way.}\label{fig:parikh-automata}
-    \Description[An 2-state automaton (left) and a slightly larger automaton with counters (right).]{
-      The rightmost automaton is the starts with an initial state with a self-loop on any character that increments both registers by one, followed by an epsilon transition to a state with a self-loop on c that increments the lower register by one.
-From that state, the automaton has an epsilon-transition into the accepting state, as well as a transition on a into a state with only a transition back again, both incrementing the lower register by one.
-The accepting state has a self-loop on any character that does nothing to the registers.
-      }
-    \end{figure}
-
-    Applied to \eqref{eq:substring}, the decision procedure in
-\cite{ostrich-plus} will construct a pre-image of $\Language_2$ under
-the substring operation~$\cdot[n:m]$, and check whether this pre-image
-is consistent with the constraint~$x \in \Language_1$. Because the
-substring operation depends on the values of the integer
-variables~$n, m$, a Parikh automaton of dimension~$2$ is a suitable
-formalism to describe the pre-image, resulting
-in the automaton in Fig.~\ref{fig:parikh-automata} (right). The first
-component of the increment vectors is used to count the number of
-letters eliminated in the beginning of the string (value~$n$), while
-the second component records the beginning of the eliminated suffix
-(value~$m$).
-
-Denoting the language described by Fig.~\ref{fig:parikh-automata}
-(right) as $\Language_{pre}$, we can then replace \eqref{eq:substring}
-with an equi-satisfiable formula that no longer contains any explicit
-substring operation:
-%
-\begin{equation}
-  \label{eq:parikh-constraint}
-  p \in \Language_{pre} \wedge~
-  \pi_\Alphabet(p) \in \Language_1
-  \wedge~~ \pi_D(p) =
-  \begin{bmatrix}
-    n \\ m
-  \end{bmatrix}~~
-  \wedge 0 \leq n \leq m \leq |p|
-\end{equation}
-%
-To check the satisfiability of \eqref{eq:parikh-constraint}, we need a
-decision procedure that can process intersections of regular languages
-(in this case, of $\Language_{pre}$ and $\Language_1$, synchronising
-on $\Alphabet$), while imposing the side
-condition~$0 \leq n \leq m \leq |p|$ on the increment sum. In
-\cite{ostrich-plus}, this decision procedure turned out to be main
-bottleneck of the string solver, which was one of the motivations to
-develop the lazy algorithm proposed in this paper.
-
-\iffalse
-\begin{figure}[t]
-  \centering
-      \includegraphics[width=0.5\linewidth]{parikh_automaton}
-      \caption{The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
-      $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
-      formula containing the constraints on the final register values cannot be
-      visualised.}\label{fig:parikh-automaton}
-    \end{figure}
-\fi
-
-\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
-
-We start by defining our calculus, \Calculus{}, for one automaton, and only
-extend it to products of automata in \cref{sec:multiple}. Assume an NFA
-$\Automaton = \AutomatonTuple$ with $\NrTransitions$ transitions $\Transitions =
-\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. For convenience, we then
-introduce the following supporting notations:
-
-\begin{definition}
-  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
-  transitions $\Transitions$ represents a path through $\Automaton$ using
-  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
-  \State_{k+1})$ holds), passing zero or more labels $\Label_1, \ldots,
-  \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
-  \InitialState$. The end state, $\State_n$, is not necessarily accepting. If it
-  is, we say the path is also \textit{accepting}.
-  \end{definition}
-
-\begin{definition}
-  Moreover, we talk about the \textit{set of paths} of an automaton,
-  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
-  valid paths through $\Automaton$. Additionally, we use the
-  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
-  state~$\State$.
-\end{definition}
-
-\begin{definition}
-  For a path $\Path = \Tuple{\State_0 \Label_1 \State_1 \RepeatSum{,}
-  \State_n}$, its \textit{word} $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat}
-  \Label_n$ is the word read out on its labels.
-\end{definition}
-
-\begin{definition}
-  The \textit{states} of a path $\Path = \Tuple{\State_0 \Label_1 \State_1
-  \RepeatSum{,} \State_n}$, $\StatesOf(\Path) = \Set{\State_0\RepeatSum{,}
-  \State_n}$ are the states visited along $\Path$. Note that $\InitialState \in
-  \StatesOf(\Path)$ for every path since all paths start in the initial state.
-\end{definition}
-
-\begin{definition}
-  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$ to state
-  $\State$, written $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of
-  transitions such that every accepting path $p$ where $\State \in
-  \StatesOf(\Path)$ contains a transition $\Transition \in C$. A cut does not
-  exist for every state; notably it never exists for $\InitialState$.
-\end{definition}
-
-\begin{definition}
- The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
- number of times a transition $\Transition =
- \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
- $p$.
-\end{definition}
-
-We then introduce the two predicates into our calculus with the following
-definitions:
-
-\begin{definition}\label{def:single-image}
-  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
-  $\Automaton = \AutomatonTuple$, modulo some map $\Map : \Alphabet^* \to \Monoid$ to a
-  commutative monoid $\Monoid$ as described in \cref{sec:generalised}
-  and with a transition selection function
-  $\Filter:\: \Transitions \to \Naturals$ holds when
-  $\MonoidElement \in \Monoid$ is an element of the Parikh image of
-  $\Language(\Automaton)$
-  modulo $\Map$, or more formally when there is an accepting path
-  $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
-  that $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
-  for all $t \in \Transitions$, and $\MonoidElement = \Map(\WordOf(\Path))$.
-\end{definition}
-
-\begin{definition}
-  $\Connected(\Automaton, \Filter)$ for some automaton $\Automaton =
-  \AutomatonTuple$ holds when for every $\Transition =
-  \FromLabelTo{\State}{}{\State'} \in \Transitions$, $\Filter(\Transition) > 0
-  \implies \exists \Path \in \Paths(\Automaton)$ $\Filter(\Transition_i) > 0$
-  for $\Transition_i \in p$, and $\State \in \StatesOf(p)$, or in words that
-  there exists some $\Filter$-selected valid path that reaches $\Transition$'s
-  starting state, $\State$. Intuitively, it represents the condition that
-  $\Automaton$ is connected with respect to the selection function $\Filter$ for
-  every transition. It is redundant to $\Image$ by design.
-\end{definition}
-
-We present the rules of \Calculus{} for one automaton in
-\cref{tbl:rules:single}. The rules operate on sets of formulas, and
-can be interpreted as rules of a one-sided sequent calculus, in
-which all formulas are located in the antecedent~\cite{Fitting96a}.
-Each of the rules should be read and
-applied bottom-up, and relates
-premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
-conclusion~$\SomeClause$. When constructing a proof, we start from some
-root~$\SomeClause$, and then apply proof rules to the goals of the proof
-until a goal can be closed, or no more rule is applicable.
-
-A proof in which all proof goals could be closed shows that the formulas
-in the root~$\SomeClause$ of the proof are inconsistent, and do not
-have any solutions. A proof goal that could not be closed, but to which no
-more rules are applicable, gives rise to  a solution of the formulas in the
-root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic, so that a solution can be computed using
-standard algorithms~\cite{Fitting96a}.
-
-We use the convention of splitting the
-formulas into linear (in)equalities ($\SomeInequalities$)
-and other formulas ($\SomeClause$), and assume that predicates
-$\Image$ and $\Connected$ only occur positively.
- Since
-our rules operate by adding and matching linear (in)equalities in a proof goal,
-we use the shorthand of listing the matched inequalities as antecedents
-e.g., in \Propagate{}).
-The filtering function~$\Filter$ is evaluated symbolically, and can in practice
-be read as a function from transitions to $\Naturals$-valued terms
-(e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{}, described in
-\cref{sec:implementation}, $\Filter$~is a vector of fresh variables with the
-same size as~$\Transitions$.
-
-We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
-transitions of an automaton~$\Automaton$. Additionally, for an
-automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
-like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
-\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
-$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
-instance, the basis for the matched linear inequalities is
-implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
-\left(X;\MonoidOp;0_{\Monoid}\right)$ and the map into it $\Map : \Strings
-\rightarrow X$, we also allow mapping over transitions:
-$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$.
-
-We require that rules can only be applied when they add new formulas
-on every created branch (the notion of \emph{regularity} of a proof is
-required~\cite{Fitting96a}). For example, this means that \Split{} can
-only be applied to proof goals that contain neither
-$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
-never be applied to split on the same term twice on the same branch.
-This suggests a proof strategy where you \Propagate{} when you can, \Split{}
-when you must, and \Subsume{} when neither is possible anymore.
-
-\begin{table}
-\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-  Name & \text{Rule} & Side conditions\\
-  \midrule
-
-  % EXPAND
-  \Expand & 
-    \inferrule
-  {\Connected(\Automaton, \Filter) ,~ \FlowEq(\Automaton, \Filter) ,~ \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
-  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
-  None \\[4ex]
-
-  % SPLIT
-  \Split & 
-  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  if $\Transition \in \Transitions_\Automaton$ \\[4ex]
-
-  % PROPAGATE
-  \Propagate &
-  \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
-  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[4ex]
-
-  % SUBSUME
-  \Subsume &
-  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  \Split{} and \Propagate{} cannot be applied \\
-  \bottomrule
-  \end{tabular}
-  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
-\end{table}
-We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
-of existentially quantified linear inequalities with the following definition,
-where we assign fresh, existentially quantified variables to
-$\FinalStateVar_\State, \Filter(\Transition)$ for every $\State \in
-\AcceptingStates, \Transition \in \Transitions$:
-\[
-\begin{aligned}
-  & \FlowEq(\Automaton, \Filter) = \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
-  \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}\\
-  & \Sink(\State) = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
-  & \In(\State, \Filter) = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
-  & \StartFlow(\State)  = 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
-  & \Out(\State, \Filter) = \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
-\end{aligned}
-\]
-
-In addition to \cref{tbl:rules:single}, we assume the existence of a rule
-\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
-formulae, and for the elements of~$\Monoid$.
-
-The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
-$\Automaton$. It states that we are only allowed to use transitions attached to
-a reachable state, and is necessary to ensure connectedness in the presence of
-cycles in~$\Automaton$.
-
-\textsc{Expand} expands the predicate into its most basic rules; one set of
-linear equations synchronising the transitions mentioned by~$\Filter$ to the
-corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
-the standard Parikh image formulation, as described by~\FlowEq. Since
-$\Connected$ and $\Image$ are partially redundant and the difference is covered
-by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
-this sense, we split the semantics of the $\Image$~predicate into its counting
-aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
-$\Connected$).
-
-Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
-contested transition from a potential solution before concluding that it must be
-included. Intuitively, this is what guarantees our ability to make forward
-progress by eliminating paths through~$\Automaton$.
-
-A decision procedure for our predicate in a tableau-based automated theorem
-prover would start by expanding the predicate using the $\Expand{}$~rule. A
-theorem prover would perform algebraic substitution on the underlying constants
-of~$\Filter$, boiling them down to choices of branches, which depend on one
-single variable, and loop transitions. This logic corresponds to the placement
-of counters for optimally edge-profiling the CFG of a program, making up a
-minimum-spanning tree of the automaton~\cite{path-profiling}.
-
-In order to make the examples below tractable, we will assume the existence of a
-rule \EquationReasoning{} that allows us to perform standard algebraic reasoning
-on linear inequalities. This rule is not necessary for correctness or
-completeness, but shortens the examples considerably.
-
-\subsection{An Example}
-
-Starting with~$\AcaOrBc{}$, where $\Map$ is the length function, in effect
-$\Transition \mapsto 1$ for transitions, and the constraints that the length is
-odd using the same trick as in \cref{sec:motivation}, we have the
-definitions in \cref{fig:example:single:equivalences} (omitting existential
-quantifiers and $x \geq 0$ for every variable to avoid clutter).
-
-\begin{figure}[ht]
-  We define $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$ as in the following two
-  equations, and then apply \EquationReasoning{} (under the implicit assumption
-  that every RHS is $\geq 0$) to obtain the equivalent third definition :
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        & \Filter(\FromLabelTo{S}{a}{A}) & = \TransitionVar_1 \\
-        & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_2 \\
-        & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        & \Filter(\FromLabelTo{B}{b}{S}) & = \TransitionVar_4 \\
-        & \Filter(\FromLabelTo{A}{a}{F}) & = \TransitionVar_5 \\
-        & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \hspace{0.5cm}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        % S
-        &  \TransitionVar_4 = \TransitionVar_5 + \TransitionVar_2 - 1 \\
-        % A
-        & \TransitionVar_1 = \TransitionVar_5 \\
-        % B
-        & \TransitionVar_2 = \TransitionVar_4 + \TransitionVar_6 \\
-        % F
-        & \TransitionVar_5 + \TransitionVar_6 = \FinalStateVar_1 \\
-        & \FinalStateVar_1 = 1 \\
-      \end{aligned}  
-    \end{equation*}    
-  \end{minipage}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \caption{Equivalences defining $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$
-  respectively.}\label{fig:example:single:equivalences}
-  \end{figure}
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo[]{
-    \begin{matrix}
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 = 0 \land \\
-      2 + \TransitionVar_4 > 0 \land \\
-      k = 1 + \TransitionVar_4 
-    \end{matrix}
-  }
-  \infer1[\Subsume{}]{
-    \begin{matrix}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 = 0 \land \\
-      2 + \TransitionVar_4 > 0 \land \\
-      k = 1 + \TransitionVar_4 
-    \end{matrix}
-  }
-  \infer1[\Propagate]{
-    \begin{matrix}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 \geq 0 \land \\
-      2 + \TransitionVar_4 \geq 0 \land \\
-      2k = 2 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{matrix}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{matrix}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{matrix}
-  }
-  % BRANCH:  = 0'
-  \hypo{
-    \begin{matrix}
-      \TransitionVar_3 > 0 \land \\
-      \TransitionVar_6 = 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + \TransitionVar_3
-    \end{matrix}
-  }
-  \infer1[\Subsume{}]{
-    \begin{matrix}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_3 > = 0 \land \\
-      \TransitionVar_6 = 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + \TransitionVar_3
-    \end{matrix}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{matrix}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0\\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{matrix}
-  }
-  \infer1[\Propagate{}]{
-    \begin{matrix}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{matrix}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{matrix}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{matrix} % BRANCH: > 0
-  } % SPLIT
-  \infer2[\Split{} $1 - \TransitionVar_6$]{ \Connected(\AcaOrBc{}, \Filter) \land 2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6 }
-  \infer1[\EquationReasoning{}]{
-    \Connected(\AcaOrBc{}, \Filter) \land
-    2k + 1 =
-    \sum\limits_{\Transition \in \Transitions} (\Transition \mapsto 1)(\Label) \cdot \Filter(\Transition)
-  }
-  \infer1[\Expand{}]{\Image{}_{\AcaOrBc{}, \Transition \mapsto 1}(\Filter, 2k + 1)}
-\end{prooftree}
-\caption{A derivation for \Calculus{} computing odd lengths in $\AcaOrBc{}$.}\label{fig:derivation:single}
-\end{figure}
-
-In \cref{fig:derivation:single}, we see how we start by expanding the predicate
-using the (simplified) flow equations. We can also see how interleaving
-reasoning on the corresponding linear equations helps the deduction. In both
-branches the reasoning is similar: we conclude that when choosing either path at
-the starting state we must avoid the other one, perform propagation based on
-that fact, use some algebraic reasoning to derive the expected forms of the
-upper and lower bounds to propagate the now disconnected transitions, and
-finally subsume, removing the $\Connected{}$ predicate when we are unable to use
-either of the rules, leaving only a set of linear equations. To obtain a solution from the
-proof, one would need to perform standard model generation on the remaining
-constraints to obtain values for the sought variable $\MonoidElement$. The full
-Parikh image is the disjunction of the constraints left at the two
-leaves after quantifier elimination.
-
-Following the values of $\Filter$, we can read out the path corresponding to our solution in the product, and with that a word.
-In our case, for the right branch we have $\Filter(\FromLabelTo{S}{b}{B}) = 0,
-\Filter(\FromLabelTo{A}{c}{A})  > 0,
-\Filter(\FromLabelTo{B}{b}{S}) = 0,
-\Filter(\FromLabelTo{S}{a}{A}) = 1 ,
-\Filter(\FromLabelTo{A}{a}{F}) = 1 ,
-\Filter(\FromLabelTo{B}{c}{F}) = 0$.
-We can read out a word along it: aca, accca, etc, and the corresponding constraint: $\exists k : \TransitionVar_3 = 2k -1 \land \TransitionVar_3 > 0 \land k \geq 0$.
-
-\subsection{Correctness of \Calculus{}}\label{sec:single:correct}
-
-Our correctness proof of \Calculus{} consists of two main parts: we
-first show that the construction of a proof always terminates, and
-then that each of the proof rules in \cref{tbl:rules:single} is
-an equivalence transformation, i.e., does not change the set of
-satisfying assignments of a formula. In combination, those two results
-immediately imply that \Calculus{} gives rise to a decision procedure.
-
-\subsubsection{\Calculus{} terminates}
-\begin{lemma}\label{lma:single-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the predicates
-  $\Image$ and $\Connected$ only occur positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:single}.
-\end{lemma}
-
-\begin{proof}
-  The rule~\Expand{} can only be applied finitely often, since each
-  application removes one $\Image$ predicate, and none of the rules
-  introduce new instances of the predicate. The rule~\Subsume{} can
-  only be applied finitely often, since it strictly decreases the
-  combined number of $\Image$ and $\Connected$ predicates in sets of
-  formulas, and none of the rules increases that number.
-
-  To show termination of \Split{} and \Propagate{}, observe that the
-  $\Filter$ in a predicate~$\Connected(\Automaton, \Filter)$ is never
-  updated on a proof branch, which means that the set of terms
-  $\Filter(t)$ for $t \in \Automaton$ on every branch is finite. Each
-  application of \Split{} and \Propagate{} adds a new
-  formula~$\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$
-  to a proof goal, which can only happen finitely often.
-\end{proof}
-
-\subsubsection{The rules in \cref{tbl:rules:single} are solution-preserving}
-
-\begin{lemma}\label{lma:single-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:single}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
-  conclusion~$\SomeClause$ if and only if it satisfies one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  This property has to be shown by analysing the possible applications
-  of each proof rule.
-
-  \Expand{} unfolds the definition of the $\Image$ predicate. To show
-  that the rule is solution-preserving we prove the equivalence of the
-  upper and lower sets of formulas:
-  \begin{itemize}
-  \item Assume that $\beta$ satisfies the conclusion, which means that
-    there is some accepting path
-    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ with
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$ and $\val_\beta(\MonoidElement) =
-    \Map(\WordOf(\Path))$. Since immediately implies that $\beta$
-    satisfies $\Connected(\Automaton, \Filter)$, since a path is
-    connected, and $\FlowEq(\Automaton, \Filter)$ since an accepting
-    path satisfies the flow equations. The
-    equation~$\MonoidElement = \sum_{\Transition \in
-      \Transitions_\Automaton}\Filter(\Transition) \cdot
-    \Map(\Transition)$ holds because of
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$.
-  \item Assume that $\beta$ satisfies the premise, which implies that
-    $\val_\beta(\Filter)$ describes a consistent, connected flow of
-    the automaton. By the same argument as in
-    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
-    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
-    that each transition~$\Transition$ occurs on $\Path$ exactly
-    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
-  \end{itemize}
-
-  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
-  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
-  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
-  be satisfied, implying the property.
-
-  For \Propagate{}, suppose that
-  $\SeparatingCut(C, \Automaton, \State)$, which means that every
-  accepting path visiting~$q$ contains at least one of the transitions
-  in $C$. For a $\beta$ satisfying $\Connected(\Automaton, \Filter)$
-  and $\Filter(\Transition') = 0$ for $\Transition' \in C$, this means
-  that also $\val_\beta(\Filter(\Transition)) = 0$ has to hold.
-
-  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
-  then a goal must contain~$\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
-  formulas in $\SomeInequalities$ are inconsistent, an application of
-  \Subsume{} is trivially solution-preserving; therefore assume that
-  $\SomeInequalities$ is consistent, which means that it contains
-  exactly one of $\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for each $\Transition$. Since
-  \Propagate{} is not applicable, the transitions~$\Transition$ with
-  $\Filter(\Transition) > 0$ must form a connect sub-graph of the
-  automaton; this means that $\Connected(\Automaton, \Filter)$ is
-  redundant as it is implied by $\SomeInequalities$.
-\end{proof}
-
-\section{Parikh Images from Products of Automata}\label{sec:multiple}
-
-We now generalise our calculus to natively work with intersections of
-regular languages, or equivalently products of automata. For this
-extension, we change the main predicate~$\Image$ to be indexed by a
-vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
-simplicity, we assume that the sets of states of the $k$ automata (and
-therefore also the transition sets) are pairwise disjoint. 
-\begin{definition}\label{def:multiple}
-  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
-  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
-  $\Monoid$,
-  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
-  is a transition selection function, and
-  $\MonoidElement \in \Monoid$. The predicate
-  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
-      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
-  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
-  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
-  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
-  \begin{itemize}
-  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
-    $\Filter$, that is,
-    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
-  \item the automata all accept the same
-    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
-  \item the accepted word is mapped to
-    $\MonoidElement = \Map(\WordOf(\Path_i)) =
-    \Map(\WordOf(\Path_1))$.
-  \end{itemize}
-\end{definition}
-
-  \begin{table}[t]
-    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-      Name & \text{Rule} & Side conditions\\
-      \midrule
-
-      % EXPAND
-      \ExpandM & 
-      \inferrule
-      {
-       {\begin{array}{c}
-           \left\{
-           \FlowEq(\Automaton_i, \Filter),~~
-           \Connected(\Automaton_i, \Filter),~~
-           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
-           \right\}_{i=1}^k,
-           \\[4ex]
-          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
-          \SomeInequalities, \SomeClause
-        \end{array}}
-        }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
-      None \\[5ex]
-      \Materialise &
-      \inferrule
-      {{
-         \begin{array}{c}
-           \FlowEq(\Automaton', \Filter'),~
-           \Connected(\Automaton', \Filter'),~
-           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
-           \\\ImagePredicate{
-                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
-                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
-                     \Automaton'},~ \SomeInequalities, \SomeClause
-           }{\Map}{\Filter'}{\MonoidElement}
-           \end{array}}
-           }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
-      $\begin{array}{@{}l@{}}
-         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
-         1 \leq i < j \leq k, \\
-      \Automaton' = \Automaton_i \times \Automaton_j,\\
-      \end{array}$ \\
-      \bottomrule
-      \end{tabular}
-      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
-    \end{table}
-
-    For the calculus (\cref{tbl:rules:multi}), we first extend
-    $\Expand$ to generate flow equations and instances of $\Connected$
-    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
-    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
-    since it is needed to keep track of the currently considered
-    partial products.
-
-The rule~$\Materialise$ introduces the product of two individual
-automata~$\Automaton_i, \Automaton_j$; this step eliminates
-$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
-instead adds the product~$\Automaton_i \times \Automaton_j$, while
-also introducing the flow equations and the $\Connected$ predicate.
-
-The rule~$\Materialise$ also has to connect the newly introduced
-product~$\Automaton_i \times \Automaton_j$ to the previous automata
-$\Automaton_i, \Automaton_j$. This is done by extending the
-selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
-to fresh variables~$x_{\Transition}$:
-\begin{equation*}
-  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
-  ~~=~~
-  \begin{cases}
-    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
-    \\
-    \Filter(t) & \text{otherwise}
-  \end{cases}
-\end{equation*}
-
-The multiplicity of transitions in the product then has to be related
-to the multiplicities in the individual automata, modelled using the
-$\BindingSum$ predicate. The predicate expresses that the multiplicity
-of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
-$\Automaton_i$ has to coincide with the sum of the multiplicities of
-transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
-and similarly for $\Automaton_j$:
-%
-  $$
-  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
-  \begin{aligned}
-  & \Set{ 
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
-  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
-  & \Set{
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
-  }
-  \end{aligned}
-$$
-
-\iffalse
-Note that this definition implies that $\Filter(\Transition) =
-\Filter(\Transition')$ whenever two transitions $\Transition \in
-\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
-produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
-\times \Automaton_2}$. This corresponds to our intuition that the terms of the
-product must agree on the value they accept. As before, we implicitly map
-$\Filter$ to fresh terms for each transition in the product.
-\fi
-
-For instances of precisely one automaton, neither rule applies and we
-perform the calculus as before.
-
-\subsection{An Example}\label{sec:multiple:example}
-
-We return again to our example in \cref{sec:introduction:parikh}, where we
-compute the whole Parikh image of the product of $\AcaOrBc{}$ and
-$\SomethingCSomething$ under the constraint that there are more instances of
-letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
-element-wise addition, and $\Map$ that maps each transition to the corresponding
-increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.
-
-In the interest of space, we refer back to
-\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
-$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
-substitutions as follows. Note the expansion of the $\Sigma$ labels:
-    \begin{equation*}
-      \begin{aligned}
-        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-        % Other automaton
-        & \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
-        & \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        & \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
-      \end{aligned}
-    \end{equation*}
-
-In the derivation tree of \cref{fig:derivation:multi} we start as before, but
-with the product version of the $\Image$ predicate. The only possible rule here
-is $\ExpandM$, which we use to add the corresponding constraints on each
-automata of the product as we would have had in the single-automaton version. We
-perform algebraic reasoning on those equations, along with the constraint on
-$\MonoidElement$ to determine bounds on transition variables that will enable us
-to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
-finished doing so, we use \Materialise{} to compute the product of the now
-filtered automata. Further algebraic reasoning allows us to remove the final
-instance of $\Connected$. We then perform \Expand{} to get rid of the final
-$\Image$ instance, and are immediately able to $\Subsume$ the resulting
-$\Connected{}$ predicate. This, again, leaves us with a set of linear
-inequalities at the leaf that we can use to obtain a model of our final
-$\MonoidElement$ values (now a vector). This time we can read out aca, or
-$\VectorLiteral{2, 0, 1}$.
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo{
-    \begin{matrix}
-      1 = \TransitionVar_{10} \land 
-      \TransitionVar_3 = \TransitionVar_{11} \land \\
-      1 = \TransitionVar_{12} \land 
-      1 = \TransitionVar_{11} \land 
-      2 > 0
-    \end{matrix}  
-  }
-  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
-    \begin{matrix}
-      1 = \TransitionVar_{10} \land 
-      \TransitionVar_3 = \TransitionVar_{11} \land  \\
-      1 = \TransitionVar_{12} \land 
-      1 = \TransitionVar_{11} \land \\
-      \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{2, 0, 1}) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land
-        2 > 0
-    \end{matrix}  
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{matrix}
-      1 = \TransitionVar_{10} \land
-      \TransitionVar_3 = \TransitionVar_{11} 
-      1 = \TransitionVar_{12} \land
-      1 = \TransitionVar_{11} \\
-      \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \\
-        \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-        1 - \TransitionVar_6 > 0 \land
-        \TransitionVar_6 = 0 \land 
-        \TransitionVar_4 = 0 \land
-        \TransitionVar_3 > 0 \land
-        2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-          \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} \land \\
-      \Connected(\SomethingCSomething{}, \Filter) \land
-      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{matrix}  
-  }
-  \infer1[\Materialise]{
-    \begin{matrix}
-      \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-      1 - \TransitionVar_6 > 0 \land
-      \TransitionVar_6 = 0 \land 
-      \TransitionVar_4 = 0 \land
-      \TransitionVar_3 > 0 \land
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} \land \\
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{matrix}
-  }
-  \infer1[\Subsume{}]{
-  \begin{matrix}
-    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-    1 - \TransitionVar_6 > 0 \land
-    \TransitionVar_6 = 0 \land 
-    \TransitionVar_4 = 0 \land
-    \TransitionVar_3 > 0 \land
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-    \VectorLiteral{
-      \TransitionVar_{7a} + \TransitionVar_{9a},
-      0,
-      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-      = \VectorLiteral{
-        2,
-        0,
-        \TransitionVar_3} \land \\
-    \Connected(\AcaOrBc{}, \Filter) \land 
-    \Connected(\SomethingCSomething{}, \Filter) \land \\
-    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-  \end{matrix}
-  }
-  \infer1[\EquationReasoning]{
-    \begin{matrix}
-      \VectorLiteral{a, b, c} = \VectorLiteral{
-          2 - 2\TransitionVar_6,
-          2\TransitionVar_4 + 3\TransitionVar_6,
-          \TransitionVar_3 + \TransitionVar_6
-        } \land
-        \\
-        \VectorLiteral{a, b, c} = \VectorLiteral{
-            \TransitionVar_{7a} + \TransitionVar_{9a},
-            \TransitionVar_{7b} + \TransitionVar_{9b},
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          } \land \\
-      \Connected(\AcaOrBc{}, \Filter) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{a, b, c}) \land a > b
-    \end{matrix}
-  }
-  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
-\end{prooftree}
-\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
-\end{figure}
-
-\subsection{Correctness of \Calculus{} for Products of Automata}
-
-Since \cref{tbl:rules:multi} only extends the existing rules of
-\cref{tbl:rules:single}, we focus on the differences compared
-to the calculus for a single automaton.
-
-%\subsubsection{\Calculus{} for products of automata terminates}
-\begin{lemma}\label{lma:multi-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the product
-  version of $\Image$ only occurs positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
-\end{lemma}
-
-\begin{proof}
-  The rule \Materialise{} can similarly only be used finitely many times, as
-  each application reduces the number of automata in the product of $\Image$ by
-  one automaton, until only one remains and \cref{lma:single-terminates} for
-  single-automaton instances applies.
-
-  This implies that also the rule~\ExpandM{} can only be applied
-  finitely often since its side condition only allows applying it once
-  per instance of an $\Image$ predicate containing a product, and only finitely
-  many instances of $\Image$ can be introduced on each branch.
-\end{proof}
-
-%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
-
-Since our calculus now includes a rule introducing new variables, the
-\Materialise{} rule, we have to slightly generalise the notion of
-solution-preservation:
-%
-\begin{lemma}\label{lma:multi-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:multi}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
-  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
-  there is an extension~$\beta' of \beta$ satisfying one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  We have to consider the two new rules in \cref{tbl:rules:multi}. The
-  result is immediate for \ExpandM{}, since this rule does not remove the
-  $\Image$ predicate from a proof goal, and the newly introduced formulas
-  are all implied by the $\Image$ predicate.
-
-  For \Materialise{}, observe that the existence of an accepting path
-  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
-  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
-  same word. The path in the product will satisfy the flow equations
-  and connectedness, and it will be related to the individual paths as
-  stipulated by the \BindingSum{} predicate.
-\end{proof}
-
-\section{Extensions}\label{sec:extensions}
-
-This section describes a number of possible extensions to the calculus. Note
-that some are partially or fully implemented in \Catra{} already. In particular,
-we have some level of support for symbolic transitions over Unicode alphabets to
-keep practical automata under a reasonable size, though we do not allow a full Boolean algebra over symbols as described e.g. in~\cite{symbolic-automata}.
-
-\subsection{Backjumping and Learning No-Goods}\label{sec:ext:backjumping}
-
-\Calculus{} can be accelerated for some instances by adding rules for
-backjumping. In particular, central connectedness constraints for an automaton
-can be learnt. In that case, when discovering that a state $\State$ of an
-automaton $\Automaton$ under a certain transition variable $\Filter$ has become
-unreachable, we can learn the clause $\Connected(\Automaton, \Filter) \land
-\sum_{\Transition \in \SeparatingCut(\Automaton, \State)} \Filter(\Transition) =
-0 \implies \sum_{\Transition' = \FromLabelTo{\State}{}{} \in
-\Transitions_{\Automaton}} = 0$. A similar rule can be used when a state becomes
-backwards-unreachable (but with a corresponding backwards cut). At the moment,
-only forward-cut learning is implemented in \Catra, and has provided a slight
-improvement in performance on some instances.
-
-The other source of clauses to learn is the \Materialise{} rule. Whenever an
-attempt to materialise a product of two ($\Filter$-filtered) automata
-$\Automaton_1 \times \Automaton_2$ produces an empty product, we can determine
-the cause of the failure with respect to the automata and their respective
-transition variables, and learn no-good combinations that can never be part of a
-model. In order to be able to do this, we need a few semantic predicates to
-record the state of the calculation, as well as a system of disambiguating
-automata, since it is possible to arrive at the same automaton by multiple
-combinations of decisions and products. In \Catra{}, we use a number of
-additional predicates to record the status of the materialisation of products,
-to separate instances of our main predicate, and to register the mapping between
-automata and their respective transition $\Filter$ terms, described in
-\cref{sec:implementation}. However, at this point only rudimentary no-good
-learning is implemented.
-
-\subsection{Symbolic Automata}\label{sec:ext:symbolic}
-
-Extending \Calculus{} to support fully symbolic automata is possible within the
-framework, depending on your choice of $\Map$. The difficulty consists in
-handling the mapping of the homomorphism $\Map$ over symbolic labels, assuming
-it maps to finitely many monoid elements. This is not always straightforward, as
-seen in the example of \cref{sec:multiple:example}. This complexity is inherent
-in computing the full Parikh image, and stems from the fact that we need to
-differentiate between the possible interpretations of the $\Sigma$ transitions
-without knowing ahead of times which ones will actually be materialied in the
-product. If, on the other hand, if our $\Map$ had been length-counting, which
-does not differentiate between values, it would have required no adaptation at
-all. With a somewhat liberal interpretation of what we are allowed to map to
-(e.g. fresh terms), it is possible to use $\Map$ to represent choice operations
-like the ones in a range label. In \Catra{}, we frontload this problem by
-requiring the user to encode their input as a Parikh automaton. For an example of how automata can be encoded, see \cref{sec:parikh-automata}.
-
-\subsection{Transducers}\label{sec:ext:transducers}
-
-Another interesting application is applying \Calculus{} to automata with
-multiple tracks, e.g. transducers. One application of such a calculus would be
-to represent replace operations and other functions on regular languages, and to
-be able to answer questions like "does this operation change the length of the
-string". The difficulty in implementing it for transducers is first to perform
-the mapping on the labels, which for both the length and Parikh cases is
-straightforward; just do the same thing in two dimensions. The more complicated
-operation is defining the product of transducers. In some cases it would
-probably be desirable to perform synchronisation (e.g. requiring overlapping
-transitions) on only some transitions, for example the first track.
-
-Implementing such a calculus is straightforward in \Calculus, since the
-definition of products was left out of the definition. All that would be needed
-is an appropriate update of the definitions of products. Simiarly, \Catra was
-written with modularity in mind, and it should be straightforward to extend both
-the input grammar and automata implementations to accommodate multi-track
-automata with arbitrary synchronisation.
-
-\subsection{Finding the Presburger representation of a homomorphic image}\label{sec:finding-the-image}
-
-To find the Presburger form of the homomorphic image efficiently, we adapt the
-quantifier elimination approach of~\cite{qe} to our problem domain. The core
-method is the same: we incrementally use \Calculus{} to find models,
-\Generalise{} the models we find into a quantifier-free Presburger formula with
-the $\MonoidElement$ as the only free variable (algorithm~\ref{alg:generalise}), add the
-negated formula as a constraint, and continue enumerating models until none
-remain. The disjunction of the generalised models we enumerated is now our
-image.
-
-\begin{algorithm}
-  \caption{$\Generalise{}(\Automaton, \Filter, a)$ will generalise a final product $\Automaton$ and a model $a$ under the homomorphism $\Map$.}\label{alg:generalise}
-  \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-  \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-  \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-
-  $\Automaton' \gets \Tuple{\Automaton_\States, \Automaton_{\InitialState}, \Automaton_\AcceptingStates, \Set{\Transition \in \Automaton_\Transitions \SuchThat a(\Filter(\Transition)) \neq 0}}$
-
-  \KwRet{\EliminateQuantifiers{$\Map(\ParikhMap(\Automaton'))$}}
-  \end{algorithm}
-
-  \begin{algorithm}
-    \DontPrintSemicolon
-    \caption{$\FindImage{}(\Automaton_1 \times \ldots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
-    \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-    \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-    \SetKwFunction{NewTheoremProver}{newTheoremProver}
-    \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-    \SetKwFunction{FreshVariable}{freshVariable}
-    \SetKwFunction{Assert}{assert}
-    \SetKwFunction{GetModel}{getModel}
-    \SetKwData{ImageVar}{image}
-
-$p \gets \NewTheoremProver{}$\;
-$\Filter(\Transition) := \FreshVariable{p}$ for every $\Transition \in \Transitions_{\Automaton}$\;
-$\MonoidElement \gets$ \FreshVariable{$p$}\;
-\Assert{$p, \exists \MonoidElement, \Filter(\Transition) \text{ for every } \Transition \in \Transitions_\Automaton \HoldsThat \ImagePredicate{\Automaton}{\Map}{\Filter}{\MonoidElement} \land \AndComp{\Transition \in \Transitions}{\Filter(\Transition) \geq 0}$}\;
-$\ImageVar \gets \bot$\;
-\While{$p$ has more models}{
-  $\Tuple{\Automaton, a} \gets \GetModel(p)$\;
-  $G \gets \Generalise{}(\Automaton, \Filter, a)$\;
-  $\ImageVar \gets G \lor \Image$\;
-  \Assert{$p, \lnot G$}\;
-  }
-    \KwRet{\ImageVar}
-    \end{algorithm}
-
-    The \GetModel{} function is nonstandard in that it returns both the model
-    and its associated automaton. Since algorithm~\ref{alg:find-image} is
-    essentially a form of quantifier elimination and therefore an internal affair
-    to the theorem prover, this should be considered fair game.
-
-    A preliminary implementation of this approach is available as part
-    of~\Catra, but has been neither optimised nor fully tested.
-
-\section{Implementation}\label{sec:implementation}
-
-We implement \Calculus{} for Parikh automata as described in
-\cref{sec:parikh-automata}. The artefact submitted along with this paper is a
-program that reads an instance file with one or more products of one or more
-Parikh automaton with transition labels defined as ranges of Unicode characters,
-along with a set of constraints on the final values of their registers expressed
-as Presburger arithmetic in a C-like syntax. We call this program
-\Catra.\footnote{If you really must read it as an acronym, please read it as
-CAtegory Theory on Register Automata, or if you object to the somewhat
-nonstandard use of register automata and category theory, as Check Assignments
-of The Registers Afterwards. Or alternatively if you find it all to be too much
-of a theoretical exercise, as Can Anyone Think of A Real Application.}
-
-\Catra{} is written in Scala, with the calculus described in this paper
-implemented as a theory plug-in for the \Princess{} automated theorem
-prover~\cite{princess}, which also performs the Presburger reasoning. For
-comparison, we also provide an implementation of the baseline method
-from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
-symbolic model checker~\cite{nuxmv} to solve our constraints, and the
-approximation described in~\cite{approximate-parikh} on top of the standard
-baseline back-end. An example of an input file corresponding to our running
-example introduced in \cref{sec:motivation} can be seen in
-\cref{lst:input-example}.
-
-\Catra{} uses symbolic labels for automata. A symbolic label is defined as a
-finite range of Unicode code points. This allows succinct representation of many
-regular expression patterns such as \lstinline{(a-z).*} which would have
-otherwise required $27$~transitions. The transition for the range would be
-written \lstinline{a -> b [97, 122]}.
-
-In satisfaction mode, supported by all backends, \Catra{} tries to satisfy the
-constraints expressed by the input file, reporting \Sat{} with register
-assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
-Additionally, baseline and \Calculus{} also support generation of the Presburger
-formula describing the constraints of the input file. Baseline uses standard
-quantifier elimination, and \Calculus{} uses the method described in
-\cref{sec:finding-the-image}.
-
-Since \Princess{} does not support multiple-arity predicates like the ones we
-use in \Calculus{}, we have implemented variable-length arguments using
-additional helper predicates. These are $\Unused{}(\Automaton)$, which marks an
-automaton as unused in any product, and $\TransitionMask{}(\Automaton,
-\Transition, \Filter(\Transition))$ which associates a transition $\Transition$
-and automaton $\Automaton$ with its corresponding transition variable.
-Additionally, we associate each of our predicates with an instance variable in
-order to differentiate instances of the predicates.
-
-\subsection{Implementing the Baseline}\label{sec:implementing-baseline}
-
-We baseline using the same Presburger solver (\Princess{}), input file parser,
-and automaton implementation as \Catra. We do this in order to better analyse
-the impact of the calculus rules themselves. Using the formula of
-\eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
-successive term and add them to \Princess. We compute the product incrementally
-term by term, checking satisfiability at each step. We use a priority queue to
-select automata for each step, and order it by the number of transitions as a
-heuristic for the size of the automata. We use this heuristic to avoid computing
-large (and therefore slow) products until we have to, banking our hopes on
-computing an empty intermittent product early. The pseudocode for our
-implementation can be seen in \cref{alg:baseline}.
-
-\begin{algorithm}
-  \caption{How we implement the baseline approach}\label{alg:baseline}
-  \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
-  \KwResult{\textsc{Sat} or \textsc{Unsat}}
-  \SetKwFunction{NewTheoremProver}{newTheoremProver}
-  \SetKwFunction{NewPriorityQueue}{newPriorityQueue}
-  \SetKwFunction{Dequeue}{dequeue}
-  \SetKwFunction{Enqueue}{enqueue}
-  \SetKwFunction{Assert}{assert}
-
-  $p \gets \NewTheoremProver{}$
-
-  \Assert{$p$, $\SomeClause$}
-
-  \ForEach{$\Automaton_i$}{
-    \Assert{$p, \ParikhMap(\Automaton_i)$}
-
-    \If{$p$ is \textsc{Unsat}}{break}
-
-  }
-
-  $q \gets \NewPriorityQueue{}$
-
-  \While{$p$ not \textsc{Unsat} and $|q| > 1$}{
-    $\Automaton, \Automaton' \gets \Dequeue{q}$
-
-    \Assert{$p, \ParikhMap(\Automaton \times \Automaton')$}
-
-    \Enqueue{$q, \Automaton \times \Automaton'$}
-  }
-
-  \KwRet{$p$'s SAT status}
-
-  \end{algorithm}
-
-As an optimisation, our automata (including intermittent products) are created
-forward- and backward- reachable-minimal. Any automaton we produce only contains
-states that are both reachable from the initial state and has a path to an
-accepting state. We never perform any other minimisation on the automata for
-either backend. More complex minimisation was left out since performing
-minimisation on automata with counters is non-trivial, and minimising symbolic
-automata risks exponential blowup \cite{minimising-symbolic}.
-
-\subsection{Heuristics and search strategies}
-
-There are a number of choices left unspecified in \Calculus{} as described in
-\cref{sec:calculus,sec:multiple}. For example, the order of materialisation of
-intermediate products and the order of splitting. In this section we describe
-additional implementation details and techniques used to enhance \Catra.
-
-\subsubsection{Splitting, Materialisation, and Propagation}
-
-We order our rule applications so that we first propagate connectedness if
-possible, then perform materialisation if tractable as defined below, then
-finally resort to splitting if we must.
-
-In addition to applying \Split{} as described in \cref{tbl:rules:single} to
-randomly selected transitions, we prefer splitting to sever a strongly connected
-from the initial state. We randomly select an automaton where we can compute a
-cut between an SCC and the initial state, that is where the SCC does not contain
-the initial state and where the sum of the transition variables of the
-transitions in the cut is not known to be positive. If there are multiple such
-strongly connected components we choose one randomly. We then proceed to split
-on the sum of the transition variables of the cut as if it were a regular
-transition, e.g. its sum being zero or nonzero. In this way we drive \Calculus{}
-towards applying \Propagate{}.
-
-The implementation of the connectedness constraint is opportunistic and straightforward.
-We compute a set of dead states by performing forward and backwards reachability computations on an automaton, where we disregard any transition whose associated variable is known to be zero.
-After that we add clauses ensuring any transition variable associated with a transition starting in a dead state is zero.
-
-Product materialisation is the final piece of the puzzle. In the current
-implementation we put off computing intermediate products until at least all but
-two transition variables of one of the automata is known to be either present or
-absent. The number was chosen experimentally, and we observe a consistent trend
-towards lower numbers being better. The other automaton for the product is
-selected randomly.
-
-\subsubsection{Clause Learning}
-
-\Catra{} enables clause learning by default when using our backend, as it has
-been experimentally shown to increase the performance in aggregate (though not
-strictly). We do not currently implement all the proposed features of
-\cref{sec:ext:backjumping}, but we do implement forward-reachability cut
-learning. No sophisticated clause learning for products has been implemented.
-
-\subsubsection{Random Restarts}
-
-Finally, we perform restarts scaled by the Luby series~\cite{luby}. Experimental
-results have shown this to have a large improvement in performance, which is
-unsurprising given how many random choices we make during solving and how
-tail-heavy our problem is.
-
-\section{Evaluation}\label{sec:experiments}
-
-We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
-automata intersection problems generated by the \OstrichPlus{} string constraint
-solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
-initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
-five seconds by baseline.
-We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial.
-For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s.
-The benchmarks are run on commit~\texttt{\commit}.
-
-The benchmarks are executed in parallel using GNU Parallel~\cite{parallel},
-since they are mostly single-threaded and initial results showed negligible
-interference on performance, on \BenchmarkRig{}. We compiled the code using
-Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a
-maximum heap of~\MaxHeapSize{}. We used \Nuxmv{} version~\NuxmvVersion{} invoked
-as a subprocess for each instance. Instances were executed in batches of
-\BatchSize{}, each given a fresh JVM. We believe this represents a realistic use
-case where \Calculus{} is used to support e.g a string solver. Experiments were
-executed in random order for all backends. Each instance got a time budget
-of~\RuntimeTimeout.
-
-All runtimes are measured in wall-clock time as observed by the JVM when
-executing the instance, and exclude time spent parsing (usually far below
-\numprint{0.1}s).
-
-\subsection{Execution Time and Ability to Solve Instances}\label{sec:runtime}
-
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
-the respective back-ends could solve and with which status. A full summary of
-their outcomes is also available in \cref{tab:solve-status}. Note that we lack
-ground truth for all instances. We see that \Calculus{} generally outperforms
-\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
-instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
-kind of instance, but most of all on satisfiable instances, and \Calculus{}
-outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
-\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
-caused by a non-deterministically triggered bug in the clause learning of
-\Calculus{}, but have not had time to further investigate it. We have however
-validated all \numprint{3908} satisfying assignments from \Calculus{} with
-\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
-our implementation.
-
-Baseline performs worse on satisfiable instances because it executes a heuristic
-meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances. The heuristic is enabled since it is unrealistic to assume that we
-know in advance whether an instance is satisfiable or not, but we do know that
-unsatisfiable instances are more common.
-
-\begin{table}
-  \centering
-  \input{graphs/solved_pivot_table.tex}
-  \caption{The result of running the respective back-ends with a timeout of
-  \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
-  from the table. }\label{tab:solve-status}
-\end{table}
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
-    \caption{The division of statuses per backend.}
-    \label{fig:solve-division}
-    \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
-Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
-    \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
-    \label{fig:runtime-boxplot}
-    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline.
-The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy.
-On the other hand, lazy has  a lot of outliers.}
-  \end{subfigure}
-  \vfill
-  \begin{subfigure}[b]{0.75\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
-    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
-    \label{fig:cactus}
-    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
-Lazy solves much more instances than nuxmv.}
-  \end{subfigure}
-\end{figure}
-
-\subsubsection{Scalability}\label{sec:scaling}
-
-A cactus plot showing the number of instances solved within a given timeout for
-each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
-\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
-long-term scaling properties.
-
-\iffalse
-\subsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
-
-For baseline and \Calculus{}, \Catra{} offers the ability to find the equivalent Presburger formula representing a given instance.
-For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} we use the specially tailored approach described in \cref{sec:finding-the-image}.
-For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in \cref{sec:scaling,sec:runtime}.
-
-To make sure baseline puts up as much competition as possible, we disable
-checking intermittent satisfiability and configure \Catra{} to run in the
-maximally eager mode where the product is first computed before any
-satisifiability check is performed. We run the experiments with a timeout
-of~\ImageTimeout{}. The results of the experiment is summarised in
-\cref{fig:cactus:image} and \cref{tab:image-results}. \Fudge{We see here that
-something happens}.
-
-\begin{figure}[ht]
-  \caption{The number of instances \Catra{} was able to find the Presburger form of the image for within a given number of seconds per backend.}
-  \label{fig:cactus:image}
-\end{figure}
-\fi
-
-\subsection{Threats to Validity}
-
-The most obvious threats to validity would be poor benchmarking or poor
-implementation, e.g. if the method described in \cref{sec:calculus,sec:multiple}
-deviate from what is actually benchmarked in section \cref{sec:experiments}, or
-if the methods used for benchmarking would be unsound.  To increase the
-probability that our results are representative both in the sense of
-representing  on expected inputs and in the statistical sense for a given run of
-\Catra{} despite the use of randomness in our implementation we execute many
-experiments. Moreover, to address the issue of correctness we have validated all
-reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
-is indeed sound.
-
-Another threat to validity would be if the competition (baseline and \Nuxmv)
-would be disadvantaged in our comparison. For \Nuxmv{} we use the default
-configuration which we believe should be performant (or it should not be the
-default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
-easy for artefact reviewers. For baseline, our best argument is that the
-solution we use is close to the one found in the \Ostrich{} string solver and
-that it therefore should be realistic.
-
-The greatest threat to validity is our choice of implementation platform and
-automata library. There are some signs that product computation is inefficient,
-notably the good performance of low-threshold automata materialisation that
-prioritises computing smaller products. This means that \Calculus{} makes less
-heavy use of product computation than baseline does, instead relying more on
-\Princess{}'. This situation would unfairly advantage \Calculus{} if our
-automata library was the bottleneck due to our automata library being
-unnaturally poor. We believe this is unlikely since similar performance issues
-have been reported for string solvers. Additionally, profiling suggests that
-both \Princess{}-based backends spend most of their time in \Princess{},
-suggesting that the automaton implementation is not the bottleneck.
-
-\section{Conclusion}
-
-In this paper we have introduced a calculus to compute commutating operations on
-intersections of regular languages that we call \Calculus{}. We have evaluated
-it on \NrBenchmarks{} Parikh automata intersection problems generated by the
-\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
-suite \cite{pyex} using our Parikh automata solver \Catra{}.
-
-Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
-usage and solve-time compared to the baseline approach laid out in
-\cite{generate-parikh-image} when implemented on the same underlying automated
-theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
-\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
-and generally outperforming it for timeouts under 30 seconds with its advantage
-increasing drastically for even shorter timeouts. 30 seconds would generally be
-considered a long timeout for our intended use as supporting infrastructure to a
-string constraint solver.
-
-Future investigations involve two tracks. The first one is integration into
-existing string solvers (wich \Ostrich{} being a particularly promising
-candidate due to its shared use of \Princess{}), and further adaptation to that
-use case. Closer inspection of the instances where we currently time out should
-be useful to further improve our heuristics.
-
-The second track for future improvements is the extension into other problem
-domains, including other logics, model checking problems, as well as to more
-powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata, due to our use of a full
-automated theorem prover which allows adding arbitrary constraints in addition
-to the expected Presburger formulae.
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-%% Bibliography
-\bibliography{bibliography}
-
-%% Appendix
-\appendix
-\section{Appendix}
-\subsection{Per-instance comparison between backends}
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter.pdf}
-    \caption{Per-instance runtime between baseline and \Calculus{}}
-    \label{fig:duels:baseline}
-    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and baseline]{The plot shows most dots at the bottom of the plot, with some plots collected at the right-hand side.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter.pdf}
-    \caption{Per-instance runtime between baseline and \Calculus{}}
-    \label{fig:duels:nuxmv}
-    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and \Nuxmv{}]{The plot shows flames of dots between the bottom and top, with a line of dots near the right-hand side corresponding to timeouts.}
-  \end{subfigure}
-\end{figure}
-\subsection{Example instance for \Catra{}}
-\begin{lstlisting}[caption={An example input file for \Catra{} for the problem introduced in \cref{sec:motivation}, illustrating every major syntax element.
-From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]
-  // So far we only support integer counters. Note that we have individual
-  // counters for each automaton to avoid surprises for product construction.
-  counter int l_a, l_b, l_c, r_c;
-
-  synchronised {
-  automaton aca_or_bc {
-    init S;
-
-    // We use ASCII values here, this is for lowercase a
-    S -> A [97] { l_a += 1 };
-    S -> B [98] { l_b += 1 };
-
-    B -> S  [98] { l_b += 1 };
-    B -> final [99] { l_c += 1 };
-
-    A -> A [99] { l_c += 1 };
-    A -> final [97] { l_a += 1 };
-
-    accepting final;
-  };
-
-  automaton something_c_something {
-      init S;
-
-      // Special short-hand value for the whole Unicode alphabet.
-      S -> S [any] ;
-      S -> F [99] { r_c += 1 };
-
-      F -> F [any];
-
-      accepting F;
-  };
-  };
-
-  // Constrain the number of a:s to be larger than the number of c:s. Since the
-  // automata are synchronised on every transition and counters are guaranteed
-  // to be consistent these constraints are sufficient without also constraining
-  // r_c.
-  constraint l_a > l_c;
-\end{lstlisting}
-
-% Text of appendix \ldots
-
-\end{document}

### 2023-05-17 main.tex
Slightly refactored version of the paper

@@
 
-\section{Introduction}
-
-The Parikh image is a characterisation of formal languages in terms of
-their character counts. Given a language over an
-alphabet~$\{a_1, \ldots, a_k\}$, the Parikh image is a set of
-$k$-dimensional vectors that contains some
-vector~$\VectorLiteral{m_1, m_2, \ldots, m_k}$ if and only if the
-formal language contains a word in which each $a_i$ occurs $m_i$
-times. It is a classical result that the Parikh image of every
-context-free language (and, thus, also of every regular language) is a
-semilinear set~\cite{parikh-theorem}, i.e., Presburger-definable.
-
-Parikh images play a central role in many automata-based algorithms,
-for instance and notably in today's string solvers, which often have
-to process constraints that combine regular language membership with
-word length. To decide whether a simple formula like
-$x \in \Language_1 \wedge y \in \Language_2 \wedge |x| > |y|$,
-with string variables~$x, y$ and regular
-languages~$\Language_1, \Language_2$, is satisfiable, it is
-necessary to reason about the sets of word lengths induced by
-$\Language_1, \Language_2$, which are special cases of the Parikh
-image.  This required combined reasoning about strings and string
-length has long been identified as a major bottleneck in string
-solvers
-\cite{DBLP:conf/cav/AbdullaACHRRS15,length-aware-solver,approximate-parikh,DBLP:journals/corr/BerzishZG17}.
-Other string solvers make use of Parikh automata~\cite{parikh-automata}, and thus
-Parikh images in the general case, to handle operations including
-\verb!str.substr!  and \verb!str.at!, which comes at an even higher
-price in terms of computational complexity~\cite{ostrich-plus}.
-
-\iffalse
-It appears naturally as part of \Fudge{many operations} in model
-checking and solving string constraints in automata-based solvers such
-as \Ostrich{} \cite{ostrich}, notably in representing constraints on
-string lengths. The enhanced \OstrichPlus{} solver~\cite{ostrich-plus}
-makes even more extensive use of Parikh images.
-\fi
-
-It is possible to compute an existential Presburger formula describing
-the Parikh image of any context-free language in linear
-time~\cite{generate-parikh-image}; for the special case of regular
-languages, this result was also stated in
-\cite{muscholl-linear}. While theoretically elegant, this construction
-has several disadvantages, often making it unpractical for integration
-into algorithms. Firstly, the constructed Presburger formula contains
-a linear number of existential quantifiers in the size of the
-considered grammar, as well as complex Boolean structure, which is
-needed to express the connectedness of sets of productions considered
-in the construction. Eliminating those quantifiers to obtain a
-quantifier-free representation of the Parikh image has exponential
-complexity~\cite{DBLP:conf/issac/Weispfenning97}, and is often impossible in reasonable
-time. Just solving the Parikh image membership problem is NP-complete,
-as it corresponds to computing a satisfying assignment of the
-existential Presburger formula, and taxing for solvers as
-well~\cite{ostrich-plus}.
-
-\iffalse
-Later improvements have produced a construction taking at
-most linear time to produce~\cite{muscholl-linear}. However, the
-resulting existentially quantified clauses are costly to eliminate as
-the number of variables increases, in practice making many real-world
-problems intractable.
-\fi
-
-Secondly, in applications involving regular languages, it is typically
-necessary to consider the Parikh image not only of a single automaton,
-but of the intersection of multiple automata. This problem arises in
-string solvers in particular, as conjunctions of string constraints
-lead to the computation of length images of products (intersections)
-of regular languages represented as finite automata. Applying the
-approach in~\cite{generate-parikh-image} would in this case require
-the eager computation of the product before its length image, and
-result in an existential Presburger formula of exponential size (in
-the number of automata). In several instances we have observed while
-solving real-world string constraints, the computation of the product
-of automata exhausts the memory of any machine available due to the
-exponential blow-up in size of the product, quickly becoming
-intractable as the number of automata in the product increases. The
-current best published mitigation for this problem is an
-over-approximation that works by approximating the Parikh image of a
-product of automata to be the conjunction of the image of the
-individual automata of the product \cite{approximate-parikh}. This
-approach only works for unsatisfiable instances, and comes with a
-harsh penalty for satisfiable instances.
-
-Addressing these concerns, we have developed a calculus for Parikh images of
-products of regular languages that we call \Calculus{}. It allows us to
-interleave the computation of arbitrarily deep products of automata with the
-product's Parikh image, and is generalised to an arbitrary homomorphism over
-automata labels, including string lengths. This enables us to let both
-calculations inform each other, eliminating unnecessary work, and pruning the
-size of the partial products considered in the computation for a smaller memory
-footprint. Moreover, the method can be used iteratively to tackle smaller chunks
-of the product incrementally, thereby decreasing the memory footprint.
-
-The key ideas of \Calculus{} is a combination of problem-aware case splitting,
-lazy enforcement of automata connectivity, and lazy computation of products.
-
-We implement \Calculus{} as a theory for the \Princess{} automated theorem
-prover in the tool \Catra. \Catra{} also supports the approximate method of
-\cite{approximate-parikh}, its fall-back variant adapted
-from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
-checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
-back-ends on \NrBenchmarks{} Parikh automata intersection problems generated by
-\OstrichPlus{} when solving the PyEx string constraint benchmark suite involving
-string length constraints \cite{pyex}.
-
-In summary, we contribute:
-\begin{itemize}
-\item The \Calculus{} calculus to efficiently compute (a homomorphism on) the
-Parikh image of products of Parikh automata.
-\item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future study.
-\item The \Catra{} tool for solving such instances, containing an implementation of \Calculus{}, the over-approximation described in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model checker~\cite{nuxmv}.
-\item Suggestions for how to efficiently implement \Calculus{} in a modern automated theorem prover, including strategies for case splitting, clause learning, and constraint propagation for connectedness.
-\end{itemize}
-
-\subsection{Related Work}
-
-The problem of computing constraints on Parikh images over products of regular
-languages under a given commutative homomorphism amounts to solving products of
-Parikh automata. Parikh automata are regular automata extended with integer
-counters with given increments and decrements for each transition, where we
-allow checking a set of linear constraints on the final values of the counters
-(but not their intermittent values) \cite{parikh-automata}. Parikh automata
-without constraints on the final values on their registers are also sometimes
-called cost-enriched automata, weighted automata or counter automata, depending
-on exact definitions and side-constraints. The decision problem tackled in this
-paper, determining the emptiness of an intersection of Parikh automata, was
-recently shown to be PSPACE-complete~\cite{graph-queries}.
-
-Parikh image computations as well as Parikh automata feature extensively in
-string solvers, including as mentioned above \Ostrich{} and \OstrichPlus{}
-\cite{ostrich,ostrich-plus}, but also forms the basis of Trau~\cite{trau-pldi},
-and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images frequently appear when
-introducing cardinality constraints like length or string indexing. The
-state-of-the-art approach to handling Parikh image computation is to
-over-approximate the Parikh image of a product of $k$~automata
-$\ParikhMap(\Language(\Automaton_1) \cap \ldots \cap \Language(\Automaton_k))$ with the
-conjunction of the automata's parikh maps, $\bigwedge_{i=1}^{k}
-\ParikhMap(\Automaton_i)$. This approach works only for unsatisfiable instances,
-and will require falling back to computing the product of the automata before
-using the standard approach for finding its image originally presented
-in~\cite{generate-parikh-image}.
-
-Outside of string solvers, Parikh automata have been proposed as the basis of
-queries~\cite{graph-queries}, and for solving cardinalities in model checking
-problems involving epistemic logic~\cite{epistemic-logic}.
-
-Many other generalisations of the Parikh image than the projections we use here have been
-studied. Prominent examples include generalising the Parikh map to segments of a
-fixed length \cite{KARHUMAKI1980155} and the more general Parikh matrix, which
-gives more information about a word than the standard Parikh image
-\cite{parikh-matrix}. Another notable generalisation is the p-vector, introduced
-in~\cite{infinite-words}, which denotes the position of each letter in the word
-rather than the number of their occurrences and allows for generalisations into
-infinite alphabets. All of these in some sense extend the Parikh map. By
-contrast, the main utility of the formulation introduced here is that it allows
-us to \emph{remove} something, thereby potentialy obtaining an easier problem.
+\section{Introduction}\label{sec:introduction}
+\input{introduction.tex}
 

### 2023-05-25 introduction.tex
Add new examples

@@
 
-\iffalse
-It appears naturally as part of \Fudge{many operations} in model
-checking and solving string constraints in automata-based solvers such
-as \Ostrich{} \cite{ostrich}, notably in representing constraints on
-string lengths. The enhanced \OstrichPlus{} solver~\cite{ostrich-plus}
-makes even more extensive use of Parikh images.
-\fi
-
 It is possible to compute an existential Presburger formula describing
@@
 well~\cite{ostrich-plus}.
-
-\iffalse
-Later improvements have produced a construction taking at
-most linear time to produce~\cite{muscholl-linear}. However, the
-resulting existentially quantified clauses are costly to eliminate as
-the number of variables increases, in practice making many real-world
-problems intractable.
-\fi
 
@@
 us to \emph{remove} something, thereby potentialy obtaining an easier problem.
+
+\subsection{An intuition for our method}
+
+\begin{example}\label{ex:string-constraints}
+    Consider the following set of string constraints:
+\begin{constraints}
+    \item\label{const:x-in-c-dd} $x \in \Regex{c\RegexOr{}dd}$
+    \item\label{const:y-in-b} $y \in \Language(B)$, the language accepted by automaton~$B$ of \cref{fig:aut_b}.
+    \item\label{const:x-substring} $x = \Substring(y, l, r)$
+    \item\label{const:more-a-than-b} $\CountOf{\Char{a}}(y) > \CountOf{\Char{b}}(y)$, that is $y$ contains more letters \Char{a} than letters \Char{b}.
+\end{constraints}
+\end{example}
+
+\cref{const:x-in-c-dd,const:y-in-b,const:x-substring} can be translated to the product of Parikh automata $B \times A'$, where $A'$ (\cref{fig:aut_a}) is the Parikh automata obtained by applying the encoding of $\Substring$ described in \cite{ostrich-plus}.
+Here follows an intuition for our method on $B \times A'$ under the global constraint \cref{const:more-a-than-b}.
+Note that the cost registers of \cref{fig:aut_b} are simply counting letters.
+
+Each transition is associated with a fresh variable detailing how many times it is used
+
+Counters are the sum of their increments times number of times the transition was used
+
+TODO: rewrite the input and log it through CATRA!
+
+\begin{figure}[h]
+    \centering 
+\begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \includegraphics[scale=\autscale]{b}
+    \caption{The automaton $B$, where the Parikh registers count the number of occurrences of each letter.}\label{fig:aut_b}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \includegraphics[scale=\autscale]{a}
+    \caption{The automaton $A'$, whose Parikh registers count both the number of letters and the start ($l$) and length ($r$) of the substring.}\label{fig:aut_a}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}
+\end{figure}s

### 2023-05-25 introduction.tex
More work on introduction

@@
 
+In this section we will introduce an intuition for how a string constraint problem in an automata-based solver like \OstrichPlus{} is translated to a Parikh automata intersection problem and solved efficiently using \Calculus{}.
+First, by a Parikh automaton we mean a standard Non-deterministic finite automaton (NFA) with a set of Parikh registers that are incremented (or decremented) at each transition.
+In the automata of \cref{fig:examples}, we will use the notation $a / \begin{bmatrix} x_+ \end{bmatrix}$ to mean that a transition reads an input character $a$ and increments register $x$ by one.
+We will omit any zero-valued increments and assume that all registers are scoped to their automaton.
+
+Furthermore, we use $\CountOf{c}(A)$ to refer to the number of letters $c$ in $A$, and allow this expression to be used both in the input language of \cref{ex:string-constraints} and in the langauge of constraints that we translate to.
+Where $A$ is apparent, we will omit it.
+
 \begin{example}\label{ex:string-constraints}
@@
   \end{subfigure}
-\end{figure}s
+  \caption{The collection of automata we use as running examples}\label{fig:examples}
+\end{figure}

### 2023-05-29 introduction.tex
Introduction

@@
 
-\subsection{An intuition for our method}
+\subsection{An Intuition for Our Approach}\label{sec:intuition}
+
+\begin{figure}[h]
+    \centering 
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+    \includegraphics[scale=\autscale]{a}
+    \caption{The automaton $A'$, whose Parikh registers count both the number of letters and the start ($l$) and length ($r$) of the substring matching $\Regex{c\RegexOr{}dd}$.}\label{fig:aut_a}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+    \includegraphics[scale=\autscale]{b}
+    \caption{The automaton $B$, where the Parikh registers count the number of occurrences of each letter.}\label{fig:aut_b}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}
+  \caption{The collection of automata we use as running examples}\label{fig:examples}
+\end{figure}
 
@@
 We will omit any zero-valued increments and assume that all registers are scoped to their automaton.
+The increments are in reality a vector (hence the notation), but as it is mostly sparse here, we instead use the symbolic notation $\begin{bmatrix} x_{1+} \end{bmatrix}$ rather than the more cumbersome $\begin{bmatrix} 0 \\ 1 \\ 0 \\  \vdots \\ 0 \end{bmatrix}$.
 
-Furthermore, we use $\CountOf{c}(A)$ to refer to the number of letters $c$ in $A$, and allow this expression to be used both in the input language of \cref{ex:string-constraints} and in the langauge of constraints that we translate to.
-Where $A$ is apparent, we will omit it.
+Furthermore, we use $\CountOf{c}(A)$ to refer to the number of letters $c$ in $\Language(A)$, and allow this expression to be used both in linear constraints of the input language of \cref{ex:string-constraints} and in the langauge of constraints that we translate to.
+Where the automaton $A$ is apparent, we will omit it.
 
@@
 
-Each transition is associated with a fresh variable detailing how many times it is used
+Our approach to the computation of the Parikh image (and therefore the solution to the intersection problem posed by \cref{ex:string-constraints}) consists of two applications of laziness; late computation of products of automata, and lazy enforcement of connectivity constraints.
+Both are in contrast to the eager approach to finding the Parikh image described in~\cite{generate-parikh-image}.
 
-Counters are the sum of their increments times number of times the transition was used
+Starting with an interactive theorem prover and the goal of solving the constraint $\CountOf{\Char{a}}(B \times A') > \CountOf{\Char{b}}(B \times A')$ we associate each transition of $A'$ and $B$ with a fresh integer variable, $\TransitionVar_\Transition$, identically to \cite{generate-parikh-image}.
+These variables represent how many times each transition is taken.
+Hence, the final counter value for each automaton is the element-wise sum:
+\begin{equation}
+\begin{bmatrix} \CountOf{\Char{a}} \\ \CountOf{\Char{b}} \\ \CountOf{\Char{c}} \\ \CountOf{\Char{d}} \end{bmatrix} = \sum_\Transition \TransitionVar_\Transition \cdot \IncrementVec_\Transition \text{ where $\IncrementVec_\Transition$ are the increments of transition $\Transition$}.
+\end{equation}
 
-TODO: rewrite the input and log it through CATRA!
+We start by adding linear constraints requiring transition variables to represent a flow through either automata by requiring that the number of incoming transitions is equal to the number of outgoing ones, and use the automated theorem prover to annotate each transition $\Transition$ with a symbolic representation of its associated integer variable:
+\begin{figure}[h]
+    \centering 
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+    \includegraphics[scale=\autscale]{a_annotated}
+    \caption{Note: unused self-loops at $I_{A'}$ and $F_{A'}$ have been pruned for readability.}\label{fig:aut_a_annotated}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+    \includegraphics[scale=\autscale]{b_annotated}
+    \caption{}\label{fig:aut_b_annotated}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}
+  \caption{The starting automata, without their counter increments and with their transition variables in symbolic form.}\label{fig:propagated}
+\end{figure}
+
+Note that several transitions are already marked as unuseable (have a transition value of $0$).
+This includes any transition for the letter $\Char{d}$, as all counter increments for that variable in $B$ is zero and hence it must also be zero in the product of both automata.
+By removing these transitions to obtain new automata and computing their product, we obtain the following:
 
@@
     \centering 
-\begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{b}
-    \caption{The automaton $B$, where the Parikh registers count the number of occurrences of each letter.}\label{fig:aut_b}
+    \includegraphics[scale=\autscale]{ab}
+  \caption{An automaton encoding the constraint $\CountOf{\Char{a}}(B \times A') > \CountOf{\Char{b}}(B \times A')$.}\label{fig:product}
       %\Description[SHORT]{LONG}
-  \end{subfigure}
+\end{figure}
 
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{a}
-    \caption{The automaton $A'$, whose Parikh registers count both the number of letters and the start ($l$) and length ($r$) of the substring.}\label{fig:aut_a}
-      %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \caption{The collection of automata we use as running examples}\label{fig:examples}
-\end{figure}
+By putting off computing the product $B \times A'$ until after performing linear reasoning on the number of times each transition must be taken, we have constructed a much smaller product than we would have originally.
+Moreover, since this automaton is a stick, all runs through it can be described by the flow-preservation equations described above without the need for costly connectivity-enforcing constraints as in~\cite{generate-parikh-image}.
+This illustrates the advantage of our second laziness: deferred enforcement of connectivity constraints.

### 2023-05-29 main.tex
Add new examples
More work on introduction
Introduction

@@
 
+\usepackage{enumitem}
+
 \theoremstyle{definition}
 \newtheorem{definition}{Definition}[section]
+\newtheorem{example}{Example}[section]
+
+\newlist{constraints}{enumerate}{3}
+\setlist[constraints,1]{%
+    label=\sffamily{(\roman*):},
+    ref=\normalfont{Constraint (\roman*)},
+    wide,itemsep=0pt,topsep=0pt
+}    
+\crefname{constraintsi}{}{}
+\Crefname{constraintsi}{}{}
 
@@
 \subsection{Example instance for \Catra{}}
-\begin{lstlisting}[caption={An example input file for \Catra{} for the problem introduced in \cref{sec:motivation}, illustrating every major syntax element.
+\begin{lstlisting}[caption={An example input file for \Catra{} for \cref{ex:string-constraints} in \cref{sec:intuition}, illustrating every major syntax element.
 From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]
-  // So far we only support integer counters. Note that we have individual
-  // counters for each automaton to avoid surprises for product construction.
-  counter int l_a, l_b, l_c, r_c;
-
-  synchronised {
-  automaton aca_or_bc {
-    init S;
-
-    // We use ASCII values here, this is for lowercase a
-    S -> A [97] { l_a += 1 };
-    S -> B [98] { l_b += 1 };
-
-    B -> S  [98] { l_b += 1 };
-    B -> final [99] { l_c += 1 };
-
-    A -> A [99] { l_c += 1 };
-    A -> final [97] { l_a += 1 };
-
-    accepting final;
-  };
-
-  automaton something_c_something {
-      init S;
-
-      // Special short-hand value for the whole Unicode alphabet.
-      S -> S [any] ;
-      S -> F [99] { r_c += 1 };
-
-      F -> F [any];
-
-      accepting F;
-  };
-  };
-
-  // Constrain the number of a:s to be larger than the number of c:s. Since the
-  // automata are synchronised on every transition and counters are guaranteed
-  // to be consistent these constraints are sufficient without also constraining
-  // r_c.
-  constraint l_a > l_c;
+\input{example.par}
 \end{lstlisting}

### 2023-06-05 introduction.tex
Introduction

@@
-The Parikh image is a characterisation of formal languages in terms of
-their character counts. Given a language over an
-alphabet~$\{a_1, \ldots, a_k\}$, the Parikh image is a set of
-$k$-dimensional vectors that contains some
-vector~$\VectorLiteral{m_1, m_2, \ldots, m_k}$ if and only if the
-formal language contains a word in which each $a_i$ occurs $m_i$
-times. It is a classical result that the Parikh image of every
-context-free language (and, thus, also of every regular language) is a
-semilinear set~\cite{parikh-theorem}, i.e., Presburger-definable.
+The Parikh image is a characterisation of formal languages in terms of their
+character counts. Given a language over an alphabet~$\{a_1, \ldots, a_k\}$, the
+Parikh image is a set of $k$-dimensional vectors that contains some
+vector~$\VectorLiteral{m_1, m_2, \ldots, m_k}$ if and only if the formal
+language contains a word in which each $a_i$ occurs $m_i$ times. It is a
+classical result that the Parikh image of every context-free language (and,
+thus, also of every regular language) is a semilinear set~\cite{parikh-theorem},
+i.e., Presburger-definable.
 
-Parikh images play a central role in many automata-based algorithms,
-for instance and notably in today's string solvers, which often have
-to process constraints that combine regular language membership with
-word length. To decide whether a simple formula like
-$x \in \Language_1 \wedge y \in \Language_2 \wedge |x| > |y|$,
-with string variables~$x, y$ and regular
-languages~$\Language_1, \Language_2$, is satisfiable, it is
-necessary to reason about the sets of word lengths induced by
-$\Language_1, \Language_2$, which are special cases of the Parikh
-image.  This required combined reasoning about strings and string
-length has long been identified as a major bottleneck in string
-solvers
+Parikh images play a central role in many automata-based algorithms, for
+instance and notably in today's string solvers, which often have to process
+constraints that combine regular language membership with word length. To decide
+whether a simple formula like $x \in \Language_1 \wedge y \in \Language_2 \wedge
+|x| > |y|$, with string variables~$x, y$ and regular languages~$\Language_1,
+\Language_2$, is satisfiable, it is necessary to reason about the sets of word
+lengths induced by $\Language_1, \Language_2$, which are special cases of the
+Parikh image.  This required combined reasoning about strings and string length
+has long been identified as a major bottleneck in string solvers
 \cite{DBLP:conf/cav/AbdullaACHRRS15,length-aware-solver,approximate-parikh,DBLP:journals/corr/BerzishZG17}.
-Other string solvers make use of Parikh automata~\cite{parikh-automata}, and thus
-Parikh images in the general case, to handle operations including
-\verb!str.substr!  and \verb!str.at!, which comes at an even higher
-price in terms of computational complexity~\cite{ostrich-plus}.
+Other string solvers make use of Parikh automata~\cite{parikh-automata}, and
+thus Parikh images in the general case, to handle operations including
+\verb!str.substr!  and \verb!str.at!, which comes at an even higher price in
+terms of computational complexity~\cite{ostrich-plus}.
 
-It is possible to compute an existential Presburger formula describing
-the Parikh image of any context-free language in linear
-time~\cite{generate-parikh-image}; for the special case of regular
-languages, this result was also stated in
-\cite{muscholl-linear}. While theoretically elegant, this construction
-has several disadvantages, often making it unpractical for integration
-into algorithms. Firstly, the constructed Presburger formula contains
-a linear number of existential quantifiers in the size of the
-considered grammar, as well as complex Boolean structure, which is
-needed to express the connectedness of sets of productions considered
-in the construction. Eliminating those quantifiers to obtain a
-quantifier-free representation of the Parikh image has exponential
-complexity~\cite{DBLP:conf/issac/Weispfenning97}, and is often impossible in reasonable
-time. Just solving the Parikh image membership problem is NP-complete,
-as it corresponds to computing a satisfying assignment of the
-existential Presburger formula, and taxing for solvers as
+It is possible to compute an existential Presburger formula describing the
+Parikh image of any context-free language in linear
+time~\cite{generate-parikh-image}; for the special case of regular languages,
+this result was also stated in \cite{muscholl-linear}. While theoretically
+elegant, this construction has several disadvantages, often making it
+unpractical for integration into algorithms. Firstly, the constructed Presburger
+formula contains a linear number of existential quantifiers in the size of the
+considered grammar, as well as complex Boolean structure, which is needed to
+express the connectedness of sets of productions considered in the construction.
+Eliminating those quantifiers to obtain a quantifier-free representation of the
+Parikh image has exponential complexity~\cite{DBLP:conf/issac/Weispfenning97},
+and is often impossible in reasonable time. Just solving the Parikh image
+membership problem is NP-complete, as it corresponds to computing a satisfying
+assignment of the existential Presburger formula, and taxing for solvers as
 well~\cite{ostrich-plus}.
 
-Secondly, in applications involving regular languages, it is typically
-necessary to consider the Parikh image not only of a single automaton,
-but of the intersection of multiple automata. This problem arises in
-string solvers in particular, as conjunctions of string constraints
-lead to the computation of length images of products (intersections)
-of regular languages represented as finite automata. Applying the
-approach in~\cite{generate-parikh-image} would in this case require
-the eager computation of the product before its length image, and
-result in an existential Presburger formula of exponential size (in
-the number of automata). In several instances we have observed while
-solving real-world string constraints, the computation of the product
-of automata exhausts the memory of any machine available due to the
-exponential blow-up in size of the product, quickly becoming
-intractable as the number of automata in the product increases. The
-current best published mitigation for this problem is an
-over-approximation that works by approximating the Parikh image of a
-product of automata to be the conjunction of the image of the
-individual automata of the product \cite{approximate-parikh}. This
-approach only works for unsatisfiable instances, and comes with a
-harsh penalty for satisfiable instances.
+Secondly, in applications involving regular languages, it is typically necessary
+to consider the Parikh image not only of a single automaton, but of the
+intersection of multiple automata. This problem arises in string solvers in
+particular, as conjunctions of string constraints lead to the computation of
+length images of products (intersections) of regular languages represented as
+finite automata. Applying the approach in~\cite{generate-parikh-image} would in
+this case require the eager computation of the product before its length image,
+and result in an existential Presburger formula of exponential size (in the
+number of automata). In several instances we have observed while solving
+real-world string constraints, the computation of the product of automata
+exhausts the memory of any machine available due to the exponential blow-up in
+size of the product, quickly becoming intractable as the number of automata in
+the product increases. The current best published mitigation for this problem is
+an over-approximation that works by approximating the Parikh image of a product
+of automata to be the conjunction of the image of the individual automata of the
+product \cite{approximate-parikh}. This approach only works for unsatisfiable
+instances, and comes with a harsh penalty for satisfiable instances.
 
@@
 The key ideas of \Calculus{} is a combination of problem-aware case splitting,
-lazy enforcement of automata connectivity, and lazy computation of products.
+lazy enforcement of automata connectivity, and lazy computation of products. In
+essence, we use constraint-based methods to efficiently enforce the constraints
+associated with computing Parikh images of regular languages.
 
-We implement \Calculus{} as a theory for the \Princess{} automated theorem
-prover in the tool \Catra. \Catra{} also supports the approximate method of
+We implement \Calculus{} as a plug-in theory for the \Princess{} automated
+theorem prover. \Fudge{Others} have integrated it into the similarly
+\Princess{}-backed string constraint solver \OstrichPlus{}, where it is used to
+handle constraints that use cost-enriched automata, as described in
+\cite{ostrich-plus}.
+
+We additionally wrap the Parikh image solver as a stand-alone tool, \Catra.
+\Catra{} also supports a variant of the approximate method of
 \cite{approximate-parikh}, its fall-back variant adapted
@@
 checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
-back-ends on \NrBenchmarks{} Parikh automata intersection problems generated by
-\OstrichPlus{} when solving the PyEx string constraint benchmark suite involving
-string length constraints \cite{pyex}.
+back-ends on \NrBenchmarks{} distinct Parikh automata intersection problems
+generated by \OstrichPlus{} when solving the PyEx string constraint benchmark
+suite involving string length constraints \cite{pyex}.
 
@@
 \begin{itemize}
-    \item The \Calculus{} calculus to efficiently compute (a homomorphism on) the
-          Parikh image of products of Parikh automata.
-    \item Experiments illustrating the performance of \Calculus{} on real-world examples from string solving, including \NrBenchmarks{} instances in a standardised format made available for future study.
-    \item The \Catra{} tool for solving such instances, containing an implementation of \Calculus{}, the over-approximation described in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model checker~\cite{nuxmv}.
-    \item Suggestions for how to efficiently implement \Calculus{} in a modern automated theorem prover, including strategies for case splitting, clause learning, and constraint propagation for connectedness.
+    \item The \Calculus{} calculus to efficiently compute (a homomorphism on)
+          the Parikh image of products of Parikh automata.
+    \item Experiments illustrating the performance of \Calculus{} on real-world
+    examples from string solving, including \NrBenchmarks{} instances in a
+    standardised format made available for future study.
+    \item The \Catra{} tool for solving such instances, containing an
+    implementation of \Calculus{}, the over-approximation described
+    in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model
+    checker~\cite{nuxmv}.
+    \item Suggestions for how to efficiently implement \Calculus{} in a modern
+    automated theorem prover, including strategies for case splitting, clause
+    learning, and constraint propagation for connectedness.
 \end{itemize}
@@
 over-approximate the Parikh image of a product of $k$~automata
-$\ParikhMap(\Language(\Automaton_1) \cap \ldots \cap \Language(\Automaton_k))$ with the
-conjunction of the automata's parikh maps, $\bigwedge_{i=1}^{k}
+$\ParikhMap(\Language(\Automaton_1) \cap \ldots \cap \Language(\Automaton_k))$
+with the conjunction of the automata's parikh maps, $\bigwedge_{i=1}^{k}
 \ParikhMap(\Automaton_i)$. This approach works only for unsatisfiable instances,
@@
 
-Many other generalisations of the Parikh image than the projections we use here have been
-studied. Prominent examples include generalising the Parikh map to segments of a
-fixed length \cite{KARHUMAKI1980155} and the more general Parikh matrix, which
-gives more information about a word than the standard Parikh image
+Many other generalisations of the Parikh image than the projections we use here
+have been studied. Prominent examples include generalising the Parikh map to
+segments of a fixed length \cite{KARHUMAKI1980155} and the more general Parikh
+matrix, which gives more information about a word than the standard Parikh image
 \cite{parikh-matrix}. Another notable generalisation is the p-vector, introduced
@@
     \includegraphics[scale=\autscale]{a}
-    \caption{The automaton $A'$, whose Parikh registers count both the number of letters and the start ($l$) and length ($r$) of the substring matching $\Regex{c\RegexOr{}dd}$.}\label{fig:aut_a}
+    \caption{The automaton $A'$, whose Parikh registers count both the number of
+    letters and the start ($l$) and length ($r$) of the substring matching
+    $\Regex{c\RegexOr{}dd}$.}\label{fig:aut_a}
       %\Description[SHORT]{LONG}
@@
     \includegraphics[scale=\autscale]{b}
-    \caption{The automaton $B$, where the Parikh registers count the number of occurrences of each letter.}\label{fig:aut_b}
+    \caption{The automaton $B$, where the Parikh registers count the number of
+    occurrences of each letter.}\label{fig:aut_b}
       %\Description[SHORT]{LONG}
   \end{subfigure}
-  \caption{The collection of automata we use as running examples}\label{fig:examples}
+  \caption{The collection of automata we use as running
+  examples}\label{fig:examples}
 \end{figure}
 
-In this section we will introduce an intuition for how a string constraint problem in an automata-based solver like \OstrichPlus{} is translated to a Parikh automata intersection problem and solved efficiently using \Calculus{}.
-First, by a Parikh automaton we mean a standard Non-deterministic finite automaton (NFA) with a set of Parikh registers that are incremented (or decremented) at each transition.
-In the automata of \cref{fig:examples}, we will use the notation $a / \begin{bmatrix} x_+ \end{bmatrix}$ to mean that a transition reads an input character $a$ and increments register $x$ by one.
-We will omit any zero-valued increments and assume that all registers are scoped to their automaton.
-The increments are in reality a vector (hence the notation), but as it is mostly sparse here, we instead use the symbolic notation $\begin{bmatrix} x_{1+} \end{bmatrix}$ rather than the more cumbersome $\begin{bmatrix} 0 \\ 1 \\ 0 \\  \vdots \\ 0 \end{bmatrix}$.
+In this section we will introduce an intuition for how a string constraint
+problem in an automata-based solver like \OstrichPlus{} is translated to a
+Parikh automata intersection problem and solved efficiently using \Calculus{}.
+First, by a Parikh automaton we mean a standard non-deterministic finite
+automaton (NFA) with a set of Parikh registers that are incremented (or
+decremented) at each transition. In the automata of \cref{fig:examples}, we will
+use the notation $a / \begin{bmatrix} x_+ \end{bmatrix}$ to mean that a
+transition reads an input character $a$ and increments register $x$ by one. We
+will omit any zero-valued increments and assume that all registers are scoped to
+their automaton. The increments are usually represented as a vector (hence the
+notation), but as it is mostly sparse here, we instead use the symbolic notation
+$\begin{bmatrix} x_{1+} \end{bmatrix}$ rather than the more cumbersome $\left[
+0, 1,  0 , \ldots, 0 \right]$.
 
-Furthermore, we use $\CountOf{c}(A)$ to refer to the number of letters $c$ in $\Language(A)$, and allow this expression to be used both in linear constraints of the input language of \cref{ex:string-constraints} and in the langauge of constraints that we translate to.
-Where the automaton $A$ is apparent, we will omit it.
+Furthermore, we use $\CountOf{c}(A)$ to refer to the number of letters $c$ in
+$\Language(A)$, and allow this expression to be used both in linear constraints
+of the input language of \cref{ex:string-constraints} and in the langauge of
+constraints that we translate to. Where the automaton $A$ is apparent, we will
+omit it.
 
-\begin{example}\label{ex:string-constraints}
-    Consider the following set of string constraints:
+We use standard PCRE regular expressions here and throughout the paper, writing
+them $\Regex{\texttt{in brackets}}$ to make that apparent. As is common,
+$\RegexOr{}$ is alternation, $\mathtt{*}$ the Kleene star, and $\mathtt{.}$,
+matches any single character.
+
+\begin{example}\label{ex:string-constraints} Consider the following set of
+    string constraints:
 \begin{constraints}
-    \item\label{const:x-in-c-dd} $x \in \Regex{c\RegexOr{}dd}$
-    \item\label{const:y-in-b} $y \in \Language(B)$, the language accepted by automaton~$B$ of \cref{fig:aut_b}.
-    \item\label{const:x-substring} $x = \Substring(y, l, r)$
-    \item\label{const:more-a-than-b} $\CountOf{\Char{a}}(y) > \CountOf{\Char{b}}(y)$, that is $y$ contains more letters \Char{a} than letters \Char{b}.
+    \item\label{const:x-in-c-dd} $x \in \Language(\Regex{c\RegexOr{}dd})$.
+    \item\label{const:y-in-b} $y \in \Language(B)$, the language accepted by
+    automaton~$B$ of \cref{fig:aut_b}.
+    \item\label{const:x-substring} $x = \Substring(y, i, n)$, that is $x$ is an
+    $n$-length substring of $y$ starting at offset $i$.
+    \item\label{const:more-a-than-b} $i = n$, that is the substring is as long as its starting offset.
 \end{constraints}
@@
 
-\cref{const:x-in-c-dd,const:y-in-b,const:x-substring} can be translated to the product of Parikh automata $B \times A'$, where $A'$ (\cref{fig:aut_a}) is the Parikh automata obtained by applying the encoding of $\Substring$ described in \cite{ostrich-plus}.
-Here follows an intuition for our method on $B \times A'$ under the global constraint \cref{const:more-a-than-b}.
-Note that the cost registers of \cref{fig:aut_b} are simply counting letters.
+\cref{const:x-in-c-dd,const:y-in-b,const:x-substring} can be translated to the
+product of Parikh automata $B \times A'$, where $A'$ (\cref{fig:aut_a}) is the
+Parikh automata obtained by applying the encoding of $\Substring$ described in
+\cite{ostrich-plus}. That description can then be used to obtain values for
+$\CountOf{\Char{a}}(y)$ and $\CountOf{\Char{b}}(y)$. Here follows an intuition
+for our method on $B \times A'$ under the global constraint
+\cref{const:more-a-than-b}. Note that the cost registers of \cref{fig:aut_b} are
+simply counting letters, while the extra registers $l, r$ in $A'$ also keeps
+track of the matched substring. However, these registers can be mostly
+disregarded here, since we have no constraints on their values.
 
-Our approach to the computation of the Parikh image (and therefore the solution to the intersection problem posed by \cref{ex:string-constraints}) consists of two applications of laziness; late computation of products of automata, and lazy enforcement of connectivity constraints.
-Both are in contrast to the eager approach to finding the Parikh image described in~\cite{generate-parikh-image}.
+Our approach to the computation of the Parikh image (and therefore the solution
+to the intersection problem posed by \cref{ex:string-constraints}) consists of
+two applications of laziness; late computation of products of automata, and lazy
+enforcement of connectivity constraints. Both are in contrast to the eager
+approach to finding the Parikh image described in~\cite{generate-parikh-image},
+that would have started by computing the product $A' \times B$, translated it to
+a Presburger formulae with $\CountOf{\Char{a}}$ etc as free variables, and then
+plugged in the constraint \cref{const:more-a-than-b}.
 
-Starting with an interactive theorem prover and the goal of solving the constraint $\CountOf{\Char{a}}(B \times A') > \CountOf{\Char{b}}(B \times A')$ we associate each transition of $A'$ and $B$ with a fresh integer variable, $\TransitionVar_\Transition$, identically to \cite{generate-parikh-image}.
-These variables represent how many times each transition is taken.
-Hence, the final counter value for each automaton is the element-wise sum:
+Starting with an interactive theorem prover and the goal of solving the
+constraint $\CountOf{\Char{a}}(B \times A') > \CountOf{\Char{b}}(B \times A')$
+we associate each transition of $A'$ and $B$ with a fresh integer variable,
+$\TransitionVar_\Transition$, identically to \cite{generate-parikh-image}. These
+variables represent how many times each transition is taken. Hence, the final
+counter value for each automaton is the element-wise sum:
 \begin{equation}
@@
 
-We start by adding linear constraints requiring transition variables to represent a flow through either automata by requiring that the number of incoming transitions is equal to the number of outgoing ones, and use the automated theorem prover to annotate each transition $\Transition$ with a symbolic representation of its associated integer variable:
 \begin{figure}[h]
@@
     \includegraphics[scale=\autscale]{a_annotated}
-    \caption{Note: unused self-loops at $I_{A'}$ and $F_{A'}$ have been pruned for readability.}\label{fig:aut_a_annotated}
+    \caption{ $A'$ with its associated transition variables in symbolic form.
+    Note: unused ($0$) self-loops at $I_{A'}$ and $F_{A'}$ have been pruned for
+    readability.}\label{fig:aut_a_annotated}
       %\Description[SHORT]{LONG}
@@
     \includegraphics[scale=\autscale]{b_annotated}
-    \caption{}\label{fig:aut_b_annotated}
+    \caption{$B$ with its associated transition variables in symbolic forms.
+    Note that state $B$ is now completely dead.}\label{fig:aut_b_annotated}
       %\Description[SHORT]{LONG}
   \end{subfigure}
-  \caption{The starting automata, without their counter increments and with their transition variables in symbolic form.}\label{fig:propagated}
+  \caption{The starting automata, without their counter increments and with
+  their transition variables in symbolic form.}\label{fig:propagated}
 \end{figure}
 
-Note that several transitions are already marked as unuseable (have a transition value of $0$).
-This includes any transition for the letter $\Char{d}$, as all counter increments for that variable in $B$ is zero and hence it must also be zero in the product of both automata.
-By removing these transitions to obtain new automata and computing their product, we obtain the following:
+We proceed by adding linear constraints requiring transition variables to
+represent a flow through either automata by requiring that the number of
+incoming transitions is equal to the number of outgoing ones. E.g. state $B$
+would have the sum $\cancel{x_{\Tuple{B, \Char{c}, B}}} + x_{\Tuple{I_B,
+\Char{b}, B}} = \cancel{x_{\Tuple{B, \Char{c}, B}}}  + x_{\Tuple{B, \Char{c},
+F_B}} + x_{\Tuple{B, \Char{b}, I_B}}$ where we let $x_{\Tuple{q, l, c'}}$ refer
+to the integer variable associated with a transition from state $q$ to state
+$q'$ with label $l$. Note that the self-loop cancels itself out. This means that
+for automata where a loop can become unreachable, these constraints do not fully
+capture the constraints posed by the automaton, but requires additional
+reachability constraints.
+
+Performing symbolic reasoning on these equations and plugging the corresponding
+symbolic representation of the constraints on each corresponding integer
+variable back as an annotation on its transition, we get obtain the two automata
+in \cref{fig:propagated}.
+
+Several transitions are already marked as unuseable (have a transition variable
+value of $0$). This includes any transition for the letter $\Char{d}$, as
+$\CountOf{\Char{d}}(B) = \sum_{\Transition} 0 = 0$, since the register for
+$\Char{d}$ is zero for all transitions in $B$. By removing such transitions to
+obtain new automata and computing their product, we obtain the product automaton
+of \cref{fig:product}.
 
@@
     \includegraphics[scale=\autscale]{ab}
-  \caption{An automaton encoding the constraint $\CountOf{\Char{a}}(B \times A') > \CountOf{\Char{b}}(B \times A')$.}\label{fig:product}
+  \caption{An automaton encoding the constraint $\CountOf{\Char{a}}(B \times A')
+  > \CountOf{\Char{b}}(B \times A')$.}\label{fig:product}
       %\Description[SHORT]{LONG}
@@
 
-By putting off computing the product $B \times A'$ until after performing linear reasoning on the number of times each transition must be taken, we have constructed a much smaller product than we would have originally.
-Moreover, since this automaton is a stick, all runs through it can be described by the flow-preservation equations described above without the need for costly connectivity-enforcing constraints as in~\cite{generate-parikh-image}.
-This illustrates the advantage of our second laziness: deferred enforcement of connectivity constraints.
+By putting off computing the product $B \times A'$ until after performing linear
+reasoning on the number of times each transition must be taken and using that to
+prune the terms, we have computed a smaller product than we would have with a
+na\"ive approach. Moreover, since this automaton contains only one loop along
+one main path, all runs through it can be described by the flow-preservation
+equations described above without the need for costly connectivity-enforcing
+constraints as in~\cite{generate-parikh-image}. This means that we can now fully
+represent all possible solutions to \cref{ex:string-constraints} by the linear
+equation $\CountOf{\Char{a}} = 2 \land \CountOf{\Char{c}} > 0 \land
+\CountOf{\Char{b}} = 0 \land \CountOf{\Char{d}} = 0$, derived from applying the
+same flow encoding as before.

### 2023-06-05 main.tex
Introduction

@@
 \usepackage{ebproof}
+\usepackage[makeroom]{cancel}
 
@@
 \input{introduction.tex}
-
-\section{A demonstration of our approach}\label{sec:motivation}
-
-To illustrate the key features of \Calculus, we will informally present our
-reasoning. We begin by considering the following set of string and other
-constraints, which we want to solve for all possible values of $\MonoidElement$,
-a value in some commutative monoid and the string $s$, where
-$\varphi(\MonoidElement)$ is some set of constraints on $\MonoidElement$:
-\[
-   s \in \Language(\mathtt{/}\SndRegex{}\mathtt{/})  \land
-   s \in \Language(\mathtt{/}\FstRegex{}\mathtt{/})  \land
-   \MonoidElement = f(s)  \land
-   \varphi(\MonoidElement)
-\]
-
-We will represent the strings using the nondeterministic finite automata
- $\SomethingCSomething{}$ (\cref{fig:something-c-something}) and $\AcaOrBc{}$
- (\cref{fig:aca-or-bc}) respectively.
-
- Throughout these examples we will present our three key ideas:
- \emph{automata-aware splitting}, \emph{lazy enforcement of automata
- connectivity}, and \emph{incremental materialisation of products}.
-
-\subsection{Counting String Lengths}
-
-To keep things simple, in this first example we let $f$ be the function that
-gives the length of a given string; $f\left(\text{curious and curiouser
-still}\right) = 27$, and $f\left(\text{homo}\right) +
-f\left(\text{morphism}\right) = f\left(\text{homomorphism}\right) = 12$, and we
-let $\varphi(\MonoidElement) = \exists k \HoldsThat m = 2k + 1 \land k \geq 0$,
-i.e. require that the length of the string is odd.
-
-\begin{figure}[p]
-  \centering 
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{aca_or_bc}
-    \caption{An automaton recognising the regular expression
-      $\FstRegex{}$.}\label{fig:aca-or-bc}
-      \Description[An NFA recognising $\FstRegex{}$]{An NFA recognising $\FstRegex{}$. It has 4 states.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth} 
-    \centering
-    \includegraphics[scale=\autscale]{something_c_something}
-    \caption{An automaton recognising the regular expression
-    $\SndRegex{}$.}\label{fig:something-c-something}
-    \Description[An NFA recognising $\SndRegex{}$]{An NFA recognising $\FstRegex{}$. It has 2 states.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{aca_or_bc_simplified}
-    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
-    simplification of their equations.}\label{fig:example-simplify-1}
-    \Description[The automaton $\AcaOrBc{}$ labelled with variables counting the number of times each transition is taken, after having simplified the equations.]{The starting state shows an initial flow of 1, with two outgoing transitions to $A$ ($1-l_c'$ on letter a, and $l_b' + l_c'$ on b respectively).
-$A$ has a flow of $l_c$, and its outgoing transition into the final state, $F$ is the same as the incoming one from $S$, the initial state.
-The $B$ transition has an outgoing transition labeled $c/ l_c'$ to the final state $F$, and one backwards transition $b / l_b'$ to the starting state.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{something_c_something_length_simplified}
-    \caption{$\SomethingCSomething{}$ after using linear reasoning over lengths to replace a variable.}\label{fig:example-length-reduced}
-    \Description[The automaton $\SomethingCSomething{}$ after reasoning over lengths.]{The same automaton as before for $\SomethingCSomething{}$, but with the transition from start to final annotated with a constant number 1, and the leftmost self-loop transition replaced with the equation $1 + l_c + 2l_b' -
-    r_\Sigma'$.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-      \includegraphics[scale=\autscale]{aca_or_bc_length_split_1}
-      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ is not
-      used.}\label{fig:aca-or-bc-length-split-1}
-      \Description[The automaton \AcaOrBc{} where we have removed the transition
-      from $S$ to $A$.]{ The state $A$ hangs free without incoming transitions,
-      and only the lower part of the automaton remains.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-      \includegraphics[scale=\autscale]{aca_or_bc_length_split_2}
-      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ \emph{is} used.}\label{fig:aca-or-bc-length-split-2}
-      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ when we keep the transition from $S$ to $A$.
-This removes the transition between $B$ and $F$, but otherwise maintains the automaton.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-      \includegraphics[scale=\autscale]{length_split_product}
-      \caption{The product of $\AcaOrBc$ and $\SomethingCSomething$ after
-      case splitting and removal of dead states.}\label{fig:length-split-product}
-      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ with the transition from $B$ to $F$ removed. There now remains only one path to an accepting state.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-      \includegraphics[scale=\autscale]{aca_or_bc_simplified_3}
-      \caption{$\AcaOrBc{}$ after further constraint propagation.}\label{fig:example-simplify-3}
-      \Description[$\AcaOrBc$ without the lower section.]{$\AcaOrBc$ with only its upper section left and the $a$ transitions now set to static 1.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.6\textwidth}
-    \centering
-        \includegraphics[scale=\autscale]{aca_or_bc_times_something_c_something}
-        \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ under the constraint
-        $\TransitionCount{a} > \TransitionCount{b}$.}\label{fig:final-example} \Description[The
-        final product, still a boring stick automaton with only one loop for an
-        arbitrary number of c's.]{A product of the intermittent automata, which
-        accepts \lstinline{ac.*ca}}
-  \end{subfigure}
-  \caption{The collection of automata we use as running examples in this paper, with annotations illustrating intermittent states of our calculus.}    
-  \end{figure}
-
-The na\"ive approach would be to first compute the product of the automata, then
-use the approach from \cite{generate-parikh-image} to find its image under $f$,
-and then solve the constraints in an automated theorem prover. However, we can
-take advantage of both the constraints and definitions of $f$ to prune the
-automata before computing the product and help reduce the blow-up.
-
-Similar to \cite{generate-parikh-image} we associate each transition
-$\Transition$ with an integer variable $\TransitionVar_\Transition \geq 0$
-representing how many times $\Transition$ is used that we leave implicitly
-existentially quantified. These variables are shown as annotations under the
-transition labels in \cref{fig:aca-or-bc,fig:something-c-something}. We then
-apply graph-like flow reasoning by adding linear equations (e.g. $1 + {l_b}' =
-l_a + l_b$, or $l_a + l_c' = 1$) that equate the incoming transitions into a
-state to the ones going out. For the initial and accepting states we add an
-incoming and outgoing flow of 1, respectively. We will use these linear
-equations to perform reasoning to reduce the size of the product before we
-compute it. Note that we have already performed some linear reasoning to
-eliminate some of the variables in the graphs above, for example to introduce
-the constant number $1$ in \cref{fig:something-c-something}, which comes from $1
-+ r_\Sigma = x + r_\Sigma$ solved for $x$.
-
-We sum the individual transition variables to get the length $\MonoidElement$ we
-are actually interested in. For example the image of $f$ on
-$\SomethingCSomething$ is given by $\exists_{r_\Sigma, r_\Sigma'} \: 2k + 1 =
-r_\Sigma + 1 + r_\Sigma' \land k \geq 0$. For $\SomethingCSomething{}$, this is
-all we need to produce the image since all transitions are always accessible
-from any path from the initial state to the accepting state. The difficult
-connectivity constraint is not necessary to enforce here, since any path through
-$\SomethingCSomething{}$ is connected, and therefore the flow equations
-described above are sufficient to capture its image.
-
-The story gets more complicated with $\AcaOrBc{}$, where we have a choice of two
-paths, up or down, and where that choice affects the possibility of taking
-either of the two loops. In that case, we would have the image $2l_a + l_c + l_c' + l_b + l_b'$.
-If we apply standard linear reasoning on the equations (e.g. $1 + {l_b}' = l_a +
-l_b$, or $l_a + l_c' = 1$), we can reduce the number of bound
-($\exists$-quantified) variables somewhat, obtaining the relations in
-\cref{fig:example-simplify-1}, and the corresponding over-approximation on
-the $f$-image $2 + l_c + 2l_b'$. It is an over-approximation because it only
-considers the number of times each transition is taken and the existence of a
-path from an initial to an accepting state, but omits the connectivity of that
-path in the presence of loops. For example, no valid path would have both $l_c >
-0$ and $1 - l_c' = 0$, but no such restriction exists in the formula above.
-
-Still, this approximate view allows us to perform further reasoning. Since we
-know that the length of a string accepted by the intersection of the two
-automata's languages must be the same, we also know that their images must be
-the same under $f$, e.g. that $r_\Sigma + 1 + r_\Sigma' = 2 + l_c + 2l_b' \iff
-r_\Sigma + r_\Sigma' = 1 + l_c + 2l_b'$. This in turn allows us to replace one
-additional variable, e.g. $r_\Sigma = 1 + l_c + 2l_b' - r_\Sigma'$ (see
-\cref{fig:example-length-reduced}). Note that the two automata are now in
-contact, allowing us to propagate reasoning on one to the other.
-
-Since we cannot take a transition a negative number of times, we have an
-implicit constraint on any transition variable that it must be at least $0$.
-This means that we have $1 + l_c + 2l_b' - r_\Sigma'$, implying an upper bound
-on $r_\Sigma'$: $r_\Sigma' \leq 1 + l_c + 2l_b'$.
-
-This is as far as we can get with this reasoning. We must now choose: either we
-compute (materialise) the product and continue our reasoning, or we perform a
-case split. Let us try a case split in order to put off computing a potentially
-large product.
-
-We apply the principle of splitting and select an early transition in the
-largest automaton with loops, in our case $\FromLabelTo{S}{a}{A}$ of $\AcaOrBc$.
-This creates two cases: $1-l_c' = 0$ and $1-l_c' > 0$. We start with the first
-case, and after propagating the now known values and removing any transition
-that becomes zero, we get the automaton of \cref{fig:aca-or-bc-length-split-1}.
-
-We immediately notice that the self-loop of state $A$ is now disconnected from
-the initial state and must be removed. This condition can be detected
-efficiently using standard forwards/backwards reachability from the initial
-and accepting states respectively. In other words, $2k+1 = 0$ for this case
-and is implied by the splitting constraint. However, no choice of integer $k$
-will satisfy this equation. Therefore, we close this branch and proceed with
-the other one -- where $1-l_c' > 0$. Note that we do so without ever computing
-a potentially expensive product of automata.
-
-Plugging in the assumption $1-l_c' > 0$ (and its consequence that $l_c' = 0$),
-we also get a reduced variant of $\AcaOrBc$, as seen in
-\cref{fig:aca-or-bc-length-split-2}. Note that there is now no way to select any
-transition variable to disconnect a loop from a path between $S$ and $F$,
-and we can therefore stop worrying about enforcing connectivity at all.
-
-Before continuing, we need to tie the inequalities for the transitions of the
-automata of the product to the new transition counting variables we introduced
-in the product automaton to ensure that they are consistent. In our case, we
-obtain the following equations from the principle that a transition variable in
-either automaton must be equal to the sum of variables in the resulting edges of
-the product, which gives the following two interesting inequalities from
-$\AcaOrBc{}$'s transition $\FromLabelTo{A}{c}{A}$: $2k + 1 = r_1 + r_2 + 1$
-
-Note how this conserves the requirement that the number of loop iterations be
-odd even when the same original loop appears three times in the product, twice
-as a self-loop and once as a regular state transition. Using the same principle
-as before for computing the length by adding up the transition counters, we
-obtain the quantified formula $1 + 2k + 1 + 1 + r_b + r_b= 3 + 2k + 2r_b$ for
-the length. Eliminating the final quantifier gives us the image of $f(\AcaOrBc
-\times \SomethingCSomething)$ modulo odd lengths: $3 + 2k$, for some integer
-$k$.
-
-\subsection{Counting Letters}\label{sec:introduction:parikh}
-
-Now assume that rather than the length (e.g. the count of the characters) we
-want to find the whole Parikh image of the intersection of $\AcaOrBc{}$ and
-$\SomethingCSomething{}$. In other words, now $f$ is a mapping to a 3-vector
-counting how often each character occurs in a string. Since extracting values
-from a vector is tiresome, we will use the shorthand notation
-$\TransitionCount{a}$ to refer to the number of letters a that appear. E.g.
-$\TransitionCount{a}(\text{curious and curiouser still}) = 1$. Note that we
-still have the property that $\TransitionCount{o}(\text{homomorphism}) =
-\TransitionCount{o}(\text{homo}) + \TransitionCount{o}(\text{morphism}) = 2 + 1
-= 3$. If you treat the addition as element-wise vector addition, this approach
-works for the vector version as well.
-
-For our constraint, we will use $\TransitionCount{a} > \TransitionCount{b}$,
-that is finding a count of characters in a string accepted by both automata where
-there are more a:s than b:s. Since the underlying calculus is the same, we start immediately from \cref{fig:example-simplify-1}.
-
-  We now need to relate the number of times a transition was taken to the
-  character counts we are really interested in, as opposed to just the length in
-  the previous example. For the $\Sigma$ labels, since we do not know which
-  letter to use, we introduce a sum $r_{a1} + r_{b1} + r_{c1}$ per $\Sigma$
-  transition.
-
-  As in the previous example, we sum the occurrences for each transition to
-  obtain the counts, here with the element-wise vector addition unpacked:
-  \[
-\MonoidElement =   \begin{bmatrix}
-  r_{a1} + r_{a2} \\
-  r_{b1} + r_{b2} \\
-  r_{c1} + r_{c2} + 1 \\
-  \end{bmatrix} =
-  \begin{bmatrix}
-    2 -2l_c' \\
-    2l_b' + l_c' \\
-    l_c + l_c' \\
-    \end{bmatrix} =
-    \begin{bmatrix}
-      \TransitionCount{a} \\
-      \TransitionCount{b} \\
-      \TransitionCount{c}
-      \end{bmatrix}
-  \]
-
-These equalities can be combined with the constraint that $\TransitionCount{a} >
-\TransitionCount{b}$ to obtain $2 - l_c' > l_b'$. Since we have $1-l_c' \geq 0$
-from one of the transition labels, we know $l_c' \geq 0$, and therefore that $1
-> l_b' \implies l_b' = 0$. Plugging in the same inequality again on the same
-principle, we get $2 - 2l_c' > l_c' \iff l_c' = 0$. This gives us the much
-smaller automaton of \cref{fig:example-simplify-3}.
-
-% \begin{figure}[t]
-%   \centering 
-%     \includegraphics[width=0.70\linewidth]{aca_or_bc_simplified_2}
-%     \caption{$\AcaOrBc{}$ after removing the now unused b-transition.}\label{fig:example-simplify-2}
-%     \Description[$\AcaOrBc$ without the returning transition from $B$ to $S$]{$\AcaOrBc$ without the returning transition from $B$ to $S$, which now makes the transition from $S$ to $B$ also have the same associated variable $l_c'$.}
-%   \end{figure}
-
-This will produce a straightforward product since both automata now have the
-same shape, where we essentially only need to expand the self-loops of
-$\SomethingCSomething$ and reduce its range labels to only capture a's, as seen
-in \cref{fig:final-example}. Note that we got this result by only performing
-linear inequality reasoning on the automata of the product versus the number of
-times each transition would be used.
-
-%   \begin{figure}[t]
-% \centering
-%     \includegraphics[width=0.75\linewidth]{original_product}
-%     \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ as it would have
-%     appeared if we had computed it from the initial
-%     automata.}\label{fig:original-product}
-%   \end{figure}
-
-To obtain the final Parikh image, we again need to tie the inequalities for the
-transitions of the automata of the product to the new transition counters we
-introduced in the product automaton. The only interesting transition is the one
-carrying the c, so we will choose that one for illustration. In that case, the
-bridging equation is $r_2 = l_c \land r_2 = r_{\Sigma}$ (and for all others
-$=1$), which gives the Parikh image $\exists_{r_2} \TransitionCount{a} = 1 \land
-\TransitionCount{b} = 0 \land \TransitionCount{c} = r_2 + 1$, and after
-existence-elimination using reasoning on lower bounds (e.g. $r_2 \geq 0$) we
-arrive at $\TransitionCount{c} \geq 1$.
 

### 2023-06-05 introduction.tex
Satisfying assignment

@@
 
+Intuitively, we see that the constraints are satisfiable.
+\cref{const:more-a-than-b} implies that the starting offset must be either $1$
+or $2$, that is we need to fill out $y$ with either one or two symbols before
+the substring $x$ starts. The state $D$ is dead, as $B$ contains no transitions
+with label $\Char{d}$, and hence none of the transitions into $D$ can appear in
+the intersection of the languages. An example satisfying assignment is $y =
+\Char{acca}$, which gives $x = \Char{c}$, $i = 1$, $n = 1$, but others are
+possible.
+
+To solve \cref{ex:string-constraints} using \Calculus{},
 \cref{const:x-in-c-dd,const:y-in-b,const:x-substring} can be translated to the
@@
 \begin{equation}
-\begin{bmatrix} \CountOf{\Char{a}} \\ \CountOf{\Char{b}} \\ \CountOf{\Char{c}} \\ \CountOf{\Char{d}} \end{bmatrix} = \sum_\Transition \TransitionVar_\Transition \cdot \IncrementVec_\Transition \text{ where $\IncrementVec_\Transition$ are the increments of transition $\Transition$}.
+\begin{bmatrix} 
+  \CountOf{\Char{a}} \\ \CountOf{\Char{b}} 
+  \\ \CountOf{\Char{c}} \\ \CountOf{\Char{d}} 
+\end{bmatrix} = \sum_\Transition \TransitionVar_\Transition \cdot 
+  \IncrementVec_\Transition \text{ where $\IncrementVec_\Transition$ are the increments of transition $\Transition$}.
 \end{equation}

### 2023-06-06 introduction.tex
SMT-LIB example

@@
 Parikh automata obtained by applying the encoding of $\Substring$ described in
-\cite{ostrich-plus}. That description can then be used to obtain values for
-$\CountOf{\Char{a}}(y)$ and $\CountOf{\Char{b}}(y)$. Here follows an intuition
-for our method on $B \times A'$ under the global constraint
-\cref{const:more-a-than-b}. Note that the cost registers of \cref{fig:aut_b} are
-simply counting letters, while the extra registers $l, r$ in $A'$ also keeps
-track of the matched substring. However, these registers can be mostly
-disregarded here, since we have no constraints on their values.
+\cite{ostrich-plus}. That description can then be used to obtain values for the
+subsstring offsets $i, n$. Here follows an intuition for our method on $B \times
+A'$ under the global constraint on the substring \cref{const:more-a-than-b}.
 
@@
 that would have started by computing the product $A' \times B$, translated it to
-a Presburger formulae with $\CountOf{\Char{a}}$ etc as free variables, and then
-plugged in the constraint \cref{const:more-a-than-b}.
+a Presburger formula with $i, n$ etc as free variables, and then plugged in the
+constraint \cref{const:more-a-than-b}.
 
@@
 symbolic representation of the constraints on each corresponding integer
-variable back as an annotation on its transition, we get obtain the two automata
-in \cref{fig:propagated}.
+variable back as an annotation on its transition, we obtain the two automata in
+\cref{fig:propagated}.
+
+% FIXME B-loopen inte nollad här! Uppdatera!
 
@@
 $\CountOf{\Char{d}}(B) = \sum_{\Transition} 0 = 0$, since the register for
-$\Char{d}$ is zero for all transitions in $B$. By removing such transitions to
-obtain new automata and computing their product, we obtain the product automaton
-of \cref{fig:product}.
+$\Char{d}$ is zero for all transitions in $B$.
+
+Note that the self-loop on state $B$ is nonzero due to the above mentioned
+self-balancing. However, a reachability computation will conclude that the state
+is unreachable, and allow us to conclude that all its outgoing transitions must
+be zero.
+
+After removing all transitions concluded to be unuseable to produce new,
+equivalent automata, we compute the product in \cref{fig:product}.
 
@@
 equations described above without the need for costly connectivity-enforcing
-constraints as in~\cite{generate-parikh-image}. This means that we can now fully
-represent all possible solutions to \cref{ex:string-constraints} by the linear
-equation $\CountOf{\Char{a}} = 2 \land \CountOf{\Char{c}} > 0 \land
-\CountOf{\Char{b}} = 0 \land \CountOf{\Char{d}} = 0$, derived from applying the
-same flow encoding as before.
+constraints as in~\cite{generate-parikh-image}. We can now obtain a generalised
+solution to our constraints by inspection.

### 2023-06-06 main.tex
SMT-LIB example

@@
 \end{figure}
+
 \subsection{Example instance for \Catra{}}
-\begin{lstlisting}[caption={An example input file for \Catra{} for \cref{ex:string-constraints} in \cref{sec:intuition}, illustrating every major syntax element.
-From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]
-\input{example.par}
-\end{lstlisting}
-
-% Text of appendix \ldots
+\lstinputlisting[caption={An example input file for \Catra{} for \cref{ex:string-constraints} in \cref{sec:intuition}, illustrating every major syntax element.
+From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]{example.par}
+
+\subsection{SMT-LIB encoding of the string constraint example}
+\lstinputlisting[caption={An SMT-LIB encoding of \cref{ex:string-constraints}.}, label=lst:string-constraints]{example.smt2}
 

### 2023-06-06 introduction.tex
Use a nice green colour rather than slashes to denote regexes
Closes: #41

@@
 We use standard PCRE regular expressions here and throughout the paper, writing
-them $\Regex{\texttt{in brackets}}$ to make that apparent. As is common,
-$\RegexOr{}$ is alternation, $\mathtt{*}$ the Kleene star, and $\mathtt{.}$,
-matches any single character.
+them $\Regex{\texttt{like this}}$. This means that $\RegexOr{}$ is alternation,
+$\mathtt{*}$ the Kleene star, and $\mathtt{.}$, matches any single character.
 

### 2023-06-06 appendix.tex
Split out the example into its own file

--- 2023-06-06 main.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2023}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\usepackage{amsmath,empheq,fancybox}
-\usepackage{paralist}
-\usepackage{url}
-\usepackage{color}
-\usepackage{textcomp,listings}
-\usepackage{array}
-\usepackage{mymacros}
-\usepackage{microtype}
-\usepackage{listings}
-\usepackage{csquotes}
-\usepackage{proof}
-\usepackage[capitalise]{cleveref}
-\usepackage{algorithm2e}      
-\usepackage{multirow}
-\usepackage{mathpartir}
-\usepackage{amsthm}
-\usepackage{numprint}
-\usepackage{ebproof}
-\usepackage[makeroom]{cancel}
-
-\usepackage{mathtools} % Bonus
-
-\usepackage{enumitem}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{example}{Example}[section]
-
-\newlist{constraints}{enumerate}{3}
-\setlist[constraints,1]{%
-    label=\sffamily{(\roman*):},
-    ref=\normalfont{Constraint (\roman*)},
-    wide,itemsep=0pt,topsep=0pt
-}    
-\crefname{constraintsi}{}{}
-\Crefname{constraintsi}{}{}
-
-\lstset{
-    columns=fullflexible,
-    showspaces=false,
-    showtabs=false,
-    breaklines=true,
-    showstringspaces=false,
-    breakatwhitespace=true,
-    escapeinside={(*@}{@*)},
-    commentstyle=\color{greencomments},
-    keywordstyle=\color{bluekeywords},
-    stringstyle=\color{redstrings},
-    numberstyle=\color{graynumbers},
-    basicstyle=\ttfamily\small,
-    framesep=12pt,
-    xleftmargin=12pt,
-    tabsize=4,
-    captionpos=b
-}
-
-\newif\ifcomments
-\commentstrue
-\newif\ifoutline
-\outlinetrue
-
-\newcommand{\contents}[1]{\ifoutline{\color{blue}
-    \begin{itemize}
-    #1
-    \end{itemize}
-  }\fi}
-
-\allowdisplaybreaks[1]
-
-\begin{document}
-
-%% Title information
-\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
-   counts in strings of a given language. Automata-based string solvers
-   frequently need to compute the Parikh image of large products (intersections)
-   of nondeterministic automata, which in many operations is both prohibitively
-   slow and memory-intensive. We contribute a novel understanding of how Parikh
-   maps can be tackled as a constraint solving problem to solve real-world
-   constraints stemming from functions on regular languages, most notably the
-   length constraint. Furthermore, we show how this formulation can be
-   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
-   prover supporting Presburger logic.  We implement \Calculus{} in a tool
-   called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving Parikh automata
-   intersection problems produced when solving standard string constraint
-   benchmarks involving string length constraints. We show that our solution
-   strictly outperforms the standard approach described
-   in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wide
-   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
-   model checker.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-  <ccs2012>
-  <concept>
-  <concept_id>10003752.10003766.10003773.10003775</concept_id>
-  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003766.10003776</concept_id>
-  <concept_desc>Theory of computation~Regular languages</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10003794</concept_id>
-  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10010124.10010138.10010142</concept_id>
-  <concept_desc>Theory of computation~Program verification</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10011192</concept_id>
-  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  </ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Quantitative automata}
-\ccsdesc[300]{Theory of computation~Regular languages}
-\ccsdesc[500]{Theory of computation~Automated reasoning}
-\ccsdesc[300]{Theory of computation~Program verification}
-\ccsdesc[300]{Theory of computation~Verification by model checking}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}\label{sec:introduction}
-\input{introduction.tex}
-
-\section{Preliminaries}
-
-\subsection{Monoids}
-
-A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
-consisting of the non-empty carrier set of elements, $X$, an associative binary operation
-$X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,
-c \in X$, $(a \MonoidOp b) \MonoidOp c = a \MonoidOp (b \MonoidOp c)$. Finally,
-$\Monoid$ must have an identity element $0_{\Monoid} \in X$ such that
-$0_{\Monoid} \MonoidOp a = a \MonoidOp 0_{\Monoid} =   a$ for every $a \in X$.
-We sometimes use integer multiplication to represent repeated application of
-$\MonoidOp$, e.g. $3a = a \MonoidOp a \MonoidOp a$, for $a \in X$. $\Monoid$ is
-called \textit{commutative} if $\MonoidOp$ also commutes, that is if $a
-\MonoidOp b = b \MonoidOp a$ for all $a, b \in X$.
-
-Finally, a \textit{homomorphism} is a structure-preserving map between two monoids
-$\Monoid_1, \Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
-\MonoidOp_1 b) = \Map(a) \MonoidOp_2 \Map(b)$, where $S_1, S_2$ are the carrier
-sets of $\Monoid_1$ and $\Monoid_2$ respectively, and $\MonoidOp_1, \MonoidOp_2$
-their binary operations.
-
-\subsection{Languages, Finite-state Automata and their Products}
-
-We define an alphabet as a finite set of symbols $\Alphabet$ with words $\Strings$, and
-the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
-Note that $\Strings = \Tuple{\Alphabet;\Concat{};\epsilon}$, is a
-non-commutative monoid, referred to as the free monoid. The string length
-function, $\Length{s}$ is an example of a homomorphism between $\Strings$
-and~$\mathbb{Z}$.
-
-An automaton~$\Automaton$ with alphabet~$\Alphabet$ is
-$\AutomatonTuple$ where $\Transitions \subseteq \States \times \Alphabet \times
-\States$, $\States$ is its states, $\InitialState$ its
-initial state, and $\AcceptingStates$ is its set of accepting states.  We
-write a transition $\Transition = \Tuple{\State, \Label, \State'} \in
-\Transitions$ as $\Transition = \FromLabelTo{\State}{\Label}{\State'}$.
-Similarly, we use the notation $\FromLabelTo{\State}{}{}$ to refer to the set of
-transitions starting in $\State$, and $\FromLabelTo{}{}{\State}$ to refer to the
-set of transitions coming into $\State$, whenever the automaton is clear from
-the context.
-
-We will let variables $\Transition, \Transition', \Transition_1, \ldots,
-\Transition_n$ etc describe transitions, $\State, \ldots, \State_n$ states, and
-$\Automaton, \ldots, \Automaton_n$ automata, and use subscript indexing
-($\Transitions_\Automaton$) to refer to the transitions, states, etc of a given
-automaton.
-
-By a \emph{product} of two automata $\Automaton_1, \Automaton_2$, written
-$\Automaton_1 \times \Automaton_2$, we mean an automaton constructed to run
-$\Automaton_1$ and $\Automaton_2$ in parallel on an input and only accept the
-input if both automata would do so.
-
-We refer to the resulting product states as tuples, $\Tuple{\State, \State'}$,
-which represent the state of the product automaton where $\Automaton_1$ would be
-in $\State$ and $\Automaton_2$ would be in $\State'$. Note that since we use
-ordered tuples the product is technically (but w.l.o.g) not commutative; the
-left-hand-side must come from the left-hand term. The sole purpose of this
-matching is to allow us to speak with precision about the origins of components
-in a product.
-
-\subsection{The Parikh Map and its Image}
-Formally, the \textit{Parikh map} over an alphabet $\Alphabet=
-\left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:
-$$
-\begin{aligned}
-& \ParikhMap: \MapFromTo{\Strings}{\natural^k} \\
-& \ParikhMap(s) = \VectorLiteral{\#a_1(s), \#a_2(s), \ldots, \#a_k(s)}
-\end{aligned}
-$$
-
-That is, $\ParikhMap(s)$ is a vector of the number of occurrences of each
-character in the language for a given string $s$. For example, for  $\Alphabet =
-\Set{a, b}$, we would have $\ParikhMap(abb) = \VectorLiteral{1, 2}$.
-
-We define the image of this map, the \textit{Parikh image}, of some subset of
-the language $\Language \subseteq \Strings$ as:
-\[
-\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
-\]
-
-Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
-1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
-notation $\#l(w)$ to talk about an individual letter $l$ in a word $w$. For
-example, for the Parikh vector above, we would have $\CountOf{a} = 1$. You have
-already seen this in \cref{sec:introduction:parikh}.
-
-Parikh's theorem states that any context-free language has a letter-equivalent
-regular language (c.f.~\cite{construction} for a construction of such automata
-from context-free grammars and~\cite{bounds} for bounds on its size). However,
-there are languages that are not context-free that also have semilinear images
-under~$\ParikhMap$ (e.g. $\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) =
-\ParikhMap((abc)^*) = \CountOf{a} = \CountOf{b} = \CountOf{c} \land \CountOf{a}
-\geq 0$). This means they can be represented as a quantifier-free Presburger
-formula.
-
-Note that while Parikh's theorem applies to context-free languages, in this
-paper we focus only on regular languages.
-
-\subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
-\label{sec:verma}
-
-Since Parikh images are semilinear, any Parikh image can be written as a set of linear
-equations. The following construction for an NFA $\Automaton = \AutomatonTuple$ is based on the version for context-free grammars presented in \cite{generate-parikh-image}:
-\begin{equation}
-\begin{aligned}
-\ParikhMap(\Automaton) := 
-& \AndComp{\Letter \in \Alphabet}{\LetterVar_{\Letter} = \sum_{\Transition = \FromLabelTo{}{\Letter}{} \in \Transitions} \TransitionVar_{\Transition}
-}\\
-& \AndComp{\State \in \AcceptingStates}{\FinalStateVar_{\State} > 0 \longrightarrow  \StateVar_{\State} > 0}\\
-& \AndComp{\State \in \States}{
-  \left(\FinalStateVar_{\State} \text{ if } \State \in \AcceptingStates \text{ otherwise } 0 \right) +
-  \sum_{\Transition \in \FromLabelTo{\State}{}{}} \Filter(\Transition) - \sum_{\Transition \in \FromLabelTo{}{}{\State}} \Filter(\Transition)
-= \begin{cases}
-    1 \text{  if $\State = \InitialState$} \\
-    0 \text{ otherwise}
-  \end{cases}
-}\\
-& \AndComp{\Transition = \FromLabelTo{\State'}{}{\State} \in \Transitions}{
-  \TransitionVar_{\Transition} > 0 \longrightarrow \StateVar_{\State} > 0
-} \\
-& \AndComp{\State \in F}{
-  \StateVar_{\State} > 0 \longrightarrow \FinalStateVar_{\State} = 1
-} \\
-&\AndComp{\State \in \States}{
-  \StateVar_{\State} > 0 \longrightarrow
-  \OrComp{\Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions}{
-    \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
-    \TransitionVar_{\Transition} \geq 1 \land
-  \StateVar_{\State'} \geq 1
-    }
-}
-\end{aligned}
-\label{eq:generate-parikh}
-\end{equation}
-
-All variables $\TransitionVar_\Transition, \StateVar_\State, \FinalStateVar_\State$ are
-existentially quantified and the free variables $\LetterVar_\Letter$ make up the
-image. $\StateVar_\State$ represents the distance of state $\State$ from
-$\InitialState$ in a spanning tree, $\TransitionVar_\Transition$ how many times
-a transition $\Transition$ is used, and $\FinalStateVar_\State$ whether a given
-accepting state $\State \in \AcceptingStates$ is the actually used final state.
-
-In this paper we refer to this model as the baseline approach, though we also
-apply optimisations as described in \cref{sec:implementing-baseline}. The
-calculus introduced in this paper, by contrast, lazily enforces the
-connectedness constraint of this encoding (the final three clauses) while also
-interleaving the computation of products of automata and propagating information
-between the steps to reduce the amount of work that needs to be done.
-
-\subsection{Theorem Proving}
-
-We assume basic familiarity with first-order logic, Presburger
-arithmetic, and the classical sequent calculus. For reference, see
-e.g. \cite{Fitting96a}.
-
-\section{Projections on Parikh Images}\label{sec:generalised}
-
-It is easy to see that the Parikh map~$\ParikhMap$ represents a
-homomorphism from the (free) non-commutative monoid~$\Strings$ to the
-(free) commutative monoid~$\Naturals^k$. As we are often not
-interested in the full Parikh image, but only in some projections of
-it. We consider \emph{arbitrary} homomorphisms
-$\Map:\: \Sigma^* \to \Monoid$, where
-$\Monoid = \left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative
-monoid. We give several
-examples of such projections on Parikh images later in this section.
-
-Observe that every homomorphism~$\Map:\: \Sigma^* \to \Monoid$ can be
-represented as the composition~$h' \circ \ParikhMap$, for some
-homomorphism~$h' : \Naturals^k \to M$. One of the insights underlying our
-approach is that it is more efficient to directly compute a
-projection on the Parikh image~$\Map(\Language)$, than to first compute the
-standard image~$\ParikhMap(\Language)$ followed by projection to
-some property of interest.
-
-%\subsection{Applications of generalised Parikh images}\label{sec:applications}
-
-\subsection{Example~1: Reasoning about String Length}
-
-One such simplifying homomorphism can express string
-length, the problem that originally motivated our study of the Parikh
-map. This mapping is relevant when solving constraints that combine
-language membership as well as word length, for instance the formula
-\begin{equation}\label{eq:stringLength}
-x \in \Language_1 \wedge y \in \Language_2 \wedge |x| > |y|
-\end{equation}
-mentioned in the introduction. To solve this formula, let
-$\Monoid = \Naturals$, and define the homomorphism~$L$ by $L(a) = 1$
-for all characters~$a \in \Alphabet$. The length of a string
-$s = s_1 \RepeatSum{\Concat} s_n$ is given by
-$L(s) = \sum L(s_i) = 1 \RepeatSum{+} 1 = n$, and to solve
-\eqref{eq:stringLength} we can instead solve the equi-satisfiable
-formula~$\alpha \in L(\Language_1) \wedge \beta \in L(\Language_2)
-\wedge \alpha > \beta$. This paper proposes efficient native
-procedures to reason about membership constraints
-like~$\alpha \in L(\Language_1)$, avoiding the computation of
-the complete image~$L(\Language_1)$.
-
-\subsection{Example~2: Generalised String Constraints with Integer
-  Datatype}\label{sec:parikh-automata}
-
-Parikh images are also applicable for deciding more general classes of
-string constraints~\cite{ostrich-plus}. Consider, for instance, a
-constraint involving the substring operation:
-\begin{equation}
-  \label{eq:substring}
-  x \in \Language_1 \wedge \; 0 \leq n \leq m \leq |x| \wedge x[n:m] \in \Language_2
-\end{equation}
-in which $x$ is a string variable, $n, m$ are integer variables, and
-$x[n:m]$ denotes the substring of $x$ starting at position~$n$ running
-until position~$m$. This constraint belongs to an expressive fragment
-of string logic that cannot be decided by most state-of-the-art string
-solvers. To illustrate the decision procedure proposed in
-\cite{ostrich-plus} for string constraints of this kind, suppose that
-$\Language_2$ is defined by the regular
-expression~$\mathtt{(ab|c)*}$. An automaton recognising this language
-is shown in \cref{fig:parikh-automata} (left).
-
-We can model \eqref{eq:substring} using the notion of a \emph{Parikh
-  automaton}~\cite{parikh-automata,expressiveness}. A Parikh automaton
-is an automaton in which transitions are labelled both with characters
-from the alphabet~$\Alphabet$, and with offset vectors defining the
-increments of a finite number of counters. This means that Parikh
-automata recognise words over an extended
-alphabet~$\Alphabet \times D$, where $D \subseteq \Naturals^d$ is a
-finite set of increment vectors (notation as
-in~\cite{expressiveness}). We use the symbols~$\pi_\Alphabet, \pi_D$
-to denote projections to the first and the second component of a
-composite letter~$(a, \Vector{v})$, respectively, and extend those
-projections to words:
-\begin{equation*}
-  \pi_\Alphabet((a_1, \Vector{v}_1) \RepeatSum{\Concat} (a_k, \Vector{v}_k))
-  ~=~ a_1 \RepeatSum{\Concat} a_k,
-  \qquad
-  \pi_D((a_1, \Vector{v}_1) \RepeatSum{\Concat} (a_k, \Vector{v}_k))
-  ~=~ \Vector{v}_1 \RepeatSum{+} \Vector{v}_k.
-\end{equation*}
-
-\begin{definition}\label{def:parikh-automata} A Parikh automaton of dimension $d
-  \geq 0$ is a pair $\Tuple{\Automaton, C}$, where
-  $C \subseteq \Naturals^d$ is a semi-linear set (or, equivalently, a
-  Presburger formula), and $\Automaton$ is a finite automaton with the
-  alphabet $\Alphabet \times D$, where $D \subseteq \Naturals^d$. We
-  say that $\Tuple{\Automaton, \varphi}$ recognises a
-  word~$w \in \Alphabet^*$ if and only if the automaton has a run
-  accepting an extended word~$w' \in (\Alphabet \times D)^*$ such that
-  $\pi_\Alphabet(w') = w$ and $\pi_D(w') \in C$.
-\end{definition}
-
-\begin{figure}[t]
-  \centering
-  \raisebox{5ex}{
-    \includegraphics[scale=0.9]{counter_automaton_pre}
-    }
-  \hfill
-  \includegraphics[scale=0.9]{counter_automaton}
-  \caption{An automaton recognising the language $\mathtt{(ab|c)*}$
-    (left), and a Parikh automaton describing the pre-image of this
-    language under the substring operation~$\cdot[n:m]$ (right). For
-    sake of presentation, the Parikh automaton contains
-    $\epsilon$-transitions, which could be eliminated in the standard
-    way.}\label{fig:parikh-automata}
-    \Description[An 2-state automaton (left) and a slightly larger automaton with counters (right).]{
-      The rightmost automaton is the starts with an initial state with a self-loop on any character that increments both registers by one, followed by an epsilon transition to a state with a self-loop on c that increments the lower register by one.
-From that state, the automaton has an epsilon-transition into the accepting state, as well as a transition on a into a state with only a transition back again, both incrementing the lower register by one.
-The accepting state has a self-loop on any character that does nothing to the registers.
-      }
-    \end{figure}
-
-    Applied to \eqref{eq:substring}, the decision procedure in
-\cite{ostrich-plus} will construct a pre-image of $\Language_2$ under
-the substring operation~$\cdot[n:m]$, and check whether this pre-image
-is consistent with the constraint~$x \in \Language_1$. Because the
-substring operation depends on the values of the integer
-variables~$n, m$, a Parikh automaton of dimension~$2$ is a suitable
-formalism to describe the pre-image, resulting
-in the automaton in Fig.~\ref{fig:parikh-automata} (right). The first
-component of the increment vectors is used to count the number of
-letters eliminated in the beginning of the string (value~$n$), while
-the second component records the beginning of the eliminated suffix
-(value~$m$).
-
-Denoting the language described by Fig.~\ref{fig:parikh-automata}
-(right) as $\Language_{pre}$, we can then replace \eqref{eq:substring}
-with an equi-satisfiable formula that no longer contains any explicit
-substring operation:
-%
-\begin{equation}
-  \label{eq:parikh-constraint}
-  p \in \Language_{pre} \wedge~
-  \pi_\Alphabet(p) \in \Language_1
-  \wedge~~ \pi_D(p) =
-  \begin{bmatrix}
-    n \\ m
-  \end{bmatrix}~~
-  \wedge 0 \leq n \leq m \leq |p|
-\end{equation}
-%
-To check the satisfiability of \eqref{eq:parikh-constraint}, we need a
-decision procedure that can process intersections of regular languages
-(in this case, of $\Language_{pre}$ and $\Language_1$, synchronising
-on $\Alphabet$), while imposing the side
-condition~$0 \leq n \leq m \leq |p|$ on the increment sum. In
-\cite{ostrich-plus}, this decision procedure turned out to be main
-bottleneck of the string solver, which was one of the motivations to
-develop the lazy algorithm proposed in this paper.
-
-\iffalse
-\begin{figure}[t]
-  \centering
-      \includegraphics[width=0.5\linewidth]{parikh_automaton}
-      \caption{The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
-      $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
-      formula containing the constraints on the final register values cannot be
-      visualised.}\label{fig:parikh-automaton}
-    \end{figure}
-\fi
-
-\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
-
-We start by defining our calculus, \Calculus{}, for one automaton, and only
-extend it to products of automata in \cref{sec:multiple}. Assume an NFA
-$\Automaton = \AutomatonTuple$ with $\NrTransitions$ transitions $\Transitions =
-\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. For convenience, we then
-introduce the following supporting notations:
-
-\begin{definition}
-  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
-  transitions $\Transitions$ represents a path through $\Automaton$ using
-  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
-  \State_{k+1})$ holds), passing zero or more labels $\Label_1, \ldots,
-  \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
-  \InitialState$. The end state, $\State_n$, is not necessarily accepting. If it
-  is, we say the path is also \textit{accepting}.
-  \end{definition}
-
-\begin{definition}
-  Moreover, we talk about the \textit{set of paths} of an automaton,
-  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
-  valid paths through $\Automaton$. Additionally, we use the
-  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
-  state~$\State$.
-\end{definition}
-
-\begin{definition}
-  For a path $\Path = \Tuple{\State_0 \Label_1 \State_1 \RepeatSum{,}
-  \State_n}$, its \textit{word} $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat}
-  \Label_n$ is the word read out on its labels.
-\end{definition}
-
-\begin{definition}
-  The \textit{states} of a path $\Path = \Tuple{\State_0 \Label_1 \State_1
-  \RepeatSum{,} \State_n}$, $\StatesOf(\Path) = \Set{\State_0\RepeatSum{,}
-  \State_n}$ are the states visited along $\Path$. Note that $\InitialState \in
-  \StatesOf(\Path)$ for every path since all paths start in the initial state.
-\end{definition}
-
-\begin{definition}
-  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$ to state
-  $\State$, written $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of
-  transitions such that every accepting path $p$ where $\State \in
-  \StatesOf(\Path)$ contains a transition $\Transition \in C$. A cut does not
-  exist for every state; notably it never exists for $\InitialState$.
-\end{definition}
-
-\begin{definition}
- The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
- number of times a transition $\Transition =
- \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
- $p$.
-\end{definition}
-
-We then introduce the two predicates into our calculus with the following
-definitions:
-
-\begin{definition}\label{def:single-image}
-  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
-  $\Automaton = \AutomatonTuple$, modulo some map $\Map : \Alphabet^* \to \Monoid$ to a
-  commutative monoid $\Monoid$ as described in \cref{sec:generalised}
-  and with a transition selection function
-  $\Filter:\: \Transitions \to \Naturals$ holds when
-  $\MonoidElement \in \Monoid$ is an element of the Parikh image of
-  $\Language(\Automaton)$
-  modulo $\Map$, or more formally when there is an accepting path
-  $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
-  that $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
-  for all $t \in \Transitions$, and $\MonoidElement = \Map(\WordOf(\Path))$.
-\end{definition}
-
-\begin{definition}
-  $\Connected(\Automaton, \Filter)$ for some automaton $\Automaton =
-  \AutomatonTuple$ holds when for every $\Transition =
-  \FromLabelTo{\State}{}{\State'} \in \Transitions$, $\Filter(\Transition) > 0
-  \implies \exists \Path \in \Paths(\Automaton)$ $\Filter(\Transition_i) > 0$
-  for $\Transition_i \in p$, and $\State \in \StatesOf(p)$, or in words that
-  there exists some $\Filter$-selected valid path that reaches $\Transition$'s
-  starting state, $\State$. Intuitively, it represents the condition that
-  $\Automaton$ is connected with respect to the selection function $\Filter$ for
-  every transition. It is redundant to $\Image$ by design.
-\end{definition}
-
-We present the rules of \Calculus{} for one automaton in
-\cref{tbl:rules:single}. The rules operate on sets of formulas, and
-can be interpreted as rules of a one-sided sequent calculus, in
-which all formulas are located in the antecedent~\cite{Fitting96a}.
-Each of the rules should be read and
-applied bottom-up, and relates
-premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
-conclusion~$\SomeClause$. When constructing a proof, we start from some
-root~$\SomeClause$, and then apply proof rules to the goals of the proof
-until a goal can be closed, or no more rule is applicable.
-
-A proof in which all proof goals could be closed shows that the formulas
-in the root~$\SomeClause$ of the proof are inconsistent, and do not
-have any solutions. A proof goal that could not be closed, but to which no
-more rules are applicable, gives rise to  a solution of the formulas in the
-root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic, so that a solution can be computed using
-standard algorithms~\cite{Fitting96a}.
-
-We use the convention of splitting the
-formulas into linear (in)equalities ($\SomeInequalities$)
-and other formulas ($\SomeClause$), and assume that predicates
-$\Image$ and $\Connected$ only occur positively.
- Since
-our rules operate by adding and matching linear (in)equalities in a proof goal,
-we use the shorthand of listing the matched inequalities as antecedents
-e.g., in \Propagate{}).
-The filtering function~$\Filter$ is evaluated symbolically, and can in practice
-be read as a function from transitions to $\Naturals$-valued terms
-(e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{}, described in
-\cref{sec:implementation}, $\Filter$~is a vector of fresh variables with the
-same size as~$\Transitions$.
-
-We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
-transitions of an automaton~$\Automaton$. Additionally, for an
-automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
-like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
-\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
-$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
-instance, the basis for the matched linear inequalities is
-implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
-\left(X;\MonoidOp;0_{\Monoid}\right)$ and the map into it $\Map : \Strings
-\rightarrow X$, we also allow mapping over transitions:
-$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$.
-
-We require that rules can only be applied when they add new formulas
-on every created branch (the notion of \emph{regularity} of a proof is
-required~\cite{Fitting96a}). For example, this means that \Split{} can
-only be applied to proof goals that contain neither
-$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
-never be applied to split on the same term twice on the same branch.
-This suggests a proof strategy where you \Propagate{} when you can, \Split{}
-when you must, and \Subsume{} when neither is possible anymore.
-
-\begin{table}
-\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-  Name & \text{Rule} & Side conditions\\
-  \midrule
-
-  % EXPAND
-  \Expand & 
-    \inferrule
-  {\Connected(\Automaton, \Filter) ,~ \FlowEq(\Automaton, \Filter) ,~ \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
-  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
-  None \\[4ex]
-
-  % SPLIT
-  \Split & 
-  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  if $\Transition \in \Transitions_\Automaton$ \\[4ex]
-
-  % PROPAGATE
-  \Propagate &
-  \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
-  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[4ex]
-
-  % SUBSUME
-  \Subsume &
-  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  \Split{} and \Propagate{} cannot be applied \\
-  \bottomrule
-  \end{tabular}
-  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
-\end{table}
-We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
-of existentially quantified linear inequalities with the following definition,
-where we assign fresh, existentially quantified variables to
-$\FinalStateVar_\State, \Filter(\Transition)$ for every $\State \in
-\AcceptingStates, \Transition \in \Transitions$:
-\[
-\begin{aligned}
-  \FlowEq(\Automaton, \Filter) &= \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
-  \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}\\
-  \Sink(\State) & = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
-  \In(\State, \Filter) & = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
-  \StartFlow(\State)  &= 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
-  \Out(\State, \Filter) &= \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
-\end{aligned}
-\]
-
-In addition to \cref{tbl:rules:single}, we assume the existence of a rule
-\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
-formulae, and for the elements of~$\Monoid$.
-
-The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
-$\Automaton$. It states that we are only allowed to use transitions attached to
-a reachable state, and is necessary to ensure connectedness in the presence of
-cycles in~$\Automaton$.
-
-\textsc{Expand} expands the predicate into its most basic rules; one set of
-linear equations synchronising the transitions mentioned by~$\Filter$ to the
-corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
-the standard Parikh image formulation, as described by~\FlowEq. Since
-$\Connected$ and $\Image$ are partially redundant and the difference is covered
-by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
-this sense, we split the semantics of the $\Image$~predicate into its counting
-aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
-$\Connected$).
-
-Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
-contested transition from a potential solution before concluding that it must be
-included. Intuitively, this is what guarantees our ability to make forward
-progress by eliminating paths through~$\Automaton$.
-
-A decision procedure for our predicate in a tableau-based automated theorem
-prover would start by expanding the predicate using the $\Expand{}$~rule. A
-theorem prover would perform algebraic substitution on the underlying constants
-of~$\Filter$, boiling them down to choices of branches, which depend on one
-single variable, and loop transitions. This logic corresponds to the placement
-of counters for optimally edge-profiling the CFG of a program, making up a
-minimum-spanning tree of the automaton~\cite{path-profiling}.
-
-In order to make the examples below tractable, we will assume the existence of a
-rule \EquationReasoning{} that allows us to perform standard algebraic reasoning
-on linear inequalities. This rule is not necessary for correctness or
-completeness, but shortens the examples considerably.
-
-\subsection{An Example}
-
-Starting with~$\AcaOrBc{}$, where $\Map$ is the length function, in effect
-$\Transition \mapsto 1$ for transitions, and the constraints that the length is
-odd using the same trick as in \cref{sec:motivation}, we have the
-definitions in \cref{fig:example:single:equivalences} (omitting existential
-quantifiers and $x \geq 0$ for every variable to avoid clutter).
-
-\begin{figure}[ht]
-  We define $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$ as in the following two
-  equations, and then apply \EquationReasoning{} (under the implicit assumption
-  that every RHS is $\geq 0$) to obtain the third definition:
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = \TransitionVar_1 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_2 \\
-        \Filter(\FromLabelTo{A}{c}{A}) & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = \TransitionVar_5 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \hspace{0.5cm}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        % S
-        \TransitionVar_4 &= \TransitionVar_5 + \TransitionVar_2 - 1 \\
-        % A
-        \TransitionVar_1 &= \TransitionVar_5 \\
-        % B
-        \TransitionVar_2 &= \TransitionVar_4 + \TransitionVar_6 \\
-        % F
-        \TransitionVar_5 + \TransitionVar_6 &= \FinalStateVar_1 \\
-        \FinalStateVar_1 &= 1 \\
-      \end{aligned}  
-    \end{equation*}    
-  \end{minipage}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \caption{Equivalences defining $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$
-  respectively.}\label{fig:example:single:equivalences}
-  \end{figure}
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo[]{
-{    \begin{array}{lll}
-      1 - \TransitionVar_6 = 0 & \land \TransitionVar_6 = 1 & \land  \TransitionVar_4 + 1 > 0 \\
-      \land \TransitionVar_3 = 0 &\land  2 + \TransitionVar_4 > 0 & \land k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-{    \begin{array}{lll}
-      \Connected(\AcaOrBc{}, \Filter) & \land 1 - \TransitionVar_6 = 0 \\
-      \land \TransitionVar_6 = 1 & \land \TransitionVar_4 + 1 > 0 \land \\
-      \land \TransitionVar_3 = 0 & \land 2 + \TransitionVar_4 > 0 \land \\
-      k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Propagate]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 \geq 0 \land \\
-      2 + \TransitionVar_4 \geq 0 \land \\
-      2k = 2 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{aligned}
-  }
-  % BRANCH:  = 0'
-  \hypo{
-    {\begin{array}{lll}
-      \TransitionVar_3 > 0 & \land 
-      \TransitionVar_6 = 0 \\ \land
-      2\TransitionVar_6 + \TransitionVar_4 = 0 & \land
-      1 - \TransitionVar_6 > 0\\  \land 
-      2k = 1 + \TransitionVar_3
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_3 > = 0 \land \\
-      \TransitionVar_6 = 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0\\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\Propagate{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{aligned} % BRANCH: > 0
-  } % SPLIT
-  \infer2[\Split{} $1 - \TransitionVar_6$]{ \Connected(\AcaOrBc{}, \Filter) \land 2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6 }
-  \infer1[\EquationReasoning{}]{
-    \Connected(\AcaOrBc{}, \Filter) \land
-    2k + 1 =
-    \sum\limits_{\Transition \in \Transitions} (\Transition \mapsto 1)(\Label) \cdot \Filter(\Transition)
-  }
-  \infer1[\Expand{}]{\Image{}_{\AcaOrBc{}, \Transition \mapsto 1}(\Filter, 2k + 1)}
-\end{prooftree}
-\caption{A derivation for \Calculus{} computing odd lengths in $\AcaOrBc{}$.}\label{fig:derivation:single}
-\end{figure}
-
-In \cref{fig:derivation:single}, we see how we start by expanding the predicate
-using the (simplified) flow equations. We can also see how interleaving
-reasoning on the corresponding linear equations helps the deduction. In both
-branches the reasoning is similar: we conclude that when choosing either path at
-the starting state we must avoid the other one, perform propagation based on
-that fact, use some algebraic reasoning to derive the expected forms of the
-upper and lower bounds to propagate the now disconnected transitions, and
-finally subsume, removing the $\Connected{}$ predicate when we are unable to use
-either of the rules, leaving only a set of linear equations. To obtain a solution from the
-proof, one would need to perform standard model generation on the remaining
-constraints to obtain values for the sought variable $\MonoidElement$. The full
-Parikh image is the disjunction of the constraints left at the two
-leaves after quantifier elimination.
-
-Following the values of $\Filter$, we can read out the path corresponding to our solution in the product, and with that a word.
-In our case, for the right branch we have $\Filter(\FromLabelTo{S}{b}{B}) = 0,
-\Filter(\FromLabelTo{A}{c}{A})  > 0,
-\Filter(\FromLabelTo{B}{b}{S}) = 0,
-\Filter(\FromLabelTo{S}{a}{A}) = 1 ,
-\Filter(\FromLabelTo{A}{a}{F}) = 1 ,
-\Filter(\FromLabelTo{B}{c}{F}) = 0$.
-We can read out a word along it: aca, accca, etc, and the corresponding constraint: $\exists k : \TransitionVar_3 = 2k -1 \land \TransitionVar_3 > 0 \land k \geq 0$.
-
-\subsection{Correctness of \Calculus{}}\label{sec:single:correct}
-
-Our correctness proof of \Calculus{} consists of two main parts: we
-first show that the construction of a proof always terminates, and
-then that each of the proof rules in \cref{tbl:rules:single} is
-an equivalence transformation, i.e., does not change the set of
-satisfying assignments of a formula. In combination, those two results
-immediately imply that \Calculus{} gives rise to a decision procedure.
-
-\subsubsection{\Calculus{} terminates}
-\begin{lemma}\label{lma:single-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the predicates
-  $\Image$ and $\Connected$ only occur positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:single}.
-\end{lemma}
-
-\begin{proof}
-  The rule~\Expand{} can only be applied finitely often, since each
-  application removes one $\Image$ predicate, and none of the rules
-  introduce new instances of the predicate. The rule~\Subsume{} can
-  only be applied finitely often, since it strictly decreases the
-  combined number of $\Image$ and $\Connected$ predicates in sets of
-  formulas, and none of the rules increases that number.
-
-  To show termination of \Split{} and \Propagate{}, observe that the
-  $\Filter$ in a predicate~$\Connected(\Automaton, \Filter)$ is never
-  updated on a proof branch, which means that the set of terms
-  $\Filter(t)$ for $t \in \Automaton$ on every branch is finite. Each
-  application of \Split{} and \Propagate{} adds a new
-  formula~$\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$
-  to a proof goal, which can only happen finitely often.
-\end{proof}
-
-\subsubsection{The rules in \cref{tbl:rules:single} are solution-preserving}
-
-\begin{lemma}\label{lma:single-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:single}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
-  conclusion~$\SomeClause$ if and only if it satisfies one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  This property has to be shown by analysing the possible applications
-  of each proof rule.
-
-  \Expand{} unfolds the definition of the $\Image$ predicate. To show
-  that the rule is solution-preserving we prove the equivalence of the
-  upper and lower sets of formulas:
-  \begin{itemize}
-  \item Assume that $\beta$ satisfies the conclusion, which means that
-    there is some accepting path
-    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ with
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$ and $\val_\beta(\MonoidElement) =
-    \Map(\WordOf(\Path))$. Since immediately implies that $\beta$
-    satisfies $\Connected(\Automaton, \Filter)$, since a path is
-    connected, and $\FlowEq(\Automaton, \Filter)$ since an accepting
-    path satisfies the flow equations. The
-    equation~$\MonoidElement = \sum_{\Transition \in
-      \Transitions_\Automaton}\Filter(\Transition) \cdot
-    \Map(\Transition)$ holds because of
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$.
-  \item Assume that $\beta$ satisfies the premise, which implies that
-    $\val_\beta(\Filter)$ describes a consistent, connected flow of
-    the automaton. By the same argument as in
-    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
-    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
-    that each transition~$\Transition$ occurs on $\Path$ exactly
-    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
-  \end{itemize}
-
-  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
-  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
-  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
-  be satisfied, implying the property.
-
-  For \Propagate{}, suppose that
-  $\SeparatingCut(C, \Automaton, \State)$, which means that every
-  accepting path visiting~$q$ contains at least one of the transitions
-  in $C$. For a $\beta$ satisfying $\Connected(\Automaton, \Filter)$
-  and $\Filter(\Transition') = 0$ for $\Transition' \in C$, this means
-  that also $\val_\beta(\Filter(\Transition)) = 0$ has to hold.
-
-  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
-  then a goal must contain~$\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
-  formulas in $\SomeInequalities$ are inconsistent, an application of
-  \Subsume{} is trivially solution-preserving; therefore assume that
-  $\SomeInequalities$ is consistent, which means that it contains
-  exactly one of $\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for each $\Transition$. Since
-  \Propagate{} is not applicable, the transitions~$\Transition$ with
-  $\Filter(\Transition) > 0$ must form a connect sub-graph of the
-  automaton; this means that $\Connected(\Automaton, \Filter)$ is
-  redundant as it is implied by $\SomeInequalities$.
-\end{proof}
-
-\section{Parikh Images from Products of Automata}\label{sec:multiple}
-
-We now generalise our calculus to natively work with intersections of
-regular languages, or equivalently products of automata. For this
-extension, we change the main predicate~$\Image$ to be indexed by a
-vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
-simplicity, we assume that the sets of states of the $k$ automata (and
-therefore also the transition sets) are pairwise disjoint. 
-\begin{definition}\label{def:multiple}
-  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
-  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
-  $\Monoid$,
-  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
-  is a transition selection function, and
-  $\MonoidElement \in \Monoid$. The predicate
-  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
-      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
-  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
-  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
-  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
-  \begin{itemize}
-  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
-    $\Filter$, that is,
-    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
-  \item the automata all accept the same
-    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
-  \item the accepted word is mapped to
-    $\MonoidElement = \Map(\WordOf(\Path_i)) =
-    \Map(\WordOf(\Path_1))$.
-  \end{itemize}
-\end{definition}
-
-  \begin{table}[t]
-    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-      Name & \text{Rule} & Side conditions\\
-      \midrule
-
-      % EXPAND
-      \ExpandM & 
-      \inferrule
-      {
-       {\begin{array}{c}
-           \left\{
-           \FlowEq(\Automaton_i, \Filter),~~
-           \Connected(\Automaton_i, \Filter),~~
-           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
-           \right\}_{i=1}^k,
-           \\[4ex]
-          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
-          \SomeInequalities, \SomeClause
-        \end{array}}
-        }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
-      None \\[5ex]
-      \Materialise &
-      \inferrule
-      {{
-         \begin{array}{c}
-           \FlowEq(\Automaton', \Filter'),~
-           \Connected(\Automaton', \Filter'),~
-           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
-           \\\ImagePredicate{
-                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
-                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
-                     \Automaton'},~ \SomeInequalities, \SomeClause
-           }{\Map}{\Filter'}{\MonoidElement}
-           \end{array}}
-           }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
-      $\begin{array}{@{}l@{}}
-         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
-         1 \leq i < j \leq k, \\
-      \Automaton' = \Automaton_i \times \Automaton_j,\\
-      \end{array}$ \\
-      \bottomrule
-      \end{tabular}
-      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
-    \end{table}
-
-    For the calculus (\cref{tbl:rules:multi}), we first extend
-    $\Expand$ to generate flow equations and instances of $\Connected$
-    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
-    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
-    since it is needed to keep track of the currently considered
-    partial products.
-
-The rule~$\Materialise$ introduces the product of two individual
-automata~$\Automaton_i, \Automaton_j$; this step eliminates
-$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
-instead adds the product~$\Automaton_i \times \Automaton_j$, while
-also introducing the flow equations and the $\Connected$ predicate.
-
-The rule~$\Materialise$ also has to connect the newly introduced
-product~$\Automaton_i \times \Automaton_j$ to the previous automata
-$\Automaton_i, \Automaton_j$. This is done by extending the
-selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
-to fresh variables~$x_{\Transition}$:
-\begin{equation*}
-  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
-  ~~=~~
-  \begin{cases}
-    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
-    \\
-    \Filter(t) & \text{otherwise}
-  \end{cases}
-\end{equation*}
-
-The multiplicity of transitions in the product then has to be related
-to the multiplicities in the individual automata, modelled using the
-$\BindingSum$ predicate. The predicate expresses that the multiplicity
-of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
-$\Automaton_i$ has to coincide with the sum of the multiplicities of
-transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
-and similarly for $\Automaton_j$:
-%
-  $$
-  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
-  \begin{aligned}
-  & \Set{ 
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
-  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
-  & \Set{
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
-  }
-  \end{aligned}
-$$
-
-\iffalse
-Note that this definition implies that $\Filter(\Transition) =
-\Filter(\Transition')$ whenever two transitions $\Transition \in
-\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
-produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
-\times \Automaton_2}$. This corresponds to our intuition that the terms of the
-product must agree on the value they accept. As before, we implicitly map
-$\Filter$ to fresh terms for each transition in the product.
-\fi
-
-For instances of precisely one automaton, neither rule applies and we
-perform the calculus as before.
-
-\subsection{An Example}\label{sec:multiple:example}
-
-We return again to our example in \cref{sec:introduction:parikh}, where we
-compute the whole Parikh image of the product of $\AcaOrBc{}$ and
-$\SomethingCSomething$ under the constraint that there are more instances of
-letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
-element-wise addition, and $\Map$ that maps each transition to the corresponding
-increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.
-
-In the interest of space, we refer back to
-\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
-$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
-substitutions as follows. Note the expansion of the $\Sigma$ labels:
-    \begin{equation*}
-      \begin{aligned}
-        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-        % Other automaton
-        \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
-        \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
-      \end{aligned}
-    \end{equation*}
-
-In the derivation tree of \cref{fig:derivation:multi} we start as before, but
-with the product version of the $\Image$ predicate. The only possible rule here
-is $\ExpandM$, which we use to add the corresponding constraints on each
-automata of the product as we would have had in the single-automaton version. We
-perform algebraic reasoning on those equations, along with the constraint on
-$\MonoidElement$ to determine bounds on transition variables that will enable us
-to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
-finished doing so, we use \Materialise{} to compute the product of the now
-filtered automata. Further algebraic reasoning allows us to remove the final
-instance of $\Connected$. We then perform \Expand{} to get rid of the final
-$\Image$ instance, and are immediately able to $\Subsume$ the resulting
-$\Connected{}$ predicate. This, again, leaves us with a set of linear
-inequalities at the leaf that we can use to obtain a model of our final
-$\MonoidElement$ values (now a vector). This time we can read out aca, or
-$\VectorLiteral{2, 0, 1}$.
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo{
-    {\begin{array}{lll}
-      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
-      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
-    \end{array}}  
-  }
-  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
-    {\begin{array}{ll}
-      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
-      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
-    \end{array}}  
-  }
-  \infer1[\EquationReasoning{}]{
-  {  \begin{array}{ll}
-      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
-      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
-        \TransitionVar_6 + \TransitionVar_4 = 0 \\
-        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
-        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
-        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
-          \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1} 
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} \\ \land 
-      \Connected(\SomethingCSomething{}, \Filter) &
-      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array} } 
-  }
-  \infer1[\Materialise]{
-    {\begin{array}{ll}
-      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
-      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
-      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
-      \land 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} & \land
-      \Connected(\SomethingCSomething{}, \Filter) \\
-      \land 
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-  {\begin{array}{ll}
-    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-    1 - \TransitionVar_6 > 0 &
-    \land
-    \TransitionVar_6 = 0 \land 
-    \TransitionVar_4 = 0 \\
-    \land
-    \TransitionVar_3 > 0 & \land
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
-    \land
-    \VectorLiteral{
-      \TransitionVar_{7a} + \TransitionVar_{9a},
-      0,
-      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-      = \VectorLiteral{
-        2,
-        0,
-        \TransitionVar_3} & \land 
-    \Connected(\AcaOrBc{}, \Filter) \\
-    \land 
-    \Connected(\SomethingCSomething{}, \Filter) \\
-    \land 
-    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-  \end{array}}
-  }
-  \infer1[\EquationReasoning]{
-    \begin{aligned}
-      \VectorLiteral{a, b, c} = \VectorLiteral{
-          2 - 2\TransitionVar_6,
-          2\TransitionVar_4 + 3\TransitionVar_6,
-          \TransitionVar_3 + \TransitionVar_6
-        } \land
-        \\
-        \VectorLiteral{a, b, c} = \VectorLiteral{
-            \TransitionVar_{7a} + \TransitionVar_{9a},
-            \TransitionVar_{7b} + \TransitionVar_{9b},
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          } \land \\
-      \Connected(\AcaOrBc{}, \Filter) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{a, b, c}) \land a > b
-    \end{aligned}
-  }
-  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
-\end{prooftree}
-\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
-\end{figure}
-
-\subsection{Correctness of \Calculus{} for Products of Automata}
-
-Since \cref{tbl:rules:multi} only extends the existing rules of
-\cref{tbl:rules:single}, we focus on the differences compared
-to the calculus for a single automaton.
-
-%\subsubsection{\Calculus{} for products of automata terminates}
-\begin{lemma}\label{lma:multi-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the product
-  version of $\Image$ only occurs positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
-\end{lemma}
-
-\begin{proof}
-  The rule \Materialise{} can similarly only be used finitely many times, as
-  each application reduces the number of automata in the product of $\Image$ by
-  one automaton, until only one remains and \cref{lma:single-terminates} for
-  single-automaton instances applies.
-
-  This implies that also the rule~\ExpandM{} can only be applied
-  finitely often since its side condition only allows applying it once
-  per instance of an $\Image$ predicate containing a product, and only finitely
-  many instances of $\Image$ can be introduced on each branch.
-\end{proof}
-
-%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
-
-Since our calculus now includes a rule introducing new variables, the
-\Materialise{} rule, we have to slightly generalise the notion of
-solution-preservation:
-%
-\begin{lemma}\label{lma:multi-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:multi}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
-  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
-  there is an extension~$\beta' of \beta$ satisfying one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  We have to consider the two new rules in \cref{tbl:rules:multi}. The
-  result is immediate for \ExpandM{}, since this rule does not remove the
-  $\Image$ predicate from a proof goal, and the newly introduced formulas
-  are all implied by the $\Image$ predicate.
-
-  For \Materialise{}, observe that the existence of an accepting path
-  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
-  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
-  same word. The path in the product will satisfy the flow equations
-  and connectedness, and it will be related to the individual paths as
-  stipulated by the \BindingSum{} predicate.
-\end{proof}
-
-\section{Extensions}\label{sec:extensions}
-
-This section describes a number of possible extensions to the calculus. Note
-that some are partially or fully implemented in \Catra{} already. In particular,
-we have some level of support for symbolic transitions over Unicode alphabets to
-keep practical automata under a reasonable size, though we do not allow a full Boolean algebra over symbols as described e.g. in~\cite{symbolic-automata}.
-
-\subsection{Backjumping and Learning No-Goods}\label{sec:ext:backjumping}
-
-\Calculus{} can be accelerated for some instances by adding rules for
-backjumping. In particular, central connectedness constraints for an automaton
-can be learnt. In that case, when discovering that a state $\State$ of an
-automaton $\Automaton$ under a certain transition variable $\Filter$ has become
-unreachable, we can learn the clause $\Connected(\Automaton, \Filter) \land
-\sum_{\Transition \in \SeparatingCut(\Automaton, \State)} \Filter(\Transition) =
-0 \implies \sum_{\Transition' = \FromLabelTo{\State}{}{} \in
-\Transitions_{\Automaton}} = 0$. A similar rule can be used when a state becomes
-backwards-unreachable (but with a corresponding backwards cut). At the moment,
-only forward-cut learning is implemented in \Catra, and has provided a slight
-improvement in performance on some instances.
-
-The other source of clauses to learn is the \Materialise{} rule. Whenever an
-attempt to materialise a product of two ($\Filter$-filtered) automata
-$\Automaton_1 \times \Automaton_2$ produces an empty product, we can determine
-the cause of the failure with respect to the automata and their respective
-transition variables, and learn no-good combinations that can never be part of a
-model. In order to be able to do this, we need a few semantic predicates to
-record the state of the calculation, as well as a system of disambiguating
-automata, since it is possible to arrive at the same automaton by multiple
-combinations of decisions and products. In \Catra{}, we use a number of
-additional predicates to record the status of the materialisation of products,
-to separate instances of our main predicate, and to register the mapping between
-automata and their respective transition $\Filter$ terms, described in
-\cref{sec:implementation}. However, at this point only rudimentary no-good
-learning is implemented.
-
-\subsection{Symbolic Automata}\label{sec:ext:symbolic}
-
-Extending \Calculus{} to support fully symbolic automata is possible within the
-framework, depending on your choice of $\Map$. The difficulty consists in
-handling the mapping of the homomorphism $\Map$ over symbolic labels, assuming
-it maps to finitely many monoid elements. This is not always straightforward, as
-seen in the example of \cref{sec:multiple:example}. This complexity is inherent
-in computing the full Parikh image, and stems from the fact that we need to
-differentiate between the possible interpretations of the $\Sigma$ transitions
-without knowing ahead of times which ones will actually be materialied in the
-product. If, on the other hand, if our $\Map$ had been length-counting, which
-does not differentiate between values, it would have required no adaptation at
-all. With a somewhat liberal interpretation of what we are allowed to map to
-(e.g. fresh terms), it is possible to use $\Map$ to represent choice operations
-like the ones in a range label. In \Catra{}, we frontload this problem by
-requiring the user to encode their input as a Parikh automaton. For an example of how automata can be encoded, see \cref{sec:parikh-automata}.
-
-\subsection{Transducers}\label{sec:ext:transducers}
-
-Another interesting application is applying \Calculus{} to automata with
-multiple tracks, e.g. transducers. One application of such a calculus would be
-to represent replace operations and other functions on regular languages, and to
-be able to answer questions like "does this operation change the length of the
-string". The difficulty in implementing it for transducers is first to perform
-the mapping on the labels, which for both the length and Parikh cases is
-straightforward; just do the same thing in two dimensions. The more complicated
-operation is defining the product of transducers. In some cases it would
-probably be desirable to perform synchronisation (e.g. requiring overlapping
-transitions) on only some transitions, for example the first track.
-
-Implementing such a calculus is straightforward in \Calculus, since the
-definition of products was left out of the definition. All that would be needed
-is an appropriate update of the definitions of products. Simiarly, \Catra was
-written with modularity in mind, and it should be straightforward to extend both
-the input grammar and automata implementations to accommodate multi-track
-automata with arbitrary synchronisation.
-
-\subsection{Finding the Presburger representation of a homomorphic image}\label{sec:finding-the-image}
-
-To find the Presburger form of the homomorphic image efficiently, we adapt the
-quantifier elimination approach of~\cite{qe} to our problem domain. The core
-method is the same: we incrementally use \Calculus{} to find models,
-\Generalise{} the models we find into a quantifier-free Presburger formula with
-the $\MonoidElement$ as the only free variable (algorithm~\ref{alg:generalise}), add the
-negated formula as a constraint, and continue enumerating models until none
-remain. The disjunction of the generalised models we enumerated is now our
-image.
-
-\begin{algorithm}
-  \caption{$\Generalise{}(\Automaton, \Filter, a)$ will generalise a final product $\Automaton$ and a model $a$ under the homomorphism $\Map$.}\label{alg:generalise}
-  \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-  \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-  \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-
-  $\Automaton' \gets \Tuple{\Automaton_\States, \Automaton_{\InitialState}, \Automaton_\AcceptingStates, \Set{\Transition \in \Automaton_\Transitions \SuchThat a(\Filter(\Transition)) \neq 0}}$
-
-  \KwRet{\EliminateQuantifiers{$\Map(\ParikhMap(\Automaton'))$}}
-  \end{algorithm}
-
-  \begin{algorithm}
-    \DontPrintSemicolon
-    \caption{$\FindImage{}(\Automaton_1 \times \ldots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
-    \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-    \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-    \SetKwFunction{NewTheoremProver}{newTheoremProver}
-    \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-    \SetKwFunction{FreshVariable}{freshVariable}
-    \SetKwFunction{Assert}{assert}
-    \SetKwFunction{GetModel}{getModel}
-    \SetKwData{ImageVar}{image}
-
-$p \gets \NewTheoremProver{}$\;
-$\Filter(\Transition) := \FreshVariable{p}$ for every $\Transition \in \Transitions_{\Automaton}$\;
-$\MonoidElement \gets$ \FreshVariable{$p$}\;
-\Assert{$p, \exists \MonoidElement, \Filter(\Transition) \text{ for every } \Transition \in \Transitions_\Automaton \HoldsThat \ImagePredicate{\Automaton}{\Map}{\Filter}{\MonoidElement} \land \AndComp{\Transition \in \Transitions}{\Filter(\Transition) \geq 0}$}\;
-$\ImageVar \gets \bot$\;
-\While{$p$ has more models}{
-  $\Tuple{\Automaton, a} \gets \GetModel(p)$\;
-  $G \gets \Generalise{}(\Automaton, \Filter, a)$\;
-  $\ImageVar \gets G \lor \Image$\;
-  \Assert{$p, \lnot G$}\;
-  }
-    \KwRet{\ImageVar}
-    \end{algorithm}
-
-    The \GetModel{} function is nonstandard in that it returns both the model
-    and its associated automaton. Since algorithm~\ref{alg:find-image} is
-    essentially a form of quantifier elimination and therefore an internal affair
-    to the theorem prover, this should be considered fair game.
-
-    A preliminary implementation of this approach is available as part
-    of~\Catra, but has been neither optimised nor fully tested.
-
-\section{Implementation}\label{sec:implementation}
-
-We implement \Calculus{} for Parikh automata as described in
-\cref{sec:parikh-automata}. The artefact submitted along with this paper is a
-program that reads an instance file with one or more products of one or more
-Parikh automaton with transition labels defined as ranges of Unicode characters,
-along with a set of constraints on the final values of their registers expressed
-as Presburger arithmetic in a C-like syntax. We call this program
-\Catra.
-% \footnote{If you really must read it as an acronym, please read it as
-%CAtegory Theory on Register Automata, or if you object to the somewhat
-%nonstandard use of register automata and category theory, as Check Assignments
-%of The Registers Afterwards. Or alternatively if you find it all to be too much
-%of a theoretical exercise, as Can Anyone Think of A Real Application.}
-
-\Catra{} is written in Scala, with the calculus described in this paper
-implemented as a theory plug-in for the \Princess{} automated theorem
-prover~\cite{princess}, which also performs the Presburger reasoning. For
-comparison, we also provide an implementation of the baseline method
-from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
-symbolic model checker~\cite{nuxmv} to solve our constraints, and the
-approximation described in~\cite{approximate-parikh} on top of the standard
-baseline back-end. An example of an input file corresponding to our running
-example introduced in \cref{sec:motivation} can be seen in
-\cref{lst:input-example}.
-
-\Catra{} uses symbolic labels for automata. A symbolic label is defined as a
-finite range of Unicode code points. This allows succinct representation of many
-regular expression patterns such as \lstinline{(a-z).*} which would have
-otherwise required $27$~transitions. The transition for the range would be
-written \lstinline{a -> b [97, 122]}.
-
-In satisfaction mode, supported by all backends, \Catra{} tries to satisfy the
-constraints expressed by the input file, reporting \Sat{} with register
-assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
-Additionally, baseline and \Calculus{} also support generation of the Presburger
-formula describing the constraints of the input file. Baseline uses standard
-quantifier elimination, and \Calculus{} uses the method described in
-\cref{sec:finding-the-image}.
-
-Since \Princess{} does not support multiple-arity predicates like the ones we
-use in \Calculus{}, we have implemented variable-length arguments using
-additional helper predicates. These are $\Unused{}(\Automaton)$, which marks an
-automaton as unused in any product, and $\TransitionMask{}(\Automaton,
-\Transition, \Filter(\Transition))$ which associates a transition $\Transition$
-and automaton $\Automaton$ with its corresponding transition variable.
-Additionally, we associate each of our predicates with an instance variable in
-order to differentiate instances of the predicates.
-
-\subsection{Implementing the Baseline}\label{sec:implementing-baseline}
-
-We baseline using the same Presburger solver (\Princess{}), input file parser,
-and automaton implementation as \Catra. We do this in order to better analyse
-the impact of the calculus rules themselves. Using the formula of
-\eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
-successive term and add them to \Princess. We compute the product incrementally
-term by term, checking satisfiability at each step. We use a priority queue to
-select automata for each step, and order it by the number of transitions as a
-heuristic for the size of the automata. We use this heuristic to avoid computing
-large (and therefore slow) products until we have to, banking our hopes on
-computing an empty intermittent product early. The pseudocode for our
-implementation can be seen in \cref{alg:baseline}.
-
-\begin{algorithm}
-  \caption{How we implement the baseline approach}\label{alg:baseline}
-  \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
-  \KwResult{\textsc{Sat} or \textsc{Unsat}}
-  \SetKwFunction{NewTheoremProver}{newTheoremProver}
-  \SetKwFunction{NewPriorityQueue}{newPriorityQueue}
-  \SetKwFunction{Dequeue}{dequeue}
-  \SetKwFunction{Enqueue}{enqueue}
-  \SetKwFunction{Assert}{assert}
-
-  $p \gets \NewTheoremProver{}$
-
-  \Assert{$p$, $\SomeClause$}
-
-  \ForEach{$\Automaton_i$}{
-    \Assert{$p, \ParikhMap(\Automaton_i)$}
-
-    \If{$p$ is \textsc{Unsat}}{break}
-
-  }
-
-  $q \gets \NewPriorityQueue{}$
-
-  \While{$p$ not \textsc{Unsat} and $|q| > 1$}{
-    $\Automaton, \Automaton' \gets \Dequeue{q}$
-
-    \Assert{$p, \ParikhMap(\Automaton \times \Automaton')$}
-
-    \Enqueue{$q, \Automaton \times \Automaton'$}
-  }
-
-  \KwRet{$p$'s SAT status}
-
-  \end{algorithm}
-
-As an optimisation, our automata (including intermittent products) are created
-forward- and backward- reachable-minimal. Any automaton we produce only contains
-states that are both reachable from the initial state and has a path to an
-accepting state. We never perform any other minimisation on the automata for
-either backend. More complex minimisation was left out since performing
-minimisation on automata with counters is non-trivial, and minimising symbolic
-automata risks exponential blowup \cite{minimising-symbolic}.
-
-\subsection{Heuristics and search strategies}
-
-There are a number of choices left unspecified in \Calculus{} as described in
-\cref{sec:calculus,sec:multiple}. For example, the order of materialisation of
-intermediate products and the order of splitting. In this section we describe
-additional implementation details and techniques used to enhance \Catra.
-
-\subsubsection{Splitting, Materialisation, and Propagation}
-
-We order our rule applications so that we first propagate connectedness if
-possible, then perform materialisation if tractable as defined below, then
-finally resort to splitting if we must.
-
-In addition to applying \Split{} as described in \cref{tbl:rules:single} to
-randomly selected transitions, we prefer splitting to sever a strongly connected
-from the initial state. We randomly select an automaton where we can compute a
-cut between an SCC and the initial state, that is where the SCC does not contain
-the initial state and where the sum of the transition variables of the
-transitions in the cut is not known to be positive. If there are multiple such
-strongly connected components we choose one randomly. We then proceed to split
-on the sum of the transition variables of the cut as if it were a regular
-transition, e.g. its sum being zero or nonzero. In this way we drive \Calculus{}
-towards applying \Propagate{}.
-
-The implementation of the connectedness constraint is opportunistic and straightforward.
-We compute a set of dead states by performing forward and backwards reachability computations on an automaton, where we disregard any transition whose associated variable is known to be zero.
-After that we add clauses ensuring any transition variable associated with a transition starting in a dead state is zero.
-
-Product materialisation is the final piece of the puzzle. In the current
-implementation we put off computing intermediate products until at least all but
-two transition variables of one of the automata is known to be either present or
-absent. The number was chosen experimentally, and we observe a consistent trend
-towards lower numbers being better. The other automaton for the product is
-selected randomly.
-
-\subsubsection{Clause Learning}
-
-\Catra{} enables clause learning by default when using our backend, as it has
-been experimentally shown to increase the performance in aggregate (though not
-strictly). We do not currently implement all the proposed features of
-\cref{sec:ext:backjumping}, but we do implement forward-reachability cut
-learning. No sophisticated clause learning for products has been implemented.
-
-\subsubsection{Random Restarts}
-
-Finally, we perform restarts scaled by the Luby series~\cite{luby}. Experimental
-results have shown this to have a large improvement in performance, which is
-unsurprising given how many random choices we make during solving and how
-tail-heavy our problem is.
-
-\section{Evaluation}\label{sec:experiments}
-
-We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
-automata intersection problems generated by the \OstrichPlus{} string constraint
-solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
-initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
-five seconds by baseline.
-We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial.
-For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s.
-The benchmarks are run on commit~\texttt{\commit}.
-
-The benchmarks are executed in parallel using GNU Parallel~\cite{parallel},
-since they are mostly single-threaded and initial results showed negligible
-interference on performance, on \BenchmarkRig{}. We compiled the code using
-Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a
-maximum heap of~\MaxHeapSize{}. We used \Nuxmv{} version~\NuxmvVersion{} invoked
-as a subprocess for each instance. Instances were executed in batches of
-\BatchSize{}, each given a fresh JVM. We believe this represents a realistic use
-case where \Calculus{} is used to support e.g a string solver. Experiments were
-executed in random order for all backends. Each instance got a time budget
-of~\RuntimeTimeout.
-
-All runtimes are measured in wall-clock time as observed by the JVM when
-executing the instance, and exclude time spent parsing (usually far below
-\numprint{0.1}s).
-
-\subsection{Execution Time and Ability to Solve Instances}\label{sec:runtime}
-
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
-the respective back-ends could solve and with which status. A full summary of
-their outcomes is also available in \cref{tab:solve-status}. Note that we lack
-ground truth for all instances. We see that \Calculus{} generally outperforms
-\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
-instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
-kind of instance, but most of all on satisfiable instances, and \Calculus{}
-outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
-\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
-caused by a non-deterministically triggered bug in the clause learning of
-\Calculus{}, but have not had time to further investigate it. We have however
-validated all \numprint{3908} satisfying assignments from \Calculus{} with
-\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
-our implementation.
-
-Baseline performs worse on satisfiable instances because it executes a heuristic
-meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances. The heuristic is enabled since it is unrealistic to assume that we
-know in advance whether an instance is satisfiable or not, but we do know that
-unsatisfiable instances are more common.
-
-\begin{table}
-  \centering
-  \input{graphs/solved_pivot_table.tex}
-  \caption{The result of running the respective back-ends with a timeout of
-  \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
-  from the table. }\label{tab:solve-status}
-\end{table}
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
-    \caption{The division of statuses per backend.}
-    \label{fig:solve-division}
-    \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
-Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
-    \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
-    \label{fig:runtime-boxplot}
-    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline.
-The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy.
-On the other hand, lazy has  a lot of outliers.}
-  \end{subfigure}
-  \vfill
-  \begin{subfigure}[b]{0.75\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
-    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
-    \label{fig:cactus}
-    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
-Lazy solves much more instances than nuxmv.}
-  \end{subfigure}
-\end{figure}
-
-\subsubsection{Scalability}\label{sec:scaling}
-
-A cactus plot showing the number of instances solved within a given timeout for
-each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
-\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
-long-term scaling properties.
-
-\iffalse
-\subsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
-
-For baseline and \Calculus{}, \Catra{} offers the ability to find the equivalent Presburger formula representing a given instance.
-For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} we use the specially tailored approach described in \cref{sec:finding-the-image}.
-For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in \cref{sec:scaling,sec:runtime}.
-
-To make sure baseline puts up as much competition as possible, we disable
-checking intermittent satisfiability and configure \Catra{} to run in the
-maximally eager mode where the product is first computed before any
-satisifiability check is performed. We run the experiments with a timeout
-of~\ImageTimeout{}. The results of the experiment is summarised in
-\cref{fig:cactus:image} and \cref{tab:image-results}. \Fudge{We see here that
-something happens}.
-
-\begin{figure}[ht]
-  \caption{The number of instances \Catra{} was able to find the Presburger form of the image for within a given number of seconds per backend.}
-  \label{fig:cactus:image}
-\end{figure}
-\fi
-
-\subsection{Threats to Validity}
-
-The most obvious threats to validity would be poor benchmarking or poor
-implementation, e.g. if the method described in \cref{sec:calculus,sec:multiple}
-deviate from what is actually benchmarked in section \cref{sec:experiments}, or
-if the methods used for benchmarking would be unsound.  To increase the
-probability that our results are representative both in the sense of
-representing  on expected inputs and in the statistical sense for a given run of
-\Catra{} despite the use of randomness in our implementation we execute many
-experiments. Moreover, to address the issue of correctness we have validated all
-reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
-is indeed sound.
-
-Another threat to validity would be if the competition (baseline and \Nuxmv)
-would be disadvantaged in our comparison. For \Nuxmv{} we use the default
-configuration which we believe should be performant (or it should not be the
-default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
-easy for artefact reviewers. For baseline, our best argument is that the
-solution we use is close to the one found in the \Ostrich{} string solver and
-that it therefore should be realistic.
-
-The greatest threat to validity is our choice of implementation platform and
-automata library. There are some signs that product computation is inefficient,
-notably the good performance of low-threshold automata materialisation that
-prioritises computing smaller products. This means that \Calculus{} makes less
-heavy use of product computation than baseline does, instead relying more on
-\Princess{}'. This situation would unfairly advantage \Calculus{} if our
-automata library was the bottleneck due to our automata library being
-unnaturally poor. We believe this is unlikely since similar performance issues
-have been reported for string solvers. Additionally, profiling suggests that
-both \Princess{}-based backends spend most of their time in \Princess{},
-suggesting that the automaton implementation is not the bottleneck.
-
-\section{Conclusion}
-
-In this paper we have introduced a calculus to compute commutating operations on
-intersections of regular languages that we call \Calculus{}. We have evaluated
-it on \NrBenchmarks{} Parikh automata intersection problems generated by the
-\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
-suite \cite{pyex} using our Parikh automata solver \Catra{}.
-
-Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
-usage and solve-time compared to the baseline approach laid out in
-\cite{generate-parikh-image} when implemented on the same underlying automated
-theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
-\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
-and generally outperforming it for timeouts under 30 seconds with its advantage
-increasing drastically for even shorter timeouts. 30 seconds would generally be
-considered a long timeout for our intended use as supporting infrastructure to a
-string constraint solver.
-
-Future investigations involve two tracks. The first one is integration into
-existing string solvers (wich \Ostrich{} being a particularly promising
-candidate due to its shared use of \Princess{}), and further adaptation to that
-use case. Closer inspection of the instances where we currently time out should
-be useful to further improve our heuristics.
-
-The second track for future improvements is the extension into other problem
-domains, including other logics, model checking problems, as well as to more
-powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata, due to our use of a full
-automated theorem prover which allows adding arbitrary constraints in addition
-to the expected Presburger formulae.
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-%% Bibliography
-\bibliography{bibliography}
-
-%% Appendix
-\appendix
-\section{Appendix}
 \subsection{Per-instance comparison between backends}
@@
 
-\subsection{SMT-LIB encoding of the string constraint example}
-\lstinputlisting[caption={An SMT-LIB encoding of \cref{ex:string-constraints}.}, label=lst:string-constraints]{example.smt2}
-
-\end{document}
+%\subsection{SMT-LIB encoding of the string constraint example}
+%\lstinputlisting[caption={An SMT-LIB encoding of \cref{ex:string-constraints}.}, label=lst:string-constraints]{example.smt2}

### 2023-06-06 example.tex
Split out the example into its own file

--- 2023-06-06 introduction.tex
@@
-The Parikh image is a characterisation of formal languages in terms of their
-character counts. Given a language over an alphabet~$\{a_1, \ldots, a_k\}$, the
-Parikh image is a set of $k$-dimensional vectors that contains some
-vector~$\VectorLiteral{m_1, m_2, \ldots, m_k}$ if and only if the formal
-language contains a word in which each $a_i$ occurs $m_i$ times. It is a
-classical result that the Parikh image of every context-free language (and,
-thus, also of every regular language) is a semilinear set~\cite{parikh-theorem},
-i.e., Presburger-definable.
-
-Parikh images play a central role in many automata-based algorithms, for
-instance and notably in today's string solvers, which often have to process
-constraints that combine regular language membership with word length. To decide
-whether a simple formula like $x \in \Language_1 \wedge y \in \Language_2 \wedge
-|x| > |y|$, with string variables~$x, y$ and regular languages~$\Language_1,
-\Language_2$, is satisfiable, it is necessary to reason about the sets of word
-lengths induced by $\Language_1, \Language_2$, which are special cases of the
-Parikh image.  This required combined reasoning about strings and string length
-has long been identified as a major bottleneck in string solvers
-\cite{DBLP:conf/cav/AbdullaACHRRS15,length-aware-solver,approximate-parikh,DBLP:journals/corr/BerzishZG17}.
-Other string solvers make use of Parikh automata~\cite{parikh-automata}, and
-thus Parikh images in the general case, to handle operations including
-\verb!str.substr!  and \verb!str.at!, which comes at an even higher price in
-terms of computational complexity~\cite{ostrich-plus}.
-
-It is possible to compute an existential Presburger formula describing the
-Parikh image of any context-free language in linear
-time~\cite{generate-parikh-image}; for the special case of regular languages,
-this result was also stated in \cite{muscholl-linear}. While theoretically
-elegant, this construction has several disadvantages, often making it
-unpractical for integration into algorithms. Firstly, the constructed Presburger
-formula contains a linear number of existential quantifiers in the size of the
-considered grammar, as well as complex Boolean structure, which is needed to
-express the connectedness of sets of productions considered in the construction.
-Eliminating those quantifiers to obtain a quantifier-free representation of the
-Parikh image has exponential complexity~\cite{DBLP:conf/issac/Weispfenning97},
-and is often impossible in reasonable time. Just solving the Parikh image
-membership problem is NP-complete, as it corresponds to computing a satisfying
-assignment of the existential Presburger formula, and taxing for solvers as
-well~\cite{ostrich-plus}.
-
-Secondly, in applications involving regular languages, it is typically necessary
-to consider the Parikh image not only of a single automaton, but of the
-intersection of multiple automata. This problem arises in string solvers in
-particular, as conjunctions of string constraints lead to the computation of
-length images of products (intersections) of regular languages represented as
-finite automata. Applying the approach in~\cite{generate-parikh-image} would in
-this case require the eager computation of the product before its length image,
-and result in an existential Presburger formula of exponential size (in the
-number of automata). In several instances we have observed while solving
-real-world string constraints, the computation of the product of automata
-exhausts the memory of any machine available due to the exponential blow-up in
-size of the product, quickly becoming intractable as the number of automata in
-the product increases. The current best published mitigation for this problem is
-an over-approximation that works by approximating the Parikh image of a product
-of automata to be the conjunction of the image of the individual automata of the
-product \cite{approximate-parikh}. This approach only works for unsatisfiable
-instances, and comes with a harsh penalty for satisfiable instances.
-
-Addressing these concerns, we have developed a calculus for Parikh images of
-products of regular languages that we call \Calculus{}. It allows us to
-interleave the computation of arbitrarily deep products of automata with the
-product's Parikh image, and is generalised to an arbitrary homomorphism over
-automata labels, including string lengths. This enables us to let both
-calculations inform each other, eliminating unnecessary work, and pruning the
-size of the partial products considered in the computation for a smaller memory
-footprint. Moreover, the method can be used iteratively to tackle smaller chunks
-of the product incrementally, thereby decreasing the memory footprint.
-
-The key ideas of \Calculus{} is a combination of problem-aware case splitting,
-lazy enforcement of automata connectivity, and lazy computation of products. In
-essence, we use constraint-based methods to efficiently enforce the constraints
-associated with computing Parikh images of regular languages.
-
-We implement \Calculus{} as a plug-in theory for the \Princess{} automated
-theorem prover. \Fudge{Others} have integrated it into the similarly
-\Princess{}-backed string constraint solver \OstrichPlus{}, where it is used to
-handle constraints that use cost-enriched automata, as described in
-\cite{ostrich-plus}.
-
-We additionally wrap the Parikh image solver as a stand-alone tool, \Catra.
-\Catra{} also supports a variant of the approximate method of
-\cite{approximate-parikh}, its fall-back variant adapted
-from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
-checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
-back-ends on \NrBenchmarks{} distinct Parikh automata intersection problems
-generated by \OstrichPlus{} when solving the PyEx string constraint benchmark
-suite involving string length constraints \cite{pyex}.
-
-In summary, we contribute:
-\begin{itemize}
-    \item The \Calculus{} calculus to efficiently compute (a homomorphism on)
-          the Parikh image of products of Parikh automata.
-    \item Experiments illustrating the performance of \Calculus{} on real-world
-    examples from string solving, including \NrBenchmarks{} instances in a
-    standardised format made available for future study.
-    \item The \Catra{} tool for solving such instances, containing an
-    implementation of \Calculus{}, the over-approximation described
-    in~\cite{approximate-parikh}, and an adapter for the~\Nuxmv{} model
-    checker~\cite{nuxmv}.
-    \item Suggestions for how to efficiently implement \Calculus{} in a modern
-    automated theorem prover, including strategies for case splitting, clause
-    learning, and constraint propagation for connectedness.
-\end{itemize}
-
-\subsection{Related Work}
-
-The problem of computing constraints on Parikh images over products of regular
-languages under a given commutative homomorphism amounts to solving products of
-Parikh automata. Parikh automata are regular automata extended with integer
-counters with given increments and decrements for each transition, where we
-allow checking a set of linear constraints on the final values of the counters
-(but not their intermittent values) \cite{parikh-automata}. Parikh automata
-without constraints on the final values on their registers are also sometimes
-called cost-enriched automata, weighted automata or counter automata, depending
-on exact definitions and side-constraints. The decision problem tackled in this
-paper, determining the emptiness of an intersection of Parikh automata, was
-recently shown to be PSPACE-complete~\cite{graph-queries}.
-
-Parikh image computations as well as Parikh automata feature extensively in
-string solvers, including as mentioned above \Ostrich{} and \OstrichPlus{}
-\cite{ostrich,ostrich-plus}, but also forms the basis of Trau~\cite{trau-pldi},
-and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images frequently appear when
-introducing cardinality constraints like length or string indexing. The
-state-of-the-art approach to handling Parikh image computation is to
-over-approximate the Parikh image of a product of $k$~automata
-$\ParikhMap(\Language(\Automaton_1) \cap \ldots \cap \Language(\Automaton_k))$
-with the conjunction of the automata's parikh maps, $\bigwedge_{i=1}^{k}
-\ParikhMap(\Automaton_i)$. This approach works only for unsatisfiable instances,
-and will require falling back to computing the product of the automata before
-using the standard approach for finding its image originally presented
-in~\cite{generate-parikh-image}.
-
-Outside of string solvers, Parikh automata have been proposed as the basis of
-queries~\cite{graph-queries}, and for solving cardinalities in model checking
-problems involving epistemic logic~\cite{epistemic-logic}.
-
-Many other generalisations of the Parikh image than the projections we use here
-have been studied. Prominent examples include generalising the Parikh map to
-segments of a fixed length \cite{KARHUMAKI1980155} and the more general Parikh
-matrix, which gives more information about a word than the standard Parikh image
-\cite{parikh-matrix}. Another notable generalisation is the p-vector, introduced
-in~\cite{infinite-words}, which denotes the position of each letter in the word
-rather than the number of their occurrences and allows for generalisations into
-infinite alphabets. All of these in some sense extend the Parikh map. By
-contrast, the main utility of the formulation introduced here is that it allows
-us to \emph{remove} something, thereby potentialy obtaining an easier problem.
-
-\subsection{An Intuition for Our Approach}\label{sec:intuition}
-
 \begin{figure}[h]
@@
     \includegraphics[scale=\autscale]{a}
-    \caption{The automaton $A'$, whose Parikh registers count both the number of
-    letters and the start ($l$) and length ($r$) of the substring matching
+    \caption{The automaton $A'$, whose Parikh registers count the length of the
+    string accepted by the automaton (register $l_A$), the start offset of the
+    substring ($i$) and length ($n$) of the substring matching
     $\Regex{c\RegexOr{}dd}$.}\label{fig:aut_a}

### 2023-06-06 introduction.tex
Split out the example into its own file

@@
 us to \emph{remove} something, thereby potentialy obtaining an easier problem.
-
-\subsection{An Intuition for Our Approach}\label{sec:intuition}
-
-\begin{figure}[h]
-    \centering 
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{a}
-    \caption{The automaton $A'$, whose Parikh registers count both the number of
-    letters and the start ($l$) and length ($r$) of the substring matching
-    $\Regex{c\RegexOr{}dd}$.}\label{fig:aut_a}
-      %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{b}
-    \caption{The automaton $B$, where the Parikh registers count the number of
-    occurrences of each letter.}\label{fig:aut_b}
-      %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \caption{The collection of automata we use as running
-  examples}\label{fig:examples}
-\end{figure}
-
-In this section we will introduce an intuition for how a string constraint
-problem in an automata-based solver like \OstrichPlus{} is translated to a
-Parikh automata intersection problem and solved efficiently using \Calculus{}.
-First, by a Parikh automaton we mean a standard non-deterministic finite
-automaton (NFA) with a set of Parikh registers that are incremented (or
-decremented) at each transition. In the automata of \cref{fig:examples}, we will
-use the notation $a / \begin{bmatrix} x_+ \end{bmatrix}$ to mean that a
-transition reads an input character $a$ and increments register $x$ by one. We
-will omit any zero-valued increments and assume that all registers are scoped to
-their automaton. The increments are usually represented as a vector (hence the
-notation), but as it is mostly sparse here, we instead use the symbolic notation
-$\begin{bmatrix} x_{1+} \end{bmatrix}$ rather than the more cumbersome $\left[
-0, 1,  0 , \ldots, 0 \right]$.
-
-Furthermore, we use $\CountOf{c}(A)$ to refer to the number of letters $c$ in
-$\Language(A)$, and allow this expression to be used both in linear constraints
-of the input language of \cref{ex:string-constraints} and in the langauge of
-constraints that we translate to. Where the automaton $A$ is apparent, we will
-omit it.
-
-We use standard PCRE regular expressions here and throughout the paper, writing
-them $\Regex{\texttt{like this}}$. This means that $\RegexOr{}$ is alternation,
-$\mathtt{*}$ the Kleene star, and $\mathtt{.}$, matches any single character.
-
-\begin{example}\label{ex:string-constraints} Consider the following set of
-    string constraints:
-\begin{constraints}
-    \item\label{const:x-in-c-dd} $x \in \Language(\Regex{c\RegexOr{}dd})$.
-    \item\label{const:y-in-b} $y \in \Language(B)$, the language accepted by
-    automaton~$B$ of \cref{fig:aut_b}.
-    \item\label{const:x-substring} $x = \Substring(y, i, n)$, that is $x$ is an
-    $n$-length substring of $y$ starting at offset $i$.
-    \item\label{const:more-a-than-b} $i = n$, that is the substring is as long as its starting offset.
-\end{constraints}
-\end{example}
-
-Intuitively, we see that the constraints are satisfiable.
-\cref{const:more-a-than-b} implies that the starting offset must be either $1$
-or $2$, that is we need to fill out $y$ with either one or two symbols before
-the substring $x$ starts. The state $D$ is dead, as $B$ contains no transitions
-with label $\Char{d}$, and hence none of the transitions into $D$ can appear in
-the intersection of the languages. An example satisfying assignment is $y =
-\Char{acca}$, which gives $x = \Char{c}$, $i = 1$, $n = 1$, but others are
-possible.
-
-To solve \cref{ex:string-constraints} using \Calculus{},
-\cref{const:x-in-c-dd,const:y-in-b,const:x-substring} can be translated to the
-product of Parikh automata $B \times A'$, where $A'$ (\cref{fig:aut_a}) is the
-Parikh automata obtained by applying the encoding of $\Substring$ described in
-\cite{ostrich-plus}. That description can then be used to obtain values for the
-subsstring offsets $i, n$. Here follows an intuition for our method on $B \times
-A'$ under the global constraint on the substring \cref{const:more-a-than-b}.
-
-Our approach to the computation of the Parikh image (and therefore the solution
-to the intersection problem posed by \cref{ex:string-constraints}) consists of
-two applications of laziness; late computation of products of automata, and lazy
-enforcement of connectivity constraints. Both are in contrast to the eager
-approach to finding the Parikh image described in~\cite{generate-parikh-image},
-that would have started by computing the product $A' \times B$, translated it to
-a Presburger formula with $i, n$ etc as free variables, and then plugged in the
-constraint \cref{const:more-a-than-b}.
-
-Starting with an interactive theorem prover and the goal of solving the
-constraint $\CountOf{\Char{a}}(B \times A') > \CountOf{\Char{b}}(B \times A')$
-we associate each transition of $A'$ and $B$ with a fresh integer variable,
-$\TransitionVar_\Transition$, identically to \cite{generate-parikh-image}. These
-variables represent how many times each transition is taken. Hence, the final
-counter value for each automaton is the element-wise sum:
-\begin{equation}
-\begin{bmatrix} 
-  \CountOf{\Char{a}} \\ \CountOf{\Char{b}} 
-  \\ \CountOf{\Char{c}} \\ \CountOf{\Char{d}} 
-\end{bmatrix} = \sum_\Transition \TransitionVar_\Transition \cdot 
-  \IncrementVec_\Transition \text{ where $\IncrementVec_\Transition$ are the increments of transition $\Transition$}.
-\end{equation}
-
-\begin{figure}[h]
-    \centering 
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{a_annotated}
-    \caption{ $A'$ with its associated transition variables in symbolic form.
-    Note: unused ($0$) self-loops at $I_{A'}$ and $F_{A'}$ have been pruned for
-    readability.}\label{fig:aut_a_annotated}
-      %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{b_annotated}
-    \caption{$B$ with its associated transition variables in symbolic forms.
-    Note that state $B$ is now completely dead.}\label{fig:aut_b_annotated}
-      %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \caption{The starting automata, without their counter increments and with
-  their transition variables in symbolic form.}\label{fig:propagated}
-\end{figure}
-
-We proceed by adding linear constraints requiring transition variables to
-represent a flow through either automata by requiring that the number of
-incoming transitions is equal to the number of outgoing ones. E.g. state $B$
-would have the sum $\cancel{x_{\Tuple{B, \Char{c}, B}}} + x_{\Tuple{I_B,
-\Char{b}, B}} = \cancel{x_{\Tuple{B, \Char{c}, B}}}  + x_{\Tuple{B, \Char{c},
-F_B}} + x_{\Tuple{B, \Char{b}, I_B}}$ where we let $x_{\Tuple{q, l, c'}}$ refer
-to the integer variable associated with a transition from state $q$ to state
-$q'$ with label $l$. Note that the self-loop cancels itself out. This means that
-for automata where a loop can become unreachable, these constraints do not fully
-capture the constraints posed by the automaton, but requires additional
-reachability constraints.
-
-Performing symbolic reasoning on these equations and plugging the corresponding
-symbolic representation of the constraints on each corresponding integer
-variable back as an annotation on its transition, we obtain the two automata in
-\cref{fig:propagated}.
-
-% FIXME B-loopen inte nollad här! Uppdatera!
-
-Several transitions are already marked as unuseable (have a transition variable
-value of $0$). This includes any transition for the letter $\Char{d}$, as
-$\CountOf{\Char{d}}(B) = \sum_{\Transition} 0 = 0$, since the register for
-$\Char{d}$ is zero for all transitions in $B$.
-
-Note that the self-loop on state $B$ is nonzero due to the above mentioned
-self-balancing. However, a reachability computation will conclude that the state
-is unreachable, and allow us to conclude that all its outgoing transitions must
-be zero.
-
-After removing all transitions concluded to be unuseable to produce new,
-equivalent automata, we compute the product in \cref{fig:product}.
-
-\begin{figure}[h]
-    \centering 
-    \includegraphics[scale=\autscale]{ab}
-  \caption{An automaton encoding the constraint $\CountOf{\Char{a}}(B \times A')
-  > \CountOf{\Char{b}}(B \times A')$.}\label{fig:product}
-      %\Description[SHORT]{LONG}
-\end{figure}
-
-By putting off computing the product $B \times A'$ until after performing linear
-reasoning on the number of times each transition must be taken and using that to
-prune the terms, we have computed a smaller product than we would have with a
-na\"ive approach. Moreover, since this automaton contains only one loop along
-one main path, all runs through it can be described by the flow-preservation
-equations described above without the need for costly connectivity-enforcing
-constraints as in~\cite{generate-parikh-image}. We can now obtain a generalised
-solution to our constraints by inspection.

### 2023-06-06 main.tex
Split out the example into its own file

--- 2023-06-05 main.tex
@@
 
-\section{Preliminaries}
+\section{An Intuition for Our Approach}\label{sec:intuition}
+\input{example.tex}
+
+\section{Preliminaries}\label{sec:preliminaries}
 
@@
 \section{Appendix}
-\subsection{Per-instance comparison between backends}
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter.pdf}
-    \caption{Per-instance runtime between baseline and \Calculus{}}
-    \label{fig:duels:baseline}
-    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and baseline]{The plot shows most dots at the bottom of the plot, with some plots collected at the right-hand side.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter.pdf}
-    \caption{Per-instance runtime between baseline and \Calculus{}}
-    \label{fig:duels:nuxmv}
-    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and \Nuxmv{}]{The plot shows flames of dots between the bottom and top, with a line of dots near the right-hand side corresponding to timeouts.}
-  \end{subfigure}
-\end{figure}
-\subsection{Example instance for \Catra{}}
-\begin{lstlisting}[caption={An example input file for \Catra{} for \cref{ex:string-constraints} in \cref{sec:intuition}, illustrating every major syntax element.
-From beginning to end: synchronised (product) automata using the keyword \texttt{synchronised} (automata A and B), labels (except those with ranges), register increments, and constraints on the final values of their counters.}, label=lst:input-example]
-\input{example.par}
-\end{lstlisting}
-
-% Text of appendix \ldots
-
+\input{appendix.tex}
 \end{document}

### 2023-06-06 introduction.tex
Cite Symbolic Boolean Derivatives as suggested
@pruemmer Please verify that I am not lying here!

Closes: the paper should cite a recent paper in the similar spirit: Symbolic Boolean Derivatives PLDI 21 #24

@@
 
+% @Philipp: please verify that this is not a lie!
+The problem and our calculus, \Calculus{}, takes a similar approach to the work
+of \citeauthor{symbolic-boolean-derivatives}, who present an orthogonal calculus
+for handling a similar blow-up in complexity resulting from Boolean combinations
+of regular automata in SMT string solvers \cite{symbolic-boolean-derivatives}.
+In addition to presenting a decision procedure for lazily dispatching
+constraints, we similarly also allow for symbolic labelling of automata to
+handle large Unicode alphabets.
+
 Outside of string solvers, Parikh automata have been proposed as the basis of

### 2023-06-12 example.tex
Use a new, unsatisfiable example
This version uses a new example which is hopefully easier to follow. The linear algebra is slightly breakneck, but there is little I can do about that.

This section is now, hopefully, less jargon-heavy!

Closes: #7, #31, #26, #30,  #29, #28.

@@
     substring ($i$) and length ($n$) of the substring matching
-    $\Regex{c\RegexOr{}dd}$.}\label{fig:aut_a}
+    $\Regex{c\RegexOr{}dd}$. Note the symbolic transitions in the starting and
+    accepting states matching any symbol in the alphabet!}\label{fig:aut_a}
       %\Description[SHORT]{LONG}
@@
     \includegraphics[scale=\autscale]{b}
-    \caption{The automaton $B$, where the Parikh registers count the number of
-    occurrences of each letter.}\label{fig:aut_b}
+    \caption{The automaton $B$, where the Parikh registers count the length of
+    the string in register $l_B$.}\label{fig:aut_b}
       %\Description[SHORT]{LONG}
@@
 problem in an automata-based solver like \OstrichPlus{} is translated to a
-Parikh automata intersection problem and solved efficiently using \Calculus{}.
-First, by a Parikh automaton we mean a standard non-deterministic finite
-automaton (NFA) with a set of Parikh registers that are incremented (or
+Parikh automata intersection problem and solved using \Calculus{}. First, by a
+Parikh automaton we mean a standard non-deterministic finite automaton (NFA)
+with a set of integer registers (Parikh registers) that are incremented (or
 decremented) at each transition. In the automata of \cref{fig:examples}, we will
@@
 transition reads an input character $a$ and increments register $x$ by one. We
-will omit any zero-valued increments and assume that all registers are scoped to
+will omit zero-valued increments and assume that all registers are scoped to
 their automaton. The increments are usually represented as a vector (hence the
-notation), but as it is mostly sparse here, we instead use the symbolic notation
+brackets), but as it is mostly sparse here, we use the symbolic notation
 $\begin{bmatrix} x_{1+} \end{bmatrix}$ rather than the more cumbersome $\left[
@@
 
-Furthermore, we use $\CountOf{c}(A)$ to refer to the number of letters $c$ in
-$\Language(A)$, and allow this expression to be used both in linear constraints
-of the input language of \cref{ex:string-constraints} and in the langauge of
-constraints that we translate to. Where the automaton $A$ is apparent, we will
-omit it.
+Note that the labels of transitions can be ranges of characters with $\Sigma$
+representing any character in the alphabet.
 
-We use standard PCRE regular expressions here and throughout the paper, writing
+We use PCRE regular expression notation here and throughout the paper, writing
 them $\Regex{\texttt{like this}}$. This means that $\RegexOr{}$ is alternation,
 $\mathtt{*}$ the Kleene star, and $\mathtt{.}$, matches any single character.
+
+For the length of a string $s$, we write $\Length{s}$.
 
@@
 \begin{constraints}
-    \item\label{const:x-in-c-dd} $x \in \Language(\Regex{c\RegexOr{}dd})$.
-    \item\label{const:y-in-b} $y \in \Language(B)$, the language accepted by
+    \item\label{const:s1-in-c-dd} $s_1 \in \Language(\Regex{c\RegexOr{}dd})$.
+    \item\label{const:s2-in-b} $s_2 \in \Language(B)$, the language accepted by
     automaton~$B$ of \cref{fig:aut_b}.
-    \item\label{const:x-substring} $x = \Substring(y, i, n)$, that is $x$ is an
-    $n$-length substring of $y$ starting at offset $i$.
-    \item\label{const:more-a-than-b} $i = n$, that is the substring is as long as its starting offset.
+    \item\label{const:s1-substring} $s_1 = \Substring(s_2, i, n)$, that is $s_1$ is an
+    $n$-length substring of $s_2$ starting at offset $i$.
+    \item\label{const:more-inside-than-before} $n > i$, that is what comes
+    before $s_1$ in $s_2$ is shorter than $s_1$.
+    \item\label{const:something-before-and-after} $i > 0 \land \Length{s_2} -i -n > 0$, that
+    is there is at least one character in $s_2$ after $s_1$.
 \end{constraints}
@@
 
-Intuitively, we see that the constraints are satisfiable.
-\cref{const:more-a-than-b} implies that the starting offset must be either $1$
-or $2$, that is we need to fill out $y$ with either one or two symbols before
-the substring $x$ starts. The state $D$ is dead, as $B$ contains no transitions
-with label $\Char{d}$, and hence none of the transitions into $D$ can appear in
-the intersection of the languages. An example satisfying assignment is $y =
-\Char{acca}$, which gives $x = \Char{c}$, $i = 1$, $n = 1$, but others are
-possible.
+To solve \cref{ex:string-constraints} using \Calculus{} in a theorem proving
+context, \cref{const:s1-in-c-dd,const:s2-in-b,const:s1-substring} can be
+translated to the product of Parikh automata $B \times A'$, where $A'$
+(\cref{fig:aut_a}) is the Parikh automata obtained by applying the encoding of
+$\Substring$ described in \cite{ostrich-plus}. That description can then be used
+to obtain values for the subsstring offsets $i, n$.
 
-To solve \cref{ex:string-constraints} using \Calculus{},
-\cref{const:x-in-c-dd,const:y-in-b,const:x-substring} can be translated to the
-product of Parikh automata $B \times A'$, where $A'$ (\cref{fig:aut_a}) is the
-Parikh automata obtained by applying the encoding of $\Substring$ described in
-\cite{ostrich-plus}. That description can then be used to obtain values for the
-subsstring offsets $i, n$. Here follows an intuition for our method on $B \times
-A'$ under the global constraint on the substring \cref{const:more-a-than-b}.
+Intuitively, \cref{ex:string-constraints} are unsatisfiable. By inspecting the
+automata, we realise that the path using $\Char{dd}$ in $A'$ is unusable since
+are no $\Char{d}$-labelled transitions in $B$. This means that $s_1 = \Char{c}$,
+and thus $n = 1$. That means that \cref{const:more-inside-than-before} implies
+that $i = 0$. However, no path through $B$ passes by a $\Char{c}$ without a
+preceding character. Therefore, already
+\cref{const:s1-in-c-dd,const:s1-in-c-dd,const:s1-substring,const:more-inside-than-before}
+together are unsatisfiable.
 
@@
 that would have started by computing the product $A' \times B$, translated it to
-a Presburger formula with $i, n$ etc as free variables, and then plugged in the
-constraint \cref{const:more-a-than-b}.
+a Presburger formula with $i, n$ etc as free variables, and then plugged in
+\cref{const:more-inside-than-before,const:something-before-and-after}.
 
-Starting with an interactive theorem prover and the goal of solving the
-constraint $\CountOf{\Char{a}}(B \times A') > \CountOf{\Char{b}}(B \times A')$
-we associate each transition of $A'$ and $B$ with a fresh integer variable,
+Starting with an interactive theorem prover and the goal of solving the Parikh
+automata intersection problem, we associate each transition $\Transition$ of
+$A'$ and $B$ with a fresh nonnegative integer variable from,
 $\TransitionVar_\Transition$, identically to \cite{generate-parikh-image}. These
 variables represent how many times each transition is taken. Hence, the final
-counter value for each automaton is the element-wise sum:
-\begin{equation}
+counter value for counters $r_1, \ldots, r_k$ of each automaton is the
+element-wise sum:
+\begin{equation}\label{eq:counter-sums}
 \begin{bmatrix} 
-  \CountOf{\Char{a}} \\ \CountOf{\Char{b}} 
-  \\ \CountOf{\Char{c}} \\ \CountOf{\Char{d}} 
+  r_1 \\
+  \vdots\\
+  r_k \\
 \end{bmatrix} = \sum_\Transition \TransitionVar_\Transition \cdot 
@@
 We proceed by adding linear constraints requiring transition variables to
-represent a flow through either automata by requiring that the number of
-incoming transitions is equal to the number of outgoing ones. E.g. state $B$
-would have the sum $\cancel{x_{\Tuple{B, \Char{c}, B}}} + x_{\Tuple{I_B,
-\Char{b}, B}} = \cancel{x_{\Tuple{B, \Char{c}, B}}}  + x_{\Tuple{B, \Char{c},
-F_B}} + x_{\Tuple{B, \Char{b}, I_B}}$ where we let $x_{\Tuple{q, l, c'}}$ refer
-to the integer variable associated with a transition from state $q$ to state
-$q'$ with label $l$. Note that the self-loop cancels itself out. This means that
-for automata where a loop can become unreachable, these constraints do not fully
-capture the constraints posed by the automaton, but requires additional
-reachability constraints.
+represent a flow through their automaton by adding linear constraints stating
+that the number of incoming transitions is equal to the number of outgoing ones.
+E.g. state $B$ would have the sum $\cancel{x_{\Tuple{B, \Char{c}, B}}} +
+x_{\Tuple{I_B, \Char{b}, B}} = \cancel{x_{\Tuple{B, \Char{c}, B}}}  +
+x_{\Tuple{B, \Char{c}, F_B}} + x_{\Tuple{B, \Char{b}, I_B}}$ where we let
+$x_{\Tuple{q, l, c'}}$ refer to the integer variable associated with a
+transition from state $q$ to state $q'$ with label $l$. Note that the self-loop
+cancels itself out. This means that for automata where a loop can become
+unreachable, additional constraints would be required to ensure reachability.
+However, being lazy, we put off that work until later.
 
@@
 
-% FIXME B-loopen inte nollad här! Uppdatera!
+Several transitions are already marked as unuseable (have a transition variable
+value of $0$) or always present (have a transition variable with a constant
+value greater than zero) by just applying standard linear equation solving to
+the flow equations described above. Note that this includes
+\cref{const:more-inside-than-before,const:something-before-and-after}, which are
+related to the transition variables through the sums described in
+\cref{eq:counter-sums}. This means that we effectively can prune $A'$ of its
+upper path when continuing the proof, as it is incongruent with the flow
+equations and \cref{ex:string-constraints}. Note that the per-automata length
+counting registers have been assigned the same solver variable $l_A$. Since they
+have to be equal, either one can be used in both automata.
 
-Several transitions are already marked as unuseable (have a transition variable
-value of $0$). This includes any transition for the letter $\Char{d}$, as
-$\CountOf{\Char{d}}(B) = \sum_{\Transition} 0 = 0$, since the register for
-$\Char{d}$ is zero for all transitions in $B$.
+Notably, one such conclusion is that $i > 0$ from
+\cref{const:something-before-and-after}. From flow preservation we have that the
+$F_{A'}$ to $F_{A'}$ transition is represented by $l_A - n - i$.
+\footnote{Intuitively, the total length is made up of what is before, inside and
+after the substring.} This in turn through \cref{const:more-inside-than-before}
+forces $n \geq 2$, which in turn (through flow-preservation) forces a path from
+$I_{A'}$ to $F_{A'}$ through $D$, rather than the single-transition path from
+$I_{A'}$ to $F_{A'}$. This immediately implies that the problem is
+unsatisfiable, since we know this transition has no correspondance in automaton
+$B$.
 
-Note that the self-loop on state $B$ is nonzero due to the above mentioned
-self-balancing. However, a reachability computation will conclude that the state
-is unreachable, and allow us to conclude that all its outgoing transitions must
-be zero.
+However, \Calculus{} does not at this point, and we press on with concluding
+that none of the loops of $A'$ at this stage can become disconnected since
+they are at the initial and final states, both of which are trivially reachable
+for any assignments of the non-constant transition variables. This means that
+all valid counter assignments of $A'$ are described by \cref{eq:counter-sums}
+and the flow equations alone without the need for expensive connectivity
+constraints.
 
-After removing all transitions concluded to be unuseable to produce new,
-equivalent automata, we compute the product in \cref{fig:product}.
+After concluding this, we have a choice of two paths through automaton $B$; the
+upper through state $A$ or the lower through state $B$. We case split on a
+problem-aware basis by selecting a transition variable that would disconnect
+some strongly connected component of automaton $B$, in this case the transition
+form $I_B$ to state $B$. We divide it into the cases used ($> 0$) or unused ($ =
+0$) since those are the threshold values of the calculus, preferring the latter
+option on the first-fail principle.
 
-\begin{figure}[h]
-    \centering 
-    \includegraphics[scale=\autscale]{ab}
-  \caption{An automaton encoding the constraint $\CountOf{\Char{a}}(B \times A')
-  > \CountOf{\Char{b}}(B \times A')$.}\label{fig:product}
-      %\Description[SHORT]{LONG}
-\end{figure}
+At this point we have two small, stick-like automata after discounting all
+transitions with a corresponding transition variable that is zero, and can try
+to compute their product. Doing so, we will immediately notice that the
+$\Char{d}$-labelled transition of automaton $A'$ has no correspondence in $B$,
+leading to an empty product. We close the proof goal and backtrack with the
+opposite constraint, choosing the upper $I_B$ to $A$ transition this time, the
+only potentially productive choice we have left. However, proceding with
+calculation of the product from the almost identical terms will be similarly
+improductive, and we must conclude that the problem is unsatisfiable.
 
@@
 equations described above without the need for costly connectivity-enforcing
-constraints as in~\cite{generate-parikh-image}. We can now obtain a generalised
-solution to our constraints by inspection.
+constraints as in~\cite{generate-parikh-image}.

### 2023-06-12 introduction.tex
Use a new, unsatisfiable example
This version uses a new example which is hopefully easier to follow. The linear algebra is slightly breakneck, but there is little I can do about that.

This section is now, hopefully, less jargon-heavy!

Closes: #7, #31, #26, #30,  #29, #28.

@@
 $\ParikhMap(\Language(\Automaton_1) \cap \ldots \cap \Language(\Automaton_k))$
-with the conjunction of the automata's parikh maps, $\bigwedge_{i=1}^{k}
+with the conjunction of the automata's Parikh maps, $\bigwedge_{i=1}^{k}
 \ParikhMap(\Automaton_i)$. This approach works only for unsatisfiable instances,

### 2023-06-12 example.tex
More step by step example

@@
 
-\begin{figure}[h]
-    \centering 
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{a_annotated}
-    \caption{ $A'$ with its associated transition variables in symbolic form.
-    Note: unused ($0$) self-loops at $I_{A'}$ and $F_{A'}$ have been pruned for
-    readability.}\label{fig:aut_a_annotated}
-      %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \includegraphics[scale=\autscale]{b_annotated}
-    \caption{$B$ with its associated transition variables in symbolic forms.
-    Note that state $B$ is now completely dead.}\label{fig:aut_b_annotated}
-      %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \caption{The starting automata, without their counter increments and with
-  their transition variables in symbolic form.}\label{fig:propagated}
-\end{figure}
-
 We proceed by adding linear constraints requiring transition variables to
@@
 
-Performing symbolic reasoning on these equations and plugging the corresponding
-symbolic representation of the constraints on each corresponding integer
-variable back as an annotation on its transition, we obtain the two automata in
-\cref{fig:propagated}.
+\begin{figure}[h]
+  \centering 
+  \includegraphics[scale=\autscale]{a_1}
+  \caption{ $A'$ with its associated transition variables in symbolic form after
+  minimal elminiation of variables using linear algebra. Note the incoming flow
+  of $1$ in the starting state and its balancing of the outgoing transitions!
+  }\label{fig:a_1}
+\end{figure}
 
-Several transitions are already marked as unuseable (have a transition variable
-value of $0$) or always present (have a transition variable with a constant
-value greater than zero) by just applying standard linear equation solving to
-the flow equations described above. Note that this includes
-\cref{const:more-inside-than-before,const:something-before-and-after}, which are
-related to the transition variables through the sums described in
-\cref{eq:counter-sums}. This means that we effectively can prune $A'$ of its
-upper path when continuing the proof, as it is incongruent with the flow
-equations and \cref{ex:string-constraints}. Note that the per-automata length
-counting registers have been assigned the same solver variable $l_A$. Since they
-have to be equal, either one can be used in both automata.
+As an example, we start with reasoning about $A'$. Performing symbolic reasoning
+on these equations and plugging the corresponding symbolic representation of the
+constraints on each corresponding integer variable back as an annotation on its
+transition, we obtain the automaton in \cref{fig:a_1}. After slightly more
+linear algebra to remove variables through equality elimination, we have
+\cref{fig:a_2}, whose transitions are only described in terms of the free
+variables we care about representing the length of the automaton ($l_A$), the
+start of the substring ($i$), and the length of the substring ($n$).
 
-Notably, one such conclusion is that $i > 0$ from
-\cref{const:something-before-and-after}. From flow preservation we have that the
-$F_{A'}$ to $F_{A'}$ transition is represented by $l_A - n - i$.
-\footnote{Intuitively, the total length is made up of what is before, inside and
-after the substring.} This in turn through \cref{const:more-inside-than-before}
-forces $n \geq 2$, which in turn (through flow-preservation) forces a path from
-$I_{A'}$ to $F_{A'}$ through $D$, rather than the single-transition path from
-$I_{A'}$ to $F_{A'}$. This immediately implies that the problem is
+\begin{figure}[h]
+  \centering 
+  \includegraphics[scale=\autscale]{a_2}
+  \caption{ $A'$ after even more linear algebra, with most transitions expressed
+  in terms of $n$. Intuitively, this version captures the fact that $n$ is given
+  by distributing the incoming flow of $1$ across the two outgoing transitions
+  from $I_{A'}$, the initial state.}
+  \label{fig:a_2}
+\end{figure}
+
+Having obtained this representation, we conclude that $1 < n \leq 2$ from
+\cref{const:something-before-and-after,const:something-before-and-after}. The
+upper bound is obtained by the following reasoning. Since all transition
+variables are nonnegative integers (a transition cannot be used a negative
+amount of times), $n \leq 2$ from the $I_A$ to $F_A$ transition. Therefore, it
+follows that $n=2$, which implies that the $I_A$ to $F_A$ transition can never
+be used under these constraints and that we must use the two $\Char{d}$-labelled
+transitions, both now taken $n-1 =1$ times. This immediately implies that the problem is
 unsatisfiable, since we know this transition has no correspondance in automaton
@@
 
-However, \Calculus{} does not at this point, and we press on with concluding
-that none of the loops of $A'$ at this stage can become disconnected since
-they are at the initial and final states, both of which are trivially reachable
-for any assignments of the non-constant transition variables. This means that
-all valid counter assignments of $A'$ are described by \cref{eq:counter-sums}
-and the flow equations alone without the need for expensive connectivity
-constraints.
+We repeat similar methods to arrive at the automata in \cref{fig:propagated}.
+
+\begin{figure}[h]
+    \centering 
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+    \includegraphics[scale=\autscale]{a_annotated}
+    \caption{ $A'$ with its associated transition variables in symbolic
+    form.}\label{fig:aut_a_annotated}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+    \includegraphics[scale=\autscale]{b_annotated}
+    \caption{$B$ with its associated transition variables in symbolic forms.
+    Note the large number of implicitly existence-quantified bound variables on
+    transitions, suggesting that this automaton has a more complex structure
+    with relation to its (free) target variable representing the string length
+    than $A'$. The variables used here are fresh, and do not correspond to the
+    ones previously used.}\label{fig:aut_b_annotated}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}
+  \caption{The automata after applying linear algebra and equality elimination
+  to prune them.}\label{fig:propagated}
+\end{figure}
+
+Note that the per-automata length counting registers have been assigned the same
+solver variable $l_A$. Since they have to be equal, either one can be used in
+both automata through similar applications of equality elimination rules.
+
+At this point we know we have an unsatisfiable solution, \Calculus{} does not.
+We press on with concluding that none of the loops of $A'$ at this stage can
+become disconnected since they are at the initial and final states, both of
+which are trivially reachable for any assignments of the non-constant transition
+variables. This means that all valid counter assignments of $A'$ are described
+by \cref{eq:counter-sums} and the flow equations alone without the need for
+expensive connectivity constraints.
 

### 2023-06-16 example.tex
example udpate

@@
-\begin{figure}[h]
+In this section we will introduce an intuition for how a string constraint
+problem in an automata-based solver like \OstrichPlus{} is translated to a
+Parikh automata intersection problem and solved using \Calculus{}.
+
+We use PCRE regular expression notation here and throughout the paper, writing
+them $\Regex{\texttt{like this}}$. This means that $\RegexOr{}$ is alternation,
+$\mathtt{*}$ the Kleene star, and $\mathtt{.}$, matches any single character.
+For the length of a string $s$, we write $\Length{s}$.
+
+\begin{example}\label{ex:string-constraints} Consider the following set of
+    string constraints:
+\begin{constraints}
+    \item\label{const:s1-in-c-dd} $s_1 \in \Language(\Regex{c\RegexOr{}dd})$.
+    \item\label{const:s2-in-b} $s_2 \in \Language(B)$, the language accepted by
+    automaton~$B$ of \cref{fig:aut_b}.
+    \item\label{const:s1-substring} $s_1 = \Substring(s_2, i, n)$, that is $s_1$ is an
+    $n$-length substring of $s_2$ starting at offset $i$.
+    \item\label{const:more-inside-than-before} $n > i$, that is what comes
+    before $s_1$ in $s_2$ is shorter than $s_1$.
+    \item\label{const:something-before-and-after} $i > 0 \land \Length{s_2} -i -n > 0$, that
+    is there is at least one character in $s_2$ after $s_1$.
+\end{constraints}
+\end{example}
+
+To solve these constraints, we will first perform translation into Parikh
+automata, then break down how the constraints will be handled by \Calculus{}.
+
+\subsection{Translating the constraints into a Parikh automata intersection problem}
+
+By a Parikh automaton we mean a standard non-deterministic finite automaton
+(NFA) with a set of integer registers (Parikh registers) that are incremented
+(or decremented) at each transition.
+
+To solve \cref{ex:string-constraints} using \Calculus{} in a theorem proving
+context, \cref{const:s1-in-c-dd,const:s2-in-b,const:s1-substring} can be
+translated to the product of Parikh automata $B \times A'$, where $A'$
+(\cref{fig:aut_a}) is the Parikh automata obtained by applying the encoding of
+$\Substring$ described in \cite{ostrich-plus}. That description can then be used
+to obtain values for the subsstring offsets $i, n$. The corresponding automata
+can be found in \cref{fig:examples}.
+
+In the automata of \cref{fig:examples}, we will use the notation $a /
+\begin{bmatrix} x_+ \end{bmatrix}$ to mean that a transition reads an input
+character $a$ and increments register $x$ by one. We will omit zero-valued
+increments and assume that all registers are scoped to their automaton. The
+increments are usually represented as a vector (hence the brackets), but as it
+is mostly sparse here, we use the symbolic notation $\begin{bmatrix} x_{1+}
+\end{bmatrix}$ rather than the more cumbersome $\left[ 0, 1,  0 , \ldots, 0
+\right]$.
+
+Note that the labels of transitions can be ranges of characters with $\Sigma$
+representing any character in the alphabet.
+
+\begin{figure}[ht]
     \centering 
@@
   \caption{The collection of automata we use as running
-  examples}\label{fig:examples}
+  examples, both derived from \cref{ex:string-constraints}.}\label{fig:examples}
 \end{figure}
-
-In this section we will introduce an intuition for how a string constraint
-problem in an automata-based solver like \OstrichPlus{} is translated to a
-Parikh automata intersection problem and solved using \Calculus{}. First, by a
-Parikh automaton we mean a standard non-deterministic finite automaton (NFA)
-with a set of integer registers (Parikh registers) that are incremented (or
-decremented) at each transition. In the automata of \cref{fig:examples}, we will
-use the notation $a / \begin{bmatrix} x_+ \end{bmatrix}$ to mean that a
-transition reads an input character $a$ and increments register $x$ by one. We
-will omit zero-valued increments and assume that all registers are scoped to
-their automaton. The increments are usually represented as a vector (hence the
-brackets), but as it is mostly sparse here, we use the symbolic notation
-$\begin{bmatrix} x_{1+} \end{bmatrix}$ rather than the more cumbersome $\left[
-0, 1,  0 , \ldots, 0 \right]$.
-
-Note that the labels of transitions can be ranges of characters with $\Sigma$
-representing any character in the alphabet.
-
-We use PCRE regular expression notation here and throughout the paper, writing
-them $\Regex{\texttt{like this}}$. This means that $\RegexOr{}$ is alternation,
-$\mathtt{*}$ the Kleene star, and $\mathtt{.}$, matches any single character.
-
-For the length of a string $s$, we write $\Length{s}$.
-
-\begin{example}\label{ex:string-constraints} Consider the following set of
-    string constraints:
-\begin{constraints}
-    \item\label{const:s1-in-c-dd} $s_1 \in \Language(\Regex{c\RegexOr{}dd})$.
-    \item\label{const:s2-in-b} $s_2 \in \Language(B)$, the language accepted by
-    automaton~$B$ of \cref{fig:aut_b}.
-    \item\label{const:s1-substring} $s_1 = \Substring(s_2, i, n)$, that is $s_1$ is an
-    $n$-length substring of $s_2$ starting at offset $i$.
-    \item\label{const:more-inside-than-before} $n > i$, that is what comes
-    before $s_1$ in $s_2$ is shorter than $s_1$.
-    \item\label{const:something-before-and-after} $i > 0 \land \Length{s_2} -i -n > 0$, that
-    is there is at least one character in $s_2$ after $s_1$.
-\end{constraints}
-\end{example}
-
-To solve \cref{ex:string-constraints} using \Calculus{} in a theorem proving
-context, \cref{const:s1-in-c-dd,const:s2-in-b,const:s1-substring} can be
-translated to the product of Parikh automata $B \times A'$, where $A'$
-(\cref{fig:aut_a}) is the Parikh automata obtained by applying the encoding of
-$\Substring$ described in \cite{ostrich-plus}. That description can then be used
-to obtain values for the subsstring offsets $i, n$.
 
@@
 
-Our approach to the computation of the Parikh image (and therefore the solution
-to the intersection problem posed by \cref{ex:string-constraints}) consists of
-two applications of laziness; late computation of products of automata, and lazy
-enforcement of connectivity constraints. Both are in contrast to the eager
-approach to finding the Parikh image described in~\cite{generate-parikh-image},
-that would have started by computing the product $A' \times B$, translated it to
-a Presburger formula with $i, n$ etc as free variables, and then plugged in
-\cref{const:more-inside-than-before,const:something-before-and-after}.
+We will now proceed to show how we prove this using the three principles of
+\Calculus{}: problem-aware case splitting (sometimes called branching), lazy
+enforcement of connectivity constraints (essentially consisting of a constraint
+programmgin-style propagator), and lazy computation of products of automata.
 
-Starting with an interactive theorem prover and the goal of solving the Parikh
-automata intersection problem, we associate each transition $\Transition$ of
-$A'$ and $B$ with a fresh nonnegative integer variable from,
-$\TransitionVar_\Transition$, identically to \cite{generate-parikh-image}. These
-variables represent how many times each transition is taken. Hence, the final
-counter value for counters $r_1, \ldots, r_k$ of each automaton is the
-element-wise sum:
+Both lazinesses are in contrast to the eager approach to finding the Parikh
+image described in~\cite{generate-parikh-image}, that would have started by
+computing the product $A' \times B$, translated it to a Presburger formula with
+$i, n$ etc as free variables, and then plugged in
+\cref{const:more-inside-than-before,const:something-before-and-after} to the
+resulting set of linear inequalitites.
+
+\subsection{Solving the Parikh automata intersection problem using \Calculus{}}
+
+The approach consists of four steps: first we approximately describe each
+automaton as a transition-counting flow from the initial state to its accepting
+states where we relate its transitions to integer variables and register values
+to their transitions. Then we will use the linear algebra part of the theorem
+prover to discover that some paths through one of the automata are incompatible
+by \cref{const:more-inside-than-before,const:something-before-and-after}. This
+will allow us to discover an unreachable state, which allows further
+elimination. Finally, we will use case splitting to further subdivide the
+problem into bite-sized products of automata, both of which are empty, thus
+concluding that \cref{ex:string-constraints} is unsatisfiable.
+
+\subsubsection{Eliminating paths through automata using flow analysis}
+
+We associate each transition $\Transition$ of $A'$ and $B$ with a fresh
+nonnegative integer variable from, $\TransitionVar_\Transition$, identically to
+\cite{generate-parikh-image}. These variables represent how many times each
+transition is taken. Hence, the final counter value for counters $r_1, \ldots,
+r_k$ of each automaton is the element-wise sum:
 \begin{equation}\label{eq:counter-sums}
@@
 
+\subsubsection{Using linear algebra to eliminate flow}
 As an example, we start with reasoning about $A'$. Performing symbolic reasoning
@@
 
+\subsubsection{Propagation of connectivity constraints}
 At this point we know we have an unsatisfiable solution, \Calculus{} does not.
@@
 
+\subsubsection{Case Splitting}
 After concluding this, we have a choice of two paths through automaton $B$; the
@@
 
-At this point we have two small, stick-like automata after discounting all
+\subsubsection{COMPUTING PRODUCTS}
+Now we have two small, stick-like automata after discounting all
 transitions with a corresponding transition variable that is zero, and can try

### 2023-06-16 introduction.tex
Introduction is now less focused on Parikh Automata

@@
-The Parikh image is a characterisation of formal languages in terms of their
-character counts. Given a language over an alphabet~$\{a_1, \ldots, a_k\}$, the
-Parikh image is a set of $k$-dimensional vectors that contains some
-vector~$\VectorLiteral{m_1, m_2, \ldots, m_k}$ if and only if the formal
-language contains a word in which each $a_i$ occurs $m_i$ times. It is a
-classical result that the Parikh image of every context-free language (and,
-thus, also of every regular language) is a semilinear set~\cite{parikh-theorem},
-i.e., Presburger-definable.
+Extending automated theorem provers ("SMT solvers") with support for rich string
+constraints is important for program analysis, in particular to detect
+cross-site scripting vulnerabilities and other string manipulation bugs.
 
-Parikh images play a central role in many automata-based algorithms, for
-instance and notably in today's string solvers, which often have to process
-constraints that combine regular language membership with word length. To decide
-whether a simple formula like $x \in \Language_1 \wedge y \in \Language_2 \wedge
-|x| > |y|$, with string variables~$x, y$ and regular languages~$\Language_1,
-\Language_2$, is satisfiable, it is necessary to reason about the sets of word
-lengths induced by $\Language_1, \Language_2$, which are special cases of the
-Parikh image.  This required combined reasoning about strings and string length
-has long been identified as a major bottleneck in string solvers
+To decide whether a formula like $x \in \Language_1 \wedge y \in \Language_2
+\wedge |x| > |y|$, with string variables~$x, y$ and regular
+languages~$\Language_1, \Language_2$, is satisfiable, it is necessary to reason
+about the sets of word lengths induced by $\Language_1, \Language_2$, which are
+special cases of the Parikh image. This required combined reasoning about
+strings and string length has long been identified as a major bottleneck in
+string solvers
 \cite{DBLP:conf/cav/AbdullaACHRRS15,length-aware-solver,approximate-parikh,DBLP:journals/corr/BerzishZG17}.
@@
 
+The Parikh image, more broadly, is a characterisation of formal languages in
+terms of their character counts. Given a language over an alphabet~$\{a_1,
+\ldots, a_k\}$, the Parikh image is a set of $k$-dimensional vectors that
+contains some vector~$\VectorLiteral{m_1, m_2, \ldots, m_k}$ if and only if the
+formal language contains a word in which each $a_i$ occurs $m_i$ times. It is a
+classical result that the Parikh image of every context-free language (and,
+thus, also of every regular language) is a semilinear set~\cite{parikh-theorem},
+i.e., Presburger-definable.
+
 It is possible to compute an existential Presburger formula describing the
@@
 elegant, this construction has several disadvantages, often making it
-unpractical for integration into algorithms. Firstly, the constructed Presburger
+impractical for integration into algorithms. Firstly, the constructed Presburger
 formula contains a linear number of existential quantifiers in the size of the

### 2023-06-22 example.tex
Tweak example, introduction
Closes: #27

@@
 \subsubsection{Propagation of connectivity constraints}
-At this point we know we have an unsatisfiable solution, \Calculus{} does not.
-We press on with concluding that none of the loops of $A'$ at this stage can
-become disconnected since they are at the initial and final states, both of
-which are trivially reachable for any assignments of the non-constant transition
-variables. This means that all valid counter assignments of $A'$ are described
-by \cref{eq:counter-sums} and the flow equations alone without the need for
+At this point we can conclude that none of the loops of $A'$ can become
+disconnected since they are at the initial and final states, both of which are
+reachable for any future assignments of the transition variables. This means
+that all valid counter assignments of $A'$ are described by
+\cref{eq:counter-sums} and the flow equations alone without the need for
 expensive connectivity constraints.
@@
 
-\subsubsection{COMPUTING PRODUCTS}
+\subsubsection{Computing products}
 Now we have two small, stick-like automata after discounting all

### 2023-06-22 introduction.tex
Tweak example, introduction
Closes: #27

@@
 segments of a fixed length \cite{KARHUMAKI1980155} and the more general Parikh
-matrix, which gives more information about a word than the standard Parikh image
-\cite{parikh-matrix}. Another notable generalisation is the p-vector, introduced
+matrix, which contains not only the Parikh vector, but also information about
+the order of letters. Another notable generalisation is the p-vector, introduced
 in~\cite{infinite-words}, which denotes the position of each letter in the word

### 2023-06-22 main.tex
example udpate
Ditch the Verma implementation
Closes: #34
Closes: #2

@@
 \section{Preliminaries}\label{sec:preliminaries}
-
-\subsection{Monoids}
-
-A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
-consisting of the non-empty carrier set of elements, $X$, an associative binary operation
-$X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,
-c \in X$, $(a \MonoidOp b) \MonoidOp c = a \MonoidOp (b \MonoidOp c)$. Finally,
-$\Monoid$ must have an identity element $0_{\Monoid} \in X$ such that
-$0_{\Monoid} \MonoidOp a = a \MonoidOp 0_{\Monoid} =   a$ for every $a \in X$.
-We sometimes use integer multiplication to represent repeated application of
-$\MonoidOp$, e.g. $3a = a \MonoidOp a \MonoidOp a$, for $a \in X$. $\Monoid$ is
-called \textit{commutative} if $\MonoidOp$ also commutes, that is if $a
-\MonoidOp b = b \MonoidOp a$ for all $a, b \in X$.
-
-Finally, a \textit{homomorphism} is a structure-preserving map between two monoids
-$\Monoid_1, \Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
-\MonoidOp_1 b) = \Map(a) \MonoidOp_2 \Map(b)$, where $S_1, S_2$ are the carrier
-sets of $\Monoid_1$ and $\Monoid_2$ respectively, and $\MonoidOp_1, \MonoidOp_2$
-their binary operations.
-
-\subsection{Languages, Finite-state Automata and their Products}
-
-We define an alphabet as a finite set of symbols $\Alphabet$ with words $\Strings$, and
-the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
-Note that $\Strings = \Tuple{\Alphabet;\Concat{};\epsilon}$, is a
-non-commutative monoid, referred to as the free monoid. The string length
-function, $\Length{s}$ is an example of a homomorphism between $\Strings$
-and~$\mathbb{Z}$.
-
-An automaton~$\Automaton$ with alphabet~$\Alphabet$ is
-$\AutomatonTuple$ where $\Transitions \subseteq \States \times \Alphabet \times
-\States$, $\States$ is its states, $\InitialState$ its
-initial state, and $\AcceptingStates$ is its set of accepting states.  We
-write a transition $\Transition = \Tuple{\State, \Label, \State'} \in
-\Transitions$ as $\Transition = \FromLabelTo{\State}{\Label}{\State'}$.
-Similarly, we use the notation $\FromLabelTo{\State}{}{}$ to refer to the set of
-transitions starting in $\State$, and $\FromLabelTo{}{}{\State}$ to refer to the
-set of transitions coming into $\State$, whenever the automaton is clear from
-the context.
-
-We will let variables $\Transition, \Transition', \Transition_1, \ldots,
-\Transition_n$ etc describe transitions, $\State, \ldots, \State_n$ states, and
-$\Automaton, \ldots, \Automaton_n$ automata, and use subscript indexing
-($\Transitions_\Automaton$) to refer to the transitions, states, etc of a given
-automaton.
-
-By a \emph{product} of two automata $\Automaton_1, \Automaton_2$, written
-$\Automaton_1 \times \Automaton_2$, we mean an automaton constructed to run
-$\Automaton_1$ and $\Automaton_2$ in parallel on an input and only accept the
-input if both automata would do so.
-
-We refer to the resulting product states as tuples, $\Tuple{\State, \State'}$,
-which represent the state of the product automaton where $\Automaton_1$ would be
-in $\State$ and $\Automaton_2$ would be in $\State'$. Note that since we use
-ordered tuples the product is technically (but w.l.o.g) not commutative; the
-left-hand-side must come from the left-hand term. The sole purpose of this
-matching is to allow us to speak with precision about the origins of components
-in a product.
-
-\subsection{The Parikh Map and its Image}
-Formally, the \textit{Parikh map} over an alphabet $\Alphabet=
-\left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:
-$$
-\begin{aligned}
-& \ParikhMap: \MapFromTo{\Strings}{\natural^k} \\
-& \ParikhMap(s) = \VectorLiteral{\#a_1(s), \#a_2(s), \ldots, \#a_k(s)}
-\end{aligned}
-$$
-
-That is, $\ParikhMap(s)$ is a vector of the number of occurrences of each
-character in the language for a given string $s$. For example, for  $\Alphabet =
-\Set{a, b}$, we would have $\ParikhMap(abb) = \VectorLiteral{1, 2}$.
-
-We define the image of this map, the \textit{Parikh image}, of some subset of
-the language $\Language \subseteq \Strings$ as:
-\[
-\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
-\]
-
-Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
-1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
-notation $\#l(w)$ to talk about an individual letter $l$ in a word $w$. For
-example, for the Parikh vector above, we would have $\CountOf{a} = 1$. You have
-already seen this in \cref{sec:introduction:parikh}.
-
-Parikh's theorem states that any context-free language has a letter-equivalent
-regular language (c.f.~\cite{construction} for a construction of such automata
-from context-free grammars and~\cite{bounds} for bounds on its size). However,
-there are languages that are not context-free that also have semilinear images
-under~$\ParikhMap$ (e.g. $\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) =
-\ParikhMap((abc)^*) = \CountOf{a} = \CountOf{b} = \CountOf{c} \land \CountOf{a}
-\geq 0$). This means they can be represented as a quantifier-free Presburger
-formula.
-
-Note that while Parikh's theorem applies to context-free languages, in this
-paper we focus only on regular languages.
-
-\subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
-\label{sec:verma}
-
-Since Parikh images are semilinear, any Parikh image can be written as a set of linear
-equations. The following construction for an NFA $\Automaton = \AutomatonTuple$ is based on the version for context-free grammars presented in \cite{generate-parikh-image}:
-\begin{equation}
-\begin{aligned}
-\ParikhMap(\Automaton) := 
-& \AndComp{\Letter \in \Alphabet}{\LetterVar_{\Letter} = \sum_{\Transition = \FromLabelTo{}{\Letter}{} \in \Transitions} \TransitionVar_{\Transition}
-}\\
-& \AndComp{\State \in \AcceptingStates}{\FinalStateVar_{\State} > 0 \longrightarrow  \StateVar_{\State} > 0}\\
-& \AndComp{\State \in \States}{
-  \left(\FinalStateVar_{\State} \text{ if } \State \in \AcceptingStates \text{ otherwise } 0 \right) +
-  \sum_{\Transition \in \FromLabelTo{\State}{}{}} \Filter(\Transition) - \sum_{\Transition \in \FromLabelTo{}{}{\State}} \Filter(\Transition)
-= \begin{cases}
-    1 \text{  if $\State = \InitialState$} \\
-    0 \text{ otherwise}
-  \end{cases}
-}\\
-& \AndComp{\Transition = \FromLabelTo{\State'}{}{\State} \in \Transitions}{
-  \TransitionVar_{\Transition} > 0 \longrightarrow \StateVar_{\State} > 0
-} \\
-& \AndComp{\State \in F}{
-  \StateVar_{\State} > 0 \longrightarrow \FinalStateVar_{\State} = 1
-} \\
-&\AndComp{\State \in \States}{
-  \StateVar_{\State} > 0 \longrightarrow
-  \OrComp{\Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions}{
-    \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
-    \TransitionVar_{\Transition} \geq 1 \land
-  \StateVar_{\State'} \geq 1
-    }
-}
-\end{aligned}
-\label{eq:generate-parikh}
-\end{equation}
-
-All variables $\TransitionVar_\Transition, \StateVar_\State, \FinalStateVar_\State$ are
-existentially quantified and the free variables $\LetterVar_\Letter$ make up the
-image. $\StateVar_\State$ represents the distance of state $\State$ from
-$\InitialState$ in a spanning tree, $\TransitionVar_\Transition$ how many times
-a transition $\Transition$ is used, and $\FinalStateVar_\State$ whether a given
-accepting state $\State \in \AcceptingStates$ is the actually used final state.
-
-In this paper we refer to this model as the baseline approach, though we also
-apply optimisations as described in \cref{sec:implementing-baseline}. The
-calculus introduced in this paper, by contrast, lazily enforces the
-connectedness constraint of this encoding (the final three clauses) while also
-interleaving the computation of products of automata and propagating information
-between the steps to reduce the amount of work that needs to be done.
-
-\subsection{Theorem Proving}
-
-We assume basic familiarity with first-order logic, Presburger
-arithmetic, and the classical sequent calculus. For reference, see
-e.g. \cite{Fitting96a}.
+\input{preliminaries.tex}
 
@@
 We start by defining our calculus, \Calculus{}, for one automaton, and only
-extend it to products of automata in \cref{sec:multiple}. Assume an NFA
-$\Automaton = \AutomatonTuple$ with $\NrTransitions$ transitions $\Transitions =
+extend it to products of automata in \cref{sec:multiple}. We assume basic
+familiarity with first-order logic, Presburger arithmetic, and the classical
+sequent calculus. For reference, see e.g. \cite{Fitting96a}.
+
+First, assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
+transitions $\Transitions =
 \Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. For convenience, we then
@@
 heavy use of product computation than baseline does, instead relying more on
-\Princess{}'. This situation would unfairly advantage \Calculus{} if our
+\Princess{}. This situation would unfairly advantage \Calculus{} if our
 automata library was the bottleneck due to our automata library being

### 2023-06-22 preliminaries.tex
Ditch the Verma implementation
Closes: #34
Closes: #2

\subsection{Monoids}
A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
consisting of the non-empty carrier set of elements, $X$, an associative binary operation
$X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,
c \in X$, $(a \MonoidOp b) \MonoidOp c = a \MonoidOp (b \MonoidOp c)$. Finally,
$\Monoid$ must have an identity element $0_{\Monoid} \in X$ such that
$0_{\Monoid} \MonoidOp a = a \MonoidOp 0_{\Monoid} =   a$ for every $a \in X$.
We sometimes use integer multiplication to represent repeated application of
$\MonoidOp$, e.g. $3a = a \MonoidOp a \MonoidOp a$, for $a \in X$. $\Monoid$ is
called \textit{commutative} if $\MonoidOp$ also commutes, that is if $a
\MonoidOp b = b \MonoidOp a$ for all $a, b \in X$.

Finally, a \textit{homomorphism} is a structure-preserving map between two monoids
$\Monoid_1, \Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
\MonoidOp_1 b) = \Map(a) \MonoidOp_2 \Map(b)$, where $S_1, S_2$ are the carrier
sets of $\Monoid_1$ and $\Monoid_2$ respectively, and $\MonoidOp_1, \MonoidOp_2$
their binary operations.

\subsection{Languages, Finite-state Automata and their Products}

We define an alphabet as a finite set of symbols $\Alphabet$ with words $\Strings$, and
the concatenation operation as $s_1 \Concat{} s_2$ over two strings $s_1, s_2$.
Note that $\Strings = \Tuple{\Alphabet;\Concat{};\epsilon}$, is a
non-commutative monoid, referred to as the free monoid. The string length
function, $\Length{s}$ is an example of a homomorphism between $\Strings$
and~$\mathbb{Z}$.

An automaton~$\Automaton$ with alphabet~$\Alphabet$ is
$\AutomatonTuple$ where $\Transitions \subseteq \States \times \Alphabet \times
\States$, $\States$ is its states, $\InitialState$ its
initial state, and $\AcceptingStates$ is its set of accepting states.  We
write a transition $\Transition = \Tuple{\State, \Label, \State'} \in
\Transitions$ as $\Transition = \FromLabelTo{\State}{\Label}{\State'}$.
Similarly, we use the notation $\FromLabelTo{\State}{}{}$ to refer to the set of
transitions starting in $\State$, and $\FromLabelTo{}{}{\State}$ to refer to the
set of transitions coming into $\State$, whenever the automaton is clear from
the context.

We will let variables $\Transition, \Transition', \Transition_1, \ldots,
\Transition_n$ etc describe transitions, $\State, \ldots, \State_n$ states, and
$\Automaton, \ldots, \Automaton_n$ automata, and use subscript indexing
($\Transitions_\Automaton$) to refer to the transitions, states, etc of a given
automaton.

By a \emph{product} of two automata $\Automaton_1, \Automaton_2$, written
$\Automaton_1 \times \Automaton_2$, we mean an automaton constructed to run
$\Automaton_1$ and $\Automaton_2$ in parallel on an input and only accept the
input if both automata would do so.

We refer to the resulting product states as tuples, $\Tuple{\State, \State'}$,
which represent the state of the product automaton where $\Automaton_1$ would be
in $\State$ and $\Automaton_2$ would be in $\State'$. Note that since we use
ordered tuples the product is technically (but w.l.o.g) not commutative; the
left-hand-side must come from the left-hand term. The sole purpose of this
matching is to allow us to speak with precision about the origins of components
in a product.

\subsection{The Parikh Map and its Image}
Formally, the \textit{Parikh map} over an alphabet $\Alphabet=
\left\{a_1, \ldots, a_k \right\}$ is defined as in \cite{kozen}:
$$
\begin{aligned}
& \ParikhMap: \MapFromTo{\Strings}{\natural^k} \\
& \ParikhMap(s) = \VectorLiteral{\#a_1(s), \#a_2(s), \ldots, \#a_k(s)}
\end{aligned}
$$

That is, $\ParikhMap(s)$ is a vector of the number of occurrences of each
character in the language for a given string $s$. For example, for  $\Alphabet =
\Set{a, b}$, we would have $\ParikhMap(abb) = \VectorLiteral{1, 2}$.

We define the image of this map, the \textit{Parikh image}, of some subset of
the language $\Language \subseteq \Strings$ as:
\[
\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
\]

Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
notation $\#l(w)$ to talk about an individual letter $l$ in a word $w$. For
example, for the Parikh vector above, we would have $\CountOf{a} = 1$. You have
already seen this in \cref{sec:introduction:parikh}.

Parikh's theorem states that any context-free language has a letter-equivalent
regular language (c.f.~\cite{construction} for a construction of such automata
from context-free grammars and~\cite{bounds} for bounds on its size). However,
there are languages that are not context-free that also have semilinear images
under~$\ParikhMap$ (e.g. $\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) =
\ParikhMap((abc)^*) = \CountOf{a} = \CountOf{b} = \CountOf{c} \land \CountOf{a}
\geq 0$). This means they can be represented as a quantifier-free Presburger
formula.

Note that while Parikh's theorem applies to context-free languages, in this
paper we focus only on regular languages.

\subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
\label{sec:verma}

Since Parikh images are semilinear, any Parikh image can be written as a set of
linear equations. The version for context-free grammars presented in
\cite{generate-parikh-image} can be straightforwardly adapted for use with a
product of regular languages or Parikh automta. For an intuition, the approach
consists of first computing the product, then assigning each state and
transition an existentially quantified non-negative integer variable, then
describing all paths through the automaton through two sets of equations: one
that preserves the incoming flow across the variables of a state's incoming
transitions is equal to the flow across its outgoing transitions, and one set of
constraints that enforces coherence in the presence of loops by ordering states
by distance in a spanning tree rooted in the initial state.

We refer to this model as the baseline approach, though we also apply
optimisations as described in \cref{sec:implementing-baseline}. The calculus
introduced in this paper, by contrast, lazily enforces the connectedness
constraint while also interleaving the computation of products of automata and
propagating information between the steps to reduce the amount of work that
needs to be done.

### 2023-06-22 generalised.tex
Work on section 4. Still WIP!
Closes: #35
Closes: Ta bort 4.1 #13

--- 2023-06-22 main.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2023}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\usepackage{amsmath,empheq,fancybox}
-\usepackage{paralist}
-\usepackage{url}
-\usepackage{color}
-\usepackage{textcomp,listings}
-\usepackage{array}
-\usepackage{mymacros}
-\usepackage{microtype}
-\usepackage{listings}
-\usepackage{csquotes}
-\usepackage{proof}
-\usepackage[capitalise]{cleveref}
-\usepackage{algorithm2e}      
-\usepackage{multirow}
-\usepackage{mathpartir}
-\usepackage{amsthm}
-\usepackage{numprint}
-\usepackage{ebproof}
-\usepackage[makeroom]{cancel}
-
-\usepackage{mathtools} % Bonus
-
-\usepackage{enumitem}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{example}{Example}[section]
-
-\newlist{constraints}{enumerate}{3}
-\setlist[constraints,1]{%
-    label=\sffamily{(\roman*):},
-    ref=\normalfont{Constraint (\roman*)},
-    wide,itemsep=0pt,topsep=0pt
-}    
-\crefname{constraintsi}{}{}
-\Crefname{constraintsi}{}{}
-
-\lstset{
-    columns=fullflexible,
-    showspaces=false,
-    showtabs=false,
-    breaklines=true,
-    showstringspaces=false,
-    breakatwhitespace=true,
-    escapeinside={(*@}{@*)},
-    commentstyle=\color{greencomments},
-    keywordstyle=\color{bluekeywords},
-    stringstyle=\color{redstrings},
-    numberstyle=\color{graynumbers},
-    basicstyle=\ttfamily\small,
-    framesep=12pt,
-    xleftmargin=12pt,
-    tabsize=4,
-    captionpos=b
-}
-
-\newif\ifcomments
-\commentstrue
-\newif\ifoutline
-\outlinetrue
-
-\newcommand{\contents}[1]{\ifoutline{\color{blue}
-    \begin{itemize}
-    #1
-    \end{itemize}
-  }\fi}
-
-\allowdisplaybreaks[1]
-
-\begin{document}
-
-%% Title information
-\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
-   counts in strings of a given language. Automata-based string solvers
-   frequently need to compute the Parikh image of large products (intersections)
-   of nondeterministic automata, which in many operations is both prohibitively
-   slow and memory-intensive. We contribute a novel understanding of how Parikh
-   maps can be tackled as a constraint solving problem to solve real-world
-   constraints stemming from functions on regular languages, most notably the
-   length constraint. Furthermore, we show how this formulation can be
-   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
-   prover supporting Presburger logic.  We implement \Calculus{} in a tool
-   called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving Parikh automata
-   intersection problems produced when solving standard string constraint
-   benchmarks involving string length constraints. We show that our solution
-   strictly outperforms the standard approach described
-   in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wide
-   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
-   model checker.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-  <ccs2012>
-  <concept>
-  <concept_id>10003752.10003766.10003773.10003775</concept_id>
-  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003766.10003776</concept_id>
-  <concept_desc>Theory of computation~Regular languages</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10003794</concept_id>
-  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10010124.10010138.10010142</concept_id>
-  <concept_desc>Theory of computation~Program verification</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10011192</concept_id>
-  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  </ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Quantitative automata}
-\ccsdesc[300]{Theory of computation~Regular languages}
-\ccsdesc[500]{Theory of computation~Automated reasoning}
-\ccsdesc[300]{Theory of computation~Program verification}
-\ccsdesc[300]{Theory of computation~Verification by model checking}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}\label{sec:introduction}
-\input{introduction.tex}
-
-\section{An Intuition for Our Approach}\label{sec:intuition}
-\input{example.tex}
-
-\section{Preliminaries}\label{sec:preliminaries}
-\input{preliminaries.tex}
-
-\section{Projections on Parikh Images}\label{sec:generalised}
-
-It is easy to see that the Parikh map~$\ParikhMap$ represents a
-homomorphism from the (free) non-commutative monoid~$\Strings$ to the
-(free) commutative monoid~$\Naturals^k$. As we are often not
-interested in the full Parikh image, but only in some projections of
-it. We consider \emph{arbitrary} homomorphisms
-$\Map:\: \Sigma^* \to \Monoid$, where
-$\Monoid = \left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative
-monoid. We give several
+The Parikh map~$\ParikhMap$ represents a homomorphism from the (free)
+non-commutative monoid~$\Strings$ to the (free) commutative
+monoid~$\Naturals^k$. We are, however, often interested in projections of the
+Parikh map, rather than the full image. We therefore consider \emph{arbitrary}
+homomorphisms $\Map:\: \Sigma^* \to \Monoid$, where $\Monoid =
+\left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative monoid. We give several
 examples of such projections on Parikh images later in this section.
@@
 
-%\subsection{Applications of generalised Parikh images}\label{sec:applications}
-
-\subsection{Example~1: Reasoning about String Length}
-
+\begin{example}[String Length]\label{ex:length}
 One such simplifying homomorphism can express string
@@
 map. This mapping is relevant when solving constraints that combine
-language membership as well as word length, for instance the formula
+language membership with string length, for instance the following constraint:
 \begin{equation}\label{eq:stringLength}
@@
 \end{equation}
-mentioned in the introduction. To solve this formula, let
+
+To solve this formula, let
 $\Monoid = \Naturals$, and define the homomorphism~$L$ by $L(a) = 1$
@@
 like~$\alpha \in L(\Language_1)$, avoiding the computation of
-the complete image~$L(\Language_1)$.
+the complete image~$L(\Language_1)$. This encoding can be seen applied to automaton $B$ in
+\cref{ex:string-constraints} (see \cref{fig:aut_b}).
+\end{example}
 
-\subsection{Example~2: Generalised String Constraints with Integer
-  Datatype}\label{sec:parikh-automata}
+\subsection{Integer Datatype Constraints on Strings}\label{sec:parikh-automata}
 
@@
 string constraints~\cite{ostrich-plus}. Consider, for instance, a
-constraint involving the substring operation:
-\begin{equation}
-  \label{eq:substring}
-  x \in \Language_1 \wedge \; 0 \leq n \leq m \leq |x| \wedge x[n:m] \in \Language_2
+constraint involving the substring operation of \cref{const:s1-substring}:
+\begin{equation}\label{eq:substring}
+    s_1 = \Substring(s_2, i, n)
 \end{equation}
-in which $x$ is a string variable, $n, m$ are integer variables, and
-$x[n:m]$ denotes the substring of $x$ starting at position~$n$ running
-until position~$m$. This constraint belongs to an expressive fragment
-of string logic that cannot be decided by most state-of-the-art string
-solvers. To illustrate the decision procedure proposed in
-\cite{ostrich-plus} for string constraints of this kind, suppose that
-$\Language_2$ is defined by the regular
-expression~$\mathtt{(ab|c)*}$. An automaton recognising this language
-is shown in \cref{fig:parikh-automata} (left).
 
-We can model \eqref{eq:substring} using the notion of a \emph{Parikh
+That is $s_1$ is an $n$-length substring of $s_2$ starting at offset $i$ (in
+some variants , $n$ is $i + |n|$, i.e. the ending index). That constraint
+belongs to an expressive fragment of string logic that cannot be decided by most
+state-of-the-art string solvers.
+
+We can model \cref{eq:substring} using the notion of a \emph{Parikh
   automaton}~\cite{parikh-automata,expressiveness}. A Parikh automaton
-is an automaton in which transitions are labelled both with characters
+is a standard automaton is extended such that transitions are labelled both with characters
 from the alphabet~$\Alphabet$, and with offset vectors defining the
@@
 \end{definition}
-
-\begin{figure}[t]
-  \centering
-  \raisebox{5ex}{
-    \includegraphics[scale=0.9]{counter_automaton_pre}
-    }
-  \hfill
-  \includegraphics[scale=0.9]{counter_automaton}
-  \caption{An automaton recognising the language $\mathtt{(ab|c)*}$
-    (left), and a Parikh automaton describing the pre-image of this
-    language under the substring operation~$\cdot[n:m]$ (right). For
-    sake of presentation, the Parikh automaton contains
-    $\epsilon$-transitions, which could be eliminated in the standard
-    way.}\label{fig:parikh-automata}
-    \Description[An 2-state automaton (left) and a slightly larger automaton with counters (right).]{
-      The rightmost automaton is the starts with an initial state with a self-loop on any character that increments both registers by one, followed by an epsilon transition to a state with a self-loop on c that increments the lower register by one.
-From that state, the automaton has an epsilon-transition into the accepting state, as well as a transition on a into a state with only a transition back again, both incrementing the lower register by one.
-The accepting state has a self-loop on any character that does nothing to the registers.
-      }
-    \end{figure}
 
@@
 \fi
-
-\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
-
-We start by defining our calculus, \Calculus{}, for one automaton, and only
-extend it to products of automata in \cref{sec:multiple}. We assume basic
-familiarity with first-order logic, Presburger arithmetic, and the classical
-sequent calculus. For reference, see e.g. \cite{Fitting96a}.
-
-First, assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
-transitions $\Transitions =
-\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. For convenience, we then
-introduce the following supporting notations:
-
-\begin{definition}
-  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
-  transitions $\Transitions$ represents a path through $\Automaton$ using
-  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
-  \State_{k+1})$ holds), passing zero or more labels $\Label_1, \ldots,
-  \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
-  \InitialState$. The end state, $\State_n$, is not necessarily accepting. If it
-  is, we say the path is also \textit{accepting}.
-  \end{definition}
-
-\begin{definition}
-  Moreover, we talk about the \textit{set of paths} of an automaton,
-  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
-  valid paths through $\Automaton$. Additionally, we use the
-  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
-  state~$\State$.
-\end{definition}
-
-\begin{definition}
-  For a path $\Path = \Tuple{\State_0 \Label_1 \State_1 \RepeatSum{,}
-  \State_n}$, its \textit{word} $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat}
-  \Label_n$ is the word read out on its labels.
-\end{definition}
-
-\begin{definition}
-  The \textit{states} of a path $\Path = \Tuple{\State_0 \Label_1 \State_1
-  \RepeatSum{,} \State_n}$, $\StatesOf(\Path) = \Set{\State_0\RepeatSum{,}
-  \State_n}$ are the states visited along $\Path$. Note that $\InitialState \in
-  \StatesOf(\Path)$ for every path since all paths start in the initial state.
-\end{definition}
-
-\begin{definition}
-  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$ to state
-  $\State$, written $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of
-  transitions such that every accepting path $p$ where $\State \in
-  \StatesOf(\Path)$ contains a transition $\Transition \in C$. A cut does not
-  exist for every state; notably it never exists for $\InitialState$.
-\end{definition}
-
-\begin{definition}
- The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
- number of times a transition $\Transition =
- \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
- $p$.
-\end{definition}
-
-We then introduce the two predicates into our calculus with the following
-definitions:
-
-\begin{definition}\label{def:single-image}
-  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
-  $\Automaton = \AutomatonTuple$, modulo some map $\Map : \Alphabet^* \to \Monoid$ to a
-  commutative monoid $\Monoid$ as described in \cref{sec:generalised}
-  and with a transition selection function
-  $\Filter:\: \Transitions \to \Naturals$ holds when
-  $\MonoidElement \in \Monoid$ is an element of the Parikh image of
-  $\Language(\Automaton)$
-  modulo $\Map$, or more formally when there is an accepting path
-  $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
-  that $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
-  for all $t \in \Transitions$, and $\MonoidElement = \Map(\WordOf(\Path))$.
-\end{definition}
-
-\begin{definition}
-  $\Connected(\Automaton, \Filter)$ for some automaton $\Automaton =
-  \AutomatonTuple$ holds when for every $\Transition =
-  \FromLabelTo{\State}{}{\State'} \in \Transitions$, $\Filter(\Transition) > 0
-  \implies \exists \Path \in \Paths(\Automaton)$ $\Filter(\Transition_i) > 0$
-  for $\Transition_i \in p$, and $\State \in \StatesOf(p)$, or in words that
-  there exists some $\Filter$-selected valid path that reaches $\Transition$'s
-  starting state, $\State$. Intuitively, it represents the condition that
-  $\Automaton$ is connected with respect to the selection function $\Filter$ for
-  every transition. It is redundant to $\Image$ by design.
-\end{definition}
-
-We present the rules of \Calculus{} for one automaton in
-\cref{tbl:rules:single}. The rules operate on sets of formulas, and
-can be interpreted as rules of a one-sided sequent calculus, in
-which all formulas are located in the antecedent~\cite{Fitting96a}.
-Each of the rules should be read and
-applied bottom-up, and relates
-premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
-conclusion~$\SomeClause$. When constructing a proof, we start from some
-root~$\SomeClause$, and then apply proof rules to the goals of the proof
-until a goal can be closed, or no more rule is applicable.
-
-A proof in which all proof goals could be closed shows that the formulas
-in the root~$\SomeClause$ of the proof are inconsistent, and do not
-have any solutions. A proof goal that could not be closed, but to which no
-more rules are applicable, gives rise to  a solution of the formulas in the
-root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic, so that a solution can be computed using
-standard algorithms~\cite{Fitting96a}.
-
-We use the convention of splitting the
-formulas into linear (in)equalities ($\SomeInequalities$)
-and other formulas ($\SomeClause$), and assume that predicates
-$\Image$ and $\Connected$ only occur positively.
- Since
-our rules operate by adding and matching linear (in)equalities in a proof goal,
-we use the shorthand of listing the matched inequalities as antecedents
-e.g., in \Propagate{}).
-The filtering function~$\Filter$ is evaluated symbolically, and can in practice
-be read as a function from transitions to $\Naturals$-valued terms
-(e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{}, described in
-\cref{sec:implementation}, $\Filter$~is a vector of fresh variables with the
-same size as~$\Transitions$.
-
-We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
-transitions of an automaton~$\Automaton$. Additionally, for an
-automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
-like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
-\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
-$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
-instance, the basis for the matched linear inequalities is
-implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
-\left(X;\MonoidOp;0_{\Monoid}\right)$ and the map into it $\Map : \Strings
-\rightarrow X$, we also allow mapping over transitions:
-$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$.
-
-We require that rules can only be applied when they add new formulas
-on every created branch (the notion of \emph{regularity} of a proof is
-required~\cite{Fitting96a}). For example, this means that \Split{} can
-only be applied to proof goals that contain neither
-$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
-never be applied to split on the same term twice on the same branch.
-This suggests a proof strategy where you \Propagate{} when you can, \Split{}
-when you must, and \Subsume{} when neither is possible anymore.
-
-\begin{table}
-\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-  Name & \text{Rule} & Side conditions\\
-  \midrule
-
-  % EXPAND
-  \Expand & 
-    \inferrule
-  {\Connected(\Automaton, \Filter) ,~ \FlowEq(\Automaton, \Filter) ,~ \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
-  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
-  None \\[4ex]
-
-  % SPLIT
-  \Split & 
-  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  if $\Transition \in \Transitions_\Automaton$ \\[4ex]
-
-  % PROPAGATE
-  \Propagate &
-  \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
-  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[4ex]
-
-  % SUBSUME
-  \Subsume &
-  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  \Split{} and \Propagate{} cannot be applied \\
-  \bottomrule
-  \end{tabular}
-  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
-\end{table}
-We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
-of existentially quantified linear inequalities with the following definition,
-where we assign fresh, existentially quantified variables to
-$\FinalStateVar_\State, \Filter(\Transition)$ for every $\State \in
-\AcceptingStates, \Transition \in \Transitions$:
-\[
-\begin{aligned}
-  \FlowEq(\Automaton, \Filter) &= \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
-  \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}\\
-  \Sink(\State) & = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
-  \In(\State, \Filter) & = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
-  \StartFlow(\State)  &= 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
-  \Out(\State, \Filter) &= \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
-\end{aligned}
-\]
-
-In addition to \cref{tbl:rules:single}, we assume the existence of a rule
-\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
-formulae, and for the elements of~$\Monoid$.
-
-The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
-$\Automaton$. It states that we are only allowed to use transitions attached to
-a reachable state, and is necessary to ensure connectedness in the presence of
-cycles in~$\Automaton$.
-
-\textsc{Expand} expands the predicate into its most basic rules; one set of
-linear equations synchronising the transitions mentioned by~$\Filter$ to the
-corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
-the standard Parikh image formulation, as described by~\FlowEq. Since
-$\Connected$ and $\Image$ are partially redundant and the difference is covered
-by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
-this sense, we split the semantics of the $\Image$~predicate into its counting
-aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
-$\Connected$).
-
-Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
-contested transition from a potential solution before concluding that it must be
-included. Intuitively, this is what guarantees our ability to make forward
-progress by eliminating paths through~$\Automaton$.
-
-A decision procedure for our predicate in a tableau-based automated theorem
-prover would start by expanding the predicate using the $\Expand{}$~rule. A
-theorem prover would perform algebraic substitution on the underlying constants
-of~$\Filter$, boiling them down to choices of branches, which depend on one
-single variable, and loop transitions. This logic corresponds to the placement
-of counters for optimally edge-profiling the CFG of a program, making up a
-minimum-spanning tree of the automaton~\cite{path-profiling}.
-
-In order to make the examples below tractable, we will assume the existence of a
-rule \EquationReasoning{} that allows us to perform standard algebraic reasoning
-on linear inequalities. This rule is not necessary for correctness or
-completeness, but shortens the examples considerably.
-
-\subsection{An Example}
-
-Starting with~$\AcaOrBc{}$, where $\Map$ is the length function, in effect
-$\Transition \mapsto 1$ for transitions, and the constraints that the length is
-odd using the same trick as in \cref{sec:motivation}, we have the
-definitions in \cref{fig:example:single:equivalences} (omitting existential
-quantifiers and $x \geq 0$ for every variable to avoid clutter).
-
-\begin{figure}[ht]
-  We define $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$ as in the following two
-  equations, and then apply \EquationReasoning{} (under the implicit assumption
-  that every RHS is $\geq 0$) to obtain the third definition:
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = \TransitionVar_1 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_2 \\
-        \Filter(\FromLabelTo{A}{c}{A}) & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = \TransitionVar_5 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \hspace{0.5cm}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        % S
-        \TransitionVar_4 &= \TransitionVar_5 + \TransitionVar_2 - 1 \\
-        % A
-        \TransitionVar_1 &= \TransitionVar_5 \\
-        % B
-        \TransitionVar_2 &= \TransitionVar_4 + \TransitionVar_6 \\
-        % F
-        \TransitionVar_5 + \TransitionVar_6 &= \FinalStateVar_1 \\
-        \FinalStateVar_1 &= 1 \\
-      \end{aligned}  
-    \end{equation*}    
-  \end{minipage}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \caption{Equivalences defining $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$
-  respectively.}\label{fig:example:single:equivalences}
-  \end{figure}
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo[]{
-{    \begin{array}{lll}
-      1 - \TransitionVar_6 = 0 & \land \TransitionVar_6 = 1 & \land  \TransitionVar_4 + 1 > 0 \\
-      \land \TransitionVar_3 = 0 &\land  2 + \TransitionVar_4 > 0 & \land k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-{    \begin{array}{lll}
-      \Connected(\AcaOrBc{}, \Filter) & \land 1 - \TransitionVar_6 = 0 \\
-      \land \TransitionVar_6 = 1 & \land \TransitionVar_4 + 1 > 0 \land \\
-      \land \TransitionVar_3 = 0 & \land 2 + \TransitionVar_4 > 0 \land \\
-      k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Propagate]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 \geq 0 \land \\
-      2 + \TransitionVar_4 \geq 0 \land \\
-      2k = 2 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{aligned}
-  }
-  % BRANCH:  = 0'
-  \hypo{
-    {\begin{array}{lll}
-      \TransitionVar_3 > 0 & \land 
-      \TransitionVar_6 = 0 \\ \land
-      2\TransitionVar_6 + \TransitionVar_4 = 0 & \land
-      1 - \TransitionVar_6 > 0\\  \land 
-      2k = 1 + \TransitionVar_3
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_3 > = 0 \land \\
-      \TransitionVar_6 = 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0\\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\Propagate{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{aligned} % BRANCH: > 0
-  } % SPLIT
-  \infer2[\Split{} $1 - \TransitionVar_6$]{ \Connected(\AcaOrBc{}, \Filter) \land 2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6 }
-  \infer1[\EquationReasoning{}]{
-    \Connected(\AcaOrBc{}, \Filter) \land
-    2k + 1 =
-    \sum\limits_{\Transition \in \Transitions} (\Transition \mapsto 1)(\Label) \cdot \Filter(\Transition)
-  }
-  \infer1[\Expand{}]{\Image{}_{\AcaOrBc{}, \Transition \mapsto 1}(\Filter, 2k + 1)}
-\end{prooftree}
-\caption{A derivation for \Calculus{} computing odd lengths in $\AcaOrBc{}$.}\label{fig:derivation:single}
-\end{figure}
-
-In \cref{fig:derivation:single}, we see how we start by expanding the predicate
-using the (simplified) flow equations. We can also see how interleaving
-reasoning on the corresponding linear equations helps the deduction. In both
-branches the reasoning is similar: we conclude that when choosing either path at
-the starting state we must avoid the other one, perform propagation based on
-that fact, use some algebraic reasoning to derive the expected forms of the
-upper and lower bounds to propagate the now disconnected transitions, and
-finally subsume, removing the $\Connected{}$ predicate when we are unable to use
-either of the rules, leaving only a set of linear equations. To obtain a solution from the
-proof, one would need to perform standard model generation on the remaining
-constraints to obtain values for the sought variable $\MonoidElement$. The full
-Parikh image is the disjunction of the constraints left at the two
-leaves after quantifier elimination.
-
-Following the values of $\Filter$, we can read out the path corresponding to our solution in the product, and with that a word.
-In our case, for the right branch we have $\Filter(\FromLabelTo{S}{b}{B}) = 0,
-\Filter(\FromLabelTo{A}{c}{A})  > 0,
-\Filter(\FromLabelTo{B}{b}{S}) = 0,
-\Filter(\FromLabelTo{S}{a}{A}) = 1 ,
-\Filter(\FromLabelTo{A}{a}{F}) = 1 ,
-\Filter(\FromLabelTo{B}{c}{F}) = 0$.
-We can read out a word along it: aca, accca, etc, and the corresponding constraint: $\exists k : \TransitionVar_3 = 2k -1 \land \TransitionVar_3 > 0 \land k \geq 0$.
-
-\subsection{Correctness of \Calculus{}}\label{sec:single:correct}
-
-Our correctness proof of \Calculus{} consists of two main parts: we
-first show that the construction of a proof always terminates, and
-then that each of the proof rules in \cref{tbl:rules:single} is
-an equivalence transformation, i.e., does not change the set of
-satisfying assignments of a formula. In combination, those two results
-immediately imply that \Calculus{} gives rise to a decision procedure.
-
-\subsubsection{\Calculus{} terminates}
-\begin{lemma}\label{lma:single-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the predicates
-  $\Image$ and $\Connected$ only occur positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:single}.
-\end{lemma}
-
-\begin{proof}
-  The rule~\Expand{} can only be applied finitely often, since each
-  application removes one $\Image$ predicate, and none of the rules
-  introduce new instances of the predicate. The rule~\Subsume{} can
-  only be applied finitely often, since it strictly decreases the
-  combined number of $\Image$ and $\Connected$ predicates in sets of
-  formulas, and none of the rules increases that number.
-
-  To show termination of \Split{} and \Propagate{}, observe that the
-  $\Filter$ in a predicate~$\Connected(\Automaton, \Filter)$ is never
-  updated on a proof branch, which means that the set of terms
-  $\Filter(t)$ for $t \in \Automaton$ on every branch is finite. Each
-  application of \Split{} and \Propagate{} adds a new
-  formula~$\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$
-  to a proof goal, which can only happen finitely often.
-\end{proof}
-
-\subsubsection{The rules in \cref{tbl:rules:single} are solution-preserving}
-
-\begin{lemma}\label{lma:single-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:single}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
-  conclusion~$\SomeClause$ if and only if it satisfies one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  This property has to be shown by analysing the possible applications
-  of each proof rule.
-
-  \Expand{} unfolds the definition of the $\Image$ predicate. To show
-  that the rule is solution-preserving we prove the equivalence of the
-  upper and lower sets of formulas:
-  \begin{itemize}
-  \item Assume that $\beta$ satisfies the conclusion, which means that
-    there is some accepting path
-    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ with
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$ and $\val_\beta(\MonoidElement) =
-    \Map(\WordOf(\Path))$. Since immediately implies that $\beta$
-    satisfies $\Connected(\Automaton, \Filter)$, since a path is
-    connected, and $\FlowEq(\Automaton, \Filter)$ since an accepting
-    path satisfies the flow equations. The
-    equation~$\MonoidElement = \sum_{\Transition \in
-      \Transitions_\Automaton}\Filter(\Transition) \cdot
-    \Map(\Transition)$ holds because of
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$.
-  \item Assume that $\beta$ satisfies the premise, which implies that
-    $\val_\beta(\Filter)$ describes a consistent, connected flow of
-    the automaton. By the same argument as in
-    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
-    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
-    that each transition~$\Transition$ occurs on $\Path$ exactly
-    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
-  \end{itemize}
-
-  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
-  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
-  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
-  be satisfied, implying the property.
-
-  For \Propagate{}, suppose that
-  $\SeparatingCut(C, \Automaton, \State)$, which means that every
-  accepting path visiting~$q$ contains at least one of the transitions
-  in $C$. For a $\beta$ satisfying $\Connected(\Automaton, \Filter)$
-  and $\Filter(\Transition') = 0$ for $\Transition' \in C$, this means
-  that also $\val_\beta(\Filter(\Transition)) = 0$ has to hold.
-
-  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
-  then a goal must contain~$\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
-  formulas in $\SomeInequalities$ are inconsistent, an application of
-  \Subsume{} is trivially solution-preserving; therefore assume that
-  $\SomeInequalities$ is consistent, which means that it contains
-  exactly one of $\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for each $\Transition$. Since
-  \Propagate{} is not applicable, the transitions~$\Transition$ with
-  $\Filter(\Transition) > 0$ must form a connect sub-graph of the
-  automaton; this means that $\Connected(\Automaton, \Filter)$ is
-  redundant as it is implied by $\SomeInequalities$.
-\end{proof}
-
-\section{Parikh Images from Products of Automata}\label{sec:multiple}
-
-We now generalise our calculus to natively work with intersections of
-regular languages, or equivalently products of automata. For this
-extension, we change the main predicate~$\Image$ to be indexed by a
-vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
-simplicity, we assume that the sets of states of the $k$ automata (and
-therefore also the transition sets) are pairwise disjoint. 
-\begin{definition}\label{def:multiple}
-  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
-  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
-  $\Monoid$,
-  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
-  is a transition selection function, and
-  $\MonoidElement \in \Monoid$. The predicate
-  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
-      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
-  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
-  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
-  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
-  \begin{itemize}
-  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
-    $\Filter$, that is,
-    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
-  \item the automata all accept the same
-    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
-  \item the accepted word is mapped to
-    $\MonoidElement = \Map(\WordOf(\Path_i)) =
-    \Map(\WordOf(\Path_1))$.
-  \end{itemize}
-\end{definition}
-
-  \begin{table}[t]
-    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-      Name & \text{Rule} & Side conditions\\
-      \midrule
-
-      % EXPAND
-      \ExpandM & 
-      \inferrule
-      {
-       {\begin{array}{c}
-           \left\{
-           \FlowEq(\Automaton_i, \Filter),~~
-           \Connected(\Automaton_i, \Filter),~~
-           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
-           \right\}_{i=1}^k,
-           \\[4ex]
-          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
-          \SomeInequalities, \SomeClause
-        \end{array}}
-        }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
-      None \\[5ex]
-      \Materialise &
-      \inferrule
-      {{
-         \begin{array}{c}
-           \FlowEq(\Automaton', \Filter'),~
-           \Connected(\Automaton', \Filter'),~
-           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
-           \\\ImagePredicate{
-                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
-                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
-                     \Automaton'},~ \SomeInequalities, \SomeClause
-           }{\Map}{\Filter'}{\MonoidElement}
-           \end{array}}
-           }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
-      $\begin{array}{@{}l@{}}
-         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
-         1 \leq i < j \leq k, \\
-      \Automaton' = \Automaton_i \times \Automaton_j,\\
-      \end{array}$ \\
-      \bottomrule
-      \end{tabular}
-      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
-    \end{table}
-
-    For the calculus (\cref{tbl:rules:multi}), we first extend
-    $\Expand$ to generate flow equations and instances of $\Connected$
-    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
-    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
-    since it is needed to keep track of the currently considered
-    partial products.
-
-The rule~$\Materialise$ introduces the product of two individual
-automata~$\Automaton_i, \Automaton_j$; this step eliminates
-$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
-instead adds the product~$\Automaton_i \times \Automaton_j$, while
-also introducing the flow equations and the $\Connected$ predicate.
-
-The rule~$\Materialise$ also has to connect the newly introduced
-product~$\Automaton_i \times \Automaton_j$ to the previous automata
-$\Automaton_i, \Automaton_j$. This is done by extending the
-selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
-to fresh variables~$x_{\Transition}$:
-\begin{equation*}
-  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
-  ~~=~~
-  \begin{cases}
-    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
-    \\
-    \Filter(t) & \text{otherwise}
-  \end{cases}
-\end{equation*}
-
-The multiplicity of transitions in the product then has to be related
-to the multiplicities in the individual automata, modelled using the
-$\BindingSum$ predicate. The predicate expresses that the multiplicity
-of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
-$\Automaton_i$ has to coincide with the sum of the multiplicities of
-transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
-and similarly for $\Automaton_j$:
-%
-  $$
-  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
-  \begin{aligned}
-  & \Set{ 
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
-  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
-  & \Set{
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
-  }
-  \end{aligned}
-$$
-
-\iffalse
-Note that this definition implies that $\Filter(\Transition) =
-\Filter(\Transition')$ whenever two transitions $\Transition \in
-\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
-produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
-\times \Automaton_2}$. This corresponds to our intuition that the terms of the
-product must agree on the value they accept. As before, we implicitly map
-$\Filter$ to fresh terms for each transition in the product.
-\fi
-
-For instances of precisely one automaton, neither rule applies and we
-perform the calculus as before.
-
-\subsection{An Example}\label{sec:multiple:example}
-
-We return again to our example in \cref{sec:introduction:parikh}, where we
-compute the whole Parikh image of the product of $\AcaOrBc{}$ and
-$\SomethingCSomething$ under the constraint that there are more instances of
-letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
-element-wise addition, and $\Map$ that maps each transition to the corresponding
-increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.
-
-In the interest of space, we refer back to
-\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
-$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
-substitutions as follows. Note the expansion of the $\Sigma$ labels:
-    \begin{equation*}
-      \begin{aligned}
-        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-        % Other automaton
-        \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
-        \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
-      \end{aligned}
-    \end{equation*}
-
-In the derivation tree of \cref{fig:derivation:multi} we start as before, but
-with the product version of the $\Image$ predicate. The only possible rule here
-is $\ExpandM$, which we use to add the corresponding constraints on each
-automata of the product as we would have had in the single-automaton version. We
-perform algebraic reasoning on those equations, along with the constraint on
-$\MonoidElement$ to determine bounds on transition variables that will enable us
-to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
-finished doing so, we use \Materialise{} to compute the product of the now
-filtered automata. Further algebraic reasoning allows us to remove the final
-instance of $\Connected$. We then perform \Expand{} to get rid of the final
-$\Image$ instance, and are immediately able to $\Subsume$ the resulting
-$\Connected{}$ predicate. This, again, leaves us with a set of linear
-inequalities at the leaf that we can use to obtain a model of our final
-$\MonoidElement$ values (now a vector). This time we can read out aca, or
-$\VectorLiteral{2, 0, 1}$.
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo{
-    {\begin{array}{lll}
-      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
-      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
-    \end{array}}  
-  }
-  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
-    {\begin{array}{ll}
-      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
-      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
-    \end{array}}  
-  }
-  \infer1[\EquationReasoning{}]{
-  {  \begin{array}{ll}
-      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
-      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
-        \TransitionVar_6 + \TransitionVar_4 = 0 \\
-        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
-        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
-        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
-          \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1} 
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} \\ \land 
-      \Connected(\SomethingCSomething{}, \Filter) &
-      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array} } 
-  }
-  \infer1[\Materialise]{
-    {\begin{array}{ll}
-      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
-      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
-      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
-      \land 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} & \land
-      \Connected(\SomethingCSomething{}, \Filter) \\
-      \land 
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-  {\begin{array}{ll}
-    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-    1 - \TransitionVar_6 > 0 &
-    \land
-    \TransitionVar_6 = 0 \land 
-    \TransitionVar_4 = 0 \\
-    \land
-    \TransitionVar_3 > 0 & \land
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
-    \land
-    \VectorLiteral{
-      \TransitionVar_{7a} + \TransitionVar_{9a},
-      0,
-      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-      = \VectorLiteral{
-        2,
-        0,
-        \TransitionVar_3} & \land 
-    \Connected(\AcaOrBc{}, \Filter) \\
-    \land 
-    \Connected(\SomethingCSomething{}, \Filter) \\
-    \land 
-    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-  \end{array}}
-  }
-  \infer1[\EquationReasoning]{
-    \begin{aligned}
-      \VectorLiteral{a, b, c} = \VectorLiteral{
-          2 - 2\TransitionVar_6,
-          2\TransitionVar_4 + 3\TransitionVar_6,
-          \TransitionVar_3 + \TransitionVar_6
-        } \land
-        \\
-        \VectorLiteral{a, b, c} = \VectorLiteral{
-            \TransitionVar_{7a} + \TransitionVar_{9a},
-            \TransitionVar_{7b} + \TransitionVar_{9b},
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          } \land \\
-      \Connected(\AcaOrBc{}, \Filter) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{a, b, c}) \land a > b
-    \end{aligned}
-  }
-  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
-\end{prooftree}
-\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
-\end{figure}
-
-\subsection{Correctness of \Calculus{} for Products of Automata}
-
-Since \cref{tbl:rules:multi} only extends the existing rules of
-\cref{tbl:rules:single}, we focus on the differences compared
-to the calculus for a single automaton.
-
-%\subsubsection{\Calculus{} for products of automata terminates}
-\begin{lemma}\label{lma:multi-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the product
-  version of $\Image$ only occurs positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
-\end{lemma}
-
-\begin{proof}
-  The rule \Materialise{} can similarly only be used finitely many times, as
-  each application reduces the number of automata in the product of $\Image$ by
-  one automaton, until only one remains and \cref{lma:single-terminates} for
-  single-automaton instances applies.
-
-  This implies that also the rule~\ExpandM{} can only be applied
-  finitely often since its side condition only allows applying it once
-  per instance of an $\Image$ predicate containing a product, and only finitely
-  many instances of $\Image$ can be introduced on each branch.
-\end{proof}
-
-%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
-
-Since our calculus now includes a rule introducing new variables, the
-\Materialise{} rule, we have to slightly generalise the notion of
-solution-preservation:
-%
-\begin{lemma}\label{lma:multi-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:multi}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
-  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
-  there is an extension~$\beta' of \beta$ satisfying one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  We have to consider the two new rules in \cref{tbl:rules:multi}. The
-  result is immediate for \ExpandM{}, since this rule does not remove the
-  $\Image$ predicate from a proof goal, and the newly introduced formulas
-  are all implied by the $\Image$ predicate.
-
-  For \Materialise{}, observe that the existence of an accepting path
-  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
-  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
-  same word. The path in the product will satisfy the flow equations
-  and connectedness, and it will be related to the individual paths as
-  stipulated by the \BindingSum{} predicate.
-\end{proof}
-
-\section{Extensions}\label{sec:extensions}
-
-This section describes a number of possible extensions to the calculus. Note
-that some are partially or fully implemented in \Catra{} already. In particular,
-we have some level of support for symbolic transitions over Unicode alphabets to
-keep practical automata under a reasonable size, though we do not allow a full Boolean algebra over symbols as described e.g. in~\cite{symbolic-automata}.
-
-\subsection{Backjumping and Learning No-Goods}\label{sec:ext:backjumping}
-
-\Calculus{} can be accelerated for some instances by adding rules for
-backjumping. In particular, central connectedness constraints for an automaton
-can be learnt. In that case, when discovering that a state $\State$ of an
-automaton $\Automaton$ under a certain transition variable $\Filter$ has become
-unreachable, we can learn the clause $\Connected(\Automaton, \Filter) \land
-\sum_{\Transition \in \SeparatingCut(\Automaton, \State)} \Filter(\Transition) =
-0 \implies \sum_{\Transition' = \FromLabelTo{\State}{}{} \in
-\Transitions_{\Automaton}} = 0$. A similar rule can be used when a state becomes
-backwards-unreachable (but with a corresponding backwards cut). At the moment,
-only forward-cut learning is implemented in \Catra, and has provided a slight
-improvement in performance on some instances.
-
-The other source of clauses to learn is the \Materialise{} rule. Whenever an
-attempt to materialise a product of two ($\Filter$-filtered) automata
-$\Automaton_1 \times \Automaton_2$ produces an empty product, we can determine
-the cause of the failure with respect to the automata and their respective
-transition variables, and learn no-good combinations that can never be part of a
-model. In order to be able to do this, we need a few semantic predicates to
-record the state of the calculation, as well as a system of disambiguating
-automata, since it is possible to arrive at the same automaton by multiple
-combinations of decisions and products. In \Catra{}, we use a number of
-additional predicates to record the status of the materialisation of products,
-to separate instances of our main predicate, and to register the mapping between
-automata and their respective transition $\Filter$ terms, described in
-\cref{sec:implementation}. However, at this point only rudimentary no-good
-learning is implemented.
-
-\subsection{Symbolic Automata}\label{sec:ext:symbolic}
-
-Extending \Calculus{} to support fully symbolic automata is possible within the
-framework, depending on your choice of $\Map$. The difficulty consists in
-handling the mapping of the homomorphism $\Map$ over symbolic labels, assuming
-it maps to finitely many monoid elements. This is not always straightforward, as
-seen in the example of \cref{sec:multiple:example}. This complexity is inherent
-in computing the full Parikh image, and stems from the fact that we need to
-differentiate between the possible interpretations of the $\Sigma$ transitions
-without knowing ahead of times which ones will actually be materialied in the
-product. If, on the other hand, if our $\Map$ had been length-counting, which
-does not differentiate between values, it would have required no adaptation at
-all. With a somewhat liberal interpretation of what we are allowed to map to
-(e.g. fresh terms), it is possible to use $\Map$ to represent choice operations
-like the ones in a range label. In \Catra{}, we frontload this problem by
-requiring the user to encode their input as a Parikh automaton. For an example of how automata can be encoded, see \cref{sec:parikh-automata}.
-
-\subsection{Transducers}\label{sec:ext:transducers}
-
-Another interesting application is applying \Calculus{} to automata with
-multiple tracks, e.g. transducers. One application of such a calculus would be
-to represent replace operations and other functions on regular languages, and to
-be able to answer questions like "does this operation change the length of the
-string". The difficulty in implementing it for transducers is first to perform
-the mapping on the labels, which for both the length and Parikh cases is
-straightforward; just do the same thing in two dimensions. The more complicated
-operation is defining the product of transducers. In some cases it would
-probably be desirable to perform synchronisation (e.g. requiring overlapping
-transitions) on only some transitions, for example the first track.
-
-Implementing such a calculus is straightforward in \Calculus, since the
-definition of products was left out of the definition. All that would be needed
-is an appropriate update of the definitions of products. Simiarly, \Catra was
-written with modularity in mind, and it should be straightforward to extend both
-the input grammar and automata implementations to accommodate multi-track
-automata with arbitrary synchronisation.
-
-\subsection{Finding the Presburger representation of a homomorphic image}\label{sec:finding-the-image}
-
-To find the Presburger form of the homomorphic image efficiently, we adapt the
-quantifier elimination approach of~\cite{qe} to our problem domain. The core
-method is the same: we incrementally use \Calculus{} to find models,
-\Generalise{} the models we find into a quantifier-free Presburger formula with
-the $\MonoidElement$ as the only free variable (algorithm~\ref{alg:generalise}), add the
-negated formula as a constraint, and continue enumerating models until none
-remain. The disjunction of the generalised models we enumerated is now our
-image.
-
-\begin{algorithm}
-  \caption{$\Generalise{}(\Automaton, \Filter, a)$ will generalise a final product $\Automaton$ and a model $a$ under the homomorphism $\Map$.}\label{alg:generalise}
-  \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-  \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-  \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-
-  $\Automaton' \gets \Tuple{\Automaton_\States, \Automaton_{\InitialState}, \Automaton_\AcceptingStates, \Set{\Transition \in \Automaton_\Transitions \SuchThat a(\Filter(\Transition)) \neq 0}}$
-
-  \KwRet{\EliminateQuantifiers{$\Map(\ParikhMap(\Automaton'))$}}
-  \end{algorithm}
-
-  \begin{algorithm}
-    \DontPrintSemicolon
-    \caption{$\FindImage{}(\Automaton_1 \times \ldots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
-    \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-    \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-    \SetKwFunction{NewTheoremProver}{newTheoremProver}
-    \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-    \SetKwFunction{FreshVariable}{freshVariable}
-    \SetKwFunction{Assert}{assert}
-    \SetKwFunction{GetModel}{getModel}
-    \SetKwData{ImageVar}{image}
-
-$p \gets \NewTheoremProver{}$\;
-$\Filter(\Transition) := \FreshVariable{p}$ for every $\Transition \in \Transitions_{\Automaton}$\;
-$\MonoidElement \gets$ \FreshVariable{$p$}\;
-\Assert{$p, \exists \MonoidElement, \Filter(\Transition) \text{ for every } \Transition \in \Transitions_\Automaton \HoldsThat \ImagePredicate{\Automaton}{\Map}{\Filter}{\MonoidElement} \land \AndComp{\Transition \in \Transitions}{\Filter(\Transition) \geq 0}$}\;
-$\ImageVar \gets \bot$\;
-\While{$p$ has more models}{
-  $\Tuple{\Automaton, a} \gets \GetModel(p)$\;
-  $G \gets \Generalise{}(\Automaton, \Filter, a)$\;
-  $\ImageVar \gets G \lor \Image$\;
-  \Assert{$p, \lnot G$}\;
-  }
-    \KwRet{\ImageVar}
-    \end{algorithm}
-
-    The \GetModel{} function is nonstandard in that it returns both the model
-    and its associated automaton. Since algorithm~\ref{alg:find-image} is
-    essentially a form of quantifier elimination and therefore an internal affair
-    to the theorem prover, this should be considered fair game.
-
-    A preliminary implementation of this approach is available as part
-    of~\Catra, but has been neither optimised nor fully tested.
-
-\section{Implementation}\label{sec:implementation}
-
-We implement \Calculus{} for Parikh automata as described in
-\cref{sec:parikh-automata}. The artefact submitted along with this paper is a
-program that reads an instance file with one or more products of one or more
-Parikh automaton with transition labels defined as ranges of Unicode characters,
-along with a set of constraints on the final values of their registers expressed
-as Presburger arithmetic in a C-like syntax. We call this program
-\Catra.
-% \footnote{If you really must read it as an acronym, please read it as
-%CAtegory Theory on Register Automata, or if you object to the somewhat
-%nonstandard use of register automata and category theory, as Check Assignments
-%of The Registers Afterwards. Or alternatively if you find it all to be too much
-%of a theoretical exercise, as Can Anyone Think of A Real Application.}
-
-\Catra{} is written in Scala, with the calculus described in this paper
-implemented as a theory plug-in for the \Princess{} automated theorem
-prover~\cite{princess}, which also performs the Presburger reasoning. For
-comparison, we also provide an implementation of the baseline method
-from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
-symbolic model checker~\cite{nuxmv} to solve our constraints, and the
-approximation described in~\cite{approximate-parikh} on top of the standard
-baseline back-end. An example of an input file corresponding to our running
-example introduced in \cref{sec:motivation} can be seen in
-\cref{lst:input-example}.
-
-\Catra{} uses symbolic labels for automata. A symbolic label is defined as a
-finite range of Unicode code points. This allows succinct representation of many
-regular expression patterns such as \lstinline{(a-z).*} which would have
-otherwise required $27$~transitions. The transition for the range would be
-written \lstinline{a -> b [97, 122]}.
-
-In satisfaction mode, supported by all backends, \Catra{} tries to satisfy the
-constraints expressed by the input file, reporting \Sat{} with register
-assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
-Additionally, baseline and \Calculus{} also support generation of the Presburger
-formula describing the constraints of the input file. Baseline uses standard
-quantifier elimination, and \Calculus{} uses the method described in
-\cref{sec:finding-the-image}.
-
-Since \Princess{} does not support multiple-arity predicates like the ones we
-use in \Calculus{}, we have implemented variable-length arguments using
-additional helper predicates. These are $\Unused{}(\Automaton)$, which marks an
-automaton as unused in any product, and $\TransitionMask{}(\Automaton,
-\Transition, \Filter(\Transition))$ which associates a transition $\Transition$
-and automaton $\Automaton$ with its corresponding transition variable.
-Additionally, we associate each of our predicates with an instance variable in
-order to differentiate instances of the predicates.
-
-\subsection{Implementing the Baseline}\label{sec:implementing-baseline}
-
-We baseline using the same Presburger solver (\Princess{}), input file parser,
-and automaton implementation as \Catra. We do this in order to better analyse
-the impact of the calculus rules themselves. Using the formula of
-\eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
-successive term and add them to \Princess. We compute the product incrementally
-term by term, checking satisfiability at each step. We use a priority queue to
-select automata for each step, and order it by the number of transitions as a
-heuristic for the size of the automata. We use this heuristic to avoid computing
-large (and therefore slow) products until we have to, banking our hopes on
-computing an empty intermittent product early. The pseudocode for our
-implementation can be seen in \cref{alg:baseline}.
-
-\begin{algorithm}
-  \caption{How we implement the baseline approach}\label{alg:baseline}
-  \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
-  \KwResult{\textsc{Sat} or \textsc{Unsat}}
-  \SetKwFunction{NewTheoremProver}{newTheoremProver}
-  \SetKwFunction{NewPriorityQueue}{newPriorityQueue}
-  \SetKwFunction{Dequeue}{dequeue}
-  \SetKwFunction{Enqueue}{enqueue}
-  \SetKwFunction{Assert}{assert}
-
-  $p \gets \NewTheoremProver{}$
-
-  \Assert{$p$, $\SomeClause$}
-
-  \ForEach{$\Automaton_i$}{
-    \Assert{$p, \ParikhMap(\Automaton_i)$}
-
-    \If{$p$ is \textsc{Unsat}}{break}
-
-  }
-
-  $q \gets \NewPriorityQueue{}$
-
-  \While{$p$ not \textsc{Unsat} and $|q| > 1$}{
-    $\Automaton, \Automaton' \gets \Dequeue{q}$
-
-    \Assert{$p, \ParikhMap(\Automaton \times \Automaton')$}
-
-    \Enqueue{$q, \Automaton \times \Automaton'$}
-  }
-
-  \KwRet{$p$'s SAT status}
-
-  \end{algorithm}
-
-As an optimisation, our automata (including intermittent products) are created
-forward- and backward- reachable-minimal. Any automaton we produce only contains
-states that are both reachable from the initial state and has a path to an
-accepting state. We never perform any other minimisation on the automata for
-either backend. More complex minimisation was left out since performing
-minimisation on automata with counters is non-trivial, and minimising symbolic
-automata risks exponential blowup \cite{minimising-symbolic}.
-
-\subsection{Heuristics and search strategies}
-
-There are a number of choices left unspecified in \Calculus{} as described in
-\cref{sec:calculus,sec:multiple}. For example, the order of materialisation of
-intermediate products and the order of splitting. In this section we describe
-additional implementation details and techniques used to enhance \Catra.
-
-\subsubsection{Splitting, Materialisation, and Propagation}
-
-We order our rule applications so that we first propagate connectedness if
-possible, then perform materialisation if tractable as defined below, then
-finally resort to splitting if we must.
-
-In addition to applying \Split{} as described in \cref{tbl:rules:single} to
-randomly selected transitions, we prefer splitting to sever a strongly connected
-from the initial state. We randomly select an automaton where we can compute a
-cut between an SCC and the initial state, that is where the SCC does not contain
-the initial state and where the sum of the transition variables of the
-transitions in the cut is not known to be positive. If there are multiple such
-strongly connected components we choose one randomly. We then proceed to split
-on the sum of the transition variables of the cut as if it were a regular
-transition, e.g. its sum being zero or nonzero. In this way we drive \Calculus{}
-towards applying \Propagate{}.
-
-The implementation of the connectedness constraint is opportunistic and straightforward.
-We compute a set of dead states by performing forward and backwards reachability computations on an automaton, where we disregard any transition whose associated variable is known to be zero.
-After that we add clauses ensuring any transition variable associated with a transition starting in a dead state is zero.
-
-Product materialisation is the final piece of the puzzle. In the current
-implementation we put off computing intermediate products until at least all but
-two transition variables of one of the automata is known to be either present or
-absent. The number was chosen experimentally, and we observe a consistent trend
-towards lower numbers being better. The other automaton for the product is
-selected randomly.
-
-\subsubsection{Clause Learning}
-
-\Catra{} enables clause learning by default when using our backend, as it has
-been experimentally shown to increase the performance in aggregate (though not
-strictly). We do not currently implement all the proposed features of
-\cref{sec:ext:backjumping}, but we do implement forward-reachability cut
-learning. No sophisticated clause learning for products has been implemented.
-
-\subsubsection{Random Restarts}
-
-Finally, we perform restarts scaled by the Luby series~\cite{luby}. Experimental
-results have shown this to have a large improvement in performance, which is
-unsurprising given how many random choices we make during solving and how
-tail-heavy our problem is.
-
-\section{Evaluation}\label{sec:experiments}
-
-We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
-automata intersection problems generated by the \OstrichPlus{} string constraint
-solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
-initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
-five seconds by baseline.
-We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial.
-For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s.
-The benchmarks are run on commit~\texttt{\commit}.
-
-The benchmarks are executed in parallel using GNU Parallel~\cite{parallel},
-since they are mostly single-threaded and initial results showed negligible
-interference on performance, on \BenchmarkRig{}. We compiled the code using
-Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a
-maximum heap of~\MaxHeapSize{}. We used \Nuxmv{} version~\NuxmvVersion{} invoked
-as a subprocess for each instance. Instances were executed in batches of
-\BatchSize{}, each given a fresh JVM. We believe this represents a realistic use
-case where \Calculus{} is used to support e.g a string solver. Experiments were
-executed in random order for all backends. Each instance got a time budget
-of~\RuntimeTimeout.
-
-All runtimes are measured in wall-clock time as observed by the JVM when
-executing the instance, and exclude time spent parsing (usually far below
-\numprint{0.1}s).
-
-\subsection{Execution Time and Ability to Solve Instances}\label{sec:runtime}
-
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
-the respective back-ends could solve and with which status. A full summary of
-their outcomes is also available in \cref{tab:solve-status}. Note that we lack
-ground truth for all instances. We see that \Calculus{} generally outperforms
-\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
-instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
-kind of instance, but most of all on satisfiable instances, and \Calculus{}
-outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
-\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
-caused by a non-deterministically triggered bug in the clause learning of
-\Calculus{}, but have not had time to further investigate it. We have however
-validated all \numprint{3908} satisfying assignments from \Calculus{} with
-\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
-our implementation.
-
-Baseline performs worse on satisfiable instances because it executes a heuristic
-meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances. The heuristic is enabled since it is unrealistic to assume that we
-know in advance whether an instance is satisfiable or not, but we do know that
-unsatisfiable instances are more common.
-
-\begin{table}
-  \centering
-  \input{graphs/solved_pivot_table.tex}
-  \caption{The result of running the respective back-ends with a timeout of
-  \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
-  from the table. }\label{tab:solve-status}
-\end{table}
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
-    \caption{The division of statuses per backend.}
-    \label{fig:solve-division}
-    \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
-Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
-    \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
-    \label{fig:runtime-boxplot}
-    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline.
-The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy.
-On the other hand, lazy has  a lot of outliers.}
-  \end{subfigure}
-  \vfill
-  \begin{subfigure}[b]{0.75\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
-    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
-    \label{fig:cactus}
-    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
-Lazy solves much more instances than nuxmv.}
-  \end{subfigure}
-\end{figure}
-
-\subsubsection{Scalability}\label{sec:scaling}
-
-A cactus plot showing the number of instances solved within a given timeout for
-each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
-\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
-long-term scaling properties.
-
-\iffalse
-\subsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
-
-For baseline and \Calculus{}, \Catra{} offers the ability to find the equivalent Presburger formula representing a given instance.
-For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} we use the specially tailored approach described in \cref{sec:finding-the-image}.
-For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in \cref{sec:scaling,sec:runtime}.
-
-To make sure baseline puts up as much competition as possible, we disable
-checking intermittent satisfiability and configure \Catra{} to run in the
-maximally eager mode where the product is first computed before any
-satisifiability check is performed. We run the experiments with a timeout
-of~\ImageTimeout{}. The results of the experiment is summarised in
-\cref{fig:cactus:image} and \cref{tab:image-results}. \Fudge{We see here that
-something happens}.
-
-\begin{figure}[ht]
-  \caption{The number of instances \Catra{} was able to find the Presburger form of the image for within a given number of seconds per backend.}
-  \label{fig:cactus:image}
-\end{figure}
-\fi
-
-\subsection{Threats to Validity}
-
-The most obvious threats to validity would be poor benchmarking or poor
-implementation, e.g. if the method described in \cref{sec:calculus,sec:multiple}
-deviate from what is actually benchmarked in section \cref{sec:experiments}, or
-if the methods used for benchmarking would be unsound.  To increase the
-probability that our results are representative both in the sense of
-representing  on expected inputs and in the statistical sense for a given run of
-\Catra{} despite the use of randomness in our implementation we execute many
-experiments. Moreover, to address the issue of correctness we have validated all
-reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
-is indeed sound.
-
-Another threat to validity would be if the competition (baseline and \Nuxmv)
-would be disadvantaged in our comparison. For \Nuxmv{} we use the default
-configuration which we believe should be performant (or it should not be the
-default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
-easy for artefact reviewers. For baseline, our best argument is that the
-solution we use is close to the one found in the \Ostrich{} string solver and
-that it therefore should be realistic.
-
-The greatest threat to validity is our choice of implementation platform and
-automata library. There are some signs that product computation is inefficient,
-notably the good performance of low-threshold automata materialisation that
-prioritises computing smaller products. This means that \Calculus{} makes less
-heavy use of product computation than baseline does, instead relying more on
-\Princess{}. This situation would unfairly advantage \Calculus{} if our
-automata library was the bottleneck due to our automata library being
-unnaturally poor. We believe this is unlikely since similar performance issues
-have been reported for string solvers. Additionally, profiling suggests that
-both \Princess{}-based backends spend most of their time in \Princess{},
-suggesting that the automaton implementation is not the bottleneck.
-
-\section{Conclusion}
-
-In this paper we have introduced a calculus to compute commutating operations on
-intersections of regular languages that we call \Calculus{}. We have evaluated
-it on \NrBenchmarks{} Parikh automata intersection problems generated by the
-\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
-suite \cite{pyex} using our Parikh automata solver \Catra{}.
-
-Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
-usage and solve-time compared to the baseline approach laid out in
-\cite{generate-parikh-image} when implemented on the same underlying automated
-theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
-\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
-and generally outperforming it for timeouts under 30 seconds with its advantage
-increasing drastically for even shorter timeouts. 30 seconds would generally be
-considered a long timeout for our intended use as supporting infrastructure to a
-string constraint solver.
-
-Future investigations involve two tracks. The first one is integration into
-existing string solvers (wich \Ostrich{} being a particularly promising
-candidate due to its shared use of \Princess{}), and further adaptation to that
-use case. Closer inspection of the instances where we currently time out should
-be useful to further improve our heuristics.
-
-The second track for future improvements is the extension into other problem
-domains, including other logics, model checking problems, as well as to more
-powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata, due to our use of a full
-automated theorem prover which allows adding arbitrary constraints in addition
-to the expected Presburger formulae.
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-%% Bibliography
-\bibliography{bibliography}
-
-%% Appendix
-\appendix
-\section{Appendix}
-\input{appendix.tex}
-\end{document}

### 2023-06-22 main.tex
Work on section 4. Still WIP!
Closes: #35
Closes: Ta bort 4.1 #13

@@
 \section{Projections on Parikh Images}\label{sec:generalised}
-
-It is easy to see that the Parikh map~$\ParikhMap$ represents a
-homomorphism from the (free) non-commutative monoid~$\Strings$ to the
-(free) commutative monoid~$\Naturals^k$. As we are often not
-interested in the full Parikh image, but only in some projections of
-it. We consider \emph{arbitrary} homomorphisms
-$\Map:\: \Sigma^* \to \Monoid$, where
-$\Monoid = \left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative
-monoid. We give several
-examples of such projections on Parikh images later in this section.
-
-Observe that every homomorphism~$\Map:\: \Sigma^* \to \Monoid$ can be
-represented as the composition~$h' \circ \ParikhMap$, for some
-homomorphism~$h' : \Naturals^k \to M$. One of the insights underlying our
-approach is that it is more efficient to directly compute a
-projection on the Parikh image~$\Map(\Language)$, than to first compute the
-standard image~$\ParikhMap(\Language)$ followed by projection to
-some property of interest.
-
-%\subsection{Applications of generalised Parikh images}\label{sec:applications}
-
-\subsection{Example~1: Reasoning about String Length}
-
-One such simplifying homomorphism can express string
-length, the problem that originally motivated our study of the Parikh
-map. This mapping is relevant when solving constraints that combine
-language membership as well as word length, for instance the formula
-\begin{equation}\label{eq:stringLength}
-x \in \Language_1 \wedge y \in \Language_2 \wedge |x| > |y|
-\end{equation}
-mentioned in the introduction. To solve this formula, let
-$\Monoid = \Naturals$, and define the homomorphism~$L$ by $L(a) = 1$
-for all characters~$a \in \Alphabet$. The length of a string
-$s = s_1 \RepeatSum{\Concat} s_n$ is given by
-$L(s) = \sum L(s_i) = 1 \RepeatSum{+} 1 = n$, and to solve
-\eqref{eq:stringLength} we can instead solve the equi-satisfiable
-formula~$\alpha \in L(\Language_1) \wedge \beta \in L(\Language_2)
-\wedge \alpha > \beta$. This paper proposes efficient native
-procedures to reason about membership constraints
-like~$\alpha \in L(\Language_1)$, avoiding the computation of
-the complete image~$L(\Language_1)$.
-
-\subsection{Example~2: Generalised String Constraints with Integer
-  Datatype}\label{sec:parikh-automata}
-
-Parikh images are also applicable for deciding more general classes of
-string constraints~\cite{ostrich-plus}. Consider, for instance, a
-constraint involving the substring operation:
-\begin{equation}
-  \label{eq:substring}
-  x \in \Language_1 \wedge \; 0 \leq n \leq m \leq |x| \wedge x[n:m] \in \Language_2
-\end{equation}
-in which $x$ is a string variable, $n, m$ are integer variables, and
-$x[n:m]$ denotes the substring of $x$ starting at position~$n$ running
-until position~$m$. This constraint belongs to an expressive fragment
-of string logic that cannot be decided by most state-of-the-art string
-solvers. To illustrate the decision procedure proposed in
-\cite{ostrich-plus} for string constraints of this kind, suppose that
-$\Language_2$ is defined by the regular
-expression~$\mathtt{(ab|c)*}$. An automaton recognising this language
-is shown in \cref{fig:parikh-automata} (left).
-
-We can model \eqref{eq:substring} using the notion of a \emph{Parikh
-  automaton}~\cite{parikh-automata,expressiveness}. A Parikh automaton
-is an automaton in which transitions are labelled both with characters
-from the alphabet~$\Alphabet$, and with offset vectors defining the
-increments of a finite number of counters. This means that Parikh
-automata recognise words over an extended
-alphabet~$\Alphabet \times D$, where $D \subseteq \Naturals^d$ is a
-finite set of increment vectors (notation as
-in~\cite{expressiveness}). We use the symbols~$\pi_\Alphabet, \pi_D$
-to denote projections to the first and the second component of a
-composite letter~$(a, \Vector{v})$, respectively, and extend those
-projections to words:
-\begin{equation*}
-  \pi_\Alphabet((a_1, \Vector{v}_1) \RepeatSum{\Concat} (a_k, \Vector{v}_k))
-  ~=~ a_1 \RepeatSum{\Concat} a_k,
-  \qquad
-  \pi_D((a_1, \Vector{v}_1) \RepeatSum{\Concat} (a_k, \Vector{v}_k))
-  ~=~ \Vector{v}_1 \RepeatSum{+} \Vector{v}_k.
-\end{equation*}
-
-\begin{definition}\label{def:parikh-automata} A Parikh automaton of dimension $d
-  \geq 0$ is a pair $\Tuple{\Automaton, C}$, where
-  $C \subseteq \Naturals^d$ is a semi-linear set (or, equivalently, a
-  Presburger formula), and $\Automaton$ is a finite automaton with the
-  alphabet $\Alphabet \times D$, where $D \subseteq \Naturals^d$. We
-  say that $\Tuple{\Automaton, \varphi}$ recognises a
-  word~$w \in \Alphabet^*$ if and only if the automaton has a run
-  accepting an extended word~$w' \in (\Alphabet \times D)^*$ such that
-  $\pi_\Alphabet(w') = w$ and $\pi_D(w') \in C$.
-\end{definition}
-
-\begin{figure}[t]
-  \centering
-  \raisebox{5ex}{
-    \includegraphics[scale=0.9]{counter_automaton_pre}
-    }
-  \hfill
-  \includegraphics[scale=0.9]{counter_automaton}
-  \caption{An automaton recognising the language $\mathtt{(ab|c)*}$
-    (left), and a Parikh automaton describing the pre-image of this
-    language under the substring operation~$\cdot[n:m]$ (right). For
-    sake of presentation, the Parikh automaton contains
-    $\epsilon$-transitions, which could be eliminated in the standard
-    way.}\label{fig:parikh-automata}
-    \Description[An 2-state automaton (left) and a slightly larger automaton with counters (right).]{
-      The rightmost automaton is the starts with an initial state with a self-loop on any character that increments both registers by one, followed by an epsilon transition to a state with a self-loop on c that increments the lower register by one.
-From that state, the automaton has an epsilon-transition into the accepting state, as well as a transition on a into a state with only a transition back again, both incrementing the lower register by one.
-The accepting state has a self-loop on any character that does nothing to the registers.
-      }
-    \end{figure}
-
-    Applied to \eqref{eq:substring}, the decision procedure in
-\cite{ostrich-plus} will construct a pre-image of $\Language_2$ under
-the substring operation~$\cdot[n:m]$, and check whether this pre-image
-is consistent with the constraint~$x \in \Language_1$. Because the
-substring operation depends on the values of the integer
-variables~$n, m$, a Parikh automaton of dimension~$2$ is a suitable
-formalism to describe the pre-image, resulting
-in the automaton in Fig.~\ref{fig:parikh-automata} (right). The first
-component of the increment vectors is used to count the number of
-letters eliminated in the beginning of the string (value~$n$), while
-the second component records the beginning of the eliminated suffix
-(value~$m$).
-
-Denoting the language described by Fig.~\ref{fig:parikh-automata}
-(right) as $\Language_{pre}$, we can then replace \eqref{eq:substring}
-with an equi-satisfiable formula that no longer contains any explicit
-substring operation:
-%
-\begin{equation}
-  \label{eq:parikh-constraint}
-  p \in \Language_{pre} \wedge~
-  \pi_\Alphabet(p) \in \Language_1
-  \wedge~~ \pi_D(p) =
-  \begin{bmatrix}
-    n \\ m
-  \end{bmatrix}~~
-  \wedge 0 \leq n \leq m \leq |p|
-\end{equation}
-%
-To check the satisfiability of \eqref{eq:parikh-constraint}, we need a
-decision procedure that can process intersections of regular languages
-(in this case, of $\Language_{pre}$ and $\Language_1$, synchronising
-on $\Alphabet$), while imposing the side
-condition~$0 \leq n \leq m \leq |p|$ on the increment sum. In
-\cite{ostrich-plus}, this decision procedure turned out to be main
-bottleneck of the string solver, which was one of the motivations to
-develop the lazy algorithm proposed in this paper.
-
-\iffalse
-\begin{figure}[t]
-  \centering
-      \includegraphics[width=0.5\linewidth]{parikh_automaton}
-      \caption{The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
-      $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
-      formula containing the constraints on the final register values cannot be
-      visualised.}\label{fig:parikh-automaton}
-    \end{figure}
-\fi
+\input{generalised.tex}
 

### 2023-06-22 calculus.tex
WIP: Rename the Cut predicate and LazyParikh!
Closes: #51
Closes: #14

--- 2023-06-22 main.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2023}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\usepackage{amsmath,empheq,fancybox}
-\usepackage{paralist}
-\usepackage{url}
-\usepackage{color}
-\usepackage{textcomp,listings}
-\usepackage{array}
-\usepackage{mymacros}
-\usepackage{microtype}
-\usepackage{listings}
-\usepackage{csquotes}
-\usepackage{proof}
-\usepackage[capitalise]{cleveref}
-\usepackage{algorithm2e}      
-\usepackage{multirow}
-\usepackage{mathpartir}
-\usepackage{amsthm}
-\usepackage{numprint}
-\usepackage{ebproof}
-\usepackage[makeroom]{cancel}
-
-\usepackage{mathtools} % Bonus
-
-\usepackage{enumitem}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{example}{Example}[section]
-
-\newlist{constraints}{enumerate}{3}
-\setlist[constraints,1]{%
-    label=\sffamily{(\roman*):},
-    ref=\normalfont{Constraint (\roman*)},
-    wide,itemsep=0pt,topsep=0pt
-}    
-\crefname{constraintsi}{}{}
-\Crefname{constraintsi}{}{}
-
-\lstset{
-    columns=fullflexible,
-    showspaces=false,
-    showtabs=false,
-    breaklines=true,
-    showstringspaces=false,
-    breakatwhitespace=true,
-    escapeinside={(*@}{@*)},
-    commentstyle=\color{greencomments},
-    keywordstyle=\color{bluekeywords},
-    stringstyle=\color{redstrings},
-    numberstyle=\color{graynumbers},
-    basicstyle=\ttfamily\small,
-    framesep=12pt,
-    xleftmargin=12pt,
-    tabsize=4,
-    captionpos=b
-}
-
-\newif\ifcomments
-\commentstrue
-\newif\ifoutline
-\outlinetrue
-
-\newcommand{\contents}[1]{\ifoutline{\color{blue}
-    \begin{itemize}
-    #1
-    \end{itemize}
-  }\fi}
-
-\allowdisplaybreaks[1]
-
-\begin{document}
-
-%% Title information
-\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
-   counts in strings of a given language. Automata-based string solvers
-   frequently need to compute the Parikh image of large products (intersections)
-   of nondeterministic automata, which in many operations is both prohibitively
-   slow and memory-intensive. We contribute a novel understanding of how Parikh
-   maps can be tackled as a constraint solving problem to solve real-world
-   constraints stemming from functions on regular languages, most notably the
-   length constraint. Furthermore, we show how this formulation can be
-   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
-   prover supporting Presburger logic.  We implement \Calculus{} in a tool
-   called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving Parikh automata
-   intersection problems produced when solving standard string constraint
-   benchmarks involving string length constraints. We show that our solution
-   strictly outperforms the standard approach described
-   in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wide
-   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
-   model checker.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-  <ccs2012>
-  <concept>
-  <concept_id>10003752.10003766.10003773.10003775</concept_id>
-  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003766.10003776</concept_id>
-  <concept_desc>Theory of computation~Regular languages</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10003794</concept_id>
-  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10010124.10010138.10010142</concept_id>
-  <concept_desc>Theory of computation~Program verification</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10011192</concept_id>
-  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  </ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Quantitative automata}
-\ccsdesc[300]{Theory of computation~Regular languages}
-\ccsdesc[500]{Theory of computation~Automated reasoning}
-\ccsdesc[300]{Theory of computation~Program verification}
-\ccsdesc[300]{Theory of computation~Verification by model checking}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}\label{sec:introduction}
-\input{introduction.tex}
-
-\section{An Intuition for Our Approach}\label{sec:intuition}
-\input{example.tex}
-
-\section{Preliminaries}\label{sec:preliminaries}
-\input{preliminaries.tex}
-
-\section{Projections on Parikh Images}\label{sec:generalised}
-\input{generalised.tex}
-
-\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
-
 We start by defining our calculus, \Calculus{}, for one automaton, and only
@@
 \begin{definition}
-  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$ to state
-  $\State$, written $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of
-  transitions such that every accepting path $p$ where $\State \in
-  \StatesOf(\Path)$ contains a transition $\Transition \in C$. A cut does not
-  exist for every state; notably it never exists for $\InitialState$.
+  A \textit{dominating set of transitions}, $C$ of an automaton~$\Automaton$ of
+  a transition $\Transition$, written $\Dominates(C, \Automaton, \Transition)$,
+  is a set of transitions from $\Automaton$ such that every accepting path $p
+  \in \Automaton$ where $\Transition \in p$ contains at least one transition
+  from $C$. Notably, $\Dominates(\emptyset, \Automaton, \Transition')$ for every
+  dead transition $\Transition'$ and $\Dominates(Set{\Transition''}, \Automaton,
+  \Transition)$.
 \end{definition}
@@
   % PROPAGATE
-  \Propagate &
-  \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
-  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[4ex]
+  \Propagate & \inferrule{\Connected(\Automaton, \Filter),
+  \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C},
+  \SomeInequalities, \SomeClause, \Filter(\Transition) =
+  0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat
+  \Transition' \in C}, \SomeInequalities, \SomeClause} & if $\Dominates(C,
+  \Automaton, \Transition) t \in \Transitions_\Automaton, $ \\[4ex]
 
@@
     the automaton. By the same argument as in
-    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
+    \cite{generate-parikh-image}, this flow
     can be mapped to an accepting path~$\Path$ of $\Automaton$ such
@@
 
-  For \Propagate{}, suppose that
-  $\SeparatingCut(C, \Automaton, \State)$, which means that every
-  accepting path visiting~$q$ contains at least one of the transitions
-  in $C$. For a $\beta$ satisfying $\Connected(\Automaton, \Filter)$
-  and $\Filter(\Transition') = 0$ for $\Transition' \in C$, this means
-  that also $\val_\beta(\Filter(\Transition)) = 0$ has to hold.
+  For \Propagate{}, suppose that $\Dominates(C, \Automaton, \Transition)$, which
+  means that every accepting path containing~$\Transition$ contains at least one
+  of the transitions in $C$. For a $\beta$ satisfying $\Connected(\Automaton,
+  \Filter)$. This means that every accepting path $p$ where $\Transition \in p$
+  has at least one transition $\Transition' \in C$ such that
+  $\Filter(\Transition') = 0$. This means that also
+  $\val_\beta(\Filter(\Transition)) = 0$ has to hold since no unbroken path
+  containing $\Transition$ exists.
 
@@
 \end{proof}
-
-\section{Parikh Images from Products of Automata}\label{sec:multiple}
-
-We now generalise our calculus to natively work with intersections of
-regular languages, or equivalently products of automata. For this
-extension, we change the main predicate~$\Image$ to be indexed by a
-vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
-simplicity, we assume that the sets of states of the $k$ automata (and
-therefore also the transition sets) are pairwise disjoint. 
-\begin{definition}\label{def:multiple}
-  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
-  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
-  $\Monoid$,
-  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
-  is a transition selection function, and
-  $\MonoidElement \in \Monoid$. The predicate
-  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
-      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
-  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
-  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
-  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
-  \begin{itemize}
-  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
-    $\Filter$, that is,
-    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
-  \item the automata all accept the same
-    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
-  \item the accepted word is mapped to
-    $\MonoidElement = \Map(\WordOf(\Path_i)) =
-    \Map(\WordOf(\Path_1))$.
-  \end{itemize}
-\end{definition}
-
-  \begin{table}[t]
-    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-      Name & \text{Rule} & Side conditions\\
-      \midrule
-
-      % EXPAND
-      \ExpandM & 
-      \inferrule
-      {
-       {\begin{array}{c}
-           \left\{
-           \FlowEq(\Automaton_i, \Filter),~~
-           \Connected(\Automaton_i, \Filter),~~
-           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
-           \right\}_{i=1}^k,
-           \\[4ex]
-          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
-          \SomeInequalities, \SomeClause
-        \end{array}}
-        }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
-      None \\[5ex]
-      \Materialise &
-      \inferrule
-      {{
-         \begin{array}{c}
-           \FlowEq(\Automaton', \Filter'),~
-           \Connected(\Automaton', \Filter'),~
-           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
-           \\\ImagePredicate{
-                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
-                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
-                     \Automaton'},~ \SomeInequalities, \SomeClause
-           }{\Map}{\Filter'}{\MonoidElement}
-           \end{array}}
-           }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
-      $\begin{array}{@{}l@{}}
-         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
-         1 \leq i < j \leq k, \\
-      \Automaton' = \Automaton_i \times \Automaton_j,\\
-      \end{array}$ \\
-      \bottomrule
-      \end{tabular}
-      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
-    \end{table}
-
-    For the calculus (\cref{tbl:rules:multi}), we first extend
-    $\Expand$ to generate flow equations and instances of $\Connected$
-    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
-    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
-    since it is needed to keep track of the currently considered
-    partial products.
-
-The rule~$\Materialise$ introduces the product of two individual
-automata~$\Automaton_i, \Automaton_j$; this step eliminates
-$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
-instead adds the product~$\Automaton_i \times \Automaton_j$, while
-also introducing the flow equations and the $\Connected$ predicate.
-
-The rule~$\Materialise$ also has to connect the newly introduced
-product~$\Automaton_i \times \Automaton_j$ to the previous automata
-$\Automaton_i, \Automaton_j$. This is done by extending the
-selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
-to fresh variables~$x_{\Transition}$:
-\begin{equation*}
-  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
-  ~~=~~
-  \begin{cases}
-    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
-    \\
-    \Filter(t) & \text{otherwise}
-  \end{cases}
-\end{equation*}
-
-The multiplicity of transitions in the product then has to be related
-to the multiplicities in the individual automata, modelled using the
-$\BindingSum$ predicate. The predicate expresses that the multiplicity
-of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
-$\Automaton_i$ has to coincide with the sum of the multiplicities of
-transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
-and similarly for $\Automaton_j$:
-%
-  $$
-  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
-  \begin{aligned}
-  & \Set{ 
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
-  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
-  & \Set{
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
-  }
-  \end{aligned}
-$$
-
-\iffalse
-Note that this definition implies that $\Filter(\Transition) =
-\Filter(\Transition')$ whenever two transitions $\Transition \in
-\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
-produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
-\times \Automaton_2}$. This corresponds to our intuition that the terms of the
-product must agree on the value they accept. As before, we implicitly map
-$\Filter$ to fresh terms for each transition in the product.
-\fi
-
-For instances of precisely one automaton, neither rule applies and we
-perform the calculus as before.
-
-\subsection{An Example}\label{sec:multiple:example}
-
-We return again to our example in \cref{sec:introduction:parikh}, where we
-compute the whole Parikh image of the product of $\AcaOrBc{}$ and
-$\SomethingCSomething$ under the constraint that there are more instances of
-letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
-element-wise addition, and $\Map$ that maps each transition to the corresponding
-increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.
-
-In the interest of space, we refer back to
-\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
-$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
-substitutions as follows. Note the expansion of the $\Sigma$ labels:
-    \begin{equation*}
-      \begin{aligned}
-        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-        % Other automaton
-        \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
-        \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
-      \end{aligned}
-    \end{equation*}
-
-In the derivation tree of \cref{fig:derivation:multi} we start as before, but
-with the product version of the $\Image$ predicate. The only possible rule here
-is $\ExpandM$, which we use to add the corresponding constraints on each
-automata of the product as we would have had in the single-automaton version. We
-perform algebraic reasoning on those equations, along with the constraint on
-$\MonoidElement$ to determine bounds on transition variables that will enable us
-to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
-finished doing so, we use \Materialise{} to compute the product of the now
-filtered automata. Further algebraic reasoning allows us to remove the final
-instance of $\Connected$. We then perform \Expand{} to get rid of the final
-$\Image$ instance, and are immediately able to $\Subsume$ the resulting
-$\Connected{}$ predicate. This, again, leaves us with a set of linear
-inequalities at the leaf that we can use to obtain a model of our final
-$\MonoidElement$ values (now a vector). This time we can read out aca, or
-$\VectorLiteral{2, 0, 1}$.
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo{
-    {\begin{array}{lll}
-      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
-      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
-    \end{array}}  
-  }
-  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
-    {\begin{array}{ll}
-      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
-      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
-    \end{array}}  
-  }
-  \infer1[\EquationReasoning{}]{
-  {  \begin{array}{ll}
-      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
-      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
-        \TransitionVar_6 + \TransitionVar_4 = 0 \\
-        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
-        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
-        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
-          \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1} 
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} \\ \land 
-      \Connected(\SomethingCSomething{}, \Filter) &
-      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array} } 
-  }
-  \infer1[\Materialise]{
-    {\begin{array}{ll}
-      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
-      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
-      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
-      \land 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} & \land
-      \Connected(\SomethingCSomething{}, \Filter) \\
-      \land 
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-  {\begin{array}{ll}
-    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-    1 - \TransitionVar_6 > 0 &
-    \land
-    \TransitionVar_6 = 0 \land 
-    \TransitionVar_4 = 0 \\
-    \land
-    \TransitionVar_3 > 0 & \land
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
-    \land
-    \VectorLiteral{
-      \TransitionVar_{7a} + \TransitionVar_{9a},
-      0,
-      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-      = \VectorLiteral{
-        2,
-        0,
-        \TransitionVar_3} & \land 
-    \Connected(\AcaOrBc{}, \Filter) \\
-    \land 
-    \Connected(\SomethingCSomething{}, \Filter) \\
-    \land 
-    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-  \end{array}}
-  }
-  \infer1[\EquationReasoning]{
-    \begin{aligned}
-      \VectorLiteral{a, b, c} = \VectorLiteral{
-          2 - 2\TransitionVar_6,
-          2\TransitionVar_4 + 3\TransitionVar_6,
-          \TransitionVar_3 + \TransitionVar_6
-        } \land
-        \\
-        \VectorLiteral{a, b, c} = \VectorLiteral{
-            \TransitionVar_{7a} + \TransitionVar_{9a},
-            \TransitionVar_{7b} + \TransitionVar_{9b},
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          } \land \\
-      \Connected(\AcaOrBc{}, \Filter) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{a, b, c}) \land a > b
-    \end{aligned}
-  }
-  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
-\end{prooftree}
-\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
-\end{figure}
-
-\subsection{Correctness of \Calculus{} for Products of Automata}
-
-Since \cref{tbl:rules:multi} only extends the existing rules of
-\cref{tbl:rules:single}, we focus on the differences compared
-to the calculus for a single automaton.
-
-%\subsubsection{\Calculus{} for products of automata terminates}
-\begin{lemma}\label{lma:multi-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the product
-  version of $\Image$ only occurs positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
-\end{lemma}
-
-\begin{proof}
-  The rule \Materialise{} can similarly only be used finitely many times, as
-  each application reduces the number of automata in the product of $\Image$ by
-  one automaton, until only one remains and \cref{lma:single-terminates} for
-  single-automaton instances applies.
-
-  This implies that also the rule~\ExpandM{} can only be applied
-  finitely often since its side condition only allows applying it once
-  per instance of an $\Image$ predicate containing a product, and only finitely
-  many instances of $\Image$ can be introduced on each branch.
-\end{proof}
-
-%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
-
-Since our calculus now includes a rule introducing new variables, the
-\Materialise{} rule, we have to slightly generalise the notion of
-solution-preservation:
-%
-\begin{lemma}\label{lma:multi-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:multi}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
-  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
-  there is an extension~$\beta' of \beta$ satisfying one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  We have to consider the two new rules in \cref{tbl:rules:multi}. The
-  result is immediate for \ExpandM{}, since this rule does not remove the
-  $\Image$ predicate from a proof goal, and the newly introduced formulas
-  are all implied by the $\Image$ predicate.
-
-  For \Materialise{}, observe that the existence of an accepting path
-  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
-  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
-  same word. The path in the product will satisfy the flow equations
-  and connectedness, and it will be related to the individual paths as
-  stipulated by the \BindingSum{} predicate.
-\end{proof}
-
-\section{Extensions}\label{sec:extensions}
-
-This section describes a number of possible extensions to the calculus. Note
-that some are partially or fully implemented in \Catra{} already. In particular,
-we have some level of support for symbolic transitions over Unicode alphabets to
-keep practical automata under a reasonable size, though we do not allow a full Boolean algebra over symbols as described e.g. in~\cite{symbolic-automata}.
-
-\subsection{Backjumping and Learning No-Goods}\label{sec:ext:backjumping}
-
-\Calculus{} can be accelerated for some instances by adding rules for
-backjumping. In particular, central connectedness constraints for an automaton
-can be learnt. In that case, when discovering that a state $\State$ of an
-automaton $\Automaton$ under a certain transition variable $\Filter$ has become
-unreachable, we can learn the clause $\Connected(\Automaton, \Filter) \land
-\sum_{\Transition \in \SeparatingCut(\Automaton, \State)} \Filter(\Transition) =
-0 \implies \sum_{\Transition' = \FromLabelTo{\State}{}{} \in
-\Transitions_{\Automaton}} = 0$. A similar rule can be used when a state becomes
-backwards-unreachable (but with a corresponding backwards cut). At the moment,
-only forward-cut learning is implemented in \Catra, and has provided a slight
-improvement in performance on some instances.
-
-The other source of clauses to learn is the \Materialise{} rule. Whenever an
-attempt to materialise a product of two ($\Filter$-filtered) automata
-$\Automaton_1 \times \Automaton_2$ produces an empty product, we can determine
-the cause of the failure with respect to the automata and their respective
-transition variables, and learn no-good combinations that can never be part of a
-model. In order to be able to do this, we need a few semantic predicates to
-record the state of the calculation, as well as a system of disambiguating
-automata, since it is possible to arrive at the same automaton by multiple
-combinations of decisions and products. In \Catra{}, we use a number of
-additional predicates to record the status of the materialisation of products,
-to separate instances of our main predicate, and to register the mapping between
-automata and their respective transition $\Filter$ terms, described in
-\cref{sec:implementation}. However, at this point only rudimentary no-good
-learning is implemented.
-
-\subsection{Symbolic Automata}\label{sec:ext:symbolic}
-
-Extending \Calculus{} to support fully symbolic automata is possible within the
-framework, depending on your choice of $\Map$. The difficulty consists in
-handling the mapping of the homomorphism $\Map$ over symbolic labels, assuming
-it maps to finitely many monoid elements. This is not always straightforward, as
-seen in the example of \cref{sec:multiple:example}. This complexity is inherent
-in computing the full Parikh image, and stems from the fact that we need to
-differentiate between the possible interpretations of the $\Sigma$ transitions
-without knowing ahead of times which ones will actually be materialied in the
-product. If, on the other hand, if our $\Map$ had been length-counting, which
-does not differentiate between values, it would have required no adaptation at
-all. With a somewhat liberal interpretation of what we are allowed to map to
-(e.g. fresh terms), it is possible to use $\Map$ to represent choice operations
-like the ones in a range label. In \Catra{}, we frontload this problem by
-requiring the user to encode their input as a Parikh automaton. For an example of how automata can be encoded, see \cref{sec:parikh-automata}.
-
-\subsection{Transducers}\label{sec:ext:transducers}
-
-Another interesting application is applying \Calculus{} to automata with
-multiple tracks, e.g. transducers. One application of such a calculus would be
-to represent replace operations and other functions on regular languages, and to
-be able to answer questions like "does this operation change the length of the
-string". The difficulty in implementing it for transducers is first to perform
-the mapping on the labels, which for both the length and Parikh cases is
-straightforward; just do the same thing in two dimensions. The more complicated
-operation is defining the product of transducers. In some cases it would
-probably be desirable to perform synchronisation (e.g. requiring overlapping
-transitions) on only some transitions, for example the first track.
-
-Implementing such a calculus is straightforward in \Calculus, since the
-definition of products was left out of the definition. All that would be needed
-is an appropriate update of the definitions of products. Simiarly, \Catra was
-written with modularity in mind, and it should be straightforward to extend both
-the input grammar and automata implementations to accommodate multi-track
-automata with arbitrary synchronisation.
-
-\subsection{Finding the Presburger representation of a homomorphic image}\label{sec:finding-the-image}
-
-To find the Presburger form of the homomorphic image efficiently, we adapt the
-quantifier elimination approach of~\cite{qe} to our problem domain. The core
-method is the same: we incrementally use \Calculus{} to find models,
-\Generalise{} the models we find into a quantifier-free Presburger formula with
-the $\MonoidElement$ as the only free variable (algorithm~\ref{alg:generalise}), add the
-negated formula as a constraint, and continue enumerating models until none
-remain. The disjunction of the generalised models we enumerated is now our
-image.
-
-\begin{algorithm}
-  \caption{$\Generalise{}(\Automaton, \Filter, a)$ will generalise a final product $\Automaton$ and a model $a$ under the homomorphism $\Map$.}\label{alg:generalise}
-  \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-  \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-  \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-
-  $\Automaton' \gets \Tuple{\Automaton_\States, \Automaton_{\InitialState}, \Automaton_\AcceptingStates, \Set{\Transition \in \Automaton_\Transitions \SuchThat a(\Filter(\Transition)) \neq 0}}$
-
-  \KwRet{\EliminateQuantifiers{$\Map(\ParikhMap(\Automaton'))$}}
-  \end{algorithm}
-
-  \begin{algorithm}
-    \DontPrintSemicolon
-    \caption{$\FindImage{}(\Automaton_1 \times \ldots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
-    \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-    \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-    \SetKwFunction{NewTheoremProver}{newTheoremProver}
-    \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-    \SetKwFunction{FreshVariable}{freshVariable}
-    \SetKwFunction{Assert}{assert}
-    \SetKwFunction{GetModel}{getModel}
-    \SetKwData{ImageVar}{image}
-
-$p \gets \NewTheoremProver{}$\;
-$\Filter(\Transition) := \FreshVariable{p}$ for every $\Transition \in \Transitions_{\Automaton}$\;
-$\MonoidElement \gets$ \FreshVariable{$p$}\;
-\Assert{$p, \exists \MonoidElement, \Filter(\Transition) \text{ for every } \Transition \in \Transitions_\Automaton \HoldsThat \ImagePredicate{\Automaton}{\Map}{\Filter}{\MonoidElement} \land \AndComp{\Transition \in \Transitions}{\Filter(\Transition) \geq 0}$}\;
-$\ImageVar \gets \bot$\;
-\While{$p$ has more models}{
-  $\Tuple{\Automaton, a} \gets \GetModel(p)$\;
-  $G \gets \Generalise{}(\Automaton, \Filter, a)$\;
-  $\ImageVar \gets G \lor \Image$\;
-  \Assert{$p, \lnot G$}\;
-  }
-    \KwRet{\ImageVar}
-    \end{algorithm}
-
-    The \GetModel{} function is nonstandard in that it returns both the model
-    and its associated automaton. Since algorithm~\ref{alg:find-image} is
-    essentially a form of quantifier elimination and therefore an internal affair
-    to the theorem prover, this should be considered fair game.
-
-    A preliminary implementation of this approach is available as part
-    of~\Catra, but has been neither optimised nor fully tested.
-
-\section{Implementation}\label{sec:implementation}
-
-We implement \Calculus{} for Parikh automata as described in
-\cref{sec:parikh-automata}. The artefact submitted along with this paper is a
-program that reads an instance file with one or more products of one or more
-Parikh automaton with transition labels defined as ranges of Unicode characters,
-along with a set of constraints on the final values of their registers expressed
-as Presburger arithmetic in a C-like syntax. We call this program
-\Catra.
-% \footnote{If you really must read it as an acronym, please read it as
-%CAtegory Theory on Register Automata, or if you object to the somewhat
-%nonstandard use of register automata and category theory, as Check Assignments
-%of The Registers Afterwards. Or alternatively if you find it all to be too much
-%of a theoretical exercise, as Can Anyone Think of A Real Application.}
-
-\Catra{} is written in Scala, with the calculus described in this paper
-implemented as a theory plug-in for the \Princess{} automated theorem
-prover~\cite{princess}, which also performs the Presburger reasoning. For
-comparison, we also provide an implementation of the baseline method
-from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
-symbolic model checker~\cite{nuxmv} to solve our constraints, and the
-approximation described in~\cite{approximate-parikh} on top of the standard
-baseline back-end. An example of an input file corresponding to our running
-example introduced in \cref{sec:motivation} can be seen in
-\cref{lst:input-example}.
-
-\Catra{} uses symbolic labels for automata. A symbolic label is defined as a
-finite range of Unicode code points. This allows succinct representation of many
-regular expression patterns such as \lstinline{(a-z).*} which would have
-otherwise required $27$~transitions. The transition for the range would be
-written \lstinline{a -> b [97, 122]}.
-
-In satisfaction mode, supported by all backends, \Catra{} tries to satisfy the
-constraints expressed by the input file, reporting \Sat{} with register
-assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
-Additionally, baseline and \Calculus{} also support generation of the Presburger
-formula describing the constraints of the input file. Baseline uses standard
-quantifier elimination, and \Calculus{} uses the method described in
-\cref{sec:finding-the-image}.
-
-Since \Princess{} does not support multiple-arity predicates like the ones we
-use in \Calculus{}, we have implemented variable-length arguments using
-additional helper predicates. These are $\Unused{}(\Automaton)$, which marks an
-automaton as unused in any product, and $\TransitionMask{}(\Automaton,
-\Transition, \Filter(\Transition))$ which associates a transition $\Transition$
-and automaton $\Automaton$ with its corresponding transition variable.
-Additionally, we associate each of our predicates with an instance variable in
-order to differentiate instances of the predicates.
-
-\subsection{Implementing the Baseline}\label{sec:implementing-baseline}
-
-We baseline using the same Presburger solver (\Princess{}), input file parser,
-and automaton implementation as \Catra. We do this in order to better analyse
-the impact of the calculus rules themselves. Using the formula of
-\eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
-successive term and add them to \Princess. We compute the product incrementally
-term by term, checking satisfiability at each step. We use a priority queue to
-select automata for each step, and order it by the number of transitions as a
-heuristic for the size of the automata. We use this heuristic to avoid computing
-large (and therefore slow) products until we have to, banking our hopes on
-computing an empty intermittent product early. The pseudocode for our
-implementation can be seen in \cref{alg:baseline}.
-
-\begin{algorithm}
-  \caption{How we implement the baseline approach}\label{alg:baseline}
-  \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
-  \KwResult{\textsc{Sat} or \textsc{Unsat}}
-  \SetKwFunction{NewTheoremProver}{newTheoremProver}
-  \SetKwFunction{NewPriorityQueue}{newPriorityQueue}
-  \SetKwFunction{Dequeue}{dequeue}
-  \SetKwFunction{Enqueue}{enqueue}
-  \SetKwFunction{Assert}{assert}
-
-  $p \gets \NewTheoremProver{}$
-
-  \Assert{$p$, $\SomeClause$}
-
-  \ForEach{$\Automaton_i$}{
-    \Assert{$p, \ParikhMap(\Automaton_i)$}
-
-    \If{$p$ is \textsc{Unsat}}{break}
-
-  }
-
-  $q \gets \NewPriorityQueue{}$
-
-  \While{$p$ not \textsc{Unsat} and $|q| > 1$}{
-    $\Automaton, \Automaton' \gets \Dequeue{q}$
-
-    \Assert{$p, \ParikhMap(\Automaton \times \Automaton')$}
-
-    \Enqueue{$q, \Automaton \times \Automaton'$}
-  }
-
-  \KwRet{$p$'s SAT status}
-
-  \end{algorithm}
-
-As an optimisation, our automata (including intermittent products) are created
-forward- and backward- reachable-minimal. Any automaton we produce only contains
-states that are both reachable from the initial state and has a path to an
-accepting state. We never perform any other minimisation on the automata for
-either backend. More complex minimisation was left out since performing
-minimisation on automata with counters is non-trivial, and minimising symbolic
-automata risks exponential blowup \cite{minimising-symbolic}.
-
-\subsection{Heuristics and search strategies}
-
-There are a number of choices left unspecified in \Calculus{} as described in
-\cref{sec:calculus,sec:multiple}. For example, the order of materialisation of
-intermediate products and the order of splitting. In this section we describe
-additional implementation details and techniques used to enhance \Catra.
-
-\subsubsection{Splitting, Materialisation, and Propagation}
-
-We order our rule applications so that we first propagate connectedness if
-possible, then perform materialisation if tractable as defined below, then
-finally resort to splitting if we must.
-
-In addition to applying \Split{} as described in \cref{tbl:rules:single} to
-randomly selected transitions, we prefer splitting to sever a strongly connected
-from the initial state. We randomly select an automaton where we can compute a
-cut between an SCC and the initial state, that is where the SCC does not contain
-the initial state and where the sum of the transition variables of the
-transitions in the cut is not known to be positive. If there are multiple such
-strongly connected components we choose one randomly. We then proceed to split
-on the sum of the transition variables of the cut as if it were a regular
-transition, e.g. its sum being zero or nonzero. In this way we drive \Calculus{}
-towards applying \Propagate{}.
-
-The implementation of the connectedness constraint is opportunistic and straightforward.
-We compute a set of dead states by performing forward and backwards reachability computations on an automaton, where we disregard any transition whose associated variable is known to be zero.
-After that we add clauses ensuring any transition variable associated with a transition starting in a dead state is zero.
-
-Product materialisation is the final piece of the puzzle. In the current
-implementation we put off computing intermediate products until at least all but
-two transition variables of one of the automata is known to be either present or
-absent. The number was chosen experimentally, and we observe a consistent trend
-towards lower numbers being better. The other automaton for the product is
-selected randomly.
-
-\subsubsection{Clause Learning}
-
-\Catra{} enables clause learning by default when using our backend, as it has
-been experimentally shown to increase the performance in aggregate (though not
-strictly). We do not currently implement all the proposed features of
-\cref{sec:ext:backjumping}, but we do implement forward-reachability cut
-learning. No sophisticated clause learning for products has been implemented.
-
-\subsubsection{Random Restarts}
-
-Finally, we perform restarts scaled by the Luby series~\cite{luby}. Experimental
-results have shown this to have a large improvement in performance, which is
-unsurprising given how many random choices we make during solving and how
-tail-heavy our problem is.
-
-\section{Evaluation}\label{sec:experiments}
-
-We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
-automata intersection problems generated by the \OstrichPlus{} string constraint
-solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
-initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
-five seconds by baseline.
-We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial.
-For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s.
-The benchmarks are run on commit~\texttt{\commit}.
-
-The benchmarks are executed in parallel using GNU Parallel~\cite{parallel},
-since they are mostly single-threaded and initial results showed negligible
-interference on performance, on \BenchmarkRig{}. We compiled the code using
-Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a
-maximum heap of~\MaxHeapSize{}. We used \Nuxmv{} version~\NuxmvVersion{} invoked
-as a subprocess for each instance. Instances were executed in batches of
-\BatchSize{}, each given a fresh JVM. We believe this represents a realistic use
-case where \Calculus{} is used to support e.g a string solver. Experiments were
-executed in random order for all backends. Each instance got a time budget
-of~\RuntimeTimeout.
-
-All runtimes are measured in wall-clock time as observed by the JVM when
-executing the instance, and exclude time spent parsing (usually far below
-\numprint{0.1}s).
-
-\subsection{Execution Time and Ability to Solve Instances}\label{sec:runtime}
-
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
-the respective back-ends could solve and with which status. A full summary of
-their outcomes is also available in \cref{tab:solve-status}. Note that we lack
-ground truth for all instances. We see that \Calculus{} generally outperforms
-\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
-instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
-kind of instance, but most of all on satisfiable instances, and \Calculus{}
-outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
-\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
-caused by a non-deterministically triggered bug in the clause learning of
-\Calculus{}, but have not had time to further investigate it. We have however
-validated all \numprint{3908} satisfying assignments from \Calculus{} with
-\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
-our implementation.
-
-Baseline performs worse on satisfiable instances because it executes a heuristic
-meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances. The heuristic is enabled since it is unrealistic to assume that we
-know in advance whether an instance is satisfiable or not, but we do know that
-unsatisfiable instances are more common.
-
-\begin{table}
-  \centering
-  \input{graphs/solved_pivot_table.tex}
-  \caption{The result of running the respective back-ends with a timeout of
-  \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
-  from the table. }\label{tab:solve-status}
-\end{table}
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
-    \caption{The division of statuses per backend.}
-    \label{fig:solve-division}
-    \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
-Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
-    \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
-    \label{fig:runtime-boxplot}
-    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline.
-The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy.
-On the other hand, lazy has  a lot of outliers.}
-  \end{subfigure}
-  \vfill
-  \begin{subfigure}[b]{0.75\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
-    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
-    \label{fig:cactus}
-    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
-Lazy solves much more instances than nuxmv.}
-  \end{subfigure}
-\end{figure}
-
-\subsubsection{Scalability}\label{sec:scaling}
-
-A cactus plot showing the number of instances solved within a given timeout for
-each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
-\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
-long-term scaling properties.
-
-\iffalse
-\subsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
-
-For baseline and \Calculus{}, \Catra{} offers the ability to find the equivalent Presburger formula representing a given instance.
-For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} we use the specially tailored approach described in \cref{sec:finding-the-image}.
-For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in \cref{sec:scaling,sec:runtime}.
-
-To make sure baseline puts up as much competition as possible, we disable
-checking intermittent satisfiability and configure \Catra{} to run in the
-maximally eager mode where the product is first computed before any
-satisifiability check is performed. We run the experiments with a timeout
-of~\ImageTimeout{}. The results of the experiment is summarised in
-\cref{fig:cactus:image} and \cref{tab:image-results}. \Fudge{We see here that
-something happens}.
-
-\begin{figure}[ht]
-  \caption{The number of instances \Catra{} was able to find the Presburger form of the image for within a given number of seconds per backend.}
-  \label{fig:cactus:image}
-\end{figure}
-\fi
-
-\subsection{Threats to Validity}
-
-The most obvious threats to validity would be poor benchmarking or poor
-implementation, e.g. if the method described in \cref{sec:calculus,sec:multiple}
-deviate from what is actually benchmarked in section \cref{sec:experiments}, or
-if the methods used for benchmarking would be unsound.  To increase the
-probability that our results are representative both in the sense of
-representing  on expected inputs and in the statistical sense for a given run of
-\Catra{} despite the use of randomness in our implementation we execute many
-experiments. Moreover, to address the issue of correctness we have validated all
-reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
-is indeed sound.
-
-Another threat to validity would be if the competition (baseline and \Nuxmv)
-would be disadvantaged in our comparison. For \Nuxmv{} we use the default
-configuration which we believe should be performant (or it should not be the
-default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
-easy for artefact reviewers. For baseline, our best argument is that the
-solution we use is close to the one found in the \Ostrich{} string solver and
-that it therefore should be realistic.
-
-The greatest threat to validity is our choice of implementation platform and
-automata library. There are some signs that product computation is inefficient,
-notably the good performance of low-threshold automata materialisation that
-prioritises computing smaller products. This means that \Calculus{} makes less
-heavy use of product computation than baseline does, instead relying more on
-\Princess{}. This situation would unfairly advantage \Calculus{} if our
-automata library was the bottleneck due to our automata library being
-unnaturally poor. We believe this is unlikely since similar performance issues
-have been reported for string solvers. Additionally, profiling suggests that
-both \Princess{}-based backends spend most of their time in \Princess{},
-suggesting that the automaton implementation is not the bottleneck.
-
-\section{Conclusion}
-
-In this paper we have introduced a calculus to compute commutating operations on
-intersections of regular languages that we call \Calculus{}. We have evaluated
-it on \NrBenchmarks{} Parikh automata intersection problems generated by the
-\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
-suite \cite{pyex} using our Parikh automata solver \Catra{}.
-
-Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
-usage and solve-time compared to the baseline approach laid out in
-\cite{generate-parikh-image} when implemented on the same underlying automated
-theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
-\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
-and generally outperforming it for timeouts under 30 seconds with its advantage
-increasing drastically for even shorter timeouts. 30 seconds would generally be
-considered a long timeout for our intended use as supporting infrastructure to a
-string constraint solver.
-
-Future investigations involve two tracks. The first one is integration into
-existing string solvers (wich \Ostrich{} being a particularly promising
-candidate due to its shared use of \Princess{}), and further adaptation to that
-use case. Closer inspection of the instances where we currently time out should
-be useful to further improve our heuristics.
-
-The second track for future improvements is the extension into other problem
-domains, including other logics, model checking problems, as well as to more
-powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata, due to our use of a full
-automated theorem prover which allows adding arbitrary constraints in addition
-to the expected Presburger formulae.
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-%% Bibliography
-\bibliography{bibliography}
-
-%% Appendix
-\appendix
-\section{Appendix}
-\input{appendix.tex}
-\end{document}

### 2023-06-22 extensions.tex
WIP: Rename the Cut predicate and LazyParikh!
Closes: #51
Closes: #14

--- 2023-06-22 main.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2023}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\usepackage{amsmath,empheq,fancybox}
-\usepackage{paralist}
-\usepackage{url}
-\usepackage{color}
-\usepackage{textcomp,listings}
-\usepackage{array}
-\usepackage{mymacros}
-\usepackage{microtype}
-\usepackage{listings}
-\usepackage{csquotes}
-\usepackage{proof}
-\usepackage[capitalise]{cleveref}
-\usepackage{algorithm2e}      
-\usepackage{multirow}
-\usepackage{mathpartir}
-\usepackage{amsthm}
-\usepackage{numprint}
-\usepackage{ebproof}
-\usepackage[makeroom]{cancel}
-
-\usepackage{mathtools} % Bonus
-
-\usepackage{enumitem}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{example}{Example}[section]
-
-\newlist{constraints}{enumerate}{3}
-\setlist[constraints,1]{%
-    label=\sffamily{(\roman*):},
-    ref=\normalfont{Constraint (\roman*)},
-    wide,itemsep=0pt,topsep=0pt
-}    
-\crefname{constraintsi}{}{}
-\Crefname{constraintsi}{}{}
-
-\lstset{
-    columns=fullflexible,
-    showspaces=false,
-    showtabs=false,
-    breaklines=true,
-    showstringspaces=false,
-    breakatwhitespace=true,
-    escapeinside={(*@}{@*)},
-    commentstyle=\color{greencomments},
-    keywordstyle=\color{bluekeywords},
-    stringstyle=\color{redstrings},
-    numberstyle=\color{graynumbers},
-    basicstyle=\ttfamily\small,
-    framesep=12pt,
-    xleftmargin=12pt,
-    tabsize=4,
-    captionpos=b
-}
-
-\newif\ifcomments
-\commentstrue
-\newif\ifoutline
-\outlinetrue
-
-\newcommand{\contents}[1]{\ifoutline{\color{blue}
-    \begin{itemize}
-    #1
-    \end{itemize}
-  }\fi}
-
-\allowdisplaybreaks[1]
-
-\begin{document}
-
-%% Title information
-\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
-   counts in strings of a given language. Automata-based string solvers
-   frequently need to compute the Parikh image of large products (intersections)
-   of nondeterministic automata, which in many operations is both prohibitively
-   slow and memory-intensive. We contribute a novel understanding of how Parikh
-   maps can be tackled as a constraint solving problem to solve real-world
-   constraints stemming from functions on regular languages, most notably the
-   length constraint. Furthermore, we show how this formulation can be
-   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
-   prover supporting Presburger logic.  We implement \Calculus{} in a tool
-   called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving Parikh automata
-   intersection problems produced when solving standard string constraint
-   benchmarks involving string length constraints. We show that our solution
-   strictly outperforms the standard approach described
-   in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wide
-   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
-   model checker.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-  <ccs2012>
-  <concept>
-  <concept_id>10003752.10003766.10003773.10003775</concept_id>
-  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003766.10003776</concept_id>
-  <concept_desc>Theory of computation~Regular languages</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10003794</concept_id>
-  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10010124.10010138.10010142</concept_id>
-  <concept_desc>Theory of computation~Program verification</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10011192</concept_id>
-  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  </ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Quantitative automata}
-\ccsdesc[300]{Theory of computation~Regular languages}
-\ccsdesc[500]{Theory of computation~Automated reasoning}
-\ccsdesc[300]{Theory of computation~Program verification}
-\ccsdesc[300]{Theory of computation~Verification by model checking}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}\label{sec:introduction}
-\input{introduction.tex}
-
-\section{An Intuition for Our Approach}\label{sec:intuition}
-\input{example.tex}
-
-\section{Preliminaries}\label{sec:preliminaries}
-\input{preliminaries.tex}
-
-\section{Projections on Parikh Images}\label{sec:generalised}
-\input{generalised.tex}
-
-\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
-
-We start by defining our calculus, \Calculus{}, for one automaton, and only
-extend it to products of automata in \cref{sec:multiple}. We assume basic
-familiarity with first-order logic, Presburger arithmetic, and the classical
-sequent calculus. For reference, see e.g. \cite{Fitting96a}.
-
-First, assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
-transitions $\Transitions =
-\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. For convenience, we then
-introduce the following supporting notations:
-
-\begin{definition}
-  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
-  transitions $\Transitions$ represents a path through $\Automaton$ using
-  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
-  \State_{k+1})$ holds), passing zero or more labels $\Label_1, \ldots,
-  \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
-  \InitialState$. The end state, $\State_n$, is not necessarily accepting. If it
-  is, we say the path is also \textit{accepting}.
-  \end{definition}
-
-\begin{definition}
-  Moreover, we talk about the \textit{set of paths} of an automaton,
-  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
-  valid paths through $\Automaton$. Additionally, we use the
-  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
-  state~$\State$.
-\end{definition}
-
-\begin{definition}
-  For a path $\Path = \Tuple{\State_0 \Label_1 \State_1 \RepeatSum{,}
-  \State_n}$, its \textit{word} $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat}
-  \Label_n$ is the word read out on its labels.
-\end{definition}
-
-\begin{definition}
-  The \textit{states} of a path $\Path = \Tuple{\State_0 \Label_1 \State_1
-  \RepeatSum{,} \State_n}$, $\StatesOf(\Path) = \Set{\State_0\RepeatSum{,}
-  \State_n}$ are the states visited along $\Path$. Note that $\InitialState \in
-  \StatesOf(\Path)$ for every path since all paths start in the initial state.
-\end{definition}
-
-\begin{definition}
-  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$ to state
-  $\State$, written $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of
-  transitions such that every accepting path $p$ where $\State \in
-  \StatesOf(\Path)$ contains a transition $\Transition \in C$. A cut does not
-  exist for every state; notably it never exists for $\InitialState$.
-\end{definition}
-
-\begin{definition}
- The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
- number of times a transition $\Transition =
- \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
- $p$.
-\end{definition}
-
-We then introduce the two predicates into our calculus with the following
-definitions:
-
-\begin{definition}\label{def:single-image}
-  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
-  $\Automaton = \AutomatonTuple$, modulo some map $\Map : \Alphabet^* \to \Monoid$ to a
-  commutative monoid $\Monoid$ as described in \cref{sec:generalised}
-  and with a transition selection function
-  $\Filter:\: \Transitions \to \Naturals$ holds when
-  $\MonoidElement \in \Monoid$ is an element of the Parikh image of
-  $\Language(\Automaton)$
-  modulo $\Map$, or more formally when there is an accepting path
-  $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
-  that $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
-  for all $t \in \Transitions$, and $\MonoidElement = \Map(\WordOf(\Path))$.
-\end{definition}
-
-\begin{definition}
-  $\Connected(\Automaton, \Filter)$ for some automaton $\Automaton =
-  \AutomatonTuple$ holds when for every $\Transition =
-  \FromLabelTo{\State}{}{\State'} \in \Transitions$, $\Filter(\Transition) > 0
-  \implies \exists \Path \in \Paths(\Automaton)$ $\Filter(\Transition_i) > 0$
-  for $\Transition_i \in p$, and $\State \in \StatesOf(p)$, or in words that
-  there exists some $\Filter$-selected valid path that reaches $\Transition$'s
-  starting state, $\State$. Intuitively, it represents the condition that
-  $\Automaton$ is connected with respect to the selection function $\Filter$ for
-  every transition. It is redundant to $\Image$ by design.
-\end{definition}
-
-We present the rules of \Calculus{} for one automaton in
-\cref{tbl:rules:single}. The rules operate on sets of formulas, and
-can be interpreted as rules of a one-sided sequent calculus, in
-which all formulas are located in the antecedent~\cite{Fitting96a}.
-Each of the rules should be read and
-applied bottom-up, and relates
-premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
-conclusion~$\SomeClause$. When constructing a proof, we start from some
-root~$\SomeClause$, and then apply proof rules to the goals of the proof
-until a goal can be closed, or no more rule is applicable.
-
-A proof in which all proof goals could be closed shows that the formulas
-in the root~$\SomeClause$ of the proof are inconsistent, and do not
-have any solutions. A proof goal that could not be closed, but to which no
-more rules are applicable, gives rise to  a solution of the formulas in the
-root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic, so that a solution can be computed using
-standard algorithms~\cite{Fitting96a}.
-
-We use the convention of splitting the
-formulas into linear (in)equalities ($\SomeInequalities$)
-and other formulas ($\SomeClause$), and assume that predicates
-$\Image$ and $\Connected$ only occur positively.
- Since
-our rules operate by adding and matching linear (in)equalities in a proof goal,
-we use the shorthand of listing the matched inequalities as antecedents
-e.g., in \Propagate{}).
-The filtering function~$\Filter$ is evaluated symbolically, and can in practice
-be read as a function from transitions to $\Naturals$-valued terms
-(e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{}, described in
-\cref{sec:implementation}, $\Filter$~is a vector of fresh variables with the
-same size as~$\Transitions$.
-
-We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
-transitions of an automaton~$\Automaton$. Additionally, for an
-automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
-like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
-\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
-$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
-instance, the basis for the matched linear inequalities is
-implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
-\left(X;\MonoidOp;0_{\Monoid}\right)$ and the map into it $\Map : \Strings
-\rightarrow X$, we also allow mapping over transitions:
-$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$.
-
-We require that rules can only be applied when they add new formulas
-on every created branch (the notion of \emph{regularity} of a proof is
-required~\cite{Fitting96a}). For example, this means that \Split{} can
-only be applied to proof goals that contain neither
-$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
-never be applied to split on the same term twice on the same branch.
-This suggests a proof strategy where you \Propagate{} when you can, \Split{}
-when you must, and \Subsume{} when neither is possible anymore.
-
-\begin{table}
-\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-  Name & \text{Rule} & Side conditions\\
-  \midrule
-
-  % EXPAND
-  \Expand & 
-    \inferrule
-  {\Connected(\Automaton, \Filter) ,~ \FlowEq(\Automaton, \Filter) ,~ \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
-  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
-  None \\[4ex]
-
-  % SPLIT
-  \Split & 
-  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  if $\Transition \in \Transitions_\Automaton$ \\[4ex]
-
-  % PROPAGATE
-  \Propagate &
-  \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
-  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[4ex]
-
-  % SUBSUME
-  \Subsume &
-  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  \Split{} and \Propagate{} cannot be applied \\
-  \bottomrule
-  \end{tabular}
-  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
-\end{table}
-We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
-of existentially quantified linear inequalities with the following definition,
-where we assign fresh, existentially quantified variables to
-$\FinalStateVar_\State, \Filter(\Transition)$ for every $\State \in
-\AcceptingStates, \Transition \in \Transitions$:
-\[
-\begin{aligned}
-  \FlowEq(\Automaton, \Filter) &= \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
-  \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}\\
-  \Sink(\State) & = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
-  \In(\State, \Filter) & = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
-  \StartFlow(\State)  &= 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
-  \Out(\State, \Filter) &= \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
-\end{aligned}
-\]
-
-In addition to \cref{tbl:rules:single}, we assume the existence of a rule
-\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
-formulae, and for the elements of~$\Monoid$.
-
-The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
-$\Automaton$. It states that we are only allowed to use transitions attached to
-a reachable state, and is necessary to ensure connectedness in the presence of
-cycles in~$\Automaton$.
-
-\textsc{Expand} expands the predicate into its most basic rules; one set of
-linear equations synchronising the transitions mentioned by~$\Filter$ to the
-corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
-the standard Parikh image formulation, as described by~\FlowEq. Since
-$\Connected$ and $\Image$ are partially redundant and the difference is covered
-by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
-this sense, we split the semantics of the $\Image$~predicate into its counting
-aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
-$\Connected$).
-
-Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
-contested transition from a potential solution before concluding that it must be
-included. Intuitively, this is what guarantees our ability to make forward
-progress by eliminating paths through~$\Automaton$.
-
-A decision procedure for our predicate in a tableau-based automated theorem
-prover would start by expanding the predicate using the $\Expand{}$~rule. A
-theorem prover would perform algebraic substitution on the underlying constants
-of~$\Filter$, boiling them down to choices of branches, which depend on one
-single variable, and loop transitions. This logic corresponds to the placement
-of counters for optimally edge-profiling the CFG of a program, making up a
-minimum-spanning tree of the automaton~\cite{path-profiling}.
-
-In order to make the examples below tractable, we will assume the existence of a
-rule \EquationReasoning{} that allows us to perform standard algebraic reasoning
-on linear inequalities. This rule is not necessary for correctness or
-completeness, but shortens the examples considerably.
-
-\subsection{An Example}
-
-Starting with~$\AcaOrBc{}$, where $\Map$ is the length function, in effect
-$\Transition \mapsto 1$ for transitions, and the constraints that the length is
-odd using the same trick as in \cref{sec:motivation}, we have the
-definitions in \cref{fig:example:single:equivalences} (omitting existential
-quantifiers and $x \geq 0$ for every variable to avoid clutter).
-
-\begin{figure}[ht]
-  We define $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$ as in the following two
-  equations, and then apply \EquationReasoning{} (under the implicit assumption
-  that every RHS is $\geq 0$) to obtain the third definition:
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = \TransitionVar_1 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_2 \\
-        \Filter(\FromLabelTo{A}{c}{A}) & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = \TransitionVar_5 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \hspace{0.5cm}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        % S
-        \TransitionVar_4 &= \TransitionVar_5 + \TransitionVar_2 - 1 \\
-        % A
-        \TransitionVar_1 &= \TransitionVar_5 \\
-        % B
-        \TransitionVar_2 &= \TransitionVar_4 + \TransitionVar_6 \\
-        % F
-        \TransitionVar_5 + \TransitionVar_6 &= \FinalStateVar_1 \\
-        \FinalStateVar_1 &= 1 \\
-      \end{aligned}  
-    \end{equation*}    
-  \end{minipage}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \caption{Equivalences defining $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$
-  respectively.}\label{fig:example:single:equivalences}
-  \end{figure}
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo[]{
-{    \begin{array}{lll}
-      1 - \TransitionVar_6 = 0 & \land \TransitionVar_6 = 1 & \land  \TransitionVar_4 + 1 > 0 \\
-      \land \TransitionVar_3 = 0 &\land  2 + \TransitionVar_4 > 0 & \land k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-{    \begin{array}{lll}
-      \Connected(\AcaOrBc{}, \Filter) & \land 1 - \TransitionVar_6 = 0 \\
-      \land \TransitionVar_6 = 1 & \land \TransitionVar_4 + 1 > 0 \land \\
-      \land \TransitionVar_3 = 0 & \land 2 + \TransitionVar_4 > 0 \land \\
-      k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Propagate]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 \geq 0 \land \\
-      2 + \TransitionVar_4 \geq 0 \land \\
-      2k = 2 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{aligned}
-  }
-  % BRANCH:  = 0'
-  \hypo{
-    {\begin{array}{lll}
-      \TransitionVar_3 > 0 & \land 
-      \TransitionVar_6 = 0 \\ \land
-      2\TransitionVar_6 + \TransitionVar_4 = 0 & \land
-      1 - \TransitionVar_6 > 0\\  \land 
-      2k = 1 + \TransitionVar_3
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_3 > = 0 \land \\
-      \TransitionVar_6 = 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0\\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\Propagate{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{aligned} % BRANCH: > 0
-  } % SPLIT
-  \infer2[\Split{} $1 - \TransitionVar_6$]{ \Connected(\AcaOrBc{}, \Filter) \land 2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6 }
-  \infer1[\EquationReasoning{}]{
-    \Connected(\AcaOrBc{}, \Filter) \land
-    2k + 1 =
-    \sum\limits_{\Transition \in \Transitions} (\Transition \mapsto 1)(\Label) \cdot \Filter(\Transition)
-  }
-  \infer1[\Expand{}]{\Image{}_{\AcaOrBc{}, \Transition \mapsto 1}(\Filter, 2k + 1)}
-\end{prooftree}
-\caption{A derivation for \Calculus{} computing odd lengths in $\AcaOrBc{}$.}\label{fig:derivation:single}
-\end{figure}
-
-In \cref{fig:derivation:single}, we see how we start by expanding the predicate
-using the (simplified) flow equations. We can also see how interleaving
-reasoning on the corresponding linear equations helps the deduction. In both
-branches the reasoning is similar: we conclude that when choosing either path at
-the starting state we must avoid the other one, perform propagation based on
-that fact, use some algebraic reasoning to derive the expected forms of the
-upper and lower bounds to propagate the now disconnected transitions, and
-finally subsume, removing the $\Connected{}$ predicate when we are unable to use
-either of the rules, leaving only a set of linear equations. To obtain a solution from the
-proof, one would need to perform standard model generation on the remaining
-constraints to obtain values for the sought variable $\MonoidElement$. The full
-Parikh image is the disjunction of the constraints left at the two
-leaves after quantifier elimination.
-
-Following the values of $\Filter$, we can read out the path corresponding to our solution in the product, and with that a word.
-In our case, for the right branch we have $\Filter(\FromLabelTo{S}{b}{B}) = 0,
-\Filter(\FromLabelTo{A}{c}{A})  > 0,
-\Filter(\FromLabelTo{B}{b}{S}) = 0,
-\Filter(\FromLabelTo{S}{a}{A}) = 1 ,
-\Filter(\FromLabelTo{A}{a}{F}) = 1 ,
-\Filter(\FromLabelTo{B}{c}{F}) = 0$.
-We can read out a word along it: aca, accca, etc, and the corresponding constraint: $\exists k : \TransitionVar_3 = 2k -1 \land \TransitionVar_3 > 0 \land k \geq 0$.
-
-\subsection{Correctness of \Calculus{}}\label{sec:single:correct}
-
-Our correctness proof of \Calculus{} consists of two main parts: we
-first show that the construction of a proof always terminates, and
-then that each of the proof rules in \cref{tbl:rules:single} is
-an equivalence transformation, i.e., does not change the set of
-satisfying assignments of a formula. In combination, those two results
-immediately imply that \Calculus{} gives rise to a decision procedure.
-
-\subsubsection{\Calculus{} terminates}
-\begin{lemma}\label{lma:single-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the predicates
-  $\Image$ and $\Connected$ only occur positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:single}.
-\end{lemma}
-
-\begin{proof}
-  The rule~\Expand{} can only be applied finitely often, since each
-  application removes one $\Image$ predicate, and none of the rules
-  introduce new instances of the predicate. The rule~\Subsume{} can
-  only be applied finitely often, since it strictly decreases the
-  combined number of $\Image$ and $\Connected$ predicates in sets of
-  formulas, and none of the rules increases that number.
-
-  To show termination of \Split{} and \Propagate{}, observe that the
-  $\Filter$ in a predicate~$\Connected(\Automaton, \Filter)$ is never
-  updated on a proof branch, which means that the set of terms
-  $\Filter(t)$ for $t \in \Automaton$ on every branch is finite. Each
-  application of \Split{} and \Propagate{} adds a new
-  formula~$\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$
-  to a proof goal, which can only happen finitely often.
-\end{proof}
-
-\subsubsection{The rules in \cref{tbl:rules:single} are solution-preserving}
-
-\begin{lemma}\label{lma:single-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:single}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
-  conclusion~$\SomeClause$ if and only if it satisfies one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  This property has to be shown by analysing the possible applications
-  of each proof rule.
-
-  \Expand{} unfolds the definition of the $\Image$ predicate. To show
-  that the rule is solution-preserving we prove the equivalence of the
-  upper and lower sets of formulas:
-  \begin{itemize}
-  \item Assume that $\beta$ satisfies the conclusion, which means that
-    there is some accepting path
-    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ with
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$ and $\val_\beta(\MonoidElement) =
-    \Map(\WordOf(\Path))$. Since immediately implies that $\beta$
-    satisfies $\Connected(\Automaton, \Filter)$, since a path is
-    connected, and $\FlowEq(\Automaton, \Filter)$ since an accepting
-    path satisfies the flow equations. The
-    equation~$\MonoidElement = \sum_{\Transition \in
-      \Transitions_\Automaton}\Filter(\Transition) \cdot
-    \Map(\Transition)$ holds because of
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$.
-  \item Assume that $\beta$ satisfies the premise, which implies that
-    $\val_\beta(\Filter)$ describes a consistent, connected flow of
-    the automaton. By the same argument as in
-    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
-    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
-    that each transition~$\Transition$ occurs on $\Path$ exactly
-    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
-  \end{itemize}
-
-  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
-  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
-  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
-  be satisfied, implying the property.
-
-  For \Propagate{}, suppose that
-  $\SeparatingCut(C, \Automaton, \State)$, which means that every
-  accepting path visiting~$q$ contains at least one of the transitions
-  in $C$. For a $\beta$ satisfying $\Connected(\Automaton, \Filter)$
-  and $\Filter(\Transition') = 0$ for $\Transition' \in C$, this means
-  that also $\val_\beta(\Filter(\Transition)) = 0$ has to hold.
-
-  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
-  then a goal must contain~$\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
-  formulas in $\SomeInequalities$ are inconsistent, an application of
-  \Subsume{} is trivially solution-preserving; therefore assume that
-  $\SomeInequalities$ is consistent, which means that it contains
-  exactly one of $\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for each $\Transition$. Since
-  \Propagate{} is not applicable, the transitions~$\Transition$ with
-  $\Filter(\Transition) > 0$ must form a connect sub-graph of the
-  automaton; this means that $\Connected(\Automaton, \Filter)$ is
-  redundant as it is implied by $\SomeInequalities$.
-\end{proof}
-
-\section{Parikh Images from Products of Automata}\label{sec:multiple}
-
-We now generalise our calculus to natively work with intersections of
-regular languages, or equivalently products of automata. For this
-extension, we change the main predicate~$\Image$ to be indexed by a
-vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
-simplicity, we assume that the sets of states of the $k$ automata (and
-therefore also the transition sets) are pairwise disjoint. 
-\begin{definition}\label{def:multiple}
-  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
-  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
-  $\Monoid$,
-  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
-  is a transition selection function, and
-  $\MonoidElement \in \Monoid$. The predicate
-  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
-      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
-  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
-  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
-  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
-  \begin{itemize}
-  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
-    $\Filter$, that is,
-    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
-  \item the automata all accept the same
-    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
-  \item the accepted word is mapped to
-    $\MonoidElement = \Map(\WordOf(\Path_i)) =
-    \Map(\WordOf(\Path_1))$.
-  \end{itemize}
-\end{definition}
-
-  \begin{table}[t]
-    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-      Name & \text{Rule} & Side conditions\\
-      \midrule
-
-      % EXPAND
-      \ExpandM & 
-      \inferrule
-      {
-       {\begin{array}{c}
-           \left\{
-           \FlowEq(\Automaton_i, \Filter),~~
-           \Connected(\Automaton_i, \Filter),~~
-           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
-           \right\}_{i=1}^k,
-           \\[4ex]
-          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
-          \SomeInequalities, \SomeClause
-        \end{array}}
-        }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
-      None \\[5ex]
-      \Materialise &
-      \inferrule
-      {{
-         \begin{array}{c}
-           \FlowEq(\Automaton', \Filter'),~
-           \Connected(\Automaton', \Filter'),~
-           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
-           \\\ImagePredicate{
-                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
-                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
-                     \Automaton'},~ \SomeInequalities, \SomeClause
-           }{\Map}{\Filter'}{\MonoidElement}
-           \end{array}}
-           }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
-      $\begin{array}{@{}l@{}}
-         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
-         1 \leq i < j \leq k, \\
-      \Automaton' = \Automaton_i \times \Automaton_j,\\
-      \end{array}$ \\
-      \bottomrule
-      \end{tabular}
-      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
-    \end{table}
-
-    For the calculus (\cref{tbl:rules:multi}), we first extend
-    $\Expand$ to generate flow equations and instances of $\Connected$
-    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
-    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
-    since it is needed to keep track of the currently considered
-    partial products.
-
-The rule~$\Materialise$ introduces the product of two individual
-automata~$\Automaton_i, \Automaton_j$; this step eliminates
-$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
-instead adds the product~$\Automaton_i \times \Automaton_j$, while
-also introducing the flow equations and the $\Connected$ predicate.
-
-The rule~$\Materialise$ also has to connect the newly introduced
-product~$\Automaton_i \times \Automaton_j$ to the previous automata
-$\Automaton_i, \Automaton_j$. This is done by extending the
-selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
-to fresh variables~$x_{\Transition}$:
-\begin{equation*}
-  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
-  ~~=~~
-  \begin{cases}
-    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
-    \\
-    \Filter(t) & \text{otherwise}
-  \end{cases}
-\end{equation*}
-
-The multiplicity of transitions in the product then has to be related
-to the multiplicities in the individual automata, modelled using the
-$\BindingSum$ predicate. The predicate expresses that the multiplicity
-of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
-$\Automaton_i$ has to coincide with the sum of the multiplicities of
-transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
-and similarly for $\Automaton_j$:
-%
-  $$
-  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
-  \begin{aligned}
-  & \Set{ 
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
-  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
-  & \Set{
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
-  }
-  \end{aligned}
-$$
-
-\iffalse
-Note that this definition implies that $\Filter(\Transition) =
-\Filter(\Transition')$ whenever two transitions $\Transition \in
-\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
-produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
-\times \Automaton_2}$. This corresponds to our intuition that the terms of the
-product must agree on the value they accept. As before, we implicitly map
-$\Filter$ to fresh terms for each transition in the product.
-\fi
-
-For instances of precisely one automaton, neither rule applies and we
-perform the calculus as before.
-
-\subsection{An Example}\label{sec:multiple:example}
-
-We return again to our example in \cref{sec:introduction:parikh}, where we
-compute the whole Parikh image of the product of $\AcaOrBc{}$ and
-$\SomethingCSomething$ under the constraint that there are more instances of
-letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
-element-wise addition, and $\Map$ that maps each transition to the corresponding
-increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.
-
-In the interest of space, we refer back to
-\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
-$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
-substitutions as follows. Note the expansion of the $\Sigma$ labels:
-    \begin{equation*}
-      \begin{aligned}
-        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-        % Other automaton
-        \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
-        \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
-      \end{aligned}
-    \end{equation*}
-
-In the derivation tree of \cref{fig:derivation:multi} we start as before, but
-with the product version of the $\Image$ predicate. The only possible rule here
-is $\ExpandM$, which we use to add the corresponding constraints on each
-automata of the product as we would have had in the single-automaton version. We
-perform algebraic reasoning on those equations, along with the constraint on
-$\MonoidElement$ to determine bounds on transition variables that will enable us
-to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
-finished doing so, we use \Materialise{} to compute the product of the now
-filtered automata. Further algebraic reasoning allows us to remove the final
-instance of $\Connected$. We then perform \Expand{} to get rid of the final
-$\Image$ instance, and are immediately able to $\Subsume$ the resulting
-$\Connected{}$ predicate. This, again, leaves us with a set of linear
-inequalities at the leaf that we can use to obtain a model of our final
-$\MonoidElement$ values (now a vector). This time we can read out aca, or
-$\VectorLiteral{2, 0, 1}$.
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo{
-    {\begin{array}{lll}
-      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
-      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
-    \end{array}}  
-  }
-  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
-    {\begin{array}{ll}
-      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
-      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
-    \end{array}}  
-  }
-  \infer1[\EquationReasoning{}]{
-  {  \begin{array}{ll}
-      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
-      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
-        \TransitionVar_6 + \TransitionVar_4 = 0 \\
-        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
-        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
-        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
-          \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1} 
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} \\ \land 
-      \Connected(\SomethingCSomething{}, \Filter) &
-      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array} } 
-  }
-  \infer1[\Materialise]{
-    {\begin{array}{ll}
-      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
-      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
-      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
-      \land 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} & \land
-      \Connected(\SomethingCSomething{}, \Filter) \\
-      \land 
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-  {\begin{array}{ll}
-    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-    1 - \TransitionVar_6 > 0 &
-    \land
-    \TransitionVar_6 = 0 \land 
-    \TransitionVar_4 = 0 \\
-    \land
-    \TransitionVar_3 > 0 & \land
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
-    \land
-    \VectorLiteral{
-      \TransitionVar_{7a} + \TransitionVar_{9a},
-      0,
-      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-      = \VectorLiteral{
-        2,
-        0,
-        \TransitionVar_3} & \land 
-    \Connected(\AcaOrBc{}, \Filter) \\
-    \land 
-    \Connected(\SomethingCSomething{}, \Filter) \\
-    \land 
-    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-  \end{array}}
-  }
-  \infer1[\EquationReasoning]{
-    \begin{aligned}
-      \VectorLiteral{a, b, c} = \VectorLiteral{
-          2 - 2\TransitionVar_6,
-          2\TransitionVar_4 + 3\TransitionVar_6,
-          \TransitionVar_3 + \TransitionVar_6
-        } \land
-        \\
-        \VectorLiteral{a, b, c} = \VectorLiteral{
-            \TransitionVar_{7a} + \TransitionVar_{9a},
-            \TransitionVar_{7b} + \TransitionVar_{9b},
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          } \land \\
-      \Connected(\AcaOrBc{}, \Filter) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{a, b, c}) \land a > b
-    \end{aligned}
-  }
-  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
-\end{prooftree}
-\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
-\end{figure}
-
-\subsection{Correctness of \Calculus{} for Products of Automata}
-
-Since \cref{tbl:rules:multi} only extends the existing rules of
-\cref{tbl:rules:single}, we focus on the differences compared
-to the calculus for a single automaton.
-
-%\subsubsection{\Calculus{} for products of automata terminates}
-\begin{lemma}\label{lma:multi-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the product
-  version of $\Image$ only occurs positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
-\end{lemma}
-
-\begin{proof}
-  The rule \Materialise{} can similarly only be used finitely many times, as
-  each application reduces the number of automata in the product of $\Image$ by
-  one automaton, until only one remains and \cref{lma:single-terminates} for
-  single-automaton instances applies.
-
-  This implies that also the rule~\ExpandM{} can only be applied
-  finitely often since its side condition only allows applying it once
-  per instance of an $\Image$ predicate containing a product, and only finitely
-  many instances of $\Image$ can be introduced on each branch.
-\end{proof}
-
-%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
-
-Since our calculus now includes a rule introducing new variables, the
-\Materialise{} rule, we have to slightly generalise the notion of
-solution-preservation:
-%
-\begin{lemma}\label{lma:multi-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:multi}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
-  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
-  there is an extension~$\beta' of \beta$ satisfying one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  We have to consider the two new rules in \cref{tbl:rules:multi}. The
-  result is immediate for \ExpandM{}, since this rule does not remove the
-  $\Image$ predicate from a proof goal, and the newly introduced formulas
-  are all implied by the $\Image$ predicate.
-
-  For \Materialise{}, observe that the existence of an accepting path
-  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
-  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
-  same word. The path in the product will satisfy the flow equations
-  and connectedness, and it will be related to the individual paths as
-  stipulated by the \BindingSum{} predicate.
-\end{proof}
-
-\section{Extensions}\label{sec:extensions}
-
 This section describes a number of possible extensions to the calculus. Note
@@
 \Calculus{} can be accelerated for some instances by adding rules for
-backjumping. In particular, central connectedness constraints for an automaton
-can be learnt. In that case, when discovering that a state $\State$ of an
+backjumping. In particular, key connectedness constraints for an automaton can
+be learnt. In that case, when discovering that a transition $\Transition$ of an
 automaton $\Automaton$ under a certain transition variable $\Filter$ has become
 unreachable, we can learn the clause $\Connected(\Automaton, \Filter) \land
-\sum_{\Transition \in \SeparatingCut(\Automaton, \State)} \Filter(\Transition) =
-0 \implies \sum_{\Transition' = \FromLabelTo{\State}{}{} \in
-\Transitions_{\Automaton}} = 0$. A similar rule can be used when a state becomes
-backwards-unreachable (but with a corresponding backwards cut). At the moment,
-only forward-cut learning is implemented in \Catra, and has provided a slight
-improvement in performance on some instances.
+\sum_{\Transition' \in C} \Filter(\Transition') = 0 \implies \Transition = 0$ where
+$\Dominates(C, \Automaton, \Transition)$ for some minimal $C$. Note that this
+captures both forward (unrachable from the initial state) and
+backward (does not reach an accepting state) unreachability. Bidirectional cut
+learning is implemented in \Catra, and has provided a slight improvement in
+performance on some instances.
 
@@
     of~\Catra, but has been neither optimised nor fully tested.
-
-\section{Implementation}\label{sec:implementation}
-
-We implement \Calculus{} for Parikh automata as described in
-\cref{sec:parikh-automata}. The artefact submitted along with this paper is a
-program that reads an instance file with one or more products of one or more
-Parikh automaton with transition labels defined as ranges of Unicode characters,
-along with a set of constraints on the final values of their registers expressed
-as Presburger arithmetic in a C-like syntax. We call this program
-\Catra.
-% \footnote{If you really must read it as an acronym, please read it as
-%CAtegory Theory on Register Automata, or if you object to the somewhat
-%nonstandard use of register automata and category theory, as Check Assignments
-%of The Registers Afterwards. Or alternatively if you find it all to be too much
-%of a theoretical exercise, as Can Anyone Think of A Real Application.}
-
-\Catra{} is written in Scala, with the calculus described in this paper
-implemented as a theory plug-in for the \Princess{} automated theorem
-prover~\cite{princess}, which also performs the Presburger reasoning. For
-comparison, we also provide an implementation of the baseline method
-from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
-symbolic model checker~\cite{nuxmv} to solve our constraints, and the
-approximation described in~\cite{approximate-parikh} on top of the standard
-baseline back-end. An example of an input file corresponding to our running
-example introduced in \cref{sec:motivation} can be seen in
-\cref{lst:input-example}.
-
-\Catra{} uses symbolic labels for automata. A symbolic label is defined as a
-finite range of Unicode code points. This allows succinct representation of many
-regular expression patterns such as \lstinline{(a-z).*} which would have
-otherwise required $27$~transitions. The transition for the range would be
-written \lstinline{a -> b [97, 122]}.
-
-In satisfaction mode, supported by all backends, \Catra{} tries to satisfy the
-constraints expressed by the input file, reporting \Sat{} with register
-assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
-Additionally, baseline and \Calculus{} also support generation of the Presburger
-formula describing the constraints of the input file. Baseline uses standard
-quantifier elimination, and \Calculus{} uses the method described in
-\cref{sec:finding-the-image}.
-
-Since \Princess{} does not support multiple-arity predicates like the ones we
-use in \Calculus{}, we have implemented variable-length arguments using
-additional helper predicates. These are $\Unused{}(\Automaton)$, which marks an
-automaton as unused in any product, and $\TransitionMask{}(\Automaton,
-\Transition, \Filter(\Transition))$ which associates a transition $\Transition$
-and automaton $\Automaton$ with its corresponding transition variable.
-Additionally, we associate each of our predicates with an instance variable in
-order to differentiate instances of the predicates.
-
-\subsection{Implementing the Baseline}\label{sec:implementing-baseline}
-
-We baseline using the same Presburger solver (\Princess{}), input file parser,
-and automaton implementation as \Catra. We do this in order to better analyse
-the impact of the calculus rules themselves. Using the formula of
-\eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
-successive term and add them to \Princess. We compute the product incrementally
-term by term, checking satisfiability at each step. We use a priority queue to
-select automata for each step, and order it by the number of transitions as a
-heuristic for the size of the automata. We use this heuristic to avoid computing
-large (and therefore slow) products until we have to, banking our hopes on
-computing an empty intermittent product early. The pseudocode for our
-implementation can be seen in \cref{alg:baseline}.
-
-\begin{algorithm}
-  \caption{How we implement the baseline approach}\label{alg:baseline}
-  \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
-  \KwResult{\textsc{Sat} or \textsc{Unsat}}
-  \SetKwFunction{NewTheoremProver}{newTheoremProver}
-  \SetKwFunction{NewPriorityQueue}{newPriorityQueue}
-  \SetKwFunction{Dequeue}{dequeue}
-  \SetKwFunction{Enqueue}{enqueue}
-  \SetKwFunction{Assert}{assert}
-
-  $p \gets \NewTheoremProver{}$
-
-  \Assert{$p$, $\SomeClause$}
-
-  \ForEach{$\Automaton_i$}{
-    \Assert{$p, \ParikhMap(\Automaton_i)$}
-
-    \If{$p$ is \textsc{Unsat}}{break}
-
-  }
-
-  $q \gets \NewPriorityQueue{}$
-
-  \While{$p$ not \textsc{Unsat} and $|q| > 1$}{
-    $\Automaton, \Automaton' \gets \Dequeue{q}$
-
-    \Assert{$p, \ParikhMap(\Automaton \times \Automaton')$}
-
-    \Enqueue{$q, \Automaton \times \Automaton'$}
-  }
-
-  \KwRet{$p$'s SAT status}
-
-  \end{algorithm}
-
-As an optimisation, our automata (including intermittent products) are created
-forward- and backward- reachable-minimal. Any automaton we produce only contains
-states that are both reachable from the initial state and has a path to an
-accepting state. We never perform any other minimisation on the automata for
-either backend. More complex minimisation was left out since performing
-minimisation on automata with counters is non-trivial, and minimising symbolic
-automata risks exponential blowup \cite{minimising-symbolic}.
-
-\subsection{Heuristics and search strategies}
-
-There are a number of choices left unspecified in \Calculus{} as described in
-\cref{sec:calculus,sec:multiple}. For example, the order of materialisation of
-intermediate products and the order of splitting. In this section we describe
-additional implementation details and techniques used to enhance \Catra.
-
-\subsubsection{Splitting, Materialisation, and Propagation}
-
-We order our rule applications so that we first propagate connectedness if
-possible, then perform materialisation if tractable as defined below, then
-finally resort to splitting if we must.
-
-In addition to applying \Split{} as described in \cref{tbl:rules:single} to
-randomly selected transitions, we prefer splitting to sever a strongly connected
-from the initial state. We randomly select an automaton where we can compute a
-cut between an SCC and the initial state, that is where the SCC does not contain
-the initial state and where the sum of the transition variables of the
-transitions in the cut is not known to be positive. If there are multiple such
-strongly connected components we choose one randomly. We then proceed to split
-on the sum of the transition variables of the cut as if it were a regular
-transition, e.g. its sum being zero or nonzero. In this way we drive \Calculus{}
-towards applying \Propagate{}.
-
-The implementation of the connectedness constraint is opportunistic and straightforward.
-We compute a set of dead states by performing forward and backwards reachability computations on an automaton, where we disregard any transition whose associated variable is known to be zero.
-After that we add clauses ensuring any transition variable associated with a transition starting in a dead state is zero.
-
-Product materialisation is the final piece of the puzzle. In the current
-implementation we put off computing intermediate products until at least all but
-two transition variables of one of the automata is known to be either present or
-absent. The number was chosen experimentally, and we observe a consistent trend
-towards lower numbers being better. The other automaton for the product is
-selected randomly.
-
-\subsubsection{Clause Learning}
-
-\Catra{} enables clause learning by default when using our backend, as it has
-been experimentally shown to increase the performance in aggregate (though not
-strictly). We do not currently implement all the proposed features of
-\cref{sec:ext:backjumping}, but we do implement forward-reachability cut
-learning. No sophisticated clause learning for products has been implemented.
-
-\subsubsection{Random Restarts}
-
-Finally, we perform restarts scaled by the Luby series~\cite{luby}. Experimental
-results have shown this to have a large improvement in performance, which is
-unsurprising given how many random choices we make during solving and how
-tail-heavy our problem is.
-
-\section{Evaluation}\label{sec:experiments}
-
-We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
-automata intersection problems generated by the \OstrichPlus{} string constraint
-solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
-initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
-five seconds by baseline.
-We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial.
-For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s.
-The benchmarks are run on commit~\texttt{\commit}.
-
-The benchmarks are executed in parallel using GNU Parallel~\cite{parallel},
-since they are mostly single-threaded and initial results showed negligible
-interference on performance, on \BenchmarkRig{}. We compiled the code using
-Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a
-maximum heap of~\MaxHeapSize{}. We used \Nuxmv{} version~\NuxmvVersion{} invoked
-as a subprocess for each instance. Instances were executed in batches of
-\BatchSize{}, each given a fresh JVM. We believe this represents a realistic use
-case where \Calculus{} is used to support e.g a string solver. Experiments were
-executed in random order for all backends. Each instance got a time budget
-of~\RuntimeTimeout.
-
-All runtimes are measured in wall-clock time as observed by the JVM when
-executing the instance, and exclude time spent parsing (usually far below
-\numprint{0.1}s).
-
-\subsection{Execution Time and Ability to Solve Instances}\label{sec:runtime}
-
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
-the respective back-ends could solve and with which status. A full summary of
-their outcomes is also available in \cref{tab:solve-status}. Note that we lack
-ground truth for all instances. We see that \Calculus{} generally outperforms
-\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
-instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
-kind of instance, but most of all on satisfiable instances, and \Calculus{}
-outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
-\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
-caused by a non-deterministically triggered bug in the clause learning of
-\Calculus{}, but have not had time to further investigate it. We have however
-validated all \numprint{3908} satisfying assignments from \Calculus{} with
-\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
-our implementation.
-
-Baseline performs worse on satisfiable instances because it executes a heuristic
-meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances. The heuristic is enabled since it is unrealistic to assume that we
-know in advance whether an instance is satisfiable or not, but we do know that
-unsatisfiable instances are more common.
-
-\begin{table}
-  \centering
-  \input{graphs/solved_pivot_table.tex}
-  \caption{The result of running the respective back-ends with a timeout of
-  \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
-  from the table. }\label{tab:solve-status}
-\end{table}
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
-    \caption{The division of statuses per backend.}
-    \label{fig:solve-division}
-    \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
-Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
-    \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
-    \label{fig:runtime-boxplot}
-    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline.
-The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy.
-On the other hand, lazy has  a lot of outliers.}
-  \end{subfigure}
-  \vfill
-  \begin{subfigure}[b]{0.75\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
-    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
-    \label{fig:cactus}
-    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
-Lazy solves much more instances than nuxmv.}
-  \end{subfigure}
-\end{figure}
-
-\subsubsection{Scalability}\label{sec:scaling}
-
-A cactus plot showing the number of instances solved within a given timeout for
-each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
-\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
-long-term scaling properties.
-
-\iffalse
-\subsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
-
-For baseline and \Calculus{}, \Catra{} offers the ability to find the equivalent Presburger formula representing a given instance.
-For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} we use the specially tailored approach described in \cref{sec:finding-the-image}.
-For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in \cref{sec:scaling,sec:runtime}.
-
-To make sure baseline puts up as much competition as possible, we disable
-checking intermittent satisfiability and configure \Catra{} to run in the
-maximally eager mode where the product is first computed before any
-satisifiability check is performed. We run the experiments with a timeout
-of~\ImageTimeout{}. The results of the experiment is summarised in
-\cref{fig:cactus:image} and \cref{tab:image-results}. \Fudge{We see here that
-something happens}.
-
-\begin{figure}[ht]
-  \caption{The number of instances \Catra{} was able to find the Presburger form of the image for within a given number of seconds per backend.}
-  \label{fig:cactus:image}
-\end{figure}
-\fi
-
-\subsection{Threats to Validity}
-
-The most obvious threats to validity would be poor benchmarking or poor
-implementation, e.g. if the method described in \cref{sec:calculus,sec:multiple}
-deviate from what is actually benchmarked in section \cref{sec:experiments}, or
-if the methods used for benchmarking would be unsound.  To increase the
-probability that our results are representative both in the sense of
-representing  on expected inputs and in the statistical sense for a given run of
-\Catra{} despite the use of randomness in our implementation we execute many
-experiments. Moreover, to address the issue of correctness we have validated all
-reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
-is indeed sound.
-
-Another threat to validity would be if the competition (baseline and \Nuxmv)
-would be disadvantaged in our comparison. For \Nuxmv{} we use the default
-configuration which we believe should be performant (or it should not be the
-default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
-easy for artefact reviewers. For baseline, our best argument is that the
-solution we use is close to the one found in the \Ostrich{} string solver and
-that it therefore should be realistic.
-
-The greatest threat to validity is our choice of implementation platform and
-automata library. There are some signs that product computation is inefficient,
-notably the good performance of low-threshold automata materialisation that
-prioritises computing smaller products. This means that \Calculus{} makes less
-heavy use of product computation than baseline does, instead relying more on
-\Princess{}. This situation would unfairly advantage \Calculus{} if our
-automata library was the bottleneck due to our automata library being
-unnaturally poor. We believe this is unlikely since similar performance issues
-have been reported for string solvers. Additionally, profiling suggests that
-both \Princess{}-based backends spend most of their time in \Princess{},
-suggesting that the automaton implementation is not the bottleneck.
-
-\section{Conclusion}
-
-In this paper we have introduced a calculus to compute commutating operations on
-intersections of regular languages that we call \Calculus{}. We have evaluated
-it on \NrBenchmarks{} Parikh automata intersection problems generated by the
-\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
-suite \cite{pyex} using our Parikh automata solver \Catra{}.
-
-Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
-usage and solve-time compared to the baseline approach laid out in
-\cite{generate-parikh-image} when implemented on the same underlying automated
-theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
-\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
-and generally outperforming it for timeouts under 30 seconds with its advantage
-increasing drastically for even shorter timeouts. 30 seconds would generally be
-considered a long timeout for our intended use as supporting infrastructure to a
-string constraint solver.
-
-Future investigations involve two tracks. The first one is integration into
-existing string solvers (wich \Ostrich{} being a particularly promising
-candidate due to its shared use of \Princess{}), and further adaptation to that
-use case. Closer inspection of the instances where we currently time out should
-be useful to further improve our heuristics.
-
-The second track for future improvements is the extension into other problem
-domains, including other logics, model checking problems, as well as to more
-powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata, due to our use of a full
-automated theorem prover which allows adding arbitrary constraints in addition
-to the expected Presburger formulae.
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-%% Bibliography
-\bibliography{bibliography}
-
-%% Appendix
-\appendix
-\section{Appendix}
-\input{appendix.tex}
-\end{document}

### 2023-06-22 main.tex
WIP: Rename the Cut predicate and LazyParikh!
Closes: #51
Closes: #14

@@
 \section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
-
-We start by defining our calculus, \Calculus{}, for one automaton, and only
-extend it to products of automata in \cref{sec:multiple}. We assume basic
-familiarity with first-order logic, Presburger arithmetic, and the classical
-sequent calculus. For reference, see e.g. \cite{Fitting96a}.
-
-First, assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
-transitions $\Transitions =
-\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. For convenience, we then
-introduce the following supporting notations:
-
-\begin{definition}
-  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
-  transitions $\Transitions$ represents a path through $\Automaton$ using
-  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
-  \State_{k+1})$ holds), passing zero or more labels $\Label_1, \ldots,
-  \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
-  \InitialState$. The end state, $\State_n$, is not necessarily accepting. If it
-  is, we say the path is also \textit{accepting}.
-  \end{definition}
-
-\begin{definition}
-  Moreover, we talk about the \textit{set of paths} of an automaton,
-  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
-  valid paths through $\Automaton$. Additionally, we use the
-  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
-  state~$\State$.
-\end{definition}
-
-\begin{definition}
-  For a path $\Path = \Tuple{\State_0 \Label_1 \State_1 \RepeatSum{,}
-  \State_n}$, its \textit{word} $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat}
-  \Label_n$ is the word read out on its labels.
-\end{definition}
-
-\begin{definition}
-  The \textit{states} of a path $\Path = \Tuple{\State_0 \Label_1 \State_1
-  \RepeatSum{,} \State_n}$, $\StatesOf(\Path) = \Set{\State_0\RepeatSum{,}
-  \State_n}$ are the states visited along $\Path$. Note that $\InitialState \in
-  \StatesOf(\Path)$ for every path since all paths start in the initial state.
-\end{definition}
-
-\begin{definition}
-  A \textit{cut}, $C$ of an automaton~$\Automaton = \AutomatonTuple$ to state
-  $\State$, written $\SeparatingCut(C, \Automaton, \State)$, is a minimal set of
-  transitions such that every accepting path $p$ where $\State \in
-  \StatesOf(\Path)$ contains a transition $\Transition \in C$. A cut does not
-  exist for every state; notably it never exists for $\InitialState$.
-\end{definition}
-
-\begin{definition}
- The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
- number of times a transition $\Transition =
- \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
- $p$.
-\end{definition}
-
-We then introduce the two predicates into our calculus with the following
-definitions:
-
-\begin{definition}\label{def:single-image}
-  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
-  $\Automaton = \AutomatonTuple$, modulo some map $\Map : \Alphabet^* \to \Monoid$ to a
-  commutative monoid $\Monoid$ as described in \cref{sec:generalised}
-  and with a transition selection function
-  $\Filter:\: \Transitions \to \Naturals$ holds when
-  $\MonoidElement \in \Monoid$ is an element of the Parikh image of
-  $\Language(\Automaton)$
-  modulo $\Map$, or more formally when there is an accepting path
-  $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
-  that $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
-  for all $t \in \Transitions$, and $\MonoidElement = \Map(\WordOf(\Path))$.
-\end{definition}
-
-\begin{definition}
-  $\Connected(\Automaton, \Filter)$ for some automaton $\Automaton =
-  \AutomatonTuple$ holds when for every $\Transition =
-  \FromLabelTo{\State}{}{\State'} \in \Transitions$, $\Filter(\Transition) > 0
-  \implies \exists \Path \in \Paths(\Automaton)$ $\Filter(\Transition_i) > 0$
-  for $\Transition_i \in p$, and $\State \in \StatesOf(p)$, or in words that
-  there exists some $\Filter$-selected valid path that reaches $\Transition$'s
-  starting state, $\State$. Intuitively, it represents the condition that
-  $\Automaton$ is connected with respect to the selection function $\Filter$ for
-  every transition. It is redundant to $\Image$ by design.
-\end{definition}
-
-We present the rules of \Calculus{} for one automaton in
-\cref{tbl:rules:single}. The rules operate on sets of formulas, and
-can be interpreted as rules of a one-sided sequent calculus, in
-which all formulas are located in the antecedent~\cite{Fitting96a}.
-Each of the rules should be read and
-applied bottom-up, and relates
-premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
-conclusion~$\SomeClause$. When constructing a proof, we start from some
-root~$\SomeClause$, and then apply proof rules to the goals of the proof
-until a goal can be closed, or no more rule is applicable.
-
-A proof in which all proof goals could be closed shows that the formulas
-in the root~$\SomeClause$ of the proof are inconsistent, and do not
-have any solutions. A proof goal that could not be closed, but to which no
-more rules are applicable, gives rise to  a solution of the formulas in the
-root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic, so that a solution can be computed using
-standard algorithms~\cite{Fitting96a}.
-
-We use the convention of splitting the
-formulas into linear (in)equalities ($\SomeInequalities$)
-and other formulas ($\SomeClause$), and assume that predicates
-$\Image$ and $\Connected$ only occur positively.
- Since
-our rules operate by adding and matching linear (in)equalities in a proof goal,
-we use the shorthand of listing the matched inequalities as antecedents
-e.g., in \Propagate{}).
-The filtering function~$\Filter$ is evaluated symbolically, and can in practice
-be read as a function from transitions to $\Naturals$-valued terms
-(e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{}, described in
-\cref{sec:implementation}, $\Filter$~is a vector of fresh variables with the
-same size as~$\Transitions$.
-
-We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
-transitions of an automaton~$\Automaton$. Additionally, for an
-automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
-like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
-\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
-$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
-instance, the basis for the matched linear inequalities is
-implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
-\left(X;\MonoidOp;0_{\Monoid}\right)$ and the map into it $\Map : \Strings
-\rightarrow X$, we also allow mapping over transitions:
-$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$.
-
-We require that rules can only be applied when they add new formulas
-on every created branch (the notion of \emph{regularity} of a proof is
-required~\cite{Fitting96a}). For example, this means that \Split{} can
-only be applied to proof goals that contain neither
-$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
-never be applied to split on the same term twice on the same branch.
-This suggests a proof strategy where you \Propagate{} when you can, \Split{}
-when you must, and \Subsume{} when neither is possible anymore.
-
-\begin{table}
-\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-  Name & \text{Rule} & Side conditions\\
-  \midrule
-
-  % EXPAND
-  \Expand & 
-    \inferrule
-  {\Connected(\Automaton, \Filter) ,~ \FlowEq(\Automaton, \Filter) ,~ \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
-  {\SinglePredicateInstance, \SomeInequalities, \SomeClause} & 
-  None \\[4ex]
-
-  % SPLIT
-  \Split & 
-  \inferrule{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) = 0 \mid \Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause, \Filter(\Transition) > 0}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  if $\Transition \in \Transitions_\Automaton$ \\[4ex]
-
-  % PROPAGATE
-  \Propagate &
-  \inferrule{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause, \Filter(\Transition) = 0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat \Transition' \in C}, \SomeInequalities, \SomeClause} &
-  if $t = \FromLabelTo{\State}{}{\State'} \in \Transitions_\Automaton, \SeparatingCut(C, \Automaton, \State)$ \\[4ex]
-
-  % SUBSUME
-  \Subsume &
-  \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  \Split{} and \Propagate{} cannot be applied \\
-  \bottomrule
-  \end{tabular}
-  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
-\end{table}
-We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
-of existentially quantified linear inequalities with the following definition,
-where we assign fresh, existentially quantified variables to
-$\FinalStateVar_\State, \Filter(\Transition)$ for every $\State \in
-\AcceptingStates, \Transition \in \Transitions$:
-\[
-\begin{aligned}
-  \FlowEq(\Automaton, \Filter) &= \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
-  \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}\\
-  \Sink(\State) & = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
-  \In(\State, \Filter) & = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
-  \StartFlow(\State)  &= 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
-  \Out(\State, \Filter) &= \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
-\end{aligned}
-\]
-
-In addition to \cref{tbl:rules:single}, we assume the existence of a rule
-\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
-formulae, and for the elements of~$\Monoid$.
-
-The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
-$\Automaton$. It states that we are only allowed to use transitions attached to
-a reachable state, and is necessary to ensure connectedness in the presence of
-cycles in~$\Automaton$.
-
-\textsc{Expand} expands the predicate into its most basic rules; one set of
-linear equations synchronising the transitions mentioned by~$\Filter$ to the
-corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
-the standard Parikh image formulation, as described by~\FlowEq. Since
-$\Connected$ and $\Image$ are partially redundant and the difference is covered
-by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
-this sense, we split the semantics of the $\Image$~predicate into its counting
-aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
-$\Connected$).
-
-Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
-contested transition from a potential solution before concluding that it must be
-included. Intuitively, this is what guarantees our ability to make forward
-progress by eliminating paths through~$\Automaton$.
-
-A decision procedure for our predicate in a tableau-based automated theorem
-prover would start by expanding the predicate using the $\Expand{}$~rule. A
-theorem prover would perform algebraic substitution on the underlying constants
-of~$\Filter$, boiling them down to choices of branches, which depend on one
-single variable, and loop transitions. This logic corresponds to the placement
-of counters for optimally edge-profiling the CFG of a program, making up a
-minimum-spanning tree of the automaton~\cite{path-profiling}.
-
-In order to make the examples below tractable, we will assume the existence of a
-rule \EquationReasoning{} that allows us to perform standard algebraic reasoning
-on linear inequalities. This rule is not necessary for correctness or
-completeness, but shortens the examples considerably.
-
-\subsection{An Example}
-
-Starting with~$\AcaOrBc{}$, where $\Map$ is the length function, in effect
-$\Transition \mapsto 1$ for transitions, and the constraints that the length is
-odd using the same trick as in \cref{sec:motivation}, we have the
-definitions in \cref{fig:example:single:equivalences} (omitting existential
-quantifiers and $x \geq 0$ for every variable to avoid clutter).
-
-\begin{figure}[ht]
-  We define $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$ as in the following two
-  equations, and then apply \EquationReasoning{} (under the implicit assumption
-  that every RHS is $\geq 0$) to obtain the third definition:
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = \TransitionVar_1 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_2 \\
-        \Filter(\FromLabelTo{A}{c}{A}) & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = \TransitionVar_5 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \hspace{0.5cm}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        % S
-        \TransitionVar_4 &= \TransitionVar_5 + \TransitionVar_2 - 1 \\
-        % A
-        \TransitionVar_1 &= \TransitionVar_5 \\
-        % B
-        \TransitionVar_2 &= \TransitionVar_4 + \TransitionVar_6 \\
-        % F
-        \TransitionVar_5 + \TransitionVar_6 &= \FinalStateVar_1 \\
-        \FinalStateVar_1 &= 1 \\
-      \end{aligned}  
-    \end{equation*}    
-  \end{minipage}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \caption{Equivalences defining $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$
-  respectively.}\label{fig:example:single:equivalences}
-  \end{figure}
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo[]{
-{    \begin{array}{lll}
-      1 - \TransitionVar_6 = 0 & \land \TransitionVar_6 = 1 & \land  \TransitionVar_4 + 1 > 0 \\
-      \land \TransitionVar_3 = 0 &\land  2 + \TransitionVar_4 > 0 & \land k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-{    \begin{array}{lll}
-      \Connected(\AcaOrBc{}, \Filter) & \land 1 - \TransitionVar_6 = 0 \\
-      \land \TransitionVar_6 = 1 & \land \TransitionVar_4 + 1 > 0 \land \\
-      \land \TransitionVar_3 = 0 & \land 2 + \TransitionVar_4 > 0 \land \\
-      k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Propagate]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 \geq 0 \land \\
-      2 + \TransitionVar_4 \geq 0 \land \\
-      2k = 2 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{aligned}
-  }
-  % BRANCH:  = 0'
-  \hypo{
-    {\begin{array}{lll}
-      \TransitionVar_3 > 0 & \land 
-      \TransitionVar_6 = 0 \\ \land
-      2\TransitionVar_6 + \TransitionVar_4 = 0 & \land
-      1 - \TransitionVar_6 > 0\\  \land 
-      2k = 1 + \TransitionVar_3
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_3 > = 0 \land \\
-      \TransitionVar_6 = 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0\\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\Propagate{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
-  }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{aligned} % BRANCH: > 0
-  } % SPLIT
-  \infer2[\Split{} $1 - \TransitionVar_6$]{ \Connected(\AcaOrBc{}, \Filter) \land 2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6 }
-  \infer1[\EquationReasoning{}]{
-    \Connected(\AcaOrBc{}, \Filter) \land
-    2k + 1 =
-    \sum\limits_{\Transition \in \Transitions} (\Transition \mapsto 1)(\Label) \cdot \Filter(\Transition)
-  }
-  \infer1[\Expand{}]{\Image{}_{\AcaOrBc{}, \Transition \mapsto 1}(\Filter, 2k + 1)}
-\end{prooftree}
-\caption{A derivation for \Calculus{} computing odd lengths in $\AcaOrBc{}$.}\label{fig:derivation:single}
-\end{figure}
-
-In \cref{fig:derivation:single}, we see how we start by expanding the predicate
-using the (simplified) flow equations. We can also see how interleaving
-reasoning on the corresponding linear equations helps the deduction. In both
-branches the reasoning is similar: we conclude that when choosing either path at
-the starting state we must avoid the other one, perform propagation based on
-that fact, use some algebraic reasoning to derive the expected forms of the
-upper and lower bounds to propagate the now disconnected transitions, and
-finally subsume, removing the $\Connected{}$ predicate when we are unable to use
-either of the rules, leaving only a set of linear equations. To obtain a solution from the
-proof, one would need to perform standard model generation on the remaining
-constraints to obtain values for the sought variable $\MonoidElement$. The full
-Parikh image is the disjunction of the constraints left at the two
-leaves after quantifier elimination.
-
-Following the values of $\Filter$, we can read out the path corresponding to our solution in the product, and with that a word.
-In our case, for the right branch we have $\Filter(\FromLabelTo{S}{b}{B}) = 0,
-\Filter(\FromLabelTo{A}{c}{A})  > 0,
-\Filter(\FromLabelTo{B}{b}{S}) = 0,
-\Filter(\FromLabelTo{S}{a}{A}) = 1 ,
-\Filter(\FromLabelTo{A}{a}{F}) = 1 ,
-\Filter(\FromLabelTo{B}{c}{F}) = 0$.
-We can read out a word along it: aca, accca, etc, and the corresponding constraint: $\exists k : \TransitionVar_3 = 2k -1 \land \TransitionVar_3 > 0 \land k \geq 0$.
-
-\subsection{Correctness of \Calculus{}}\label{sec:single:correct}
-
-Our correctness proof of \Calculus{} consists of two main parts: we
-first show that the construction of a proof always terminates, and
-then that each of the proof rules in \cref{tbl:rules:single} is
-an equivalence transformation, i.e., does not change the set of
-satisfying assignments of a formula. In combination, those two results
-immediately imply that \Calculus{} gives rise to a decision procedure.
-
-\subsubsection{\Calculus{} terminates}
-\begin{lemma}\label{lma:single-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the predicates
-  $\Image$ and $\Connected$ only occur positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:single}.
-\end{lemma}
-
-\begin{proof}
-  The rule~\Expand{} can only be applied finitely often, since each
-  application removes one $\Image$ predicate, and none of the rules
-  introduce new instances of the predicate. The rule~\Subsume{} can
-  only be applied finitely often, since it strictly decreases the
-  combined number of $\Image$ and $\Connected$ predicates in sets of
-  formulas, and none of the rules increases that number.
-
-  To show termination of \Split{} and \Propagate{}, observe that the
-  $\Filter$ in a predicate~$\Connected(\Automaton, \Filter)$ is never
-  updated on a proof branch, which means that the set of terms
-  $\Filter(t)$ for $t \in \Automaton$ on every branch is finite. Each
-  application of \Split{} and \Propagate{} adds a new
-  formula~$\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$
-  to a proof goal, which can only happen finitely often.
-\end{proof}
-
-\subsubsection{The rules in \cref{tbl:rules:single} are solution-preserving}
-
-\begin{lemma}\label{lma:single-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:single}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ satisfies the
-  conclusion~$\SomeClause$ if and only if it satisfies one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  This property has to be shown by analysing the possible applications
-  of each proof rule.
-
-  \Expand{} unfolds the definition of the $\Image$ predicate. To show
-  that the rule is solution-preserving we prove the equivalence of the
-  upper and lower sets of formulas:
-  \begin{itemize}
-  \item Assume that $\beta$ satisfies the conclusion, which means that
-    there is some accepting path
-    $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ with
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$ and $\val_\beta(\MonoidElement) =
-    \Map(\WordOf(\Path))$. Since immediately implies that $\beta$
-    satisfies $\Connected(\Automaton, \Filter)$, since a path is
-    connected, and $\FlowEq(\Automaton, \Filter)$ since an accepting
-    path satisfies the flow equations. The
-    equation~$\MonoidElement = \sum_{\Transition \in
-      \Transitions_\Automaton}\Filter(\Transition) \cdot
-    \Map(\Transition)$ holds because of
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$.
-  \item Assume that $\beta$ satisfies the premise, which implies that
-    $\val_\beta(\Filter)$ describes a consistent, connected flow of
-    the automaton. By the same argument as in
-    \cite{generate-parikh-image} (\cref{sec:verma}), this flow
-    can be mapped to an accepting path~$\Path$ of $\Automaton$ such
-    that each transition~$\Transition$ occurs on $\Path$ exactly
-    $\val_\beta(\Filter(\Transition))$ times. Together with the equation
-    $\val_\beta(\Filter(\Transition)) = \TransitionCount(\Transition,
-    \Path)$, this implies that $\beta$ satisfies $\SinglePredicateInstance$.
-  \end{itemize}
-
-  In \Split{}, we make use of the fact that $\Filter(\Transition)$ is
-  $\Naturals$-valued by definition. For any $\beta$, clearly exactly
-  one of $\Filter(\Transition) = 0$ or $\Filter(\Transition) > 0$ will
-  be satisfied, implying the property.
-
-  For \Propagate{}, suppose that
-  $\SeparatingCut(C, \Automaton, \State)$, which means that every
-  accepting path visiting~$q$ contains at least one of the transitions
-  in $C$. For a $\beta$ satisfying $\Connected(\Automaton, \Filter)$
-  and $\Filter(\Transition') = 0$ for $\Transition' \in C$, this means
-  that also $\val_\beta(\Filter(\Transition)) = 0$ has to hold.
-
-  Finally, for \Subsume{}, observe that if \Split{} cannot be applied,
-  then a goal must contain~$\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for every $\Transition$. In case the
-  formulas in $\SomeInequalities$ are inconsistent, an application of
-  \Subsume{} is trivially solution-preserving; therefore assume that
-  $\SomeInequalities$ is consistent, which means that it contains
-  exactly one of $\Filter(\Transition) = 0$ or
-  $\Filter(\Transition) > 0$ for each $\Transition$. Since
-  \Propagate{} is not applicable, the transitions~$\Transition$ with
-  $\Filter(\Transition) > 0$ must form a connect sub-graph of the
-  automaton; this means that $\Connected(\Automaton, \Filter)$ is
-  redundant as it is implied by $\SomeInequalities$.
-\end{proof}
+\input{calculus.tex}
 
@@
 \section{Extensions}\label{sec:extensions}
-
-This section describes a number of possible extensions to the calculus. Note
-that some are partially or fully implemented in \Catra{} already. In particular,
-we have some level of support for symbolic transitions over Unicode alphabets to
-keep practical automata under a reasonable size, though we do not allow a full Boolean algebra over symbols as described e.g. in~\cite{symbolic-automata}.
-
-\subsection{Backjumping and Learning No-Goods}\label{sec:ext:backjumping}
-
-\Calculus{} can be accelerated for some instances by adding rules for
-backjumping. In particular, central connectedness constraints for an automaton
-can be learnt. In that case, when discovering that a state $\State$ of an
-automaton $\Automaton$ under a certain transition variable $\Filter$ has become
-unreachable, we can learn the clause $\Connected(\Automaton, \Filter) \land
-\sum_{\Transition \in \SeparatingCut(\Automaton, \State)} \Filter(\Transition) =
-0 \implies \sum_{\Transition' = \FromLabelTo{\State}{}{} \in
-\Transitions_{\Automaton}} = 0$. A similar rule can be used when a state becomes
-backwards-unreachable (but with a corresponding backwards cut). At the moment,
-only forward-cut learning is implemented in \Catra, and has provided a slight
-improvement in performance on some instances.
-
-The other source of clauses to learn is the \Materialise{} rule. Whenever an
-attempt to materialise a product of two ($\Filter$-filtered) automata
-$\Automaton_1 \times \Automaton_2$ produces an empty product, we can determine
-the cause of the failure with respect to the automata and their respective
-transition variables, and learn no-good combinations that can never be part of a
-model. In order to be able to do this, we need a few semantic predicates to
-record the state of the calculation, as well as a system of disambiguating
-automata, since it is possible to arrive at the same automaton by multiple
-combinations of decisions and products. In \Catra{}, we use a number of
-additional predicates to record the status of the materialisation of products,
-to separate instances of our main predicate, and to register the mapping between
-automata and their respective transition $\Filter$ terms, described in
-\cref{sec:implementation}. However, at this point only rudimentary no-good
-learning is implemented.
-
-\subsection{Symbolic Automata}\label{sec:ext:symbolic}
-
-Extending \Calculus{} to support fully symbolic automata is possible within the
-framework, depending on your choice of $\Map$. The difficulty consists in
-handling the mapping of the homomorphism $\Map$ over symbolic labels, assuming
-it maps to finitely many monoid elements. This is not always straightforward, as
-seen in the example of \cref{sec:multiple:example}. This complexity is inherent
-in computing the full Parikh image, and stems from the fact that we need to
-differentiate between the possible interpretations of the $\Sigma$ transitions
-without knowing ahead of times which ones will actually be materialied in the
-product. If, on the other hand, if our $\Map$ had been length-counting, which
-does not differentiate between values, it would have required no adaptation at
-all. With a somewhat liberal interpretation of what we are allowed to map to
-(e.g. fresh terms), it is possible to use $\Map$ to represent choice operations
-like the ones in a range label. In \Catra{}, we frontload this problem by
-requiring the user to encode their input as a Parikh automaton. For an example of how automata can be encoded, see \cref{sec:parikh-automata}.
-
-\subsection{Transducers}\label{sec:ext:transducers}
-
-Another interesting application is applying \Calculus{} to automata with
-multiple tracks, e.g. transducers. One application of such a calculus would be
-to represent replace operations and other functions on regular languages, and to
-be able to answer questions like "does this operation change the length of the
-string". The difficulty in implementing it for transducers is first to perform
-the mapping on the labels, which for both the length and Parikh cases is
-straightforward; just do the same thing in two dimensions. The more complicated
-operation is defining the product of transducers. In some cases it would
-probably be desirable to perform synchronisation (e.g. requiring overlapping
-transitions) on only some transitions, for example the first track.
-
-Implementing such a calculus is straightforward in \Calculus, since the
-definition of products was left out of the definition. All that would be needed
-is an appropriate update of the definitions of products. Simiarly, \Catra was
-written with modularity in mind, and it should be straightforward to extend both
-the input grammar and automata implementations to accommodate multi-track
-automata with arbitrary synchronisation.
-
-\subsection{Finding the Presburger representation of a homomorphic image}\label{sec:finding-the-image}
-
-To find the Presburger form of the homomorphic image efficiently, we adapt the
-quantifier elimination approach of~\cite{qe} to our problem domain. The core
-method is the same: we incrementally use \Calculus{} to find models,
-\Generalise{} the models we find into a quantifier-free Presburger formula with
-the $\MonoidElement$ as the only free variable (algorithm~\ref{alg:generalise}), add the
-negated formula as a constraint, and continue enumerating models until none
-remain. The disjunction of the generalised models we enumerated is now our
-image.
-
-\begin{algorithm}
-  \caption{$\Generalise{}(\Automaton, \Filter, a)$ will generalise a final product $\Automaton$ and a model $a$ under the homomorphism $\Map$.}\label{alg:generalise}
-  \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-  \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-  \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-
-  $\Automaton' \gets \Tuple{\Automaton_\States, \Automaton_{\InitialState}, \Automaton_\AcceptingStates, \Set{\Transition \in \Automaton_\Transitions \SuchThat a(\Filter(\Transition)) \neq 0}}$
-
-  \KwRet{\EliminateQuantifiers{$\Map(\ParikhMap(\Automaton'))$}}
-  \end{algorithm}
-
-  \begin{algorithm}
-    \DontPrintSemicolon
-    \caption{$\FindImage{}(\Automaton_1 \times \ldots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
-    \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}
-    \KwResult{a quantifier-free Presburger formula $P$ representing a partial $\Map$-homomorphic image}
-    \SetKwFunction{NewTheoremProver}{newTheoremProver}
-    \SetKwFunction{EliminateQuantifiers}{eliminateQuantifiers}
-    \SetKwFunction{FreshVariable}{freshVariable}
-    \SetKwFunction{Assert}{assert}
-    \SetKwFunction{GetModel}{getModel}
-    \SetKwData{ImageVar}{image}
-
-$p \gets \NewTheoremProver{}$\;
-$\Filter(\Transition) := \FreshVariable{p}$ for every $\Transition \in \Transitions_{\Automaton}$\;
-$\MonoidElement \gets$ \FreshVariable{$p$}\;
-\Assert{$p, \exists \MonoidElement, \Filter(\Transition) \text{ for every } \Transition \in \Transitions_\Automaton \HoldsThat \ImagePredicate{\Automaton}{\Map}{\Filter}{\MonoidElement} \land \AndComp{\Transition \in \Transitions}{\Filter(\Transition) \geq 0}$}\;
-$\ImageVar \gets \bot$\;
-\While{$p$ has more models}{
-  $\Tuple{\Automaton, a} \gets \GetModel(p)$\;
-  $G \gets \Generalise{}(\Automaton, \Filter, a)$\;
-  $\ImageVar \gets G \lor \Image$\;
-  \Assert{$p, \lnot G$}\;
-  }
-    \KwRet{\ImageVar}
-    \end{algorithm}
-
-    The \GetModel{} function is nonstandard in that it returns both the model
-    and its associated automaton. Since algorithm~\ref{alg:find-image} is
-    essentially a form of quantifier elimination and therefore an internal affair
-    to the theorem prover, this should be considered fair game.
-
-    A preliminary implementation of this approach is available as part
-    of~\Catra, but has been neither optimised nor fully tested.
+\input{extensions.tex}
 

### 2023-06-22 preliminaries.tex
Work on section 4. Still WIP!
Closes: #35
Closes: Ta bort 4.1 #13
WIP: Rename the Cut predicate and LazyParikh!
Closes: #51
Closes: #14

@@
 \subsection{Monoids}
+
+% @Philipp: reviewer comment #54 complains that this is basic, are they right?
 A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
@@
 notation $\#l(w)$ to talk about an individual letter $l$ in a word $w$. For
-example, for the Parikh vector above, we would have $\CountOf{a} = 1$. You have
-already seen this in \cref{sec:introduction:parikh}.
+example, for the Parikh vector above, we would have $\CountOf{a} = 1$.
 

### 2023-06-22 example.tex
WIP: Rename the Cut predicate and LazyParikh!
Closes: #51
Closes: #14
Tweak figures
Much tighter integration between sections

@@
 automata, then break down how the constraints will be handled by \Calculus{}.
+This example will recur in a more formal fashion in \cref{sec:generalised},
+which can be read in parallel if (even) more detail is desired.
 
@@
 (NFA) with a set of integer registers (Parikh registers) that are incremented
-(or decremented) at each transition.
+(or decremented) at each transition. A formal definition can be found in
+\cref{def:parikh-automata}.
 
@@
 (\cref{fig:aut_a}) is the Parikh automata obtained by applying the encoding of
-$\Substring$ described in \cite{ostrich-plus}. That description can then be used
-to obtain values for the subsstring offsets $i, n$. The corresponding automata
-can be found in \cref{fig:examples}.
+$\Substring$ described in \cite{ostrich-plus} to
+$\Language(\Regex{c\RegexOr{}dd})$. That description can then be used to obtain
+values for the substring offsets $i, n$. The corresponding automata can be found
+in \cref{fig:examples}.
 
@@
 
+\newcommand{\autscale}[0]{0.48}
+
 \begin{figure}[ht]
     \centering 
-  \begin{subfigure}[b]{0.5\textwidth}
+  \begin{subfigure}[b]{\autscale\textwidth}
     \centering
-    \includegraphics[scale=\autscale]{a}
+    \includegraphics[width=\textwidth]{a}
     \caption{The automaton $A'$, whose Parikh registers count the length of the
@@
       %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\textwidth}
+  \end{subfigure}%
+  \hfill%
+  \begin{subfigure}[b]{\autscale\textwidth}
     \centering
-    \includegraphics[scale=\autscale]{b}
+    \includegraphics[width=\textwidth]{b}
     \caption{The automaton $B$, where the Parikh registers count the length of
@@
   \centering 
-  \includegraphics[scale=\autscale]{a_1}
+  \includegraphics[width=0.6\textwidth]{a_1}
   \caption{ $A'$ with its associated transition variables in symbolic form after
@@
   \centering 
-  \includegraphics[scale=\autscale]{a_2}
+  \includegraphics[width=0.6\textwidth]{a_2}
   \caption{ $A'$ after even more linear algebra, with most transitions expressed
@@
     \centering 
-  \begin{subfigure}[b]{0.5\textwidth}
+  \begin{subfigure}[b]{\autscale\textwidth}
     \centering
-    \includegraphics[scale=\autscale]{a_annotated}
+    \includegraphics[width=\textwidth]{a_annotated}
     \caption{ $A'$ with its associated transition variables in symbolic
@@
       %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\textwidth}
+  \end{subfigure}\hfill%
+  \begin{subfigure}[b]{\autscale\textwidth}
     \centering
-    \includegraphics[scale=\autscale]{b_annotated}
+    \includegraphics[width=\textwidth]{b_annotated}
     \caption{$B$ with its associated transition variables in symbolic forms.

### 2023-06-22 generalised.tex
Much tighter integration between sections

@@
 
-\subsection{Integer Datatype Constraints on Strings}\label{sec:parikh-automata}
+\subsection{Integer Constraints on Strings}\label{sec:parikh-automata}
 
-Parikh images are also applicable for deciding more general classes of
-string constraints~\cite{ostrich-plus}. Consider, for instance, a
-constraint involving the substring operation of \cref{const:s1-substring}:
-\begin{equation}\label{eq:substring}
-    s_1 = \Substring(s_2, i, n)
-\end{equation}
+Parikh images are also applicable for deciding more general classes of string
+constraints~\cite{ostrich-plus}. Consider the substring constraint
+\cref{const:s1-substring} of \cref{ex:string-constraints}: $ s_1 =
+\Substring(s_2, i, n)$, that is $s_1$ is an $n$-length substring of $s_2$
+starting at offset $i$ (in some variants , $n$ is $i + |n|$, i.e. the ending
+index). That constraint belongs to an expressive fragment of string logic that
+cannot be decided by most state-of-the-art string solvers.
 
-That is $s_1$ is an $n$-length substring of $s_2$ starting at offset $i$ (in
-some variants , $n$ is $i + |n|$, i.e. the ending index). That constraint
-belongs to an expressive fragment of string logic that cannot be decided by most
-state-of-the-art string solvers.
+In \cref{sec:intuition} we hand-wavily modelled \cref{const:s1-substring} (and
+  the other constraints) using \emph{Parikh
+  automata}~\cite{parikh-automata,expressiveness}. A Parikh automaton is a
+  standard NFA extended such that transitions are additionally labelled both
+  with offset vectors defining the increments of a finite number of counters.
+  This means that Parikh automata recognise words over an extended
+  alphabet~$\Alphabet \times D$, where $D \subseteq \Naturals^d$ is a finite set
+  of increment vectors (notation as in~\cite{expressiveness}) and $\Alphabet$ is
+  the alphabet of the original NFA.
 
-We can model \cref{eq:substring} using the notion of a \emph{Parikh
-  automaton}~\cite{parikh-automata,expressiveness}. A Parikh automaton
-is a standard automaton is extended such that transitions are labelled both with characters
-from the alphabet~$\Alphabet$, and with offset vectors defining the
-increments of a finite number of counters. This means that Parikh
-automata recognise words over an extended
-alphabet~$\Alphabet \times D$, where $D \subseteq \Naturals^d$ is a
-finite set of increment vectors (notation as
-in~\cite{expressiveness}). We use the symbols~$\pi_\Alphabet, \pi_D$
-to denote projections to the first and the second component of a
-composite letter~$(a, \Vector{v})$, respectively, and extend those
-projections to words:
+  We use the symbols~$\pi_\Alphabet, \pi_D$ to denote projections to the first
+  and the second component of a composite letter~$(a, \Vector{v})$,
+  respectively, and extend those projections to words:
 \begin{equation*}
@@
 
-    Applied to \eqref{eq:substring}, the decision procedure in
-\cite{ostrich-plus} will construct a pre-image of $\Language_2$ under
-the substring operation~$\cdot[n:m]$, and check whether this pre-image
-is consistent with the constraint~$x \in \Language_1$. Because the
-substring operation depends on the values of the integer
-variables~$n, m$, a Parikh automaton of dimension~$2$ is a suitable
-formalism to describe the pre-image, resulting
-in the automaton in Fig.~\ref{fig:parikh-automata} (right). The first
-component of the increment vectors is used to count the number of
-letters eliminated in the beginning of the string (value~$n$), while
-the second component records the beginning of the eliminated suffix
-(value~$m$).
+Applied to \cref{const:s1-substring}, the decision procedure in
+\cite{ostrich-plus} will construct a pre-image of $s_1$ under the substring
+operation, and check whether this pre-image is consistent with the constraint~$x
+\in \Language_1$, corresponding to the full product seen in
+\cref{ex:string-constraints}. Because the substring operation depends on the
+values of the integer variables~$i, n$, a Parikh automaton of dimension~$2$,
+shown in \cref{fig:aut_a}, models the pre-image. Intuitively, this construction
+accumulates any symbol (for the prefix of the substring operation), incrementing
+$i$ to mark the start of the substring, followed by the automaton representing
+the substring itself, modified to add a counter to increment $n$ at each
+transition, followed by a state that performs no increments and only accepts any
+suffix after the matched substring.
 
-Denoting the language described by Fig.~\ref{fig:parikh-automata}
-(right) as $\Language_{pre}$, we can then replace \eqref{eq:substring}
-with an equi-satisfiable formula that no longer contains any explicit
-substring operation:
+% @Philipp: I am quite unsure about which intersection we are looking for here,
+% please verify I'm not wrong!
+Denoting the language described by \cref{fig:aut_a} as $\Language_{pre}$, we can
+then replace $s_1 = \Substring(s_2, i, n)$ and $s_2 \in
+\Language(B)$ (of \cref{const:s1-in-c-dd}) with an
+equi-satisfiable formula that no longer contains any explicit substring
+operation:
 %
@@
   p \in \Language_{pre} \wedge~
-  \pi_\Alphabet(p) \in \Language_1
+  \pi_\Alphabet(p) \in \Language(B)
   \wedge~~ \pi_D(p) =
   \begin{bmatrix}
-    n \\ m
+    i \\ n
   \end{bmatrix}~~
-  \wedge 0 \leq n \leq m \leq |p|
+  \wedge 0 \leq i \leq i + n \leq |p|
 \end{equation}
 %
-To check the satisfiability of \eqref{eq:parikh-constraint}, we need a
+To check the satisfiability of \cref{eq:parikh-constraint}, we need a
 decision procedure that can process intersections of regular languages
-(in this case, of $\Language_{pre}$ and $\Language_1$, synchronising
+(in this case, of $\Language_{pre}$ and $\Language(B)$, synchronising
 on $\Alphabet$), while imposing the side
@@
 develop the lazy algorithm proposed in this paper.
-
-\iffalse
-\begin{figure}[t]
-  \centering
-      \includegraphics[width=0.5\linewidth]{parikh_automaton}
-      \caption{The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
-      $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
-      formula containing the constraints on the final register values cannot be
-      visualised.}\label{fig:parikh-automaton}
-    \end{figure}
-\fi

### 2023-06-22 extensions.tex
Various small fixes
Closes: #43
Closes: #903
Closes: #32

@@
 differentiate between the possible interpretations of the $\Sigma$ transitions
-without knowing ahead of times which ones will actually be materialied in the
+without knowing ahead of times which ones will actually be materialised in the
 product. If, on the other hand, if our $\Map$ had been length-counting, which

### 2023-06-22 preliminaries.tex
Various small fixes
Closes: #43
Closes: #903
Closes: #32
All numbers are natural
(might require application in other places!)

Closes: #50

@@
 
-Parikh's theorem states that any context-free language has a letter-equivalent
+% @Philipp: Reviewers suggest we drop this:
+Parikh's theorem states that any context-free language has a Parikh-equivalent
 regular language (c.f.~\cite{construction} for a construction of such automata

### 2023-06-24 calculus.tex
Work on example 5.1!

@@
 
-\begin{definition}
-  $\Connected(\Automaton, \Filter)$ for some automaton $\Automaton =
-  \AutomatonTuple$ holds when for every $\Transition =
-  \FromLabelTo{\State}{}{\State'} \in \Transitions$, $\Filter(\Transition) > 0
-  \implies \exists \Path \in \Paths(\Automaton)$ $\Filter(\Transition_i) > 0$
-  for $\Transition_i \in p$, and $\State \in \StatesOf(p)$, or in words that
-  there exists some $\Filter$-selected valid path that reaches $\Transition$'s
-  starting state, $\State$. Intuitively, it represents the condition that
-  $\Automaton$ is connected with respect to the selection function $\Filter$ for
-  every transition. It is redundant to $\Image$ by design.
+\begin{definition}\label{def:connected} $\Connected(\Automaton, \Filter)$ for
+  some automaton $\Automaton = \AutomatonTuple$ holds when for every
+  $\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions$,
+  $\Filter(\Transition) > 0 \implies \exists \Path \in \Paths(\Automaton)$
+  $\Filter(\Transition_i) > 0$ for $\Transition_i \in p$, and $\State \in
+  \StatesOf(p)$, or in words that there exists some $\Filter$-selected valid
+  path that reaches $\Transition$'s starting state, $\State$. It represents the
+  condition that $\Automaton$ is connected with respect to the selection
+  function $\Filter$ for every transition. It is redundant to $\Image$ by
+  design.
 \end{definition}
@@
 which all formulas are located in the antecedent~\cite{Fitting96a}.
-Each of the rules should be read and
-applied bottom-up, and relates
+
+Each rule should be read and applied bottom-up, and relates
 premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
 conclusion~$\SomeClause$. When constructing a proof, we start from some
-root~$\SomeClause$, and then apply proof rules to the goals of the proof
-until a goal can be closed, or no more rule is applicable.
+root~$\SomeClause$, and then apply proof rules to the goals of the proof until a
+goal can be closed, or no more rule is applicable.
 
@@
 root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic, so that a solution can be computed using
+in Presburger arithmetic allowing a solution to be computed using
 standard algorithms~\cite{Fitting96a}.
 
-We use the convention of splitting the
-formulas into linear (in)equalities ($\SomeInequalities$)
-and other formulas ($\SomeClause$), and assume that predicates
-$\Image$ and $\Connected$ only occur positively.
- Since
-our rules operate by adding and matching linear (in)equalities in a proof goal,
-we use the shorthand of listing the matched inequalities as antecedents
-e.g., in \Propagate{}).
+We use the convention of splitting the formulas into linear (in)equalities
+($\SomeInequalities$) and other formulas ($\SomeClause$), and assume that
+predicates $\Image$ and $\Connected$ only occur positively. Since our rules
+operate by adding and matching linear (in)equalities in a proof goal, we use the
+shorthand of listing the matched inequalities as antecedents e.g., in
+\Propagate{}).
+
 The filtering function~$\Filter$ is evaluated symbolically, and can in practice
@@
 \subsection{An Example}
+Here we will return to $A'$ from \cref{sec:intuition} and performs the steps of
+\cref{sec:intuition:algebra,sec:intuition:prop,sec:intuition:split} with the
+formal calculus we just established.
 
-Starting with~$\AcaOrBc{}$, where $\Map$ is the length function, in effect
-$\Transition \mapsto 1$ for transitions, and the constraints that the length is
-odd using the same trick as in \cref{sec:motivation}, we have the
-definitions in \cref{fig:example:single:equivalences} (omitting existential
-quantifiers and $x \geq 0$ for every variable to avoid clutter).
-
-\begin{figure}[ht]
-  We define $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$ as in the following two
-  equations, and then apply \EquationReasoning{} (under the implicit assumption
-  that every RHS is $\geq 0$) to obtain the third definition:
-  \begin{minipage}[b]{0.3\linewidth}
+Starting with~$A'$, $\Map$ extracts the increments of a transition (see
+\cref{fig:a_1}), $h(\FromLabelTo{q_1}{c / i}{q_2}) = \Vec{v}$ for the increment
+vector $\Vec{v}$, concretely $h(\FromLabelTo{I_A}{\Sigma / \left[ l_{A+}, i_+
+\right]}{I_A}) = \left[ l_{A+}, i_+ \right]$. We use the same compact notation
+here as in \cref{sec:intuition} to represent what is essentially a sparse vector
+of $1$ and $0$ coefficients. Initially, we let $\Filter$ map to fresh variables
+to obtain:
 \begin{equation*}
     \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = \TransitionVar_1 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_2 \\
-        \Filter(\FromLabelTo{A}{c}{A}) & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = \TransitionVar_5 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
+      \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A}) & = \TransitionVar_1    & 
+                \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) & = \TransitionVar_2 \\
+      \Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) & =  \TransitionVar_3 & 
+                \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) & = \TransitionVar_4 \\
+      \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) & = \TransitionVar_5 & 
+      \FlowEq(A', \Filter) & = \begin{aligned}
+        1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_3 \\
+        \TransitionVar_4 &= \TransitionVar_1 \\
+        \TransitionVar_3 + \TransitionVar_1 + \cancel{\TransitionVar_2} &\geq \cancel{\TransitionVar_2}  \\
+      \end{aligned} \\
     \end{aligned}
 \end{equation*}
-  \end{minipage}
-  \hspace{0.5cm}
-  \begin{minipage}[b]{0.3\linewidth}
+
+% RUN EXPAND GET:
+$i = \TransitionVar_5, l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k, n = \TransitionVar_4 + \TransitionVar_1$
+
+  If we apply reasoning by equivalence as in \cref{sec:a_1}, we end up with the
+  following definitions, corresponding to \cref{fig:a_1} at the end of
+  \cref{sec:a_1}:
   \begin{equation*}
-      \begin{aligned}
-        % S
-        \TransitionVar_4 &= \TransitionVar_5 + \TransitionVar_2 - 1 \\
-        % A
-        \TransitionVar_1 &= \TransitionVar_5 \\
-        % B
-        \TransitionVar_2 &= \TransitionVar_4 + \TransitionVar_6 \\
-        % F
-        \TransitionVar_5 + \TransitionVar_6 &= \FinalStateVar_1 \\
-        \FinalStateVar_1 &= 1 \\
+    \begin{aligned} % Post-propagation
+      \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A}) = \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) & = \TransitionVar_1\\
+      \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) & = l_A  - i - n\\
+      \Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) & =  1 - \TransitionVar_1\\
+      \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) & = i\\
     \end{aligned}
   \end{equation*}
-  \end{minipage}
-  \begin{minipage}[b]{0.3\linewidth}
-    \begin{equation*}
-      \begin{aligned}
-        \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-      \end{aligned}
-    \end{equation*}    
-  \end{minipage}
-  \caption{Equivalences defining $\Filter$ and $\FlowEq(\AcaOrBc{}, \Filter)$
-  respectively.}\label{fig:example:single:equivalences}
-  \end{figure}
 

### 2023-06-24 example.tex
Various small fixes
Closes: #43
Closes: #903
Closes: #32
All numbers are natural
(might require application in other places!)

Closes: #50
Work on example 5.1!

@@
 $\mathtt{*}$ the Kleene star, and $\mathtt{.}$, matches any single character.
-For the length of a string $s$, we write $\Length{s}$.
+For the length of a string $s$, we write $\Length{s}$. All number variables in
+this example are in $\Naturals$ with $0$.
 
@@
 
-\subsubsection{Eliminating paths through automata using flow analysis}
+\subsubsection{Eliminating paths through automata using flow analysis}\label{sec:a_1}
 
@@
 
-\subsubsection{Using linear algebra to eliminate flow}
+\subsubsection{Using linear algebra to eliminate flow}\label{sec:intuition:algebra}
 As an example, we start with reasoning about $A'$. Performing symbolic reasoning
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering 
@@
 
-\subsubsection{Propagation of connectivity constraints}
+\subsubsection{Propagation of connectivity constraints}\label{sec:intuition:prop}
 At this point we can conclude that none of the loops of $A'$ can become
@@
 
-\subsubsection{Case Splitting}
+\subsubsection{Case Splitting}\label{sec:intuition:split}
 After concluding this, we have a choice of two paths through automaton $B$; the
@@
 
-\subsubsection{Computing products}
+\subsubsection{Computing products}\label{sec:intuition:materialise}
 Now we have two small, stick-like automata after discounting all

### 2023-06-24 calculus.tex
An acceptable (but hardly great) example
This also adds propagation to 2.1, which I had forgotten.

Punching out from today after 5h 30 min.

@@
 \cref{sec:intuition:algebra,sec:intuition:prop,sec:intuition:split} with the
-formal calculus we just established.
+formal calculus we just established, but exclude \cref{const:s2-in-b} and
+therefore the entire automaton $B$ since we introduce support for products of
+automata in \cref{sec:multiple}. Note that this example then becomes
+satisfiable, as $s_1 = \Char{dd}, i = 1, s_2 = \Char{adda}, l_A = 4$ is a
+satisfying assignment to
+\cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring}.
 
@@
 here as in \cref{sec:intuition} to represent what is essentially a sparse vector
-of $1$ and $0$ coefficients. Initially, we let $\Filter$ map to fresh variables
-to obtain:
-\begin{equation*}
+of $1$ and $0$ coefficients, i.e. = $\left[ l_{A+}, i_+ \right] = \left[1, 1,
+0\right]$. Initially, we let $\Filter$ map to fresh variables to obtain:
+\begin{equation}
     \begin{aligned}
       \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A}) & = \TransitionVar_1    & 
-                \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) & = \TransitionVar_2 \\
-      \Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) & =  \TransitionVar_3 & 
-                \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) & = \TransitionVar_4 \\
-      \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) & = \TransitionVar_5 & 
-      \FlowEq(A', \Filter) & = \begin{aligned}
-        1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_3 \\
+                \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) & = \TransitionVar_2 &       \FlowEq(A', \Filter) & = \begin{aligned}
+                    1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_3 + \TransitionVar_1\\
                     \TransitionVar_4 &= \TransitionVar_1 \\
@@
                   \end{aligned} \\
+      \Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) & =  \TransitionVar_3 & 
+                \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) & = \TransitionVar_4 &\\
+      \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) & = \TransitionVar_5 & 
     \end{aligned}
-\end{equation*}
+\end{equation}\label{eq:single-filter}
 
-% RUN EXPAND GET:
-$i = \TransitionVar_5, l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k, n = \TransitionVar_4 + \TransitionVar_1$
+Like in \cref{sec:a_1,sec:intuition:algebra}, we apply reasoning by equivalence
+  to end up with fixed values for several of the transition variables
+  $\TransitionVar_k$, and eventually we use \cref{const:more-inside-than-before}
+  and \cref{const:something-before-and-after} to conclude that the transition
+  directly from state $I_{A}$ to $F_{A}$ is incompatible with the constraint.
 
-  If we apply reasoning by equivalence as in \cref{sec:a_1}, we end up with the
-  following definitions, corresponding to \cref{fig:a_1} at the end of
-  \cref{sec:a_1}:
-  \begin{equation*}
-    \begin{aligned} % Post-propagation
-      \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A}) = \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) & = \TransitionVar_1\\
-      \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) & = l_A  - i - n\\
-      \Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) & =  1 - \TransitionVar_1\\
-      \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) & = i\\
-    \end{aligned}
-  \end{equation*}
+  Then, following \cref{sec:intuition:subsume}, we remove the $\Connected{}$
+   predicate using \Subsume{} since no transition can can become disconnected
+   (and in fact never could for $A'$, since all looping transitions are attached
+   to the initial and single accepting state and therefore always must be
+   reachable from any accepting path). After this, the rest of the reasoning can
+   be continued without any part of our calculus, since all values of the
+   existentially quantified transition variables
+   $\TransitionVar_1,\ldots,\TransitionVar_5$ compatible with
+   \cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring,const:something-before-and-after}
+   are represented by the flow equations introduced by $\Expand{}$. The
+   derivation is shown in \cref{fig:derivation:single}.
 
@@
 \begin{prooftree}
-  \hypo[]{
-{    \begin{array}{lll}
-      1 - \TransitionVar_6 = 0 & \land \TransitionVar_6 = 1 & \land  \TransitionVar_4 + 1 > 0 \\
-      \land \TransitionVar_3 = 0 &\land  2 + \TransitionVar_4 > 0 & \land k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-{    \begin{array}{lll}
-      \Connected(\AcaOrBc{}, \Filter) & \land 1 - \TransitionVar_6 = 0 \\
-      \land \TransitionVar_6 = 1 & \land \TransitionVar_4 + 1 > 0 \land \\
-      \land \TransitionVar_3 = 0 & \land 2 + \TransitionVar_4 > 0 \land \\
-      k = 1 + \TransitionVar_4 
-    \end{array}}
-  }
-  \infer1[\Propagate]{
+  \hypo[]{{\text{Regular solving finishes the proof}}}
+  \infer1[]{
     \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      \TransitionVar_6 = 1 \land \\
-      \TransitionVar_4 + 1 > 0 \land \\
-      \TransitionVar_3 \geq 0 \land \\
-      2 + \TransitionVar_4 \geq 0 \land \\
-      2k = 2 + 2\TransitionVar_4 + \TransitionVar_3
+    \TransitionVar_1 = 1 \land 
+    i = \TransitionVar_5 \land 
+    n = 2 \land \\
+    2 \geq i \land l_A - i - 2 > 0 
+    \land \TransitionVar_4 = 1 
+    \land \TransitionVar_2 = l_A - i - 2 
+    \land \TransitionVar_3 = 0 \\
+    \end{aligned}}
+  \infer1[\Subsume: \cref{sec:intuition:subsume}]{
+    \begin{aligned}
+    \Connected(A', \Filter) \land 
+    \TransitionVar_1 = 1 \land 
+    i = \TransitionVar_5 \land 
+    n = 2 \land 
+    2 \geq i \land l_A - i - 2 > 0  \\
+    \land \TransitionVar_4 = 1 
+    \land \TransitionVar_2 = l_A - i - 2 
+    \land \TransitionVar_3 = 0 \\
     \end{aligned}
   }
-  \infer1[\EquationReasoning{}]{
+  \infer1[\EquationReasoning{}: $\TransitionVar_1 = 1$ \cref{sec:a_1}]{
     \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 = 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
+    \Connected(A', \Filter) \land \\
+    i = \TransitionVar_5 \land 
+    n = 2\TransitionVar_1 \land n \geq i \land l_A - i - n > 0 
+    \land \TransitionVar_4 = \TransitionVar_1 \\
+    \land \TransitionVar_2 = l_A - i - n 
+    \land \TransitionVar_3 = 1 - \TransitionVar_1 
     \end{aligned}
   }
-  % BRANCH:  = 0'
-  \hypo{
-    {\begin{array}{lll}
-      \TransitionVar_3 > 0 & \land 
-      \TransitionVar_6 = 0 \\ \land
-      2\TransitionVar_6 + \TransitionVar_4 = 0 & \land
-      1 - \TransitionVar_6 > 0\\  \land 
-      2k = 1 + \TransitionVar_3
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
+  \infer1[\EquationReasoning{}: $=$-elimination]{
     \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_3 > = 0 \land \\
-      \TransitionVar_6 = 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + \TransitionVar_3
+    \Connected(A', \Filter) \land \\
+    i = \TransitionVar_5 \land 
+    n = \TransitionVar_4 + \TransitionVar_1 \land n \geq i \land l_A - i - n > 0 \\
+    \land \TransitionVar_4 = \TransitionVar_1 
+    \land \TransitionVar_2 = l_A - i - n 
+    \land \TransitionVar_3 = 1 - \TransitionVar_1 
     \end{aligned}
   }
-  \infer1[\EquationReasoning{}]{
+  \infer1[\EquationReasoning{}: expand sum]{
     \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2\TransitionVar_6 + \TransitionVar_4 = 0\\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
+        i = \TransitionVar_5, l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k \land n = \TransitionVar_4 + \TransitionVar_1 \land n \geq i \land l_A - i - n > 0
+    \land \Connected(A', \Filter)
     \end{aligned}
   }
-  \infer1[\Propagate{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      \TransitionVar_6 = 0 \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3
-    \end{aligned}
+  \infer1[\EquationReasoning{}: expand sum]{
+    \Connected(A', \Filter) \land
+    \left[l_A, i, n\right] =
+    \sum\limits_{\Transition \in \Transitions_{A'} } \Map(\Transition) \cdot \Filter(\Transition)
   }
-  \infer1[\EquationReasoning{}]{
-    \begin{aligned}
-      \Connected(\AcaOrBc{}, \Filter) \land \\
-      1 - \TransitionVar_6 > 0 \land \\
-      2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6
-    \end{aligned} % BRANCH: > 0
-  } % SPLIT
-  \infer2[\Split{} $1 - \TransitionVar_6$]{ \Connected(\AcaOrBc{}, \Filter) \land 2k = 1 + 2\TransitionVar_4 + \TransitionVar_3 + \TransitionVar_6 }
-  \infer1[\EquationReasoning{}]{
-    \Connected(\AcaOrBc{}, \Filter) \land
-    2k + 1 =
-    \sum\limits_{\Transition \in \Transitions} (\Transition \mapsto 1)(\Label) \cdot \Filter(\Transition)
-  }
-  \infer1[\Expand{}]{\Image{}_{\AcaOrBc{}, \Transition \mapsto 1}(\Filter, 2k + 1)}
+  \infer1[\Expand{}]{\Image{}_{A', \Map}(\Filter, \left[l_A, i, n\right]) \land n \geq i \land l_A - i - n > 0}
 \end{prooftree}
-\caption{A derivation for \Calculus{} computing odd lengths in $\AcaOrBc{}$.}\label{fig:derivation:single}
+\caption{A derivation tree for an application of
+\cref{const:s1-in-c-dd,const:s1-substring,const:something-before-and-after,const:more-inside-than-before}
+from \cref{ex:string-constraints}, corresponding to handling the Parikh
+automaton $A'$ of \cref{fig:aut_a}. Note that the flow equations of
+\cref{eq:single-filter} are left out of the tree in order to keep it
+manageable.}\label{fig:derivation:single}
 \end{figure}
-
-In \cref{fig:derivation:single}, we see how we start by expanding the predicate
-using the (simplified) flow equations. We can also see how interleaving
-reasoning on the corresponding linear equations helps the deduction. In both
-branches the reasoning is similar: we conclude that when choosing either path at
-the starting state we must avoid the other one, perform propagation based on
-that fact, use some algebraic reasoning to derive the expected forms of the
-upper and lower bounds to propagate the now disconnected transitions, and
-finally subsume, removing the $\Connected{}$ predicate when we are unable to use
-either of the rules, leaving only a set of linear equations. To obtain a solution from the
-proof, one would need to perform standard model generation on the remaining
-constraints to obtain values for the sought variable $\MonoidElement$. The full
-Parikh image is the disjunction of the constraints left at the two
-leaves after quantifier elimination.
-
-Following the values of $\Filter$, we can read out the path corresponding to our solution in the product, and with that a word.
-In our case, for the right branch we have $\Filter(\FromLabelTo{S}{b}{B}) = 0,
-\Filter(\FromLabelTo{A}{c}{A})  > 0,
-\Filter(\FromLabelTo{B}{b}{S}) = 0,
-\Filter(\FromLabelTo{S}{a}{A}) = 1 ,
-\Filter(\FromLabelTo{A}{a}{F}) = 1 ,
-\Filter(\FromLabelTo{B}{c}{F}) = 0$.
-We can read out a word along it: aca, accca, etc, and the corresponding constraint: $\exists k : \TransitionVar_3 = 2k -1 \land \TransitionVar_3 > 0 \land k \geq 0$.
 

### 2023-06-24 example.tex
An acceptable (but hardly great) example
This also adds propagation to 2.1, which I had forgotten.

Punching out from today after 5h 30 min.

@@
 
-\subsubsection{Propagation of connectivity constraints}\label{sec:intuition:prop}
+\subsubsection{Subsumption of connectivity constraints}\label{sec:intuition:subsume}
 At this point we can conclude that none of the loops of $A'$ can become
@@
 
-\subsubsection{Case Splitting}\label{sec:intuition:split}
+\subsubsection{Case Splitting and propagation of connectivity constraints}\label{sec:intuition:split}
 After concluding this, we have a choice of two paths through automaton $B$; the
@@
 option on the first-fail principle.
+
+Having performed the split, we can conclude that the path from $I_B$ to state
+$B$ is now unreachable without passing through a zero-times-used transition.
+This means that the self-loop from state $B$ is also unusable, and thus must
+also be used zero times.
 

### 2023-06-27 calculus.tex
Acceptable version of 5.1

@@
 
-\begin{figure}
+\begin{figure}[ht]
   \centering
@@
     \begin{aligned}
-    \TransitionVar_1 = 1 \land 
     i = \TransitionVar_5 \land 
@@
     2 \geq i \land l_A - i - 2 > 0 
-    \land \TransitionVar_4 = 1 
     \land \TransitionVar_2 = l_A - i - 2 
@@
     2 \geq i \land l_A - i - 2 > 0  \\
-    \land \TransitionVar_4 = 1 
     \land \TransitionVar_2 = l_A - i - 2 
@@
     n = 2\TransitionVar_1 \land n \geq i \land l_A - i - n > 0 
-    \land \TransitionVar_4 = \TransitionVar_1 \\
     \land \TransitionVar_2 = l_A - i - n 
@@
     \begin{aligned}
-    \Connected(A', \Filter) \land \\
-    i = \TransitionVar_5 \land 
-    n = \TransitionVar_4 + \TransitionVar_1 \land n \geq i \land l_A - i - n > 0 \\
+    \Connected(A', \Filter)  
+    \land i = \TransitionVar_5 
+    \land n = \TransitionVar_4 + \TransitionVar_1 
+    \land n \geq i 
+    \land l_A - i - n > 0 \\
     \land \TransitionVar_4 = \TransitionVar_1 
@@
     \land \TransitionVar_3 = 1 - \TransitionVar_1
+    \land 1 = \TransitionVar_3 + \TransitionVar_1 
+    \land \TransitionVar_4 = \TransitionVar_1 
     \end{aligned}
   }
-  \infer1[\EquationReasoning{}: expand sum]{
+  \infer1[\EquationReasoning{}: expand $\Sigma, \FlowEq$]{
     \begin{aligned}
-        i = \TransitionVar_5, l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k \land n = \TransitionVar_4 + \TransitionVar_1 \land n \geq i \land l_A - i - n > 0
+    i = \TransitionVar_5 
+    \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k 
+    \land n \geq i \land l_A - i - n > 0
+    \land n = \TransitionVar_4 + \TransitionVar_1\\
     \land \Connected(A', \Filter) 
+    \land \FlowEq(A', \Filter)
     \end{aligned}
@@
     \Connected(A', \Filter) \land
+    \FlowEq(A', \Filter) \land
     \left[l_A, i, n\right] =
@@
 from \cref{ex:string-constraints}, corresponding to handling the Parikh
-automaton $A'$ of \cref{fig:aut_a}. Note that the flow equations of
-\cref{eq:single-filter} are left out of the tree in order to keep it
-manageable.}\label{fig:derivation:single}
+automaton $A'$ of \cref{fig:aut_a}. Note the liberal use of removal of equal
+variables (e.g. if $x = y$ appears as a statement, we substitute $y$ for $x$
+everywhere and drop $\ldots \land x=y \land \ldots$) to keep the size of the
+equations managable.}\label{fig:derivation:single}
 \end{figure}
+
+To obtain the automaton corresponding to the final model, one can evaluate
+$\Filter(\Transition) = k$ for each transition $\Transition$ and construct a
+path with each such transition taken $k$ times.
 

### 2023-06-27 evaluation.tex
Evaluation is now a separate file

--- 2023-06-22 main.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2023}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\usepackage{amsmath,empheq,fancybox}
-\usepackage{paralist}
-\usepackage{url}
-\usepackage{color}
-\usepackage{textcomp,listings}
-\usepackage{array}
-\usepackage{mymacros}
-\usepackage{microtype}
-\usepackage{listings}
-\usepackage{csquotes}
-\usepackage{proof}
-\usepackage[capitalise]{cleveref}
-\usepackage{algorithm2e}      
-\usepackage{multirow}
-\usepackage{mathpartir}
-\usepackage{amsthm}
-\usepackage{numprint}
-\usepackage{ebproof}
-\usepackage[makeroom]{cancel}
-
-\usepackage{mathtools} % Bonus
-
-\usepackage{enumitem}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{example}{Example}[section]
-
-\newlist{constraints}{enumerate}{3}
-\setlist[constraints,1]{%
-    label=\sffamily{(\roman*):},
-    ref=\normalfont{Constraint (\roman*)},
-    wide,itemsep=0pt,topsep=0pt
-}    
-\crefname{constraintsi}{}{}
-\Crefname{constraintsi}{}{}
-
-\lstset{
-    columns=fullflexible,
-    showspaces=false,
-    showtabs=false,
-    breaklines=true,
-    showstringspaces=false,
-    breakatwhitespace=true,
-    escapeinside={(*@}{@*)},
-    commentstyle=\color{greencomments},
-    keywordstyle=\color{bluekeywords},
-    stringstyle=\color{redstrings},
-    numberstyle=\color{graynumbers},
-    basicstyle=\ttfamily\small,
-    framesep=12pt,
-    xleftmargin=12pt,
-    tabsize=4,
-    captionpos=b
-}
-
-\newif\ifcomments
-\commentstrue
-\newif\ifoutline
-\outlinetrue
-
-\newcommand{\contents}[1]{\ifoutline{\color{blue}
-    \begin{itemize}
-    #1
-    \end{itemize}
-  }\fi}
-
-\allowdisplaybreaks[1]
-
-\begin{document}
-
-%% Title information
-\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
-   counts in strings of a given language. Automata-based string solvers
-   frequently need to compute the Parikh image of large products (intersections)
-   of nondeterministic automata, which in many operations is both prohibitively
-   slow and memory-intensive. We contribute a novel understanding of how Parikh
-   maps can be tackled as a constraint solving problem to solve real-world
-   constraints stemming from functions on regular languages, most notably the
-   length constraint. Furthermore, we show how this formulation can be
-   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
-   prover supporting Presburger logic.  We implement \Calculus{} in a tool
-   called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving Parikh automata
-   intersection problems produced when solving standard string constraint
-   benchmarks involving string length constraints. We show that our solution
-   strictly outperforms the standard approach described
-   in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wide
-   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
-   model checker.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-  <ccs2012>
-  <concept>
-  <concept_id>10003752.10003766.10003773.10003775</concept_id>
-  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003766.10003776</concept_id>
-  <concept_desc>Theory of computation~Regular languages</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10003794</concept_id>
-  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10010124.10010138.10010142</concept_id>
-  <concept_desc>Theory of computation~Program verification</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10011192</concept_id>
-  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  </ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Quantitative automata}
-\ccsdesc[300]{Theory of computation~Regular languages}
-\ccsdesc[500]{Theory of computation~Automated reasoning}
-\ccsdesc[300]{Theory of computation~Program verification}
-\ccsdesc[300]{Theory of computation~Verification by model checking}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}\label{sec:introduction}
-\input{introduction.tex}
-
-\section{An Intuition for Our Approach}\label{sec:intuition}
-\input{example.tex}
-
-\section{Preliminaries}\label{sec:preliminaries}
-\input{preliminaries.tex}
-
-\section{Projections on Parikh Images}\label{sec:generalised}
-\input{generalised.tex}
-
-\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
-\input{calculus.tex}
-
-\section{Parikh Images from Products of Automata}\label{sec:multiple}
-
-We now generalise our calculus to natively work with intersections of
-regular languages, or equivalently products of automata. For this
-extension, we change the main predicate~$\Image$ to be indexed by a
-vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
-simplicity, we assume that the sets of states of the $k$ automata (and
-therefore also the transition sets) are pairwise disjoint. 
-\begin{definition}\label{def:multiple}
-  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
-  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
-  $\Monoid$,
-  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
-  is a transition selection function, and
-  $\MonoidElement \in \Monoid$. The predicate
-  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
-      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
-  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
-  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
-  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
-  \begin{itemize}
-  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
-    $\Filter$, that is,
-    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
-  \item the automata all accept the same
-    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
-  \item the accepted word is mapped to
-    $\MonoidElement = \Map(\WordOf(\Path_i)) =
-    \Map(\WordOf(\Path_1))$.
-  \end{itemize}
-\end{definition}
-
-  \begin{table}[t]
-    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-      Name & \text{Rule} & Side conditions\\
-      \midrule
-
-      % EXPAND
-      \ExpandM & 
-      \inferrule
-      {
-       {\begin{array}{c}
-           \left\{
-           \FlowEq(\Automaton_i, \Filter),~~
-           \Connected(\Automaton_i, \Filter),~~
-           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
-           \right\}_{i=1}^k,
-           \\[4ex]
-          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
-          \SomeInequalities, \SomeClause
-        \end{array}}
-        }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
-      None \\[5ex]
-      \Materialise &
-      \inferrule
-      {{
-         \begin{array}{c}
-           \FlowEq(\Automaton', \Filter'),~
-           \Connected(\Automaton', \Filter'),~
-           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
-           \\\ImagePredicate{
-                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
-                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
-                     \Automaton'},~ \SomeInequalities, \SomeClause
-           }{\Map}{\Filter'}{\MonoidElement}
-           \end{array}}
-           }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
-      $\begin{array}{@{}l@{}}
-         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
-         1 \leq i < j \leq k, \\
-      \Automaton' = \Automaton_i \times \Automaton_j,\\
-      \end{array}$ \\
-      \bottomrule
-      \end{tabular}
-      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
-    \end{table}
-
-    For the calculus (\cref{tbl:rules:multi}), we first extend
-    $\Expand$ to generate flow equations and instances of $\Connected$
-    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
-    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
-    since it is needed to keep track of the currently considered
-    partial products.
-
-The rule~$\Materialise$ introduces the product of two individual
-automata~$\Automaton_i, \Automaton_j$; this step eliminates
-$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
-instead adds the product~$\Automaton_i \times \Automaton_j$, while
-also introducing the flow equations and the $\Connected$ predicate.
-
-The rule~$\Materialise$ also has to connect the newly introduced
-product~$\Automaton_i \times \Automaton_j$ to the previous automata
-$\Automaton_i, \Automaton_j$. This is done by extending the
-selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
-to fresh variables~$x_{\Transition}$:
-\begin{equation*}
-  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
-  ~~=~~
-  \begin{cases}
-    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
-    \\
-    \Filter(t) & \text{otherwise}
-  \end{cases}
-\end{equation*}
-
-The multiplicity of transitions in the product then has to be related
-to the multiplicities in the individual automata, modelled using the
-$\BindingSum$ predicate. The predicate expresses that the multiplicity
-of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
-$\Automaton_i$ has to coincide with the sum of the multiplicities of
-transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
-and similarly for $\Automaton_j$:
-%
-  $$
-  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
-  \begin{aligned}
-  & \Set{ 
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
-  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
-  & \Set{
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
-  }
-  \end{aligned}
-$$
-
-\iffalse
-Note that this definition implies that $\Filter(\Transition) =
-\Filter(\Transition')$ whenever two transitions $\Transition \in
-\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
-produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
-\times \Automaton_2}$. This corresponds to our intuition that the terms of the
-product must agree on the value they accept. As before, we implicitly map
-$\Filter$ to fresh terms for each transition in the product.
-\fi
-
-For instances of precisely one automaton, neither rule applies and we
-perform the calculus as before.
-
-\subsection{An Example}\label{sec:multiple:example}
-
-We return again to our example in \cref{sec:introduction:parikh}, where we
-compute the whole Parikh image of the product of $\AcaOrBc{}$ and
-$\SomethingCSomething$ under the constraint that there are more instances of
-letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
-element-wise addition, and $\Map$ that maps each transition to the corresponding
-increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.
-
-In the interest of space, we refer back to
-\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
-$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
-substitutions as follows. Note the expansion of the $\Sigma$ labels:
-    \begin{equation*}
-      \begin{aligned}
-        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-        % Other automaton
-        \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
-        \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
-      \end{aligned}
-    \end{equation*}
-
-In the derivation tree of \cref{fig:derivation:multi} we start as before, but
-with the product version of the $\Image$ predicate. The only possible rule here
-is $\ExpandM$, which we use to add the corresponding constraints on each
-automata of the product as we would have had in the single-automaton version. We
-perform algebraic reasoning on those equations, along with the constraint on
-$\MonoidElement$ to determine bounds on transition variables that will enable us
-to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
-finished doing so, we use \Materialise{} to compute the product of the now
-filtered automata. Further algebraic reasoning allows us to remove the final
-instance of $\Connected$. We then perform \Expand{} to get rid of the final
-$\Image$ instance, and are immediately able to $\Subsume$ the resulting
-$\Connected{}$ predicate. This, again, leaves us with a set of linear
-inequalities at the leaf that we can use to obtain a model of our final
-$\MonoidElement$ values (now a vector). This time we can read out aca, or
-$\VectorLiteral{2, 0, 1}$.
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo{
-    {\begin{array}{lll}
-      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
-      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
-    \end{array}}  
-  }
-  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
-    {\begin{array}{ll}
-      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
-      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
-    \end{array}}  
-  }
-  \infer1[\EquationReasoning{}]{
-  {  \begin{array}{ll}
-      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
-      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
-        \TransitionVar_6 + \TransitionVar_4 = 0 \\
-        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
-        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
-        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
-          \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1} 
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} \\ \land 
-      \Connected(\SomethingCSomething{}, \Filter) &
-      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array} } 
-  }
-  \infer1[\Materialise]{
-    {\begin{array}{ll}
-      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
-      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
-      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
-      \land 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} & \land
-      \Connected(\SomethingCSomething{}, \Filter) \\
-      \land 
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-  {\begin{array}{ll}
-    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-    1 - \TransitionVar_6 > 0 &
-    \land
-    \TransitionVar_6 = 0 \land 
-    \TransitionVar_4 = 0 \\
-    \land
-    \TransitionVar_3 > 0 & \land
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
-    \land
-    \VectorLiteral{
-      \TransitionVar_{7a} + \TransitionVar_{9a},
-      0,
-      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-      = \VectorLiteral{
-        2,
-        0,
-        \TransitionVar_3} & \land 
-    \Connected(\AcaOrBc{}, \Filter) \\
-    \land 
-    \Connected(\SomethingCSomething{}, \Filter) \\
-    \land 
-    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-  \end{array}}
-  }
-  \infer1[\EquationReasoning]{
-    \begin{aligned}
-      \VectorLiteral{a, b, c} = \VectorLiteral{
-          2 - 2\TransitionVar_6,
-          2\TransitionVar_4 + 3\TransitionVar_6,
-          \TransitionVar_3 + \TransitionVar_6
-        } \land
-        \\
-        \VectorLiteral{a, b, c} = \VectorLiteral{
-            \TransitionVar_{7a} + \TransitionVar_{9a},
-            \TransitionVar_{7b} + \TransitionVar_{9b},
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          } \land \\
-      \Connected(\AcaOrBc{}, \Filter) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{a, b, c}) \land a > b
-    \end{aligned}
-  }
-  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
-\end{prooftree}
-\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
-\end{figure}
-
-\subsection{Correctness of \Calculus{} for Products of Automata}
-
-Since \cref{tbl:rules:multi} only extends the existing rules of
-\cref{tbl:rules:single}, we focus on the differences compared
-to the calculus for a single automaton.
-
-%\subsubsection{\Calculus{} for products of automata terminates}
-\begin{lemma}\label{lma:multi-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the product
-  version of $\Image$ only occurs positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
-\end{lemma}
-
-\begin{proof}
-  The rule \Materialise{} can similarly only be used finitely many times, as
-  each application reduces the number of automata in the product of $\Image$ by
-  one automaton, until only one remains and \cref{lma:single-terminates} for
-  single-automaton instances applies.
-
-  This implies that also the rule~\ExpandM{} can only be applied
-  finitely often since its side condition only allows applying it once
-  per instance of an $\Image$ predicate containing a product, and only finitely
-  many instances of $\Image$ can be introduced on each branch.
-\end{proof}
-
-%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
-
-Since our calculus now includes a rule introducing new variables, the
-\Materialise{} rule, we have to slightly generalise the notion of
-solution-preservation:
-%
-\begin{lemma}\label{lma:multi-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:multi}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
-  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
-  there is an extension~$\beta' of \beta$ satisfying one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  We have to consider the two new rules in \cref{tbl:rules:multi}. The
-  result is immediate for \ExpandM{}, since this rule does not remove the
-  $\Image$ predicate from a proof goal, and the newly introduced formulas
-  are all implied by the $\Image$ predicate.
-
-  For \Materialise{}, observe that the existence of an accepting path
-  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
-  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
-  same word. The path in the product will satisfy the flow equations
-  and connectedness, and it will be related to the individual paths as
-  stipulated by the \BindingSum{} predicate.
-\end{proof}
-
-\section{Extensions}\label{sec:extensions}
-\input{extensions.tex}
-
-\section{Implementation}\label{sec:implementation}
-
-We implement \Calculus{} for Parikh automata as described in
-\cref{sec:parikh-automata}. The artefact submitted along with this paper is a
-program that reads an instance file with one or more products of one or more
-Parikh automaton with transition labels defined as ranges of Unicode characters,
-along with a set of constraints on the final values of their registers expressed
-as Presburger arithmetic in a C-like syntax. We call this program
-\Catra.
-% \footnote{If you really must read it as an acronym, please read it as
-%CAtegory Theory on Register Automata, or if you object to the somewhat
-%nonstandard use of register automata and category theory, as Check Assignments
-%of The Registers Afterwards. Or alternatively if you find it all to be too much
-%of a theoretical exercise, as Can Anyone Think of A Real Application.}
-
-\Catra{} is written in Scala, with the calculus described in this paper
-implemented as a theory plug-in for the \Princess{} automated theorem
-prover~\cite{princess}, which also performs the Presburger reasoning. For
-comparison, we also provide an implementation of the baseline method
-from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
-symbolic model checker~\cite{nuxmv} to solve our constraints, and the
-approximation described in~\cite{approximate-parikh} on top of the standard
-baseline back-end. An example of an input file corresponding to our running
-example introduced in \cref{sec:motivation} can be seen in
-\cref{lst:input-example}.
-
-\Catra{} uses symbolic labels for automata. A symbolic label is defined as a
-finite range of Unicode code points. This allows succinct representation of many
-regular expression patterns such as \lstinline{(a-z).*} which would have
-otherwise required $27$~transitions. The transition for the range would be
-written \lstinline{a -> b [97, 122]}.
-
-In satisfaction mode, supported by all backends, \Catra{} tries to satisfy the
-constraints expressed by the input file, reporting \Sat{} with register
-assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
-Additionally, baseline and \Calculus{} also support generation of the Presburger
-formula describing the constraints of the input file. Baseline uses standard
-quantifier elimination, and \Calculus{} uses the method described in
-\cref{sec:finding-the-image}.
-
-Since \Princess{} does not support multiple-arity predicates like the ones we
-use in \Calculus{}, we have implemented variable-length arguments using
-additional helper predicates. These are $\Unused{}(\Automaton)$, which marks an
-automaton as unused in any product, and $\TransitionMask{}(\Automaton,
-\Transition, \Filter(\Transition))$ which associates a transition $\Transition$
-and automaton $\Automaton$ with its corresponding transition variable.
-Additionally, we associate each of our predicates with an instance variable in
-order to differentiate instances of the predicates.
-
-\subsection{Implementing the Baseline}\label{sec:implementing-baseline}
-
-We baseline using the same Presburger solver (\Princess{}), input file parser,
-and automaton implementation as \Catra. We do this in order to better analyse
-the impact of the calculus rules themselves. Using the formula of
-\eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
-successive term and add them to \Princess. We compute the product incrementally
-term by term, checking satisfiability at each step. We use a priority queue to
-select automata for each step, and order it by the number of transitions as a
-heuristic for the size of the automata. We use this heuristic to avoid computing
-large (and therefore slow) products until we have to, banking our hopes on
-computing an empty intermittent product early. The pseudocode for our
-implementation can be seen in \cref{alg:baseline}.
-
-\begin{algorithm}
-  \caption{How we implement the baseline approach}\label{alg:baseline}
-  \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
-  \KwResult{\textsc{Sat} or \textsc{Unsat}}
-  \SetKwFunction{NewTheoremProver}{newTheoremProver}
-  \SetKwFunction{NewPriorityQueue}{newPriorityQueue}
-  \SetKwFunction{Dequeue}{dequeue}
-  \SetKwFunction{Enqueue}{enqueue}
-  \SetKwFunction{Assert}{assert}
-
-  $p \gets \NewTheoremProver{}$
-
-  \Assert{$p$, $\SomeClause$}
-
-  \ForEach{$\Automaton_i$}{
-    \Assert{$p, \ParikhMap(\Automaton_i)$}
-
-    \If{$p$ is \textsc{Unsat}}{break}
-
-  }
-
-  $q \gets \NewPriorityQueue{}$
-
-  \While{$p$ not \textsc{Unsat} and $|q| > 1$}{
-    $\Automaton, \Automaton' \gets \Dequeue{q}$
-
-    \Assert{$p, \ParikhMap(\Automaton \times \Automaton')$}
-
-    \Enqueue{$q, \Automaton \times \Automaton'$}
-  }
-
-  \KwRet{$p$'s SAT status}
-
-  \end{algorithm}
-
-As an optimisation, our automata (including intermittent products) are created
-forward- and backward- reachable-minimal. Any automaton we produce only contains
-states that are both reachable from the initial state and has a path to an
-accepting state. We never perform any other minimisation on the automata for
-either backend. More complex minimisation was left out since performing
-minimisation on automata with counters is non-trivial, and minimising symbolic
-automata risks exponential blowup \cite{minimising-symbolic}.
-
-\subsection{Heuristics and search strategies}
-
-There are a number of choices left unspecified in \Calculus{} as described in
-\cref{sec:calculus,sec:multiple}. For example, the order of materialisation of
-intermediate products and the order of splitting. In this section we describe
-additional implementation details and techniques used to enhance \Catra.
-
-\subsubsection{Splitting, Materialisation, and Propagation}
-
-We order our rule applications so that we first propagate connectedness if
-possible, then perform materialisation if tractable as defined below, then
-finally resort to splitting if we must.
-
-In addition to applying \Split{} as described in \cref{tbl:rules:single} to
-randomly selected transitions, we prefer splitting to sever a strongly connected
-from the initial state. We randomly select an automaton where we can compute a
-cut between an SCC and the initial state, that is where the SCC does not contain
-the initial state and where the sum of the transition variables of the
-transitions in the cut is not known to be positive. If there are multiple such
-strongly connected components we choose one randomly. We then proceed to split
-on the sum of the transition variables of the cut as if it were a regular
-transition, e.g. its sum being zero or nonzero. In this way we drive \Calculus{}
-towards applying \Propagate{}.
-
-The implementation of the connectedness constraint is opportunistic and straightforward.
-We compute a set of dead states by performing forward and backwards reachability computations on an automaton, where we disregard any transition whose associated variable is known to be zero.
-After that we add clauses ensuring any transition variable associated with a transition starting in a dead state is zero.
-
-Product materialisation is the final piece of the puzzle. In the current
-implementation we put off computing intermediate products until at least all but
-two transition variables of one of the automata is known to be either present or
-absent. The number was chosen experimentally, and we observe a consistent trend
-towards lower numbers being better. The other automaton for the product is
-selected randomly.
-
-\subsubsection{Clause Learning}
-
-\Catra{} enables clause learning by default when using our backend, as it has
-been experimentally shown to increase the performance in aggregate (though not
-strictly). We do not currently implement all the proposed features of
-\cref{sec:ext:backjumping}, but we do implement forward-reachability cut
-learning. No sophisticated clause learning for products has been implemented.
-
-\subsubsection{Random Restarts}
-
-Finally, we perform restarts scaled by the Luby series~\cite{luby}. Experimental
-results have shown this to have a large improvement in performance, which is
-unsurprising given how many random choices we make during solving and how
-tail-heavy our problem is.
-
-\section{Evaluation}\label{sec:experiments}
-
 We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
@@
 suggesting that the automaton implementation is not the bottleneck.
-
-\section{Conclusion}
-
-In this paper we have introduced a calculus to compute commutating operations on
-intersections of regular languages that we call \Calculus{}. We have evaluated
-it on \NrBenchmarks{} Parikh automata intersection problems generated by the
-\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
-suite \cite{pyex} using our Parikh automata solver \Catra{}.
-
-Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
-usage and solve-time compared to the baseline approach laid out in
-\cite{generate-parikh-image} when implemented on the same underlying automated
-theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
-\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
-and generally outperforming it for timeouts under 30 seconds with its advantage
-increasing drastically for even shorter timeouts. 30 seconds would generally be
-considered a long timeout for our intended use as supporting infrastructure to a
-string constraint solver.
-
-Future investigations involve two tracks. The first one is integration into
-existing string solvers (wich \Ostrich{} being a particularly promising
-candidate due to its shared use of \Princess{}), and further adaptation to that
-use case. Closer inspection of the instances where we currently time out should
-be useful to further improve our heuristics.
-
-The second track for future improvements is the extension into other problem
-domains, including other logics, model checking problems, as well as to more
-powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata, due to our use of a full
-automated theorem prover which allows adding arbitrary constraints in addition
-to the expected Presburger formulae.
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-%% Bibliography
-\bibliography{bibliography}
-
-%% Appendix
-\appendix
-\section{Appendix}
-\input{appendix.tex}
-\end{document}

### 2023-06-27 main.tex
Various small fixes
Closes: #43
Closes: #903
Closes: #32
Evaluation is now a separate file

@@
 large (and therefore slow) products until we have to, banking our hopes on
-computing an empty intermittent product early. The pseudocode for our
-implementation can be seen in \cref{alg:baseline}.
+computing an empty intermittent product early.
 
@@
 \section{Evaluation}\label{sec:experiments}
-
-We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
-automata intersection problems generated by the \OstrichPlus{} string constraint
-solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
-initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
-five seconds by baseline.
-We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial.
-For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s.
-The benchmarks are run on commit~\texttt{\commit}.
-
-The benchmarks are executed in parallel using GNU Parallel~\cite{parallel},
-since they are mostly single-threaded and initial results showed negligible
-interference on performance, on \BenchmarkRig{}. We compiled the code using
-Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a
-maximum heap of~\MaxHeapSize{}. We used \Nuxmv{} version~\NuxmvVersion{} invoked
-as a subprocess for each instance. Instances were executed in batches of
-\BatchSize{}, each given a fresh JVM. We believe this represents a realistic use
-case where \Calculus{} is used to support e.g a string solver. Experiments were
-executed in random order for all backends. Each instance got a time budget
-of~\RuntimeTimeout.
-
-All runtimes are measured in wall-clock time as observed by the JVM when
-executing the instance, and exclude time spent parsing (usually far below
-\numprint{0.1}s).
-
-\subsection{Execution Time and Ability to Solve Instances}\label{sec:runtime}
-
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
-the respective back-ends could solve and with which status. A full summary of
-their outcomes is also available in \cref{tab:solve-status}. Note that we lack
-ground truth for all instances. We see that \Calculus{} generally outperforms
-\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
-instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
-kind of instance, but most of all on satisfiable instances, and \Calculus{}
-outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
-\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
-caused by a non-deterministically triggered bug in the clause learning of
-\Calculus{}, but have not had time to further investigate it. We have however
-validated all \numprint{3908} satisfying assignments from \Calculus{} with
-\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
-our implementation.
-
-Baseline performs worse on satisfiable instances because it executes a heuristic
-meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances. The heuristic is enabled since it is unrealistic to assume that we
-know in advance whether an instance is satisfiable or not, but we do know that
-unsatisfiable instances are more common.
-
-\begin{table}
-  \centering
-  \input{graphs/solved_pivot_table.tex}
-  \caption{The result of running the respective back-ends with a timeout of
-  \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
-  from the table. }\label{tab:solve-status}
-\end{table}
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
-    \caption{The division of statuses per backend.}
-    \label{fig:solve-division}
-    \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
-Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
-    \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
-    \label{fig:runtime-boxplot}
-    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline.
-The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy.
-On the other hand, lazy has  a lot of outliers.}
-  \end{subfigure}
-  \vfill
-  \begin{subfigure}[b]{0.75\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
-    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
-    \label{fig:cactus}
-    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
-Lazy solves much more instances than nuxmv.}
-  \end{subfigure}
-\end{figure}
-
-\subsubsection{Scalability}\label{sec:scaling}
-
-A cactus plot showing the number of instances solved within a given timeout for
-each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
-\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
-long-term scaling properties.
-
-\iffalse
-\subsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
-
-For baseline and \Calculus{}, \Catra{} offers the ability to find the equivalent Presburger formula representing a given instance.
-For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} we use the specially tailored approach described in \cref{sec:finding-the-image}.
-For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in \cref{sec:scaling,sec:runtime}.
-
-To make sure baseline puts up as much competition as possible, we disable
-checking intermittent satisfiability and configure \Catra{} to run in the
-maximally eager mode where the product is first computed before any
-satisifiability check is performed. We run the experiments with a timeout
-of~\ImageTimeout{}. The results of the experiment is summarised in
-\cref{fig:cactus:image} and \cref{tab:image-results}. \Fudge{We see here that
-something happens}.
-
-\begin{figure}[ht]
-  \caption{The number of instances \Catra{} was able to find the Presburger form of the image for within a given number of seconds per backend.}
-  \label{fig:cactus:image}
-\end{figure}
-\fi
-
-\subsection{Threats to Validity}
-
-The most obvious threats to validity would be poor benchmarking or poor
-implementation, e.g. if the method described in \cref{sec:calculus,sec:multiple}
-deviate from what is actually benchmarked in section \cref{sec:experiments}, or
-if the methods used for benchmarking would be unsound.  To increase the
-probability that our results are representative both in the sense of
-representing  on expected inputs and in the statistical sense for a given run of
-\Catra{} despite the use of randomness in our implementation we execute many
-experiments. Moreover, to address the issue of correctness we have validated all
-reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
-is indeed sound.
-
-Another threat to validity would be if the competition (baseline and \Nuxmv)
-would be disadvantaged in our comparison. For \Nuxmv{} we use the default
-configuration which we believe should be performant (or it should not be the
-default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
-easy for artefact reviewers. For baseline, our best argument is that the
-solution we use is close to the one found in the \Ostrich{} string solver and
-that it therefore should be realistic.
-
-The greatest threat to validity is our choice of implementation platform and
-automata library. There are some signs that product computation is inefficient,
-notably the good performance of low-threshold automata materialisation that
-prioritises computing smaller products. This means that \Calculus{} makes less
-heavy use of product computation than baseline does, instead relying more on
-\Princess{}. This situation would unfairly advantage \Calculus{} if our
-automata library was the bottleneck due to our automata library being
-unnaturally poor. We believe this is unlikely since similar performance issues
-have been reported for string solvers. Additionally, profiling suggests that
-both \Princess{}-based backends spend most of their time in \Princess{},
-suggesting that the automaton implementation is not the bottleneck.
+\input{evaluation.tex}
 

### 2023-06-27 main.tex
Multiple automata are now a section

@@
 \section{Parikh Images from Products of Automata}\label{sec:multiple}
-
-We now generalise our calculus to natively work with intersections of
-regular languages, or equivalently products of automata. For this
-extension, we change the main predicate~$\Image$ to be indexed by a
-vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
-simplicity, we assume that the sets of states of the $k$ automata (and
-therefore also the transition sets) are pairwise disjoint. 
-\begin{definition}\label{def:multiple}
-  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
-  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
-  $\Monoid$,
-  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
-  is a transition selection function, and
-  $\MonoidElement \in \Monoid$. The predicate
-  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
-      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
-  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
-  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
-  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
-  \begin{itemize}
-  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
-    $\Filter$, that is,
-    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
-  \item the automata all accept the same
-    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
-  \item the accepted word is mapped to
-    $\MonoidElement = \Map(\WordOf(\Path_i)) =
-    \Map(\WordOf(\Path_1))$.
-  \end{itemize}
-\end{definition}
-
-  \begin{table}[t]
-    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
-      Name & \text{Rule} & Side conditions\\
-      \midrule
-
-      % EXPAND
-      \ExpandM & 
-      \inferrule
-      {
-       {\begin{array}{c}
-           \left\{
-           \FlowEq(\Automaton_i, \Filter),~~
-           \Connected(\Automaton_i, \Filter),~~
-           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
-           \right\}_{i=1}^k,
-           \\[4ex]
-          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
-          \SomeInequalities, \SomeClause
-        \end{array}}
-        }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
-      None \\[5ex]
-      \Materialise &
-      \inferrule
-      {{
-         \begin{array}{c}
-           \FlowEq(\Automaton', \Filter'),~
-           \Connected(\Automaton', \Filter'),~
-           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
-           \\\ImagePredicate{
-                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
-                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
-                     \Automaton'},~ \SomeInequalities, \SomeClause
-           }{\Map}{\Filter'}{\MonoidElement}
-           \end{array}}
-           }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
-      $\begin{array}{@{}l@{}}
-         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
-         1 \leq i < j \leq k, \\
-      \Automaton' = \Automaton_i \times \Automaton_j,\\
-      \end{array}$ \\
-      \bottomrule
-      \end{tabular}
-      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
-    \end{table}
-
-    For the calculus (\cref{tbl:rules:multi}), we first extend
-    $\Expand$ to generate flow equations and instances of $\Connected$
-    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
-    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
-    since it is needed to keep track of the currently considered
-    partial products.
-
-The rule~$\Materialise$ introduces the product of two individual
-automata~$\Automaton_i, \Automaton_j$; this step eliminates
-$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
-instead adds the product~$\Automaton_i \times \Automaton_j$, while
-also introducing the flow equations and the $\Connected$ predicate.
-
-The rule~$\Materialise$ also has to connect the newly introduced
-product~$\Automaton_i \times \Automaton_j$ to the previous automata
-$\Automaton_i, \Automaton_j$. This is done by extending the
-selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
-to fresh variables~$x_{\Transition}$:
-\begin{equation*}
-  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
-  ~~=~~
-  \begin{cases}
-    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
-    \\
-    \Filter(t) & \text{otherwise}
-  \end{cases}
-\end{equation*}
-
-The multiplicity of transitions in the product then has to be related
-to the multiplicities in the individual automata, modelled using the
-$\BindingSum$ predicate. The predicate expresses that the multiplicity
-of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
-$\Automaton_i$ has to coincide with the sum of the multiplicities of
-transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
-and similarly for $\Automaton_j$:
-%
-  $$
-  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
-  \begin{aligned}
-  & \Set{ 
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
-  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
-  & \Set{
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
-  }
-  \end{aligned}
-$$
-
-\iffalse
-Note that this definition implies that $\Filter(\Transition) =
-\Filter(\Transition')$ whenever two transitions $\Transition \in
-\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
-produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
-\times \Automaton_2}$. This corresponds to our intuition that the terms of the
-product must agree on the value they accept. As before, we implicitly map
-$\Filter$ to fresh terms for each transition in the product.
-\fi
-
-For instances of precisely one automaton, neither rule applies and we
-perform the calculus as before.
-
-\subsection{An Example}\label{sec:multiple:example}
-
-We return again to our example in \cref{sec:introduction:parikh}, where we
-compute the whole Parikh image of the product of $\AcaOrBc{}$ and
-$\SomethingCSomething$ under the constraint that there are more instances of
-letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
-element-wise addition, and $\Map$ that maps each transition to the corresponding
-increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.
-
-In the interest of space, we refer back to
-\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
-$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
-substitutions as follows. Note the expansion of the $\Sigma$ labels:
-    \begin{equation*}
-      \begin{aligned}
-        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-        % Other automaton
-        \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
-        \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
-      \end{aligned}
-    \end{equation*}
-
-In the derivation tree of \cref{fig:derivation:multi} we start as before, but
-with the product version of the $\Image$ predicate. The only possible rule here
-is $\ExpandM$, which we use to add the corresponding constraints on each
-automata of the product as we would have had in the single-automaton version. We
-perform algebraic reasoning on those equations, along with the constraint on
-$\MonoidElement$ to determine bounds on transition variables that will enable us
-to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
-finished doing so, we use \Materialise{} to compute the product of the now
-filtered automata. Further algebraic reasoning allows us to remove the final
-instance of $\Connected$. We then perform \Expand{} to get rid of the final
-$\Image$ instance, and are immediately able to $\Subsume$ the resulting
-$\Connected{}$ predicate. This, again, leaves us with a set of linear
-inequalities at the leaf that we can use to obtain a model of our final
-$\MonoidElement$ values (now a vector). This time we can read out aca, or
-$\VectorLiteral{2, 0, 1}$.
-
-\begin{figure}
-  \centering
-\begin{prooftree}
-  \hypo{
-    {\begin{array}{lll}
-      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
-      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
-    \end{array}}  
-  }
-  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
-    {\begin{array}{ll}
-      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
-      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
-    \end{array}}  
-  }
-  \infer1[\EquationReasoning{}]{
-  {  \begin{array}{ll}
-      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
-      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
-        \TransitionVar_6 + \TransitionVar_4 = 0 \\
-        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
-        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
-        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
-          \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1} 
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} \\ \land 
-      \Connected(\SomethingCSomething{}, \Filter) &
-      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array} } 
-  }
-  \infer1[\Materialise]{
-    {\begin{array}{ll}
-      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
-      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
-      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
-      \land 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} & \land
-      \Connected(\SomethingCSomething{}, \Filter) \\
-      \land 
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-  {\begin{array}{ll}
-    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-    1 - \TransitionVar_6 > 0 &
-    \land
-    \TransitionVar_6 = 0 \land 
-    \TransitionVar_4 = 0 \\
-    \land
-    \TransitionVar_3 > 0 & \land
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
-    \land
-    \VectorLiteral{
-      \TransitionVar_{7a} + \TransitionVar_{9a},
-      0,
-      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-      = \VectorLiteral{
-        2,
-        0,
-        \TransitionVar_3} & \land 
-    \Connected(\AcaOrBc{}, \Filter) \\
-    \land 
-    \Connected(\SomethingCSomething{}, \Filter) \\
-    \land 
-    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-  \end{array}}
-  }
-  \infer1[\EquationReasoning]{
-    \begin{aligned}
-      \VectorLiteral{a, b, c} = \VectorLiteral{
-          2 - 2\TransitionVar_6,
-          2\TransitionVar_4 + 3\TransitionVar_6,
-          \TransitionVar_3 + \TransitionVar_6
-        } \land
-        \\
-        \VectorLiteral{a, b, c} = \VectorLiteral{
-            \TransitionVar_{7a} + \TransitionVar_{9a},
-            \TransitionVar_{7b} + \TransitionVar_{9b},
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          } \land \\
-      \Connected(\AcaOrBc{}, \Filter) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{a, b, c}) \land a > b
-    \end{aligned}
-  }
-  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
-\end{prooftree}
-\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
-\end{figure}
-
-\subsection{Correctness of \Calculus{} for Products of Automata}
-
-Since \cref{tbl:rules:multi} only extends the existing rules of
-\cref{tbl:rules:single}, we focus on the differences compared
-to the calculus for a single automaton.
-
-%\subsubsection{\Calculus{} for products of automata terminates}
-\begin{lemma}\label{lma:multi-terminates}
-  Suppose $\SomeClause{}$ is a set of formulas in which the product
-  version of $\Image$ only occurs positively. There is no
-  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
-  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
-  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
-\end{lemma}
-
-\begin{proof}
-  The rule \Materialise{} can similarly only be used finitely many times, as
-  each application reduces the number of automata in the product of $\Image$ by
-  one automaton, until only one remains and \cref{lma:single-terminates} for
-  single-automaton instances applies.
-
-  This implies that also the rule~\ExpandM{} can only be applied
-  finitely often since its side condition only allows applying it once
-  per instance of an $\Image$ predicate containing a product, and only finitely
-  many instances of $\Image$ can be introduced on each branch.
-\end{proof}
-
-%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
-
-Since our calculus now includes a rule introducing new variables, the
-\Materialise{} rule, we have to slightly generalise the notion of
-solution-preservation:
-%
-\begin{lemma}\label{lma:multi-correct}
-  Consider an application of one of the rules in
-  \cref{tbl:rules:multi}, with
-  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
-  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
-  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
-  there is an extension~$\beta' of \beta$ satisfying one of the
-  premises~$\SomeClause_i$.
-\end{lemma}
-
-\begin{proof}
-  We have to consider the two new rules in \cref{tbl:rules:multi}. The
-  result is immediate for \ExpandM{}, since this rule does not remove the
-  $\Image$ predicate from a proof goal, and the newly introduced formulas
-  are all implied by the $\Image$ predicate.
-
-  For \Materialise{}, observe that the existence of an accepting path
-  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
-  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
-  same word. The path in the product will satisfy the flow equations
-  and connectedness, and it will be related to the individual paths as
-  stipulated by the \BindingSum{} predicate.
-\end{proof}
+\input{multiple.tex}
 

### 2023-06-27 multiple.tex
Multiple automata are now a section

We now generalise our calculus to natively work with intersections of
regular languages, or equivalently products of automata. For this
extension, we change the main predicate~$\Image$ to be indexed by a
vector of automata~$\Tuple{\Automaton_1, \ldots, \Automaton_k}$.  For
simplicity, we assume that the sets of states of the $k$ automata (and
therefore also the transition sets) are pairwise disjoint. 
\begin{definition}\label{def:multiple}
  Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
  $\Monoid$,
  $\Filter: \bigcup_{i=1}^k \Transitions_{\Automaton_i} \to \Naturals$
  is a transition selection function, and
  $\MonoidElement \in \Monoid$. The predicate
  $\ImagePredicate{\Tuple{\Automaton_1, \ldots,
      \Automaton_k}}{\Map}{\Filter}{\MonoidElement}$ is true exactly
  when there are accepting paths~$\Path_1, \ldots, \Path_k$ of the
  respective automata, such that for each $i \in \{1, \ldots, k\}$ and
  $\Transition \in \Transitions_{\Automaton_i}$ it holds that
  \begin{itemize}
  \item the multiplicity of $\Transition$ on $\Path_i$ is consistent with
    $\Filter$, that is,
    $\Filter(\Transition) = \TransitionCount(\Transition, \Path_i)$,
  \item the automata all accept the same
    word~$\WordOf(\Path_i) = \WordOf(\Path_1)$, and
  \item the accepted word is mapped to
    $\MonoidElement = \Map(\WordOf(\Path_i)) =
    \Map(\WordOf(\Path_1))$.
  \end{itemize}
\end{definition}

  \begin{table}[t]
    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
      Name & \text{Rule} & Side conditions\\
      \midrule

      % EXPAND
      \ExpandM & 
      \inferrule
      {
       {\begin{array}{c}
           \left\{
           \FlowEq(\Automaton_i, \Filter),~~
           \Connected(\Automaton_i, \Filter),~~
           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
           \right\}_{i=1}^k,
           \\[4ex]
          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
          \SomeInequalities, \SomeClause
        \end{array}}
        }
      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
      None \\[5ex]
      \Materialise &
      \inferrule
      {{
         \begin{array}{c}
           \FlowEq(\Automaton', \Filter'),~
           \Connected(\Automaton', \Filter'),~
           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
           \\\ImagePredicate{
                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
                     \Automaton'},~ \SomeInequalities, \SomeClause
           }{\Map}{\Filter'}{\MonoidElement}
           \end{array}}
           }
      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
      $\begin{array}{@{}l@{}}
         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
         1 \leq i < j \leq k, \\
      \Automaton' = \Automaton_i \times \Automaton_j,\\
      \end{array}$ \\
      \bottomrule
      \end{tabular}
      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
    \end{table}

    For the calculus (\cref{tbl:rules:multi}), we first extend
    $\Expand$ to generate flow equations and instances of $\Connected$
    for each automaton, resulting in a new rule $\ExpandM$.  Unlike
    \Expand{}, \ExpandM{} does not remove the $\Image$~predicate,
    since it is needed to keep track of the currently considered
    partial products.

The rule~$\Materialise$ introduces the product of two individual
automata~$\Automaton_i, \Automaton_j$; this step eliminates
$\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
instead adds the product~$\Automaton_i \times \Automaton_j$, while
also introducing the flow equations and the $\Connected$ predicate.

The rule~$\Materialise$ also has to connect the newly introduced
product~$\Automaton_i \times \Automaton_j$ to the previous automata
$\Automaton_i, \Automaton_j$. This is done by extending the
selection function~$\Filter$ to $\Filter'$, mapping the transitions of the product
to fresh variables~$x_{\Transition}$:
\begin{equation*}
  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
  ~~=~~
  \begin{cases}
    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
    \\
    \Filter(t) & \text{otherwise}
  \end{cases}
\end{equation*}

The multiplicity of transitions in the product then has to be related
to the multiplicities in the individual automata, modelled using the
$\BindingSum$ predicate. The predicate expresses that the multiplicity
of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
$\Automaton_i$ has to coincide with the sum of the multiplicities of
transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
and similarly for $\Automaton_j$:
%
  $$
  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
  \begin{aligned}
  & \Set{ 
    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
  & \Set{
    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
  }
  \end{aligned}
$$

\iffalse
Note that this definition implies that $\Filter(\Transition) =
\Filter(\Transition')$ whenever two transitions $\Transition \in
\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
\times \Automaton_2}$. This corresponds to our intuition that the terms of the
product must agree on the value they accept. As before, we implicitly map
$\Filter$ to fresh terms for each transition in the product.
\fi

For instances of precisely one automaton, neither rule applies and we
perform the calculus as before.

\subsection{An Example}\label{sec:multiple:example}

We return again to our example in \cref{sec:introduction:parikh}, where we
compute the whole Parikh image of the product of $\AcaOrBc{}$ and
$\SomethingCSomething$ under the constraint that there are more instances of
letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
element-wise addition, and $\Map$ that maps each transition to the corresponding
increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.

In the interest of space, we refer back to
\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
substitutions as follows. Note the expansion of the $\Sigma$ labels:
    \begin{equation*}
      \begin{aligned}
        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
        % Other automaton
        \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
        \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
        \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
      \end{aligned}
    \end{equation*}

In the derivation tree of \cref{fig:derivation:multi} we start as before, but
with the product version of the $\Image$ predicate. The only possible rule here
is $\ExpandM$, which we use to add the corresponding constraints on each
automata of the product as we would have had in the single-automaton version. We
perform algebraic reasoning on those equations, along with the constraint on
$\MonoidElement$ to determine bounds on transition variables that will enable us
to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
finished doing so, we use \Materialise{} to compute the product of the now
filtered automata. Further algebraic reasoning allows us to remove the final
instance of $\Connected$. We then perform \Expand{} to get rid of the final
$\Image$ instance, and are immediately able to $\Subsume$ the resulting
$\Connected{}$ predicate. This, again, leaves us with a set of linear
inequalities at the leaf that we can use to obtain a model of our final
$\MonoidElement$ values (now a vector). This time we can read out aca, or
$\VectorLiteral{2, 0, 1}$.

\begin{figure}
  \centering
\begin{prooftree}
  \hypo{
    {\begin{array}{lll}
      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
    \end{array}}  
  }
  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
    {\begin{array}{ll}
      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
    \end{array}}  
  }
  \infer1[\EquationReasoning{}]{
  {  \begin{array}{ll}
      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
      \land \Image{}_{\Automaton', \Map}(\Filter, 
      \VectorLiteral{
        \TransitionVar_{7a} + \TransitionVar_{9a},
        0,
        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
        \TransitionVar_6 + \TransitionVar_4 = 0 \\
        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
          \VectorLiteral{
        \TransitionVar_{7a} + \TransitionVar_{9a},
        0,
        \TransitionVar_{7c} + \TransitionVar_{9c} + 1} 
        = \VectorLiteral{
          2,
          0,
          \TransitionVar_3} \\ \land 
      \Connected(\SomethingCSomething{}, \Filter) &
      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
    \end{array} } 
  }
  \infer1[\Materialise]{
    {\begin{array}{ll}
      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
      \land 
      \VectorLiteral{
        \TransitionVar_{7a} + \TransitionVar_{9a},
        0,
        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
        = \VectorLiteral{
          2,
          0,
          \TransitionVar_3} & \land
      \Connected(\SomethingCSomething{}, \Filter) \\
      \land 
      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
    \end{array}}
  }
  \infer1[\Subsume{}]{
  {\begin{array}{ll}
    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
    1 - \TransitionVar_6 > 0 &
    \land
    \TransitionVar_6 = 0 \land 
    \TransitionVar_4 = 0 \\
    \land
    \TransitionVar_3 > 0 & \land
    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
    \land
    \VectorLiteral{
      \TransitionVar_{7a} + \TransitionVar_{9a},
      0,
      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
      = \VectorLiteral{
        2,
        0,
        \TransitionVar_3} & \land 
    \Connected(\AcaOrBc{}, \Filter) \\
    \land 
    \Connected(\SomethingCSomething{}, \Filter) \\
    \land 
    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
  \end{array}}
  }
  \infer1[\EquationReasoning]{
    \begin{aligned}
      \VectorLiteral{a, b, c} = \VectorLiteral{
          2 - 2\TransitionVar_6,
          2\TransitionVar_4 + 3\TransitionVar_6,
          \TransitionVar_3 + \TransitionVar_6
        } \land
        \\
        \VectorLiteral{a, b, c} = \VectorLiteral{
            \TransitionVar_{7a} + \TransitionVar_{9a},
            \TransitionVar_{7b} + \TransitionVar_{9b},
            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
          } \land \\
      \Connected(\AcaOrBc{}, \Filter) \land 
      \Connected(\SomethingCSomething{}, \Filter) \land \\
      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
      \VectorLiteral{a, b, c}) \land a > b
    \end{aligned}
  }
  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
\end{prooftree}
\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
\end{figure}

\subsection{Correctness of \Calculus{} for Products of Automata}

Since \cref{tbl:rules:multi} only extends the existing rules of
\cref{tbl:rules:single}, we focus on the differences compared
to the calculus for a single automaton.

%\subsubsection{\Calculus{} for products of automata terminates}
\begin{lemma}\label{lma:multi-terminates}
  Suppose $\SomeClause{}$ is a set of formulas in which the product
  version of $\Image$ only occurs positively. There is no
  infinite sequence of proofs~$P_0, P_1, P_2, \ldots$ in which $P_0$
  has $\SomeClause{}$ as root, and each $P_{i+1}$ is derived from
  $P_i$ by applying one of the rules in \cref{tbl:rules:multi}.
\end{lemma}

\begin{proof}
  The rule \Materialise{} can similarly only be used finitely many times, as
  each application reduces the number of automata in the product of $\Image$ by
  one automaton, until only one remains and \cref{lma:single-terminates} for
  single-automaton instances applies.

  This implies that also the rule~\ExpandM{} can only be applied
  finitely often since its side condition only allows applying it once
  per instance of an $\Image$ predicate containing a product, and only finitely
  many instances of $\Image$ can be introduced on each branch.
\end{proof}

%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}

Since our calculus now includes a rule introducing new variables, the
\Materialise{} rule, we have to slightly generalise the notion of
solution-preservation:
%
\begin{lemma}\label{lma:multi-correct}
  Consider an application of one of the rules in
  \cref{tbl:rules:multi}, with
  premises~$\SomeClause_1, \ldots, \SomeClause_k$ and
  conclusion~$\SomeClause$. An assignment~$\beta$ (over the symbols in
  $\SomeClause$) satisfies the conclusion~$\SomeClause$ if and only if
  there is an extension~$\beta' of \beta$ satisfying one of the
  premises~$\SomeClause_i$.
\end{lemma}

\begin{proof}
  We have to consider the two new rules in \cref{tbl:rules:multi}. The
  result is immediate for \ExpandM{}, since this rule does not remove the
  $\Image$ predicate from a proof goal, and the newly introduced formulas
  are all implied by the $\Image$ predicate.

  For \Materialise{}, observe that the existence of an accepting path
  in $\Automaton_i \times \Automaton_j$ is equivalent to the existence
  of individual paths in $\Automaton_i, \Automaton_j$ accepting the
  same word. The path in the product will satisfy the flow equations
  and connectedness, and it will be related to the individual paths as
  stipulated by the \BindingSum{} predicate.
\end{proof}

### 2023-06-27 evaluation.tex
Evaluation

@@
 initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
-five seconds by baseline.
-We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial.
-For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s.
-The benchmarks are run on commit~\texttt{\commit}.
+five seconds by baseline. We also attempted to benchmark instances generated by
+\Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len
+(\numprint{791} instances) benchmark suites, but had to discard them since they
+were all trivial. For instance, on a laptop, \Calculus{} could solve every
+instance from kaluza-len in under five seconds, mean about 0.1s. The benchmarks
+are run on commit~\texttt{\commit}.
 
-The benchmarks are executed in parallel using GNU Parallel~\cite{parallel},
-since they are mostly single-threaded and initial results showed negligible
-interference on performance, on \BenchmarkRig{}. We compiled the code using
-Scala~\ScalaVersion{}, and executed the experiments on~\JvmVersion{} with a
-maximum heap of~\MaxHeapSize{}. We used \Nuxmv{} version~\NuxmvVersion{} invoked
-as a subprocess for each instance. Instances were executed in batches of
-\BatchSize{}, each given a fresh JVM. We believe this represents a realistic use
-case where \Calculus{} is used to support e.g a string solver. Experiments were
-executed in random order for all backends. Each instance got a time budget
-of~\RuntimeTimeout.
+The benchmarks are executed in parallel across a cluster of \Fudge{identical}
+machines, \BenchmarkRig{}, but as the sole job for each machine.  We compiled
+the code using Scala~\ScalaVersion{}, and executed the experiments
+on~\JvmVersion{} with a maximum heap of~\MaxHeapSize{}. We used \Nuxmv{}
+version~\NuxmvVersion{} invoked as a subprocess for each instance. Instances
+were executed in batches of \BatchSize{}, each given a fresh JVM. Each JVM was
+warmed up for \numprint{10}s on a random benchmark from the set before starting
+to execute. We believe this represents a realistic use case where \Calculus{} is
+used to support e.g a string solver. Experiments were executed in random order
+for all backends. Each instance got a time budget of~\RuntimeTimeout.
 
@@
 
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
-the respective back-ends could solve and with which status. A full summary of
-their outcomes is also available in \cref{tab:solve-status}. Note that we lack
-ground truth for all instances. We see that \Calculus{} generally outperforms
-\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
-instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
-kind of instance, but most of all on satisfiable instances, and \Calculus{}
-outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
-\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
-caused by a non-deterministically triggered bug in the clause learning of
-\Calculus{}, but have not had time to further investigate it. We have however
-validated all \numprint{3908} satisfying assignments from \Calculus{} with
-\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
-our implementation.
+In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances
+the respective back-end could solve. A summary of their outcomes by instance
+type is also available in \cref{tab:solve-status}. Note that many instances lack
+a corum ground truth as they are solved by only one backend. We see that
+\Calculus{} generally outperforms \Nuxmv{} on determining unsatisfiability,
+while being similar at satisfiable instances. Both \Nuxmv{} and \Calculus{}
+outperform baseline on satisfiable instances, and \Calculus{} outperforms
+\Nuxmv{}.
 
@@
 meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances. The heuristic is enabled since it is unrealistic to assume that we
-know in advance whether an instance is satisfiable or not, but we do know that
-unsatisfiable instances are more common.
+instances roughly similar to \cite{approximate-parikh}. The heuristic is enabled
+since we suspect that unsatisifiable instances are more common; at least we have
+seen more solved unsatisfiable instances.
 
@@
   \input{graphs/solved_pivot_table.tex}
-  \caption{The result of running the respective back-ends with a timeout of
-  \RuntimeTimeout. Instances solved by no backend within the timeout are omitted
-  from the table. }\label{tab:solve-status}
+  \caption{Number of successful results within a timeout of \RuntimeTimeout{}.
+  Instances solved by no backend within the timeout (about half of the set) are
+  omitted from the table.}\label{tab:solve-status}
 \end{table}
@@
 
-\subsubsection{Scalability}\label{sec:scaling}
+\subsubsection{Scalability}\label{sec:scaling}%
+To get an idea of how the different backends scale, we additionally execute
+\numprint{2000} different randomly sampled benchmarks with a
+\numprint{120}-second timeout. In this set we also include two additional
+variants of \Calculus{}: one with clause learning and restarts disabled, and one
+with only clause learning disabled. This is meant to show the impact of these
+extra features.
 
 A cactus plot showing the number of instances solved within a given timeout for
-each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
-\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
-long-term scaling properties.
+each backend configuration can be seen in \cref{fig:cactus}. Here, we see
+clearly that \Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{}
+might have better long-term scaling properties. Notably, for machines with
+better single-core performance \Nuxmv{}'s scaling benefits kicks in faster,
+though the graph looks similar in shape.
 
@@
 reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
-is indeed sound.
+is indeed sound. A previous version contained a race condition triggered by
+random restarts during product materialisation which would in about
+\numprint{0.7} of instances lead to non-deterministically unsound behaviour. We
+have observed no instances of this bug or any other soundness issue since fixing
+that bug, and have in addition to running all \NrBenchmarks{} we have
+re-executed the same instances multiple times for long durations to capture
+other soundness issues without any issue.
 
@@
 
-The greatest threat to validity is our choice of implementation platform and
-automata library. There are some signs that product computation is inefficient,
-notably the good performance of low-threshold automata materialisation that
-prioritises computing smaller products. This means that \Calculus{} makes less
-heavy use of product computation than baseline does, instead relying more on
-\Princess{}. This situation would unfairly advantage \Calculus{} if our
-automata library was the bottleneck due to our automata library being
-unnaturally poor. We believe this is unlikely since similar performance issues
-have been reported for string solvers. Additionally, profiling suggests that
-both \Princess{}-based backends spend most of their time in \Princess{},
-suggesting that the automaton implementation is not the bottleneck.
+The greatest threat to validity remaining is our choice of implementation
+platform and automata library. We There are some signs that product computation
+is inefficient, notably the good performance of low-threshold automata
+materialisation that prioritises computing smaller products. This means that
+\Calculus{} makes less heavy use of product computation than baseline does,
+instead relying more on \Princess{}. This situation would unfairly advantage
+\Calculus{} if our automata library was the bottleneck due to our automata
+library being unnaturally poor. We believe this is unlikely since similar
+performance issues have been reported for string solvers. Additionally,
+profiling suggests that both \Princess{}-based backends spend most of their time
+in \Princess{}, suggesting that the automaton implementation is not the
+bottleneck. Finally, we have compared aggregate results for both native backends
+when optimising the automata construction procedures and observed that the trend
+in performance differences has been robust as the automata library improved
+vastly.

### 2023-06-27 multiple.tex
Evaluation

@@
 
-We return again to our example in \cref{sec:introduction:parikh}, where we
-compute the whole Parikh image of the product of $\AcaOrBc{}$ and
-$\SomethingCSomething$ under the constraint that there are more instances of
-letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
-element-wise addition, and $\Map$ that maps each transition to the corresponding
-increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.
+We return again to our example in \cref{sec:intuition} and start by performing
+the steps of \cref{sec:single:example} shown in \cref{fig:derivation:single},
+since the steps to propagate \cref{ex:string-constraints} across the automaton
+$A'$ still applies for the product. We also reuse the definition of $\Map$ as
+the simple projection to extract the counter increments, so in this case,
+$\Map(\Transition) = \left[l_{B+}\right]$ for every transition~$\Transition$
+since we are only counting the length.
 
-In the interest of space, we refer back to
-\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
-$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
-substitutions as follows. Note the expansion of the $\Sigma$ labels:
-    \begin{equation*}
+We reuse the transition variable associations of our previous
+\cref{sec:single:example}, from \cref{eq:single-filter}, but add the definitions
+for $B$ shown in \cref{eq:multi-filter}:
+\begin{equation}\label{eq:multi-filter}
     \begin{aligned}
-        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
-        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
-        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
-        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
-        % Other automaton
-        \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
-        \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
-        \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
+        \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6\\
+        \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7\\
+        \Filter(\FromLabelTo{B}{b / l_{B+}}{S})= \TransitionVar_8 \\
+        \Filter(\FromLabelTo{B}{c / l_{B+}}{B})= \TransitionVar_9\\
+        \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}\\
+        \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}\\
+        \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}\\
+        \FlowEq(B, \Filter) =  % FIXME FLOWS FOR B \\
     \end{aligned}
-    \end{equation*}
+\end{equation}
 
@@
 is $\ExpandM$, which we use to add the corresponding constraints on each
-automata of the product as we would have had in the single-automaton version. We
-perform algebraic reasoning on those equations, along with the constraint on
+automata of the product as we would have had in the single-automaton version.
+
+We perform algebraic reasoning on those equations, along with the constraint on
 $\MonoidElement$ to determine bounds on transition variables that will enable us
@@
   }
-  \infer1[\ExpandM]{\Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
+  \infer1[\ExpandM]{\Image{}_{\Tuple{A', B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})}
 \end{prooftree}

### 2023-06-27 evaluation.tex
Some implementation

@@
 variants of \Calculus{}: one with clause learning and restarts disabled, and one
-with only clause learning disabled. This is meant to show the impact of these
-extra features.
+with only clause learning disabled to show the impact of
+\cref{sec:random-restarts}, and \cref{sec:clause-learning} respectively.
 

### 2023-06-27 implementation.tex
Some implementation

--- 2023-06-27 main.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2023}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\usepackage{amsmath,empheq,fancybox}
-\usepackage{paralist}
-\usepackage{url}
-\usepackage{color}
-\usepackage{textcomp,listings}
-\usepackage{array}
-\usepackage{mymacros}
-\usepackage{microtype}
-\usepackage{listings}
-\usepackage{csquotes}
-\usepackage{proof}
-\usepackage[capitalise]{cleveref}
-\usepackage{algorithm2e}      
-\usepackage{multirow}
-\usepackage{mathpartir}
-\usepackage{amsthm}
-\usepackage{numprint}
-\usepackage{ebproof}
-\usepackage[makeroom]{cancel}
-
-\usepackage{mathtools} % Bonus
-
-\usepackage{enumitem}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{example}{Example}[section]
-
-\newlist{constraints}{enumerate}{3}
-\setlist[constraints,1]{%
-    label=\sffamily{(\roman*):},
-    ref=\normalfont{Constraint (\roman*)},
-    wide,itemsep=0pt,topsep=0pt
-}    
-\crefname{constraintsi}{}{}
-\Crefname{constraintsi}{}{}
-
-\lstset{
-    columns=fullflexible,
-    showspaces=false,
-    showtabs=false,
-    breaklines=true,
-    showstringspaces=false,
-    breakatwhitespace=true,
-    escapeinside={(*@}{@*)},
-    commentstyle=\color{greencomments},
-    keywordstyle=\color{bluekeywords},
-    stringstyle=\color{redstrings},
-    numberstyle=\color{graynumbers},
-    basicstyle=\ttfamily\small,
-    framesep=12pt,
-    xleftmargin=12pt,
-    tabsize=4,
-    captionpos=b
-}
-
-\newif\ifcomments
-\commentstrue
-\newif\ifoutline
-\outlinetrue
-
-\newcommand{\contents}[1]{\ifoutline{\color{blue}
-    \begin{itemize}
-    #1
-    \end{itemize}
-  }\fi}
-
-\allowdisplaybreaks[1]
-
-\begin{document}
-
-%% Title information
-\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
-   counts in strings of a given language. Automata-based string solvers
-   frequently need to compute the Parikh image of large products (intersections)
-   of nondeterministic automata, which in many operations is both prohibitively
-   slow and memory-intensive. We contribute a novel understanding of how Parikh
-   maps can be tackled as a constraint solving problem to solve real-world
-   constraints stemming from functions on regular languages, most notably the
-   length constraint. Furthermore, we show how this formulation can be
-   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
-   prover supporting Presburger logic.  We implement \Calculus{} in a tool
-   called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving Parikh automata
-   intersection problems produced when solving standard string constraint
-   benchmarks involving string length constraints. We show that our solution
-   strictly outperforms the standard approach described
-   in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wide
-   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
-   model checker.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-  <ccs2012>
-  <concept>
-  <concept_id>10003752.10003766.10003773.10003775</concept_id>
-  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003766.10003776</concept_id>
-  <concept_desc>Theory of computation~Regular languages</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10003794</concept_id>
-  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10010124.10010138.10010142</concept_id>
-  <concept_desc>Theory of computation~Program verification</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10011192</concept_id>
-  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  </ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Quantitative automata}
-\ccsdesc[300]{Theory of computation~Regular languages}
-\ccsdesc[500]{Theory of computation~Automated reasoning}
-\ccsdesc[300]{Theory of computation~Program verification}
-\ccsdesc[300]{Theory of computation~Verification by model checking}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}\label{sec:introduction}
-\input{introduction.tex}
-
-\section{An Intuition for Our Approach}\label{sec:intuition}
-\input{example.tex}
-
-\section{Preliminaries}\label{sec:preliminaries}
-\input{preliminaries.tex}
-
-\section{Projections on Parikh Images}\label{sec:generalised}
-\input{generalised.tex}
-
-\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
-\input{calculus.tex}
-
-\section{Parikh Images from Products of Automata}\label{sec:multiple}
-\input{multiple.tex}
-
-\section{Extensions}\label{sec:extensions}
-\input{extensions.tex}
-
-\section{Implementation}\label{sec:implementation}
-
 We implement \Calculus{} for Parikh automata as described in
@@
 
-\subsubsection{Clause Learning}
+\subsubsection{Clause Learning}\label{sec:clause-learning}
 
@@
 
-\subsubsection{Random Restarts}
+\subsubsection{Random Restarts}\label{sec:random-restarts}
 
@@
 tail-heavy our problem is.
-
-\section{Evaluation}\label{sec:experiments}
-\input{evaluation.tex}
-
-\section{Conclusion}
-
-In this paper we have introduced a calculus to compute commutating operations on
-intersections of regular languages that we call \Calculus{}. We have evaluated
-it on \NrBenchmarks{} Parikh automata intersection problems generated by the
-\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
-suite \cite{pyex} using our Parikh automata solver \Catra{}.
-
-Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
-usage and solve-time compared to the baseline approach laid out in
-\cite{generate-parikh-image} when implemented on the same underlying automated
-theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
-\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
-and generally outperforming it for timeouts under 30 seconds with its advantage
-increasing drastically for even shorter timeouts. 30 seconds would generally be
-considered a long timeout for our intended use as supporting infrastructure to a
-string constraint solver.
-
-Future investigations involve two tracks. The first one is integration into
-existing string solvers (wich \Ostrich{} being a particularly promising
-candidate due to its shared use of \Princess{}), and further adaptation to that
-use case. Closer inspection of the instances where we currently time out should
-be useful to further improve our heuristics.
-
-The second track for future improvements is the extension into other problem
-domains, including other logics, model checking problems, as well as to more
-powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata, due to our use of a full
-automated theorem prover which allows adding arbitrary constraints in addition
-to the expected Presburger formulae.
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-%% Bibliography
-\bibliography{bibliography}
-
-%% Appendix
-\appendix
-\section{Appendix}
-\input{appendix.tex}
-\end{document}

### 2023-06-27 main.tex
Evaluation
Some implementation

@@
 
-\newif\ifcomments
-\commentstrue
-\newif\ifoutline
-\outlinetrue
+\newif\ifcomments%
+\commentstrue%
+\newif\ifoutline%
+\outlinetrue{}
 
@@
 \section{Implementation}\label{sec:implementation}
-
-We implement \Calculus{} for Parikh automata as described in
-\cref{sec:parikh-automata}. The artefact submitted along with this paper is a
-program that reads an instance file with one or more products of one or more
-Parikh automaton with transition labels defined as ranges of Unicode characters,
-along with a set of constraints on the final values of their registers expressed
-as Presburger arithmetic in a C-like syntax. We call this program
-\Catra.
-% \footnote{If you really must read it as an acronym, please read it as
-%CAtegory Theory on Register Automata, or if you object to the somewhat
-%nonstandard use of register automata and category theory, as Check Assignments
-%of The Registers Afterwards. Or alternatively if you find it all to be too much
-%of a theoretical exercise, as Can Anyone Think of A Real Application.}
-
-\Catra{} is written in Scala, with the calculus described in this paper
-implemented as a theory plug-in for the \Princess{} automated theorem
-prover~\cite{princess}, which also performs the Presburger reasoning. For
-comparison, we also provide an implementation of the baseline method
-from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
-symbolic model checker~\cite{nuxmv} to solve our constraints, and the
-approximation described in~\cite{approximate-parikh} on top of the standard
-baseline back-end. An example of an input file corresponding to our running
-example introduced in \cref{sec:motivation} can be seen in
-\cref{lst:input-example}.
-
-\Catra{} uses symbolic labels for automata. A symbolic label is defined as a
-finite range of Unicode code points. This allows succinct representation of many
-regular expression patterns such as \lstinline{(a-z).*} which would have
-otherwise required $27$~transitions. The transition for the range would be
-written \lstinline{a -> b [97, 122]}.
-
-In satisfaction mode, supported by all backends, \Catra{} tries to satisfy the
-constraints expressed by the input file, reporting \Sat{} with register
-assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
-Additionally, baseline and \Calculus{} also support generation of the Presburger
-formula describing the constraints of the input file. Baseline uses standard
-quantifier elimination, and \Calculus{} uses the method described in
-\cref{sec:finding-the-image}.
-
-Since \Princess{} does not support multiple-arity predicates like the ones we
-use in \Calculus{}, we have implemented variable-length arguments using
-additional helper predicates. These are $\Unused{}(\Automaton)$, which marks an
-automaton as unused in any product, and $\TransitionMask{}(\Automaton,
-\Transition, \Filter(\Transition))$ which associates a transition $\Transition$
-and automaton $\Automaton$ with its corresponding transition variable.
-Additionally, we associate each of our predicates with an instance variable in
-order to differentiate instances of the predicates.
-
-\subsection{Implementing the Baseline}\label{sec:implementing-baseline}
-
-We baseline using the same Presburger solver (\Princess{}), input file parser,
-and automaton implementation as \Catra. We do this in order to better analyse
-the impact of the calculus rules themselves. Using the formula of
-\eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
-successive term and add them to \Princess. We compute the product incrementally
-term by term, checking satisfiability at each step. We use a priority queue to
-select automata for each step, and order it by the number of transitions as a
-heuristic for the size of the automata. We use this heuristic to avoid computing
-large (and therefore slow) products until we have to, banking our hopes on
-computing an empty intermittent product early.
-
-\begin{algorithm}
-  \caption{How we implement the baseline approach}\label{alg:baseline}
-  \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
-  \KwResult{\textsc{Sat} or \textsc{Unsat}}
-  \SetKwFunction{NewTheoremProver}{newTheoremProver}
-  \SetKwFunction{NewPriorityQueue}{newPriorityQueue}
-  \SetKwFunction{Dequeue}{dequeue}
-  \SetKwFunction{Enqueue}{enqueue}
-  \SetKwFunction{Assert}{assert}
-
-  $p \gets \NewTheoremProver{}$
-
-  \Assert{$p$, $\SomeClause$}
-
-  \ForEach{$\Automaton_i$}{
-    \Assert{$p, \ParikhMap(\Automaton_i)$}
-
-    \If{$p$ is \textsc{Unsat}}{break}
-
-  }
-
-  $q \gets \NewPriorityQueue{}$
-
-  \While{$p$ not \textsc{Unsat} and $|q| > 1$}{
-    $\Automaton, \Automaton' \gets \Dequeue{q}$
-
-    \Assert{$p, \ParikhMap(\Automaton \times \Automaton')$}
-
-    \Enqueue{$q, \Automaton \times \Automaton'$}
-  }
-
-  \KwRet{$p$'s SAT status}
-
-  \end{algorithm}
-
-As an optimisation, our automata (including intermittent products) are created
-forward- and backward- reachable-minimal. Any automaton we produce only contains
-states that are both reachable from the initial state and has a path to an
-accepting state. We never perform any other minimisation on the automata for
-either backend. More complex minimisation was left out since performing
-minimisation on automata with counters is non-trivial, and minimising symbolic
-automata risks exponential blowup \cite{minimising-symbolic}.
-
-\subsection{Heuristics and search strategies}
-
-There are a number of choices left unspecified in \Calculus{} as described in
-\cref{sec:calculus,sec:multiple}. For example, the order of materialisation of
-intermediate products and the order of splitting. In this section we describe
-additional implementation details and techniques used to enhance \Catra.
-
-\subsubsection{Splitting, Materialisation, and Propagation}
-
-We order our rule applications so that we first propagate connectedness if
-possible, then perform materialisation if tractable as defined below, then
-finally resort to splitting if we must.
-
-In addition to applying \Split{} as described in \cref{tbl:rules:single} to
-randomly selected transitions, we prefer splitting to sever a strongly connected
-from the initial state. We randomly select an automaton where we can compute a
-cut between an SCC and the initial state, that is where the SCC does not contain
-the initial state and where the sum of the transition variables of the
-transitions in the cut is not known to be positive. If there are multiple such
-strongly connected components we choose one randomly. We then proceed to split
-on the sum of the transition variables of the cut as if it were a regular
-transition, e.g. its sum being zero or nonzero. In this way we drive \Calculus{}
-towards applying \Propagate{}.
-
-The implementation of the connectedness constraint is opportunistic and straightforward.
-We compute a set of dead states by performing forward and backwards reachability computations on an automaton, where we disregard any transition whose associated variable is known to be zero.
-After that we add clauses ensuring any transition variable associated with a transition starting in a dead state is zero.
-
-Product materialisation is the final piece of the puzzle. In the current
-implementation we put off computing intermediate products until at least all but
-two transition variables of one of the automata is known to be either present or
-absent. The number was chosen experimentally, and we observe a consistent trend
-towards lower numbers being better. The other automaton for the product is
-selected randomly.
-
-\subsubsection{Clause Learning}
-
-\Catra{} enables clause learning by default when using our backend, as it has
-been experimentally shown to increase the performance in aggregate (though not
-strictly). We do not currently implement all the proposed features of
-\cref{sec:ext:backjumping}, but we do implement forward-reachability cut
-learning. No sophisticated clause learning for products has been implemented.
-
-\subsubsection{Random Restarts}
-
-Finally, we perform restarts scaled by the Luby series~\cite{luby}. Experimental
-results have shown this to have a large improvement in performance, which is
-unsurprising given how many random choices we make during solving and how
-tail-heavy our problem is.
+\input{implementation.tex}
 

### 2023-06-27 multiple.tex
Progress on example

@@
         \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7\\
-        \Filter(\FromLabelTo{B}{b / l_{B+}}{S})= \TransitionVar_8 \\
+        \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B})= \TransitionVar_8 \\
         \Filter(\FromLabelTo{B}{c / l_{B+}}{B})= \TransitionVar_9\\
@@
         \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}\\
-        \FlowEq(B, \Filter) =  % FIXME FLOWS FOR B \\
+        \FlowEq(B, \Filter) =  \begin{aligned}
+            1 + \TransitionVar_{12}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
+            \land \TransitionVar_{11} &= \TransitionVar_6\\
+            \land \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}
+        \end{aligned}
     \end{aligned}
@@
 \begin{prooftree}
-  \hypo{
-    {\begin{array}{lll}
-      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
-      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
-    \end{array}}  
-  }
-  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
-    {\begin{array}{ll}
-      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
-      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
-    \end{array}}  
-  }
-  \infer1[\EquationReasoning{}]{
-  {  \begin{array}{ll}
-      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
-      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
-      \land \Image{}_{\Automaton', \Map}(\Filter, 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
-        \TransitionVar_6 + \TransitionVar_4 = 0 \\
-        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
-        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
-        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
-          \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1} 
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} \\ \land 
-      \Connected(\SomethingCSomething{}, \Filter) &
-      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array} } 
-  }
-  \infer1[\Materialise]{
-    {\begin{array}{ll}
-      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
-      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
-      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
-      \land 
-      \VectorLiteral{
-        \TransitionVar_{7a} + \TransitionVar_{9a},
-        0,
-        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-        = \VectorLiteral{
-          2,
-          0,
-          \TransitionVar_3} & \land
-      \Connected(\SomethingCSomething{}, \Filter) \\
-      \land 
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-    \end{array}}
-  }
-  \infer1[\Subsume{}]{
-  {\begin{array}{ll}
-    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
-    1 - \TransitionVar_6 > 0 &
-    \land
-    \TransitionVar_6 = 0 \land 
-    \TransitionVar_4 = 0 \\
-    \land
-    \TransitionVar_3 > 0 & \land
-    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
-    \land
-    \VectorLiteral{
-      \TransitionVar_{7a} + \TransitionVar_{9a},
-      0,
-      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
-      = \VectorLiteral{
-        2,
-        0,
-        \TransitionVar_3} & \land 
-    \Connected(\AcaOrBc{}, \Filter) \\
-    \land 
-    \Connected(\SomethingCSomething{}, \Filter) \\
-    \land 
-    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
-  \end{array}}
-  }
-  \infer1[\EquationReasoning]{
+  % LEFT TREE  ->
+  \hypo{l}% here x_12 is zero
+  % RIGHT TREE ->
+  \hypo{r} % here x_12 is one
+  \infer2[\Split{} $\TransitionVar_{12} = 0$]{\text{FINAL STEP OF STUMP}}
+  \infer1[\EquationReasoning: probably something good]{
     \begin{aligned}
-      \VectorLiteral{a, b, c} = \VectorLiteral{
-          2 - 2\TransitionVar_6,
-          2\TransitionVar_4 + 3\TransitionVar_6,
-          \TransitionVar_3 + \TransitionVar_6
-        } \land
-        \\
-        \VectorLiteral{a, b, c} = \VectorLiteral{
-            \TransitionVar_{7a} + \TransitionVar_{9a},
-            \TransitionVar_{7b} + \TransitionVar_{9b},
-            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
-          } \land \\
-      \Connected(\AcaOrBc{}, \Filter) \land 
-      \Connected(\SomethingCSomething{}, \Filter) \land \\
-      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
-      \VectorLiteral{a, b, c}) \land a > b
+      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
+      \land \TransitionVar_2 = l_A - 3\\
+      \land 1 + \TransitionVar_{12}   = \TransitionVar_{11} + \TransitionVar_{12}\\
+      \land \TransitionVar_{11} = \TransitionVar_6\\
+      \land \TransitionVar_{12} = \TransitionVar_8 + \TransitionVar_{10}\\
+      \land \Connected(B, \Filter) \\
+      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
     \end{aligned}
   }
-  \infer1[\ExpandM]{\Image{}_{\Tuple{A', B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})}
+  \infer1[\EquationReasoning: expand $\FlowEq$]{
+    \begin{aligned}
+      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
+      \land \TransitionVar_2 = l_A - 3\\
+      \land \FlowEq(B, \Filter) \\
+      \land \Connected(B, \Filter) \\
+      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
+    \end{aligned}
+  }
+  \infer1[Repeat \cref{fig:derivation:single}, $l_A = l_B$.]{
+    \begin{aligned}
+      l_B = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
+      \land i = \TransitionVar_5\\
+      \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k\\
+      \land n = \TransitionVar_4 + \TransitionVar_1\\  
+      \land \FlowEq(B, \Filter) \\
+      \land \Connected(A', \Filter) \\
+      \land \Connected(B, \Filter) \\
+      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) 
+      \land l_A = l_B\\
+      \land l_A - i - n > 0\\
+      \land 0 < i < n
+    \end{aligned}
+  }
+  \infer1[\ExpandM]{
+    \Image{}_{\Tuple{A', B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})
+    \land l_A = l_B % Implied constraint
+    \land l_A - i - n > 0
+    \land 0 < i < n
+    }
 \end{prooftree}
-\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
+\caption{A derivation for \Calculus{} on the Parikh image strings for the
+constraints of \cref{ex:string-constraints}. Note the constant propagation
+for~$n$!}\label{fig:derivation:multi}
 \end{figure}

### 2023-06-27 multiple.tex
Multi-ex

@@
 
-We return again to our example in \cref{sec:intuition} and start by performing
-the steps of \cref{sec:single:example} shown in \cref{fig:derivation:single},
-since the steps to propagate \cref{ex:string-constraints} across the automaton
-$A'$ still applies for the product. We also reuse the definition of $\Map$ as
-the simple projection to extract the counter increments, so in this case,
-$\Map(\Transition) = \left[l_{B+}\right]$ for every transition~$\Transition$
-since we are only counting the length.
+We return again to our example in \cref{sec:intuition}.We also reuse the
+definition of $\Map$ as the simple projection to extract the counter increments,
+so in this case, $\Map(\Transition) = \left[l_{B+}\right]$ for every
+transition~$\Transition$ since we are only counting the length. The reader is
+strongly advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
+\cref{sec:intuition} while going through this example! In particular, note that
+the transition variables of the figures correspond to the ones used here.
 
-We reuse the transition variable associations of our previous
-\cref{sec:single:example}, from \cref{eq:single-filter}, but add the definitions
-for $B$ shown in \cref{eq:multi-filter}:
+The transition variables of \cref{sec:single:example} recur
+\cref{eq:single-filter}, but we add the definitions for $B$ shown in
+\cref{eq:multi-filter}:
 \begin{equation}\label{eq:multi-filter}
@@
         \FlowEq(B, \Filter) =  \begin{aligned}
-            1 + \TransitionVar_{12}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
+            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
             \land \TransitionVar_{11} &= \TransitionVar_6\\
@@
 
-In the derivation tree of \cref{fig:derivation:multi} we start as before, but
-with the product version of the $\Image$ predicate. The only possible rule here
-is $\ExpandM$, which we use to add the corresponding constraints on each
-automata of the product as we would have had in the single-automaton version.
+The only possible rule at the start is $\ExpandM$, which we use to add the
+corresponding constraints on each automata of the product as we would have had
+in the single-automaton version.
 
-We perform algebraic reasoning on those equations, along with the constraint on
-$\MonoidElement$ to determine bounds on transition variables that will enable us
-to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
-finished doing so, we use \Materialise{} to compute the product of the now
-filtered automata. Further algebraic reasoning allows us to remove the final
-instance of $\Connected$. We then perform \Expand{} to get rid of the final
-$\Image$ instance, and are immediately able to $\Subsume$ the resulting
-$\Connected{}$ predicate. This, again, leaves us with a set of linear
-inequalities at the leaf that we can use to obtain a model of our final
-$\MonoidElement$ values (now a vector). This time we can read out aca, or
-$\VectorLiteral{2, 0, 1}$.
+We then continue by repeating the steps of the steps of
+\cref{sec:single:example} shown in \cref{fig:derivation:single}, since the steps
+to propagate \cref{ex:string-constraints} across the automaton $A'$ still apply
+for the product. Note that this fixes the values of all transitions but one as
+constants.
+
+\Fudge{description here!}
 
@@
 \begin{prooftree}
-  % LEFT TREE  ->
-  \hypo{l}% here x_12 is zero
-  % RIGHT TREE ->
-  \hypo{r} % here x_12 is one
-  \infer2[\Split{} $\TransitionVar_{12} = 0$]{\text{FINAL STEP OF STUMP}}
-  \infer1[\EquationReasoning: probably something good]{
+  {% % LEFT TREE: here x_12 is zero
+  \hypo{PROOF CLOSED}
+  \infer1[\Propagate]{WHAT HAPPENS WHEN X_12 IS ZERO}
+  }% END LEFT TREE
+  {% RIGHT TREE % here x_12 is one
+  \hypo{r} 
+  }% END RIGHT TREE
+  \infer2[\Split{} $\TransitionVar_{12} = 0$]{
+  \begin{aligned}
+    \TransitionVar_7 = l_A - \TransitionVar_9 - 2\TransitionVar_{11}-2\TransitionVar_{12}\\
+    \land \TransitionVar_2 = l_A - 3\\
+    \land \TransitionVar_{8} = \TransitionVar_{11} + \TransitionVar_{12} - 1\\
+    \land \TransitionVar_6 = \TransitionVar_{11}\\
+    \land \TransitionVar_{10} = 1 - \TransitionVar_{11}\\
+    \land \Connected(B, \Filter) \\
+    \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
+  \end{aligned}
+  }
+  \infer1[Expand $\sum$, \EquationReasoning] {
     \begin{aligned}
@@
       \land \TransitionVar_2 = l_A - 3\\
-      \land 1 + \TransitionVar_{12}   = \TransitionVar_{11} + \TransitionVar_{12}\\
+      \land 1 + \TransitionVar_{8}   = \TransitionVar_{11} + \TransitionVar_{12}\\
       \land \TransitionVar_{11} = \TransitionVar_6\\

### 2023-06-27 multiple.tex
WOW IT FITS

@@
   {% % LEFT TREE: here x_12 is zero
-  \hypo{PROOF CLOSED}
-  \infer1[\Propagate]{WHAT HAPPENS WHEN X_12 IS ZERO}
+  \hypo{
+      \ldots \land 1 = 0 
+      \land \Image{}_{\emptyset, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
+  }
+  \infer1[\Materialise: $\emptyset$]{
+    \begin{array}{llll}
+      \TransitionVar_9 = 0 
+      \land \TransitionVar_7 = l_A - 2 \\
+      \land \TransitionVar_2 = l_A - 3
+      \land \TransitionVar_{8} = 0 \\
+      \land \TransitionVar_6 = \TransitionVar_{11} = 1
+      \land \TransitionVar_{10} = 0 \\
+      \land \Connected(B, \Filter)\\
+      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
+    \end{array}
+  }
+  \infer1[\Propagate{}]{
+    \begin{array}{llll}
+      \TransitionVar_{12} = 0 
+      \land \TransitionVar_7 = l_A - \TransitionVar_9 - 2 \\
+      \land \TransitionVar_2 = l_A - 3 
+      \land \TransitionVar_{8} = 0 \\
+      \land \TransitionVar_6 = \TransitionVar_{11} = 1 \\
+      \land \TransitionVar_{10} = 0 
+      \land \Connected(B, \Filter) \\
+      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
+    \end{array}
+    }
   }% END LEFT TREE
-  {% RIGHT TREE % here x_12 is one
-  \hypo{r} 
+  {% RIGHT TREE % here x_12 is > 0
+  \hypo{
+      \ldots \land 1 = 0 
+      \land \Image{}_{\emptyset, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
+  }
+  \infer1[\Materialise: $\emptyset$]{
+    \begin{array}{rrrr}
+      \TransitionVar_{12} > 0 \land \\
+      \TransitionVar_7 = l_A - \TransitionVar_9 - 2\TransitionVar_{11}-2\TransitionVar_{12} \land \\
+      \TransitionVar_2 = l_A - 3 \land \\
+      \TransitionVar_{8} = \TransitionVar_{11} + \TransitionVar_{12} - 1 \land \\
+      \TransitionVar_6 = \TransitionVar_{11} \land\\
+      \TransitionVar_{10} = 1 - \TransitionVar_{11} \land\\
+      \Connected(B, \Filter) \land\\
+      \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
+    \end{array}
+  }
   }% END RIGHT TREE
   \infer2[\Split{} $\TransitionVar_{12} = 0$]{
-  \begin{aligned}
-    \TransitionVar_7 = l_A - \TransitionVar_9 - 2\TransitionVar_{11}-2\TransitionVar_{12}\\
-    \land \TransitionVar_2 = l_A - 3\\
-    \land \TransitionVar_{8} = \TransitionVar_{11} + \TransitionVar_{12} - 1\\
-    \land \TransitionVar_6 = \TransitionVar_{11}\\
-    \land \TransitionVar_{10} = 1 - \TransitionVar_{11}\\
-    \land \Connected(B, \Filter) \\
+  \begin{array}{llll}
+    \TransitionVar_7 = l_A - \TransitionVar_9 - 2\TransitionVar_{11}-2\TransitionVar_{12} \\
+    \land \TransitionVar_2 = l_A - 3 \\
+    \land \TransitionVar_{8} = \TransitionVar_{11} + \TransitionVar_{12} - 1 \\
+    \land \TransitionVar_6 = \TransitionVar_{11} \\
+    \land \TransitionVar_{10} = 1 - \TransitionVar_{11} \\
+    \land \Connected(B, \Filter)\\
     \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
-  \end{aligned}
+  \end{array}
   }
   \infer1[Expand $\sum$, \EquationReasoning] {
-    \begin{aligned}
-      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
-      \land \TransitionVar_2 = l_A - 3\\
-      \land 1 + \TransitionVar_{8}   = \TransitionVar_{11} + \TransitionVar_{12}\\
-      \land \TransitionVar_{11} = \TransitionVar_6\\
-      \land \TransitionVar_{12} = \TransitionVar_8 + \TransitionVar_{10}\\
-      \land \Connected(B, \Filter) \\
+    \begin{array}{llll}
+      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k
+      \land \TransitionVar_2 = l_A - 3
+      \land 1 + \TransitionVar_{8}   = \TransitionVar_{11} + \TransitionVar_{12}
+      \land \TransitionVar_{11} = \TransitionVar_6 \\
+      \land \TransitionVar_{12} = \TransitionVar_8 + \TransitionVar_{10}
+      \land \Connected(B, \Filter) 
       \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
-    \end{aligned}
+    \end{array}
   }
   \infer1[\EquationReasoning: expand $\FlowEq$]{
-    \begin{aligned}
-      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
-      \land \TransitionVar_2 = l_A - 3\\
+    \begin{array}{llll}
+      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k
+      \land \TransitionVar_2 = l_A - 3
       \land \FlowEq(B, \Filter) \\
-      \land \Connected(B, \Filter) \\
+      \land \Connected(B, \Filter) 
       \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
-    \end{aligned}
+    \end{array}
   }
   \infer1[Repeat \cref{fig:derivation:single}, $l_A = l_B$.]{
-    \begin{aligned}
-      l_B = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
-      \land i = \TransitionVar_5\\
-      \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k\\
-      \land n = \TransitionVar_4 + \TransitionVar_1\\  
-      \land \FlowEq(B, \Filter) \\
-      \land \Connected(A', \Filter) \\
-      \land \Connected(B, \Filter) \\
-      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) 
-      \land l_A = l_B\\
-      \land l_A - i - n > 0\\
+    \begin{array}{llll}
+      l_B = \sum_{k = 6}^{k = 12} \TransitionVar_k
+      \land i = \TransitionVar_5
+      \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k \\
+      \land n = \TransitionVar_4 + \TransitionVar_1
+      \land \FlowEq(B, \Filter)
+      \land \Connected(A', \Filter)\\
+      \land \Connected(B, \Filter)
+      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) \\
+      \land l_A = l_B
+      \land l_A - i - n > 0
       \land 0 < i < n
-    \end{aligned}
+    \end{array}
   }

### 2023-06-28 calculus.tex
Evaluation
Progress on example
Clean up the equations a bit

@@
 shorthand of listing the matched inequalities as antecedents e.g., in
-\Propagate{}).
+\Propagate{}.
 
@@
 
-\subsection{An Example}
+\subsection{An Example}\label{sec:single:example}
 Here we will return to $A'$ from \cref{sec:intuition} and performs the steps of
@@
 0\right]$. Initially, we let $\Filter$ map to fresh variables to obtain:
-\begin{equation}
+\begin{equation}\label{eq:single-filter}
     \begin{aligned}
-      \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A}) & = \TransitionVar_1    & 
-                \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) & = \TransitionVar_2 &       \FlowEq(A', \Filter) & = \begin{aligned}
+      & \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A}) = \TransitionVar_1
+        & \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) = \TransitionVar_2   
+        & \FlowEq(A', \Filter)  = \begin{aligned}
                        1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_3 + \TransitionVar_1\\
@@
                     \end{aligned} \\
-      \Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) & =  \TransitionVar_3 & 
-                \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) & = \TransitionVar_4 &\\
-      \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) & = \TransitionVar_5 & 
+      &\Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) =  \TransitionVar_3 
+          & \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) = \TransitionVar_4
+          & \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) = \TransitionVar_5  
     \end{aligned}
-\end{equation}\label{eq:single-filter}
+\end{equation}
 
@@
   \hypo[]{{\text{Regular solving finishes the proof}}}
+  \infer1[\EquationReasoning{}: $i = 1$]{
+  i = 1
+  \land n = 2
+  \land \TransitionVar_2 = l_A - 3
+  \land \TransitionVar_3 = 0
+  \land \TransitionVar_5 = 1
+  \land \TransitionVar_1 = \TransitionVar_4 = 1
+  }
   \infer1[]{
@@
     n = 2 \land \\
-    2 \geq i \land l_A - i - 2 > 0 
+    0 < i < 2
+    \land l_A - i - 2 > 0 
     \land \TransitionVar_2 = l_A - i - 2 
@@
     i = \TransitionVar_5 \land 
-    n = 2 \land 
-    2 \geq i \land l_A - i - 2 > 0  \\
+    n = 2 \land \\
+    0 < i < 2 
+    \land l_A - i - 2 > 0
     \land \TransitionVar_2 = l_A - i - 2 
-    \land \TransitionVar_3 = 0 \\
+    \land \TransitionVar_3 = 0
     \end{aligned}
@@
     \begin{aligned}
-    \Connected(A', \Filter) \land \\
+    \Connected(A', \Filter) \land
     i = \TransitionVar_5 \land 
-    n = 2\TransitionVar_1 \land n \geq i \land l_A - i - n > 0 
+    n = 2\TransitionVar_1 
+    \land n > i > 0 \\
+    \land l_A - i - n > 0 
     \land \TransitionVar_2 = l_A - i - n 
@@
     \land n = \TransitionVar_4 + \TransitionVar_1 
-    \land n \geq i 
-    \land l_A - i - n > 0 \\
-    \land \TransitionVar_4 = \TransitionVar_1 
+    \land n > i > 0
+    \land l_A - i - n > 0
+    \land \TransitionVar_4 = \TransitionVar_1 \\
     \land \TransitionVar_2 = l_A - i - n 
@@
     \land 1 = \TransitionVar_3 + \TransitionVar_1 
+    \land i > 0
     \land \TransitionVar_4 = \TransitionVar_1 
@@
     \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k 
-    \land n \geq i \land l_A - i - n > 0
+    \land n > i > 0
+    \land l_A - i - n > 0
     \land n = \TransitionVar_4 + \TransitionVar_1\\
@@
   }
-  \infer1[\Expand{}]{\Image{}_{A', \Map}(\Filter, \left[l_A, i, n\right]) \land n \geq i \land l_A - i - n > 0}
+  \infer1[\Expand{}]{
+    \Image{}_{A', \Map}(\Filter, \left[l_A, i, n\right]) 
+    \land n \geq i 
+    \land l_A - i - n > 0
+    \land i > 0
+    }
 \end{prooftree}

### 2023-06-28 evaluation.tex
Clean up the equations a bit

@@
 a corum ground truth as they are solved by only one backend. We see that
-\Calculus{} generally outperforms \Nuxmv{} on determining unsatisfiability,
-while being similar at satisfiable instances. Both \Nuxmv{} and \Calculus{}
-outperform baseline on satisfiable instances, and \Calculus{} outperforms
-\Nuxmv{}.
+\Calculus{} generally outperforms \Nuxmv{} on determining unsatisfiability, as
+does baseline while being similar at satisfiable instances. Both \Calculus{} and
+baseline outperforms baseline on satisfiable instances. On satisfiable
+instances, \Nuxmv and \Calculus{} has similar performance.
 
@@
 instances roughly similar to \cite{approximate-parikh}. The heuristic is enabled
-since we suspect that unsatisifiable instances are more common; at least we have
-seen more solved unsatisfiable instances.
+since the improvement is significant compared to the extra cost and since we
+have observed more unsatisfiable instances than satisfiable ones.
 
@@
 
-A cactus plot showing the number of instances solved within a given timeout for
-each backend configuration can be seen in \cref{fig:cactus}. Here, we see
-clearly that \Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{}
-might have better long-term scaling properties. Notably, for machines with
-better single-core performance \Nuxmv{}'s scaling benefits kicks in faster,
-though the graph looks similar in shape.
-
-\iffalse
-\subsection{Finding a Presburger Formula}\label{sec:evaluation:finding-image}
-
-For baseline and \Calculus{}, \Catra{} offers the ability to find the equivalent Presburger formula representing a given instance.
-For baseline, we use the built-in quantifier elimination facilities of the underlying \Princess{} theorem prover, while for \Catra{} we use the specially tailored approach described in \cref{sec:finding-the-image}.
-For this experiment, we use only the~\NrKnownSat{} instances known to be satisfiable from the previous experiment detailed in \cref{sec:scaling,sec:runtime}.
-
-To make sure baseline puts up as much competition as possible, we disable
-checking intermittent satisfiability and configure \Catra{} to run in the
-maximally eager mode where the product is first computed before any
-satisifiability check is performed. We run the experiments with a timeout
-of~\ImageTimeout{}. The results of the experiment is summarised in
-\cref{fig:cactus:image} and \cref{tab:image-results}. \Fudge{We see here that
-something happens}.
-
-\begin{figure}[ht]
-  \caption{The number of instances \Catra{} was able to find the Presburger form of the image for within a given number of seconds per backend.}
-  \label{fig:cactus:image}
-\end{figure}
-\fi
+A cactus plot showing the number of instances solved within a given timeout can
+be seen in \cref{fig:cactus}. We see that \Calculus{} outperforms \Nuxmv{} in
+general, but that \Nuxmv{} might have better scaling properties on longer runs.
+For machines with better single-core performance than our experiment machine
+\Nuxmv{}'s scaling benefits kicks in faster, so it is likely that the amount of
+CPU cycles spent on the problem is the underlying factor. This is likely due to
+two factors. First, \Nuxmv{} is more mature than \Calculus{}, and it makes sense
+that its more general model checking methods would pay off on more difficult
+instances where more problem-specific methods offer less payoff. Second, for
+longer running calculations clause learning as described in
+\cref{sec:clause-learning} can be assumed to matter more. However, the clauses
+we learn are still relatively unpolished, and in particular generalising
+insights across product computations is nontrivial and therefore partially left
+as future work.
 

### 2023-06-28 multiple.tex
Clean up the equations a bit

@@
 
-We return again to our example in \cref{sec:intuition}.We also reuse the
+We return again to our example in \cref{sec:intuition}. We also reuse the
 definition of $\Map$ as the simple projection to extract the counter increments,
@@
     \begin{aligned}
-        \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6\\
-        \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7\\
-        \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B})= \TransitionVar_8 \\
-        \Filter(\FromLabelTo{B}{c / l_{B+}}{B})= \TransitionVar_9\\
-        \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}\\
-        \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}\\
-        \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}\\
-        \FlowEq(B, \Filter) =  \begin{aligned}
+        & \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6 
+        & \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7 
+        & \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) = \TransitionVar_8 
+        &\Filter(\FromLabelTo{B}{c / l_{B+}}{B}) = \TransitionVar_9  \\
+        & \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}
+        & \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}
+        & \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}
+        \FlowEq(B, \Filter) &=  \begin{aligned}
             1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
@@
 
-%\subsubsection{\Calculus{} for products of automata terminates}
+\subsubsection{\Calculus{} for products of automata terminates}
 \begin{lemma}\label{lma:multi-terminates}
@@
 
-%\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
+\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
 

### 2023-06-28 conclusion.tex
Conclusions and multi light touchup

--- 2023-06-27 main.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2023}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\usepackage{amsmath,empheq,fancybox}
-\usepackage{paralist}
-\usepackage{url}
-\usepackage{color}
-\usepackage{textcomp,listings}
-\usepackage{array}
-\usepackage{mymacros}
-\usepackage{microtype}
-\usepackage{listings}
-\usepackage{csquotes}
-\usepackage{proof}
-\usepackage[capitalise]{cleveref}
-\usepackage{algorithm2e}      
-\usepackage{multirow}
-\usepackage{mathpartir}
-\usepackage{amsthm}
-\usepackage{numprint}
-\usepackage{ebproof}
-\usepackage[makeroom]{cancel}
-
-\usepackage{mathtools} % Bonus
-
-\usepackage{enumitem}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{example}{Example}[section]
-
-\newlist{constraints}{enumerate}{3}
-\setlist[constraints,1]{%
-    label=\sffamily{(\roman*):},
-    ref=\normalfont{Constraint (\roman*)},
-    wide,itemsep=0pt,topsep=0pt
-}    
-\crefname{constraintsi}{}{}
-\Crefname{constraintsi}{}{}
-
-\lstset{
-    columns=fullflexible,
-    showspaces=false,
-    showtabs=false,
-    breaklines=true,
-    showstringspaces=false,
-    breakatwhitespace=true,
-    escapeinside={(*@}{@*)},
-    commentstyle=\color{greencomments},
-    keywordstyle=\color{bluekeywords},
-    stringstyle=\color{redstrings},
-    numberstyle=\color{graynumbers},
-    basicstyle=\ttfamily\small,
-    framesep=12pt,
-    xleftmargin=12pt,
-    tabsize=4,
-    captionpos=b
-}
-
-\newif\ifcomments%
-\commentstrue%
-\newif\ifoutline%
-\outlinetrue{}
-
-\newcommand{\contents}[1]{\ifoutline{\color{blue}
-    \begin{itemize}
-    #1
-    \end{itemize}
-  }\fi}
-
-\allowdisplaybreaks[1]
-
-\begin{document}
-
-%% Title information
-\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
-   counts in strings of a given language. Automata-based string solvers
-   frequently need to compute the Parikh image of large products (intersections)
-   of nondeterministic automata, which in many operations is both prohibitively
-   slow and memory-intensive. We contribute a novel understanding of how Parikh
-   maps can be tackled as a constraint solving problem to solve real-world
-   constraints stemming from functions on regular languages, most notably the
-   length constraint. Furthermore, we show how this formulation can be
-   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
-   prover supporting Presburger logic.  We implement \Calculus{} in a tool
-   called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving Parikh automata
-   intersection problems produced when solving standard string constraint
-   benchmarks involving string length constraints. We show that our solution
-   strictly outperforms the standard approach described
-   in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wide
-   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
-   model checker.
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-  <ccs2012>
-  <concept>
-  <concept_id>10003752.10003766.10003773.10003775</concept_id>
-  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003766.10003776</concept_id>
-  <concept_desc>Theory of computation~Regular languages</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10003794</concept_id>
-  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10010124.10010138.10010142</concept_id>
-  <concept_desc>Theory of computation~Program verification</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  <concept>
-  <concept_id>10003752.10003790.10011192</concept_id>
-  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
-  <concept_significance>300</concept_significance>
-  </concept>
-  </ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Quantitative automata}
-\ccsdesc[300]{Theory of computation~Regular languages}
-\ccsdesc[500]{Theory of computation~Automated reasoning}
-\ccsdesc[300]{Theory of computation~Program verification}
-\ccsdesc[300]{Theory of computation~Verification by model checking}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}\label{sec:introduction}
-\input{introduction.tex}
-
-\section{An Intuition for Our Approach}\label{sec:intuition}
-\input{example.tex}
-
-\section{Preliminaries}\label{sec:preliminaries}
-\input{preliminaries.tex}
-
-\section{Projections on Parikh Images}\label{sec:generalised}
-\input{generalised.tex}
-
-\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
-\input{calculus.tex}
-
-\section{Parikh Images from Products of Automata}\label{sec:multiple}
-\input{multiple.tex}
-
-\section{Extensions}\label{sec:extensions}
-\input{extensions.tex}
-
-\section{Implementation}\label{sec:implementation}
-\input{implementation.tex}
-
-\section{Evaluation}\label{sec:experiments}
-\input{evaluation.tex}
-
-\section{Conclusion}
-
 In this paper we have introduced a calculus to compute commutating operations on
@@
 
-Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
-usage and solve-time compared to the baseline approach laid out in
+Within \Catra{}, \Calculus{} shows astonishing performance in terms of
+solve-time compared to the baseline approach laid out in
 \cite{generate-parikh-image} when implemented on the same underlying automated
@@
 to the expected Presburger formulae.
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-%% Bibliography
-\bibliography{bibliography}
-
-%% Appendix
-\appendix
-\section{Appendix}
-\input{appendix.tex}
-\end{document}

### 2023-06-28 main.tex
Conclusions and multi light touchup

@@
 \section{Conclusion}
-
-In this paper we have introduced a calculus to compute commutating operations on
-intersections of regular languages that we call \Calculus{}. We have evaluated
-it on \NrBenchmarks{} Parikh automata intersection problems generated by the
-\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
-suite \cite{pyex} using our Parikh automata solver \Catra{}.
-
-Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
-usage and solve-time compared to the baseline approach laid out in
-\cite{generate-parikh-image} when implemented on the same underlying automated
-theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
-\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
-and generally outperforming it for timeouts under 30 seconds with its advantage
-increasing drastically for even shorter timeouts. 30 seconds would generally be
-considered a long timeout for our intended use as supporting infrastructure to a
-string constraint solver.
-
-Future investigations involve two tracks. The first one is integration into
-existing string solvers (wich \Ostrich{} being a particularly promising
-candidate due to its shared use of \Princess{}), and further adaptation to that
-use case. Closer inspection of the instances where we currently time out should
-be useful to further improve our heuristics.
-
-The second track for future improvements is the extension into other problem
-domains, including other logics, model checking problems, as well as to more
-powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata, due to our use of a full
-automated theorem prover which allows adding arbitrary constraints in addition
-to the expected Presburger formulae.
+\input{conclusion.tex}
 

### 2023-06-28 multiple.tex
Conclusions and multi light touchup
Description of what we do in 6.1

@@
 
-\iffalse
-Note that this definition implies that $\Filter(\Transition) =
-\Filter(\Transition')$ whenever two transitions $\Transition \in
-\Transitions_{\Automaton_1}, \Transition' \in \Transitions_{\Automaton_2}$
-produces a product transition $\Transition'' \in \Transitions_{\Automaton_1
-\times \Automaton_2}$. This corresponds to our intuition that the terms of the
-product must agree on the value they accept. As before, we implicitly map
-$\Filter$ to fresh terms for each transition in the product.
-\fi
-
-For instances of precisely one automaton, neither rule applies and we
-perform the calculus as before.
+For instances of precisely one automaton, neither rule applies, and we perform
+the calculus as before.
 
@@
 
-We return again to our example in \cref{sec:intuition}. We also reuse the
+We return to our example in \cref{sec:intuition}. We also reuse the
 definition of $\Map$ as the simple projection to extract the counter increments,
 so in this case, $\Map(\Transition) = \left[l_{B+}\right]$ for every
-transition~$\Transition$ since we are only counting the length. The reader is
+transition~$\Transition$ since $B$ only counts length. The reader is
 strongly advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
 \cref{sec:intuition} while going through this example! In particular, note that
-the transition variables of the figures correspond to the ones used here.
+the transition variables of the figures match the ones used here.
 
-The transition variables of \cref{sec:single:example} recur
-\cref{eq:single-filter}, but we add the definitions for $B$ shown in
-\cref{eq:multi-filter}:
+We extend \cref{eq:single-filter} from \cref{sec:single:example} with
+definitions for $B$ shown in \cref{eq:multi-filter}:
 \begin{equation}\label{eq:multi-filter}
@@
 The only possible rule at the start is $\ExpandM$, which we use to add the
-corresponding constraints on each automata of the product as we would have had
-in the single-automaton version.
+corresponding constraints on each automaton of the product as we would have had
+in the single-automaton version. After that, we expand the various sums and apply
+light equality propagation.
 
-We then continue by repeating the steps of the steps of
-\cref{sec:single:example} shown in \cref{fig:derivation:single}, since the steps
-to propagate \cref{ex:string-constraints} across the automaton $A'$ still apply
-for the product. Note that this fixes the values of all transitions but one as
-constants.
+We then continue by repeating the steps of \cref{sec:single:example} shown in
+\cref{fig:derivation:single}, since the steps to propagate
+\cref{ex:string-constraints} across the automaton $A'$ still apply for the
+product. Note that this fixes the values of all transitions but one of $A'$ as
+constants, including the free counter variables $i, n$, whose new constant
+values we propagate everywhere.
 
-\Fudge{description here!}
+Once we have performed reasoning by equality to eliminate transition variables,
+we obtain the version of $B$ shown in \cref{fig:aut_b_annotated} and represented
+by the clauses in the final node in the tree. This corresponds to the end of
+\cref{sec:intuition:algebra}. Our only options now are to either split on which
+path we take from $I_B$ by case-splitting on~$\TransitionVar_{12}$ or to
+directly invoke \Materialise{}. The latter would produce a shorter tree in this
+instance but might lead to a larger product being computed, so we (and indeed
+\Catra{}) pick \Split{}. This leads to an opportunity for propagation since
+$\TransitionVar_{12} = 0$ cuts off $\TransitionVar_{9}$ from $I_B$ on the left
+branch. Note that the same opportunity does not present itself on the right
+branch, though; the use of the upper path to state~$A$ does not preclude visiting
+state~$B$.
+
+In both cases, we have comparatively simple automata left, so we apply
+\Materialise{} to obtain their product, which for both branches is empty. An
+empty product will have no outgoing transitions from its initial state, and so
+will produce a flow equation $1 = 0$ since the right-hand balancing side will
+always be the empty sum. We can then close the proof and know that the problem
+is unsatisfiable.
 
@@
   \hypo{
-      \ldots \land 1 = 0 
+    \begin{array}{l}
+      \ldots \boxed{\land 1 = 0} \\
       \land \Image{}_{\emptyset, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
+    \end{array}
   }
-  \infer1[\Materialise: $\emptyset$]{
+  \infer1[\Materialise]{
     \begin{array}{llll}
@@
   \hypo{
-      \ldots \land 1 = 0 
+    \begin{array}{l}
+      \ldots \boxed{\land 1 = 0} \\
       \land \Image{}_{\emptyset, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
+    \end{array}
   }
-  \infer1[\Materialise: $\emptyset$]{
+  \infer1[\Materialise]{
     \begin{array}{rrrr}
@@
 for~$n$!}\label{fig:derivation:multi}
+\Description[]{}% This isn't a problem; it's all text
 \end{figure}
@@
   one automaton, until only one remains and \cref{lma:single-terminates} for
-  single-automaton instances applies.
+  single-automaton instances apply.
 

### 2023-07-04 conclusion.tex
Revise experiments, grammar check

@@
-In this paper we have introduced a calculus to compute commutating operations on
+In this paper, we have introduced a calculus to compute commutating operations on
 intersections of regular languages that we call \Calculus{}. We have evaluated
@@
 powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata, due to our use of a full
-automated theorem prover which allows adding arbitrary constraints in addition
+express stronger constraints than Parikh automata due to our use of a full
+automated theorem prover, which allows adding arbitrary constraints in addition
 to the expected Presburger formulae.

### 2023-07-04 implementation.tex
Clean up the equations a bit
Revise experiments, grammar check

@@
 \Catra.
-% \footnote{If you really must read it as an acronym, please read it as
-%CAtegory Theory on Register Automata, or if you object to the somewhat
-%nonstandard use of register automata and category theory, as Check Assignments
-%of The Registers Afterwards. Or alternatively if you find it all to be too much
-%of a theoretical exercise, as Can Anyone Think of A Real Application.}
 
@@
 baseline back-end. An example of an input file corresponding to our running
-example introduced in \cref{sec:motivation} can be seen in
+example introduced in \cref{sec:intuition} can be seen in
 \cref{lst:input-example}.
@@
 finite range of Unicode code points. This allows succinct representation of many
-regular expression patterns such as \lstinline{(a-z).*} which would have
+regular expression patterns such as \lstinline{(a-z).*}, which would have
 otherwise required $27$~transitions. The transition for the range would be
@@
 assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
-Additionally, baseline and \Calculus{} also support generation of the Presburger
+Additionally, baseline and \Calculus{} also support generating the Presburger
 formula describing the constraints of the input file. Baseline uses standard
@@
 
-Since \Princess{} does not support multiple-arity predicates like the ones we
-use in \Calculus{}, we have implemented variable-length arguments using
-additional helper predicates. These are $\Unused{}(\Automaton)$, which marks an
-automaton as unused in any product, and $\TransitionMask{}(\Automaton,
-\Transition, \Filter(\Transition))$ which associates a transition $\Transition$
-and automaton $\Automaton$ with its corresponding transition variable.
-Additionally, we associate each of our predicates with an instance variable in
-order to differentiate instances of the predicates.
+Since \Princess{} does not support the multiple-arity predicates used in
+\Calculus{}, we have implemented variable-length arguments using additional
+helper predicates. These are $\Unused{}(\Automaton)$, which marks an automaton
+as unused in any product, and $\TransitionMask{}(\Automaton, \Transition,
+\Filter(\Transition))$ which associates a transition $\Transition$ and automaton
+$\Automaton$ with its corresponding transition variable. Additionally, we
+associate each predicate with an instance variable to tie helper predicates
+to the $\Image$ predicate they originated from.
 
@@
 We baseline using the same Presburger solver (\Princess{}), input file parser,
-and automaton implementation as \Catra. We do this in order to better analyse
-the impact of the calculus rules themselves. Using the formula of
+and automaton implementation as \Catra. We do this to better analyse the impact
+of the calculus rules themselves. Using the formula of
 \eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
@@
 term by term, checking satisfiability at each step. We use a priority queue to
-select automata for each step, and order it by the number of transitions as a
-heuristic for the size of the automata. We use this heuristic to avoid computing
-large (and therefore slow) products until we have to, banking our hopes on
-computing an empty intermittent product early.
+select automata for each step and order them by their number of transitions. We
+use this heuristic to put off computing large (and therefore slow) products until
+we have to, hoping to find an empty intermittent product.
 
@@
 
-As an optimisation, our automata (including intermittent products) are created
-forward- and backward- reachable-minimal. Any automaton we produce only contains
+As an optimisation, our automata (including intermittent products) have dead
+states eliminated during construction. Any automaton we produce contains only
 states that are both reachable from the initial state and has a path to an
@@
 minimisation on automata with counters is non-trivial, and minimising symbolic
-automata risks exponential blowup \cite{minimising-symbolic}.
+automata risks exponential blowup~\cite{minimising-symbolic}.
 
@@
 
-There are a number of choices left unspecified in \Calculus{} as described in
-\cref{sec:calculus,sec:multiple}. For example, the order of materialisation of
-intermediate products and the order of splitting. In this section we describe
-additional implementation details and techniques used to enhance \Catra.
+\Calculus{} as described in \cref{sec:calculus,sec:multiple} leaves some choices
+unspecified, including the priority of rules and the order of their arguments. In
+this section, we address these and describe additional implementation details
+and techniques used to enhance \Catra{}.
 
@@
 
-We order our rule applications so that we first propagate connectedness if
+We order our rule applications as follows: first, propagate connectedness if
 possible, then perform materialisation if tractable as defined below, then
-finally resort to splitting if we must.
+finally resort to splitting as a last resort.
 
@@
 randomly selected transitions, we prefer splitting to sever a strongly connected
-from the initial state. We randomly select an automaton where we can compute a
-cut between an SCC and the initial state, that is where the SCC does not contain
-the initial state and where the sum of the transition variables of the
-transitions in the cut is not known to be positive. If there are multiple such
-strongly connected components we choose one randomly. We then proceed to split
-on the sum of the transition variables of the cut as if it were a regular
-transition, e.g. its sum being zero or nonzero. In this way we drive \Calculus{}
+component from the initial state. We randomly select an automaton where we can
+compute a cut between an SCC and the initial state, that is, where the SCC does
+not contain the initial state and where the sum of the transition variables of
+the transitions in the cut is not known to be positive. If there are multiple
+such strongly connected components, we choose one randomly. We then proceed to
+split on the sum of the transition variables of the cut as if it were a regular
+transition, e.g., its sum being zero or nonzero. In this way, we drive \Calculus{}
 towards applying \Propagate{}.
 
-The implementation of the connectedness constraint is opportunistic and straightforward.
-We compute a set of dead states by performing forward and backwards reachability computations on an automaton, where we disregard any transition whose associated variable is known to be zero.
-After that we add clauses ensuring any transition variable associated with a transition starting in a dead state is zero.
+The implementation of the connectedness constraint is opportunistic and
+straightforward. We compute a set of dead states by performing forward and
+backward reachability computations on an automaton, where we disregard any
+transition whose associated variable is known to be zero. After that we add
+clauses ensuring any transition variable associated with a transition starting
+in a dead state is zero.
 
 Product materialisation is the final piece of the puzzle. In the current
-implementation we put off computing intermediate products until at least all but
-two transition variables of one of the automata is known to be either present or
-absent. The number was chosen experimentally, and we observe a consistent trend
-towards lower numbers being better. The other automaton for the product is
+implementation we put off computing intermediate products until six transition
+variables of one of the automata is known to be either present or absent. The
+number was chosen experimentally. The other automaton for the product is
 selected randomly.

### 2023-07-04 experiments/experiments.py

@@ -2,6 +2,8 @@
 
 from collections import namedtuple
 from enum import Enum, auto
+from pandas import DataFrame
+import pandas
 import re
 
 class Status(Enum):
@@ -48,3 +50,32 @@ def parse_log(log):
             if line.startswith("RESULT"):
                 for r in parse_line(line[len("RESULT "):], configurations):
                     yield r
+                    
+def log2df(log):
+    runtimes = DataFrame(parse_log(log))
+    runtimes = runtimes[~runtimes.index.duplicated()]
+    runtimes['backend'] = runtimes['backend'].astype("category")
+    runtimes['status'] = runtimes['status'].astype("category")
+    runtimes['status'] = runtimes['status'].cat.set_categories(
+        [Status.SAT, Status.UNSAT, Status.TIMEOUT, Status.MEMORY_OUT, Status.ERROR], 
+        ordered=True)
+    return runtimes
+
+
+def solved_after(seconds, known_solved):
+    df = known_solved[known_solved['runtime'] < seconds]\
+        .groupby('backend')\
+        .size()\
+        .reset_index()
+    df['after'] = seconds
+    df = df.rename(columns = {0: 'nr solved'})
+    return df
+
+
+def prepare_cactus_plot(log, timeout_s, step_size):
+    runtimes = log2df(log)
+    known_solved = runtimes[runtimes['status'] < Status.TIMEOUT]
+    return (runtimes.instance.unique().size, 
+            pandas.concat([solved_after(i * step_size, known_solved) for i in range(int(timeout_s / step_size) + 1)], axis=0))
+
+    

### 2023-07-10 experiments/experiments.py
Ok really for real new figures

@@ -72,7 +72,6 @@ def solved_after(seconds, known_solved):
     df = df.rename(columns = {0: 'nr solved'})
     return df
 
-
 def prepare_cactus_plot(log, timeout_s, step_size):
     runtimes = log2df(log)
     known_solved = runtimes[runtimes['status'] < Status.TIMEOUT]

### 2023-07-10 generalised.tex
comments, fixes

@@
 \cite{ostrich-plus} will construct a pre-image of $s_1$ under the substring
-operation, and check whether this pre-image is consistent with the constraint~$x
-\in \Language_1$, corresponding to the full product seen in
+operation, and check whether this pre-image is consistent with the constraint~$s_2 \in \Language(B)$, corresponding to the full product seen in
 \cref{ex:string-constraints}. Because the substring operation depends on the
@@
 % please verify I'm not wrong!
+% PR: Sounds right (did some small fixes)
 Denoting the language described by \cref{fig:aut_a} as $\Language_{pre}$, we can

### 2023-07-10 preliminaries.tex
comments, bit of shortening

@@
 
-% @Philipp: reviewer comment #54 complains that this is basic, are they right?
+% AS: reviewer comment #54 complains that this is basic, are they right?
+
+% PR: sure, this is rather basic; but I think we should still include
+% those definitions
+
 A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
@@
 
-% @Philipp: Reviewers suggest we drop this:
+% AS: Reviewers suggest we drop this:
+% PR, shortened a bit
 Parikh's theorem states that any context-free language has a Parikh-equivalent
 regular language (c.f.~\cite{construction} for a construction of such automata
-from context-free grammars and~\cite{bounds} for bounds on its size). However,
-there are languages that are not context-free that also have semilinear images
-under~$\ParikhMap$ (e.g. $\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) =
-\ParikhMap((abc)^*) = \CountOf{a} = \CountOf{b} = \CountOf{c} \land \CountOf{a}
-\geq 0$). This means they can be represented as a quantifier-free Presburger
-formula.
-
-Note that while Parikh's theorem applies to context-free languages, in this
+from context-free grammars and~\cite{bounds} for bounds on its size).
+%However,
+%there are languages that are not context-free that also have semilinear images
+%under~$\ParikhMap$ (e.g. $\ParikhMap(\Set{a^nb^nc^n \SuchThat n \geq 0}) =
+%\ParikhMap((abc)^*) = \CountOf{a} = \CountOf{b} = \CountOf{c} \land \CountOf{a}
+%\geq 0$). This means they can be represented as a quantifier-free Presburger
+%formula.
+%
+While Parikh's theorem applies to arbitrary context-free languages, in this
 paper we focus only on regular languages.

### 2023-07-10 calculus.tex
Results in paper

@@
 everywhere and drop $\ldots \land x=y \land \ldots$) to keep the size of the
-equations managable.}\label{fig:derivation:single}
+equations manageable.}\label{fig:derivation:single}
 \end{figure}

### 2023-07-10 evaluation.tex
Results in paper

@@
   \vfill
-  \begin{subfigure}[b]{0.75\textwidth}
+  \begin{subfigure}[b]{\textwidth}
     \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
-    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
+    \caption{The number of instances solved as the time budget increases, simulated from one \numprint{120}s-timeout run.}
     \label{fig:cactus}
@@
 To get an idea of how the different backends scale, we additionally execute
-\numprint{2000} different randomly sampled benchmarks with a
-\numprint{120}-second timeout. In this set we also include two additional
-variants of \Calculus{}: one with clause learning and restarts disabled, and one
-with only clause learning disabled to show the impact of
-\cref{sec:random-restarts}, and \cref{sec:clause-learning} respectively.
+\numprint{2000} randomly sampled (without replacement) benchmarks with a
+\numprint{120}-second timeout. In addition to the previous back-ends, we also
+execute \Calculus{} with clause learning and restarts disabled, and with only
+clause learning disabled to show the impact of \cref{sec:random-restarts}, and
+\cref{sec:clause-learning} respectively.
 
@@
 be seen in \cref{fig:cactus}. We see that \Calculus{} outperforms \Nuxmv{} in
-general, but that \Nuxmv{} might have better scaling properties on longer runs.
-For machines with better single-core performance than our experiment machine
-\Nuxmv{}'s scaling benefits kicks in faster, so it is likely that the amount of
-CPU cycles spent on the problem is the underlying factor. This is likely due to
-two factors. First, \Nuxmv{} is more mature than \Calculus{}, and it makes sense
-that its more general model checking methods would pay off on more difficult
-instances where more problem-specific methods offer less payoff. Second, for
-longer running calculations clause learning as described in
-\cref{sec:clause-learning} can be assumed to matter more. However, the clauses
-we learn are still relatively unpolished, and in particular generalising
-insights across product computations is nontrivial and therefore partially left
-as future work.
+general but that \Nuxmv{} might scale better on very long runtimes.
+Non-systematic observations suggest that \Nuxmv{}'s scaling benefits kicks in
+faster on newer machines, so it is likely that the amount of CPU cycles spent is
+the underlying factor. This is likely due to two factors. First, \Nuxmv{} is
+more mature than \Calculus{}, and it makes sense that its more general model
+checking methods would pay off on more difficult instances where more
+problem-specific methods offer less payoff. Second, for longer running
+calculations clause learning as described in \cref{sec:clause-learning} can be
+assumed to matter more. However, the clauses we learn are still unpolished, and
+in particular, generalise poorly across product computations. Improvements are
+left as future work.
 

### 2023-07-10 implementation.tex
Results in paper

@@
 and automaton implementation as \Catra. We do this to better analyse the impact
-of the calculus rules themselves. Using the formula of
-\eqref{eq:generate-parikh}, we produce quantified Presburger formulae for each
-successive term and add them to \Princess. We compute the product incrementally
-term by term, checking satisfiability at each step. We use a priority queue to
-select automata for each step and order them by their number of transitions. We
-use this heuristic to put off computing large (and therefore slow) products until
-we have to, hoping to find an empty intermittent product.
+of the calculus rules themselves. Adapting~\cite{generate-parikh-image}, we
+produce quantified Presburger formulae for each successive term and add them to
+\Princess. We compute the product incrementally term by term, checking
+satisfiability at each step. We use a priority queue to select automata for each
+step and order them by their number of transitions. We use this heuristic to put
+off computing large (and therefore slow) products until we have to, hoping to
+find an empty intermittent product. This is roughly similar to the approach
+taken in~\cite{approximate-parikh}.
 

### 2023-07-10 implementation.tex
Drop claim about minimsation
Closes: #44 (in the bluntest possible way).

Ping @pruemmer for possibly a better claim.

@@
 either backend. More complex minimisation was left out since performing
-minimisation on automata with counters is non-trivial, and minimising symbolic
-automata risks exponential blowup~\cite{minimising-symbolic}.
+minimisation on automata with counters is non-trivial.
+
+% @ AS: I did the simple thing and dropped the claim about symbolic automata
+% here: ours are less symbolic (i.e. typically finite) than the ones in the
+% cited paper, and so are technically no worse than normal non-symbolic ones.
 

### 2023-07-11 introduction.tex
comments, bit of shortening
intro

@@
-Extending automated theorem provers ("SMT solvers") with support for rich string
-constraints is important for program analysis, in particular to detect
-cross-site scripting vulnerabilities and other string manipulation bugs.
-
+Extending automated theorem provers and SMT solvers with support for
+rich string constraints is important for program analysis, in
+particular to detect cross-site scripting vulnerabilities and other
+string manipulation bugs~\cite{DBLP:books/sp/BultanYAA17}.
+%
 To decide whether a formula like $x \in \Language_1 \wedge y \in \Language_2
@@
 about the sets of word lengths induced by $\Language_1, \Language_2$, which are
-special cases of the Parikh image. This required combined reasoning about
+special cases of the Parikh image of such languages.
+This required combined reasoning about
 strings and string length has long been identified as a major bottleneck in
@@
 Other string solvers make use of Parikh automata~\cite{parikh-automata}, and
-thus Parikh images in the general case, to handle operations including
-\verb!str.substr!  and \verb!str.at!, which comes at an even higher price in
+thus Parikh images in the general case, to handle operations that combine
+strings and integers (including
+\verb!str.substr!  and \verb!str.at!), which comes at an even higher price in
 terms of computational complexity~\cite{ostrich-plus}.
@@
 classical result that the Parikh image of every context-free language (and,
-thus, also of every regular language) is a semilinear set~\cite{parikh-theorem},
-i.e., Presburger-definable.
+thus, also of every regular language) is a semilinear set,
+i.e., Presburger-definable~\cite{parikh-theorem}. As mentioned,
+Parikh images are
+
+ Beyond the field
+of string constraint solving
 
@@
 We implement \Calculus{} as a plug-in theory for the \Princess{} automated
-theorem prover. \Fudge{Others} have integrated it into the similarly
-\Princess{}-backed string constraint solver \OstrichPlus{}, where it is used to
-handle constraints that use cost-enriched automata, as described in
-\cite{ostrich-plus}.
-
+theorem prover.
+%
 We additionally wrap the Parikh image solver as a stand-alone tool, \Catra.
@@
 
+% PR: maybe only mention this in the evaluation?
+
+% \Fudge{Others} have integrated it into the similarly
+%\Princess{}-backed string constraint solver \OstrichPlus{}, where it is used to
+%handle constraints that use cost-enriched automata, as described in
+%\cite{ostrich-plus}.
+
 In summary, we contribute:
@@
           the Parikh image of products of Parikh automata.
-    \item Experiments illustrating the performance of \Calculus{} on real-world
-    examples from string solving, including \NrBenchmarks{} instances in a
-    standardised format made available for future study.
+    \item Techniques to efficiently implement \Calculus{} in a modern
+    automated theorem prover, including strategies for case splitting, clause
+    learning, and constraint propagation for connectedness.
     \item The \Catra{} tool for solving such instances, containing an
@@
     checker~\cite{nuxmv}.
-    \item Suggestions for how to efficiently implement \Calculus{} in a modern
-    automated theorem prover, including strategies for case splitting, clause
-    learning, and constraint propagation for connectedness.
+    \item Experiments illustrating the performance of \Calculus{} on real-world
+    examples from string solving, including \NrBenchmarks{} instances in a
+    standardised format made available for future study.
 \end{itemize}
@@
 
-% @Philipp: please verify that this is not a lie!
+% AS please verify that this is not a lie!
+
+% PR from looking quickly at this article, it seems to solve a
+% completely different problem?
+
 The problem and our calculus, \Calculus{}, takes a similar approach to the work
@@
 us to \emph{remove} something, thereby potentialy obtaining an easier problem.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "main"
+%%% End:

### 2023-07-11 main.tex
intro

@@
   A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
+   linear arithmetic formula that describes all possible combinations of character
    counts in strings of a given language. Automata-based string solvers

### 2023-07-11 introduction.tex
intro

@@
 thus, also of every regular language) is a semilinear set,
-i.e., Presburger-definable~\cite{parikh-theorem}. As mentioned,
-Parikh images are
-
- Beyond the field
-of string constraint solving
+i.e., Presburger-definable~\cite{parikh-theorem}.
 
@@
 The problem of computing constraints on Parikh images over products of regular
-languages under a given commutative homomorphism amounts to solving products of
+languages under a given commutative homomorphism amounts to checking emptiness
+of products of
 Parikh automata. Parikh automata are regular automata extended with integer
@@
 without constraints on the final values on their registers are also sometimes
-called cost-enriched automata, weighted automata or counter automata, depending
-on exact definitions and side-constraints. The decision problem tackled in this
+called cost-enriched automata, weighted automata, or counter automata, depending
+on exact definitions and side constraints. The decision problem tackled in this
 paper, determining the emptiness of an intersection of Parikh automata, was
@@
 
-Parikh image computations as well as Parikh automata feature extensively in
+Parikh image computations, as well as Parikh automata, feature extensively in
 string solvers, including as mentioned above \Ostrich{} and \OstrichPlus{}
-\cite{ostrich,ostrich-plus}, but also forms the basis of Trau~\cite{trau-pldi},
+\cite{ostrich,ostrich-plus}, but also forms the basis of
+\textsc{Trau}~\cite{trau-pldi},
 and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images frequently appear when
@@
 with the conjunction of the automata's Parikh maps, $\bigwedge_{i=1}^{k}
-\ParikhMap(\Automaton_i)$. This approach works only for unsatisfiable instances,
+\ParikhMap(\Automaton_i)$. Due to the over-approximation,
+this approach works only for unsatisfiable instances,
 and will require falling back to computing the product of the automata before
@@
 
-% AS please verify that this is not a lie!
-
-% PR from looking quickly at this article, it seems to solve a
-% completely different problem?
-
-The problem and our calculus, \Calculus{}, takes a similar approach to the work
-of \citeauthor{symbolic-boolean-derivatives}, who present an orthogonal calculus
-for handling a similar blow-up in complexity resulting from Boolean combinations
-of regular automata in SMT string solvers \cite{symbolic-boolean-derivatives}.
+Our calculus~\Calculus{} is similar in spirit to the work
+of \citeauthor{symbolic-boolean-derivatives}, who tackle the exponential
+blow-up resulting from Boolean combinations
+of regular automata in SMT string solvers
+through the use of symbolic derivatives~\cite{symbolic-boolean-derivatives}.
 In addition to presenting a decision procedure for lazily dispatching
@@
 
-Outside of string solvers, Parikh automata have been proposed as the basis of
-queries~\cite{graph-queries}, and for solving cardinalities in model checking
-problems involving epistemic logic~\cite{epistemic-logic}.
+Beyond the field of string solving, Parikh image computation is used
+as an elementary building block in a variety of areas.  For instance,
+Parikh automata have been proposed as the basis of queries in graph
+databases~\cite{graph-queries}; Parikh images are used for handling
+cardinalities in parameterised model checking for epistemic
+logic~\cite{epistemic-logic}; or for handling summation constraints in
+expressive array logics~\cite{rodrigoRaya}. Our calculus~\Calculus{},
+and the stand-alone solver~\Catra{},
+are potentially useful in all such applications.
 
-Many other generalisations of the Parikh image than the projections we use here
+Other generalisations of the Parikh image than the projections we use here
 have been studied. Prominent examples include generalising the Parikh map to
-segments of a fixed length \cite{KARHUMAKI1980155} and the more general Parikh
+segments of a fixed length~\cite{KARHUMAKI1980155} and the more general Parikh
 matrix, which contains not only the Parikh vector, but also information about
@@
 infinite alphabets. All of these in some sense extend the Parikh map. By
-contrast, the main utility of the formulation introduced here is that it allows
-us to \emph{remove} something, thereby potentialy obtaining an easier problem.
+contrast, the main utility of the formulation introduced here is to reason
+about Parikh images \emph{lazily,}
+thereby potentially obtaining answers more quickly. We expect that our
+calculus~\Calculus{} can be generalised to the mentioned functions on
+formal languages as well, but leave such investigations to future work.
 

### 2023-07-11 example.tex
more refs

@@
-In this section we will introduce an intuition for how a string constraint
+In this section, we will introduce an intuition for how a string constraint
 problem in an automata-based solver like \OstrichPlus{} is translated to a
-Parikh automata intersection problem and solved using \Calculus{}.
+Parikh automata intersection problem, and solved using \Calculus{}.
 
 We use PCRE regular expression notation here and throughout the paper, writing
-them $\Regex{\texttt{like this}}$. This means that $\RegexOr{}$ is alternation,
-$\mathtt{*}$ the Kleene star, and $\mathtt{.}$, matches any single character.
+them $\Regex{\texttt{like this}}$. This means that $\Regex{\RegexOr{}}$ is alternation,
+$\Regex{\mathtt{*}}$ the Kleene star, and $\Regex{\mathtt{.}}$ matches any single character.
 For the length of a string $s$, we write $\Length{s}$. All number variables in
@@
 \begin{example}\label{ex:string-constraints} Consider the following set of
-    string constraints:
+    string constraints over string variables~$s_1, s_2$ and integer
+    variables~$n, i$:
 \begin{constraints}
@@
     \item\label{const:something-before-and-after} $i > 0 \land \Length{s_2} -i -n > 0$, that
-    is there is at least one character in $s_2$ after $s_1$.
+    is there is at least one character in $s_2$ before and after $s_1$.
 \end{constraints}
@@
 
-To solve these constraints, we will first perform translation into Parikh
+Although the constraints are simple, it should be noted that
+decidability results for string constraints involving integers are
+notoriously hard to obtain; most state-of-the-art string solvers, for
+instance Z3~\cite{Z3}, different versions of
+Z3-str~\cite{Z3-str,DBLP:conf/fm/MoraBKNG21}, or
+cvc5~\cite{cvc5}
+are not guaranteed to be complete or to terminate on constraints of this kind.
+
+A decision procedure for constraints of this kind is defined in
+\cite{ostrich-plus}.
+%
+To solve the constraints, we will first perform translation into Parikh
 automata, then break down how the constraints will be handled by \Calculus{}.
@@
 constraints as in~\cite{generate-parikh-image}.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "main"
+%%% End:

### 2023-07-11 evaluation.tex
Ostrich evaluation sketch

@@
 
+\subsection{Evaluation in \Ostrich{}}\label{}%
+\Catra{} was experimentally integrated as the Parikh automata product solver for
+an in-progress version of \Ostrich{} (which shares \Princess{} as the backing
+theorem prover) with cost-enriched automata (CEFA). \Ostrich{} (\textit{sans}
+CEFA) recently participated in the 2023~SMT-COMP, winning the single-query track
+for quantifier-free strings (\texttt{QF\_S}) as well as the entire string
+solving track on unsatisfiable instances~\cite{smt-comp-23}.
+
+The \texttt{QF\_S} track contains no Parikh automata intersection problems,
+however. For that, we need to look to the linear integer arithmetic
+(\texttt{QF\_SLIA}), where \Ostrich{} did worse in SMT-COMP. The evaluations in
+\cref{sec:scaling} use inputs generated by an older branch of CEFA-enriched
+\Ostrich{} on the PyEx benchmarks, which are part of the set used in the
+competition. In this section, we evaluated a version of \Ostrich{} using
+\Catra{} as its back-end \Fudge{on StarExec, the same cluster SMT-COMP ran on}
+on \numprint{1000}~randomly selected instances from the PyEx benchmarks. The
+results are shown in comparison \Fudge{to the other provers of the 2023
+competition on the same benchmarks, as well as \Ostrich{} using the baseline
+back-end described in \cref{sec:implementing-baseline}}.
+
+\Fudge{RESULTS HERE PLEASE!}
+
 \subsection{Threats to Validity}

### 2023-07-11 experiments/analyse-smtlib.py

@@ -1,0 +1,31 @@
+#!/usr/bin/env python3
+
+from pathlib import Path
+import sys
+import sys
+
+def classify_file(fp):
+    present = []
+    for line in fp:
+        if "str.len" in line:
+            present.append("str.len")
+        if "str.replace" in line:
+            present.append("str.replace")
+        if "str.indexof" in line:
+            present.append("str.indexof")
+        if "str.substr" in line:
+            present.append("str.substr")
+
+    return present
+
+files = []
+for file in sys.stdin:
+    file = Path(file.strip())
+    with file.open() as fp:
+        try:
+            files.append((file, classify_file(fp)))
+        except UnicodeDecodeError:
+            continue
+
+for file, constraints in sorted(files, key=lambda f: len(f[1])):
+    print("\t".join([str(file), ",".join(constraints)]))

### 2023-07-11 experiments/select-experiments.py

@@ -1,0 +1,143 @@
+#!/usr/bin/env python3
+
+import subprocess
+import sys
+import os
+import time
+
+
+TIMEOUT_S = 60
+JVM_OPTS = [
+    "-Xss20000k",
+    "-Xmx2000m",
+]
+
+
+def z3(benchmark):
+    return subprocess.Popen(
+        ["/usr/bin/time", "z3", f"-T:{TIMEOUT_S}", benchmark],
+        stdout=subprocess.PIPE,
+        stderr=subprocess.PIPE,
+        text=True,
+    )
+
+
+def cvc(benchmark):
+    return subprocess.Popen(
+        ["/usr/bin/time", "cvc4", "--strings-exp", f"--tlimit={TIMEOUT_S * 1000}", benchmark],
+        stdout=subprocess.PIPE,
+        stderr=subprocess.PIPE,
+        text=True,
+    )
+
+
+def ostrich(benchmark):
+    return subprocess.Popen(
+        [
+            "/usr/bin/time",
+            "java",
+            *JVM_OPTS,
+            "-jar",
+            os.path.expanduser("~/ostrich/target/scala-2.13/ostrich-assembly-1.3.jar"),
+            "-stringSolver=ostrich.OstrichStringTheory:",
+            "+quiet",
+            f"-timeout={TIMEOUT_S * 1000}",
+            benchmark,
+        ],
+        text=True,
+        stdout=subprocess.PIPE,
+        stderr=subprocess.PIPE,
+    )
+
+
+def ostrich_catra(benchmark):
+    return subprocess.Popen(
+        [
+            "/usr/bin/time",
+            "java",
+            *JVM_OPTS,
+            "-jar",
+            os.path.expanduser("~/ostrich/target/scala-2.13/ostrich-assembly-1.3.jar"),
+            "-stringSolver=ostrich.cesolver.stringtheory.CEStringTheory:-ceaBackend=catra",
+            "+quiet",
+            f"-timeout={TIMEOUT_S * 1000}",
+            benchmark,
+        ],
+        text=True,
+        stdout=subprocess.PIPE,
+        stderr=subprocess.PIPE,
+    )
+
+def poll_longer(proc):
+    try:
+        return proc.wait(timeout=5)
+    except subprocess.TimeoutExpired:
+        return None
+
+
+def drain_all(processes):
+    num_processes = len(processes)
+    while True:
+        finished = 0
+        for p in processes:
+            if poll_longer(p) == None:
+                p.terminate()
[…]

### 2023-07-11 example.tex
explanation

@@
     \item\label{const:s2-in-b} $s_2 \in \Language(B)$, the language accepted by
-    automaton~$B$ of \cref{fig:aut_b}.
+    automaton~$B$ of \cref{fig:aut_b} (ignoring
+    the counter~$l_B$).
     \item\label{const:s1-substring} $s_1 = \Substring(s_2, i, n)$, that is $s_1$ is an
@@
 decidability results for string constraints involving integers are
-notoriously hard to obtain; most state-of-the-art string solvers, for
+notoriously hard to obtain; many state-of-the-art string solvers, for
 instance Z3~\cite{Z3}, different versions of
@@
 cvc5~\cite{cvc5}
-are not guaranteed to be complete or to terminate on constraints of this kind.
+are not guaranteed to be complete or to terminate on such constraints
+(but, of course, they often achieve very good performance in practice).
 
-A decision procedure for constraints of this kind is defined in
+A decision procedure for a fragment of string constraints with integers,
+covering our constraints, is defined in
 \cite{ostrich-plus}.
 %
-To solve the constraints, we will first perform translation into Parikh
+We will first perform translation into Parikh
 automata, then break down how the constraints will be handled by \Calculus{}.
@@
 
-To solve \cref{ex:string-constraints} using \Calculus{} in a theorem proving
-context, \cref{const:s1-in-c-dd,const:s2-in-b,const:s1-substring} can be
-translated to the product of Parikh automata $B \times A'$, where $A'$
-(\cref{fig:aut_a}) is the Parikh automata obtained by applying the encoding of
-$\Substring$ described in \cite{ostrich-plus} to
-$\Language(\Regex{c\RegexOr{}dd})$. That description can then be used to obtain
-values for the substring offsets $i, n$. The corresponding automata can be found
-in \cref{fig:examples}.
+To solve \cref{ex:string-constraints} using \Calculus{} in a theorem
+proving context,
+\cref{const:s1-in-c-dd,const:s2-in-b,const:s1-substring} can be
+translated to the product of Parikh automata $B \times
+A'$. Automaton~$B$ is given in \cref{fig:aut_b}.  Automaton~$A'$
+(\cref{fig:aut_a}) is an automaton defining the pre-image of
+$\Language(\Regex{c\RegexOr{}dd})$ under the $\Substring$ function,
+and obtained by applying the construction described in
+\cite{ostrich-plus}. Intuitively, $A'$ describes all
+tuples~$(s_2, i, n)$ such that $\Substring(s_2, i, n)$ is in
+$\Language(\Regex{c\RegexOr{}dd})$. To this end, $A'$ contains
+registers~$l_A, i, n$, which count the length of the overall string
+read, the length of the prefix before the extracted substring, and the
+length of the substring, respectively. Registers are assumed to start
+from zero in each automaton run.
 
-In the automata of \cref{fig:examples}, we will use the notation $a /
+In the automata of \cref{fig:examples}, we use the notation $a /
 \begin{bmatrix} x_+ \end{bmatrix}$ to mean that a transition reads an input
@@
 
-Intuitively, \cref{ex:string-constraints} are unsatisfiable. By inspecting the
+Intuitively, \cref{ex:string-constraints} is unsatisfiable. By inspecting the
 automata, we realise that the path using $\Char{dd}$ in $A'$ is unusable since
@@
 enforcement of connectivity constraints (essentially consisting of a constraint
-programmgin-style propagator), and lazy computation of products of automata.
+programming-style propagator), and lazy computation of products of automata.
 

### 2023-07-11 example.tex
more polishing

@@
     automaton~$B$ of \cref{fig:aut_b} (ignoring
-    the counter~$l_B$).
+    the registers~$l_B$).
     \item\label{const:s1-substring} $s_1 = \Substring(s_2, i, n)$, that is $s_1$ is an
@@
 
-The approach consists of four steps: first we approximately describe each
+Our approach proceedings in multiple steps. Similarly as in
+\cite{generate-parikh-image}, we first approximately describe each
 automaton as a transition-counting flow from the initial state to its accepting
-states where we relate its transitions to integer variables and register values
-to their transitions. Then we will use the linear algebra part of the theorem
-prover to discover that some paths through one of the automata are incompatible
-by \cref{const:more-inside-than-before,const:something-before-and-after}. This
+states, relating the automaton transitions to integer variables.
+We use a linear integer solver to discover that some paths through the automata are incompatible
+with \cref{const:more-inside-than-before,const:something-before-and-after}. This
 will allow us to discover an unreachable state, which allows further
@@
 We associate each transition $\Transition$ of $A'$ and $B$ with a fresh
-nonnegative integer variable from, $\TransitionVar_\Transition$, identically to
-\cite{generate-parikh-image}. These variables represent how many times each
-transition is taken. Hence, the final counter value for counters $r_1, \ldots,
+variable~$\TransitionVar_\Transition$ ranging over natural numbers (i.e.,
+non-negative integers).
+These variables represent how many times each
+transition is taken. Hence, the final value for registers~$r_1, \ldots,
 r_k$ of each automaton is the element-wise sum:
@@
 that the number of incoming transitions is equal to the number of outgoing ones.
-E.g. state $B$ would have the sum $\cancel{x_{\Tuple{B, \Char{c}, B}}} +
+E.g., state $B$ of automaton~$B$ would have the sum
+%
+\begin{equation}\label{eq:flow-eqs}
+  \cancel{x_{\Tuple{B, \Char{c}, B}}} +
   x_{\Tuple{I_B, \Char{b}, B}} = \cancel{x_{\Tuple{B, \Char{c}, B}}}  +
-x_{\Tuple{B, \Char{c}, F_B}} + x_{\Tuple{B, \Char{b}, I_B}}$ where we let
+  x_{\Tuple{B, \Char{c}, F_B}} + x_{\Tuple{B, \Char{b}, I_B}}
+\end{equation}
+%
+where we let
 $x_{\Tuple{q, l, c'}}$ refer to the integer variable associated with a
-transition from state $q$ to state $q'$ with label $l$. Note that the self-loop
-cancels itself out. This means that for automata where a loop can become
+transition from state $q$ to state $q'$ with label $l$. The initial state
+of the automaton receives an additional inflow of $1$, and accepting
+states have an additional common outflow of $1$. Note that self-loops
+cancels themselves out. This means that for automata where a loop can become
 unreachable, additional constraints would be required to ensure reachability.
@@
 
-\begin{figure}[ht]
-  \centering 
-  \includegraphics[width=0.6\textwidth]{a_1}
-  \caption{ $A'$ with its associated transition variables in symbolic form after
-  minimal elminiation of variables using linear algebra. Note the incoming flow
-  of $1$ in the starting state and its balancing of the outgoing transitions!
-  }\label{fig:a_1}
-\end{figure}
+%\begin{figure}[tb]
+%  \centering 
+%  \includegraphics[width=0.6\textwidth]{a_1}
+%  \caption{ $A'$ with its associated transition variables in symbolic form after
+%  minimal elimination of variables using linear algebra. Note the incoming flow
+%  of $1$ in the starting state and its balancing of the outgoing transitions!
+%  }\label{fig:a_1}
+%\end{figure}
 
 \subsubsection{Using linear algebra to eliminate flow}\label{sec:intuition:algebra}
-As an example, we start with reasoning about $A'$. Performing symbolic reasoning
-on these equations and plugging the corresponding symbolic representation of the
-constraints on each corresponding integer variable back as an annotation on its
-transition, we obtain the automaton in \cref{fig:a_1}. After slightly more
-linear algebra to remove variables through equality elimination, we have
-\cref{fig:a_2}, whose transitions are only described in terms of the free
-variables we care about representing the length of the automaton ($l_A$), the
+
+We then use linear arithmetic reasoning on register
+equations~\eqref{eq:counter-sums} and flow
+equations~\eqref{eq:flow-eqs} to simplify the counters associated
+with transitions.
+As an example, we start with $A'$. Substituting back
+solutions to the equations for $A'$ to the automaton, we obtain the
+automaton in \cref{fig:a_2}, where the notation~$a / t$ now denotes
+a transition that accepts letter~$a$ and is taken~$t$ times.
+In this case, the transitions of the automaton
+can be specified purely in terms of the free
+variables we care about representing:  the length of the automaton ($l_A$), the
 start of the substring ($i$), and the length of the substring ($n$).
 
-\begin{figure}[ht]
+\begin{figure}[tb]
   \centering 

### 2023-07-11 evaluation.tex
Evaluation + kill your darlings
Fix numbers issues, new graph

@@
 solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
-initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
-five seconds by baseline. We also attempted to benchmark instances generated by
-\Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len
-(\numprint{791} instances) benchmark suites, but had to discard them since they
-were all trivial. For instance, on a laptop, \Calculus{} could solve every
-instance from kaluza-len in under five seconds, mean about 0.1s. The benchmarks
-are run on commit~\texttt{\commit}.
+initial~\InitialNrBenchmarks{}, we remove~\NrTrivial{} instances solved in under
+five seconds by baseline as well as~\NrInvalid{} instances that produced errors
+during experimentation. We also attempted to benchmark instances generated by
+\Ostrich{} solving the kaluza-len (\numprint{38227}~instances) and pyex-len
+(\numprint{791}~instances) benchmark suites but discarded them since they were
+all trivial. For instance, on a laptop, \Calculus{} could solve every instance
+from kaluza-len in under five seconds (mean about 0.1s). The benchmarks are run
+on commit~\texttt{\commit}.
 
@@
 machines, \BenchmarkRig{}, but as the sole job for each machine.  We compiled
-the code using Scala~\ScalaVersion{}, and executed the experiments
+the code using Scala~\ScalaVersion{} and executed the experiments
 on~\JvmVersion{} with a maximum heap of~\MaxHeapSize{}. We used \Nuxmv{}
@@
 to execute. We believe this represents a realistic use case where \Calculus{} is
-used to support e.g a string solver. Experiments were executed in random order
+used to support, e.g., a string solver. Experiments were executed in random order
 for all backends. Each instance got a time budget of~\RuntimeTimeout.
@@
 
-In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances
+In \cref{fig:solve-division}, we show how many of the~\NrBenchmarks{} instances
 the respective back-end could solve. A summary of their outcomes by instance
@@
 
+We include this section to show that \Calculus{} can be used to enhance the
+performance of an automaton-based string solver with cost-enriched automata like
+\Ostrich{}. However, significant engineering effort remains to bring the
+integration to a state where it can participate in the full SMT-COMP.
+
 \Fudge{RESULTS HERE PLEASE!}

### 2023-07-11 experiments/select-experiments.py

@@ -141,3 +141,5 @@ for benchmark in sys.argv[1:]:
     results = wait_all(benchmark)
     columns = [benchmark, *[f"{solver}={summarise_results(res)}" for solver, res in results.items()]]
     print(" || ".join(columns))
+    with open("experiments.log", "a") as logfile:
+        logfile.write(" || ".join(columns) + "\n")

### 2023-07-11 example.tex
more description

@@
 
-Our approach proceedings in multiple steps. Similarly as in
-\cite{generate-parikh-image}, we first approximately describe each
-automaton as a transition-counting flow from the initial state to its accepting
-states, relating the automaton transitions to integer variables.
-We use a linear integer solver to discover that some paths through the automata are incompatible
-with \cref{const:more-inside-than-before,const:something-before-and-after}. This
+Our approach proceedings in multiple steps:
+\begin{inparaenum}[(i)]
+\item Similarly as in \cite{generate-parikh-image}, we first describe
+  each automaton as a network-flow problem, counting how often each
+  transition is taken in a run of the automaton.
+\item   We use a linear integer solver to discover that some
+  paths through the automata are incompatible with
+  \cref{const:more-inside-than-before,const:something-before-and-after}. This
   will allow us to discover an unreachable state, which allows further
-elimination. Finally, we will use case splitting to further subdivide the
-problem into bite-sized products of automata, both of which are empty, thus
-concluding that \cref{ex:string-constraints} is unsatisfiable.
+  elimination. Finally, we will use case splitting to further
+  subdivide the problem into bite-sized products of automata, both of
+  which are empty, thus concluding that \cref{ex:string-constraints}
+  is unsatisfiable.
+\end{inparaenum}
 

### 2023-07-11 example.tex
refined description

@@
 
-We will now proceed to show how we prove this using the three principles of
-\Calculus{}: problem-aware case splitting (sometimes called branching), lazy
-enforcement of connectivity constraints (essentially consisting of a constraint
-programming-style propagator), and lazy computation of products of automata.
+%We will now proceed to show how we prove this using the three principles of
+%\Calculus{}: problem-aware case splitting (sometimes called branching), lazy
+%enforcement of connectivity constraints (essentially consisting of a constraint
+%programming-style propagator), and lazy computation of products of automata.
 
-Both lazinesses are in contrast to the eager approach to finding the Parikh
-image described in~\cite{generate-parikh-image}, that would have started by
-computing the product $A' \times B$, translated it to a Presburger formula with
-$i, n$ etc as free variables, and then plugged in
+An eager approach to finding the Parikh
+image, as described in~\cite{generate-parikh-image}, would start by
+computing the product $A' \times B$, translating it to a Presburger formula with
+$i, n$, etc., as free variables, and then adding
 \cref{const:more-inside-than-before,const:something-before-and-after} to the
-resulting set of linear inequalitites.
+resulting set of linear inequalitites. Scalability of this
+construction is limited due to the size of the product automata to be computed,
+and the complexity of the resulting Presburger formula.
 
@@
 
-Our approach proceedings in multiple steps:
+We will now proceed to show how our calculus~\Calculus{} can lazily
+prove the unsatisfiability of the running example. The calculus
+interleaves several reasoning principles, which we will later
+define precisely as calculus rules:
 \begin{inparaenum}[(i)]
 \item Similarly as in \cite{generate-parikh-image}, we first describe
-  each automaton as a network-flow problem, counting how often each
-  transition is taken in a run of the automaton.
-\item   We use a linear integer solver to discover that some
-  paths through the automata are incompatible with
-  \cref{const:more-inside-than-before,const:something-before-and-after}. This
-  will allow us to discover an unreachable state, which allows further
-  elimination. Finally, we will use case splitting to further
-  subdivide the problem into bite-sized products of automata, both of
-  which are empty, thus concluding that \cref{ex:string-constraints}
-  is unsatisfiable.
+  each automaton as a flow network, counting how often each transition
+  is taken in a run of the automaton. The flow constraints are an
+  over-approximation of the possible accepting runs of an automaton.
+\item We use a linear integer solver to simplify the flow constraints,
+  and prune away paths through the automata that are not feasible.
+\item We use a tailor-made propagation algorithm to identify disconnected
+  transitions of the automata that can never be taken, and lazily add
+  the corresponding constraints.
+\item When propagation cannot infer further constraints, we use case
+  splitting to sub-divide the problem into smaller parts.
+\item Once the constituent automata of a product have become
+  sufficiently small, we compute the precise product automaton.
 \end{inparaenum}
 
-\subsubsection{Eliminating paths through automata using flow analysis}\label{sec:a_1}
+\subsubsection{Approximating paths through automata using flow analysis}\label{sec:a_1}
 
@@
 
-\subsubsection{Using linear algebra to eliminate flow}\label{sec:intuition:algebra}
+\subsubsection{Arithmetic flow simplification}\label{sec:intuition:algebra}
 
@@
 can be specified purely in terms of the free
-variables we care about representing:  the length of the automaton ($l_A$), the
+variables we care about representing:  the length of the accepted word ($l_A$), the
 start of the substring ($i$), and the length of the substring ($n$).
@@
 
-Having obtained this representation, we conclude that $1 < n \leq 2$ from
-\cref{const:something-before-and-after,const:something-before-and-after}. The
-upper bound is obtained by the following reasoning. Since all transition
-variables are nonnegative integers (a transition cannot be used a negative
-amount of times), $n \leq 2$ from the $I_A$ to $F_A$ transition. Therefore, it
-follows that $n=2$, which implies that the $I_A$ to $F_A$ transition can never
-be used under these constraints and that we must use the two $\Char{d}$-labelled
-transitions, both now taken $n-1 =1$ times. This immediately implies that the problem is
-unsatisfiable, since we know this transition has no correspondance in automaton
-$B$.
-
-We repeat similar methods to arrive at the automata in \cref{fig:propagated}.
+Having obtained this representation, we conclude that $1 < n \leq 2$
+from
+\cref{const:more-inside-than-before,const:something-before-and-after}. The
+lower bound directly follows from
+\cref{const:more-inside-than-before,const:something-before-and-after},
+whereas the upper bound is obtained by the following reasoning. Since all
+transition variables are non-negative integers (a transition cannot be
+used a negative amount of times), $n \leq 2$ from the $I_A$ to $F_A$
+transition. Therefore, it follows that $n=2$, which implies that the
+$I_A$ to $F_A$ transition can never be used under these constraints
+and that we must use the two $\Char{d}$-labelled transitions, both now
+taken $n-1 =1$ times.
+% This immediately implies that the problem is
+%unsatisfiable, since we know this transition has no correspondance in
+%automaton $B$.
+%
+Substituting the derived value of $n$, and applying similar
+reasoning to automaton~$B$, we arrive at the simplified
+automata in \cref{fig:propagated}.
 
@@
     \centering 
-  \begin{subfigure}[b]{\autscale\textwidth}
+  \begin{subfigure}{0.42\textwidth}
     \centering
     \includegraphics[width=\textwidth]{a_annotated}
+    \vspace*{5ex}
     \caption{ $A'$ with its associated transition variables in symbolic
@@
       %\Description[SHORT]{LONG}
+    \vspace*{1.9ex}
   \end{subfigure}\hfill%
-  \begin{subfigure}[b]{\autscale\textwidth}
+  \begin{subfigure}{0.54\textwidth}
     \centering
@@
     \caption{$B$ with its associated transition variables in symbolic forms.
-    Note the large number of implicitly existence-quantified bound variables on
-    transitions, suggesting that this automaton has a more complex structure
+    The large number of implicitly existentially quantified
+    variables~$x_i$ on
+    transitions  suggests that this automaton has a more complex structure
     with relation to its (free) target variable representing the string length
-    than $A'$. The variables used here are fresh, and do not correspond to the
-    ones previously used.}\label{fig:aut_b_annotated}
+    than $A'$. %The variables used here are fresh, and do not correspond to the
+    % ones previously used.
+  }\label{fig:aut_b_annotated}
       %\Description[SHORT]{LONG}
   \end{subfigure}
-  \caption{The automata after applying linear algebra and equality elimination
-  to prune them.}\label{fig:propagated}
+  \caption{The automata after arithmetic simplification.}\label{fig:propagated}
 \end{figure}
@@
 
-\subsubsection{Subsumption of connectivity constraints}\label{sec:intuition:subsume}
-At this point we can conclude that none of the loops of $A'$ can become
-disconnected since they are at the initial and final states, both of which are
-reachable for any future assignments of the transition variables. This means
-that all valid counter assignments of $A'$ are described by
-\cref{eq:counter-sums} and the flow equations alone without the need for
-expensive connectivity constraints.
+%\subsubsection{Connectivity reasoning}\label{sec:intuition:subsume}
+%At this point we can conclude that none of the loops of $A'$ can become
+%disconnected since they are at the initial and final states, both of which are
+%reachable for any future assignments of the transition variables. This means
+%that all valid counter assignments of $A'$ are described by
+%\cref{eq:counter-sums} and the flow equations alone without the need for
+%expensive connectivity constraints.
 
-\subsubsection{Case Splitting and propagation of connectivity constraints}\label{sec:intuition:split}
-After concluding this, we have a choice of two paths through automaton $B$; the
-upper through state $A$ or the lower through state $B$. We case split on a
-problem-aware basis by selecting a transition variable that would disconnect
+\subsubsection{Case Splitting and connectivity propagation}\label{sec:intuition:split}
+We now have a choice of two paths through automaton $B$; the
+upper through state $A$ or the lower through state $B$. Since
+arithmetic reasoning and propagation are not able to resolve this choice,
+we perform a case split by selecting a transition variable that would disconnect
 some strongly connected component of automaton $B$, in this case the transition
-form $I_B$ to state $B$. We divide it into the cases used ($> 0$) or unused ($ =
-0$) since those are the threshold values of the calculus, preferring the latter
-option on the first-fail principle.
+from~$I_B$ to state $B$ guarded by variable~$x_{12}$.
+ We split the reasoning into the cases $x_{12} > 0$ (transition used)
+and $x_{12} = 0$ (transition unused). For presentation, we focus
+on the latter case, as the former can be handled in a similar way.
 
-Having performed the split, we can conclude that the path from $I_B$ to state
-$B$ is now unreachable without passing through a zero-times-used transition.
-This means that the self-loop from state $B$ is also unusable, and thus must
-also be used zero times.
+In the case~$x_{12} = 0$ we can conclude
+that the state~$B$ is now unreachable, which means that also
+transitions outgoing from state $B$ are now unusable. Propagation can
+therefore infer the additional equations $x_9 = 0$ and $x_{11} = 1$.
 
 \subsubsection{Computing products}\label{sec:intuition:materialise}
-Now we have two small, stick-like automata after discounting all
-transitions with a corresponding transition variable that is zero, and can try
-to compute their product. Doing so, we will immediately notice that the
-$\Char{d}$-labelled transition of automaton $A'$ has no correspondence in $B$,
-leading to an empty product. We close the proof goal and backtrack with the
-opposite constraint, choosing the upper $I_B$ to $A$ transition this time, the
-only potentially productive choice we have left. However, proceding with
-calculation of the product from the almost identical terms will be similarly
-improductive, and we must conclude that the problem is unsatisfiable.
-
-By putting off computing the product $B \times A'$ until after performing linear
-reasoning on the number of times each transition must be taken and using that to
-prune the terms, we have computed a smaller product than we would have with a
-na\"ive approach. Moreover, since this automaton contains only one loop along
-one main path, all runs through it can be described by the flow-preservation
-equations described above without the need for costly connectivity-enforcing
-constraints as in~\cite{generate-parikh-image}.
+After discounting all transitions that can no longer be taken, we are
+left with two (small) flat automata, and can compute their product with
+relative ease. By putting off computing the product $B \times A'$ until after
+performing linear reasoning, and using that to prune transitions, we
+have computed a smaller product than we would have with an eager
+approach. Computing the product, we will immediately notice that the
+$\Char{d}$-labelled transition of automaton $A'$ has no correspondence
+in $B$, leading to an empty product. We can close the proof goal and
+backtrack, and will eventually derive that the imposed constraints are
+unsatisfiable.
 

### 2023-07-11 preliminaries.tex
Grammar check
some clean-up

@@
 A monoid $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ is an algebraic structure
-consisting of the non-empty carrier set of elements, $X$, an associative binary operation
+consisting of the non-empty carrier set of elements, $X$; an associative binary operation
 $X \times X \rightarrow X$ denoted as $\MonoidOp$, that is where for all $a, b,
-c \in X$, $(a \MonoidOp b) \MonoidOp c = a \MonoidOp (b \MonoidOp c)$. Finally,
-$\Monoid$ must have an identity element $0_{\Monoid} \in X$ such that
+c \in X$, $(a \MonoidOp b) \MonoidOp c = a \MonoidOp (b \MonoidOp c)$; 
+and an identity element $0_{\Monoid} \in X$ such that
 $0_{\Monoid} \MonoidOp a = a \MonoidOp 0_{\Monoid} =   a$ for every $a \in X$.
-We sometimes use integer multiplication to represent repeated application of
+We sometimes use integer multiplication to represent a repeated application of
 $\MonoidOp$, e.g. $3a = a \MonoidOp a \MonoidOp a$, for $a \in X$. $\Monoid$ is
@@
 
-Finally, a \textit{homomorphism} is a structure-preserving map between two monoids
-$\Monoid_1, \Monoid_2$, that is a map $\Map : S_1 \rightarrow S_2$ such that $\Map(a
-\MonoidOp_1 b) = \Map(a) \MonoidOp_2 \Map(b)$, where $S_1, S_2$ are the carrier
-sets of $\Monoid_1$ and $\Monoid_2$ respectively, and $\MonoidOp_1, \MonoidOp_2$
-their binary operations.
+A \textit{homomorphism} is a structure-preserving map between two
+monoids $\Monoid = \Tuple{X;\MonoidOp;0_{\Monoid}}$ and
+$\Monoid' = \Tuple{X';\MonoidOp';0_{\Monoid'}}$, that is a map
+$\Map : X_1 \rightarrow X_2$ such that
+$\Map(a \MonoidOp b) = \Map(a) \MonoidOp' \Map(b)$ and
+$\Map(0_{\Monoid}) = 0_{\Monoid'}$.
 
@@
 Note that $\Strings = \Tuple{\Alphabet;\Concat{};\epsilon}$, is a
-non-commutative monoid, referred to as the free monoid. The string length
+non-commutative monoid, referred to as the free monoid on $\Alphabet$.
+The string length
 function, $\Length{s}$ is an example of a homomorphism between $\Strings$
@@
 
-An automaton~$\Automaton$ with alphabet~$\Alphabet$ is
-$\AutomatonTuple$ where $\Transitions \subseteq \States \times \Alphabet \times
-\States$, $\States$ is its states, $\InitialState$ its
-initial state, and $\AcceptingStates$ is its set of accepting states.  We
-write a transition $\Transition = \Tuple{\State, \Label, \State'} \in
-\Transitions$ as $\Transition = \FromLabelTo{\State}{\Label}{\State'}$.
-Similarly, we use the notation $\FromLabelTo{\State}{}{}$ to refer to the set of
-transitions starting in $\State$, and $\FromLabelTo{}{}{\State}$ to refer to the
-set of transitions coming into $\State$, whenever the automaton is clear from
-the context.
+A \emph{finite-state automaton}~$\Automaton$ with alphabet~$\Alphabet$ is a
+tuple~$\AutomatonTuple$, where $\States$ is the set of states,
+$\InitialState$ the initial state, $\AcceptingStates$ the set of
+accepting states, and
+$\Transitions \subseteq \States \times \Alphabet \times \States$ the
+transition relation.  We write a transition
+$\Transition = \Tuple{\State, \Label, \State'} \in \Transitions$ as
+$\Transition = \FromLabelTo{\State}{\Label}{\State'}$.  Similarly, we
+use the notation $\FromLabelTo{\State}{}{}$ to refer to the set of
+transitions starting in $\State$, and $\FromLabelTo{}{}{\State}$ to
+refer to the set of transitions coming into $\State$, whenever the
+automaton is clear from the context. An accepting run of $\Automaton$
+is a sequence
+$\InitialState = q_0 \xrightarrow{l_1} q_1 \xrightarrow{l_2} \cdots
+\xrightarrow{l_k} q_k \in \AcceptingStates$, and the accepted word is
+$l_1 l_2 \ldots l_k \in \Strings$. The set of words accepted by an
+automaton~$\Automaton$ is denoted by
+$\Language(\Automaton) \subseteq \Strings$.
 
 We will let variables $\Transition, \Transition', \Transition_1, \ldots,
-\Transition_n$ etc describe transitions, $\State, \ldots, \State_n$ states, and
+\Transition_n$, etc., denote transitions, $\State, \ldots, \State_n$ states, and
 $\Automaton, \ldots, \Automaton_n$ automata, and use subscript indexing
-($\Transitions_\Automaton$) to refer to the transitions, states, etc of a given
+($\Transitions_\Automaton$) to refer to the transitions, states, etc., of a given
 automaton.
 
-By a \emph{product} of two automata $\Automaton_1, \Automaton_2$, written
-$\Automaton_1 \times \Automaton_2$, we mean an automaton constructed to run
-$\Automaton_1$ and $\Automaton_2$ in parallel on an input and only accept the
-input if both automata would do so.
-
-We refer to the resulting product states as tuples, $\Tuple{\State, \State'}$,
-which represent the state of the product automaton where $\Automaton_1$ would be
-in $\State$ and $\Automaton_2$ would be in $\State'$. Note that since we use
-ordered tuples the product is technically (but w.l.o.g) not commutative; the
-left-hand-side must come from the left-hand term. The sole purpose of this
-matching is to allow us to speak with precision about the origins of components
-in a product.
+The \emph{product} of two
+automata~$\Automaton_1 = \Tuple{\States^1, \InitialState^1,
+  \AcceptingStates^1, \Transitions^1}$ and
+$\Automaton_2 = \Tuple{\States^2, \InitialState^2,
+  \AcceptingStates^2, \Transitions^2}$ is the automaton
+\begin{align*}
+\Automaton_1 \times \Automaton_2 &=
+\Tuple{\States_1 \times \States_2, \STuple{\InitialState^1, \InitialState^2},
+\AcceptingStates^1 \times \AcceptingStates^2,
+  \Transitions}\\
+  \text{with}\quad \Transitions &= \{\STuple{\STuple{q,q''}, l, \STuple{q',q'''}} \mid \STuple{q, l, q'} \in \Transitions^1, \STuple{q'', l, q'''} \in \Transitions^2\}~.
+\end{align*}
+The product automaton runs $\Automaton_1$ and $\Automaton_2$ in
+parallel on an input and only accepts the input if both automata would
+do so; we have $\Language(\Automaton_1 \times \Automaton_2) =
+\Language(\Automaton_1) \cap \Language(\Automaton_2)$.
 
@@
 from context-free grammars and~\cite{bounds} for bounds on its size).
+The Parikh image is therefore a semi-linear set and Presburger-definable.
 %However,
@@
 needs to be done.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "main"
+%%% End:

### 2023-07-11 example.tex
Mathcal

@@
     \item\label{const:s1-in-c-dd} $s_1 \in \Language(\Regex{c\RegexOr{}dd})$.
-    \item\label{const:s2-in-b} $s_2 \in \Language(B)$, the language accepted by
-    automaton~$B$ of \cref{fig:aut_b} (ignoring
-    the registers~$l_B$).
+    \item\label{const:s2-in-b} $s_2 \in \Language(\mathcal{B})$, the language accepted by
+    automaton~$\mathcal{B}$ of \cref{fig:aut_b} (ignoring
+    the registers~$l_\mathcal{B}$).
     \item\label{const:s1-substring} $s_1 = \Substring(s_2, i, n)$, that is $s_1$ is an
@@
 \cref{const:s1-in-c-dd,const:s2-in-b,const:s1-substring} can be
-translated to the product of Parikh automata $B \times
-A'$. Automaton~$B$ is given in \cref{fig:aut_b}.  Automaton~$A'$
+translated to the product of Parikh automata $\mathcal{B} \times
+\mathcal{A}'$. Automaton~$\mathcal{B}$ is given in \cref{fig:aut_b}.  Automaton~$\mathcal{A}'$
 (\cref{fig:aut_a}) is an automaton defining the pre-image of
@@
 and obtained by applying the construction described in
-\cite{ostrich-plus}. Intuitively, $A'$ describes all
+\cite{ostrich-plus}. Intuitively, $\mathcal{A}'$ describes all
 tuples~$(s_2, i, n)$ such that $\Substring(s_2, i, n)$ is in
-$\Language(\Regex{c\RegexOr{}dd})$. To this end, $A'$ contains
+$\Language(\Regex{c\RegexOr{}dd})$. To this end, $\mathcal{A}'$ contains
 registers~$l_A, i, n$, which count the length of the overall string
@@
 automata, we realise that the path using $\Char{dd}$ in $A'$ is unusable since
-are no $\Char{d}$-labelled transitions in $B$. This means that $s_1 = \Char{c}$,
+are no $\Char{d}$-labelled transitions in~$\mathcal{B}$. This means that $s_1 = \Char{c}$,
 and thus $n = 1$. That means that \cref{const:more-inside-than-before} implies
@@
 image, as described in~\cite{generate-parikh-image}, would start by
-computing the product $A' \times B$, translating it to a Presburger formula with
+computing the product $\mathcal{A}' \times \mathcal{B}$, translating it to a Presburger formula with
 $i, n$, etc., as free variables, and then adding
@@
 
-We associate each transition $\Transition$ of $A'$ and $B$ with a fresh
+We associate each transition $\Transition$ of $\mathcal{A}'$ and $\mathcal{B}$ with a fresh
 variable~$\TransitionVar_\Transition$ ranging over natural numbers (i.e.,
@@
 that the number of incoming transitions is equal to the number of outgoing ones.
-E.g., state $B$ of automaton~$B$ would have the sum
+E.g., state $B$ of automaton~$\mathcal{B}$ would have the sum
 %
@@
 with transitions.
-As an example, we start with $A'$. Substituting back
-solutions to the equations for $A'$ to the automaton, we obtain the
+As an example, we start with $\mathcal{A}'$. Substituting back
+solutions to the equations for $\mathcal{A}'$ to the automaton, we obtain the
 automaton in \cref{fig:a_2}, where the notation~$a / t$ now denotes
@@
 can be specified purely in terms of the free
-variables we care about representing:  the length of the accepted word ($l_A$), the
+variables we care about representing:  the length of the accepted word ($l_\mathcal{A}$), the
 start of the substring ($i$), and the length of the substring ($n$).
@@
   \includegraphics[width=0.6\textwidth]{a_2}
-  \caption{ $A'$ after even more linear algebra, with most transitions expressed
+  \caption{ $\mathcal{A}'$ after even more linear algebra, with most transitions expressed
   in terms of $n$. Intuitively, this version captures the fact that $n$ is given
@@
     \vspace*{5ex}
-    \caption{ $A'$ with its associated transition variables in symbolic
+    \caption{ $\mathcal{A}'$ with its associated transition variables in symbolic
     form.}\label{fig:aut_a_annotated}
@@
     \includegraphics[width=\textwidth]{b_annotated}
-    \caption{$B$ with its associated transition variables in symbolic forms.
+    \caption{$\mathcal{B}$ with its associated transition variables in symbolic forms.
     The large number of implicitly existentially quantified
     variables~$x_i$ on
-    transitions  suggests that this automaton has a more complex structure
+    transitions suggest that this automaton has a more complex structure
     with relation to its (free) target variable representing the string length
-    than $A'$. %The variables used here are fresh, and do not correspond to the
+    than $\mathcal{A}'$. %The variables used here are fresh, and do not correspond to the
     % ones previously used.
@@
 \subsubsection{Case Splitting and connectivity propagation}\label{sec:intuition:split}
-We now have a choice of two paths through automaton $B$; the
+We now have a choice of two paths through automaton $\mathcal{B}$; the
 upper through state $A$ or the lower through state $B$. Since
@@
 we perform a case split by selecting a transition variable that would disconnect
-some strongly connected component of automaton $B$, in this case the transition
+some strongly connected component of automaton $\mathcal{B}$, in this case the transition
 from~$I_B$ to state $B$ guarded by variable~$x_{12}$.
@@
 left with two (small) flat automata, and can compute their product with
-relative ease. By putting off computing the product $B \times A'$ until after
+relative ease. By putting off computing the product $\mathcal{B} \times \mathcal{A}'$ until after
 performing linear reasoning, and using that to prune transitions, we
@@
 approach. Computing the product, we will immediately notice that the
-$\Char{d}$-labelled transition of automaton $A'$ has no correspondence
-in $B$, leading to an empty product. We can close the proof goal and
+$\Char{d}$-labelled transition of automaton $\mathcal{A}'$ has no correspondence
+in $\mathcal{B}$, leading to an empty product. We can close the proof goal and
 backtrack, and will eventually derive that the imposed constraints are

### 2023-07-11 introduction.tex
Mathcal

@@
-Extending automated theorem provers and SMT solvers with support for
-rich string constraints is important for program analysis, in
-particular to detect cross-site scripting vulnerabilities and other
-string manipulation bugs~\cite{DBLP:books/sp/BultanYAA17}.
+Extending automated theorem provers and SMT solvers with support for rich string
+constraints is important for program analysis, particularly to detect cross-site
+scripting vulnerabilities and other string manipulation
+bugs~\cite{DBLP:books/sp/BultanYAA17}.
 %

### 2023-07-11 preliminaries.tex
polishing

@@
 
-We define the image of this map, the \textit{Parikh image}, of some subset of
-the language $\Language \subseteq \Strings$ as:
+We define the image of this map, the \textit{Parikh image}, of some
+language $\Language \subseteq \Strings$ as:
 \[
@@
 1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
-notation $\#l(w)$ to talk about an individual letter $l$ in a word $w$. For
-example, for the Parikh vector above, we would have $\CountOf{a} = 1$.
+notation $\CountOf{l}(w)$ to talk about an individual letter $l$ in a word $w$. For
+example, for the Parikh vector above, we would have $\CountOf{a}(abb) = 1$.
 
@@
 
-Since Parikh images are semilinear, any Parikh image can be written as a set of
-linear equations. The version for context-free grammars presented in
-\cite{generate-parikh-image} can be straightforwardly adapted for use with a
-product of regular languages or Parikh automta. For an intuition, the approach
-consists of first computing the product, then assigning each state and
-transition an existentially quantified non-negative integer variable, then
-describing all paths through the automaton through two sets of equations: one
-that preserves the incoming flow across the variables of a state's incoming
-transitions is equal to the flow across its outgoing transitions, and one set of
-constraints that enforces coherence in the presence of loops by ordering states
-by distance in a spanning tree rooted in the initial state.
+It is known that the Parikh image of any context-free language can be
+described by a linear-size existential Presburger
+formula~\cite{generate-parikh-image}. This representation can be
+straightforwardly adapted for use with a product of regular
+languages. For an intuition, the approach consists of first computing
+the product, then assigning each state and transition an existentially
+quantified non-negative integer variable, and then describing all
+paths through the automaton through two sets of constraints: flow
+equations relating the inflow and outflow of each automaton state, and
+constraints that enforce connectedness by ordering states by distance
+in a spanning tree rooted in the initial state.
 

### 2023-07-11 generalised.tex
comments, bit of shortening
more

@@
 approach is that it is more efficient to directly compute a
-projection on the Parikh image~$\Map(\Language)$, than to first compute the
+projection~$\Map(\Language)$ on the Parikh image, than to first compute the
 standard image~$\ParikhMap(\Language)$ followed by projection to
@@
 
-% @Philipp: I am quite unsure about which intersection we are looking for here,
+% AS: I am quite unsure about which intersection we are looking for here,
 % please verify I'm not wrong!

### 2023-07-11 evaluation.tex
Grammarly

@@
 \Calculus{} generally outperforms \Nuxmv{} on determining unsatisfiability, as
-does baseline while being similar at satisfiable instances. Both \Calculus{} and
-baseline outperforms baseline on satisfiable instances. On satisfiable
-instances, \Nuxmv and \Calculus{} has similar performance.
+does baseline, while being similar at satisfiable instances. Both \Calculus{} and
+\Nuxmv outperforms baseline on satisfiable instances. On satisfiable
+instances, \Nuxmv and \Calculus{} have similar performance.
 
@@
 deviate from what is actually benchmarked in section \cref{sec:experiments}, or
-if the methods used for benchmarking would be unsound. To increase the
-probability that our results are representative both in the sense of
-representing  on expected inputs and in the statistical sense for a given run of
-\Catra{} despite the use of randomness in our implementation we execute many
-experiments. Moreover, to address the issue of correctness we have validated all
-reported solutions made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{}
-is indeed sound. A previous version contained a race condition triggered by
-random restarts during product materialisation which would in about
-\numprint{0.7} of instances lead to non-deterministically unsound behaviour. We
-have observed no instances of this bug or any other soundness issue since fixing
-that bug, and have in addition to running all \NrBenchmarks{} we have
-re-executed the same instances multiple times for long durations to capture
-other soundness issues without any issue.
+if the methods used for benchmarking were unsound. To increase the probability
+that our results are representative both in the sense of representing expected
+inputs and, in the statistical sense, for a given run of \Catra{} despite the
+use of randomness in our implementation, we execute many experiments. Moreover,
+to address the issue of correctness we have validated all reported solutions
+made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{} is indeed sound. A
+previous version contained a race condition triggered by random restarts during
+product materialisation which would in about \numprint{0.7} of instances lead to
+non-deterministically unsound behaviour. We have observed no instances of this
+bug or any other soundness issue since fixing that bug, and in addition to
+running all \NrBenchmarks{}, we have re-executed the same instances multiple
+times for long durations to capture other soundness issues without any issues.
 
 Another threat to validity would be if the competition (baseline and \Nuxmv)
-would be disadvantaged in our comparison. For \Nuxmv{} we use the default
-configuration which we believe should be performant (or it should not be the
+would be disadvantaged in our comparison. For \Nuxmv{}, we use the default
+configuration, which we believe should be performant (or it should not be the
 default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
@@
 solution we use is close to the one found in the \Ostrich{} string solver and
-that it therefore should be realistic.
+that it, therefore, should be realistic.
 
 The greatest threat to validity remaining is our choice of implementation
-platform and automata library. We There are some signs that product computation
+platform and automata library. There are some signs that product computation
 is inefficient, notably the good performance of low-threshold automata

### 2023-07-11 generalised.tex
more

@@
 \Substring(s_2, i, n)$, that is $s_1$ is an $n$-length substring of $s_2$
-starting at offset $i$ (in some variants , $n$ is $i + |n|$, i.e. the ending
-index). That constraint belongs to an expressive fragment of string logic that
+starting at offset $i$.
+% (in some variants , $n$ is $i + |n|$, i.e. the ending index).
+That constraint belongs to an expressive fragment of string logic that
 cannot be decided by most state-of-the-art string solvers.
 
-In \cref{sec:intuition} we hand-wavily modelled \cref{const:s1-substring} (and
-  the other constraints) using \emph{Parikh
-  automata}~\cite{parikh-automata,expressiveness}. A Parikh automaton is a
-  standard NFA extended such that transitions are additionally labelled both
-  with offset vectors defining the increments of a finite number of counters.
-  This means that Parikh automata recognise words over an extended
-  alphabet~$\Alphabet \times D$, where $D \subseteq \Naturals^d$ is a finite set
-  of increment vectors (notation as in~\cite{expressiveness}) and $\Alphabet$ is
-  the alphabet of the original NFA.
+In \cref{sec:intuition} we modelled \cref{const:s1-substring} (and the
+other constraints) using \emph{Parikh
+  automata}~\cite{parikh-automata,expressiveness}. A Parikh automaton
+is a finite-state automaton extended such that transitions are
+additionally labelled with offset vectors defining the increments
+of a finite number of counters.  This means that Parikh automata
+recognise words over an extended alphabet~$\Alphabet \times D$, where
+$D \subseteq \Naturals^d$ is a finite set of increment vectors
+(notation as in~\cite{expressiveness}) and $\Alphabet$ is the alphabet
+of the original automaton.
 
@@
 operation, and check whether this pre-image is consistent with the constraint~$s_2 \in \Language(B)$, corresponding to the full product seen in
-\cref{ex:string-constraints}. Because the substring operation depends on the
-values of the integer variables~$i, n$, a Parikh automaton of dimension~$2$,
-shown in \cref{fig:aut_a}, models the pre-image. Intuitively, this construction
-accumulates any symbol (for the prefix of the substring operation), incrementing
-$i$ to mark the start of the substring, followed by the automaton representing
+\cref{ex:string-constraints}. Because $\Substring$ depends on the
+values of the integer variables~$i, n$, a Parikh automaton,
+shown in \cref{fig:aut_a}, models the pre-image. The Parikh automaton
+has dimension~$3$, as it includes also the length register~$l_{A}$, besides
+variables $i, n$.  Intuitively, this construction
+accumulates any prefix symbol, incrementing
+$i$ to mark the start of the substring; followed by the automaton representing
 the substring itself, modified to add a counter to increment $n$ at each
-transition, followed by a state that performs no increments and only accepts any
+transition; followed by a state that accepts any
 suffix after the matched substring.
@@
   \begin{bmatrix}
-    i \\ n
+    l_{A} \\ i \\ n
   \end{bmatrix}~~
-  \wedge 0 \leq i \leq i + n \leq |p|
+  \wedge 0 \leq i \leq i + n \leq l_{A}
 \end{equation}
@@
 develop the lazy algorithm proposed in this paper.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "main"
+%%% End:

### 2023-07-11 example.tex

@@
 A decision procedure for a fragment of string constraints with integers,
-covering our constraints, is defined in
+covering our constraints is defined in
 \cite{ostrich-plus}.
@@
 $\Language(\Regex{c\RegexOr{}dd})$. To this end, $\mathcal{A}'$ contains
-registers~$l_A, i, n$, which count the length of the overall string
+registers~$l_A, i, n$, which counts the length of the overall string
 read, the length of the prefix before the extracted substring, and the
@@
 \begin{inparaenum}[(i)]
-\item Similarly as in \cite{generate-parikh-image}, we first describe
+\item Similarly, as in \cite{generate-parikh-image}, we first describe
   each automaton as a flow network, counting how often each transition
@@
   over-approximation of the possible accepting runs of an automaton.
-\item We use a linear integer solver to simplify the flow constraints,
+\item We use a linear integer solver to simplify the flow constraints
   and prune away paths through the automata that are not feasible.
 \item We use a tailor-made propagation algorithm to identify disconnected
-  transitions of the automata that can never be taken, and lazily add
+  transitions of the automata that can never be taken and lazily add
   the corresponding constraints.
 \item When propagation cannot infer further constraints, we use case
-  splitting to sub-divide the problem into smaller parts.
+  splitting to subdivide the problem into smaller parts.
 \item Once the constituent automata of a product have become
@@
 
-We then use linear arithmetic reasoning on register
+We then use linear arithmetic reasoning on the register
 equations~\eqref{eq:counter-sums} and flow
@@
 %
-Substituting the derived value of $n$, and applying similar
+Substituting the derived value of $n$ and applying similar
 reasoning to automaton~$B$, we arrive at the simplified

### 2023-07-11 example.tex
Fix first example

@@
     automaton~$\mathcal{B}$ of \cref{fig:aut_b} (ignoring
-    the registers~$l_\mathcal{B}$).
+    the registers~$l_B$).
     \item\label{const:s1-substring} $s_1 = \Substring(s_2, i, n)$, that is $s_1$ is an
@@
 Intuitively, \cref{ex:string-constraints} is unsatisfiable. By inspecting the
-automata, we realise that the path using $\Char{dd}$ in $A'$ is unusable since
+automata, we realise that the path using $\Char{dd}$ in $\mathcal{A}'$ is unusable since
 are no $\Char{d}$-labelled transitions in~$\mathcal{B}$. This means that $s_1 = \Char{c}$,
@@
 where we let
-$x_{\Tuple{q, l, c'}}$ refer to the integer variable associated with a
+$x_{\Tuple{q, l, q'}}$ refer to the integer variable associated with a
 transition from state $q$ to state $q'$ with label $l$. The initial state
@@
 states have an additional common outflow of $1$. Note that self-loops
-cancels themselves out. This means that for automata where a loop can become
-unreachable, additional constraints would be required to ensure reachability.
-However, being lazy, we put off that work until later.
+cancels themselves out. Therefore, when a loop can become
+unreachable, additional constraints are required to ensure consistency.
 
@@
 can be specified purely in terms of the free
-variables we care about representing:  the length of the accepted word ($l_\mathcal{A}$), the
+variables we care about representing:  the length of the accepted word ($l_A$), the
 start of the substring ($i$), and the length of the substring ($n$).
@@
 Substituting the derived value of $n$ and applying similar
-reasoning to automaton~$B$, we arrive at the simplified
+reasoning to automaton~$\mathcal{B}$, we arrive at the simplified
 automata in \cref{fig:propagated}.
@@
 
-In the case~$x_{12} = 0$ we can conclude
-that the state~$B$ is now unreachable, which means that also
-transitions outgoing from state $B$ are now unusable. Propagation can
+In the case~$x_{12} = 0$, we can conclude
+that the state~$B$ is now unreachable, which means that its outgoing transitions are now unusable. Propagation can
 therefore infer the additional equations $x_9 = 0$ and $x_{11} = 1$.
@@
 in $\mathcal{B}$, leading to an empty product. We can close the proof goal and
-backtrack, and will eventually derive that the imposed constraints are
-unsatisfiable.
+backtrack, and we will eventually derive that the imposed constraints are
+unsatisfiable by repeating the same process on the other branch.
 

### 2023-07-11 calculus.tex
Mathcal

@@
 of~$\Filter$, boiling them down to choices of branches, which depend on one
-single variable, and loop transitions. This logic corresponds to the placement
+single variable and loop transitions. This logic corresponds to the placement
 of counters for optimally edge-profiling the CFG of a program, making up a
@@
 on linear inequalities. This rule is not necessary for correctness or
-completeness, but shortens the examples considerably.
+completeness but shortens the examples considerably.
 
 \subsection{An Example}\label{sec:single:example}
-Here we will return to $A'$ from \cref{sec:intuition} and performs the steps of
-\cref{sec:intuition:algebra,sec:intuition:prop,sec:intuition:split} with the
+Here we will return to $\mathcal{A}'$ from \cref{sec:intuition} and performs the steps of
+\cref{sec:intuition:algebra,sec:intuition:split} with the
 formal calculus we just established, but exclude \cref{const:s2-in-b} and
-therefore the entire automaton $B$ since we introduce support for products of
+therefore the entire automaton $\mathcal{B}$ since we introduce support for products of
 automata in \cref{sec:multiple}. Note that this example then becomes
@@
 
-Starting with~$A'$, $\Map$ extracts the increments of a transition (see
-\cref{fig:a_1}), $h(\FromLabelTo{q_1}{c / i}{q_2}) = \Vec{v}$ for the increment
+Starting with~$\mathcal{A}'$, $\Map$ extracts the increments of a transition, $h(\FromLabelTo{q_1}{c / i}{q_2}) = \Vec{v}$ for the increment
 vector $\Vec{v}$, concretely $h(\FromLabelTo{I_A}{\Sigma / \left[ l_{A+}, i_+
@@
 
-  Then, following \cref{sec:intuition:subsume}, we remove the $\Connected{}$
-   predicate using \Subsume{} since no transition can can become disconnected
-   (and in fact never could for $A'$, since all looping transitions are attached
-   to the initial and single accepting state and therefore always must be
-   reachable from any accepting path). After this, the rest of the reasoning can
-   be continued without any part of our calculus, since all values of the
-   existentially quantified transition variables
-   $\TransitionVar_1,\ldots,\TransitionVar_5$ compatible with
+  Then, we remove the $\Connected{}$ predicate using \Subsume{} since no
+   transition can become disconnected (and never could for $\mathcal{A}'$,
+   since all looping transitions are attached to the initial and single
+   accepting state and therefore always must be reachable from any accepting
+   path). After this, the rest of the reasoning can be continued without any
+   part of \Calculus{} since all values of the existentially quantified
+   transition variables $\TransitionVar_1,\ldots,\TransitionVar_5$ compatible
+   with
    \cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring,const:something-before-and-after}
@@
     \end{aligned}}
-  \infer1[\Subsume: \cref{sec:intuition:subsume}]{
+  \infer1[\Subsume]{
     \begin{aligned}
-    \Connected(A', \Filter) \land 
+    \Connected(\mathcal{A}', \Filter) \land 
     \TransitionVar_1 = 1 \land 
@@
     \begin{aligned}
-    \Connected(A', \Filter) \land
+    \Connected(\mathcal{A}', \Filter) \land
     i = \TransitionVar_5 \land 
@@
     \begin{aligned}
-    \Connected(A', \Filter)  
+    \Connected(\mathcal{A}', \Filter)  
     \land i = \TransitionVar_5 
@@
     \land n = \TransitionVar_4 + \TransitionVar_1\\
-    \land \Connected(A', \Filter) 
-    \land \FlowEq(A', \Filter)
+    \land \Connected(\mathcal{A}', \Filter) 
+    \land \FlowEq(\mathcal{A}', \Filter)
     \end{aligned}
@@
     \left[l_A, i, n\right] =
-    \sum\limits_{\Transition \in \Transitions_{A'} } \Map(\Transition) \cdot \Filter(\Transition)
+    \sum\limits_{\Transition \in \Transitions_{\mathcal{A}'} } \Map(\Transition) \cdot \Filter(\Transition)
   }
   \infer1[\Expand{}]{
-    \Image{}_{A', \Map}(\Filter, \left[l_A, i, n\right]) 
+    \Image{}_{\mathcal{A}', \Map}(\Filter, \left[l_A, i, n\right]) 
     \land n \geq i 
@@
 from \cref{ex:string-constraints}, corresponding to handling the Parikh
-automaton $A'$ of \cref{fig:aut_a}. Note the liberal use of removal of equal
+automaton $\mathcal{A}'$ of \cref{fig:aut_a}. Note the liberal use of removal of equal
 variables (e.g. if $x = y$ appears as a statement, we substitute $y$ for $x$
@@
 
-Our correctness proof of \Calculus{} consists of two main parts: we
-first show that the construction of a proof always terminates, and
-then that each of the proof rules in \cref{tbl:rules:single} is
-an equivalence transformation, i.e., does not change the set of
-satisfying assignments of a formula. In combination, those two results
-immediately imply that \Calculus{} gives rise to a decision procedure.
+Our correctness proof of \Calculus{} consists of two main parts: first, we show
+that the construction of a proof always terminates, and then that each of the
+proof rules in \cref{tbl:rules:single} is an equivalence transformation, i.e.,
+does not change the set of satisfying assignments of a formula. In combination,
+those two results immediately imply that \Calculus{} gives rise to a decision
+procedure.
 

### 2023-07-11 calculus.tex
Missing mathcal

@@
         & \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) = \TransitionVar_2   
-        & \FlowEq(A', \Filter)  = \begin{aligned}
+        & \FlowEq(\mathcal{A}', \Filter)  = \begin{aligned}
                        1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_3 + \TransitionVar_1\\

### 2023-07-11 multiple.tex
Math!cal!

@@
 so in this case, $\Map(\Transition) = \left[l_{B+}\right]$ for every
-transition~$\Transition$ since $B$ only counts length. The reader is
+transition~$\Transition$ since $\mathcal{B}$ only counts length. The reader is
 strongly advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
@@
         & \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}
-        \FlowEq(B, \Filter) &=  \begin{aligned}
+        \FlowEq(\mathcal{B}, \Filter) &=  \begin{aligned}
             1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
@@
 \cref{fig:derivation:single}, since the steps to propagate
-\cref{ex:string-constraints} across the automaton $A'$ still apply for the
-product. Note that this fixes the values of all transitions but one of $A'$ as
+\cref{ex:string-constraints} across the automaton $\mathcal{A}'$ still apply for the
+product. Note that this fixes the values of all transitions but one of $\mathcal{A}'$ as
 constants, including the free counter variables $i, n$, whose new constant
@@
 Once we have performed reasoning by equality to eliminate transition variables,
-we obtain the version of $B$ shown in \cref{fig:aut_b_annotated} and represented
+we obtain the version of $\mathcal{B}$ shown in \cref{fig:aut_b_annotated} and represented
 by the clauses in the final node in the tree. This corresponds to the end of
@@
       \land \Connected(B, \Filter)\\
-      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
+      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})
     \end{array}
@@
       \land \TransitionVar_{10} = 0 
-      \land \Connected(B, \Filter) \\
-      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
+      \land \Connected(\mathcal{B}, \Filter) \\
+      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
     \end{array}
@@
       \TransitionVar_{10} = 1 - \TransitionVar_{11} \land\\
-      \Connected(B, \Filter) \land\\
-      \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
+      \Connected(\mathcal{B}, \Filter) \land\\
+      \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
     \end{array}
@@
     \land \TransitionVar_{10} = 1 - \TransitionVar_{11} \\
-    \land \Connected(B, \Filter)\\
-    \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
+    \land \Connected(\mathcal{B}, \Filter)\\
+    \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
   \end{array}
@@
       \land \TransitionVar_{12} = \TransitionVar_8 + \TransitionVar_{10}
-      \land \Connected(B, \Filter) 
-      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
+      \land \Connected(\mathcal{B}, \Filter) 
+      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
     \end{array}
@@
       \land \TransitionVar_2 = l_A - 3
-      \land \FlowEq(B, \Filter) \\
-      \land \Connected(B, \Filter) 
-      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
+      \land \FlowEq(\mathcal{B}, \Filter) \\
+      \land \Connected(\mathcal{B}, \Filter) 
+      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
     \end{array}
@@
       \land n = \TransitionVar_4 + \TransitionVar_1
-      \land \FlowEq(B, \Filter)
-      \land \Connected(A', \Filter)\\
-      \land \Connected(B, \Filter)
-      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) \\
+      \land \FlowEq(\mathcal{B}, \Filter)
+      \land \Connected(\mathcal{A}', \Filter)\\
+      \land \Connected(\mathcal{B}, \Filter)
+      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) \\
       \land l_A = l_B
@@
   \infer1[\ExpandM]{
-    \Image{}_{\Tuple{A', B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})
+    \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})
     \land l_A = l_B % Implied constraint

### 2023-07-11 calculus.tex
moved basic automata definitions to section 3

--- 2023-07-10 calculus.tex
@@
 We start by defining our calculus, \Calculus{}, for one automaton, and only
-extend it to products of automata in \cref{sec:multiple}. We assume basic
-familiarity with first-order logic, Presburger arithmetic, and the classical
-sequent calculus. For reference, see e.g. \cite{Fitting96a}.
-
-First, assume an NFA $\Automaton = \AutomatonTuple$ with $\NrTransitions$
+extend it to products of automata in \cref{sec:multiple}.
+%
+Assume an automaton $\Automaton = \AutomatonTuple$ with $\NrTransitions$
 transitions $\Transitions =
-\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. For convenience, we then
-introduce the following supporting notations:
+\Set{\EllipsisSequence{\Transition}{\NrTransitions}}$. We use
+the notations introduced in \cref{sec:languages}. For convenience, we
+introduce the following additional notions:
 
 \begin{definition}
-  A \textit{path} $\Path = \PathEnumeration$ of an automaton~$\Automaton$ with
-  transitions $\Transitions$ represents a path through $\Automaton$ using
-  transitions in $\Transitions$ (i.e. $\Transitions(\State_k, \Label_{k+1},
-  \State_{k+1})$ holds), passing zero or more labels $\Label_1, \ldots,
-  \Label_n$. The path must begin in the initial state, i.e.~$\State_0 =
-  \InitialState$. The end state, $\State_n$, is not necessarily accepting. If it
-  is, we say the path is also \textit{accepting}.
+  A set~$C$ of transitions of an automaton~$\Automaton$ \emph{dominates}
+  a  transition~$\Transition$, written
+  $\Dominates(C, \Automaton, \Transition)$, if
+  every accepting path $p$ of $\Automaton$
+  with $\Transition \in p$ contains at least one transition from
+  $C$. Notably, $\Dominates(\emptyset, \Automaton, \Transition)$ for
+  every unreachable transition $\Transition$, and
+  $\Dominates(\Set{\Transition}, \Automaton, \Transition)$ for every
+  transition~$\Transition$.
 \end{definition}
@@
 \begin{definition}
-  Moreover, we talk about the \textit{set of paths} of an automaton,
-  $\Paths(\Automaton)$, a possibly infinite (if $\Automaton$ has loops) set of
-  valid paths through $\Automaton$. Additionally, we use the
-  notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
-  state~$\State$.
-\end{definition}
-
-\begin{definition}
-  For a path $\Path = \Tuple{\State_0 \Label_1 \State_1 \RepeatSum{,}
-  \State_n}$, its \textit{word} $\WordOf(\Path) = \Label_1 \RepeatSum{\Concat}
-  \Label_n$ is the word read out on its labels.
-\end{definition}
-
-\begin{definition}
-  The \textit{states} of a path $\Path = \Tuple{\State_0 \Label_1 \State_1
-  \RepeatSum{,} \State_n}$, $\StatesOf(\Path) = \Set{\State_0\RepeatSum{,}
-  \State_n}$ are the states visited along $\Path$. Note that $\InitialState \in
-  \StatesOf(\Path)$ for every path since all paths start in the initial state.
-\end{definition}
-
-\begin{definition}
-  A \textit{dominating set of transitions}, $C$ of an automaton~$\Automaton$ of
-  a transition $\Transition$, written $\Dominates(C, \Automaton, \Transition)$,
-  is a set of transitions from $\Automaton$ such that every accepting path $p
-  \in \Automaton$ where $\Transition \in p$ contains at least one transition
-  from $C$. Notably, $\Dominates(\emptyset, \Automaton, \Transition')$ for every
-  dead transition $\Transition'$ and $\Dominates(Set{\Transition''}, \Automaton,
-  \Transition)$.
-\end{definition}
-
-\begin{definition}
- The \textit{transition count}, $\TransitionCount(\Transition, \Path)$ is the
- number of times a transition $\Transition =
- \FromLabelTo{\State_1}{\Label}{\State_2} \in \Transitions$ appears in a path
- $p$.
+  The \textit{transition count},
+  $\TransitionCount(\Transition, \Path)$ is the number of times a
+  transition
+  $\Transition = \FromLabelTo{\State}{\Label}{\State'} \in
+  \Transitions$ appears on a path $p$. A \emph{transition selection
+    function} is a function~$\Filter:\: \Transitions \to \Naturals$
+  labelling every transition~$\Transition \in \Transitions$ with a
+  non-negative number.
 \end{definition}
@@
 \begin{definition}\label{def:single-image}
-  The Parikh predicate, $\SinglePredicateInstance$, for some automaton
-  $\Automaton = \AutomatonTuple$, modulo some map $\Map : \Alphabet^* \to \Monoid$ to a
-  commutative monoid $\Monoid$ as described in \cref{sec:generalised}
-  and with a transition selection function
-  $\Filter:\: \Transitions \to \Naturals$ holds when
-  $\MonoidElement \in \Monoid$ is an element of the Parikh image of
-  $\Language(\Automaton)$
-  modulo $\Map$, or more formally when there is an accepting path
+  The Parikh predicate, $\SinglePredicateInstance$ holds for some
+  automaton $\Automaton = \AutomatonTuple$, some map
+  $\Map : \Alphabet^* \to \Monoid$ to a commutative monoid $\Monoid$,
+  some transition selection function
+  $\Filter:\: \Transitions \to \Naturals$, and some monoid
+  element~$\MonoidElement \in \Monoid$ if $\MonoidElement$ is an
+  element of the Parikh image of $\Language(\Automaton)$ modulo
+  $\Map$, or more formally when there is an accepting path
   $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
   that $\Filter(\Transition) = \TransitionCount(\Transition, \Path)$
-  for all $t \in \Transitions$, and $\MonoidElement = \Map(\WordOf(\Path))$.
+  for all $t \in \Transitions$, and
+  $\MonoidElement = \Map(\WordOf(\Path))$.
 \end{definition}
 
-\begin{definition}\label{def:connected} $\Connected(\Automaton, \Filter)$ for
-  some automaton $\Automaton = \AutomatonTuple$ holds when for every
+\begin{definition}\label{def:connected} $\Connected(\Automaton, \Filter)$
+  holds for some automaton $\Automaton = \AutomatonTuple$ and
+  transition selection function
+  $\Filter:\: \Transitions \to \Naturals$
+  if for every
   $\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions$,
-  $\Filter(\Transition) > 0 \implies \exists \Path \in \Paths(\Automaton)$
-  $\Filter(\Transition_i) > 0$ for $\Transition_i \in p$, and $\State \in
-  \StatesOf(p)$, or in words that there exists some $\Filter$-selected valid
-  path that reaches $\Transition$'s starting state, $\State$. It represents the
+  if $\Filter(\Transition) > 0$ then there is some $\Filter$-selected
+  path that reaches $\Transition$'s starting state~$\State$. More formally,
+  if $\Filter(\Transition) > 0$ then there is some
+  path~$\Path \in \Paths(\Automaton)$ with
+  $\Filter(\Transition') > 0$ for every $\Transition' \in p$, and $\State \in
+  \StatesOf(p)$. Predicate~$\Connected(\Automaton, \Filter)$ represents the
   condition that $\Automaton$ is connected with respect to the selection
-  function $\Filter$ for every transition. It is redundant to $\Image$ by
-  design.
+  function $\Filter$ for every transition, and is implied by
+  $\SinglePredicateInstance$.
 \end{definition}
@@
 \end{proof}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "main"
+%%% End:

### 2023-07-11 preliminaries.tex
more
moved basic automata definitions to section 3

@@
+We first survey some of the required background on finite-state
+automata and the Parikh image. In addition, we assume basic familiarity
+with first-order logic, Presburger
+arithmetic, and the classical sequent calculus; for reference, see
+e.g. \cite{Fitting96a}.
+
 \subsection{Monoids}
@@
 \subsection{Languages, Finite-state Automata and their Products}
+\label{sec:languages}
 
@@
 refer to the set of transitions coming into $\State$, whenever the
-automaton is clear from the context. An accepting run of $\Automaton$
-is a sequence
-$\InitialState = q_0 \xrightarrow{l_1} q_1 \xrightarrow{l_2} \cdots
-\xrightarrow{l_k} q_k \in \AcceptingStates$, and the accepted word is
-$l_1 l_2 \ldots l_k \in \Strings$. The set of words accepted by an
-automaton~$\Automaton$ is denoted by
-$\Language(\Automaton) \subseteq \Strings$.
+automaton is clear from the context.
 
@@
 automaton.
+
+A \textit{path} $\Path = \PathEnumeration$ of an
+automaton~$\Automaton$ is a sequence of
+states~$\State_0, \ldots, \State_n$ interleaved with
+letters~$\Label_1, \ldots, \Label_n$ such that
+$\State_0 = \InitialState$ and
+$\FromLabelTo{\State_{i-1}}{\Label_{i}}{\State_{i}}$ for
+$i \in \{1, \ldots, n\}$.  The path is \emph{accepting} if the end
+state is accepting, $\State_n \in \AcceptingStates$. The \textit{set
+  of paths} of $\Automaton$ is denoted by
+$\Paths(\Automaton)$. Additionally, we use the
+notation~$\Paths(\Automaton, \State)$ to mean all paths ending in
+state~$\State$. We write $\Transition \in \Path$ to express that
+transition~$\Transition$ occurs on path~$\Path$. The \textit{states} of a path
+$\Path = \PathEnumeration$, denoted
+$\StatesOf(\Path) = \Set{\State_0\RepeatSum{,} \State_n}$ are the
+states visited along $\Path$. Note that
+$\InitialState \in \StatesOf(\Path)$ for every path since all paths
+start in the initial state.  The \textit{word} of a path
+$\Path = \PathEnumeration$ is the word
+$\WordOf(\Path) = \Label_1 \RepeatSum{\Concat}
+ \Label_n \in \Strings$ formed by the labels on the path. Finally, the
+ set of words accepted by an automaton~$\Automaton$, denoted by
+ $\Language(\Automaton) \subseteq \Strings$, is the set of words of
+ accepting paths.
 

### 2023-07-11 calculus.tex
bw-reach

@@
   if $\Filter(\Transition) > 0$ then there is some $\Filter$-selected
-  path that reaches $\Transition$'s starting state~$\State$. More formally,
+  accepting path that visits $\Transition$'s starting state~$\State$. More formally,
   if $\Filter(\Transition) > 0$ then there is some
@@
   $\Filter(\Transition') > 0$ for every $\Transition' \in p$, and $\State \in
-  \StatesOf(p)$. Predicate~$\Connected(\Automaton, \Filter)$ represents the
-  condition that $\Automaton$ is connected with respect to the selection
+  \StatesOf(p)$ such that $p$ ends in an accepting state $\State' \in \AcceptingStates$. The predicate~$\Connected(\Automaton, \Filter)$ represents the
+  condition that $\Automaton$ is connected under the selection
   function $\Filter$ for every transition, and is implied by

### 2023-07-11 calculus.tex
Grammar grammar

@@
   $\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions$,
-  if $\Filter(\Transition) > 0$ then there is some $\Filter$-selected
+  if $\Filter(\Transition) > 0$ there is some $\Filter$-selected
   accepting path that visits $\Transition$'s starting state~$\State$. More formally,

### 2023-07-11 calculus.tex
More grammar

@@
   element of the Parikh image of $\Language(\Automaton)$ modulo
-  $\Map$, or more formally when there is an accepting path
+  $\Map$, or more formally, when there is an accepting path
   $\Path = \PathEnumeration \in \Accepting{\Paths(\Automaton)}$ such
@@
 We present the rules of \Calculus{} for one automaton in
-\cref{tbl:rules:single}. The rules operate on sets of formulas, and
+\cref{tbl:rules:single}. The rules operate on sets of formulas and
 can be interpreted as rules of a one-sided sequent calculus, in
@@
 
-Each rule should be read and applied bottom-up, and relates
+Each rule should be read and applied bottom-up and relates
 premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
-conclusion~$\SomeClause$. When constructing a proof, we start from some
+conclusion~$\SomeClause$. When constructing a proof, we start with some
 root~$\SomeClause$, and then apply proof rules to the goals of the proof until a
@@
 A proof in which all proof goals could be closed shows that the formulas
-in the root~$\SomeClause$ of the proof are inconsistent, and do not
-have any solutions. A proof goal that could not be closed, but to which no
-more rules are applicable, gives rise to  a solution of the formulas in the
+in the root~$\SomeClause$ of the proof are inconsistent and do not
+have any solutions. A proof goal that could not be closed but to which no
+more rules are applicable gives rise to a solution of the formulas in the
 root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic allowing a solution to be computed using
+in Presburger arithmetic, allowing a solution to be computed using
 standard algorithms~\cite{Fitting96a}.
@@
 predicates $\Image$ and $\Connected$ only occur positively. Since our rules
-operate by adding and matching linear (in)equalities in a proof goal, we use the
-shorthand of listing the matched inequalities as antecedents e.g., in
+operate by adding and matching linear (in)equalities in a proof goal; we use the
+shorthand of listing the matched inequalities as antecedents, e.g., in
 \Propagate{}.
 
-The filtering function~$\Filter$ is evaluated symbolically, and can in practice
+The filtering function~$\Filter$ is evaluated symbolically and can, in practice,
 be read as a function from transitions to $\Naturals$-valued terms
@@
 transitions of an automaton~$\Automaton$. Additionally, for an
-automaton~$\Automaton = \AutomatonTuple$ we allow mapping the selection function
+automaton~$\Automaton = \AutomatonTuple$, we allow mapping the selection function
 like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
@@
 \begin{proof}
-  The rule~\Expand{} can only be applied finitely often, since each
+  The rule~\Expand{} can only be applied finitely often since each
   application removes one $\Image$ predicate, and none of the rules
   introduce new instances of the predicate. The rule~\Subsume{} can
-  only be applied finitely often, since it strictly decreases the
+  only be applied finitely often since it strictly decreases the
   combined number of $\Image$ and $\Connected$ predicates in sets of
@@
   \Expand{} unfolds the definition of the $\Image$ predicate. To show
-  that the rule is solution-preserving we prove the equivalence of the
+  that the rule is solution-preserving, we prove the equivalence of the
   upper and lower sets of formulas:

### 2023-07-11 calculus.tex
typesetting

--- 2023-07-11 calculus.tex
@@
 
-A proof in which all proof goals could be closed shows that the formulas
+A proof in which all proof goals are closed witnesses that the formulas
 in the root~$\SomeClause$ of the proof are inconsistent, and do not
-have any solutions. A proof goal that could not be closed, but to which no
+have any solutions. A proof goal that cannot be closed, but to which no
 more rules are applicable, gives rise to  a solution of the formulas in the
@@
 in Presburger arithmetic allowing a solution to be computed using
-standard algorithms~\cite{Fitting96a}.
+standard algorithms~\cite{DBLP:books/daglib/0022394}.
 
@@
   0}{\Connected(\Automaton, \Filter), \Set{\Filter(\Transition') = 0 \SuchThat
-  \Transition' \in C}, \SomeInequalities, \SomeClause} & if $\Dominates(C,
-  \Automaton, \Transition) t \in \Transitions_\Automaton, $ \\[4ex]
+  \Transition' \in C}, \SomeInequalities, \SomeClause} & if
+                                                         $t \in \Transitions_\Automaton$ and\newline $\Dominates(C,
+  \Automaton, \Transition)$ \\[4ex]
 
@@
   \inferrule{\SomeInequalities,\SomeClause}{\Connected(\Automaton, \Filter), \SomeInequalities, \SomeClause} &
-  \Split{} and \Propagate{} cannot be applied \\
+  \Split{} and \Propagate{}\newline cannot be applied \\
   \bottomrule
@@
 \begin{equation}\label{eq:single-filter}
-    \begin{aligned}
-      & \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A}) = \TransitionVar_1
-        & \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} ) = \TransitionVar_2   
-        & \FlowEq(\mathcal{A}', \Filter)  = \begin{aligned}
+  \begin{array}{rl@{\qquad}rl@{\qquad}rl}
+    \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A})
+    &= \TransitionVar_1
+    &
+    \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} )
+    &= \TransitionVar_2   
+    &
+    \Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) 
+    &=  \TransitionVar_3 
+    \\
+    \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) 
+    &= \TransitionVar_4
+    &
+    \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) 
+    &= \TransitionVar_5  \\[3ex]
+%
+  \multicolumn{6}{c}{
+    \FlowEq(\mathcal{A}', \Filter)  = \begin{aligned}
                        1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_3 + \TransitionVar_1\\
@@
                     \TransitionVar_3 + \TransitionVar_1 + \cancel{\TransitionVar_2} &\geq \cancel{\TransitionVar_2}  \\
-                    \end{aligned} \\
-      &\Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) =  \TransitionVar_3 
-          & \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) = \TransitionVar_4
-          & \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) = \TransitionVar_5  
-    \end{aligned}
+                    \end{aligned}}
+  \end{array}
 \end{equation}

### 2023-07-11 calculus.tex

--- 2023-07-11 calculus.tex
@@
 
-A proof in which all proof goals could be closed shows that the formulas
-in the root~$\SomeClause$ of the proof are inconsistent and do not
-have any solutions. A proof goal that could not be closed but to which no
-more rules are applicable gives rise to a solution of the formulas in the
+A proof in which all proof goals are closed witnesses that the formulas
+in the root~$\SomeClause$ of the proof are inconsistent, and do not
+have any solutions. A proof goal that cannot be closed, but to which no
+more rules are applicable, gives rise to  a solution of the formulas in the
 root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic, allowing a solution to be computed using
-standard algorithms~\cite{Fitting96a}.
+in Presburger arithmetic allowing a solution to be computed using
+standard algorithms~\cite{DBLP:books/daglib/0022394}.
 

### 2023-07-11 calculus.tex
restructuring

@@
 introduce the following additional notions:
-
-\begin{definition}
-  A set~$C$ of transitions of an automaton~$\Automaton$ \emph{dominates}
-  a  transition~$\Transition$, written
-  $\Dominates(C, \Automaton, \Transition)$, if
-  every accepting path $p$ of $\Automaton$
-  with $\Transition \in p$ contains at least one transition from
-  $C$. Notably, $\Dominates(\emptyset, \Automaton, \Transition)$ for
-  every unreachable transition $\Transition$, and
-  $\Dominates(\Set{\Transition}, \Automaton, \Transition)$ for every
-  transition~$\Transition$.
-\end{definition}
 
@@
 
-We then introduce the two predicates into our calculus with the following
-definitions:
-
+We introduce the two predicates that will be used by our calculus,
+with the following definitions:
+%
 \begin{definition}\label{def:single-image}
@@
 \end{definition}
-
-We present the rules of \Calculus{} for one automaton in
-\cref{tbl:rules:single}. The rules operate on sets of formulas and
-can be interpreted as rules of a one-sided sequent calculus, in
-which all formulas are located in the antecedent~\cite{Fitting96a}.
-
-Each rule should be read and applied bottom-up and relates
-premises~$\SomeClause_1, \ldots, \SomeClause_k$ with some
-conclusion~$\SomeClause$. When constructing a proof, we start with some
-root~$\SomeClause$, and then apply proof rules to the goals of the proof until a
-goal can be closed, or no more rule is applicable.
-
-A proof in which all proof goals are closed witnesses that the formulas
-in the root~$\SomeClause$ of the proof are inconsistent, and do not
-have any solutions. A proof goal that cannot be closed, but to which no
-more rules are applicable, gives rise to  a solution of the formulas in the
-root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic allowing a solution to be computed using
-standard algorithms~\cite{DBLP:books/daglib/0022394}.
-
-We use the convention of splitting the formulas into linear (in)equalities
-($\SomeInequalities$) and other formulas ($\SomeClause$), and assume that
-predicates $\Image$ and $\Connected$ only occur positively. Since our rules
-operate by adding and matching linear (in)equalities in a proof goal; we use the
-shorthand of listing the matched inequalities as antecedents, e.g., in
-\Propagate{}.
-
-The filtering function~$\Filter$ is evaluated symbolically and can, in practice,
-be read as a function from transitions to $\Naturals$-valued terms
-(e.g.~\texttt{t} or~\texttt{t+1}). In our implementation \Catra{}, described in
-\cref{sec:implementation}, $\Filter$~is a vector of fresh variables with the
-same size as~$\Transitions$.
-
-We use the shorthand notation~$\Transitions_\Automaton$ to refer to the
-transitions of an automaton~$\Automaton$. Additionally, for an
-automaton~$\Automaton = \AutomatonTuple$, we allow mapping the selection function
-like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
-\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
-$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
-instance, the basis for the matched linear inequalities is
-implicitly~$\SomeInequalities$. Similarly, for our commutative monoid $\Monoid =
-\left(X;\MonoidOp;0_{\Monoid}\right)$ and the map into it $\Map : \Strings
-\rightarrow X$, we also allow mapping over transitions:
-$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$.
-
-We require that rules can only be applied when they add new formulas
-on every created branch (the notion of \emph{regularity} of a proof is
-required~\cite{Fitting96a}). For example, this means that \Split{} can
-only be applied to proof goals that contain neither
-$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
-never be applied to split on the same term twice on the same branch.
-This suggests a proof strategy where you \Propagate{} when you can, \Split{}
-when you must, and \Subsume{} when neither is possible anymore.
 
@@
 \end{table}
-We use the symbolic function $\FlowEq(\Automaton, \Filter)$ that generates a set
-of existentially quantified linear inequalities with the following definition,
-where we assign fresh, existentially quantified variables to
-$\FinalStateVar_\State, \Filter(\Transition)$ for every $\State \in
-\AcceptingStates, \Transition \in \Transitions$:
+
+The rules of \Calculus{} for one automaton are given in
+\cref{tbl:rules:single}. The rules operate on sets of formulas and can
+be interpreted as rules of a one-sided sequent calculus, in which all
+formulas are located in the antecedent~\cite{Fitting96a}.
+%
+The rules relate premises~$\SomeClause_1, \ldots, \SomeClause_k$ with
+some conclusion~$\SomeClause$. 
+%
+When constructing a proof, we start with some root~$\SomeClause$, and
+then apply proof rules to the goals of the proof, in bottom-up
+direction, until all goals are closed, or no more rule is applicable.
+
+A proof in which all proof goals are closed shows that the formulas
+in the root~$\SomeClause$ of the proof are inconsistent, and do not
+have any solutions. A proof goal that cannot be closed, but to which no
+more rules are applicable, gives rise to  a solution of the formulas in the
+root~$\SomeClause$. Such an open proof goal will only contain formulas
+in Presburger arithmetic allowing a solution to be computed using
+standard algorithms~\cite{DBLP:books/daglib/0022394}.
+
+We use the convention of splitting the formulas in proof goals into
+linear (in)equalities ($\SomeInequalities$) and other formulas
+($\SomeClause$), and assume that predicates $\Image$ and $\Connected$
+only occur positively.
+%Since our rules operate by adding and matching
+%linear (in)equalities in a proof goal; we use the shorthand of listing
+%the matched inequalities as antecedents, e.g., in \Propagate{}.
+%
+The transition selection function~$\Filter$ is represented
+symbolically and can, in practice, be read as a function from
+transitions to $\Naturals$-valued terms (e.g.~\texttt{t}
+or~\texttt{t+1}). In our implementation \Catra{}, described in
+\cref{sec:implementation}, $\Filter$~is a vector of fresh variables
+with the same size as~$\Transitions$.
+
+To ensure termination, rules can only be applied when they add new formulas
+on every created branch (the notion of \emph{regularity} of a proof is
+required~\cite{Fitting96a}). For example, this means that \Split{} can
+only be applied to proof goals that contain neither
+$\Filter(\Transition) = 0$ nor $\Filter(\Transition) > 0$, and can
+never be applied to split on the same term twice on the same branch.
+
+\medskip
+%
+The rule~\textsc{Expand} expands an $\SinglePredicateInstance$
+predicate into the more basic predicate
+$\Connected(\Automaton, \Filter)$, as well as linear equations
+relating the transitions mentioned by~$\Filter$ with the monoid
+element~$\MonoidElement$, and linear flow equations described
+by~\FlowEq{} (below). Since $\Connected$ and $\Image$ are partially
+redundant and the difference is covered by~$\FlowEq$, we can remove
+the instance of~$\Image$ when applying~$\Expand$. In this sense, we
+split the semantics of the $\Image$~predicate into its counting aspect
+(covered by $\FlowEq$) and its connectedness aspect (covered by
+$\Connected$).
+
+In \textsc{Expand}, we use the shorthand
+notation~$\Map(\FromLabelTo{\State}{\Label}{\State'}) = \Map(\Label)$,
+i.e., we allow the homomorphism~$h$ to be applied also to
+transitions~$\Transition$.
+%
+The predicate $\FlowEq(\Automaton, \Filter)$ represents the flow
+equations to be generated when expanding
+$\SinglePredicateInstance$. We assume that each application of the
+predicate introduces fresh integer variables~$f_q$ for every accepting
+state~$q \in \AcceptingStates$, and define:
+%
 \[
 \begin{aligned}
-  \FlowEq(\Automaton, \Filter) &= \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State) \land
-  \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}\\
-  \Sink(\State) & = 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
-  \In(\State, \Filter) & = \StartFlow(\State) + \sum_{\Transition \in \FromLabelTo{\State'}{}{\State}} \Filter(\Transition)\\
-  \StartFlow(\State)  &= 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
-  \Out(\State, \Filter) &= \sum_{\Transition \in \FromLabelTo{\State}{}{\State'}} \Filter(\Transition)
+  \FlowEq(\Automaton, \Filter) &~=~
+   \AndComp{\State \in \AcceptingStates}{\FinalStateVar_\State \geq 0}
+   \land
+   \sum\limits_{\State \in \AcceptingStates} \FinalStateVar_\State = 1 \land \AndComp{\State \in \States}{\In(\State, \Filter) - \Out(\State, \Filter)} = \Sink(\State)\\
+  \Sink(\State) & ~=~ 0 \text{ if } \State \not\in \AcceptingStates, \FinalStateVar_\State \text{ otherwise.} \\
+  \In(\State, \Filter) & ~=~ \StartFlow(\State) + \sum_{\Transition \in\, (\FromLabelTo{}{}{\State})} \Filter(\Transition)\\
+  \StartFlow(\State)  &~=~ 1 \text{ if } \State = \InitialState, \text{ otherwise $0$.} \\
+  \Out(\State, \Filter) &~=~ \sum_{\Transition \in\, (\FromLabelTo{\State\,}{}{})} \Filter(\Transition)
 \end{aligned}
@@
 
-In addition to \cref{tbl:rules:single}, we assume the existence of a rule
-\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
-formulae, and for the elements of~$\Monoid$.
-
-The $\Propagate{}$ rule allows us to propagate (dis-)connectedness across
-$\Automaton$. It states that we are only allowed to use transitions attached to
-a reachable state, and is necessary to ensure connectedness in the presence of
-cycles in~$\Automaton$.
-
-\textsc{Expand} expands the predicate into its most basic rules; one set of
-linear equations synchronising the transitions mentioned by~$\Filter$ to the
-corresponding Monoid element~$\MonoidElement$, and the linear flow equations of
-the standard Parikh image formulation, as described by~\FlowEq. Since
-$\Connected$ and $\Image$ are partially redundant and the difference is covered
-by~$\FlowEq$, we can remove the instance of~$\Image$ when applying~$\Expand$. In
-this sense, we split the semantics of the $\Image$~predicate into its counting
-aspect (covered by $\FlowEq$) and its connectedness aspect (covered by
-$\Connected$).
-
-Finally, $\Split{}$ allows us to branch the proof tree by trying to exclude a
+The rule~$\Split{}$ allows us to branch the proof tree by trying to exclude a
 contested transition from a potential solution before concluding that it must be
@@
 
-A decision procedure for our predicate in a tableau-based automated theorem
-prover would start by expanding the predicate using the $\Expand{}$~rule. A
-theorem prover would perform algebraic substitution on the underlying constants
-of~$\Filter$, boiling them down to choices of branches, which depend on one
-single variable and loop transitions. This logic corresponds to the placement
-of counters for optimally edge-profiling the CFG of a program, making up a
-minimum-spanning tree of the automaton~\cite{path-profiling}.
+The $\Propagate{}$ rule allows us to propagate (dis-)connectedness
+across $\Automaton$. It states that we are only allowed to use
+transitions attached to a reachable state, and is necessary to ensure
+connectedness in the presence of cycles in~$\Automaton$. The rule
+makes use of the notion of \emph{dominating} sets of transitions, defined
+as follows:
+%
+\begin{definition}
+  A set~$C$ of transitions of an automaton~$\Automaton$ \emph{dominates}
+  a  transition~$\Transition$, written
+  $\Dominates(C, \Automaton, \Transition)$, if
+  every accepting path $p$ of $\Automaton$
+  with $\Transition \in p$ contains at least one transition from
+  $C$. Notably, $\Dominates(\emptyset, \Automaton, \Transition)$ for
+  every unreachable transition $\Transition$, and
+  $\Dominates(\Set{\Transition}, \Automaton, \Transition)$ for every
+  transition~$\Transition$.
+\end{definition}
 
-In order to make the examples below tractable, we will assume the existence of a
+Finally, the rule~\Subsume{} can be applied when the connectedness of
+an automaton has been ensured by exhaustive application of the other
+rules. This suggests a proof strategy where you \Propagate{} when you
+can, \Split{} when you must, and \Subsume{} when neither is possible
+anymore.
+
+In addition to \cref{tbl:rules:single}, we assume the existence of a rule
+\PresburgerClose{}, corresponding to a sound and complete solver for Presburger
+formulae, and for constraints over~$\Monoid$.
+
+\medskip
+%
+A decision procedure would start from one or multiple
+predicates~$\SinglePredicateInstance$ to be satisfied, possibly in
+combination with other constraints about $m$. It would then first
+expand the predicates using the $\Expand{}$~rule, and subsequently
+apply the other rules to search for a solution.
+
+As illustrated in \cref{sec:intuition}, a decision procedure can
+also perform arithmetic rewriting of the occurring terms and
+equations. Such reasoning is not necessary for correctness or
+completeness, but it shortens the examples considerably; we
+will therefore assume the existence of a
 rule \EquationReasoning{} that allows us to perform standard algebraic reasoning
-on linear inequalities. This rule is not necessary for correctness or
-completeness but shortens the examples considerably.
+on linear arithmetic constraints.
+
+%, boiling them down to choices of branches, which depend
+%on one single variable and loop transitions. This logic corresponds to
+%the placement of counters for optimally edge-profiling the CFG of a
+%program, making up a minimum-spanning tree of the
+%automaton~\cite{path-profiling}.
 
 \subsection{An Example}\label{sec:single:example}
-Here we will return to $\mathcal{A}'$ from \cref{sec:intuition} and performs the steps of
+Here we will return to $\mathcal{A}'$ from \cref{sec:intuition} and perform the steps of
 \cref{sec:intuition:algebra,sec:intuition:split} with the
 formal calculus we just established, but exclude \cref{const:s2-in-b} and
-therefore the entire automaton $\mathcal{B}$ since we introduce support for products of
-automata in \cref{sec:multiple}. Note that this example then becomes
+therefore the entire automaton $\mathcal{B}$, since we introduce support for products of
+automata in \cref{sec:multiple}. The example then becomes
 satisfiable, as $s_1 = \Char{dd}, i = 1, s_2 = \Char{adda}, l_A = 4$ is a
@@
 here as in \cref{sec:intuition} to represent what is essentially a sparse vector
-of $1$ and $0$ coefficients, i.e. = $\left[ l_{A+}, i_+ \right] = \left[1, 1,
-0\right]$. Initially, we let $\Filter$ map to fresh variables to obtain:
+of $1$ and $0$ coefficients, i.e. $\left[ l_{A+}, i_+ \right] = \left[1, 1,
+0\right]$. Initially, we let $\Filter$ map to fresh variables to obtain, after some simplifications:
 \begin{equation}\label{eq:single-filter}
   \begin{array}{rl@{\qquad}rl@{\qquad}rl}
-    \Filter(\FromLabelTo{D}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A})
+    \Filter(\FromLabelTo{D}{d/[l_{A+}, n_+]}{F_A})
     &= \TransitionVar_1
     &
-    \Filter(\FromLabelTo{F_A}{\Sigma/\begin{bmatrix} l_{A+} \end{bmatrix}}{F_A} )
+    \Filter(\FromLabelTo{F_A}{\Sigma/[ l_{A+} ]}{F_A} )
     &= \TransitionVar_2   
     &
-    \Filter(\FromLabelTo{I_A}{c/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{F_A} ) 
+    \Filter(\FromLabelTo{I_A}{c/[l_{A+}, n_+]}{F_A} ) 
     &=  \TransitionVar_3 
     \\
-    \Filter(\FromLabelTo{I_A}{d/\begin{bmatrix}l_{A+}, n_+\end{bmatrix}}{D}) 
+    \Filter(\FromLabelTo{I_A}{d/[l_{A+}, n_+]}{D}) 
     &= \TransitionVar_4
     &
-    \Filter(\FromLabelTo{I_A}{\Sigma/\begin{bmatrix} l_{A+} i_+\end{bmatrix}}{I_A}) 
+    \Filter(\FromLabelTo{I_A}{\Sigma/[ l_{A+}, i_+]}{I_A}) 
     &= \TransitionVar_5  \\[3ex]
@@
     \FlowEq(\mathcal{A}', \Filter)  = \begin{aligned}
-                       1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_3 + \TransitionVar_1\\
+                       1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_4 + \TransitionVar_3\\
                     \TransitionVar_4 &= \TransitionVar_1 \\
-                    \TransitionVar_3 + \TransitionVar_1 + \cancel{\TransitionVar_2} &\geq \cancel{\TransitionVar_2}  \\
+                    \TransitionVar_3 + \TransitionVar_1 + \cancel{\TransitionVar_2} &= \cancel{\TransitionVar_2} + 1  \\
                     \end{aligned}}
@@
 
-Like in \cref{sec:a_1,sec:intuition:algebra}, we apply reasoning by equivalence
-  to end up with fixed values for several of the transition variables
-  $\TransitionVar_k$, and eventually we use \cref{const:more-inside-than-before}
-  and \cref{const:something-before-and-after} to conclude that the transition
-  directly from state $I_{A}$ to $F_{A}$ is incompatible with the constraint.
+The proof tree is shown in \cref{fig:derivation:single}.  Like in
+\cref{sec:a_1,sec:intuition:algebra}, by arithmetic reasoning the
+calculus ends up with fixed values for several of the transition
+variables $\TransitionVar_k$, and eventually concludes that the
+transition directly from state $I_{A}$ to $F_{A}$ (variable~$x_3$)
+is incompatible with
+the constraint.  It is then possible to remove the $\Connected{}$
+predicate using \Subsume{}, since \Propagate{} is not able to infer
+further constraints, and no non-trivial applications of \Split{}
+remain to be done.  After this, by further arithmetic reasoning we can
+derive a solution~$l_A = 4, i = 1, n = 2$, and conclude that the root
+constraint is satisfiable.  To obtain values for the string
+variables~$s_1, s_2$ corresponding to the solution, one can
+construct an accepting path of the automaton with each transition~$t$
+taken $\Filter(t)$ times.
 
-  Then, we remove the $\Connected{}$ predicate using \Subsume{} since no
-   transition can become disconnected (and never could for $\mathcal{A}'$,
-   since all looping transitions are attached to the initial and single
-   accepting state and therefore always must be reachable from any accepting
-   path). After this, the rest of the reasoning can be continued without any
-   part of \Calculus{} since all values of the existentially quantified
-   transition variables $\TransitionVar_1,\ldots,\TransitionVar_5$ compatible
-   with
+\iffalse
+
+the rest of the reasoning can be
+continued without any part of \Calculus{} since all values of the
+existentially quantified transition variables
+$\TransitionVar_1,\ldots,\TransitionVar_5$ compatible with
 \cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring,const:something-before-and-after}
-   are represented by the flow equations introduced by $\Expand{}$. The
-   derivation is shown in \cref{fig:derivation:single}.
+are represented by the flow equations introduced by $\Expand{}$.
+\fi
 
-\begin{figure}[ht]
+\begin{figure}[tb]
   \centering
 \begin{prooftree}
-  \hypo[]{{\text{Regular solving finishes the proof}}}
-  \infer1[\EquationReasoning{}: $i = 1$]{
-  i = 1
-  \land n = 2
-  \land \TransitionVar_2 = l_A - 3
-  \land \TransitionVar_3 = 0
-  \land \TransitionVar_5 = 1
-  \land \TransitionVar_1 = \TransitionVar_4 = 1
+  \hypo[]{$\vdots$}
+  \infer1[]{
+      x_1 = 1 \land
+      x_2 = l_A - 3 \land
+      x_3 = 0 \land 
+      x_4 = 1 \land
+      x_5 = 1 \land
+      n = 2 \land i = 1
+    \land l_A > 3
   }
-  \infer1[]{
-    \begin{aligned}
-    i = \TransitionVar_5 \land 
-    n = 2 \land \\
-    0 < i < 2
-    \land l_A - i - 2 > 0 
-    \land \TransitionVar_2 = l_A - i - 2 
-    \land \TransitionVar_3 = 0 \\
-    \end{aligned}}
   \infer1[\Subsume]{
-    \begin{aligned}
-    \Connected(\mathcal{A}', \Filter) \land 
-    \TransitionVar_1 = 1 \land 
-    i = \TransitionVar_5 \land 
-    n = 2 \land \\
-    0 < i < 2 
-    \land l_A - i - 2 > 0
-    \land \TransitionVar_2 = l_A - i - 2 
-    \land \TransitionVar_3 = 0
-    \end{aligned}
+    \begin{array}{c}
+      \Connected(A', \Filter) \land \mbox{}\\
+      x_1 = 1 \land
+      x_2 = l_A - 3 \land
+      x_3 = 0 \land 
+      x_4 = 1 \land
+      x_5 = 1 \land
+      n = 2 \land i = 1
+    \land l_A > 3
+    \end{array}
+%    \begin{aligned}
+%    i = \TransitionVar_5 \land 
+%    n = 2 \land \\
+%    0 < i < 2
+%    \land l_A - i - 2 > 0 
+%    \land \TransitionVar_2 = l_A - i - 2 
+%    \land \TransitionVar_3 = 0 \\
+%  \end{aligned}
   }
-  \infer1[\EquationReasoning{}: $\TransitionVar_1 = 1$ \cref{sec:a_1}]{
-    \begin{aligned}
-    \Connected(\mathcal{A}', \Filter) \land
-    i = \TransitionVar_5 \land 
-    n = 2\TransitionVar_1 
-    \land n > i > 0 \\
+  \infer1[\EquationReasoning: $=$-simplification]{
+    \begin{array}{c}
+      \Connected(A', \Filter) \land\mbox{}\\
+      x_1 = n - 1 \land
+      x_2 = l_A - n - i \land
+      x_3 = 2 - n \land 
+      x_4 = n - 1 \land
+      x_5 = i \land \mbox{}\\
+      n = 2 \land i = 1
+    \land l_A > 3
+    \end{array}
+%    \begin{aligned}
+%    \Connected(\mathcal{A}', \Filter) \land 
+%    \TransitionVar_1 = 1 \land 
+%    i = \TransitionVar_5 \land 
+%    n = 2 \land \\
+%    0 < i < 2 
+%    \land l_A - i - 2 > 0
+%    \land \TransitionVar_2 = l_A - i - 2 
+%    \land \TransitionVar_3 = 0
+%    \end{aligned}
+  }
+  \infer1[\EquationReasoning{}: $>$-reasoning]{
+    \begin{array}{c}
+      \Connected(A', \Filter) \land\mbox{}\\
+      x_1 = n - 1 \land
+      x_2 = l_A - n - i \land
+      x_3 = 2 - n \land 
+      x_4 = n - 1 \land
+      x_5 = i \land \mbox{}\\
+      n > i 
     \land l_A - i - n > 0
-    \land \TransitionVar_2 = l_A - i - n 
-    \land \TransitionVar_3 = 1 - \TransitionVar_1 
-    \end{aligned}
+    \land i > 0
+    \end{array}
+%    \begin{aligned}
+%    \Connected(\mathcal{A}', \Filter) \land
+%    i = \TransitionVar_5 \land 
+%    n = 2\TransitionVar_1 
+%    \land n > i > 0 \\
+%    \land l_A - i - n > 0 
+%    \land \TransitionVar_2 = l_A - i - n 
+%    \land \TransitionVar_3 = 1 - \TransitionVar_1 
+%    \end{aligned}
   }
-  \infer1[\EquationReasoning{}: $=$-elimination]{
-    \begin{aligned}
-    \Connected(\mathcal{A}', \Filter)  
-    \land i = \TransitionVar_5 
-    \land n = \TransitionVar_4 + \TransitionVar_1 
-    \land n > i > 0
+  \infer1[\EquationReasoning{}: $=$-simplification]{
+    \begin{array}{c}
+      \Connected(A', \Filter) \land\mbox{}\\
+      1 = x_4 + x_3 \land x_4 = x_1 \land x_3 + x_1 = 1 \land \mbox{}\\
+      l_A = x_1 + x_2 + x_3 + x_4 + x_5 \land
+      i = x_5 \land 
+      n = x_1 + x_3 + x_4 \land \mbox{}\\
+      n > i 
     \land l_A - i - n > 0
-    \land \TransitionVar_4 = \TransitionVar_1 \\
-    \land \TransitionVar_2 = l_A - i - n 
-    \land \TransitionVar_3 = 1 - \TransitionVar_1
-    \land 1 = \TransitionVar_3 + \TransitionVar_1 
     \land i > 0
-    \land \TransitionVar_4 = \TransitionVar_1 
-    \end{aligned}
+    \end{array}
+%    \begin{aligned}
+%    \Connected(\mathcal{A}', \Filter)  
+%    \land i = \TransitionVar_5 
+%    \land n = \TransitionVar_4 + \TransitionVar_1 
+%    \land n > i > 0
+%    \land l_A - i - n > 0
+%    \land \TransitionVar_4 = \TransitionVar_1 \\
+%    \land \TransitionVar_2 = l_A - i - n 
+%    \land \TransitionVar_3 = 1 - \TransitionVar_1
+%    \land 1 = \TransitionVar_3 + \TransitionVar_1 
+%    \land i > 0
+%    \land \TransitionVar_4 = \TransitionVar_1 
+%    \end{aligned}
   }
-  \infer1[\EquationReasoning{}: expand $\Sigma, \FlowEq$]{
-    \begin{aligned}
-    i = \TransitionVar_5 
-    \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k 
-    \land n > i > 0
+  \infer1[(Expanding $\FlowEq$)]{
+    \begin{array}{c}
+      \Connected(A', \Filter) \land
+      \FlowEq(A', \Filter) \land \mbox{}\\
+      l_A = x_1 + x_2 + x_3 + x_4 + x_5 \land
+      i = x_5 \land 
+      n = x_1 + x_3 + x_4 \land \mbox{}\\
+      n > i 
     \land l_A - i - n > 0
-    \land n = \TransitionVar_4 + \TransitionVar_1\\
-    \land \Connected(\mathcal{A}', \Filter) 
-    \land \FlowEq(\mathcal{A}', \Filter)
-    \end{aligned}
+    \land i > 0
+    \end{array}
+%    \begin{aligned}
+%    i = \TransitionVar_5 
+%    \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k 
+%    \land n > i > 0
+%    \land l_A - i - n > 0
+%    \land n = \TransitionVar_4 + \TransitionVar_1\\
+%    \land \Connected(\mathcal{A}', \Filter) 
+%    \land \FlowEq(\mathcal{A}', \Filter)
+%    \end{aligned}
   }
-  \infer1[\EquationReasoning{}: expand sum]{
+  \infer1[(Expanding sum)]{
+    \begin{array}{c}
       \Connected(A', \Filter) \land
@@
       \left[l_A, i, n\right] =
-    \sum\limits_{\Transition \in \Transitions_{\mathcal{A}'} } \Map(\Transition) \cdot \Filter(\Transition)
+      \sum\limits_{\Transition \in \Transitions_{\mathcal{A}'} } \Map(\Transition) \cdot \Filter(\Transition) \land\mbox{}\\
+     n > i 
+    \land l_A - i - n > 0
+    \land i > 0
+    \end{array}
   }
@@
     \Image{}_{\mathcal{A}', \Map}(\Filter, \left[l_A, i, n\right]) 
-    \land n \geq i 
+    \land n > i 
     \land l_A - i - n > 0
@@
 \end{prooftree}
-\caption{A derivation tree for an application of
+\caption{A proof tree for
 \cref{const:s1-in-c-dd,const:s1-substring,const:something-before-and-after,const:more-inside-than-before}
 from \cref{ex:string-constraints}, corresponding to handling the Parikh
-automaton $\mathcal{A}'$ of \cref{fig:aut_a}. Note the liberal use of removal of equal
-variables (e.g. if $x = y$ appears as a statement, we substitute $y$ for $x$
-everywhere and drop $\ldots \land x=y \land \ldots$) to keep the size of the
-equations manageable.}\label{fig:derivation:single}
+automaton $\mathcal{A}'$ of \cref{fig:aut_a}.}\label{fig:derivation:single}
 \end{figure}
-
-To obtain the automaton corresponding to the final model, one can evaluate
-$\Filter(\Transition) = k$ for each transition $\Transition$ and construct a
-path with each such transition taken $k$ times.
 

### 2023-07-11 multiple.tex
restructuring

@@
     partial products.
+
+Additionally, for an
+automaton~$\Automaton = \AutomatonTuple$, we allow mapping the selection function
+like so: $\Filter(\Automaton) = \Tuple{\States, \InitialState, \AcceptingStates,
+\Set{\Transition \in \Transitions \SuchThat \Filter(\Transition) > 0}}$, i.e.
+$\Automaton$~with only the transitions for which~$\Filter$ is positive. In this
+instance, the basis for the matched linear inequalities is
+implicitly~$\SomeInequalities$.
 

### 2023-07-12 calculus.tex
typo

@@
 satisfying assignment to
-\cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring}.
+\cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring,const:something-before-and-after}.
 
@@
     \Path)$ and $\val_\beta(\MonoidElement) =
-    \Map(\WordOf(\Path))$. Since immediately implies that $\beta$
+    \Map(\WordOf(\Path))$. This immediately implies that $\beta$
     satisfies $\Connected(\Automaton, \Filter)$, since a path is

### 2023-07-12 multiple.tex
tried to fix Materialise

--- 2023-07-11 multiple.tex
@@
                      \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
-                     \Automaton'},~ \SomeInequalities, \SomeClause
-           }{\Map}{\Filter'}{\MonoidElement}
+                     \Automaton'}
+           }{\Map}{\Filter'}{\MonoidElement},~ \SomeInequalities, \SomeClause
            \end{array}}
@@
       $\begin{array}{@{}l@{}}
-         \Filter' = \Extend(\Automaton_i, \Automaton_j,\Filter),\\
+         \Filter' = \Extend(\Automaton',\Filter),\\
          1 \leq i < j \leq k, \\
-      \Automaton' = \Automaton_i \times \Automaton_j,\\
+      \Automaton' =
+         \Filter_{\SomeInequalities}(\Automaton_i) \times \Filter_{\SomeInequalities}(\Automaton_j)\\
       \end{array}$ \\
@@
 $\Automaton_i, \Automaton_j$ as index of the $\Image$ predicate, and
-instead adds the product~$\Automaton_i \times \Automaton_j$, while
-also introducing the flow equations and the $\Connected$ predicate.
+instead adds the product of the automata restricted to the
+transitions that can still be taken. The rule
+also introduces the flow equations and the $\Connected$ predicate.
 
-The rule~$\Materialise$ also has to connect the newly introduced
-product~$\Automaton_i \times \Automaton_j$ to the previous automata
+In $\Materialise$, we use the following notation for pruning away
+parts of an automaton based on the transition selection
+function~$\sigma$, only keeping those transitions for which~$\Filter$
+is positive:
+\begin{equation*}
+  \Filter_{\SomeInequalities}(\Automaton) =
+  \Tuple{\States, \InitialState, \AcceptingStates,
+\Set{\Transition \in \Transitions \SuchThat (\Filter(\Transition) = 0) \not\in E}}
+\end{equation*}
+This filtering operation can be optimised to also eliminate states from
+$Q$ and $F$ that become unreachable; this is kept implicit for sake of
+presentation at this point.
+
+The rule~$\Materialise$ has to connect the newly introduced
+product to the previous automata
 $\Automaton_i, \Automaton_j$. This is done by extending the
@@
 \begin{equation*}
-  \Extend(\Automaton_i, \Automaton_j,\sigma)(t)
+  \Extend(\Automaton',\sigma)(t)
   ~~=~~
   \begin{cases}
-    x_{\Transition} & t \in \Transitions_{\Automaton_i \times \Automaton_j}
+    x_{\Transition} & t \in \Transitions_{\Automaton'}
     \\
@@
 $\Automaton_i$ has to coincide with the sum of the multiplicities of
-transitions in $\Automaton_i \times \Automaton_j$ derived from $t$,
+transitions in $\Filter_{\SomeInequalities}(\Automaton_i) \times \Filter_{\SomeInequalities}(\Automaton_j)$ derived from $t$,
 and similarly for $\Automaton_j$:
@@
   & \Set{ 
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter'(\Transition')
+    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State'_R}}} \Filter'(\Transition')
   \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
   & \Set{
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
+    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State'_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
   }
@@
 transition~$\Transition$ since $\mathcal{B}$ only counts length. The reader is
-strongly advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
-\cref{sec:intuition} while going through this example! In particular, note that
-the transition variables of the figures match the ones used here.
+advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
+\cref{sec:intuition} while going through this example.
+The transition variables of the figures match the ones used here.
 

### 2023-07-12 abstract.tex
Abstract

A common problem in string constraint solvers is computing the Parikh image, a
linear arithmetic formula that describes all possible combinations of character
counts in strings of a given language. Automata-based string solvers frequently
need to compute the Parikh image of large products (intersections) of
nondeterministic automata, which in many operations is both prohibitively slow
and memory-intensive. We contribute a novel understanding of how Parikh maps can
be tackled as a constraint solving problem to solve real-world constraints
stemming from functions on regular languages, including length and indexing
constraints. Furthermore, we show how this formulation can be efficiently
implemented as a calculus, \Calculus{}, in an automated theorem prover
supporting Presburger logic.  We implement \Calculus{} in a tool called
\Catra{}, and evaluate it on constraints produced by the \OstrichPlus{} string
constraint solver when solving Parikh automata intersection problems produced
when solving standard string constraint benchmarks involving cardinality
constraints on strings. We show that our solution strictly outperforms the
standard approach described in~\citeauthor{generate-parikh-image} as well as the
over-approximating method recently described by~\citeauthor{approximate-parikh}
by a wide margin, and for realistic timeouts for constraint solving also
the~\Nuxmv{} model checker.

### 2023-07-12 evaluation.tex
more

@@
 automata intersection problems generated by the \OstrichPlus{} string constraint
-solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
+solver~\cite{ostrich-plus} when solving the PyEx benchmarks, 
+which are string constraints from symbolic execution that are known
+to be hard for many solvers~\cite{pyex}. After generating an
 initial~\InitialNrBenchmarks{}, we remove~\NrTrivial{} instances solved in under
-five seconds by baseline as well as~\NrInvalid{} instances that produced errors
-during experimentation. We also attempted to benchmark instances generated by
-\Ostrich{} solving the kaluza-len (\numprint{38227}~instances) and pyex-len
-(\numprint{791}~instances) benchmark suites but discarded them since they were
-all trivial. For instance, on a laptop, \Calculus{} could solve every instance
-from kaluza-len in under five seconds (mean about 0.1s). The benchmarks are run
-on commit~\texttt{\commit}.
+five seconds by baseline as well \Fudge{as~\NrInvalid{} instances that produced errors
+during experimentation}.
+% Better explanation needed why we removed the benchmarks
+ We also attempted to benchmark instances generated by
+\OstrichPlus{} solving the Kaluza benchmarks~\cite{Saxena10:kaluza}
+(\numprint{38227}~instances),
+\iffalse
+and pyex-len
+(\numprint{791}~instances) benchmark suites\fi
+but discarded them since they all turned out to be trivial for our tool
+\Catra{}: every instance
+could be solved in under five seconds, with a mean runtime of about 0.1s.
+\Fudge{The benchmarks are run
+on commit~\texttt{\commit}.}
+% Commit of what?
 
@@
 \begin{table}
-  \centering
+  \begin{center}
   \input{graphs/solved_pivot_table.tex}
+  \end{center}
   \caption{Number of successful results within a timeout of \RuntimeTimeout{}.
@@
 vastly.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "main"
+%%% End:

### 2023-07-12 extensions.tex
Evaluation + kill your darlings
more

@@
 unreachable, we can learn the clause $\Connected(\Automaton, \Filter) \land
-\sum_{\Transition' \in C} \Filter(\Transition') = 0 \implies \Transition = 0$ where
+\bigwedge_{\Transition' \in C} \Filter(\Transition') = 0 \implies \Transition = 0$ where
 $\Dominates(C, \Automaton, \Transition)$ for some minimal $C$. Note that this
@@
 attempt to materialise a product of two ($\Filter$-filtered) automata
-$\Automaton_1 \times \Automaton_2$ produces an empty product, we can determine
+produces an empty product, we can determine
 the cause of the failure with respect to the automata and their respective
@@
 
+\iffalse
 \subsection{Transducers}\label{sec:ext:transducers}
@@
 definition of products was left out of the definition. All that would be needed
-is an appropriate update of the definitions of products. Simiarly, \Catra was
+is an appropriate update of the definitions of products. Similarly, \Catra was
 written with modularity in mind, and it should be straightforward to extend both
@@
     of~\Catra, but has been neither optimised nor fully tested.
+\fi
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "main"
+%%% End:

### 2023-07-12 implementation.tex
more

@@
 Additionally, baseline and \Calculus{} also support generating the Presburger
-formula describing the constraints of the input file. Baseline uses standard
-quantifier elimination, and \Calculus{} uses the method described in
-\cref{sec:finding-the-image}.
+formula describing the constraints of the input file, i.e., computing
+a closed-from representation of the complete Parikh image.
+% Baseline uses standard
+%quantifier elimination, and \Calculus{} uses the method described in
+%\cref{sec:finding-the-image}.
 
-Since \Princess{} does not support the multiple-arity predicates used in
-\Calculus{}, we have implemented variable-length arguments using additional
-helper predicates. These are $\Unused{}(\Automaton)$, which marks an automaton
-as unused in any product, and $\TransitionMask{}(\Automaton, \Transition,
-\Filter(\Transition))$ which associates a transition $\Transition$ and automaton
-$\Automaton$ with its corresponding transition variable. Additionally, we
-associate each predicate with an instance variable to tie helper predicates
-to the $\Image$ predicate they originated from.
+Since \Princess{} does not support the multiple-arity predicates used
+in \Calculus{}, we have implemented variable-length arguments using
+additional helper predicates. These are $\Unused{}(\Automaton)$, which
+marks an automaton as unused in any product, and
+$\TransitionMask{}(\Automaton, \Transition, \Filter(\Transition))$
+which associates a transition $\Transition$ and automaton $\Automaton$
+with its corresponding transition
+variable~$\Filter(\Transition)$. Additionally, we associate each
+predicate with an instance variable to tie helper predicates to the
+$\Image$ predicate they originated from.
 
@@
 states eliminated during construction. Any automaton we produce contains only
-states that are both reachable from the initial state and has a path to an
+states that are both reachable from the initial state and have a path to an
 accepting state. We never perform any other minimisation on the automata for
@@
 implementation we put off computing intermediate products until six transition
-variables of one of the automata is known to be either present or absent. The
+variables of one of the automata is known to be either \Fudge{present or absent}. The
 number was chosen experimentally. The other automaton for the product is
 selected randomly.
+
+% this is unclear; don't we delay computing products until one of the
+% automata is (almost) completely defined?
 
@@
 tail-heavy our problem is.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "main"
+%%% End:

### 2023-07-12 main.tex

--- 2023-06-28 main.tex
@@
 \begin{abstract}
-  A common problem in string constraint solvers is computing the Parikh image, a
-   set of linear equations that describe all possible combinations of character
-   counts in strings of a given language. Automata-based string solvers
-   frequently need to compute the Parikh image of large products (intersections)
-   of nondeterministic automata, which in many operations is both prohibitively
-   slow and memory-intensive. We contribute a novel understanding of how Parikh
-   maps can be tackled as a constraint solving problem to solve real-world
-   constraints stemming from functions on regular languages, most notably the
-   length constraint. Furthermore, we show how this formulation can be
-   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
-   prover supporting Presburger logic.  We implement \Calculus{} in a tool
-   called \Catra{}, and evaluate it on constraints produced by the
-   \OstrichPlus{} string constraint solver when solving Parikh automata
-   intersection problems produced when solving standard string constraint
-   benchmarks involving string length constraints. We show that our solution
-   strictly outperforms the standard approach described
-   in~\citeauthor{generate-parikh-image} as well as the over-approximating
-   method recently described by~\citeauthor{approximate-parikh} by a wide
-   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
-   model checker.
+\input{abstract.tex}
 \end{abstract}

### 2023-07-12 implementation.tex
Fix implementation

@@
 Product materialisation is the final piece of the puzzle. In the current
-implementation we put off computing intermediate products until six transition
-variables of one of the automata is known to be either \Fudge{present or absent}. The
+implementation we put off computing intermediate products until at most six transition
+variables of one of the automata are not known to always be used ($> 0$) or always unused ($\leq 0$). The
 number was chosen experimentally. The other automaton for the product is
 selected randomly.
-
-% this is unclear; don't we delay computing products until one of the
-% automata is (almost) completely defined?
 

### 2023-07-12 evaluation.tex
Commit of what
They were duplicates

@@
 initial~\InitialNrBenchmarks{}, we remove~\NrTrivial{} instances solved in under
-five seconds by baseline as well \Fudge{as~\NrInvalid{} instances that produced errors
-during experimentation}.
-% Better explanation needed why we removed the benchmarks
+five seconds by baseline as well as~\NrInvalid{} instances that were duplicates of other instances in the set.
+
  We also attempted to benchmark instances generated by
@@
 could be solved in under five seconds, with a mean runtime of about 0.1s.
-\Fudge{The benchmarks are run
-on commit~\texttt{\commit}.}
-% Commit of what?
+The benchmarks are run
+on commit~\texttt{\commit} of \Catra{}\footnote{Github repository redacted for anonymity.}.
 

### 2023-07-12 evaluation.tex
results

--- 2023-07-12 evaluation.tex
@@
 \subsection{Evaluation in \Ostrich{}}\label{}%
-\Catra{} was experimentally integrated as the Parikh automata product solver for
-an in-progress version of \Ostrich{} (which shares \Princess{} as the backing
-theorem prover) with cost-enriched automata (CEFA). \Ostrich{} (\textit{sans}
-CEFA) recently participated in the 2023~SMT-COMP, winning the single-query track
-for quantifier-free strings (\texttt{QF\_S}) as well as the entire string
-solving track on unsatisfiable instances~\cite{smt-comp-23}.
 
-The \texttt{QF\_S} track contains no Parikh automata intersection problems,
-however. For that, we need to look to the linear integer arithmetic
-(\texttt{QF\_SLIA}), where \Ostrich{} did worse in SMT-COMP. The evaluations in
-\cref{sec:scaling} use inputs generated by an older branch of CEFA-enriched
-\Ostrich{} on the PyEx benchmarks, which are part of the set used in the
-competition. In this section, we evaluated a version of \Ostrich{} using
-\Catra{} as its back-end \Fudge{on StarExec, the same cluster SMT-COMP ran on}
-on \numprint{1000}~randomly selected instances from the PyEx benchmarks. The
-results are shown in comparison \Fudge{to the other provers of the 2023
-competition on the same benchmarks, as well as \Ostrich{} using the baseline
-back-end described in \cref{sec:implementing-baseline}}.
+\begin{table}
+  \begin{center}
+  \input{graphs/solved_smt_comp.tex}
+  \end{center}
+  \caption{Number of solved benchmarks, on a subset of 1000
+    PyEx benchmarks from SMT-COMP.}\label{tab:solve-status-smt-comp}
+\end{table}
+
+To evaluate the effectiveness of \Calculus{} in an actual string solver,
+\Catra{} was experimentally integrated as the Parikh automata product solver
+into the string solver \Ostrich{}~version~1.3. \Ostrich{} is an independently
+developed solver that participated in the recent SMT-COMP~2023\footnote{\url{https://smt-comp.github.io/2023/participants/ostrich}},
+ winning the single-query track
+for quantifier-free strings (\texttt{QF\_S}), as well as dominating
+other solvers on unsatisfiable string benchmarks~\cite{smt-comp-23}.
+
+At SMT-COMP, \Ostrich{} competed using three different back-ends:
+\begin{itemize}
+\item \textbf{BW-Str:} a backward-propagation-based solver, not utilising
+  Parikh automata~\cite{ostrich}.
+\item \textbf{ADT-Str:} a solver based on algebraic data-types.
+\item \textbf{CE-Str:} a re-implementation of the \OstrichPlus{}
+  algorithm~\cite{ostrichplus}., using Parikh automata and a variant
+  of our baseline method~\cite{generate-parikh-image}.
+\end{itemize}
+
+For our experiments, we modified the CE-Str solver to apply \Catra{}
+instead of the previous baseline method, resulting in a new back-end
+\textbf{CA-Str}.
+
+As benchmarks, we again focused on the PyEx benchmarks~\cite{pyex}, of
+which we picked 1000~problems uniformly at random. We ran the
+different back-ends on those 1000~problems with a 120s timeout.
+%
+%The \texttt{QF\_S} track contains no Parikh automata intersection problems,
+%however. For that, we need to look to the linear integer arithmetic
+%(\texttt{QF\_SLIA}), where \Ostrich{} did worse in SMT-COMP. The evaluations in
+%\cref{sec:scaling} use inputs generated by an older branch of CEFA-enriched
+%\Ostrich{} on the PyEx benchmarks, which are part of the set used in the
+%competition. In this section, we evaluated a version of \Ostrich{} using
+%\Catra{} as its back-end \Fudge{on StarExec, the same cluster SMT-COMP ran on}
+%on \numprint{1000}~randomly selected instances from the PyEx benchmarks. 
+The results are shown in \cref{tab:solve-status-smt-comp}.
+%
+% comparison \Fudge{to the other provers of the 2023
+%competition on the same benchmarks, as well as \Ostrich{} using the baseline
+%back-end described in \cref{sec:implementing-baseline}}.
 
@@
 performance of an automaton-based string solver with cost-enriched automata like
-\Ostrich{}. However, significant engineering effort remains to bring the
-integration to a state where it can participate in the full SMT-COMP.
-
-\Fudge{RESULTS HERE PLEASE!}
+\Ostrich{}. The results should be considered as preliminary, however, 
+as we believe that a deeper integration of our \Catra{} solver into
+string solvers can lead to significant performance gains.
 

### 2023-07-12 evaluation.tex
more

@@
 for quantifier-free strings (\texttt{QF\_S}), as well as dominating
-other solvers on unsatisfiable string benchmarks~\cite{smt-comp-23}.
+other solvers on the complete set of
+ unsatisfiable string benchmarks~\cite{smt-comp-23}.
 
-At SMT-COMP, \Ostrich{} competed using three different back-ends:
+ At SMT-COMP, \Ostrich{} competed using different
+ (automata-based) back-ends:
 \begin{itemize}
@@
   Parikh automata~\cite{ostrich}.
-\item \textbf{ADT-Str:} a solver based on algebraic data-types.
+%\item \textbf{ADT-Str:} a solver based on algebraic data-types.
 \item \textbf{CE-Str:} a re-implementation of the \OstrichPlus{}
-  algorithm~\cite{ostrichplus}., using Parikh automata and a variant
-  of our baseline method~\cite{generate-parikh-image}.
+  algorithm~\cite{ostrich-plus}, using Parikh automata and the
+  encoding from~\cite{generate-parikh-image}.
 \end{itemize}
@@
 As benchmarks, we again focused on the PyEx benchmarks~\cite{pyex}, of
-which we picked 1000~problems uniformly at random. We ran the
-different back-ends on those 1000~problems with a 120s timeout.
+which we picked 1000~problems uniformly at random. We ran
+different combinations of the
+back-ends on those 1000~problems with a 120s timeout.
 %
@@
 
-We include this section to show that \Calculus{} can be used to enhance the
-performance of an automaton-based string solver with cost-enriched automata like
-\Ostrich{}. The results should be considered as preliminary, however, 
+As can be seen from the table, integrating \Catra{} as a back-end
+leads to gains both on satisfiable and unsatisfiable problems. The
+increase is more relevant for the unsatisfiable problems, given that
+\Ostrich{} is a state-of-the-art solver in particular on such benchmarks.
+
+The results show that \Calculus{} can be used to enhance the
+performance of an automaton-based string solver.
+The results should be considered as preliminary, however, 
 as we believe that a deeper integration of our \Catra{} solver into

### 2023-10-20 evaluation.tex
Something
New ostrich

@@
 
-The benchmarks are executed in parallel across a cluster of \Fudge{identical}
+The benchmarks are executed in parallel across a cluster of identical
 machines, \BenchmarkRig{}, but as the sole job for each machine.  We compiled
@@
 does baseline, while being similar at satisfiable instances. Both \Calculus{} and
-\Nuxmv outperforms baseline on satisfiable instances. On satisfiable
+\Nuxmv{} outperforms baseline on satisfiable instances. On satisfiable
 instances, \Nuxmv and \Calculus{} have similar performance.
@@
   \begin{center}
-  \input{graphs/solved_smt_comp.tex}
+  \input{experiments/qf_slia_comparison.tex}
   \end{center}
-  \caption{Number of solved benchmarks, on a subset of 1000
-    PyEx benchmarks from SMT-COMP.}\label{tab:solve-status-smt-comp}
+  \caption{Number of solved benchmarks from the single-query QF\_SLIA track of
+  quantifier-free string linear integer arithmetic at SMT-COMP 2023. The numbers
+  for each solver are from the publicly available data files, except for
+  \Ostrich-\Catra{}, which are executed by us. }\label{tab:solve-status-smt-comp}
 \end{table}

### 2023-10-20 evaluation.tex
discussion of data

@@
 \Nuxmv{} outperforms baseline on satisfiable instances. On satisfiable
-instances, \Nuxmv and \Calculus{} have similar performance.
+instances, \Nuxmv{} and \Calculus{} have similar performance.
 
@@
   \end{center}
-  \caption{Number of solved benchmarks from the single-query QF\_SLIA track of
-  quantifier-free string linear integer arithmetic at SMT-COMP 2023. The numbers
-  for each solver are from the publicly available data files, except for
-  \Ostrich-\Catra{}, which are executed by us. }\label{tab:solve-status-smt-comp}
+  \caption{Number of solved benchmarks from the single-query QF\_SLIA
+    track of quantifier-free string linear integer arithmetic at
+    SMT-COMP 2023. The numbers for each solver are from the publicly
+    available data files, except for CA-Str, which is executed by us
+    (600s timeout, AMD Opteron 2220
+    SE), and \Ostrich+CA, which is a virtual portfolio. }
+  \label{tab:solve-status-smt-comp}
 \end{table}
@@
 
- At SMT-COMP, \Ostrich{} competed using different
- (automata-based) back-ends:
+ At SMT-COMP, \Ostrich{} competed using a portfolio of three different
+ back-ends:
 \begin{itemize}
@@
   Parikh automata~\cite{ostrich}.
-%\item \textbf{ADT-Str:} a solver based on algebraic data-types.
+\item \textbf{ADT-Str:} a solver based on algebraic data-types.
 \item \textbf{CE-Str:} a re-implementation of the \OstrichPlus{}
@@
 instead of the previous baseline method, resulting in a new back-end
-\textbf{CA-Str}.
+\textbf{CA-Str}. We created also a new virtual portfolio~\Ostrich{}+CA
+by combining the three solvers in \Ostrich{}~1.3 with \textbf{CA-Str}.
 
-As benchmarks, we again focused on the PyEx benchmarks~\cite{pyex}, of
-which we picked 1000~problems uniformly at random. We ran
-different combinations of the
-back-ends on those 1000~problems with a 120s timeout.
+%As benchmarks, we again focused on the PyEx benchmarks~\cite{pyex}, of
+%which we picked 1000~problems uniformly at random. We ran
+%different combinations of the
+%back-ends on those 1000~problems with a 120s timeout.
 %
@@
 %on \numprint{1000}~randomly selected instances from the PyEx benchmarks. 
-The results are shown in \cref{tab:solve-status-smt-comp}.
 %
@@
 
+We evaluate the different solvers in \cref{tab:solve-status-smt-comp}
+on 21\,938~benchmarks from the single-query QF\_SLIA track at SMT-COMP 2023, as
+QF\_SLIA was the most challenging track for \Ostrich{} in the
+competition.
+%
 As can be seen from the table, integrating \Catra{} as a back-end
-leads to gains both on satisfiable and unsatisfiable problems. The
-increase is more relevant for the unsatisfiable problems, given that
-\Ostrich{} is a state-of-the-art solver in particular on such benchmarks.
+leads to gains both on satisfiable and unsatisfiable problems. On
+satisfiable benchmarks, the combination \Ostrich+CA is still
+outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
+\Ostrich+CA narrowly beats the other solvers, which is promising given
+that also \Ostrich{}~1.3, cvc5, and z3alpha show very strong
+performance on this class of benchmarks.
 

### 2023-10-21 evaluation.tex
Relabel graphs etc

@@
 
-The most obvious threats to validity would be poor benchmarking or poor
-implementation, e.g. if the method described in \cref{sec:calculus,sec:multiple}
-deviate from what is actually benchmarked in section \cref{sec:experiments}, or
-if the methods used for benchmarking were unsound. To increase the probability
-that our results are representative both in the sense of representing expected
-inputs and, in the statistical sense, for a given run of \Catra{} despite the
-use of randomness in our implementation, we execute many experiments. Moreover,
-to address the issue of correctness we have validated all reported solutions
-made by \Calculus{} with \Nuxmv{} to ensure that \Calculus{} is indeed sound. A
-previous version contained a race condition triggered by random restarts during
-product materialisation which would in about \numprint{0.7} of instances lead to
-non-deterministically unsound behaviour. We have observed no instances of this
-bug or any other soundness issue since fixing that bug, and in addition to
-running all \NrBenchmarks{}, we have re-executed the same instances multiple
-times for long durations to capture other soundness issues without any issues.
+The most obvious threats to validity would be an unsound implementation. To
+address this we have validated all reported solutions made by \Calculus{} with
+\Nuxmv{}. A previous version contained a race with random restarts during
+product materialisation causing non-deterministic unsoundness in \numprint{0.7}
+of instances. Since addressing that bug we have observed no other soundness
+issues, despite re-executing the same instances multiple times for long
+durations.
 
-Another threat to validity would be if the competition (baseline and \Nuxmv)
-would be disadvantaged in our comparison. For \Nuxmv{}, we use the default
-configuration, which we believe should be performant (or it should not be the
-default). Additionally, tweaking our invocation of \Nuxmv{} is explicitly made
-easy for artefact reviewers. For baseline, our best argument is that the
-solution we use is close to the one found in the \Ostrich{} string solver and
-that it, therefore, should be realistic.
-
-The greatest threat to validity remaining is our choice of implementation
-platform and automata library. There are some signs that product computation
-is inefficient, notably the good performance of low-threshold automata
-materialisation that prioritises computing smaller products. This means that
-\Calculus{} makes less heavy use of product computation than baseline does,
-instead relying more on \Princess{}. This situation would unfairly advantage
-\Calculus{} if our automata library was the bottleneck due to our automata
-library being unnaturally poor. We believe this is unlikely since similar
-performance issues have been reported for string solvers. Additionally,
-profiling suggests that both \Princess{}-based backends spend most of their time
-in \Princess{}, suggesting that the automaton implementation is not the
-bottleneck. Finally, we have compared aggregate results for both native backends
-when optimising the automata construction procedures and observed that the trend
-in performance differences has been robust as the automata library improved
-vastly.
+The second threat to validity is our implementation of automata operations. As
+\Calculus{} by design offloads some of the product computation work onto
+\Princess{}, the baseline could be unfairly disadvantaged by a slow automata
+product implementation. We believe this is not an issue since similar
+performance issues with the baseline approach have been reported for string
+solvers. Additionally, profiling shows that baseline spends most of its time in
+\Princess{}, suggesting that the automaton implementation is not the bottleneck.
+Finally, the difference in performance between \Calculus{} and baseline has been
+robust under significant optimisation of the automata library.
 

### 2023-10-21 evaluation.tex
Banish the word lazy

@@
     \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances.
-Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
+Baseline could seemingly only solve unsatisfiable instances, PC* could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than PC*.}
   \end{subfigure}
@@
     \label{fig:runtime-boxplot}
-    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline.
-The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy.
-On the other hand, lazy has a lot of outliers.}
+    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the PC* backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline.
+The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for PC*.
+On the other hand, PC* has a lot of outliers.}
   \end{subfigure}
@@
     \label{fig:cactus}
-    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
+    \Description[A cactus plot comparing the performance of nuxmv, PC* and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that.
 Lazy solves much more instances than nuxmv.}

### 2023-10-21 evaluation.tex
Remove dangling references, add fig 6 description

@@
   \end{subfigure}
+  \caption{Evaluation of \Calculus{} solving Parikh automata problems generated by \Ostrich{}.}
 \end{figure}
@@
 \subsubsection{Scalability}\label{sec:scaling}%
-To get an idea of how the different backends scale, we additionally execute
-\numprint{2000} randomly sampled (without replacement) benchmarks with a
-\numprint{120}-second timeout. In addition to the previous back-ends, we also
-execute \Calculus{} with clause learning and restarts disabled, and with only
-clause learning disabled to show the impact of \cref{sec:random-restarts}, and
-\cref{sec:clause-learning} respectively.
+To investigate scaling we additionally execute \numprint{2000} randomly sampled
+(without replacement) benchmarks with a \numprint{120}-second timeout. In
+addition to the previous back-ends, we also execute \Calculus{} with clause
+learning and restarts disabled, and with only clause learning disabled to show
+the impact of \cref{sec:random-restarts}, and \cref{sec:clause-learning}
+respectively.
 

### 2023-10-21 evaluation.tex
Text/tightening

@@
 The benchmarks are executed in parallel across a cluster of identical
-machines, \BenchmarkRig{}, but as the sole job for each machine.  We compiled
+machines, \BenchmarkRig{}, as the sole job for each machine.  We compiled
 the code using Scala~\ScalaVersion{} and executed the experiments
@@
 type is also available in \cref{tab:solve-status}. Note that many instances lack
-a corum ground truth as they are solved by only one backend. We see that
+a ground truth as they are solved by only one backend. We see that
 \Calculus{} generally outperforms \Nuxmv{} on determining unsatisfiability, as
@@
 Baseline performs worse on satisfiable instances because it executes a heuristic
-meant to detect unsatisfiability early at a heavy penalty to satisfiable
-instances roughly similar to \cite{approximate-parikh}. The heuristic is enabled
-since the improvement is significant compared to the extra cost and since we
-have observed more unsatisfiable instances than satisfiable ones.
+meant to detect unsatisfiability early, similar to~\cite{approximate-parikh}.
+The heuristic is enabled since the improvement is significant compared to the
+extra cost.
 
@@
     \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
-    \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
+    \caption{The distribution of runtimes for solved instances per backend. Total numbers of solved instances differ.}
     \label{fig:runtime-boxplot}
@@
 To investigate scaling we additionally execute \numprint{2000} randomly sampled
-(without replacement) benchmarks with a \numprint{120}-second timeout. In
-addition to the previous back-ends, we also execute \Calculus{} with clause
-learning and restarts disabled, and with only clause learning disabled to show
-the impact of \cref{sec:random-restarts}, and \cref{sec:clause-learning}
-respectively.
+(without replacement) benchmarks with a \numprint{120}-second timeout. We add
+runs of \Calculus{} with clause learning and restarts disabled, and with only
+clause learning disabled to show the impact of \cref{sec:random-restarts}, and
+\cref{sec:clause-learning} respectively.
 
@@
 be seen in \cref{fig:cactus}. We see that \Calculus{} outperforms \Nuxmv{} in
-general but that \Nuxmv{} might scale better on very long runtimes.
-Non-systematic observations suggest that \Nuxmv{}'s scaling benefits kicks in
-faster on newer machines, so it is likely that the amount of CPU cycles spent is
-the underlying factor. This is likely due to two factors. First, \Nuxmv{} is
-more mature than \Calculus{}, and it makes sense that its more general model
-checking methods would pay off on more difficult instances where more
-problem-specific methods offer less payoff. Second, for longer running
-calculations clause learning as described in \cref{sec:clause-learning} can be
-assumed to matter more. However, the clauses we learn are still unpolished, and
-in particular, generalise poorly across product computations. Improvements are
-left as future work.
+general but that \Nuxmv{} might scale better on very long runtimes. This is
+likely due to two factors. First, \Nuxmv{} is more mature than \Calculus{}, and
+its more general model checking methods might pay off on more difficult
+instances compared to problem-specific methods. Second, for longer-running
+calculations clause learning as described in \cref{sec:clause-learning} might
+matter more. As clause learning in \Catra{} is rudimentary and generalises
+poorly across product computations, there is a performance hit.
 
@@
   \end{center}
-  \caption{Number of solved benchmarks from the single-query QF\_SLIA
-    track of quantifier-free string linear integer arithmetic at
-    SMT-COMP 2023. The numbers for each solver are from the publicly
-    available data files, except for CA-Str, which is executed by us
-    (600s timeout, AMD Opteron 2220
-    SE), and \Ostrich+CA, which is a virtual portfolio. }
+  \caption{Number of solved benchmarks in the set of quantifier-free string
+    linear integer arithmetic (QF\_SLIA) at SMT-COMP 2023. The numbers are from
+    the competition results, except for CA-Str, which is executed by us (600s
+    timeout, AMD Opteron 2220 SE), and \Ostrich+CA, a virtual portfolio. }
   \label{tab:solve-status-smt-comp}
@@
 
-To evaluate the effectiveness of \Calculus{} in an actual string solver,
-\Catra{} was experimentally integrated as the Parikh automata product solver
-into the string solver \Ostrich{}~version~1.3. \Ostrich{} is an independently
-developed solver that participated in the recent SMT-COMP~2023\footnote{\url{https://smt-comp.github.io/2023/participants/ostrich}},
- winning the single-query track
-for quantifier-free strings (\texttt{QF\_S}), as well as dominating
-other solvers on the complete set of
- unsatisfiable string benchmarks~\cite{smt-comp-23}.
+To evaluate the effectiveness of \Calculus{} in a string solver, \Catra{} was
+experimentally integrated as the Parikh automata product solver into the string
+solver \Ostrich{}~version~1.3. \Ostrich{} is an independently developed solver
+that participated in the recent
+SMT-COMP~2023\footnote{\url{https://smt-comp.github.io/2023/participants/ostrich}},
+winning the single-query track for quantifier-free strings (\texttt{QF\_S}), as
+well as dominating other solvers on unsatisfiable string
+benchmarks~\cite{smt-comp-23}.
 
@@
   algorithm~\cite{ostrich-plus}, using Parikh automata and the
-  encoding from~\cite{generate-parikh-image}.
+  encoding from~\cite{generate-parikh-image} (referred to in this paper as baseline).
 \end{itemize}
@@
 
-The results show that \Calculus{} can be used to enhance the
-performance of an automaton-based string solver.
-The results should be considered as preliminary, however, 
-as we believe that a deeper integration of our \Catra{} solver into
-string solvers can lead to significant performance gains.
+The results show that \Calculus{} can be used to enhance the performance of an
+automaton-based string solver. The results should be considered preliminary,
+however, as we believe that a deeper integration of \Catra{} into string solvers
+can lead to significant performance gains.
 
@@
 
-The most obvious threats to validity would be an unsound implementation. To
+The most obvious threat to validity would be an unsound implementation. To
 address this we have validated all reported solutions made by \Calculus{} with
@@
 product materialisation causing non-deterministic unsoundness in \numprint{0.7}
-of instances. Since addressing that bug we have observed no other soundness
-issues, despite re-executing the same instances multiple times for long
-durations.
+of instances. Since addressing that bug we have observed no further soundness
+issues.
 

### 2023-10-21 calculus.tex
corrected flow
Finally 23 pages

@@
 The rule~$\Split{}$ allows us to branch the proof tree by trying to exclude a
-contested transition from a potential solution before concluding that it must be
-included. Intuitively, this is what guarantees our ability to make forward
-progress by eliminating paths through~$\Automaton$.
+transition from a potential solution before concluding that it must be included.
+Intuitively, this is what guarantees our ability to make forward progress by
+eliminating paths through~$\Automaton$.
 
@@
 of $1$ and $0$ coefficients, i.e. $\left[ l_{A+}, i_+ \right] = \left[1, 1,
-0\right]$. Initially, we let $\Filter$ map to fresh variables to obtain, after some simplifications:
+0\right]$.  The reader is
+advised to review \cref{fig:aut_a} from
+\cref{sec:intuition} while going through this example.
+
+Initially, we let $\Filter$ map to fresh variables to obtain, after some simplifications:
 \begin{equation}\label{eq:single-filter}
@@
   \multicolumn{6}{c}{
-    \FlowEq(\mathcal{A}', \Filter)  = \begin{aligned}
-                       1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_4 + \TransitionVar_3\\
-                    \TransitionVar_4 &= \TransitionVar_1 \\
+    \FlowEq(\mathcal{A}', \Filter)  =\quad \begin{aligned}
+                       1 + \cancel{\TransitionVar_5} &= \cancel{\TransitionVar_5} + \TransitionVar_4 + \TransitionVar_3 \wedge\mbox{}\\
+                    \TransitionVar_4 &= \TransitionVar_1 \wedge\mbox{}\\
                     \TransitionVar_3 + \TransitionVar_1 + \cancel{\TransitionVar_2} &= \cancel{\TransitionVar_2} + 1  \\

### 2023-10-21 implementation.tex
Remove dangling references, add fig 6 description
Finally 23 pages

@@
 \Catra{} uses symbolic labels for automata. A symbolic label is defined as a
-finite range of Unicode code points. This allows succinct representation of many
-regular expression patterns such as \lstinline{(a-z).*}, which would have
-otherwise required $27$~transitions. The transition for the range would be
-written \lstinline{a -> b [97, 122]}.
+finite range of Unicode code points. This allows representing regular expression
+patterns like \lstinline{(a-z)} as \lstinline{a -> b [a, z]} when it would have
+otherwise required $27$~non-symbolic transitions.
 
@@
 constraints expressed by the input file, reporting \Sat{} with register
-assignments or \Unsat{} much like traditional~SAT- or SMT solvers would.
-Additionally, baseline and \Calculus{} also support generating the Presburger
-formula describing the constraints of the input file, i.e., computing
-a closed-from representation of the complete Parikh image.
+assignments or \Unsat{}. Additionally, baseline and \Calculus{} also support
+generating the Presburger formula describing the constraints of the input file,
+i.e., computing a closed-form representation of the complete Parikh image.
 % Baseline uses standard
@@
 
-Since \Princess{} does not support the multiple-arity predicates used
-in \Calculus{}, we have implemented variable-length arguments using
-additional helper predicates. These are $\Unused{}(\Automaton)$, which
-marks an automaton as unused in any product, and
-$\TransitionMask{}(\Automaton, \Transition, \Filter(\Transition))$
-which associates a transition $\Transition$ and automaton $\Automaton$
-with its corresponding transition
-variable~$\Filter(\Transition)$. Additionally, we associate each
-predicate with an instance variable to tie helper predicates to the
-$\Image$ predicate they originated from.
+% Since \Princess{} does not support the multiple-arity predicates used
+% in \Calculus{}, we have implemented variable-length arguments using
+% additional helper predicates. These are $\Unused{}(\Automaton)$, which
+% marks an automaton as unused in any product, and
+% $\TransitionMask{}(\Automaton, \Transition, \Filter(\Transition))$
+% which associates a transition $\Transition$ and automaton $\Automaton$
+% with its corresponding transition
+% variable~$\Filter(\Transition)$. Additionally, we associate each
+% predicate with an instance variable to tie helper predicates to the
+% $\Image$ predicate they originated from.
 
@@
 
-  $q \gets \NewPriorityQueue{}$
+  $q \gets \NewPriorityQueue{$\mathcal{A}_i, \ldots, \mathcal{A}_n$}$
 
@@
 been experimentally shown to increase the performance in aggregate (though not
-strictly). We do not currently implement all the proposed features of
-\cref{sec:ext:backjumping}, but we do implement forward-reachability cut
-learning. No sophisticated clause learning for products has been implemented.
+strictly). We currently only implement minimal clause learning based on
+forward-reachability cuts. No sophisticated clause learning for products has
+been implemented.
 

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/ECMARegexParser.scala
Move stuff out

@@ -1,727 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020-2022 Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ap.parser._
-import ap.basetypes.IdealInt
-
-import ecma2020regex._
-import ecma2020regex.Absyn._
-import ecma2020regex.Absyn.{Quantifier => ECMAQuantifier}
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/Exploration.scala

@@ -1,894 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2023 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ostrich.automata.{Automaton, BricsAutomaton, AutomataUtils}
-import ostrich.preop.PreOp
-
-import ap.SimpleAPI
-import SimpleAPI.ProverStatus
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OFlags.scala

@@ -1,71 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2023 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-object OFlags {
-
-  object LengthOptions extends Enumeration {
-    val Off, On, Auto = Value
-  }
-
-  object CEABackend extends Enumeration {
-    val Baseline, Unary, Catra = Value
-  }
-
-  /**
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OstrichEqualityPropagator.scala

@@ -1,122 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020-2021 Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ap.basetypes.IdealInt
-import ap.proof.theoryPlugins.Plugin
-import ap.proof.goal.Goal
-import ap.terfor.{Term, TerForConvenience}
-import ap.terfor.linearcombination.LinearCombination
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OstrichInternalPreprocessor.scala

@@ -1,191 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ap.basetypes.IdealInt
-import ap.theories.Theory
-import ap.terfor.{TermOrder, TerForConvenience}
-import ap.terfor.conjunctions.{Conjunction, ReduceWithConjunction}
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OstrichMain.scala

@@ -1,131 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020-2023 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ap.ParallelFileProver
-import ap.parameters.Param
-
-/**
- * Wrapper around <code>ap.CmdlMain</code>, adding the option
- * <code>-stringSolver=ostrich.OstrichStringTheory</code>.
- */
-object OstrichMain {
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OstrichPreprocessor.scala

@@ -1,462 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2019-2023 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ap.basetypes.IdealInt
-import ap.parser._
-import ap.theories.strings.StringTheory
-
-/**
- * Pre-processor for reducing some operators to more basic ones.
- */
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OstrichReducer.scala

@@ -1,527 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2021-2022 Riccardo de Masellis, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ostrich.automata.{AutDatabase, BricsAutomaton}
-import ap.basetypes.IdealInt
-import ap.parser.{IBoolLit, IFunApp, IIntLit, IExpression}
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OstrichSolver.scala

@@ -1,346 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ostrich.automata.{Automaton, BricsAutomaton}
-import ostrich.preop.{PreOp, ConcatPreOp}
-
-import ap.SimpleAPI
-import ap.parser.IFunction
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OstrichStringFunctionTranslator.scala

@@ -1,223 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2023 Matthew Hague, Philipp Ruemmer. All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- *
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- *
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ostrich.automata.{Regex2PFA, JavascriptPrioAutomatonBuilder,
-                         AtomicStateAutomaton, BricsTransducer}
-import ostrich.preop._
-
-import ap.parser.ITerm
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OstrichStringTheory.scala

@@ -1,479 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2023 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ostrich.automata.{AutDatabase, Transducer}
-import ostrich.preop.{PreOp, TransducerPreOp, ReversePreOp}
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/OstrichStringTheoryBuilder.scala

@@ -1,120 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2019-2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ostrich.automata.TransducerTranslator
-
-import ap.theories.strings.{StringTheory, StringTheoryBuilder, SeqStringTheory}
-import ap.util.CmdlParser
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/StrDatabase.scala

@@ -1,262 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020-2022 Riccardo de Masellis, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ostrich.automata.Regex2Aut
-
-import ap.basetypes.IdealInt
-import ap.parser.{IFunApp, IIntLit, ITerm, IExpression}
-import ap.theories.strings.StringTheory
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/UnicodeData.scala

@@ -1,269 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2021-2022 Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich
-
-import ap.util.Seqs
-
-import scala.io.Source
-import scala.collection.immutable.VectorBuilder
-import scala.collection.mutable.{HashMap => MHashMap, ArrayBuffer,
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/AnchoredLabel.scala

@@ -1,180 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2021 Zhilei Han. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-object AnchoredLabels {
-  // an anchored label is normal label conjoined
-  // with ^ and $
-  abstract class AnchoredLabel
-
-  // normal label
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/AutDatabase.scala

@@ -1,249 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020-2021 Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import ostrich.OstrichStringTheory
-
-import ap.parser._
-
-import scala.collection.mutable.{HashMap => MHashMap}
-
-object AutDatabase {
-
-  abstract sealed class NamedAutomaton (val id : Int) {
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/AutomataAdapters.scala

@@ -1,310 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import scala.collection.mutable.{HashMap => MHashMap, ArrayStack,
-                                 HashSet => MHashSet, MultiMap,
-                                 LinkedHashSet => MLinkedHashSet,
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/AutomataUtils.scala

@@ -1,609 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2021 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import ostrich.UnicodeData
-
-import scala.collection.mutable.{HashSet => MHashSet, ArrayStack,
-                                 Stack => MStack, HashMap => MHashMap,
-                                 MultiMap, ArrayBuffer, Set => MSet}
-
-/**
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/Automaton.scala

@@ -1,538 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import ap.basetypes.IdealInt
-import ap.PresburgerTools
-import ap.terfor.{Formula, Term, TerForConvenience, TermOrder, OneTerm}
-import ap.terfor.linearcombination.LinearCombination
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/BricsAutomaton.scala

@@ -1,628 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2021 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import ostrich.OFlags
-import dk.brics.automaton.{BasicAutomata, BasicOperations, RegExp, Transition, Automaton => BAutomaton, State => BState}
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/BricsTransducer.scala

@@ -1,904 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import scala.collection.mutable.{HashSet => MHashSet,
-                                 LinkedHashSet => MLinkedHashSet,
-                                 HashMap => MHashMap,
-                                 Stack => MStack,
-                                 MultiMap => MMultiMap,
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/PrioAutomaton.scala

@@ -1,1035 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020-2022 Zhilei Han. All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- *
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- *
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import ostrich.{OstrichStringTheory, ECMARegexParser}
-import StreamingTransducer._
-import AnchoredLabels._
-
-import ap.basetypes.IdealInt
-import ap.parser._
-import ap.theories.strings.StringTheory
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/PrioStreamingTransducer.scala

@@ -1,791 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020-2021 Zhilei Han. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import StreamingTransducer._
-import AnchoredLabels._
-
-import scala.collection.mutable.{HashSet => MHashSet,
-                                 LinkedHashSet => MLinkedHashSet,
-                                 HashMap => MHashMap,
-                                 Stack => MStack,
-                                 MultiMap => MMultiMap,
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/Regex2Aut.scala

@@ -1,565 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import ostrich.{OstrichStringTheory, ECMARegexParser}
-
-import ap.basetypes.IdealInt
-import ap.parser._
-import ap.theories.strings.StringTheory
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/StreamingTransducer.scala

@@ -1,152 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020 Zhilei Han. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-object StreamingTransducer {
-  // operators which update the values of string variables in PSST
-  abstract class UpdateOp
-
-  // Append a constant character
-  case class Constant(val c: Char) extends UpdateOp
-
-  // Append a character by adding offset to the input character
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/automata/TransducerTranslator.scala

@@ -1,377 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2021 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.automata
-
-import ap.parser._
-import IExpression._
-import ap.SimpleAPI
-import ap.basetypes.IdealInt
-import SimpleAPI.ProverStatus
-import ap.theories.ModuloArithmetic
-import ap.theories.strings.{StringTheory, StringTheoryBuilder}
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/CEMain.scala

@@ -1,55 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver
-
-import ap.ParallelFileProver
-import ap.parameters.Param
-
-/**
- * Wrapper around <code>ap.CmdlMain</code>, adding the option
- * <code>-stringSolver=ostrich.OstrichStringTheory</code>.
- */
-object CEMain {
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/automata/BricsAutomatonWrapper.scala

@@ -1,95 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.automata
-
-import dk.brics.automaton.{
-  BasicAutomata,
-  RegExp,
-  Automaton => BAutomaton
-}
-
-import scala.collection.JavaConverters.asScala
-
-object BricsAutomatonWrapper {
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/automata/CEAutDatabase.scala

@@ -1,66 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.automata
-
-import ostrich.OstrichStringTheory
-import ostrich.automata.AutDatabase
-import scala.collection.mutable.{HashMap => MHashMap}
-import ap.parser.ITerm
-import ostrich.automata.Automaton
-import ostrich.automata.AutomataUtils
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/automata/CEBasicOperations.scala

@@ -1,701 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.automata
-
-import dk.brics.automaton.{BasicOperations, State, Automaton => BAutomaton}
-import scala.collection.mutable.{
-  HashMap => MHashMap,
-  HashSet => MHashSet,
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/automata/CETransducer.scala

@@ -1,602 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.automata
-import scala.collection.mutable.{
-  HashMap => MHashMap,
-  HashSet => MHashSet,
-  Stack => MStack,
-  MultiMap => MMultiMap,
-  Set => MSet,
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/automata/CostEnrichedAutomaton.scala

@@ -1,36 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.automata
-
-class CostEnrichedAutomaton extends CostEnrichedAutomatonBase{
-}

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/automata/CostEnrichedAutomatonAdapter.scala

@@ -1,182 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.automata
-
-import scala.collection.mutable.{
-  HashMap => MHashMap,
-  ArrayStack,
-  HashSet => MHashSet,
-  LinkedHashSet => MLinkedHashSet
-}
-
-import ostrich.cesolver.util.TermGenerator
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/automata/CostEnrichedAutomatonBase.scala

@@ -1,414 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.automata
-
-import ostrich.automata._
-
-import scala.collection.mutable.{
-  HashMap => MHashMap,
-  Stack => MStack,
-  HashSet => MHashSet,
-  ArrayBuffer
-}
-import scala.collection.immutable.Map
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/automata/Regex2CEAut.scala

@@ -1,172 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.automata
-
-import ostrich.automata.Regex2Aut
-import ap.parser._
-import dk.brics.automaton.BasicAutomata
-import ap.basetypes.IdealInt
-import CEBasicOperations._
-import scala.collection.mutable.ArrayStack
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/convenience/CostEnrichedConvenience.scala

@@ -1,64 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.convenience
-
-import ostrich.cesolver.automata.BricsAutomatonWrapper
-import ostrich.automata.Automaton
-import ap.parser.Internal2InputAbsy
-import ap.parser.ITerm
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/Model.scala

@@ -1,73 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-
-import ap.basetypes.IdealInt
-import ap.parser.ITerm
-
-object Model {
-  sealed trait Value 
-  case class IntValue(i: IdealInt) extends Value
-  case class StringValue(str: Seq[Int]) extends Value
-}
-import Model._
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/ParikhExploration.scala

@@ -1,530 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2023 Denghang Hu, Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-import ostrich.preop.PreOp
-import ostrich.automata.Automaton
-import ostrich.StrDatabase
-import ap.SimpleAPI
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/ParikhStore.scala

@@ -1,181 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-import ostrich.automata.Automaton
-import scala.collection.mutable.{ArrayBuffer, ArrayStack, HashMap => MHashMap}
-import ostrich.cesolver.convenience.CostEnrichedConvenience._
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/finalConstraints/BaselineFinalConstraints.scala

@@ -1,58 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-import scala.collection.mutable.{HashMap => MHashMap}
-import ostrich.cesolver.util.ParikhUtil
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
-import ap.parser.ITerm
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/finalConstraints/CatraFinalConstraints.scala

@@ -1,63 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-import ap.terfor.Formula
-import scala.collection.mutable.{HashMap => MHashMap}
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
-import ap.parser.ITerm
-import ap.parser.IExpression._
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/finalConstraints/FinalConstraints.scala

@@ -1,268 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-
-import ap.api.PartialModel
-import ap.basetypes.IdealInt
-import ap.terfor.ConstantTerm
-import ap.terfor.OneTerm
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/finalConstraints/UnaryFinalConstraints.scala

@@ -1,169 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-import scala.collection.mutable.{HashMap => MHashMap}
-import ostrich.cesolver.util.ParikhUtil
-import ostrich.cesolver.automata.CEBasicOperations
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/finalConstraintsSolver/BaselineSolver.scala

@@ -1,97 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-import ap.api.SimpleAPI
-import ap.api.SimpleAPI.ProverStatus
-import ap.parser.SymbolCollector
-import ap.parser.Internal2InputAbsy._
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/finalConstraintsSolver/CatraBasedSolver.scala

@@ -1,329 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-import ap.api.SimpleAPI.ProverStatus
-import ap.parser.{SymbolCollector, PrincessLineariser}
-import ap.terfor.TerForConvenience._
-import ap.terfor.Term
-import ap.basetypes.IdealInt
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/finalConstraintsSolver/FinalConstraintsSolver.scala

@@ -1,89 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-import ap.api.SimpleAPI.ProverStatus
-import ap.basetypes.IdealInt
-import ostrich.cesolver.util.ParikhUtil.measure
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
-import ap.parser.ITerm
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/core/finalConstraintsSolver/UnaryBasedSolver.scala

@@ -1,136 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.core
-
-import ap.api.SimpleAPI
-import ap.api.SimpleAPI.ProverStatus
-import ap.parser.SymbolCollector
-import ostrich.cesolver.convenience.CostEnrichedConvenience._
-import FinalConstraints._
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preop/CEPreOp.scala

@@ -1,37 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preop
-
-import ostrich.preop.PreOp
-
-trait CEPreOp extends PreOp {}

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preop/CETransducerPreOp.scala

@@ -1,64 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preop
-
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
-import ostrich.cesolver.automata.CETransducer
-import ostrich.automata.Automaton
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preop/ConcatCEPreOp.scala

@@ -1,142 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preop
-
-import ostrich.automata.Automaton
-import ostrich.cesolver.convenience.CostEnrichedConvenience._
-import ostrich.cesolver.automata.CostEnrichedInitFinalAutomaton
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preop/IndexOfCEPreOp.scala

@@ -1,237 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preop
-
-import ostrich.automata.Automaton
-import ap.terfor.linearcombination.LinearCombination
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preop/LengthCEPreOp.scala

@@ -1,89 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preop
-
-import ostrich.automata.Automaton
-import ostrich.cesolver.automata.CostEnrichedAutomaton
-import ostrich.cesolver._
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
-import ap.parser.ITerm
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preop/PreOpUtil.scala

@@ -1,65 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preop
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
-import ostrich.cesolver.automata.BricsAutomatonWrapper.{makeEmpty, makeEmptyString}
-
-object PreOpUtil {
-    def automatonWithLen(len: Int): CostEnrichedAutomatonBase = {
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preop/ReplaceAllCEPreOp.scala

@@ -1,276 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preop
-
-import scala.collection.mutable.{HashSet => MHashSet, HashMap => MHashMap, Stack=> MStack}
-
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
-import ostrich.cesolver.automata.CETransducer
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preop/ReplaceCEPreOp.scala

@@ -1,268 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preop
-
-import scala.collection.mutable.{
-  HashMap => MHashMap,
-  HashSet => MHashSet,
-  Stack => MStack
-}
-
-import ostrich.automata.Automaton
-import ostrich.automata.Transducer
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preop/SubStringCEPreOp.scala

@@ -1,182 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preop
-
-import ostrich.automata.Automaton
-import ostrich.cesolver.automata.CostEnrichedAutomaton
-import ostrich.cesolver.automata.BricsAutomatonWrapper
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/preprocess/CEPreprocessor.scala

@@ -1,156 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu, Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.preprocess
-
-import ap.parser._
-import IExpression._
-import ap.theories.strings.StringTheory.ConcreteString
-import ostrich.cesolver.stringtheory.CEStringTheory
-import ostrich.cesolver.util.ParikhUtil
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/stringtheory/CESolver.scala

@@ -1,444 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu, Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.stringtheory
-
-import ostrich.automata.{Automaton, BricsAutomaton}
-import ostrich.preop.{PreOp, ConcatPreOp}
-
-import ap.SimpleAPI
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/stringtheory/CEStringFunctionTranslator.scala

@@ -1,161 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2023 Denghang Hu, Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.stringtheory
-
-import ostrich.automata.{Regex2PFA, JavascriptPrioAutomatonBuilder,
-                         AtomicStateAutomaton, BricsTransducer}
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/stringtheory/CEStringTheoryBuilder.scala

@@ -1,107 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2019-2023 Denghang Hu, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.stringtheory
-
-import ostrich._
-import ostrich.automata.TransducerTranslator
-
-import ap.theories.strings.{StringTheory, StringTheoryBuilder, SeqStringTheory}
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/util/Exception.scala

@@ -1,37 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.util
-
-case class UnknownException(val info: String) extends Exception
-
-case class TimeoutException(val time: Long) extends Exception

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/util/ParikhUtil.scala

@@ -1,171 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.util
-
-import scala.collection.mutable.{
-  Map => MMap,
-  HashSet => MHashSet,
-  HashMap => MHashMap,
-  ArrayStack
-}
-import ostrich.cesolver.automata.CostEnrichedAutomatonBase
-
-import ap.basetypes.IdealInt
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/cesolver/util/Writer.scala

@@ -1,84 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2023 Denghang Hu. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.cesolver.util
-
-import java.io.{File, BufferedWriter, FileWriter}
-
-trait Writer {
-
-  val filename: String
-
-  val file: File
-
-  val writer: BufferedWriter
-
-  def write(s: String) = writer.write(s)
-
-  def closeAfterWrite(s: String) = {
-    write(s)
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/ConcatPreOp.scala

@@ -1,143 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{Automaton, AtomicStateAutomaton, InitFinalAutomaton,
-                         ProductAutomaton, ConcatAutomaton}
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/ExtractPreOp.scala

@@ -1,176 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020 Zhilei Han. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{Regex2PFA, PrioStreamingTransducer, PrioAutomaton,
-                         TLabelOps, AnchoredLabels, StreamingTransducer}
-import StreamingTransducer._
-import AnchoredLabels._
-
-import scala.collection.mutable.{HashSet => MHashSet,
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/NOPPreOp.scala

@@ -1,66 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018 Matthew Hague, Philipp Ruemmer. All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- *
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- *
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{Automaton, AtomicStateAutomaton, ProductAutomaton}
-
-/**
- * Pre-Op for the No-op function.
- */
-object NOPPreOp extends PreOp {
-  override def toString = "nop"
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/PreOp.scala

@@ -1,104 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{Automaton, BricsAutomaton}
-
-import ap.terfor.{Term, Formula, TermOrder}
-import ap.terfor.conjunctions.Conjunction
-
-/**
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/ReplaceAllCGPreOp.scala

@@ -1,194 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020 Zhilei Han. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{Regex2PFA, PrioStreamingTransducer, PrioAutomaton,
-                         StreamingTransducer, TLabelOps, AnchoredLabels}
-import StreamingTransducer._
-import AnchoredLabels._
-
-import scala.collection.mutable.{HashSet => MHashSet,
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/ReplaceAllPreOp.scala

@@ -1,593 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018-2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- *
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- *
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{AtomicStateAutomaton, Transducer, Automaton,
-                         CaleyGraph, ReplaceCharAutomaton, InitFinalAutomaton,
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/ReplaceCGPreOp.scala

@@ -1,199 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020 Zhilei Han. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{Regex2PFA, StreamingTransducer,
-                         PrioStreamingTransducer, PrioAutomaton, TLabelOps,
-                         AnchoredLabels}
-import StreamingTransducer._
-import AnchoredLabels._
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/ReversePreOp.scala

@@ -1,70 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{Automaton, AtomicStateAutomaton, ReverseAutomaton,
-                         ProductAutomaton}
-
-/**
- * Pre-image computation for the reverse operator.
- */
-object ReversePreOp extends PreOp {
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/StreamingTransducerPreOp.scala

@@ -1,65 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2020 Zhilei Han. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{StreamingTransducer, Automaton, ProductAutomaton,
-                         AtomicStateAutomaton}
-
-object StreamingTransducerPreOp {
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/preop/TransducerPreOp.scala

@@ -1,73 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2018 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.preop
-
-import ostrich.automata.{Transducer, AtomicStateAutomaton, Automaton,
-                         ProductAutomaton, PostImageAutomaton}
-
-object TransducerPreOp {
-  def apply(t : Transducer) = new TransducerPreOp(t)
-}
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/proofops/FormulaBuilder.scala

@@ -1,141 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.proofops
-
-import ostrich._
-
-import ap.basetypes.IdealInt
-import ap.terfor.{VariableTerm, Formula, Term, TerForConvenience}
-import ap.types.Sort
-
-import scala.collection.mutable.{ArrayBuffer, HashMap => MHashMap}
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/proofops/OstrichNielsenSplitter.scala

@@ -1,607 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2022 Matthew Hague, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.proofops
-
-import ostrich._
-
-import ap.basetypes.IdealInt
-import ap.parameters.Param
-import ap.proof.ModelSearchProver
-import ap.proof.theoryPlugins.Plugin
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/main/scala/ostrich/proofops/OstrichPredtoEqConverter.scala

@@ -1,232 +1,0 @@
-/**
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (c) 2022 Oliver Markgraf, Philipp Ruemmer. All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- * 
- * * Redistributions of source code must retain the above copyright notice, this
- *   list of conditions and the following disclaimer.
- * 
- * * Redistributions in binary form must reproduce the above copyright notice,
- *   this list of conditions and the following disclaimer in the documentation
- *   and/or other materials provided with the distribution.
- * 
- * * Neither the name of the authors nor the names of their
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
- * FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
- * COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
- * INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
- * (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
- * SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
- * STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
- * OF THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-package ostrich.proofops
-
-import ostrich._
-
-import ap.basetypes.IdealInt
-import ap.parameters.Param
-import ap.proof.ModelSearchProver
-import ap.proof.theoryPlugins.Plugin
-import ap.proof.goal.Goal
-import ap.terfor.TerForConvenience.l
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/APITest.scala

@@ -1,122 +1,0 @@
-package ostrich
-
-import ap.SimpleAPI
-import SimpleAPI.ProverStatus
-import ap.parser._
-import ap.theories.strings.SeqStringTheory
-import ap.util.Debug
-
-import org.scalacheck.{Arbitrary, Gen, Properties}
-import org.scalacheck.Prop._
-
-object APITest extends Properties("APITest") {
-
-  Debug enableAllAssertions true
-
-  def expect[A](x : A, expected : A) : A = {
-    assert(x == expected, "expected: " + expected + ", got: " + x)
-    x
-  }
-
-  val stringTheory = new OstrichStringTheory(List(), OFlags())
-  import stringTheory._
-  import IExpression._
-
-  property("concat") =
-    Console.withErr(ap.CmdlMain.NullStream) {
-    SimpleAPI.withProver(enableAssert = true) { p =>
-      import p._
-
-      val x, y, z = createConstant(StringSort)
-      implicit val ctxt = decoderContext
-
-      scope {
-        !! (x === "abc")
-
-        expect(???, ProverStatus.Sat)
-        evalToTerm(x)
-        expect(asString(eval(x)), "abc")
-
-        !! (x ++ y ++ "xyz" === z)
- 
-        expect(???, ProverStatus.Sat)
-        assert(asString(eval(x)) + asString(eval(y)) + "xyz" == asString(eval(z)))
-
-        !! (y =/= "")
-
-        expect(???, ProverStatus.Sat)
-        assert(asString(eval(x)) + asString(eval(y)) + "xyz" == asString(eval(z)))
-
-        scope {
-          !! (y === z)
-          expect(???, ProverStatus.Unsat)
-        }
-
-        ??? == ProverStatus.Sat
-      }
-    }}
-
-  property("head-tail") =
-    Console.withErr(ap.CmdlMain.NullStream) {
-    SimpleAPI.withProver(enableAssert = true) { p =>
-      import p._
-
-      val x, y, z = createConstant(StringSort)
-      implicit val ctxt = decoderContext
-
-      scope {
-        !! (x ++ y === y ++ x)
-
-        // scope {
-        //   !! (str_len(x) === 2)
-        //   !! (str_len(y) === 5)
-        //   !! (x =/= "")
-
-        //   println(???)             // currently this cannot be handled
-        //   println("x = " + evalToTerm(x))
-        //   println("y = " + evalToTerm(y))
-
-        //   scope {
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/DenghangTests.scala

@@ -1,106 +1,0 @@
-package ostrich.cesolver
-
-import ap.CmdlMain
-import ap.DialogUtil.asString
-import org.scalacheck.Properties
-
-object DenghangTests extends Properties("DenghangTests") {
-
-  import System.lineSeparator
-
-  val timeout = 30000
-
-  def expectResult[A](expResult: String)(computation: => A): Boolean = {
-    val result = asString {
-      Console.withErr(ap.CmdlMain.NullStream) {
-        computation
-      }
-    }
-
-    expResult match {
-      case "error" =>
-        result contains "error"
-      case res =>
-        (result split lineSeparator) contains res
-    }
-  }
-
-  def checkFile(
-      filename: String,
-      result: String,
-      extractOpts: String*
-  ): Boolean =
-    expectResult(result) {
-      CEMain.main(
-        (List(
-          "+assert",
-          "-timeout=" + timeout,
-          filename
-        ) ++ extractOpts).toArray
-      )
-    }
-
-
-  // integration tests for length
-  property("length_sat.smt2") =
-    checkFile("tests/hu-benchmarks/length_sat.smt2", "sat")
-
-  // integration tests for concatenate
-  property("concat_sat.smt2") =
-    checkFile("tests/hu-benchmarks/concat_sat.smt2", "sat")
-  property("concat_unsat.smt2") =
-    checkFile("tests/hu-benchmarks/concat_unsat.smt2", "unsat")
-  property("concat_backjump_bug.smt2") = 
-    checkFile("tests/hu-benchmarks/concat_backjump_bug.smt2", "sat")
-
-  // integration tests for indexof
-  property("indexof_const_index_sat.smt2") =
-    checkFile("tests/hu-benchmarks/indexof_const_index_sat.smt2", "sat")
-  property("indexof_const_index_unsat.smt2") =
-    checkFile("tests/hu-benchmarks/indexof_const_index_unsat.smt2", "unsat")
-  property("indexof_const_startpos_sat.smt2") =
-    checkFile("tests/hu-benchmarks/indexof_const_startpos_sat.smt2", "sat")
-  property("indexof_const_startpos_unsat.smt2") =
-    checkFile("tests/hu-benchmarks/indexof_const_startpos_unsat.smt2", "unsat")
-  property("indexof_var_sat.smt2") =
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/SMTLIBTests.scala

@@ -1,442 +1,0 @@
-package ostrich
-
-import ap.CmdlMain
-import ap.DialogUtil.asString
-
-import org.scalacheck.{Arbitrary, Gen, Properties}
-import org.scalacheck.Prop._
-import ostrich.cesolver.util.ParikhUtil
-
-object SMTLIBTests extends Properties("SMTLIBTests") {
-
-  import System.lineSeparator
-
-  val timeout = 30000
-
-  def expectResult[A](expResult : String)(computation : => A) : Boolean = {
-    val result = asString {
-      Console.withErr(ap.CmdlMain.NullStream) {
-        computation
-      }
-    }
-
-    expResult match {
-      case "error" =>
-        result contains "error"
-      case res =>
-        (result split lineSeparator) contains res
-    }
-  }
-
-  def checkFile(filename : String, result : String,
-                extractOpts : String*) : Boolean =
-    expectResult(result) {
-      CmdlMain.doMain((List("+assert", "-timeout=" + timeout,
-                            "-stringSolver=ostrich.OstrichStringTheory",
-                            filename) ++ extractOpts).toArray,
-                        false)
-    }
-
-  def checkFileOpts(filename : String, result : String, ostrichOpts : String,
-                    extractOpts : String*) : Boolean =
-    expectResult(result) {
-      CmdlMain.doMain((List("+assert", "-timeout=" + timeout,
-                            "-stringSolver=ostrich.OstrichStringTheory:" +
-                               ostrichOpts,
-                            filename) ++ extractOpts).toArray,
-                        false)
-    }
-
-  property("prefix-1.smt2") =
-    checkFile("tests/prefix-1.smt2", "unsat")
-  property("prefix-2.smt2") =
-    checkFile("tests/prefix-2.smt2", "sat")
-
-  property("suffix-1.smt2") =
-    checkFile("tests/suffix-1.smt2", "unsat")
-  property("suffix-2.smt2") =
-    checkFile("tests/suffix-2.smt2", "sat")
-  property("suffix-3.smt2") =
-    checkFile("tests/suffix-3.smt2", "sat")
-  property("suffix-4.smt2") =
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/YanTests.scala

@@ -1,119 +1,0 @@
-package ostrich
-
-import ap.CmdlMain
-import ap.DialogUtil.asString
-
-import org.scalacheck.{Arbitrary, Gen, Properties}
-import org.scalacheck.Prop._
-
-object YanTests extends Properties("YanTests") {
-
-  import System.lineSeparator
-  
-  def expectResult[A](expResult : String)(computation : => A) : Boolean = {
-    val result = asString {
-      Console.withErr(ap.CmdlMain.NullStream) {
-        computation
-      }
-    }
-
-    (result split lineSeparator) contains expResult
-  }
-
-  def checkFile(filename : String, result : String,
-                extractOpts : String*) : Boolean =
-    expectResult(result) {
-      CmdlMain.doMain((List("+assert", "-timeout=10000",
-                            "-stringSolver=ostrich.OstrichStringTheory",
-                            filename) ++ extractOpts).toArray,
-                        false)
-    }
-
-  property("tests/yan-benchmarks/concat-001.smt2") =
-    checkFile("tests/yan-benchmarks/concat-001.smt2", "sat")
-  property("tests/yan-benchmarks/concat-002.smt2") =
-    checkFile("tests/yan-benchmarks/concat-002.smt2", "sat")
-  property("tests/yan-benchmarks/concat-003.smt2") =
-    checkFile("tests/yan-benchmarks/concat-003.smt2", "sat")
-  property("tests/yan-benchmarks/concat-004-unsat.smt2") =
-    checkFile("tests/yan-benchmarks/concat-004-unsat.smt2", "unsat")
-  property("tests/yan-benchmarks/concat-005-unsat.smt2") =
-    checkFile("tests/yan-benchmarks/concat-005-unsat.smt2", "unsat")
-  property("tests/yan-benchmarks/concat-006.smt2") =
-    checkFile("tests/yan-benchmarks/concat-006.smt2", "sat")
-  property("tests/yan-benchmarks/concat-007.smt2") =
-    checkFile("tests/yan-benchmarks/concat-007.smt2", "sat")
-  property("tests/yan-benchmarks/concat-008.smt2") =
-    checkFile("tests/yan-benchmarks/concat-008.smt2", "sat")
-  property("tests/yan-benchmarks/concat-009.smt2") =
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/ostrich/automata/BricsPrioTransducerSpecification.scala

@@ -1,91 +1,0 @@
-package ostrich.automata
-
-import scala.collection.mutable.HashMap
-
-import org.scalacheck.{Arbitrary, Gen, Properties}
-import org.scalacheck.Prop._
-
-import dk.brics.automaton.{Automaton => BAutomaton, State, Transition}
-
-object BricsPrioTransducerSpecification
-	extends Properties("BricsPrioTransducer"){
-  import Transducer._
-
-  def seq(s : String) = s.map(_.toInt)
-
-  val simplePrePostTran = {
-    // Transducer q0 -- [a-c], ("zz", +0, "adb") --> qf
-    val builder = BricsPrioTransducer.getBuilder
-    val q0 = builder.getNewState
-    val qf = builder.getNewState
-    builder.setAccept(qf, true)
-
-    val op = OutputOp("zz", Plus(0), "adb")
-    builder.addTransition(q0, ('a', 'c'), op, qf)
-
-    builder.setInitialState(q0)
-
-    builder.getTransducer
-  }
-
-  val aStarPrioTran = {
-    // Transducer encoding the capture group (a*)a*
-    val builder = BricsPrioTransducer.getBuilder
-    val q0 = builder.getNewState
-    val q1 = builder.getNewState
-    val qf = builder.getNewState
-    builder.setAccept(qf, true)
-
-    val op = OutputOp("", Plus(0), "")
-    val nop = OutputOp("", NOP, "")
-    builder.addTransition(q0, ('a', 'a'), op, q1)
-    builder.addTransition(qf, ('a', 'a'), nop, qf)
-    builder.addETransition(q1, nop, 0, qf)
-    builder.addETransition(q1, nop, 1, q0) // this is the prioritised transition
-
-    builder.setInitialState(q0)
-
-    builder.getTransducer
-  }
-
-  property("Simple Pre With Pre and Post") = {
-    // Automaton q1 -[z]-> q2 -[a-z]-> q3 -[b]-> q4 -- [a] --> q2
-    val q1 = new IDState(1)
-    val q2 = new IDState(2)
-    val q3 = new IDState(3)
-    val q4 = new IDState(4)
-    q4.setAccept(true)
-    q1.addTransition(new Transition('z', 'z', q2))
-    q2.addTransition(new Transition('a', 'z', q3))
-    q3.addTransition(new Transition('b', 'b', q4))
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/ostrich/automata/BricsTransducerSpecification.scala

@@ -1,468 +1,0 @@
-package ostrich.automata
-
-import scala.collection.mutable.HashMap
-
-import org.scalacheck.{Arbitrary, Gen, Properties}
-import org.scalacheck.Prop._
-
-import dk.brics.automaton.{Automaton => BAutomaton, State, Transition}
-
-object BricsTransducerSpecification
-	extends Properties("BricsTransducer"){
-  import Transducer._
-
-  def seq(s : String) = s.map(_.toInt)
-
-  val simplePrePostTran = {
-    // Transducer q0 -- [a-c], ("zz", +0, "adb") --> qf
-    val builder = BricsTransducer.getBuilder
-    val q0 = builder.getNewState
-    val qf = builder.getNewState
-    builder.setAccept(qf, true)
-
-    val op = OutputOp("zz", Plus(0), "adb")
-    builder.addTransition(q0, ('a', 'c'), op, qf)
-
-    builder.setInitialState(q0)
-
-    builder.getTransducer
-  }
-
-  val simplePrePostDel = {
-    // Transducer q0 -- [a-c], ("zz", delete, "adb") --> qf
-    val builder = BricsTransducer.getBuilder
-
-    val q0 = builder.getNewState
-    val qf = builder.getNewState
-
-    builder.setAccept(qf, true)
-    val op = OutputOp("zz", NOP, "adb")
-    builder.addTransition(q0, ('a', 'c'), op, qf)
-    builder.setInitialState(q0)
-
-    builder.getTransducer
-  }
-
-  val simplePrePostInternal = {
-    // Transducer q0 -- [a-c], ("zz", internal, "adb") --> qf
-    val builder = BricsTransducer.getBuilder
-
-    val q0 = builder.getNewState
-    val qf = builder.getNewState
-
-    builder.setAccept(qf, true)
-    val op = OutputOp("zz", Internal, "adb")
-    builder.addTransition(q0, ('a', 'c'), op, qf)
-    builder.setInitialState(q0)
-
-    builder.getTransducer
-  }
-
-  val epsilonTrans = {
-    // Transducer q0 -- epsilon, ("zz", delete, "adb") --> q1
-    //            q1 -- a, ("", copy, "") --> qf
-    val builder = BricsTransducer.getBuilder
-
-    val q0 = builder.getNewState
-    val q1 = builder.getNewState
-    val qf = builder.getNewState
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/ostrich/automata/CaleyGraphSpecification.scala

@@ -1,157 +1,0 @@
-package ostrich.automata
-
-import org.scalacheck.{Arbitrary, Gen, Properties}
-import org.scalacheck.Prop._
-import dk.brics.automaton.{Automaton => BAutomaton, State, Transition}
-import scala.collection.mutable.Set
-import scala.collection.JavaConverters.asScala
-
-class PrintableState extends State {
-  override def toString = "q" + hashCode
-}
-
-class IDState(val ident : Int) extends State {
-  override def toString =
-    asScala(getTransitions).foldLeft("q" + ident + '\n') { (s, t) =>
-      val dest = t.getDest match {
-        case d : IDState => "q" + d.ident
-        case q => q.toString
-      }
-      s + "  --[" + t.getMin + ", " + t.getMax + "]--> " + dest + '\n'
-    }
-}
-
-object CaleyGraphSpecification
-	extends Properties("CaleyGraph"){
-
-  val states = Seq.fill(10)(new PrintableState)
-
-  property("Test compose") = {
-
-    val b1 = new Box[BricsAutomaton]()
-    val b2 = new Box[BricsAutomaton]()
-
-    b1.addEdge(states(0), states(1))
-    b2.addEdge(states(1), states(2))
-
-    val box = b1 ++ b2
-
-    box(states(0)) == Set(states(2))
-  }
-
-  implicit def arbBox : Arbitrary[Box[BricsAutomaton]] = Arbitrary {
-    for (
-      items <- Gen.listOf(Gen.listOfN(2, Gen.oneOf(states)))
-    ) yield {
-      // Can't figure out how to use Box(items) due to
-      // PrintableState
-      items.foldLeft(new Box[BricsAutomaton])((box, qq) =>
-        box.addEdge(qq(0), qq(1))
-      )
-    }
-  }
-
-  def caleyGraphHasBoxes(baut : BricsAutomaton,
-               boxes : Set[Box[BricsAutomaton]]) = {
-    val cg = CaleyGraph(baut)
-    cg.getNodes == boxes
-  }
-
-  property("Serious test compose") =
-    forAll { (b1 : Box[BricsAutomaton], b2 : Box[BricsAutomaton]) =>
-      val bb = b1 ++ b2
-      val dir1 = // in bb => in b1 and b2
-        bb.getEdges.forall { case (q1, q3) =>
-          b1(q1).exists(q2 => b2.hasEdge(q2, q3))
-        }
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/ostrich/automata/PrioStreamingTransducerSpecification.scala

@@ -1,261 +1,0 @@
-/*
- * This file is part of Ostrich, an SMT solver for strings.
- * Copyright (C) 2020  Zhilei Han
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <https://www.gnu.org/licenses/>.
- */
-package ostrich.automata
-
-import StreamingTransducer._
-import AnchoredLabels._
-
-import scala.collection.mutable.HashMap
-
-import org.scalacheck.{Arbitrary, Gen, Properties}
-import org.scalacheck.Prop._
-
-import dk.brics.automaton.{Automaton => BAutomaton, State, Transition}
-
-object PrioStreamingTransducerSpecification
-  extends Properties("PrioStreamingTransducer"){
-
-  def seq(s : String) = s.map(_.toInt)
-
-  def toConstants(s : String) : Seq[UpdateOp] = s.map((c) => Constant(c))
-
-  // a trivial PSST which simulate the normal prioritised transducer
-  // q0 -- [a-c], ("zz", +0, "adb") --> qf
-  val simplePrePostTran = {
-    // have a single variable for 'result'
-    val builder = PrioStreamingTransducer.getBuilder(1)
-    val q0 = builder.getNewState
-    val qf = builder.getNewState
-    builder.setAccept(qf, true, List(RefVariable(0)))
-
-    builder.addTransition(q0, NormalLabel(('a', 'c')), List(toConstants("zz") ++ List(Offset(0)) ++ toConstants("adb")), qf)
-
-    builder.setInitialState(q0)
-
-    builder.getTransducer
-  }
-
-  property("Simple Pre With Pre and Post PreImage") = {
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/ostrich/preop/ConcatPreOpSpecification.scala

@@ -1,24 +1,0 @@
-package ostrich.preop
-
-import ostrich.automata.{BricsAutomaton, Automaton}
-
-import org.scalacheck.{Arbitrary, Gen, Properties}
-import org.scalacheck.Prop._
-
-import dk.brics.automaton.{Automaton => BAutomaton, State, Transition}
-
-object ConcatPreOpSpecification
-	extends Properties("ConcatPreOp"){
-
-  def seq(s : String) = s.map(_.toInt)
-
-  property("Simple Post") = {
-    val baut1 = BricsAutomaton.fromString("d")
-    val baut2 = BricsAutomaton.fromString("e")
-
-    val concat = ConcatPreOp.forwardApprox(Seq(Seq(baut1.asInstanceOf[Automaton]),
-                                               Seq(baut2.asInstanceOf[Automaton])))
-
-    concat(seq("de")) && !concat(seq("d"))
-  }
-}

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/ostrich/preop/ReplaceAllPreOpSpecification.scala

@@ -1,323 +1,0 @@
-package ostrich.preop
-
-import ostrich.automata.{BricsAutomaton, Automaton, IDState}
-
-import org.scalacheck.Properties
-import org.scalacheck.Prop._
-
-import dk.brics.automaton.{Automaton => BAutomaton, State, Transition}
-
-object ReplaceAllPreOpSpecification
-	extends Properties("ReplaceAllPreOp"){
-
-  val abcdAut = BricsAutomaton.fromString("abcd")
-  val bcAut = BricsAutomaton.fromString("bc")
-  val aaAut = BricsAutomaton.fromString("aa")
-  val aorbAut = BricsAutomaton.fromString("a") | BricsAutomaton.fromString("b")
-  val aaorbbAut = BricsAutomaton.fromString("aa") | BricsAutomaton.fromString("bb")
-  val abdAut = BricsAutomaton.fromString("abd")
-  val dbaAut = BricsAutomaton.fromString("dba")
-  val dAut = BricsAutomaton.fromString("d")
-  val ddAut = BricsAutomaton.fromString("dd")
-  val eAut = BricsAutomaton.fromString("")
-
-  def seq(s : String) = s.map(_.toInt)
-
-  property("Simple single char test 1") = {
-    // abcd = replaceall(x, e, bc)
-    ReplaceAllLongestPreOp('e')(Seq(Seq(), Seq(bcAut)), abcdAut)._1.exists(cons => {
-      cons(0)(seq("aed"))
-    })
-  }
-
-  property("Simple single char test 2") = {
-    // abcd = replaceall(x, e, bc)
-    ReplaceAllLongestPreOp('e')(Seq(Seq(), Seq(bcAut)), abcdAut)._1.exists(cons => {
-      cons(0)(seq("abcd"))
-    })
-  }
-
-  property("Simple single char test 3") = {
-    // abcd = replaceall(x, e, bc)
-    !ReplaceAllLongestPreOp('e')(Seq(Seq(), Seq(bcAut)), abcdAut)._1.exists(cons => {
-      cons(0)(seq("abd"))
-    })
-  }
-
-  property("Simple word test 1") = {
-    // abcd = replaceall(x, word, bc)
-    ReplaceAllLongestPreOp("word")(Seq(Seq(), Seq(bcAut)), abcdAut)._1.exists(cons => {
-      cons(0)(seq("awordd"))
-    })
-  }
-
-  property("Simple word test 2") = {
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/ostrich/preop/ReplacePreOpSpecification.scala

@@ -1,352 +1,0 @@
-package ostrich.preop
-
-import ostrich.automata.{BricsAutomaton, Automaton, IDState}
-
-import org.scalacheck.Properties
-import org.scalacheck.Prop._
-
-import dk.brics.automaton.{Automaton => BAutomaton, State, Transition}
-
-object ReplacePreOpSpecification
-    extends Properties("ReplacePreOp"){
-
-  val abcdAut = BricsAutomaton.fromString("abcd")
-  val bcAut = BricsAutomaton.fromString("bc")
-  val aaAut = BricsAutomaton.fromString("aa")
-  val aorbAut = BricsAutomaton.fromString("a") | BricsAutomaton.fromString("b")
-  val aaorbbAut = BricsAutomaton.fromString("aa") | BricsAutomaton.fromString("bb")
-  val abdAut = BricsAutomaton.fromString("abd")
-  val cAut = BricsAutomaton.fromString("c")
-  val caaAut = BricsAutomaton.fromString("caa")
-  val dbaAut = BricsAutomaton.fromString("dba")
-  val dAut = BricsAutomaton.fromString("d")
-  val ddAut = BricsAutomaton.fromString("dd")
-  val eAut = BricsAutomaton.fromString("")
-
-  def seq(s : String) = s.map(_.toInt)
-
-  property("Simple single char test 1") = {
-    // abcd = replace(x, e, bc)
-    ReplaceLongestPreOp('e')(Seq(Seq(), Seq(bcAut)), abcdAut)._1.exists(cons => {
-      cons(0)(seq("aed"))
-    })
-  }
-
-  property("Simple single char test 2") = {
-    // abcd = replace(x, e, bc)
-    ReplaceLongestPreOp('e')(Seq(Seq(), Seq(bcAut)), abcdAut)._1.exists(cons => {
-      cons(0)(seq("abcd"))
-    })
-  }
-
-  property("Simple single char test 3") = {
-    // abcd = replace(x, e, bc)
-    !ReplaceLongestPreOp('e')(Seq(Seq(), Seq(bcAut)), abcdAut)._1.exists(cons => {
-      cons(0)(seq("abd"))
-    })
-  }
-
-  property("Simple word test 1") = {
-    // abcd = replace(x, word, bc)
-    ReplaceLongestPreOp("word")(Seq(Seq(), Seq(bcAut)), abcdAut)._1.exists(cons => {
-      cons(0)(seq("awordd"))
-    })
-  }
-
[…]

### 2024-02-15 experiments/ostrich-catra/ostrich/src/test/scala/ostrich/preop/ReversePreOpSpecification.scala

@@ -1,86 +1,0 @@
-
-package ostrich.preop
-
-import ostrich.automata.{IDState, BricsAutomaton}
-
-import org.scalacheck.Properties
-import org.scalacheck.Prop._
-
-import dk.brics.automaton.{Automaton => BAutomaton, State, Transition}
-
-object ReversePreOpSpecification
-	extends Properties("ReversePreOp"){
-
-  // Automaton q1 -[a]-> q2 -[b]-> q3 -[c]-> q4 -[d]-> q5
-  val abcdAut = {
-    val q1 = new IDState(1)
-    val q2 = new IDState(2)
-    val q3 = new IDState(3)
-    val q4 = new IDState(4)
-    val q5 = new IDState(5)
-    q5.setAccept(true)
-    q1.addTransition(new Transition('a', 'a', q2))
-    q2.addTransition(new Transition('b', 'b', q3))
-    q3.addTransition(new Transition('c', 'c', q4))
-    q4.addTransition(new Transition('d', 'd', q5))
-    val aut = new BAutomaton
-    aut.setInitialState(q1)
-    new BricsAutomaton(aut)
-  }
-
-  // Automaton q1 -[a]-> q2 -[b]-> q3
-  //              -[c]-> q4
-  val aborcAut = {
-    val q1 = new IDState(1)
-    val q2 = new IDState(2)
-    val q3 = new IDState(3)
-    val q4 = new IDState(4)
-    q3.setAccept(true)
-    q4.setAccept(true)
-    q1.addTransition(new Transition('a', 'a', q2))
-    q2.addTransition(new Transition('b', 'b', q3))
-    q1.addTransition(new Transition('c', 'c', q4))
-    val aut = new BAutomaton
-    aut.setInitialState(q1)
-    new BricsAutomaton(aut)
-  }
-
-
-  def seq(s : String) = s.map(_.toInt)
-
-  property("Simple reverse abcd") = {
-    // dcba = reverse(abcd)
-    ReversePreOp(Seq(), abcdAut)._1.exists(cons => {
-      cons(0)(seq("dcba"))
-    })
-  }
-
-  property("Simple reverse abcd not accept") = {
-    // abcd != reverse(abcd)
-    !ReversePreOp(Seq(), abcdAut)._1.exists(cons => {
-      cons(0)(seq("abcd"))
-    })
-  }
-
-  property("Simple reverse with multiple final 1") = {
[…]

### 2024-02-19 evaluation.tex
Update results from SMT-COMP

@@
   \end{center}
-  \caption{Number of solved benchmarks in the set of quantifier-free string
-    linear integer arithmetic (QF\_SLIA) at SMT-COMP 2023. The numbers are from
-    the competition results, except for CA-Str, which is executed by us (600s
-    timeout, AMD Opteron 2220 SE), and \Ostrich+CA, a virtual portfolio. }
+  \caption{Number of solved benchmarks in the set of quantifier-free strings with
+    linear integer arithmetic constraints (QF\_SLIA) at SMT-COMP~2023. The numbers are from
+    the competition results, except for CA-Str, which is executed by us on the same 
+    cluster as the competition with the same resources, and the two virtual portfolio solvers
+    \Ostrich+CA and \textsc{Competition} which aggregate the best results from \Ostrich{}/CA-Str and 
+    all the competition results for non-\Ostrich{} solvers respectively. }
   \label{tab:solve-status-smt-comp}
@@
 instead of the previous baseline method, resulting in a new back-end
-\textbf{CA-Str}. We created also a new virtual portfolio~\Ostrich{}+CA
-by combining the three solvers in \Ostrich{}~1.3 with \textbf{CA-Str}.
+\textbf{CA-Str}. Combining the results from SMT-COMP for \Ostrich{}~1.3 
+with our new results running CA-Str we construct a virtual 
+portfolio~\Ostrich{}+CA that simulates running CA-Str as a fourth
+back-end to \Ostrich{}.
 
@@
 
-We evaluate the different solvers in \cref{tab:solve-status-smt-comp}
-on 21\,938~benchmarks from the single-query QF\_SLIA track at SMT-COMP 2023, as
-QF\_SLIA was the most challenging track for \Ostrich{} in the
-competition.
+We extend the results of SMT-COMP 2023 with our modified \Ostrich
+on the single-query string solving with linear integer arithmetic constraint
+track (QF\_SLIA) in~\cref{tab:solve-status-smt-comp}. We obtain
+the results by executing \Ostrich~1.3 with CA-Str using
+the same benchmarking infrastructure (the StarExec cluster)
+and configuration that ran SMT-COMP 2023, combining our new results for CA-Str with the
+published results of SMT-COMP 2023. Where avaiable, we use the revised, out-of
+competition version of the results for solvers with bugs.
+
+We picked this track since
+\Ostrich{} already performed well on the other two string solving tracks. In fact,
+every solver, including \Ostrich{}, handled every benchmark in QF\_SNIA within the timeout.
 %
-As can be seen from the table, integrating \Catra{} as a back-end
+As can be seen from \cref{tab:solve-status-smt-comp}, integrating \Catra{} as a back-end
 leads to gains both on satisfiable and unsatisfiable problems. On
-satisfiable benchmarks, the combination \Ostrich+CA is still
+satisfiable problems, the combination \Ostrich+CA is still
 outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
-\Ostrich+CA narrowly beats the other solvers, which is promising given
-that also \Ostrich{}~1.3, cvc5, and z3alpha show very strong
-performance on this class of benchmarks.
+\Ostrich+CA narrowly beats the other solvers,
+which is promising given that \Ostrich{}~1.3 (without CA-Str), cvc5, and z3alpha show very strong
+performance on this class of benchmarks. Similar improvements
+occur in QF\_S, they are much smaller (an additional 2 unsatisfiable, 4 satisfiable 
+solved out of~8\,847 in total).
 
@@
 however, as we believe that a deeper integration of \Catra{} into string solvers
-can lead to significant performance gains.
+can lead to significant performance gains. In particular, the integration layer 
+is currently too shallow to allow Ostrich to learn clauses generated by \Catra{}
+and additionally incurs subprocess-spawning overhead.
 
@@
 of instances. Since addressing that bug we have observed no further soundness
-issues.
+issues, with the exception of one erroneous \textsc{Unsat} response on the QF\_S track of 
+SMT-COMP (\texttt{QF\_S\/20230329-automatark-lu\/instance08425.smt2}) that likely occurred
+in the interaction surface between \Ostrich{} and \Catra{}.
 

### 2024-02-19 evaluation.tex
Add bold to the table

@@
 For our experiments, we modified the CE-Str solver to apply \Catra{}
+ (with \Calculus{} as its chosen backend)
 instead of the previous baseline method, resulting in a new back-end
 \textbf{CA-Str}. Combining the results from SMT-COMP for \Ostrich{}~1.3 
-with our new results running CA-Str we construct a virtual 
-portfolio~\Ostrich{}+CA that simulates running CA-Str as a fourth
+with our new results running CA-Str we constructed a virtual 
+portfolio~\Ostrich{}+CA that simulated running CA-Str as a fourth
 back-end to \Ostrich{}.
@@
 
-We extend the results of SMT-COMP 2023 with our modified \Ostrich
+We extend the results of SMT-COMP 2023 with our modified \Ostrich{}
 on the single-query string solving with linear integer arithmetic constraint
 track (QF\_SLIA) in~\cref{tab:solve-status-smt-comp}. We obtain
-the results by executing \Ostrich~1.3 with CA-Str using
+the results by executing \Ostrich{}~1.3 with CA-Str using
 the same benchmarking infrastructure (the StarExec cluster)
@@
 published results of SMT-COMP 2023. Where avaiable, we use the revised, out-of
-competition version of the results for solvers with bugs.
+competition version of the results for solvers with bugs, including \Ostrich{} and Z3-Noodler.
+Note that Z3-Noodler abstained on 70~instances, and thus has a lower total number of results.
 
-We picked this track since
+We picked QF\_SLIA since
 \Ostrich{} already performed well on the other two string solving tracks. In fact,
@@
 can lead to significant performance gains. In particular, the integration layer 
-is currently too shallow to allow Ostrich to learn clauses generated by \Catra{}
-and additionally incurs subprocess-spawning overhead.
+is currently too shallow to allow \Ostrich{} to learn clauses generated by \Catra{}
+and additionally incurs subprocess-spawning overhead and others, as well as overhead
+from serialising and deserialising the current Parikh automata problem into \Catra{}'s
+input format.
 
@@
 SMT-COMP (\texttt{QF\_S\/20230329-automatark-lu\/instance08425.smt2}) that likely occurred
-in the interaction surface between \Ostrich{} and \Catra{}.
+in the interaction surface between \Ostrich{} and \Catra{}, since we have robustly tested 
+\Catra{} separately from \Ostrich{}. For this case, as well as for the similarly buggy case
+for Z3-Noodler, we use the same rule as SMT-COMP and count the erroneous responses as error
+results.
 

### 2024-02-20 evaluation.tex
Clarify origins of results

@@
 and configuration that ran SMT-COMP 2023, combining our new results for CA-Str with the
-published results of SMT-COMP 2023. Where avaiable, we use the revised, out-of
-competition version of the results for solvers with bugs, including \Ostrich{} and Z3-Noodler.
+published results of SMT-COMP 2023.
+\footnote{StarExec Job ID 59410, "SMT-COMP 2023 single query final", 
+\url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59410}.} 
+Where avaiable, we use the revised, out-of
+competition version of the results for solvers with bugs discovered during
+the competition, including \Ostrich{} and Z3-Noodler. 
+\footnote{StarExec Job ID 59668, "SMT-COMP 2023 single query final fixed" available at
+\url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59668}.}
 Note that Z3-Noodler abstained on 70~instances, and thus has a lower total number of results.
+We used the parallel results for all solvers.
 

### 2024-02-20 evaluation.tex
More evaluation touch-up

@@
 used to support, e.g., a string solver. Experiments were executed in random order
-for all backends. Each instance got a time budget of~\RuntimeTimeout.
+for all backends. Each instance got a time budget of~\RuntimeTimeout{}.
 
@@
 
-We extend the results of SMT-COMP 2023 with our modified \Ostrich{}
+We extend the results of SMT-COMP~2023 with our modified \Ostrich{}
 on the single-query string solving with linear integer arithmetic constraint
@@
 the same benchmarking infrastructure (the StarExec cluster)
-and configuration that ran SMT-COMP 2023, combining our new results for CA-Str with the
-published results of SMT-COMP 2023.
+and configuration that ran SMT-COMP~2023, combining our new results for CA-Str with the
+published results of SMT-COMP~2023.
 \footnote{StarExec Job ID 59410, "SMT-COMP 2023 single query final", 
@@
 outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
-\Ostrich+CA narrowly beats the other solvers,
-which is promising given that \Ostrich{}~1.3 (without CA-Str), cvc5, and z3alpha show very strong
+\Ostrich+CA now narrowly beats the other solvers by squeezing past cvc5,
+which is promising given that \Ostrich{}~1.3 (without CA-Str), cvc5, and z3alpha all show very strong
 performance on this class of benchmarks. Similar improvements
@@
 of instances. Since addressing that bug we have observed no further soundness
-issues, with the exception of one erroneous \textsc{Unsat} response on the QF\_S track of 
-SMT-COMP (\texttt{QF\_S\/20230329-automatark-lu\/instance08425.smt2}) that likely occurred
-in the interaction surface between \Ostrich{} and \Catra{}, since we have robustly tested 
-\Catra{} separately from \Ostrich{}. For this case, as well as for the similarly buggy case
-for Z3-Noodler, we use the same rule as SMT-COMP and count the erroneous responses as error
-results.
+issues in \Catra{}. During extended evaluations by reviewer suggestion for the 
+camera-ready version of this paper, however, one single erroneous \textsc{Unsat} 
+response on the QF\_S track of 
+SMT-COMP (\texttt{QF\_S\/20230329-automatark-lu\/instance08425.smt2}) was discovered.
+The most likely cause is a misgenerated constraint in the interaction surface between \Ostrich{}
+and \Catra{}, since we have robustly tested 
+\Catra{} separately from \Ostrich{} and vice versa. Additionally, Z3-Noodler
+also similarly has one incorrect response, even in the bugfixed version. For both of
+these, we apply the same rule as SMT-COMP, counting the erroneous responses for Z3-Noodler
+and \Ostrich{}+CA as error results.
 
@@
 Finally, the difference in performance between \Calculus{} and baseline has been
-robust under significant optimisation of the automata library.
+robust under significant optimisation of the automata library, additionally
+strengthening this thesis.
 

### 2024-02-20 evaluation.tex
Same format

@@
 \footnote{StarExec Job ID 59410, "SMT-COMP 2023 single query final", 
-\url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59410}.} 
+available at \url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59410}.} 
 Where avaiable, we use the revised, out-of

### 2024-02-21 calculus.tex
Clarify backward propagation, dom()
Closes: #81

@@
 \end{definition}
+
+The $\Dominates$~relation can be efficiently implemented in a solver
+by using standard Ford-Fulkerson\slash{}Edmonds-Karp min-cut between a state
+and the initial state after removing transitions where $\Filter(\Transition) = 0$.
+By only performing this computation after such filtering, a solver additionally avoids
+breaking the rule of adding clauses that already appear in the formula.
+
+We here only define forward propagation. Backward propagation 
+can be straightforwardly obtained by computing reachability from
+an accepting state on the same principle as above, and indeed our
+implementation of \Calculus{} does this.
 

### 2024-02-21 multiple.tex
typesetting
corrected flow
Fix lemma 6.1 re/side-conditions
Closes: #83

@@
            \Connected(\Automaton', \Filter'),~
-           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
+           \BindingSum(\Automaton_i, \Automaton_j, \Automaton', \Filter')
            \\\ImagePredicate{
@@
   $$
-  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
+  \BindingSum(\Automaton_i, \Automaton_j, \Automaton', \Filter') ~~=~~
   \begin{aligned}
   & \Set{ 
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State'_R}}} \Filter'(\Transition')
+    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State'_R}} \in \Transitions_{\Automaton'}} \Filter'(\Transition')
   \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
   & \Set{
-    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State'_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
+    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State'_L, \State'}} \in \Transitions_{\Automaton'}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
   }
@@
 \begin{equation}\label{eq:multi-filter}
-    \begin{aligned}
-        & \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6 
-        & \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7 
-        & \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) = \TransitionVar_8 
-        &\Filter(\FromLabelTo{B}{c / l_{B+}}{B}) = \TransitionVar_9  \\
-        & \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}
-        & \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}
-        & \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}
-        \FlowEq(\mathcal{B}, \Filter) &=  \begin{aligned}
-            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
-            \land \TransitionVar_{11} &= \TransitionVar_6\\
-            \land \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}
-        \end{aligned}
-    \end{aligned}
+  \begin{array}{rl@{\qquad}rl@{\qquad}rl}
+     \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) &= \TransitionVar_6 
+     &
+     \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) &= \TransitionVar_7 
+     &
+      \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) &= \TransitionVar_8 
+     \\
+     \Filter(\FromLabelTo{B}{c / l_{B+}}{B}) &= \TransitionVar_9 
+     &
+     \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B}) &= \TransitionVar_{10}
+     &
+     \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A}) &= \TransitionVar_{11}
+     \\
+      \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B}) &= \TransitionVar_{12}
+    \\[3ex]
+% 
+    \multicolumn{6}{c}{
+        \FlowEq(\mathcal{B}, \Filter) =\quad  \begin{aligned}
+            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12} \wedge\mbox{}\\
+            \TransitionVar_{11} &= \TransitionVar_6\wedge\mbox{}\\
+            \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}\mbox{}\\
+            \TransitionVar_6 + \TransitionVar_{10} &= 1
+        \end{aligned}}
+  \end{array}
 \end{equation}
@@
 
-  This implies that also the rule~\ExpandM{} can only be applied
-  finitely often since its side condition only allows applying it once
-  per instance of an $\Image$ predicate containing a product, and only finitely
-  many instances of $\Image$ can be introduced on each branch.
+  \ExpandM{} can only be applied precisely once per $\Image$~term since
+  each application introduces an identical set of formulas and we have a generic
+  side condition that no rule may add only redundant formulas.
 \end{proof}

### 2024-02-22 appendix.tex
Fix minor reviewer issues
Closes: #77
Closes: #75
Closes: #76
Closes: #71
Closes: #72
Flesh out evaluation somewhat, move figures from appendix
Closes: #73
Closes: #74

Also addresses: #78

@@
-\subsection{Per-instance comparison between backends}
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter.pdf}
-    \caption{Per-instance runtime between baseline and \Calculus{}}
-    \label{fig:duels:baseline}
-    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and baseline]{The plot shows most dots at the bottom of the plot, with some plots collected at the right-hand side.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter.pdf}
-    \caption{Per-instance runtime between baseline and \Calculus{}}
-    \label{fig:duels:nuxmv}
-    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and \Nuxmv{}]{The plot shows flames of dots between the bottom and top, with a line of dots near the right-hand side corresponding to timeouts.}
-  \end{subfigure}
-\end{figure}
-
 \subsection{Example instance for \Catra{}}

### 2024-02-22 evaluation.tex
Flesh out evaluation somewhat, move figures from appendix
Closes: #73
Closes: #74

Also addresses: #78

@@
 
-In \cref{fig:solve-division}, we show how many of the~\NrBenchmarks{} instances
-the respective back-end could solve. A summary of their outcomes by instance
-type is also available in \cref{tab:solve-status}. Note that many instances lack
-a ground truth as they are solved by only one backend. We see that
-\Calculus{} generally outperforms \Nuxmv{} on determining unsatisfiability, as
-does baseline, while being similar at satisfiable instances. Both \Calculus{} and
-\Nuxmv{} outperforms baseline on satisfiable instances. On satisfiable
-instances, \Nuxmv{} and \Calculus{} have similar performance.
-
-Baseline performs worse on satisfiable instances because it executes a heuristic
-meant to detect unsatisfiability early, similar to~\cite{approximate-parikh}.
-The heuristic is enabled since the improvement is significant compared to the
-extra cost.
-
-\begin{table}
-  \begin{center}
-  \input{graphs/solved_pivot_table.tex}
-  \end{center}
-  \caption{Number of successful results within a timeout of \RuntimeTimeout{}.
-  Instances solved by no backend within the timeout (about half of the set) are
-  omitted from the table.}\label{tab:solve-status}
-\end{table}
-
 \begin{figure}[t]
@@
     \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
-    \caption{The distribution of runtimes for solved instances per backend. Total numbers of solved instances differ.}
+    \caption{The distribution of runtimes for solved instances per backend, assuming they were solved within the \numprint{30}s timeout.
+    The whiskers use the default placement in Pandas, \numprint{1.5} times the IQR (interquartile range).}
     \label{fig:runtime-boxplot}
@@
 \end{figure}
+
+In \cref{fig:solve-division}, we show how many of the~\NrBenchmarks{} instances
+the respective back-end could solve. A summary of their outcomes by instance
+type is also available in \cref{tab:solve-status}. Note that many instances lack
+a ground truth as they are solved by only one backend. We see that
+\Calculus{} generally outperforms \Nuxmv{} on determining unsatisfiability, as
+does baseline, while being similar at satisfiable instances. Both \Calculus{} and
+\Nuxmv{} outperforms baseline on satisfiable instances. On satisfiable
+instances, \Nuxmv{} and \Calculus{} have similar performance.
+
+Baseline performs worse on satisfiable instances because it executes a heuristic
+meant to detect unsatisfiability early, similar to~\cite{approximate-parikh}.
+The heuristic is enabled since the improvement is significant compared to the
+extra cost.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter.pdf}
+    \caption{Per-instance runtime between baseline and \Calculus{}}
+    \label{fig:duels:baseline}
+    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and baseline]{The plot shows most dots at the bottom of the plot, with some plots collected at the right-hand side.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter.pdf}
+    \caption{Per-instance runtime between \Nuxmv{} and \Calculus{}}
+    \label{fig:duels:nuxmv}
+    \Description[A scatter plot showing individual instances as dots compared to \Calculus{} and \Nuxmv{}]{The plot shows flames of dots between the bottom and top, with a line of dots near the right-hand side corresponding to timeouts.}
+  \end{subfigure}
+\end{figure}
+
+Finally, in \cref{fig:runtime-boxplot,fig:duels:baseline,fig:duels:nuxmv} and we compare
+runtimes for solved instances between the backends. We see that \Nuxmv{} has a more even spread of runtimes up to \numprint{20}s,
+while both \Calculus{} and baseline tend towards solving their
+instances quickly or not at all, though \Calculus{} does have a long tail of outlier
+instances that finish as the timeout increaseses.
+Notably, \cref{fig:duels:baseline} shows that any instance solved by baseline in \numprint{30}s is also
+solved by \Calculus{} in under \numprint{5}s.
+
+\begin{table}
+  \begin{center}
+  \input{graphs/solved_pivot_table.tex}
+  \end{center}
+  \caption{Number of successful results within a timeout of \RuntimeTimeout{}.
+  Instances solved by no backend within the timeout (about half of the set) are
+  omitted from the table.}\label{tab:solve-status}
+\end{table}
 
@@
 We extend the results of SMT-COMP~2023 with our modified \Ostrich{}
-on the single-query string solving with linear integer arithmetic constraint
+on the single-query string solving with linear integer arithmetic constraints
 track (QF\_SLIA) in~\cref{tab:solve-status-smt-comp}. We obtain

### 2024-02-22 evaluation.tex
Add a cactus plot for QF_SLIA

@@
   \begin{center}
-  \input{experiments/qf_slia_comparison.tex}
+  \input{graphs/qf_slia_comparison.tex}
   \end{center}
@@
 \end{table}
+
+\begin{figure}[t]
+  \includegraphics[width=\linewidth]{graphs/smt-comp-cactus}
+  \caption{The number of instances solved in the QF\_SLIA track of SMT-COMP 2023
+   as the time budget increases towards the competition limit of \numprint{20}~minutes.}
+  \label{fig:smt-comp:cactus}
+  \Description[A cactus plot comparing the performance of the solvers in the QF_SLIA track of SMT-COMP]%
+  {The plot shows the results for the solvers Ostrich 1.3, CA-Str, Ostrich+CA, Z3-Noodler,
+  cvc5 and z3alpha as number of solved instances (y-axis) after N seconds (x-axis). The
+  topmost solver is the virtual portfolio "Competition", representing every non-Ostrich
+  solver, followed by cvc5, after ca 200 seconds, after which it overtakes the previous leader, z3alpha.}
+\end{figure}
 
@@
 %
-As can be seen from \cref{tab:solve-status-smt-comp}, integrating \Catra{} as a back-end
+As can be seen from \cref{tab:solve-status-smt-comp} and \cref{fig:smt-comp:cactus},
+integrating \Catra{} as a back-end
 leads to gains both on satisfiable and unsatisfiable problems. On
@@
 outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
-\Ostrich+CA now narrowly beats the other solvers by squeezing past cvc5,
+\Ostrich+CA now narrowly beats the other solvers 
+(shown as another virtual portfolio solver, "Competitors") by squeezing past cvc5,
 which is promising given that \Ostrich{}~1.3 (without CA-Str), cvc5, and z3alpha all show very strong
@@
 The results show that \Calculus{} can be used to enhance the performance of an
-automaton-based string solver. The results should be considered preliminary,
+automaton-based string solver. Moreover, the cactus plot in \cref{tab:solve-status-smt-comp} 
+illustrates that
+CA-Str is immediately useful, boosting the results for the \Ostrich{} portfolio
+even at the first datapoint. These results should be considered preliminary,
 however, as we believe that a deeper integration of \Catra{} into string solvers

### 2024-02-22 evaluation.tex
Fix quotes

@@
 published results of SMT-COMP~2023.
-\footnote{StarExec Job ID 59410, "SMT-COMP 2023 single query final", 
+\footnote{StarExec Job ID 59410, \enquote{SMT-COMP 2023 single query final}, 
 available at \url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59410}.} 
@@
 the competition, including \Ostrich{} and Z3-Noodler. 
-\footnote{StarExec Job ID 59668, "SMT-COMP 2023 single query final fixed" available at
+\footnote{StarExec Job ID 59668, \enquote{SMT-COMP 2023 single query final fixed} available at
 \url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59668}.}

### 2024-02-22 evaluation.tex
A preliminary attempt at addressing the worse SAT performance
Adresses: #82

@@
 
-For our experiments, we modified the CE-Str solver to apply \Catra{}
+For our experiments, we modify the CE-Str solver to apply \Catra{}
  (with \Calculus{} as its chosen backend)
@@
 \textbf{CA-Str}. Combining the results from SMT-COMP for \Ostrich{}~1.3 
-with our new results running CA-Str we constructed a virtual 
-portfolio~\Ostrich{}+CA that simulated running CA-Str as a fourth
-back-end to \Ostrich{}.
+with our new results running CA-Str we construct a virtual 
+portfolio~\Ostrich{}+CA that simulates running CA-Str as a fourth
+back-end to \Ostrich{}. We similarly combine the results of all non-\Ostrich{} solvers
+to obtain the virtual portfolio solver~\textbf{Competition}.
 
@@
 on the single-query string solving with linear integer arithmetic constraints
-track (QF\_SLIA) in~\cref{tab:solve-status-smt-comp}. We obtain
-the results by executing \Ostrich{}~1.3 with CA-Str using
+track (QF\_SLIA). We picked QF\_SLIA since
+\Ostrich{} already performed well on the other two string solving tracks. In fact,
+every solver, including \Ostrich{}, handled every benchmark in QF\_SNIA within the timeout.
+Additionally, Parikh intersection problems would mainly be generated by \Ostrich{} when
+solving constraints involving integers, meaning that CA-Str would be of little or no
+help on the QF\_S track.
+
+We obtain the results by executing \Ostrich{}~1.3 with CA-Str using
 the same benchmarking infrastructure (the StarExec cluster)
@@
 
-We picked QF\_SLIA since
-\Ostrich{} already performed well on the other two string solving tracks. In fact,
-every solver, including \Ostrich{}, handled every benchmark in QF\_SNIA within the timeout.
-%
 As can be seen from \cref{tab:solve-status-smt-comp} and \cref{fig:smt-comp:cactus},
@@
 outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
-\Ostrich+CA now narrowly beats the other solvers 
-(shown as another virtual portfolio solver, "Competitors") by squeezing past cvc5,
+\Ostrich+CA now narrowly beats the other solvers squeezing past cvc5,
 which is promising given that \Ostrich{}~1.3 (without CA-Str), cvc5, and z3alpha all show very strong
-performance on this class of benchmarks. Similar improvements
-occur in QF\_S, they are much smaller (an additional 2 unsatisfiable, 4 satisfiable 
-solved out of~8\,847 in total).
+performance on this class of benchmarks.
+
+These results are expected. It is known that automata-based string solvers (\Ostrich{}, Z3-Noodler) 
+tend to perform better on unsatisfiable than on satisfiable benchmarks, compared to solvers
+that do not utilize automata and directly work on regular expressions (cvc5, z3alpha).
+This is not unexpected, as the computation of automata representations of regular 
+constraints can be expensive, and might be unnecessary for satisfiable formulas.
+In addition, the algorithm in \Ostrich{} has stronger theoretical completeness
+guarantees as cvc5 and Z3-Str, and ensuring completeness often has an adverse
+effect on performance in practice~\cite{ostrich,Z3-str,cvc5}.
 

### 2024-02-28 introduction.tex
Finally 23 pages
Define the problem

@@
 instances, and comes with a harsh penalty for satisfiable instances.
+
+The most basic version fo the problem considered in this paper is:
+{
+    \centering
+    \begin{tcolorbox}[colback=gray!5!white,colframe=gray!75!black,%
+        title=Satisfiability of Parikh images of a product of automata,%
+        width=0.8\linewidth]
+        Given a finite set of regular languages, $\Language_1, \ldots, \Language_n$
+        over a finite, common, alphabet $\Set{a_1, \ldots, a_k}$,
+        and a Presburger formula, $\phi(\Vec{x_1, \ldots, x_k})$, 
+        decide if there is a word, $w$ in the 
+        intersection of $\Language_1, \ldots, \Language_n$, whose Parikh vector 
+        $\Vec{m_1, \ldots, m_k}$ satisfies $\phi$.
+    \end{tcolorbox}
+  }
 
@@
 with the conjunction of the automata's Parikh maps, $\bigwedge_{i=1}^{k}
-\ParikhMap(\Automaton_i)$. Due to the over-approximation,
+\ParikhMap(\Language(\Automaton_i))$. Due to the over-approximation,
 this approach works only for unsatisfiable instances,

### 2024-02-29 evaluation.tex
Poke the introduction
Closes: #86
Closes: #88
Closes: #82
Closes: #80

@@
 \url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59668}.}
-Note that Z3-Noodler abstained on 70~instances, and thus has a lower total number of results.
+Note that Z3-Noodler abstained on \numprint{70}~instances, and thus has a lower total number of results.
 We used the parallel results for all solvers.
@@
 In addition, the algorithm in \Ostrich{} has stronger theoretical completeness
-guarantees as cvc5 and Z3-Str, and ensuring completeness often has an adverse
+guarantees than cvc5 and Z3-Str, and ensuring completeness often has an adverse
 effect on performance in practice~\cite{ostrich,Z3-str,cvc5}.
@@
 CA-Str is immediately useful, boosting the results for the \Ostrich{} portfolio
-even at the first datapoint. These results should be considered preliminary,
+even at the first datapoint (though marginally). These results should be considered preliminary,
 however, as we believe that a deeper integration of \Catra{} into string solvers
@@
 is currently too shallow to allow \Ostrich{} to learn clauses generated by \Catra{}
-and additionally incurs subprocess-spawning overhead and others, as well as overhead
-from serialising and deserialising the current Parikh automata problem into \Catra{}'s
-input format.
+and additionally incurs overhead from serialising and deserialising the current Parikh
+automata problem into \Catra{}'s input format.
 
@@
 product materialisation causing non-deterministic unsoundness in \numprint{0.7}
-of instances. Since addressing that bug we have observed no further soundness
-issues in \Catra{}. During extended evaluations by reviewer suggestion for the 
-camera-ready version of this paper, however, one single erroneous \textsc{Unsat} 
-response on the QF\_S track of 
-SMT-COMP (\texttt{QF\_S\/20230329-automatark-lu\/instance08425.smt2}) was discovered.
-The most likely cause is a misgenerated constraint in the interaction surface between \Ostrich{}
-and \Catra{}, since we have robustly tested 
-\Catra{} separately from \Ostrich{} and vice versa. Additionally, Z3-Noodler
-also similarly has one incorrect response, even in the bugfixed version. For both of
-these, we apply the same rule as SMT-COMP, counting the erroneous responses for Z3-Noodler
-and \Ostrich{}+CA as error results.
+of instances. Additionally, both we and the artefact evaluation committee independently
+discovered a soundness issue in CA-Str on one instance in QF\_S where the 
+\texttt{QF\_S\slash{}20230329-automatark-lu\slash{}instance08425.smt2} instance was
+incorrectly reported as \textsc{Unsat}. The underlying issue was that \Catra{} generated
+a too broad blocking clause when encountering unsatisifiability stemming from 
+an intermittent product accepting only the empty string during product materialisation. 
+We implemented a fix for that bug and re-evaluated all benchmarks in SMT-COMP.
+
+Since addressing those bugs we have observed no further soundness
+issues in neither \Catra{} nor CA-Str. Benchmarking results for both runs are included
+in the artefact of this paper, and show virtually identical performance characteristics.
 
@@
 product implementation. We believe this is not an issue since similar
-performance issues with the baseline approach have been reported for string
-solvers. Additionally, profiling shows that baseline spends most of its time in
+performance issues with the baseline approach have been reported for other string
+solvers, as well as for a previous implementation in \Ostrich{}. Additionally,
+profiling shows that baseline spends most of its time in
 \Princess{}, suggesting that the automaton implementation is not the bottleneck.

### 2024-02-29 introduction.tex
FO FO
Poke the introduction
Closes: #86
Closes: #88
Closes: #82
Closes: #80

--- 2023-07-11 introduction.tex
@@
 
-It is possible to compute an existential Presburger formula describing the
-Parikh image of any context-free language in linear
-time~\cite{generate-parikh-image}; for the special case of regular languages,
-this result was also stated in \cite{muscholl-linear}. While theoretically
-elegant, this construction has several disadvantages, often making it
-impractical for integration into algorithms. Firstly, the constructed Presburger
+In applications involving regular languages, it is typically necessary
+to consider the Parikh image not only of a single automaton, but of the
+intersection of multiple automata. This problem arises in string solvers in
+particular, as conjunctions of string constraints lead to the computation of
+length images of products (intersections) of regular languages represented as
+finite automata. The problem considered in this paper is thus:
+{
+    \centering
+    \begin{tcolorbox}[colback=gray!5!white,colframe=gray!75!black,%
+        title=Satisfiability of Parikh images of a product of automata,%
+        width=0.8\linewidth]
+        Given a finite set of deterministic, finite-state automata, $\Automaton_1, \ldots, \Automaton_n$
+        over a finite, common, alphabet $\Set{a_1, \ldots, a_k}$,
+        and a Presburger formula, $\phi(\VectorLiteral{x_1, \ldots, x_k})$, 
+        decide if there is a word, $w$ accepted by the 
+        product $\Automaton_1 \times \cdots \times \Automaton_n$, whose Parikh vector 
+        $\VectorLiteral{m_1, \ldots, m_k}$ satisfies $\phi$.
+    \end{tcolorbox}
+  }
+
+  The current method of handling this problem consists of computing the
+  intersection of the languages (the product of their automata),
+  constructing the Presburger representation of their Parikh image, and then
+  checking its satisfiability. It is possible to compute an existential
+   Presburger formula describing the
+  Parikh image of any \emph{context-free} language in linear
+  time~\cite{generate-parikh-image}. For our special case of regular languages,
+  this result was also stated in~\cite{muscholl-linear}. While theoretically
+  elegant, this construction has several disadvantages for us, often making it
+  impractical for integration into algorithms.
+
+  Firstly, the constructed Presburger
   formula contains a linear number of existential quantifiers in the size of the
   considered grammar, as well as complex Boolean structure, which is needed to
-express the connectedness of sets of productions considered in the construction.
-Eliminating those quantifiers to obtain a quantifier-free representation of the
+  express the connectedness of sets of productions considered in the construction.%
+  \iffalse{}Eliminating those quantifiers to obtain a quantifier-free representation of the
   Parikh image has exponential complexity~\cite{DBLP:conf/issac/Weispfenning97},
-and is often impossible in reasonable time. Just solving the Parikh image
+  and is often impossible in reasonable time.\fi{}%
+  Just solving the Parikh image
   membership problem is NP-complete, as it corresponds to computing a satisfying
@@
 
-Secondly, in applications involving regular languages, it is typically necessary
-to consider the Parikh image not only of a single automaton, but of the
-intersection of multiple automata. This problem arises in string solvers in
-particular, as conjunctions of string constraints lead to the computation of
-length images of products (intersections) of regular languages represented as
-finite automata. Applying the approach in~\cite{generate-parikh-image} would in
+  Secondly, computing the product itself is usually prohibitively expensive
+  in both memory and CPU due to exponential blow-up. Applying the approach in~\cite{generate-parikh-image} would in
   this case require the eager computation of the product before its length image,
@@
   size of the product, quickly becoming intractable as the number of automata in
-the product increases. The current best published mitigation for this problem is
-an over-approximation that works by approximating the Parikh image of a product
-of automata to be the conjunction of the image of the individual automata of the
-product \cite{approximate-parikh}. This approach only works for unsatisfiable
-instances, and comes with a harsh penalty for satisfiable instances.
+  the product increases.
+
+  The current best published mitigation for this problem relies on the observation
+  that any Presburger formula that applies to a product of automata must also hold
+  for all of the terms of that product individually. Exploiting this fact, the Parikh
+  image of a product of automata is first represented as the conjunction of the image of
+  the individual automata of the product, falling back to computing the entire product
+  for satisfiable cases~\cite{approximate-parikh}.
 
@@
 
-The key ideas of \Calculus{} is a combination of problem-aware case splitting,
-lazy enforcement of automata connectivity, and lazy computation of products. In
-essence, we use constraint-based methods to efficiently enforce the constraints
-associated with computing Parikh images of regular languages.
+Our primary influence for this work is the operation of (finite-domain) 
+constraint programming (CP)%
+~solvers~\cite{cp}. CP~solvers typically employ a combination of lazy enforcement
+of constraints through \emph{propagation},
+lazily pruning values from decision variable domains that do not satisfy 
+the constraints during solving, and \emph{branching}, splitting proofs
+by cases in a way that will encourage maximum propagation.
+
+For our problem, the constraints are automata connectivity and linear
+equations (Presburger formulas).
+We use propagation to lazily enforce connectivity of automata,
+removing transitions as they become unreachable during solving or are
+eliminated by constraints in the Presburger formula. We
+also branch on the presence or absence of transitions to
+encourage propagation of the connectivity constraint. This allows
+us to whittle down the automata
+(including intermittent products) before moving on with potentially
+ explosive product construction. The underlying observation
+is that though the problem domain and the domain of our integer variables
+is infinite, the number of simple paths
+through the automata (and their product) is finite. In summary, the key ideas are:
+
+{
+    \centering
+    \begin{tcolorbox}[colback=gray!5!white,colframe=gray!75!black,%
+        title=Key ideas of \Calculus{},%
+        width=0.8\linewidth]
+
+        \begin{itemize}
+            \item Enforce automata connectivity constraints lazily
+            \item Use \emph{propagation} of constraints on the Parikh image to prune
+            transitions from the automata that are incompatible with constraints
+            \item Intelligently \emph{branch} on the presence\slash{}absence of key
+             transitions to drive propagation of the connectivity constraint
+             \item Compute products of automata lazily, after propagation
+        \end{itemize}
+    \end{tcolorbox}
+  }
 
@@
 theorem prover.
+% Suggestions: l87-89, l102-103: Mention the main evaluation result; such as, x% more than baseline.
 %
 We additionally wrap the Parikh image solver as a stand-alone tool, \Catra.
-\Catra{} also supports a variant of the approximate method of
-\cite{approximate-parikh}, its fall-back variant adapted
+\Catra{} also supports a variant of the approximate method of~\cite{approximate-parikh}, its fall-back variant adapted
 from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
@@
 generated by \OstrichPlus{} when solving the PyEx string constraint benchmark
-suite involving string length constraints \cite{pyex}.
+suite involving string length constraints~\cite{pyex}.
 

### 2024-02-29 introduction.tex
Mention results in introduction
Closes: #87

@@
 We implement \Calculus{} as a plug-in theory for the \Princess{} automated
-theorem prover.
-% Suggestions: l87-89, l102-103: Mention the main evaluation result; such as, x% more than baseline.
-%
-We additionally wrap the Parikh image solver as a stand-alone tool, \Catra.
+theorem prover and experimentally use it as the Parikh image interesection solver for
+to the \Ostrich{} string constraint solver.
+
+We additionally wrap the Parikh image solver as a stand-alone tool, \Catra{}.
 \Catra{} also supports a variant of the approximate method of~\cite{approximate-parikh}, its fall-back variant adapted
@@
 generated by \OstrichPlus{} when solving the PyEx string constraint benchmark
-suite involving string length constraints~\cite{pyex}.
+suite involving string length constraints~\cite{pyex}, finding that \Calculus{}
+outperforms both \Nuxmv{} and the baseline method, allowing \Calculus{}
+to solve every problem solved by baseline within \numprint{30}s in under \numprint{5}s.
+\Calculus{} in particular outperforms \Nuxmv{} on unsatisfiable instances, more than 
+doubling the number of unsatisfiable results found within a \numprint{30}s timeout.
+
+We extend the \Ostrich{} string solver's backend portfolio with \Calculus{} via
+\Catra{} as the Parikh image intersection solver of the \OstrichPlus{}~algorithm,
+obtaining improvements in both \textsc{Sat}  and \textsc{Unsat} performance in
+\Ostrich{}' results from SMT-COMP 2023, crucially allowing \Ostrich{} to win
+the unsatisfiable category of the track of quantifier-free, linear constraints
+over string logic (QF\_SLIA) and increasing the number of \textsc{Sat}~results
+by \numprint{3}\%.
 

### 2024-02-29 introduction.tex
Clarify no backtrack
Nicer text

@@
 finite automata. The problem considered in this paper is thus:
+
 {
@@
 size of the partial products considered in the computation for a smaller memory
-footprint. Moreover, the method can be used iteratively to tackle smaller chunks
-of the product incrementally, thereby decreasing the memory footprint.
+footprint. Moreover, our approach allows us to do this precisely, without the expensive
+fallbacks when approximations fail as in the case of~\cite{approximate-parikh}.
 
@@
              transitions to drive propagation of the connectivity constraint
-             \item Compute products of automata lazily, after propagation
+             \item Compute products of automata lazily, after pruning transitions
+             that violate constraints using propagation
         \end{itemize}

### 2024-02-29 example.tex
Fix minor reviewer issues
Closes: #77
Closes: #75
Closes: #76
Closes: #71
Closes: #72
Ensure that figures are refered to, come before ref.
Closes: #78
Closes: #87

@@
 this example are in $\Naturals$ with $0$.
+
+\newcommand{\autscale}[0]{0.48}
+
+\begin{figure}[ht]
+    \centering 
+  \begin{subfigure}[b]{\autscale\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{a}
+    \caption{The automaton $A'$, whose Parikh registers count the length of the
+    string accepted by the automaton (register $l_A$), the start offset of the
+    substring ($i$) and length ($n$) of the substring matching
+    $\Regex{c\RegexOr{}dd}$. Note the symbolic transitions in the starting and
+    accepting states matching any symbol in the alphabet!}\label{fig:aut_a}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}%
+  \hfill%
+  \begin{subfigure}[b]{\autscale\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{b}
+    \caption{The automaton $B$, where the Parikh registers count the length of
+    the string in register $l_B$.}\label{fig:aut_b}
+      %\Description[SHORT]{LONG}
+  \end{subfigure}
+  \caption{The collection of automata we use as running
+  examples, both derived from \cref{ex:string-constraints}.}\label{fig:examples}
+\end{figure}
 
@@
 representing any character in the alphabet.
-
-\newcommand{\autscale}[0]{0.48}
-
-\begin{figure}[ht]
-    \centering 
-  \begin{subfigure}[b]{\autscale\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{a}
-    \caption{The automaton $A'$, whose Parikh registers count the length of the
-    string accepted by the automaton (register $l_A$), the start offset of the
-    substring ($i$) and length ($n$) of the substring matching
-    $\Regex{c\RegexOr{}dd}$. Note the symbolic transitions in the starting and
-    accepting states matching any symbol in the alphabet!}\label{fig:aut_a}
-      %\Description[SHORT]{LONG}
-  \end{subfigure}%
-  \hfill%
-  \begin{subfigure}[b]{\autscale\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{b}
-    \caption{The automaton $B$, where the Parikh registers count the length of
-    the string in register $l_B$.}\label{fig:aut_b}
-      %\Description[SHORT]{LONG}
-  \end{subfigure}
-  \caption{The collection of automata we use as running
-  examples, both derived from \cref{ex:string-constraints}.}\label{fig:examples}
-\end{figure}
 
@@
   \includegraphics[width=0.6\textwidth]{a_2}
-  \caption{ $\mathcal{A}'$ after even more linear algebra, with most transitions expressed
+  \caption{ $\mathcal{A}'$ after even more simplification using linear algebra,
+   with most transitions expressed
   in terms of $n$. Intuitively, this version captures the fact that $n$ is given
@@
 \end{figure}
-
-Having obtained this representation, we conclude that $1 < n \leq 2$
-from
-\cref{const:more-inside-than-before,const:something-before-and-after}. The
-lower bound directly follows from
-\cref{const:more-inside-than-before,const:something-before-and-after},
-whereas the upper bound is obtained by the following reasoning. Since all
-transition variables are non-negative integers (a transition cannot be
-used a negative amount of times), $n \leq 2$ from the $I_A$ to $F_A$
-transition. Therefore, it follows that $n=2$, which implies that the
-$I_A$ to $F_A$ transition can never be used under these constraints
-and that we must use the two $\Char{d}$-labelled transitions, both now
-taken $n-1 =1$ times.
-% This immediately implies that the problem is
-%unsatisfiable, since we know this transition has no correspondance in
-%automaton $B$.
-%
-Substituting the derived value of $n$ and applying similar
-reasoning to automaton~$\mathcal{B}$, we arrive at the simplified
-automata in \cref{fig:propagated}.
 
@@
 
+Having obtained this representation, we conclude that $1 < n \leq 2$
+from
+\cref{const:more-inside-than-before,const:something-before-and-after}. The
+lower bound directly follows from
+\cref{const:more-inside-than-before,const:something-before-and-after},
+whereas the upper bound is obtained by the following reasoning. Since all
+transition variables are non-negative integers (a transition cannot be
+used a negative amount of times), $n \leq 2$ from the $I_A$ to $F_A$
+transition. Therefore, it follows that $n=2$, which implies that the
+$I_A$ to $F_A$ transition can never be used under these constraints
+and that we must use the two $\Char{d}$-labelled transitions, both now
+taken $n-1 =1$ times.
+% This immediately implies that the problem is
+%unsatisfiable, since we know this transition has no correspondance in
+%automaton $B$.
+%
+Substituting the derived value of $n$ and applying similar
+reasoning to automaton~$\mathcal{B}$, we arrive at the simplified
+automata in \cref{fig:propagated}.
+
 Note that the per-automata length counting registers have been assigned the same
-solver variable $l_A$. Since they have to be equal, either one can be used in
+solver variable~$l_A$. Since they have to be equal, either one can be used in
 both automata through similar applications of equality elimination rules.

### 2024-02-29 introduction.tex
Apparently, this is our colour now
Ensure that figures are refered to, come before ref.
Closes: #78
Closes: #87

@@
   the individual automata of the product, falling back to computing the entire product
-  for satisfiable cases~\cite{approximate-parikh}.
+  for satisfiable cases or cases when unsatisfiability is only discoverable after
+  full- or partial product construction~\cite{approximate-parikh}.
 
@@
 of constraints through \emph{propagation},
-lazily pruning values from decision variable domains that do not satisfy 
+pruning values from decision variable domains that do not satisfy 
 the constraints during solving, and \emph{branching}, splitting proofs
@@
  explosive product construction. The underlying observation
-is that though the problem domain and the domain of our integer variables
+is that though the domain of our integer variables
 is infinite, the number of simple paths
@@
     \centering
-    \begin{tcolorbox}[colback=gray!5!white,colframe=gray!75!black,%
+    \begin{tcolorbox}[colback=ourcolour!5!white,colframe=ourcolour!75!black,%
         title=Key ideas of \Calculus{},%
@@
 over string logic (QF\_SLIA) and increasing the number of \textsc{Sat}~results
-by \numprint{3}\%.
+by \numprint{3}\% on the same track.
 

### 2024-02-29 multiple.tex
Ensure that figures are refered to, come before ref.
Closes: #78
Closes: #87

@@
 
-We return to our example in \cref{sec:intuition}. We also reuse the
-definition of $\Map$ as the simple projection to extract the counter increments,
-so in this case, $\Map(\Transition) = \left[l_{B+}\right]$ for every
-transition~$\Transition$ since $\mathcal{B}$ only counts length. The reader is
-advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
-\cref{sec:intuition} while going through this example.
-The transition variables of the figures match the ones used here.
-
-We extend \cref{eq:single-filter} from \cref{sec:single:example} with
-definitions for $B$ shown in \cref{eq:multi-filter}:
-\begin{equation}\label{eq:multi-filter}
-  \begin{array}{rl@{\qquad}rl@{\qquad}rl}
-     \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) &= \TransitionVar_6 
-     &
-     \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) &= \TransitionVar_7 
-     &
-      \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) &= \TransitionVar_8 
-     \\
-     \Filter(\FromLabelTo{B}{c / l_{B+}}{B}) &= \TransitionVar_9 
-     &
-     \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B}) &= \TransitionVar_{10}
-     &
-     \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A}) &= \TransitionVar_{11}
-     \\
-      \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B}) &= \TransitionVar_{12}
-    \\[3ex]
-% 
-    \multicolumn{6}{c}{
-        \FlowEq(\mathcal{B}, \Filter) =\quad  \begin{aligned}
-            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12} \wedge\mbox{}\\
-            \TransitionVar_{11} &= \TransitionVar_6\wedge\mbox{}\\
-            \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}\mbox{}\\
-            \TransitionVar_6 + \TransitionVar_{10} &= 1
-        \end{aligned}}
-  \end{array}
-\end{equation}
-
-The only possible rule at the start is $\ExpandM$, which we use to add the
-corresponding constraints on each automaton of the product as we would have had
-in the single-automaton version. After that, we expand the various sums and apply
-light equality propagation.
-
-We then continue by repeating the steps of \cref{sec:single:example} shown in
-\cref{fig:derivation:single}, since the steps to propagate
-\cref{ex:string-constraints} across the automaton $\mathcal{A}'$ still apply for the
-product. Note that this fixes the values of all transitions but one of $\mathcal{A}'$ as
-constants, including the free counter variables $i, n$, whose new constant
-values we propagate everywhere.
-
-Once we have performed reasoning by equality to eliminate transition variables,
-we obtain the version of $\mathcal{B}$ shown in \cref{fig:aut_b_annotated} and represented
-by the clauses in the final node in the tree. This corresponds to the end of
-\cref{sec:intuition:algebra}. Our only options now are to either split on which
-path we take from $I_B$ by case-splitting on~$\TransitionVar_{12}$ or to
-directly invoke \Materialise{}. The latter would produce a shorter tree in this
-instance but might lead to a larger product being computed, so we (and indeed
-\Catra{}) pick \Split{}. This leads to an opportunity for propagation since
-$\TransitionVar_{12} = 0$ cuts off $\TransitionVar_{9}$ from $I_B$ on the left
-branch. Note that the same opportunity does not present itself on the right
-branch, though; the use of the upper path to state~$A$ does not preclude visiting
-state~$B$.
-
-In both cases, we have comparatively simple automata left, so we apply
-\Materialise{} to obtain their product, which for both branches is empty. An
-empty product will have no outgoing transitions from its initial state, and so
-will produce a flow equation $1 = 0$ since the right-hand balancing side will
-always be the empty sum. We can then close the proof and know that the problem
-is unsatisfiable.
-
 \begin{figure}
@@
 
+We return to our example in \cref{sec:intuition}. We also reuse the
+definition of $\Map$ as the simple projection to extract the counter increments,
+so in this case, $\Map(\Transition) = \left[l_{B+}\right]$ for every
+transition~$\Transition$ since $\mathcal{B}$ only counts length. The reader is
+advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
+\cref{sec:intuition} while going through this example.
+The transition variables of the figures match the ones used here.
+
+We extend \cref{eq:single-filter} from \cref{sec:single:example} with
+definitions for $B$ shown in \cref{eq:multi-filter}:
+\begin{equation}\label{eq:multi-filter}
+  \begin{array}{rl@{\qquad}rl@{\qquad}rl}
+     \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) &= \TransitionVar_6 
+     &
+     \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) &= \TransitionVar_7 
+     &
+      \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) &= \TransitionVar_8 
+     \\
+     \Filter(\FromLabelTo{B}{c / l_{B+}}{B}) &= \TransitionVar_9 
+     &
+     \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B}) &= \TransitionVar_{10}
+     &
+     \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A}) &= \TransitionVar_{11}
+     \\
+      \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B}) &= \TransitionVar_{12}
+    \\[3ex]
+% 
+    \multicolumn{6}{c}{
+        \FlowEq(\mathcal{B}, \Filter) =\quad  \begin{aligned}
+            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12} \wedge\mbox{}\\
+            \TransitionVar_{11} &= \TransitionVar_6\wedge\mbox{}\\
+            \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}\mbox{}\\
+            \TransitionVar_6 + \TransitionVar_{10} &= 1
+        \end{aligned}}
+  \end{array}
+\end{equation}
+
+The only possible rule at the start is $\ExpandM$, which we use to add the
+corresponding constraints on each automaton of the product as we would have had
+in the single-automaton version. After that, we expand the various sums and apply
+light equality propagation.
+
+We then continue by repeating the steps of~\cref{sec:single:example}, shown 
+in~\cref{fig:derivation:single} since the steps to propagate
+\cref{ex:string-constraints} across the automaton $\mathcal{A}'$ still apply for the
+product. Note that this fixes the values of all transitions but one of $\mathcal{A}'$ as
+constants, including the free counter variables $i, n$, whose new constant
+values we propagate everywhere.
+
+Once we have performed reasoning by equality to eliminate transition variables,
+we obtain the version of $\mathcal{B}$ shown in \cref{fig:aut_b_annotated} and represented
+by the clauses in the final node in the tree. This corresponds to the end of
+\cref{sec:intuition:algebra}. Our only options now are to either split on which
+path we take from $I_B$ by case-splitting on~$\TransitionVar_{12}$ or to
+directly invoke \Materialise{}. The latter would produce a shorter tree in this
+instance but might lead to a larger product being computed, so we (and indeed
+\Catra{}) pick \Split{}. This leads to an opportunity for propagation since
+$\TransitionVar_{12} = 0$ cuts off $\TransitionVar_{9}$ from $I_B$ on the left
+branch. Note that the same opportunity does not present itself on the right
+branch, though; the use of the upper path to state~$A$ does not preclude visiting
+state~$B$.
+
+In both cases, we have comparatively simple automata left, so we apply
+\Materialise{} to obtain their product, which for both branches is empty. An
+empty product will have no outgoing transitions from its initial state, and so
+will produce a flow equation $1 = 0$ since the right-hand balancing side will
+always be the empty sum. We can then close the proof and know that the problem
+is unsatisfiable. A full derivation tree for the example can be found 
+in~\cref{fig:derivation:multi}.
+
 \subsection{Correctness of \Calculus{} for Products of Automata}

### 2024-03-01 main.tex
Relabel graphs etc
Revised results after bugfix
Apparently, this is our colour now
Enable camera-ready mode

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
+%\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
+\documentclass[acmsmall]{acmart}\settopmatter{}
 
@@
 \acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
+\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
 \acmArticle{1}
@@
 
+\usepackage{tikz}
+\usepackage[most]{tcolorbox}
+\definecolor{ourcolour}{RGB}{103,194,165}
+
 \begin{document}
@@
 %% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
+\author{Amanda Stjerna}
+%\authornote{with author1 note}          %% \authornote is optional;
                                         %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\orcid{0000-0003-4180-6118}             %% \orcid is optional
 \affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
+  %\position{Position1}
+  \department{Department of Information Technology}              %% \department is recommended
+  \institution{Uppsala university}            %% \institution is required
+  \streetaddress{Lägerhyddsvägen 1, hus 10}
+  \city{Uppsala}
+  \country{Sweden}                    %% \country is recommended
 }
-\email{first1.last1@inst1.edu}          %% \email is recommended
+\email{amanda.stjerna@it.uu.se}          %% \email is recommended
 
 %% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
+\author{Philipp Rümmer}
+%\authornote{with author2 note}          %% \authornote is optional;
                                         %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\orcid{0000-0002-2733-7098}             %% \orcid is optional
 \affiliation{
@@
 
-\section{Extensions}\label{sec:extensions}
-\input{extensions.tex}
+%\section{Extensions}\label{sec:extensions}
+%\input{extensions.tex}
 

### 2024-03-01 example.tex
Image descriptions

@@
     accepting states matching any symbol in the alphabet!}\label{fig:aut_a}
-      %\Description[SHORT]{LONG}
+      \Description[A 3-state Parikh automaton.]{%
+      A 3-state Parikh automaton with two registers counting the length
+      of a substring matching c or dd.%
+      }
   \end{subfigure}%
@@
     the string in register $l_B$.}\label{fig:aut_b}
-      %\Description[SHORT]{LONG}
+      \Description[A 4-state Parikh automaton.]{%
+      A 4-state Parikh automaton with initial state IB going to the B state,
+      which counts a number of letters c, followed by one final c before reaching 
+      the accepting state, FB. The second path from the initial state consumes an a,
+      goes into state A, consumes some number of c:s, followed by a final a to get to
+      the accepting state.}
   \end{subfigure}
@@
   from $I_{A'}$, the initial state.}
+  \Description[A 3-state automaton, with equations on its labels.]{%
+  The initial state accepts any character and is labelled with the variable i. From
+  that state there is a transition to the D state reading d with $n-1$ and one reading c
+  going directly to the final state. The second transition is labelled with the equation
+  $2-n$. The final state has one transition that consumes any character with the equation 
+  $l_A-i-n$.%
+  }
   \label{fig:a_2}
@@
   form.}\label{fig:aut_a_annotated}
-    %\Description[SHORT]{LONG}
+    %\Description[A 3-state automaton with ]{LONG}
   \vspace*{1.9ex}

### 2024-03-01 example.tex
All labels in example done

@@
   form.}\label{fig:aut_a_annotated}
-    %\Description[A 3-state automaton with ]{LONG}
+    \Description[A 3-state automaton with equations on its edges]{
+      A 3-state automaton with a self-loop on any character on the
+      initial state labelled 1, a transition on c to the accepting state
+      and one on d to a state named D, labelled 0 and 1 respectively,
+      then one transition from D to FA, the accepting state on d labelled 1.
+      The accepting state has a self-loo labelled $l_A -3$.}
   \vspace*{1.9ex}
@@
 }\label{fig:aut_b_annotated}
-    %\Description[SHORT]{LONG}
+    \Description[A 4-state automaton with its transitions annotated
+    with equations]{A 4-state automaton with transitions from the initial state to B on b,
+    one to A on a, labelled $x_{12}$ and $x_{11}$ respectively. State B has one self-loop
+    on c labelled $x_9$, one transition back to the initial state labelled 
+    b / $x_{11} + x_{12} - 1$. Finally, B has a transition to the accepting state on c,
+    labelled $1 0 x_{11}$. A has one self-loop on c labelled $l_A - x_9 - 2x_{11} - 2x_{12}$,
+    and one transition on a to the accepting state labelled $x_{11}$.}
 \end{subfigure}

### 2024-03-01 calculus.tex
Ensure that figures are refered to, come before ref.
Closes: #78
Closes: #87
Fix warnings

@@
 We use the convention of splitting the formulas in proof goals into
-linear (in)equalities ($\SomeInequalities$) and other formulas
+linear (in-) equalities ($\SomeInequalities$) and other formulas
 ($\SomeClause$), and assume that predicates $\Image$ and $\Connected$
@@
 transitions to $\Naturals$-valued terms (e.g.~\texttt{t}
-or~\texttt{t+1}). In our implementation \Catra{}, described in
-\cref{sec:implementation}, $\Filter$~is a vector of fresh variables
+or~\texttt{t+1}). In our implementation \Catra{}, described 
+in~\cref{sec:implementation}, $\Filter$~is a vector of fresh variables
 with the same size as~$\Transitions$.
@@
 
-\medskip
-%
 The rule~\textsc{Expand} expands an $\SinglePredicateInstance$
@@
 
-\medskip
-%
 A decision procedure would start from one or multiple
@@
 
-The proof tree is shown in \cref{fig:derivation:single}.  Like in
-\cref{sec:a_1,sec:intuition:algebra}, by arithmetic reasoning the
-calculus ends up with fixed values for several of the transition
-variables $\TransitionVar_k$, and eventually concludes that the
-transition directly from state $I_{A}$ to $F_{A}$ (variable~$x_3$)
-is incompatible with
-the constraint.  It is then possible to remove the $\Connected{}$
-predicate using \Subsume{}, since \Propagate{} is not able to infer
-further constraints, and no non-trivial applications of \Split{}
-remain to be done.  After this, by further arithmetic reasoning we can
-derive a solution~$l_A = 4, i = 1, n = 2$, and conclude that the root
-constraint is satisfiable.  To obtain values for the string
-variables~$s_1, s_2$ corresponding to the solution, one can
-construct an accepting path of the automaton with each transition~$t$
-taken $\Filter(t)$ times.
-
-\iffalse
-
-the rest of the reasoning can be
-continued without any part of \Calculus{} since all values of the
-existentially quantified transition variables
-$\TransitionVar_1,\ldots,\TransitionVar_5$ compatible with
-\cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring,const:something-before-and-after}
-are represented by the flow equations introduced by $\Expand{}$.
-\fi
-
-\begin{figure}[tb]
+\begin{figure}[h]
   \centering
@@
 automaton $\mathcal{A}'$ of \cref{fig:aut_a}.}\label{fig:derivation:single}
+\Description[]{}% This is fine; it's all text!
 \end{figure}
+
+The proof tree is shown in~\cref{fig:derivation:single}. Like in
+\cref{sec:a_1,sec:intuition:algebra}, by arithmetic reasoning the
+calculus ends up with fixed values for several of the transition
+variables $\TransitionVar_k$, and eventually concludes that the
+transition directly from state $I_{A}$ to $F_{A}$ (variable~$x_3$)
+is incompatible with
+the constraint.  It is then possible to remove the $\Connected{}$
+predicate using \Subsume{}, since \Propagate{} is not able to infer
+further constraints, and no non-trivial applications of \Split{}
+remain to be done.  After this, by further arithmetic reasoning we can
+derive a solution~$l_A = 4, i = 1, n = 2$, and conclude that the root
+constraint is satisfiable.  To obtain values for the string
+variables~$s_1, s_2$ corresponding to the solution, one can
+construct an accepting path of the automaton with each transition~$t$
+taken $\Filter(t)$ times.
+
+\iffalse
+
+the rest of the reasoning can be
+continued without any part of \Calculus{} since all values of the
+existentially quantified transition variables
+$\TransitionVar_1,\ldots,\TransitionVar_5$ compatible with
+\cref{const:more-inside-than-before,const:s1-in-c-dd,const:s1-substring,const:something-before-and-after}
+are represented by the flow equations introduced by $\Expand{}$.
+\fi
 

### 2024-03-06 abstract.tex
Fix units, mention SMT-COMP results in abstract

@@
 constraints on strings. We show that our solution strictly outperforms the
-standard approach described in~\citeauthor{generate-parikh-image} as well as the
+standard approach described by~\citeauthor{generate-parikh-image} as well as the
 over-approximating method recently described by~\citeauthor{approximate-parikh}
-by a wide margin, and for realistic timeouts for constraint solving also
-the~\Nuxmv{} model checker.
+by a wide margin, and for realistic timeouts (under \SI{60}{s}) also
+the~\Nuxmv{} model checker. When added as the Parikh image backend of \OstrichPlus{} to
+the \Ostrich{} string constraint solver's portfolio, it boosts its results on the
+quantifier-free strings with linear integer algebra track of SMT-COMP~20203 (QF\_SLIA)
+enough to solve the most \Unsat{} instances in that track of all competitors.

### 2024-03-06 evaluation.tex
Tables: caption above
Fix units, mention SMT-COMP results in abstract

@@
 were executed in batches of \BatchSize{}, each given a fresh JVM. Each JVM was
-warmed up for \numprint{10}s on a random benchmark from the set before starting
+warmed up for \SI{10}{s} on a random benchmark from the set before starting
 to execute. We believe this represents a realistic use case where \Calculus{} is
@@
 executing the instance, and exclude time spent parsing (usually far below
-\numprint{0.1}s).
+\SI{0.1}{s}).
 
@@
     \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
-    \caption{The distribution of runtimes for solved instances per backend, assuming they were solved within the \numprint{30}s timeout.
-    The whiskers use the default placement in Pandas, \numprint{1.5} times the IQR (interquartile range).}
+    \caption{The distribution of runtimes for solved instances per backend, assuming they were solved within the \SI{30}{s} timeout.
+    The whiskers use the default placement in Pandas, \numprint{1.5}~times the IQR (interquartile range).}
     \label{fig:runtime-boxplot}
@@
     \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
-    \caption{The number of instances solved as the time budget increases, simulated from one \numprint{120}s-timeout run.}
+    \caption{The number of instances solved as the time budget increases, simulated from one \SI{120}{s}-timeout run.}
     \label{fig:cactus}
@@
 Finally, in \cref{fig:runtime-boxplot,fig:duels:baseline,fig:duels:nuxmv} and we compare
-runtimes for solved instances between the backends. We see that \Nuxmv{} has a more even spread of runtimes up to \numprint{20}s,
+runtimes for solved instances between the backends. We see that \Nuxmv{} has a more even spread of runtimes up to \SI{20}{s},
 while both \Calculus{} and baseline tend towards solving their
@@
 instances that finish as the timeout increaseses.
-Notably, \cref{fig:duels:baseline} shows that any instance solved by baseline in \numprint{30}s is also
-solved by \Calculus{} in under \numprint{5}s.
+Notably, \cref{fig:duels:baseline} shows that any instance solved by baseline in \SI{30}{s} is also
+solved by \Calculus{} in under \SI{5}{s}.
 
 \begin{table}
+  \caption{Number of successful results within a timeout of \RuntimeTimeout{}.
+  Instances solved by no backend within the timeout (about half of the set) are
+  omitted from the table.}\label{tab:solve-status}
   \begin{center}
@@
   \end{center}
-  \caption{Number of successful results within a timeout of \RuntimeTimeout{}.
-  Instances solved by no backend within the timeout (about half of the set) are
-  omitted from the table.}\label{tab:solve-status}
 \end{table}
@@
 \begin{table}
-  \begin{center}
-  \input{graphs/qf_slia_comparison.tex}
-  \end{center}
   \caption{Number of solved benchmarks in the set of quantifier-free strings with
@@
   \Ostrich+CA and \textsc{Competition} which aggregate the best results from \Ostrich{}/CA-Str and 
-    all the competition results for non-\Ostrich{} solvers respectively. }
-  \label{tab:solve-status-smt-comp}
+  all the competition results for non-\Ostrich{} solvers respectively. }\label{tab:solve-status-smt-comp}
+  \begin{center}
+  \input{graphs/qf_slia_comparison.tex}
+  \end{center}
 \end{table}

### 2024-03-06 extensions.tex
Clarify backward propagation, dom()
Closes: #81
Fix units, mention SMT-COMP results in abstract

@@
 $\Dominates(C, \Automaton, \Transition)$ for some minimal $C$. Note that this
-captures both forward (unrachable from the initial state) and
+captures both forward (unreachable from the initial state) and
 backward (does not reach an accepting state) unreachability. Bidirectional cut
@@
 
-\iffalse
 \subsection{Transducers}\label{sec:ext:transducers}
@@
 Another interesting application is applying \Calculus{} to automata with
-multiple tracks, e.g. transducers. One application of such a calculus would be
+multiple tracks, e.g.\ transducers. One application of such a calculus would be
 to represent replace operations and other functions on regular languages, and to
-be able to answer questions like "does this operation change the length of the
-string". The difficulty in implementing it for transducers is first to perform
+be able to answer questions such as \enquote{does this operation change the length of the
+string}. The difficulty in implementing it for transducers is first to perform
 the mapping on the labels, which for both the length and Parikh cases is
@@
 operation is defining the product of transducers. In some cases it would
-probably be desirable to perform synchronisation (e.g. requiring overlapping
+probably be desirable to perform synchronisation (e.g.\ requiring overlapping
 transitions) on only some transitions, for example the first track.
@@
     of~\Catra, but has been neither optimised nor fully tested.
-\fi
 

### 2024-03-06 introduction.tex
Fix units, mention SMT-COMP results in abstract

@@
 outperforms both \Nuxmv{} and the baseline method, allowing \Calculus{}
-to solve every problem solved by baseline within \numprint{30}s in under \numprint{5}s.
+to solve every problem solved by baseline within \SI{30}{s} in under \SI{5}{s}.
 \Calculus{} in particular outperforms \Nuxmv{} on unsatisfiable instances, more than 
-doubling the number of unsatisfiable results found within a \numprint{30}s timeout.
+doubling the number of unsatisfiable results found within a \SI{30}{s}~timeout.
 

### 2024-03-06 example.tex
Title case

@@
 
-\subsection{Translating the constraints into a Parikh automata intersection problem}
+\subsection{Translating the Constraints into a Parikh Automata Intersection Problem}
 
@@
 
-\subsection{Solving the Parikh automata intersection problem using \Calculus{}}
+\subsection{Solving the Parikh Automata Intersection Problem Using \Calculus{}}
 
@@
 
-\subsubsection{Approximating paths through automata using flow analysis}\label{sec:a_1}
+\subsubsection{Approximating Paths Through Automata Using Flow Analysis}\label{sec:a_1}
 
@@
 
-\subsubsection{Arithmetic flow simplification}\label{sec:intuition:algebra}
+\subsubsection{Arithmetic Flow Simplification}\label{sec:intuition:algebra}
 
@@
 
-\subsubsection{Case Splitting and connectivity propagation}\label{sec:intuition:split}
+\subsubsection{Case Splitting and Connectivity Propagation}\label{sec:intuition:split}
 We now have a choice of two paths through automaton $\mathcal{B}$; the
@@
 
-\subsubsection{Computing products}\label{sec:intuition:materialise}
+\subsubsection{Computing Products}\label{sec:intuition:materialise}
 After discounting all transitions that can no longer be taken, we are

### 2024-03-06 data.tex
Data statement

A reproduction package featuring all relevant logs is available~\cref{artefact}.
Additionally, \Ostrich{} and \Catra{} are both available as living software
on GitHub at \url{https://github.com/amandasystems/catra} and \url{https://github.com/uuverifiers/ostrich}
respectively. A version of \Ostrich{} modified to use CA-Str is available in the \texttt{cea-catra}~branch
and comes bundled with the reproduction package.

All source code to generate the artefact and the paper itself is \emph{also} available on 
GitHub at~\url{https://github.com/amandasystems/oopsla-artefact}.

\Ostrich{} and \Catra{} are both open source under a 3-clause BSD~license.

### 2024-03-06 conclusion.tex
meta-data

@@
 domains, including other logics, model checking problems, as well as to more
-powerful automata such as transducers. In principle, we are also already able to
-express stronger constraints than Parikh automata due to our use of a full
-automated theorem prover, which allows adding arbitrary constraints in addition
-to the expected Presburger formulae.
+powerful automata such as transducers. In principle, we can already express
+stronger constraints than Parikh automata due to our use of an automated theorem
+prover which allows rich constraints on counter variables.
+
+\begin{acks}
+  This work was supported by
+    the Swedish Research Council (VR)
+    under grant~2018-04727, the Swedish Foundation for Strategic
+    Research (SSF) under the project WebSec (Ref.\ RIT17-0011), and the
+    Wallenberg project UPDATE.
+\end{acks}

### 2024-03-06 main.tex
Fix units, mention SMT-COMP results in abstract
Data statement
meta-data

@@
 \usepackage{ebproof}
+\usepackage{siunitx}
 \usepackage[makeroom]{cancel}
@@
   \department{Department of Information Technology}              %% \department is recommended
-  \institution{Uppsala university}            %% \institution is required
+  \institution{Uppsala University}            %% \institution is required
   \streetaddress{Lägerhyddsvägen 1, hus 10}
@@
 \affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
+%  \position{Position2a}
+  \department{Faculty of Informatics and Data Science}             %% \department is recommended
+  \institution{University of Regensburg}           %% \institution is required
+  \streetaddress{Bajuwarenstraße 4}
+  \city{Regensburg}
+  \postcode{93053}
+  \country{Germany}                   %% \country is recommended
 }
-\email{first2.last2@inst2a.com}         %% \email is recommended
+\email{philipp.ruemmer@ur.de}         %% \email is recommended
 \affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
+  %\position{Position1}
+  \department{Department of Information Technology}              %% \department is recommended
+  \institution{Uppsala University}            %% \institution is required
+  \streetaddress{Lägerhyddsvägen 1, hus 10}
+  \city{Uppsala}
+  \country{Sweden}                    %% \country is recommended
 }
-\email{first2.last2@inst2b.org}         %% \email is recommended
+\email{philipp.ruemmer@it.uu.se}         %% \email is recommended
 
@@
 
-%% Bibliography
-\bibliography{bibliography}
-
 %% Appendix
@@
 \input{appendix.tex}
+
+\section{Data-Availability Statement}
+\input{data.tex}
+
+%% Bibliography
+\bibliography{bibliography}
+
 \end{document}

### 2024-03-06 introduction.tex
more details in related work

@@
 paper, determining the emptiness of an intersection of Parikh automata, was
-recently shown to be PSPACE-complete~\cite{graph-queries}.
+shown to be PSPACE-complete~\cite{graph-queries}.
 
@@
 and occurs in \textsc{Sloth}~\cite{sloth}. Parikh images frequently appear when
-introducing cardinality constraints like length or string indexing. The
-state-of-the-art approach to handling Parikh image computation is to
+introducing cardinality constraints like length or string indexing.
+While relying on Parikh images (or on being able to check
+the emptiness of Parikh images under given side constraints), the mentioned
+papers do not propose any techniques to compute Parikh images.
+
+A common approach to handling Parikh image computation is to
 over-approximate the Parikh image of a product of $k$~automata
@@
 with the conjunction of the automata's Parikh maps, $\bigwedge_{i=1}^{k}
-\ParikhMap(\Language(\Automaton_i))$. Due to the over-approximation,
+\ParikhMap(\Language(\Automaton_i))$~\cite{approximate-parikh}.
+Due to the over-approximation,
 this approach works only for unsatisfiable instances,
@@
 through the use of symbolic derivatives~\cite{symbolic-boolean-derivatives}.
-In addition to presenting a decision procedure for lazily dispatching
+ The research by
+ \citeauthor{symbolic-boolean-derivatives} does not consider Parikh images,
+ however. In addition to presenting a decision procedure for lazily dispatching
 constraints, we similarly also allow for symbolic labelling of automata to

### 2024-03-06 abstract.tex
reformulated

--- 2023-07-12 abstract.tex
@@
 counts in strings of a given language. Automata-based string solvers frequently
-need to compute the Parikh image of large products (intersections) of
-nondeterministic automata, which in many operations is both prohibitively slow
-and memory-intensive. We contribute a novel understanding of how Parikh maps can
-be tackled as a constraint solving problem to solve real-world constraints
-stemming from functions on regular languages, including length and indexing
-constraints. Furthermore, we show how this formulation can be efficiently
-implemented as a calculus, \Calculus{}, in an automated theorem prover
-supporting Presburger logic.  We implement \Calculus{} in a tool called
-\Catra{}, and evaluate it on constraints produced by the \OstrichPlus{} string
-constraint solver when solving Parikh automata intersection problems produced
-when solving standard string constraint benchmarks involving cardinality
-constraints on strings. We show that our solution strictly outperforms the
-standard approach described in~\citeauthor{generate-parikh-image} as well as the
+need to compute the Parikh image of products (or intersections) of
+finite-state automata, in particular when solving string constraints that
+also include the integer data-type due to operations like string length
+and indexing. In this context, the computation of Parikh images often turns
+out to be both prohibitively slow
+and memory-intensive. This paper contributes a new understanding of how
+the reasoning about Parikh images can be cast as a constraint solving problem,
+and questions about Parikh images be answered without explicitly computing
+the product automaton or the exact Parikh image. The paper shows
+how this formulation can be efficiently implemented as a calculus,
+\Calculus{}, embedded in an automated theorem prover supporting
+Presburger logic. The resulting standalone tool
+\Catra{} is evaluate on constraints produced by the \OstrichPlus{} string
+solver when solving standard string constraint benchmarks
+involving integer operations.
+The experiments show that \Calculus{} strictly outperforms the
+standard approach by~\citeauthor{generate-parikh-image} to extract Parikh
+images from finite-state automata, as well as the
 over-approximating method recently described by~\citeauthor{approximate-parikh}
-by a wide margin, and for realistic timeouts for constraint solving also
-the~\Nuxmv{} model checker.
+by a wide margin, and for realistic timeouts (under \SI{60}{s}) also
+the~\Nuxmv{} model checker. When added as the Parikh image backend of \OstrichPlus{} to
+the \Ostrich{} string constraint solver's portfolio, it boosts its results on the
+quantifier-free strings with linear integer algebra track of SMT-COMP~2023 (QF\_SLIA)
+enough to solve the most \Unsat{} instances in that track of all competitors.

### 2024-03-06 introduction.tex
various reformulations

@@
 %
-To decide whether a formula like $x \in \Language_1 \wedge y \in \Language_2
+To check the satisfiability of a formula like $x \in \Language_1 \wedge y \in \Language_2
 \wedge |x| > |y|$, with string variables~$x, y$ and regular
-languages~$\Language_1, \Language_2$, is satisfiable, it is necessary to reason
-about the sets of word lengths induced by $\Language_1, \Language_2$, which are
-special cases of the Parikh image of such languages.
+languages~$\Language_1, \Language_2$, it is necessary to reason
+about the possible lengths of $x, y$ that are admitted by
+$\Language_1, \Language_2$, respectively. The
+set of word lengths $\{ |w| \mid w \in \Language\}$ in a language~$\Language$
+is a special case of the Parikh image of a regular language.
 This required combined reasoning about
@@
 
-The Parikh image, more broadly, is a characterisation of formal languages in
-terms of their character counts. Given a language over an alphabet~$\{a_1,
-\ldots, a_k\}$, the Parikh image is a set of $k$-dimensional vectors that
-contains some vector~$\VectorLiteral{m_1, m_2, \ldots, m_k}$ if and only if the
-formal language contains a word in which each $a_i$ occurs $m_i$ times. It is a
-classical result that the Parikh image of every context-free language (and,
-thus, also of every regular language) is a semilinear set,
-i.e., Presburger-definable~\cite{parikh-theorem}.
+The Parikh image, more broadly, is a characterisation of formal
+languages in terms of their character counts. Given a
+language~$\Language$ over an alphabet~$\{a_1, \ldots, a_k\}$, the
+Parikh image is a set of $k$-dimensional vectors that contains some
+vector~$\VectorLiteral{m_1, \ldots, m_k}$ if and only if the
+language~$\Language$ contains a word in which each $a_i$ occurs $m_i$
+times. It is a classical result that the Parikh image of every
+context-free language (and, thus, also of every regular language) is a
+semilinear set, i.e., Presburger-definable~\cite{parikh-theorem}.
+In
+fact, it is possible to compute an existential Presburger formula
+describing the Parikh image of any \emph{context-free} language in
+linear time~\cite{generate-parikh-image} in the size of the grammar
+describing the language. For the special case of
+regular languages, this result was also stated
+in~\cite{muscholl-linear}.
 
-In applications involving regular languages, it is typically necessary
-to consider the Parikh image not only of a single automaton, but of the
-intersection of multiple automata. This problem arises in string solvers in
+In applications,  it is often necessary   
+to consider the Parikh image not only of a single regular language, but of the
+intersection of multiple languages.
+This happens in string solvers in
 particular, as conjunctions of string constraints lead to the computation of
-length images of products (intersections) of regular languages represented as
-finite automata. The problem considered in this paper is thus:
+length images of intersections of regular languages represented as
+finite-state automata.
+On the other hand, we are often
+not interested in a closed-form description of the complete Parikh image,
+but rather in checking whether the Parikh image contains vectors satisfying
+some given properties. 
+\emph{The main problem considered in this paper is the following:}
 
@@
     \begin{tcolorbox}[colback=gray!5!white,colframe=gray!75!black,%
-        title=Satisfiability of Parikh images of a product of automata,%
-        width=0.8\linewidth]
-        Given a finite set of deterministic, finite-state automata, $\Automaton_1, \ldots, \Automaton_n$
-        over a finite, common, alphabet $\Set{a_1, \ldots, a_k}$,
-        and a Presburger formula, $\phi(\VectorLiteral{x_1, \ldots, x_k})$, 
-        decide if there is a word, $w$ accepted by the 
-        product $\Automaton_1 \times \cdots \times \Automaton_n$, whose Parikh vector 
+        title=Joint satisfiability of Parikh images modulo Presburger constraints,%
+        width=0.9\linewidth]
+        Given regular languages~$\Language_1, \ldots, \Language_n$
+        over a common alphabet $\Set{a_1, \ldots, a_k}$,
+        and a Presburger formula~$\phi(x_1, \ldots, x_k)$, 
+        decide if there is a word~$w \in \Language_1 \cap
+        \cdots \cap \Language_n$ whose Parikh vector 
         $\VectorLiteral{m_1, \ldots, m_k}$ satisfies $\phi$.
@@
 
-  The current method of handling this problem consists of computing the
-  intersection of the languages (the product of their automata),
-  constructing the Presburger representation of their Parikh image, and then
-  checking its satisfiability. It is possible to compute an existential
-   Presburger formula describing the
-  Parikh image of any \emph{context-free} language in linear
-  time~\cite{generate-parikh-image}. For our special case of regular languages,
-  this result was also stated in~\cite{muscholl-linear}. While theoretically
-  elegant, this construction has several disadvantages for us, often making it
-  impractical for integration into algorithms.
+The problem can be solved using the classical construction of Parikh
+images~\cite{muscholl-linear,generate-parikh-image} by first computing
+the product of finite-state automata accepting the
+languages~$\Language_1, \ldots, \Language_n$, then extracting a
+Presburger formula~$\psi$ describing the Parikh image of the product
+automaton, and finally checking the satisfiability of $\phi \wedge \psi$.
 
-  Firstly, the constructed Presburger
-  formula contains a linear number of existential quantifiers in the size of the
-  considered grammar, as well as complex Boolean structure, which is needed to
-  express the connectedness of sets of productions considered in the construction.%
-  \iffalse{}Eliminating those quantifiers to obtain a quantifier-free representation of the
-  Parikh image has exponential complexity~\cite{DBLP:conf/issac/Weispfenning97},
-  and is often impossible in reasonable time.\fi{}%
-  Just solving the Parikh image
-  membership problem is NP-complete, as it corresponds to computing a satisfying
-  assignment of the existential Presburger formula, and taxing for solvers as
-  well~\cite{ostrich-plus}.
+While theoretically elegant, this construction has several
+disadvantages that easily turn into bottlenecks when reasoning about
+Parikh images in algorithms or applications.
 
-  Secondly, computing the product itself is usually prohibitively expensive
-  in both memory and CPU due to exponential blow-up. Applying the approach in~\cite{generate-parikh-image} would in
-  this case require the eager computation of the product before its length image,
-  and result in an existential Presburger formula of exponential size (in the
-  number of automata). In several instances we have observed while solving
-  real-world string constraints, the computation of the product of automata
-  exhausts the memory of any machine available due to the exponential blow-up in
-  size of the product, quickly becoming intractable as the number of automata in
-  the product increases.
+Firstly, computing the product automaton is often prohibitively
+expensive in both memory and CPU time.  In several instances we have
+observed while solving real-world string constraints, the computation
+of the product of automata exhausts the memory of any machine
+available due to the exponential blow-up in size of the product,
+quickly becoming intractable as the number of automata in the product
+increases.
 
-  The current best published mitigation for this problem relies on the observation
-  that any Presburger formula that applies to a product of automata must also hold
-  for all of the terms of that product individually. Exploiting this fact, the Parikh
-  image of a product of automata is first represented as the conjunction of the image of
-  the individual automata of the product, falling back to computing the entire product
-  for satisfiable cases or cases when unsatisfiability is only discoverable after
-  full- or partial product construction~\cite{approximate-parikh}.
+Secondly, the constructed Presburger formula~$\psi$ contains a linear
+number of existential quantifiers in the size of the product
+automaton, as well as complex Boolean structure, which is needed to
+express the connectedness of paths considered in the construction.
+Solving formulas of this kind generally tends to be taxing for
+solvers~\cite{ostrich-plus}; in the present case, since the product
+automaton itself is exponentially big, also formula~$\psi$ has
+exponential size in the number of considered regular languages, and
+can easily become too complex for today's Presburger arithmetic
+solvers to handle.
+
+The current best published mitigation for this problem relies on the
+observation that the Parikh image of an intersection of languages can
+be over-approximated by the conjunction of the Parikh images of the
+individual languages. Such over-approximation can be used to derive
+that no word exists whose Parikh vector satisfies a formula~$\phi$,
+falling back to computing the intersection of the languages (or of
+some subset of the languages) for the satisfiable
+cases~\cite{approximate-parikh}.
 
 Addressing these concerns, we have developed a calculus for Parikh images of
-products of regular languages that we call \Calculus{}. It allows us to
+intersections of regular languages that we call \Calculus{}. It allows us to
 interleave the computation of arbitrarily deep products of automata with the
@@
     \begin{tcolorbox}[colback=ourcolour!5!white,colframe=ourcolour!75!black,%
-        title=Key ideas of \Calculus{},%
-        width=0.8\linewidth]
+        title=Key paradigms in \Calculus{},%
+        width=0.9\linewidth]
 
         \begin{itemize}
-            \item Enforce automata connectivity constraints lazily
+            \item Enforce automata connectivity constraints lazily;
             \item Use \emph{propagation} of constraints on the Parikh image to prune
-            transitions from the automata that are incompatible with constraints
-            \item Intelligently \emph{branch} on the presence\slash{}absence of key
-             transitions to drive propagation of the connectivity constraint
+            transitions from the automata that are incompatible with constraints;
+            \item Intelligently \emph{branch} on the presence or absence of key
+             transitions to drive propagation of the connectivity constraint;
              \item Compute products of automata lazily, after pruning transitions
-             that violate constraints using propagation
+             that violate constraints using propagation.
         \end{itemize}
@@
 We implement \Calculus{} as a plug-in theory for the \Princess{} automated
-theorem prover and experimentally use it as the Parikh image interesection solver for
-to the \Ostrich{} string constraint solver.
-
-We additionally wrap the Parikh image solver as a stand-alone tool, \Catra{}.
+theorem prover~\cite{princess}, and additionally wrap the Parikh image solver as a stand-alone tool, \Catra{}.
 \Catra{} also supports a variant of the approximate method of~\cite{approximate-parikh}, its fall-back variant adapted
 from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
-checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
+checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other 
 back-ends on \NrBenchmarks{} distinct Parikh automata intersection problems
@@
 
-We extend the \Ostrich{} string solver's backend portfolio with \Calculus{} via
-\Catra{} as the Parikh image intersection solver of the \OstrichPlus{}~algorithm,
-obtaining improvements in both \textsc{Sat}  and \textsc{Unsat} performance in
-\Ostrich{}' results from SMT-COMP 2023, crucially allowing \Ostrich{} to win
-the unsatisfiable category of the track of quantifier-free, linear constraints
-over string logic (QF\_SLIA) and increasing the number of \textsc{Sat}~results
-by \numprint{3}\% on the same track.
+We also extend the \Ostrich{} string solver~\cite{ostrich} with a new back-end
+based on \Catra{} as the Parikh image intersection solver of
+the \OstrichPlus{}~algorithm, obtaining improvements in both
+\textsc{Sat} and \textsc{Unsat} performance in \Ostrich{}' results
+from SMT-COMP 2023, crucially allowing \Ostrich{} to win the
+unsatisfiable category of the track of quantifier-free, linear
+constraints over string logic (QF\_SLIA) and increasing the number of
+\textsc{Sat}~results by \numprint{3}\% on the same track.
 

### 2024-03-07 implementation.tex
Fix minor reviewer issues
Closes: #77
Closes: #75
Closes: #76
Closes: #71
Closes: #72
Title case
Example input moved to artefact

@@
 baseline back-end. An example of an input file corresponding to our running
-example introduced in \cref{sec:intuition} can be seen in
-\cref{lst:input-example}.
+example introduced in \cref{sec:intuition} can be found in the root directory
+of the artefact~\cref{artefact}.
 
@@
 
-We baseline using the same Presburger solver (\Princess{}), input file parser,
-and automaton implementation as \Catra. We do this to better analyse the impact
+As a baseline, we use the same Presburger solver (\Princess{}), input file parser,
+and automaton implementation as \Catra{}. We do this to better analyse the impact
 of the calculus rules themselves. Adapting~\cite{generate-parikh-image}, we
 produce quantified Presburger formulae for each successive term and add them to
-\Princess. We compute the product incrementally term by term, checking
+\Princess{}. We compute the product incrementally term by term, checking
 satisfiability at each step. We use a priority queue to select automata for each
@@
 \begin{algorithm}
-  \caption{How we implement the baseline approach}\label{alg:baseline}
   \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
@@
 
-  $q \gets \NewPriorityQueue{$\mathcal{A}_i, \ldots, \mathcal{A}_n$}$
+  $q \gets \NewPriorityQueue{$\mathcal{A}_1, \ldots, \mathcal{A}_n$}$
 
@@
   \KwRet{$p$'s SAT status}
-
+  \caption{How we implement the baseline approach}\label{alg:baseline}
   \end{algorithm}
@@
 
-\subsection{Heuristics and search strategies}
+\subsection{Heuristics and Search Strategies}
 

### 2024-03-07 evaluation.tex
Update results

@@
 
-The benchmarks are executed in parallel across a cluster of identical
-machines, \BenchmarkRig{}, as the sole job for each machine.  We compiled
+The benchmarks for \Calculus{} and \Nuxmv{} are executed in parallel on \BenchmarkRig{}, with 4~threads
+sharing the same JVM.  The baseline implementation was executed as one single job per
+instance on StarExec, with the same configuration as CA-Str and \Ostrich{} used in SMT-COMP since
+it could not be executed reliably in less than a week on our machine.
+
+We compiled
 the code using Scala~\ScalaVersion{} and executed the experiments
@@
 clause learning disabled to show the impact of \cref{sec:random-restarts}, and
-\cref{sec:clause-learning} respectively.
+\cref{sec:clause-learning} respectively. These experiments are executed on a smaller
+machine, a six-core AMD~Ryzen~5 2600 but the same Java version. Their max
+heap size is set to \SI{20}{GB} since this system has less RAM. The baseline
+experiments are executed separately as only two threads, since the implementation is
+so RAM-hungry it is difficult to benchmark reliably under longer timeouts.
 
@@
 an intermittent product accepting only the empty string during product materialisation. 
-We implemented a fix for that bug and re-evaluated all benchmarks in SMT-COMP.
+We implemented a fix for that bug and re-evaluated all benchmarks.
 
 Since addressing those bugs we have observed no further soundness
-issues in neither \Catra{} nor CA-Str. Benchmarking results for both runs are included
-in the artefact of this paper, and show virtually identical performance characteristics.
+issues in neither \Catra{} nor in CA-Str. Benchmarking results for both runs are included
+in the artefact of this paper, and show virtually identical performance characteristics,
+with the exception for higher efficiency due to better benchmarking machines. We have
+additionally performed several of the benchmarks (in particular the \SI{120}{s} experiments)
+on machines and clusters of machines with widely different performance characteristics
+and have observed the trends to be surprisingly robust.
 

### 2024-03-07 introduction.tex
reformulations

@@
 In
-fact, it is possible to compute an existential Presburger formula
+fact, it is possible to compute an existential Presburger arithmetic formula
 describing the Parikh image of any \emph{context-free} language in
@@
     \begin{tcolorbox}[colback=gray!5!white,colframe=gray!75!black,%
-        title=Joint satisfiability of Parikh images modulo Presburger constraints,%
+        title=Joint satisfiability of Parikh images modulo Presburger arithmetic constraints,%
         width=0.9\linewidth]
@@
         over a common alphabet $\Set{a_1, \ldots, a_k}$,
-        and a Presburger formula~$\phi(x_1, \ldots, x_k)$, 
+        and a Presburger arithmetic formula~$\phi(x_1, \ldots, x_k)$, 
         decide if there is a word~$w \in \Language_1 \cap
@@
 languages~$\Language_1, \ldots, \Language_n$, then extracting a
-Presburger formula~$\psi$ describing the Parikh image of the product
+Presburger arithmetic formula~$\psi$ describing the Parikh image of the product
 automaton, and finally checking the satisfiability of $\phi \wedge \psi$.
@@
 
-Secondly, the constructed Presburger formula~$\psi$ contains a linear
+Secondly, the constructed Presburger arithmetic formula~$\psi$ contains a linear
 number of existential quantifiers in the size of the product
@@
 product's Parikh image, and is generalised to an arbitrary homomorphism over
-automata labels, including string lengths. This enables us to let both
+automata labels, including string length. This enables us to let both
 calculations inform each other, eliminating unnecessary work, and pruning the
@@
 For our problem, the constraints are automata connectivity and linear
-equations (Presburger formulas).
+equations (Presburger arithmetic formulas).
 We use propagation to lazily enforce connectivity of automata,
@@
 is infinite, the number of simple paths
-through the automata (and their product) is finite. In summary, the key ideas are:
+through the automata (and their product) is finite.
+\emph{In summary, the key ideas are:}
 
@@
 \begin{itemize}
-    \item The \Calculus{} calculus to efficiently compute (a homomorphism on)
-          the Parikh image of products of Parikh automata.
+\item The \Calculus{} calculus to efficiently check the satisfiability
+  of the Parikh image of an intersection of regular languages modulo
+  Presburger arithmetic side conditions.
     \item Techniques to efficiently implement \Calculus{} in a modern
@@
 
-The problem of computing constraints on Parikh images over products of regular
-languages under a given commutative homomorphism amounts to checking emptiness
-of products of
+The problem of satisfying Parikh images over products of regular
+languages, modulo Presburger arithmetic side conditions,
+amounts to checking emptiness of products of
 Parikh automata. Parikh automata are regular automata extended with integer
@@
 
-A common approach to handling Parikh image computation is to
+\citeauthor{muscholl-linear} and \citeauthor{generate-parikh-image}
+define a closed-form description of the Parikh image of any regular
+language as an existential Presburger arithmetic formula.
+%
+An approach to optimise the construction is to
 over-approximate the Parikh image of a product of $k$~automata
 $\ParikhMap(\Language(\Automaton_1) \cap \ldots \cap \Language(\Automaton_k))$
-with the conjunction of the automata's Parikh maps, $\bigwedge_{i=1}^{k}
+with the conjunction of the Parikh images, $\bigwedge_{i=1}^{k}
 \ParikhMap(\Language(\Automaton_i))$~\cite{approximate-parikh}.
 Due to the over-approximation,
-this approach works only for unsatisfiable instances,
-and will require falling back to computing the product of the automata before
+this approach is primarily useful for unsatisfiable instances,
+and requires falling back to computing the product of the automata before
 using the standard approach for finding its image originally presented
-in~\cite{generate-parikh-image}.
+in~\cite{generate-parikh-image}. Our calculus~\Calculus{}, in a somewhat
+similar but more fine-grained manner, utilises laziness to postpone or
+avoid the most expensive steps in the computation of Parikh images of
+the intersection of regular languages.
 
-Our calculus~\Calculus{} is similar in spirit to the work
+Our calculus~\Calculus{} is also similar in spirit to the work
 of \citeauthor{symbolic-boolean-derivatives}, who tackle the exponential
 blow-up resulting from Boolean combinations
-of regular automata in SMT string solvers
+of finite-state automata in SMT string solvers
 through the use of symbolic derivatives~\cite{symbolic-boolean-derivatives}.
@@
 constraints, we similarly also allow for symbolic labelling of automata to
-handle large Unicode alphabets.
+handle large alphabets.
 

### 2024-03-07 introduction.tex
more

@@
 languages~$\Language_1, \ldots, \Language_n$, then extracting a
-Presburger arithmetic formula~$\psi$ describing the Parikh image of the product
-automaton, and finally checking the satisfiability of $\phi \wedge \psi$.
+Presburger arithmetic formula~$\rho$ describing the Parikh image of the product
+automaton, and finally checking the satisfiability of $\phi \wedge \rho$.
 
@@
 
-Secondly, the constructed Presburger arithmetic formula~$\psi$ contains a linear
+Secondly, the constructed Presburger arithmetic formula~$\rho$ contains a linear
 number of existential quantifiers in the size of the product
@@
 solvers~\cite{ostrich-plus}; in the present case, since the product
-automaton itself is exponentially big, also formula~$\psi$ has
+automaton itself is exponentially big, also formula~$\rho$ has
 exponential size in the number of considered regular languages, and
@@
 intersections of regular languages that we call \Calculus{}. It allows us to
-interleave the computation of arbitrarily deep products of automata with the
-product's Parikh image, and is generalised to an arbitrary homomorphism over
-automata labels, including string length. This enables us to let both
-calculations inform each other, eliminating unnecessary work, and pruning the
+interleave the computation of product automata with the derivation of the
+Parikh image. This enables the two
+calculations to inform each other, eliminating unnecessary work, and pruning the
 size of the partial products considered in the computation for a smaller memory
-footprint. Moreover, our approach allows us to do this precisely, without the expensive
+footprint.
+Moreover, our approach allows us to do this precisely, without the expensive
 fallbacks when approximations fail as in the case of~\cite{approximate-parikh}.
+%
+Our calculus is applicable not only to Parikh images,
+but can be used for reasoning about the image of regular languages under
+arbitrary homomorphisms into commutative monoids. Besides Parikh images,
+this includes the length abstraction of regular languages commonly used in
+string solvers.
 

### 2024-03-07 calculus.tex
Fix hboxes
Tables: caption above
Title case
Update results
small stuff

@@
   The Parikh predicate, $\SinglePredicateInstance$ holds for some
-  automaton $\Automaton = \AutomatonTuple$, some map
+  automaton $\Automaton = \AutomatonTuple$, some homomorphism
   $\Map : \Alphabet^* \to \Monoid$ to a commutative monoid $\Monoid$,
@@
 \begin{table}
+  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
 \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
@@
   \end{tabular}
-  \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
 \end{table}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
@@
   }
-  \infer1[\EquationReasoning: $=$-simplification]{
+  \infer1[\EquationReasoning: $=$-simp.]{
     \begin{array}{c}
@@
   }
-  \infer1[\EquationReasoning{}: $=$-simplification]{
+  \infer1[\EquationReasoning{}: $=$-simp.]{
     \begin{array}{c}
@@
 
-\subsubsection{\Calculus{} terminates}
+\subsubsection{\Calculus{} Terminates}
 \begin{lemma}\label{lma:single-terminates}
@@
 
-\subsubsection{The rules in \cref{tbl:rules:single} are solution-preserving}
+\subsubsection{The Rules in \cref{tbl:rules:single} are Solution-preserving}
 

### 2024-03-07 calculus.tex
small stuff
Make automatas bigger, fix incorrect claim

@@
 
-We here only define forward propagation. Backward propagation 
-can be straightforwardly obtained by computing reachability from
-an accepting state on the same principle as above, and indeed our
-implementation of \Calculus{} does this.
-
 Finally, the rule~\Subsume{} can be applied when the connectedness of
@@
 \PresburgerClose{}, corresponding to a sound and complete solver for Presburger
-formulae, and for constraints over~$\Monoid$.
+arithmetic formulas, and for constraints over~$\Monoid$.
 

### 2024-03-07 data.tex
Add data statement
Make automatas bigger, fix incorrect claim

@@
-A reproduction package featuring all relevant logs is available~\cref{artefact}.
-Additionally, \Ostrich{} and \Catra{} are both available as living software
-on GitHub at \url{https://github.com/amandasystems/catra} and \url{https://github.com/uuverifiers/ostrich}
-respectively. A version of \Ostrich{} modified to use CA-Str is available in the \texttt{cea-catra}~branch
-and comes bundled with the reproduction package.
-
-All source code to generate the artefact and the paper itself is \emph{also} available on 
-GitHub at~\url{https://github.com/amandasystems/oopsla-artefact}.
-
-\Ostrich{} and \Catra{} are both open source under a 3-clause BSD~license.
+A reproduction package featuring all relevant logs and all benchmarked instances is available~\cref{artefact}.
+\Ostrich{} and \Catra{} are both available as living software
+under a 3-clause BSD~license on GitHub at \url{https://github.com/amandasystems/catra} and \url{https://github.com/uuverifiers/ostrich}
+respectively. A version of \Ostrich{} modified to use CA-Str is available in the \texttt{cea-catra}~branch. All source code to generate the paper itself
+is available at~\url{https://github.com/amandasystems/oopsla-artefact}.

### 2024-03-07 example.tex
Make automatas bigger, fix incorrect claim

@@
 
-\newcommand{\autscale}[0]{0.48}
+\newcommand{\autscale}[0]{0.60}
 
@@
     \centering 
-  \begin{subfigure}[b]{\autscale\textwidth}
+  \begin{subfigure}[b]{0.9\textwidth}
     \centering
@@
       which counts a number of letters c, followed by one final c before reaching 
-      the accepting state, FB. The second path from the initial state consumes an a,
+      the accepting state, FB.\@The second path from the initial state consumes an a,
       goes into state A, consumes some number of c:s, followed by a final a to get to
@@
 \begin{inparaenum}[(i)]
-\item Similarly, as in \cite{generate-parikh-image}, we first describe
+\item Similarly, as in~\cite{generate-parikh-image}, we first describe
   each automaton as a flow network, counting how often each transition
@@
   \centering 
-  \includegraphics[width=0.6\textwidth]{a_2}
+  \includegraphics[width=0.9\textwidth]{a_2}
   \caption{ $\mathcal{A}'$ after even more simplification using linear algebra,
@@
 \end{subfigure}\hfill%
-\begin{subfigure}{0.54\textwidth}
+\begin{subfigure}{0.58\textwidth}
   \centering

### 2024-03-07 extensions.tex
Make automatas bigger, fix incorrect claim

@@
   \begin{algorithm}
-    \DontPrintSemicolon
-    \caption{$\FindImage{}(\Automaton_1 \times \ldots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
+    \DontPrintSemicolon{}
+    \caption{$\FindImage{}(\Automaton_1 \times \cdots \times \Automaton_k, \Map)$ will find the Presburger form for the product $\Automaton_1 \times \ldots \times \Automaton_k$ modulo a homomorphism $\Map$ where the only free variable is/are the one(s) representing the monoid element of $\Map$.}\label{alg:find-image}
     \KwData{$\Automaton$, a product from \Calculus{}, a model $a$ assigning counts to the terms that $\Filter$ associate with each transition of $\Automaton$, and our homomorphism $\Map$ that we want to compute the image modulo.}

### 2024-03-07 multiple.tex
Fix hboxes
Tables: caption above
Title case
typesetting

@@
   Suppose $\Automaton_1, \ldots, \Automaton_k$ are automata,
-  $\Map : \Alphabet^* \to \Monoid$ is a map to a commutative monoid
+  $\Map : \Alphabet^* \to \Monoid$ is a homomorphism to a commutative monoid
   $\Monoid$,
@@
   \begin{table}[t]
-    \begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
+    \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
+    \begin{tabular}{l>{$}c<{$}l}\toprule
       Name & \text{Rule} & Side conditions\\
@@
       % EXPAND
-      \ExpandM & 
-      \inferrule
+      \ExpandM{} & 
+      \inferrule%
       {
-       {\begin{array}{c}
+       {\begin{array}{c@{}}
            \left\{
-           \FlowEq(\Automaton_i, \Filter),~~
-           \Connected(\Automaton_i, \Filter),~~
+            \begin{array}{c@{}}
+              \FlowEq(\Automaton_i, \Filter),
+              \Connected(\Automaton_i, \Filter),\\[1ex]
               \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
+            \end{array}
             \right\}_{i=1}^k,
@@
       None \\[5ex]
-      \Materialise &
-      \inferrule
+      \Materialise{} &
+      \inferrule%
       {{
-         \begin{array}{c}
-           \FlowEq(\Automaton', \Filter'),~
-           \Connected(\Automaton', \Filter'),~
-           \BindingSum(\Automaton_i, \Automaton_j, \Automaton', \Filter')
-           \\\ImagePredicate{
-                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
-                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
-                     \Automaton'}
-           }{\Map}{\Filter'}{\MonoidElement},~ \SomeInequalities, \SomeClause
+         \begin{array}{c@{}}
+           \FlowEq(\Automaton', \Filter'),
+           \Connected(\Automaton', \Filter'),
+           \BindingSum(\Automaton_i, \Automaton_j, \Automaton', \Filter'),\\
+           \ImagePredicate{
+                     \Tuple{%
+                      \Automaton_1,  \Automaton_{i-1},  \Automaton_{i+1},
+                      \ldots,  \Automaton_{j-1},  \Automaton_{j+1},
+                      \ldots,  \Automaton_k,  \Automaton'
+                     }
+           }{\Map}{\Filter'}{\MonoidElement},
+\SomeInequalities, \SomeClause
            \end{array}}
            }
-      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
-      $\begin{array}{@{}l@{}}
+      {{
+      \begin{array}{l@{}}
+      \ImagePredicate{
+        \Tuple{%
+          \Automaton_1,  \ldots,  \Automaton_i, \ldots, 
+          \Automaton_j, \ldots,  \Automaton_k
+        }}{\Map}{\Filter}{\MonoidElement}
+        \SomeInequalities, \SomeClause{}
+      \end{array}}} &%
+      $\begin{array}{l@{}}
           \Filter' = \Extend(\Automaton',\Filter),\\
@@
       \end{tabular}
-      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
     \end{table}
@@
 
-\subsubsection{\Calculus{} for products of automata terminates}
+\subsubsection{\Calculus{} for Products of Automata Terminates}
 \begin{lemma}\label{lma:multi-terminates}
@@
 
-\subsubsection{The rules in \cref{tbl:rules:multi} are solution-preserving}
+\subsubsection{The Rules in \cref{tbl:rules:multi} are Solution-preserving}
 

### 2024-03-07 data.tex
data: fel ordning

@@
 A reproduction package featuring all relevant logs and all benchmarked instances is available~\cref{artefact}.
-\Ostrich{} and \Catra{} are both available as living software
+\Catra{} and \Ostrich{} are both available as living software
 under a 3-clause BSD~license on GitHub at \url{https://github.com/amandasystems/catra} and \url{https://github.com/uuverifiers/ostrich}

### 2024-03-08 experiments/archived-results/reconstruct_starexec_logs.sh
new data

@@ -1,0 +1,17 @@
+#!/bin/sh
+job=Job62526_output_catra_baseline
+for logdir in ${job}/Catra/Catra\ 0.1.5___def/*.par
+do
+    instance=../deduped-benchmarks/$(basename "$logdir")
+    if  grep --quiet "===" "$logdir"/*.txt;
+    then
+        grep "===" "$logdir"/*.txt \
+            | sd --fixed-strings "../../benchmark/theBenchmark.par" "${instance}" \
+            | sd "^.*\t" ""
+    else
+        # Fake a timeouot log
+        echo "==== ${instance}: timeout > 30000ms run: 30.1s parse: 0.1s ===="
+    fi
+
+    #popd
+done

### 2024-03-08 experiments/reconstruct_starexec_logs.sh

@@ -1,17 +1,0 @@
-#!/bin/sh
-job=Job62526_output_catra_baseline
-for logdir in ${job}/Catra/Catra\ 0.1.5___def/*.par
-do
-    instance=../deduped-benchmarks/$(basename "$logdir")
-    if  grep --quiet "===" "$logdir"/*.txt;
-    then
-        grep "===" "$logdir"/*.txt \
-            | sd --fixed-strings "../../benchmark/theBenchmark.par" "${instance}" \
-            | sd "^.*\t" ""
-    else
-        # Fake a timeouot log
-        echo "==== ${instance}: timeout > 30000ms run: 30.1s parse: 0.1s ===="
-    fi
-
-    #popd
-done

### 2024-03-08 evaluation.tex
Add artefact, update evaluation

--- 2024-03-06 evaluation.tex
@@
 The benchmarks are run
-on commit~\texttt{\commit} of \Catra{}\footnote{Github repository redacted for anonymity.}.
+on commit~\texttt{\commit} of \Catra{}.
 
-The benchmarks are executed in parallel across a cluster of identical
-machines, \BenchmarkRig{}, as the sole job for each machine.  We compiled
+The benchmarks for \Calculus{} and \Nuxmv{} are executed in parallel on \BenchmarkRig{}, with 4~threads
+sharing the same JVM. Baseline was
+executed one JVM instance per input as 6~jobs in with a \SI{4}{GB} heap each on an Intel i5-10600, \SI{3.3}{GHz} CPU
+with 6 cores and 12 threads. Simultaneously, StarExec ran the same experiment and produced almost identical results.
+
+We compiled
 the code using Scala~\ScalaVersion{} and executed the experiments
@@
 clause learning disabled to show the impact of \cref{sec:random-restarts}, and
-\cref{sec:clause-learning} respectively.
+\cref{sec:clause-learning} respectively. These experiments are executed on a smaller
+machine, a six-core AMD~Ryzen~5 2600 but the same Java version. Their max
+heap size is set to \SI{20}{GB} since this system has less RAM. The baseline
+experiments are executed separately as only two threads.
 
@@
 discovered a soundness issue in CA-Str on one instance in QF\_S where the 
-\texttt{QF\_S\slash{}20230329-automatark-lu\slash{}instance08425.smt2} instance was
+\texttt{20230329-automatark-lu\slash{}instance08425.smt2} instance was
 incorrectly reported as \textsc{Unsat}. The underlying issue was that \Catra{} generated
-a too broad blocking clause when encountering unsatisifiability stemming from 
-an intermittent product accepting only the empty string during product materialisation. 
-We implemented a fix for that bug and re-evaluated all benchmarks in SMT-COMP.
+a too broad blocking clause for a product without transitions. We implemented a fix
+and re-evaluated all benchmarks. Performance was not measurably affected.
 
 Since addressing those bugs we have observed no further soundness
-issues in neither \Catra{} nor CA-Str. Benchmarking results for both runs are included
-in the artefact of this paper, and show virtually identical performance characteristics.
+issues in neither \Catra{} nor in CA-Str. Benchmarking results for both runs are included
+in the artefact of this paper, and show virtually identical performance characteristics. We have
+additionally executed all of the benchmarks on machines with widely different performance
+characteristics and have observed the trends to be robust.
 
@@
 \Princess{}, suggesting that the automaton implementation is not the bottleneck.
-Finally, the difference in performance between \Calculus{} and baseline has been
-robust under significant optimisation of the automata library, additionally
+Finally, the difference in performance between \Calculus{} and baseline was unaffected
+ by significant optimisation of the automata library, additionally
 strengthening this thesis.

### 2024-03-08 implementation.tex
Add artefact, update evaluation

@@
 example introduced in \cref{sec:intuition} can be found in the root directory
-of the artefact~\cref{artefact}.
+of the artefact~\cite{artefact}.
 

### 2024-03-08 evaluation.tex
small fixes

@@
 available at \url{https://www.starexec.org/starexec/secure/details/job.jsp?id=59410}.} 
-Where avaiable, we use the revised, out-of
+Where available, we use the revised, out-of
 competition version of the results for solvers with bugs discovered during
@@
 
-These results are expected. It is known that automata-based string solvers (\Ostrich{}, Z3-Noodler) 
+These results are not unexpected. It is known that automata-based string solvers (\Ostrich{}, Z3-Noodler) 
 tend to perform better on unsatisfiable than on satisfiable benchmarks, compared to solvers
 that do not utilize automata and directly work on regular expressions (cvc5, z3alpha).
-This is not unexpected, as the computation of automata representations of regular 
+The computation of automata representations of regular 
 constraints can be expensive, and might be unnecessary for satisfiable formulas.

### 2024-03-08 main.tex
Add data statement
move acks
DOI
further project
Add metadata from the ACM Author Submission
small fixes

@@
 
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
+%%% The following is specific to OOPSLA1 '24 and the paper
+%%% 'A Constraint Solving Approach to Parikh Images of Regular Languages'
+%%% by Amanda Stjerna and Philipp Rümmer.
+%%%
+\setcopyright{rightsretained}
+\acmDOI{10.1145/3649855}
+\acmYear{2024}
+\copyrightyear{2024}
+\acmSubmissionID{oopslaa24main-p181-p}
 \acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2023}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
+\acmVolume{8}
+\acmNumber{OOPSLA1}
+\acmArticle{138}
+\acmMonth{4}
+\received{21-OCT-2023}
+\received[accepted]{2024-02-24}
 
@@
 
+\section{Data-Availability Statement}
+\input{data.tex}
+
+\begin{acks}
+  We thank the reviewers for insightful comments.
+  This work was supported by
+    the Swedish Research Council (VR)
+    under grants~2018-04727 and 2021-06327, the Swedish Foundation for Strategic
+    Research (SSF) under the project WebSec (Ref.\ RIT17-0011), and the
+    Wallenberg project UPDATE.
+\end{acks}
+
 %% Acknowledgments
@@
 %% Appendix
-\appendix
-\section{Appendix}
-\input{appendix.tex}
-
-\section{Data-Availability Statement}
-\input{data.tex}
+%\appendix
+%\section{Appendix}
+%\input{appendix.tex}
 

### 2024-03-08 calculus.tex
typesetting
Shorter version

@@
   \caption{Derivation rules for one automaton.}\label{tbl:rules:single}
-\begin{tabular}{@{}l>{$}c<{$}p{3cm}@{}}\toprule
+\begin{tabular}{l>{$}c<{$}p{3cm}}\toprule
   Name & \text{Rule} & Side conditions\\
@@
 When constructing a proof, we start with some root~$\SomeClause$, and
-then apply proof rules to the goals of the proof, in bottom-up
-direction, until all goals are closed, or no more rule is applicable.
+then apply proof rules to the goals of the proof in bottom-up
+direction until all goals are closed, or no more rules are applicable.
 
 A proof in which all proof goals are closed shows that the formulas
-in the root~$\SomeClause$ of the proof are inconsistent, and do not
-have any solutions. A proof goal that cannot be closed, but to which no
-more rules are applicable, gives rise to  a solution of the formulas in the
-root~$\SomeClause$. Such an open proof goal will only contain formulas
-in Presburger arithmetic allowing a solution to be computed using
+in the root $\SomeClause$ of the proof are inconsistent (have no solutions).
+An unclosable goal to which no rules are applicable
+gives rise to a solution of the formulas in the
+root~$\SomeClause$. Such a goal will only contain formulas
+in Presburger arithmetic, allowing a solution to be computed using
 standard algorithms~\cite{DBLP:books/daglib/0022394}.

### 2024-03-08 data.tex
Add artefact, update evaluation
Shorter version

@@
-A reproduction package featuring all relevant logs and all benchmarked instances is available~\cref{artefact}.
+A reproduction package featuring all relevant logs and all benchmarked instances is available~\cite{artefact}.
 \Catra{} and \Ostrich{} are both available as living software
 under a 3-clause BSD~license on GitHub at \url{https://github.com/amandasystems/catra} and \url{https://github.com/uuverifiers/ostrich}
-respectively. A version of \Ostrich{} modified to use CA-Str is available in the \texttt{cea-catra}~branch. All source code to generate the paper itself
-is available at~\url{https://github.com/amandasystems/oopsla-artefact}.
+respectively. A version of \Ostrich{} modified to use CA-Str is available in the \texttt{cea-catra} branch. All source code to generate the paper itself
+is available at \url{https://github.com/amandasystems/oopsla-artefact}.

### 2024-03-30 data.tex
Sending bits and pieces of text to the workhouse

@@
 under a 3-clause BSD~license on GitHub at \url{https://github.com/amandasystems/catra} and \url{https://github.com/uuverifiers/ostrich}
-respectively. A version of \Ostrich{} modified to use CA-Str is available in the \texttt{cea-catra} branch. All source code to generate the paper itself
+respectively. A version of \Ostrich{} using CA-Str is available in the \texttt{cea-catra} branch. Code for the paper itself
 is available at \url{https://github.com/amandasystems/oopsla-artefact}.

### 2024-03-30 multiple.tex
Bundle a new version of acm template
Sending bits and pieces of text to the workhouse

@@
          \begin{array}{c@{}}
-           \FlowEq(\Automaton', \Filter'),
+           \FlowEq(\Automaton', \Filter'),\\
            \Connected(\Automaton', \Filter'),
@@
 
-For instances of precisely one automaton, neither rule applies, and we perform
+For precisely one automaton, neither rule applies, and we perform
 the calculus as before.
@@
 \cref{ex:string-constraints} across the automaton $\mathcal{A}'$ still apply for the
-product. Note that this fixes the values of all transitions but one of $\mathcal{A}'$ as
-constants, including the free counter variables $i, n$, whose new constant
+product. Note that this fixes the values of all transitions but one of $\mathcal{A}'$, including the free counter variables $i, n$, whose new constant
 values we propagate everywhere.
@@
 directly invoke \Materialise{}. The latter would produce a shorter tree in this
-instance but might lead to a larger product being computed, so we (and indeed
-\Catra{}) pick \Split{}. This leads to an opportunity for propagation since
+instance but might lead to a larger product being computed, so we pick \Split{}. This leads to an opportunity for propagation since
 $\TransitionVar_{12} = 0$ cuts off $\TransitionVar_{9}$ from $I_B$ on the left
 branch. Note that the same opportunity does not present itself on the right
-branch, though; the use of the upper path to state~$A$ does not preclude visiting
+branch; the use of the upper path to state~$A$ does not preclude visiting
 state~$B$.
 
-In both cases, we have comparatively simple automata left, so we apply
+In both cases the automata are now simpler, so we apply
 \Materialise{} to obtain their product, which for both branches is empty. An
 empty product will have no outgoing transitions from its initial state, and so
-will produce a flow equation $1 = 0$ since the right-hand balancing side will
-always be the empty sum. We can then close the proof and know that the problem
+will lead to a flow equation $1 = 0$. We can then close the proof and know that the problem
 is unsatisfiable. A full derivation tree for the example can be found 

### 2024-03-30 preliminaries.tex
Add logics to preliminaries
more
Sending bits and pieces of text to the workhouse

@@
 
-\subsection{Languages, Finite-state Automata and their Products}
-\label{sec:languages}
+%
+%\subsection{Logics}\label{sec:preliminaries:logic}
+%
+
+\subsection{Languages, Finite-State Automata and Their Products}\label{sec:languages}
 
@@
 
-\subsection{The Parikh Map and its Image}
+\subsection{The Parikh Map and Its Image}
 Formally, the \textit{Parikh map} over an alphabet $\Alphabet=
@@
 
-\subsection{The Parikh image of a regular language expressed in Presburger arithmetic}
+\subsection{The Parikh Image of a Regular Language Expressed in Presburger Arithmetic}
 \label{sec:verma}
